### 2015-02-28 curriculo/cv.tex
[ADD] added cv

%!TEX TS-program = xelatex
\documentclass[]{friggeri-cv}
\addbibresource{bibliography.bib}

\begin{document}
\header{Henrique}{Rocha}
       {Telecommunications and Informatics}

% In the aside, each new line forces a line break
\begin{aside}
  \section{about}
    Henrique Duarte Lopes Rocha
    ~
    Rua dos Cedros nº100
    Alto do Lagoal
    2760-025 Caxias
    Portugal
    ~
    (+351) 938781922
    ~
    \href{mailto:hdlopesrocha91@gmail.com}{hdlopesrocha91\\@gmail.com}
    \href{http://web.ist.utl.pt/ist168621}{WebSite}
    \href{http://www.youtube.com/user/hdlopesrocha}{YouTube}
  \section{languages}
    portuguese (native)
    english (fluent)
  \section{programming}
    {\color{red} $\varheartsuit$} Java
	 (Android, PlayFramework, LWJGL, MapReduce), C, C++ (OpenGL), C\# (XNA, Windows Forms, WPF, lpsolve), VB, Python, Scala, Scheme, HTML, CSS, JavaScript (jQuery, AJAX, WebGL), SQL, PHP, GLSL, HLSL.  
   \section{databases}
	MySQL, PostgreSQL, MongoDB, DynamoDB.
   \section{soft skills}
	Time Management
	Problem-Solving
	Team Player
	Self Confidence
	Adaptability
\end{aside}

\section{education}

\begin{entrylist}
   \entry
    {2013–now}
    {Master of Science}
    {Instituto Superior Técnico - Taguspark}
    {Telecommunications and Informatics Engineering}

   \entry
    {2009–2014}
    {Bachelor of Science}
    {Instituto Superior Técnico - Taguspark}
    {Communication Networks Engineering}

  \entry
    {2006–2009}
    {Secondary School}
    {Escola Secundária Luís de Freitas Branco, Paço de Arcos}
    {Science \& Technology}

\end{entrylist}

\section{experience}

\begin{entrylist}
  \entry
    {11 2014\\(ongoing)}
    {Web Development}
    {Bullray}
    {\emph{Development of an architecture for a chat application. Experience with Android, PlayFramework and MongoDB.}}
  \entry
    {04 2014\\(6 months)}
    {Scientific Initiation Scholarship}
    {INESC}
    {\emph{Frontend Development for Provide Results's Bimk project. Experience with PlayFramework, Revit API and GIT.}}
  \entry
    {09 2013\\(6 months)}
    {Scientific Initiation Scholarship}
    {CEG-IST}
    {\emph{Programming Decision Analysis Software (MACBETH). Development of an algorithm to suggest judgements based on past decisions using \emph{lpsolve}.}}
  \entry
    {08 2013\\(2 weeks)}
    {Summer Work}
    {LCG}
    {\emph{Frontend development for Decision Analysis Software using \emph{ASP.NET}. }}
  \entry
    {11 2012}
    {Pizza Night Competition}
    {Microsoft Lisbon Experience}
    {\emph{2º place on Windows Phone Category} with \emph{Infinity Edge} game.}
  \entry
    {08 2006\\(1 month)}
    {Garbage Collector}
    {Jovens em Movimento, Oeiras}
    {Cleaning and maintenance of public spaces.}

\end{entrylist}

\section{applications}

\begin{entrylist}
  \entry
    {2013}
    {OpenGlobe}
    {\href{http://web.ist.utl.pt/ist168621/?page=10}{Instituto Superior Técnico}}
    {Modeling Earth surface with GPS devices and topographic collected data. It is also possible to predict satellites position through ephemeris data.}

  \entry
    {2012}
    {Infinity Edge (Windows Phone 7)}
    {\href{http://www.windowsphone.com/en-us/store/app/infinity-edge/05516463-3a89-4351-996e-62e5b4519aeb}{windows marketplace}}
    {First Person Shooter with spaceships.}

  \entry
    {2011}
    {Era of Empires}
    {\href{http://web.ist.utl.pt/ist168621/?page=8}{Personal Project}}
    {3D environment generator and Real Time Strategy game engine.}

\end{entrylist}

\section{interests}
    3D Studio Max, Photoshop, Game Development, Web Design, Network Security, Jogging and Yoga.

%\section{publications}

%Put your publications here!

%%% This piece of code has been commented by Karol Kozioł due to biblatex errors. 
% 
%\printbibsection{article}{article in peer-reviewed journal}
%\begin{refsection}
%  \nocite{*}
%  \printbibliography[sorting=chronological, type=inproceedings, title={international peer-reviewed conferences/proceedings}, notkeyword={france}, heading=subbibliography]
%\end{refsection}
%\begin{refsection}
%  \nocite{*}
%  \printbibliography[sorting=chronological, type=inproceedings, title={local peer-reviewed conferences/proceedings}, keyword={france}, heading=subbibliography]
%\end{refsection}
%\printbibsection{misc}{other publications}
%\printbibsection{report}{research reports}

\end{document}

### 2015-03-05 tese/state.tex
[ADD] added first tex (inro on NAT), added a paper about IPv4 (references must be done carefully)

\documentclass{article}

\usepackage[brazilian]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[fleqn]{amsmath}
\usepackage{amssymb}
\usepackage{listings}
\usepackage{color}
\usepackage{makeidx}
\usepackage{float}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[hmargin=3cm,vmargin=3.5cm]{geometry}
\usepackage{indentfirst}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\usepackage{eurosym}
\usepackage{url}

\lstset{frame=tb,
  language=C,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true
  tabsize=3
}

\usepackage{titlesec}
\titleformat{\section}{\hrule \large\bfseries}{\thesection}{1em}{}

\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração potenciado pela Tecnologia WebRTC}
\author{
	Henrique Lopes Rocha \\ hdlopesrocha91@gmail.com
}

\date{Fevereiro 2015}
\begin{document}
	\maketitle
	\newpage
	\tableofcontents
	\newpage

\textbf{Abstract}
\section{Introduction}
The need to build a global comunication network in an era when, almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly know users. IPv4 limits the number o public addresses in such a way that today are scarse. One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group though their private addresses, this technique is also know as Network Address Translation (NAT).

\begin{thebibliography}{9}

\bibitem{activator} Typesafe - Build Reactive Applications on the JVM - \url{http://typesafe.com/}
\bibitem{github} GitHub - Build software better, together - \url{http://github.com/}
\bibitem{gitlab} GitLab - Version Control on your Server - \url{http://gitlab.com/}
\bibitem{gitolite} Gitolite - Hosting Git Repositories - \url{http://gitolite.com/}
\bibitem{gitweb} Gitweb - Git SCM Wiki - \url{http://git.wiki.kernel.org/index.php/Gitweb}
\bibitem{iptables} iptables - Linux man page - \url{http://linux.die.net/man/8/iptables}
\bibitem{jquery} jQuery - write less, do more - \url{http://jquery.com/}
\bibitem{json} JSON - JavaScript Object Notation - \url{http://json.org/}
\bibitem{mongodb} MongoDB - NoSQL database - \url{http://mongodb.org/}
\bibitem{rest} REST - Representational state transfer - \url{http://en.wikipedia.org/wiki/Representational_state_transfer}
\bibitem{sdk} Android SDK - Software Development Kit - \url{http://developer.android.com/sdk/index.html}
\bibitem{ubuntu} Ubuntu Server - for scale-out computing - \url{http://www.ubuntu.com/server}
\bibitem{virtualbox} VirtualBox - Oracle VM - \url{http://virtualbox.org/}

\end{thebibliography}

\end{document}

### 2015-03-05 tese/state.tex
[ADD] added more stuff about NAT

@@
-\documentclass{article}
+\documentclass[11pt,twocolumn]{article}
 
-\usepackage[brazilian]{babel}
 \usepackage[utf8x]{inputenc}
@@
 \usepackage{makeidx}
+\usepackage{blindtext}
 \usepackage{float}
@@
 
-\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração potenciado pela Tecnologia WebRTC}
+\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC}
 \author{
@@
 
-\date{Fevereiro 2015}
+\date{\today}
 \begin{document}
+
+\begingroup
+\let\onecolumn\twocolumn
+
 	\maketitle
-	\newpage
-	\tableofcontents
-	\newpage
 
-\textbf{Abstract}
+\begin{abstract}
+Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
+Etiam lobortis facilisis sem.
+Nullam nec mi et neque pharetra sollicitudin.
+Praesent imperdiet mi nec ante.
+Donec ullamcorper, felis non sodales commodo, lectus velit ultrices augue, a dignissim nibh lectus placerat pede.
+\end{abstract}
+
+\endgroup
+
 \section{Introduction}
-The need to build a global comunication network in an era when, almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly know users.
-IPv4 limits the number o public addresses in such a way that today are scarse.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group though their private addresses, this technique is also know as Network Address Translation (NAT).
+The need to build a global comunication network in an era when almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
+IPv4 limits the number of public addresses in such a way that today are scarse \cite{ipv4}.
+One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each element is identified publicly by the same IP address but different ports, this technique is also know as Network Address Translation (NAT).
+
+Initially NAT offered an alternative for address exhaustion and a false sensation of better security, asymetric NAT became a vulgar configuration on the web.
+As a direct result, problems started to appear, the amount of ports that IP disponibilizes is also small compared to our current needs, worse than that, NAT also difficults end-to-end communication, forcing most of applications that follows this model to be implemented unificiently.
+
+Applications based on multimedia and file sharing were one of the most strained.
+To enable end-to-end comunication, one could simulate it via intermediate public servers or using STUN and TURN servers to overpass NAT.
+
+Client-Server models aren't suitable for real time comunication between two private end points.
+For example, amongst multiple alternatives, a messaging service could be implemented through a publish subscriber schema or by polling a messaging server.
+Clearly this type of communication requires a more expensive infrastructure and, at most cases, more network usage, leading to a worse quality of service.
+The requirements of video comunication makes this kind of model out of question.
+
+Efforts were made by Skype and Google to provide real time comunication between two or more entities, either having public or private addresses.
 
@@
 
-\bibitem{activator} Typesafe - Build Reactive Applications on the JVM - \url{http://typesafe.com/}
-\bibitem{github} GitHub - Build software better, together - \url{http://github.com/}
-\bibitem{gitlab} GitLab - Version Control on your Server - \url{http://gitlab.com/}
-\bibitem{gitolite} Gitolite - Hosting Git Repositories - \url{http://gitolite.com/}
-\bibitem{gitweb} Gitweb - Git SCM Wiki - \url{http://git.wiki.kernel.org/index.php/Gitweb}
-\bibitem{iptables} iptables - Linux man page - \url{http://linux.die.net/man/8/iptables}
-\bibitem{jquery} jQuery - write less, do more - \url{http://jquery.com/}
-\bibitem{json} JSON - JavaScript Object Notation - \url{http://json.org/}
-\bibitem{mongodb} MongoDB - NoSQL database - \url{http://mongodb.org/}
-\bibitem{rest} REST - Representational state transfer - \url{http://en.wikipedia.org/wiki/Representational_state_transfer}
-\bibitem{sdk} Android SDK - Software Development Kit - \url{http://developer.android.com/sdk/index.html}
-\bibitem{ubuntu} Ubuntu Server - for scale-out computing - \url{http://www.ubuntu.com/server}
-\bibitem{virtualbox} VirtualBox - Oracle VM - \url{http://virtualbox.org/}
+\bibitem{ipv4} Next Generation Internet: IPv4 Address Exhaustion, Mitigation Strategies and Implications for the U.S. - An IEEE-USA White Paper - 2009
 

### 2015-03-08 tese/state.tex
[ADD] added more paepers to read

@@
 
-Applications based on multimedia and file sharing were one of the most strained.
-To enable end-to-end comunication, one could simulate it via intermediate public servers or using STUN and TURN servers to overpass NAT.
+Applications based on multimedia and file sharing were one of the most strained by NAT.
+Those kind applications requires real time communication in order to achieve the best performance.
+STUN and TURN \cite{natvoip} servers are a possible solution to overpass NAT, although, none of those an establish direct connections on multiple level NATs.
 
-Client-Server models aren't suitable for real time comunication between two private end points.
-For example, amongst multiple alternatives, a messaging service could be implemented through a publish subscriber schema or by polling a messaging server.
+Most of client-server applications aren't affected by NAT when the servers are public, but they aren't suitable for real time comunication between two private end points.
 Clearly this type of communication requires a more expensive infrastructure and, at most cases, more network usage, leading to a worse quality of service.
@@
 
-Efforts were made by Skype and Google to provide real time comunication between two or more entities, either having public or private addresses.
+On the other hand, TURN uses public servers to redirect traffic between private end points, it may use a P2P network relay to find the best peer, but after that the behaviour is much like client-server.
+Direct communication is only achieved by STUN when NAT is type \textit{full cone}.
+{\color{red} [reference to NAT types]}.
+ICE uses STUN when it's possible and TURN otherwise.
 
@@
 \bibitem{ipv4} Next Generation Internet: IPv4 Address Exhaustion, Mitigation Strategies and Implications for the U.S. - An IEEE-USA White Paper - 2009
+\bibitem{natvoip} How NAT-Compatible Are VoIP Applications? Ying-Dar Lin, Chien-Chao Tseng, Cheng-Yuan Ho, and Yu-Hsien Wu, National Chiao Tung University
 

### 2015-03-14 tese/state.tex
[UPD] updated CV
[ADD] intro to hypermedia

@@
 
+When connection is established, either direct or indirect (via TURN servers), WebRTC cames to simplify how audio and video are transmited.
+
+\section{Hypemedia}
+
+Although video and audio is almost standardized, WebRTC allows us to enrich web communications by overlaying multiple elements such as buttons, text and more complex user interfaces.
+
+In the past, multiple formats were proposed to deal with hypermedia.
+
+HyVAL\cite{hyval} was proposed for modeling composition of video.
+The definition of a format must be done with carefoul.
+When a format estabilished for a document, it is expected that a parser knows how to convert text into behaviour, formats like XML are quite limitative, they cannot reproduce actions that were not previously implemented.
+
+Our solution to overcome format limitation is to take advantage of languages like javascript, like any webpage, user interface elements are defined by html tags, more complex actions like animations can be easily implemented via CSS and javascript, without ever defining a new standard for structured hypermedia.
+
+By using this already solidified web technologies, it's possible to raise communications to a new level.
+For example, with API's like WebGL, it is now possible to manipulate a three dimensional environment in the context of an hypercall.
+Another example would be a colaborative spreadsheet using WebRTC.
+With this, hypercalls are not limited to only video, audio and text.
+
+In this project our goal is to enrich hypercalls with no limits, every user should be free to choose how it whant to be contacted.
+
+Consider this case study, say that Alice whants to call Bob.
+Alice starts a WebRTC connection with Bob, at this point Alice doesn't know how the communication will proceed.
+Bob created a personal web page where he defined how he whants to be contacted.
+Alice downloads Bob's personal webpage and finds what kind of interactions could be done, for instance she knows that she can contact Bob via video streaming or can browse past videos.
+While Bob is connected, he can add links on videos, organise its content and enable different plugins for communication.
+
 \begin{thebibliography}{9}
@@
 \bibitem{natvoip} How NAT-Compatible Are VoIP Applications? Ying-Dar Lin, Chien-Chao Tseng, Cheng-Yuan Ho, and Yu-Hsien Wu, National Chiao Tung University
+\bibitem{hyval} A Structured Document Model for Authoring Video-based Hypermedia
 

### 2015-03-18 tese/state.tex
[ADD] added more smil papers and state of art

@@
 
-\section{Hypemedia}
+\section{Hypermedia}
+
+  Since the early days of video technology, one of the problems that raised with it consisted on how to add more information onto it without generating multiple versions.
+The need to translate movies, raised the problem wether is apropriate to change the original video or audio.
+For example subtitles should be an entity independent from the video, in order to be personalized or replaced easily.
+
+  Amongst multiple formats for subtitles, SAMI, SRT and SUB are used by video players. Although those formats have styling available, they are limited to text.
+
+  SMIL was introduced to describe temporal behaviour of multimedia, for instance, it could be used to overlay subtitles on films.
+With SMIL it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a diferent audio, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
+
+  In order to create a multimedia rich hypercall, SMIL fits our goals, but it lacks on browser compatibility.
+Ambulant was one of the SMIL players that were developed for browsers, although this player implements most of SMIL 3.0 specifications, it needs to be installed on browsers as a plugin.
+
+  SmilingWeb attempts to implement SMIL 3.0 with javascript, which doesn't need to be installed and shouldn't have incompatibility issues.
+But SmillingWeb isn't fully implemented yet and their sheduler engine loads the SMIL file only once, which could raise problems when leading with SMIL changes in real time.
+
+  SVG is a format that implements part of SMIL {\color{red} [which version and functionalities?]}
 
@@
 
-In the past, multiple formats were proposed to deal with hypermedia.
-
-HyVAL\cite{hyval} was proposed for modeling composition of video.
-The definition of a format must be done with carefoul.
-When a format estabilished for a document, it is expected that a parser knows how to convert text into behaviour, formats like XML are quite limitative, they cannot reproduce actions that were not previously implemented.
-
-Our solution to overcome format limitation is to take advantage of languages like javascript, like any webpage, user interface elements are defined by html tags, more complex actions like animations can be easily implemented via CSS and javascript, without ever defining a new standard for structured hypermedia.
+Our solution to overcome format limitation is to take advantage of languages like javascript, like any webpage, user interface elements are defined by html tags, more complex actions like animations can be easily implemented via CSS and javascript.
 
@@
 
+\subsection{Trash}
+
+{\color{red}
+
 Consider this case study, say that Alice whants to call Bob.
@@
 While Bob is connected, he can add links on videos, organise its content and enable different plugins for communication.
-
+}
 \begin{thebibliography}{9}

### 2015-03-19 tese/state.tex
[ADD] added more references

--- 2015-03-08 tese/state.tex
@@
 \section{Introduction}
-The need to build a global comunication network in an era when almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
-IPv4 limits the number of public addresses in such a way that today are scarse \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each element is identified publicly by the same IP address but different ports, this technique is also know as Network Address Translation (NAT).
+The need to build a global comunications network in an era when almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
+IPv4 limits the number of public addresses in such a way that today is scarse \cite{ipv4}.
+One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same IP address but different port, this technique is also know as Network Address Translation (NAT).
 
-Initially NAT offered an alternative for address exhaustion and a false sensation of better security, asymetric NAT became a vulgar configuration on the web.
-As a direct result, problems started to appear, the amount of ports that IP disponibilizes is also small compared to our current needs, worse than that, NAT also difficults end-to-end communication, forcing most of applications that follows this model to be implemented unificiently.
+Initially NAT offered an alternative for address exhaustion and a false sensation of better security, asymmetric NAT became a vulgar configuration on the web.
+As a direct result, problems started to appear, the amount of ports that IP disponibilizes is also small compared to our current needs, worse than that, NAT also difficult end-to-end communication, forcing most of applications that follows this model to be implemented ineffectively.
 
 Applications based on multimedia and file sharing were one of the most strained by NAT.
-Those kind applications requires real time communication in order to achieve the best performance.
-STUN and TURN \cite{natvoip} servers are a possible solution to overpass NAT, although, none of those an establish direct connections on multiple level NATs.
+Those kind applications require real time communication in order to achieve the best performance.
+STUN and TURN \cite{natvoip} servers are a possible solution to overpass NAT, although, none of those can establish direct connections on multiple level NATs.
 
-Most of client-server applications aren't affected by NAT when the servers are public, but they aren't suitable for real time comunication between two private end points.
-Clearly this type of communication requires a more expensive infrastructure and, at most cases, more network usage, leading to a worse quality of service.
-The requirements of video comunication makes this kind of model out of question.
+Most of client-server applications aren't affected by NAT when the servers are public, but they're inadequate for real time communication between two private endpoints.
+Clearly this type of communication requires a more expensive infrastructure and, in most cases, more network usage, leading to a worse quality of service.
+The requirements of video communication makes this kind of model unsuitable.
 
-On the other hand, TURN uses public servers to redirect traffic between private end points, it may use a P2P network relay to find the best peer, but after that the behaviour is much like client-server.
-Direct communication is only achieved by STUN when NAT is type \textit{full cone}.
+On the other hand, TURN uses public servers to redirect traffic between private endpoints, it may use a P2P network relay to find the best peer but, after that, the behavior is much like client-server.
+Direct communication is only achieved by STUN when NAT is a type \textit{full cone}.
 {\color{red} [reference to NAT types]}.
 ICE uses STUN when it's possible and TURN otherwise.
+
+When connection is established, either in a direct or indirect way (via TURN servers), WebRTC cames to simplify how audio and video are transmited.
+
+\section{Hypermedia}
+
+  Since the early days of video technology, one of the problems that raised with it consisted of how to add more information onto it without generating multiple versions.
+Some implementations like \cite{embedded} added hypermedia information to empty space present on MPEG frames, the MPEG coder and decoder were changed in order to handle hypermedia content.
+
+  The need to translate movies, raised the problem whether it is apropriate to change the original video or audio.
+For example subtitles should be an entity independent from the video, in order to be personalized or replaced easily.
+
+  Amongst multiple formats for subtitles, SAMI, SRT and SUB are used by video players that support them. Although those formats have styling available, they are quite limited to text.
+
+  HyVAL\cite{hyval} was proposed for modeling composition and interaction of video, this kind of video is also called hypervideo, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
+This approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
+
+  SMIL was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
+With SMIL it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
+
+  In order to create a multimedia rich hypercall, SMIL fits our goals, but it lacks on browser compatibility.
+Ambulant\cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of SMIL 3.0 specifications, it needs to be installed on browsers as a plugin.
+
+  SmillingWeb attempts to implement SMIL 3.0 with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
+But SmillingWeb isn't fully implemented yet and their scheduler engine loads the SMIL file only once, which could raise problems when leading with SMIL changes in real time.
+
+  SVG is a format that implements part of SMIL, it allows dynamic changes to inner content in real time, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
+
+  Video functionalities are already embedded in HTML5, like SVG it is also possible to bind javascript functions for different kinds of events over videos.
+
+  By using solid technologies like Javascript, CSS, SVG and HTML5 in a coordinated way, it's possible to raise communications to a new level.
+For example, with API's like WebGL, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
+Another example would be a collaborative spreadsheet using WebRTC.
+With this, hypercalls are not limited to only video, audio and text.
+
+  In this project our goal is to enrich hypercalls with no limits, every user should be free to choose how it wants to be contacted.
+
+  In order to give users a personalized communication channel, each user must have a personal web page where its available plugins could be downloaded from other peers, after that they can talk in the same language whatever it is.
 
@@
 \bibitem{natvoip} How NAT-Compatible Are VoIP Applications? Ying-Dar Lin, Chien-Chao Tseng, Cheng-Yuan Ho, and Yu-Hsien Wu, National Chiao Tung University
+\bibitem{hyval} A Structured Document Model for Authoring Video-based Hypermedia
+\bibitem{embedded} Combinando TV Interativa e Hipervídeo - André Leon S. Gradvohl & Yuzo Iano, Member, IEEE
+\bibitem{ambulant} A SMIL player for any web browser - Ombretta Gaggi and Luca Danese
 

### 2015-03-19 tese/state.tex
[ADD] added more references, forgot to say that the previuos commit has already SVG with SMIL examples

@@
   In order to create a multimedia rich hypercall, SMIL fits our goals, but it lacks on browser compatibility.
-Ambulant\cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of SMIL 3.0 specifications, it needs to be installed on browsers as a plugin.
+Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of SMIL 3.0 specifications, it needs to be installed on browsers as a plugin.
 
-  SmillingWeb attempts to implement SMIL 3.0 with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
+  SmillingWeb \cite{smillingweb} attempts to implement SMIL 3.0 with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
 But SmillingWeb isn't fully implemented yet and their scheduler engine loads the SMIL file only once, which could raise problems when leading with SMIL changes in real time.
@@
 \bibitem{natvoip} How NAT-Compatible Are VoIP Applications? Ying-Dar Lin, Chien-Chao Tseng, Cheng-Yuan Ho, and Yu-Hsien Wu, National Chiao Tung University
+\bibitem{embedded} Combinando TV Interativa e Hipervídeo - André Leon S. Gradvohl \& Yuzo Iano, Member, IEEE
 \bibitem{hyval} A Structured Document Model for Authoring Video-based Hypermedia
-\bibitem{embedded} Combinando TV Interativa e Hipervídeo - André Leon S. Gradvohl & Yuzo Iano, Member, IEEE
-\bibitem{ambulant} A SMIL player for any web browser - Ombretta Gaggi and Luca Danese
+\bibitem{ambulant} Video on the Web: Experiences from SMIL and from the Ambulant Annotator - Dick C.A. Bulterman, A.J. Jansen, and Pablo Cesar
+\bibitem{smillingweb} A SMIL player for any web browser - Ombretta Gaggi and Luca Danese
 

### 2015-03-21 state.tex
[ADD] applied template

--- 2015-03-19 tese/state.tex
@@
 
-  HyVAL\cite{hyval} was proposed for modeling composition and interaction of video, this kind of video is also called hypervideo, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
+  Hypervideo is a kind of video that contains links to any kind of hypermedia, including itself to another point in time.
+An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific point in time.
+
+  HyperCafe \cite{hypercafe} was an experimental project to expose hypervideo concepts that consisted in switching between conversations inside a cafe.
+
+  Detail-on-demand is a subset o hypervideo that allow us to obtain aditional information about something that apears along the video, like obtaining information about a painting that appears in a particular segment.
+Hyper-Hitchcock\cite{hitchcock} is an editor and player of detail-on-demand video.
+
+  HyVAL\cite{hyval} was proposed for modeling composition and interaction of video with XML, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
 This approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
@@
   SMIL was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
-With SMIL it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
+With SMIL it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
 
@@
 
-  SVG is a format that implements part of SMIL, it allows dynamic changes to inner content in real time, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
+  SVG is a format that incorporates the animation feature of SMIL.
+Currently SVG allow us to add movement and animate attributes of elements.
+When embedded on HTML, it allows dynamic changes to inner content in real time, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
 
@@
 
-  By using solid technologies like Javascript, CSS, SVG and HTML5 in a coordinated way, it's possible to raise communications to a new level.
+  By using techonologies that relies only on web standards (CSS, HTML5, Javascript, SVG), it's possible to raise communications to a new level.
 For example, with API's like WebGL, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
 Another example would be a collaborative spreadsheet using WebRTC.
-With this, hypercalls are not limited to only video, audio and text.
+With this, hypercalls are not limited to only video, audio and text, but also complex user interfaces.
 
@@
 \bibitem{smillingweb} A SMIL player for any web browser - Ombretta Gaggi and Luca Danese
-
+\bibitem{hitchcock} Creating Navigable Multi-Level Video Summaries - Frank Shipman, Andreas Girgensohn, Lynn Wilcox
+\bibitem{hypercafe} Authoring and Navigating Video in Space and Time - Nitin Sawhney, David Balcom, Ian Smith
 \end{thebibliography}

### 2015-03-21 tese/template_project_latex.tex
[ADD] applied template

--- 2015-03-21 state.tex
@@
-\documentclass[11pt,twocolumn]{article}
+\documentclass{llncs}
+\RequirePackage[english]{babel} % Se fizerem o texto em ingles
+%\usepackage[portuguese]{babel}   
+\usepackage[latin1]{inputenc}
 
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage[fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{listings}
-\usepackage{color}
-\usepackage{makeidx}
-\usepackage{blindtext}
-\usepackage{float}
-\usepackage{graphicx}
-\usepackage{mathtools}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage[hmargin=3cm,vmargin=3.5cm]{geometry}
-\usepackage{indentfirst}
-\definecolor{dkgreen}{rgb}{0,0.6,0}
-\definecolor{gray}{rgb}{0.5,0.5,0.5}
-\definecolor{mauve}{rgb}{0.58,0,0.82}
-\usepackage{eurosym}
-\usepackage{url}
+\usepackage{makeidx}  % allows for indexgeneration
+\usepackage{graphicx}  % allows for indexgeneration
+\usepackage{subfig}  % allows for indexgeneration
+\usepackage{color}  % allows for indexgeneration
 
-\lstset{frame=tb,
-  language=C,
-  aboveskip=3mm,
-  belowskip=3mm,
-  showstringspaces=false,
-  columns=flexible,
-  basicstyle={\small\ttfamily},
-  numbers=none,
-  numberstyle=\tiny\color{gray},
-  keywordstyle=\color{blue},
-  commentstyle=\color{dkgreen},
-  stringstyle=\color{mauve},
-  breaklines=true,
-  breakatwhitespace=true
-  tabsize=3
+\begin{document}
+\pagestyle{plain}
+\mainmatter              % start of the contributions
+
+\title{Normas para a realização do relatório de projecto}
+\author{%
+	Henrique Lopes Rocha \\
+	email1: henrique.rocha@ist.utl.pt \\
 }
+\institute{Instituto Superior Técnico}
 
-\usepackage{titlesec}
-\titleformat{\section}{\hrule \large\bfseries}{\thesection}{1em}{}
-
-\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC}
-\author{
-	Henrique Lopes Rocha \\ hdlopesrocha91@gmail.com
-}
-
-\date{\today}
-\begin{document}
-
-\begingroup
-\let\onecolumn\twocolumn
-
-	\maketitle
+\maketitle              % typeset the title of the contribution
 
 \begin{abstract}
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Etiam lobortis facilisis sem.
-Nullam nec mi et neque pharetra sollicitudin.
-Praesent imperdiet mi nec ante.
-Donec ullamcorper, felis non sodales commodo, lectus velit ultrices augue, a dignissim nibh lectus placerat pede.
+    In here put your abstract.
+Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
+Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.
+Neque porro quisquam est, qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem.
+\keywords{keyworkd1, keyworkd2, keyworkd3}
 \end{abstract}
-
-\endgroup
 
@@
 
-\begin{thebibliography}{9}
+\subsection{Context}   % English
 
-\bibitem{ipv4} Next Generation Internet: IPv4 Address Exhaustion, Mitigation Strategies and Implications for the U.S. - An IEEE-USA White Paper - 2009
-\bibitem{natvoip} How NAT-Compatible Are VoIP Applications? Ying-Dar Lin, Chien-Chao Tseng, Cheng-Yuan Ho, and Yu-Hsien Wu, National Chiao Tung University
-\bibitem{embedded} Combinando TV Interativa e Hipervídeo - André Leon S. Gradvohl \& Yuzo Iano, Member, IEEE
-\bibitem{hyval} A Structured Document Model for Authoring Video-based Hypermedia
-\bibitem{ambulant} Video on the Web: Experiences from SMIL and from the Ambulant Annotator - Dick C.A. Bulterman, A.J. Jansen, and Pablo Cesar
-\bibitem{smillingweb} A SMIL player for any web browser - Ombretta Gaggi and Luca Danese
-\bibitem{hitchcock} Creating Navigable Multi-Level Video Summaries - Frank Shipman, Andreas Girgensohn, Lynn Wilcox
-\bibitem{hypercafe} Authoring and Navigating Video in Space and Time - Nitin Sawhney, David Balcom, Ian Smith
-\end{thebibliography}
+\subsection{Problem Statement / Solution Statement} % English
+\subsection{Thesis Contributions} % English
+
+\subsection{Article Structure} % English
+
+\section{Arquitectura da solução}
+
+\subsection{Methodology} % Section
+
+\subsection{Planned Schedule} % English
+Nesta secção devem apresentar uma calendarização esperadas das vossas actividades.
+
+\section{Conclusions} % English
+
+\subsection{Summary} % English
+Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as contribuições definidas e a forma como estas se materializaram
+
+\section{Conclusions} % English
+Nesta secção devem ser retiradas conclusões do trabalho realizado, em face dos resultados obtidos \cite{exemplo}.
+
+\bibliographystyle{splncs03}
+\bibliography{references}
 

### 2015-03-21 tese/template_project_latex.tex
[ADD] added all sections in english

@@
 IPv4 limits the number of public addresses in such a way that today is scarse \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same IP address but different port, this technique is also know as Network Address Translation (NAT).
+One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same IP address but different port, this technique is also known as Network Address Translation (NAT).
 
@@
 
-\section{Hypermedia}
+\section{\color{red}Hypermedia}
 
@@
 \subsection{Context}   % English
-
 \subsection{Problem Statement / Solution Statement} % English
 \subsection{Thesis Contributions} % English
-
 \subsection{Article Structure} % English
-
-\section{Arquitectura da solução}
-
+% \section{Proposed Architecture} % English
 \subsection{Methodology} % Section
-
 \subsection{Planned Schedule} % English
-Nesta secção devem apresentar uma calendarização esperadas das vossas actividades.
-
 \section{Conclusions} % English
-
 \subsection{Summary} % English
-Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as contribuições definidas e a forma como estas se materializaram
-
 \section{Conclusions} % English
-Nesta secção devem ser retiradas conclusões do trabalho realizado, em face dos resultados obtidos \cite{exemplo}.
-
 \bibliographystyle{splncs03}

### 2015-03-21 tese/template_project_latex.tex
[FIX] fixed reference dates [...]

@@
 \RequirePackage[english]{babel} % Se fizerem o texto em ingles
-%\usepackage[portuguese]{babel}   
+\usepackage[english]{babel}   
 \usepackage[latin1]{inputenc}
+%\usepackage[utf8x]{inputenc}
+\usepackage[T1]{fontenc}
 
+\usepackage[fleqn]{amsmath}
+\usepackage{amssymb}
 \usepackage{makeidx}  % allows for indexgeneration
@@
 \usepackage{color}  % allows for indexgeneration
+\usepackage{url}
+\usepackage{hyperref}
 
@@
 
-\title{Normas para a realização do relatório de projecto}
+\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC}
 \author{%
@@
   Since the early days of video technology, one of the problems that raised with it consisted of how to add more information onto it without generating multiple versions.
-Some implementations like \cite{embedded} added hypermedia information to empty space present on MPEG frames, the MPEG coder and decoder were changed in order to handle hypermedia content.
+Some implementations like \cite{embedded} added hypermedia information to empty space present on MPEG frames in order to provide interactive television, the MPEG coder and decoder were changed in order to handle hypermedia content.
 
@@
 
-  HyVAL\cite{hyval} was proposed for modeling composition and interaction of video with XML, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
+  In order to navigate through a dynamic video, one must be aware of time synchronization and the multiple time flows, it's important that all time, causality and behavior rules are well defined.
+
+  HyVAL\cite{hyval} is an XML based language that was proposed for modeling composition, synchronization and interaction of video, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
 This approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
 
-  SMIL was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
+  SMIL\cite{smil} was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
 With SMIL it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
@@
   In order to create a multimedia rich hypercall, SMIL fits our goals, but it lacks on browser compatibility.
-Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of SMIL 3.0 specifications, it needs to be installed on browsers as a plugin.
+Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of SMIL 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plugin.
 
@@
 Another example would be a collaborative spreadsheet using WebRTC.
-With this, hypercalls are not limited to only video, audio and text, but also complex user interfaces.
+With this, hypercalls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
 
-  In this project our goal is to enrich hypercalls with no limits, every user should be free to choose how it wants to be contacted.
+  In this project our goal is to enrich hypercalls with no limits, every user should be free to choose how it wants to be contacted and it wants to share its contents.
 
   In order to give users a personalized communication channel, each user must have a personal web page where its available plugins could be downloaded from other peers, after that they can talk in the same language whatever it is.
+
+  \subsubsection{Local Storage}
+  \subsubsection{Video Recording}
+  \subsubsection{Video Playback - Past and Present}
 

### 2015-03-22 tese/template_project_latex.tex
[ADD] added original paper of HyperCafe

@@
 
-  Hypervideo is a kind of video that contains links to any kind of hypermedia, including itself to another point in time.
+  Hypervideo is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
 An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific point in time.
-
-  HyperCafe \cite{hypercafe} was an experimental project to expose hypervideo concepts that consisted in switching between conversations inside a cafe.
+HyperCafe \cite{hypercafe} was an experimental project to expose hypervideo concepts that consisted on an interactive film by switching between conversations inside a cafe.
 

### 2015-03-30 tese/template_project_latex.tex
[ADD] added more bullets
[ADD] added stroph.js (maybe it will be a reference)
[ADD] added text about SIP and XMPP

@@
 
+Compared to Skype, WebRTC allows web browser real time communications without installing any aditional application or plugin.
+
+\section{\color{red}Signaling}
+
+  WebRTC doesn't implement signaling, different applications may need different protocols, there is no single answer that fits all problems.
+Amongst multiple options, signaling can be done by using SIP, XMPP, WebSockets, Socket.io or by implementing a custom protocol.
+
+  %https://bloggeek.me/siganling-protocol-webrtc/
+  {\color{red}[SIP]}
+
+  SIP is a signaling method, it allows web browsers to have audio, video and SMS capabilities like mobile phones. For instance, it's possible to interoperate with SIP, mobile and fixed phones.
+
+  SIP is used in VoIP applications duo its compatibility with PSTN.
+Service providers are disponibilizing access to their SIP infrastrucures through WebSockets.
+Frameworks like JSSip or sipML5 are used on client side to encode SIP messages, making SIP accessible to web based applications.
+
+  {\color{red}[XMPP]}
+
+  HTTP works on a request followed by a server response.
+Sometimes it's required that some informations are obtained in real time, as we saw, some NAT's don't suport callbacks from servers, one technique to overcome this problem is polling.
+
+  Polling consists on sending periodic messages that the server responds imediately with empty content or fresh information.
+Simple polling wastes network bandwith and energy.
+Long polling consists on making the server wait until there is fresh information or expiring after some time.
+Long polling returns answers faster than simple polling.
+
+  XMPP's bi-directional communication is achieved through BOSH, which basically consists on long polling.
+
 \section{\color{red}Hypermedia}
@@
 
-  \subsubsection{Local Storage}
-  \subsubsection{Video Recording}
-  \subsubsection{Video Playback - Past and Present}
+  {\color{red}[Local Storage]}
+
+  {\color{red}[Video Recording]}
+
+  {\color{red}[Video Playback - Past and Present]}
 

### 2015-04-01 tese/template_project_latex.tex
[ADD] added more papers to read (about scalalbility of XMPP)

@@
 
+  {\color{orange}
+  Collaboration is the basis for joint value creation.
+With increasing complexity of tasks to conduct in business, science, and administration, efficient team collaboration has become a crucial success factor.
+
+  Software tools for collaboration typically support the exchange of data and centralized storage of information, for instance in shared workspaces.
+Only in recent years, technology became available allowing real-time collaboration via the web-browser.
+This enables team partners to edit a document concurrently and let others see changes instantly.
+  }
+
   {\color{red}[Local Storage]}

### 2015-04-02 tese/template_project_latex.tex
[ADD] acronyms

@@
 \usepackage{hyperref}
+
+\usepackage{acronym}
+\newacro{TMN}{This Means Nothing}
+\newacro{VoIP}{Voice Over IP}
+\newacro{NAT}{Network Address Translation}
+\newacro{ICE}{Interactive Connectivity Establishment}
+\newacro{STUN}{Session Traversal Utilities for NAT}
+\newacro{TURN}{Traversal Using Relays around NAT}
+\newacro{SMIL}{Synchronized Multimedia Integration Language}
+\newacro{XML}{eXtensible Markup Language}
+\newacro{HTML}{HyperText Markup Language}
+\newacro{SIP}{Session Initiation Protocol}
+\newacro{SVG}{Scalable Vector Graphics}
+\newacro{HTTP}{Hypertext Transfer Protocol}
+\newacro{API}{Application Programming Interface}
+\newacro{CSS}{Cascading Style Sheets}
+\newacro{IPv4}{Internet Protocol Version 4}
+\newacro{IPv6}{Internet Protocol Version 6}
+\newacro{IP}{Internet Protocol}
+\newacro{XMPP}{Extensible Messaging and Presence Protocol}
+\newacro{SMS}{Short Message Service}
+\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
+\newacro{MPEG}{Moving Picture Experts Group}
+\newacro{SRT}{SubRip Text}
+\newacro{SAMI}{Synchronized Accessible Media Interchange}
+\newacro{SDP}{Session Description Protocol}
+
+\overfullrule=2cm
 
@@
 IPv4 limits the number of public addresses in such a way that today is scarse \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same IP address but different port, this technique is also known as Network Address Translation (NAT).
+One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same \ac{IP} address but different port, this technique is also known as \ac{NAT}.
 
-Initially NAT offered an alternative for address exhaustion and a false sensation of better security, asymmetric NAT became a vulgar configuration on the web.
-As a direct result, problems started to appear, the amount of ports that IP disponibilizes is also small compared to our current needs, worse than that, NAT also difficult end-to-end communication, forcing most of applications that follows this model to be implemented ineffectively.
+Initially \ac{NAT} offered an alternative for address exhaustion and a false sensation of better security, asymmetric \ac{NAT} became a vulgar configuration on the web.
+As a direct result, problems started to appear, the amount of ports that \ac{IP} disponibilizes is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follows this model to be implemented ineffectively.
 
-Applications based on multimedia and file sharing were one of the most strained by NAT.
+Applications based on multimedia and file sharing were one of the most strained by \ac{NAT}.
 Those kind applications require real time communication in order to achieve the best performance.
-STUN and TURN \cite{natvoip} servers are a possible solution to overpass NAT, although, none of those can establish direct connections on multiple level NATs.
+\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overpass \ac{NAT}, although, none of those can establish direct connections on multiple level \ac{NAT}s.
 
-Most of client-server applications aren't affected by NAT when the servers are public, but they're inadequate for real time communication between two private endpoints.
+Most of client-server applications aren't affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints.
 Clearly this type of communication requires a more expensive infrastructure and, in most cases, more network usage, leading to a worse quality of service.
@@
 
-On the other hand, TURN uses public servers to redirect traffic between private endpoints, it may use a P2P network relay to find the best peer but, after that, the behavior is much like client-server.
-Direct communication is only achieved by STUN when NAT is a type \textit{full cone}.
-{\color{red} [reference to NAT types]}.
-ICE uses STUN when it's possible and TURN otherwise.
+On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a P2P network relay to find the best peer but, after that, the behavior is much like client-server.
+Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}.
+{\color{red} [reference to \ac{NAT} types]}.
+\ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.
 
-When connection is established, either in a direct or indirect way (via TURN servers), WebRTC cames to simplify how audio and video are transmited.
+When connection is established, either in a direct or indirect way (via \ac{TURN} servers), WebRTC cames to simplify how audio and video are transmited.
 
@@
   WebRTC doesn't implement signaling, different applications may need different protocols, there is no single answer that fits all problems.
-Amongst multiple options, signaling can be done by using SIP, XMPP, WebSockets, Socket.io or by implementing a custom protocol.
+Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, WebSockets, Socket.io or by implementing a custom protocol.
+
+  WebRTC uses \ac{SDP} to define peer connection properties, types of supported media, codecs, protocols used and network information.
 
@@
 
-  SIP is a signaling method, it allows web browsers to have audio, video and SMS capabilities like mobile phones. For instance, it's possible to interoperate with SIP, mobile and fixed phones.
+  \ac{SIP} can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
+For instance, it's possible to interoperate with \ac{SIP}, mobile and fixed phones.
 
-  SIP is used in VoIP applications duo its compatibility with PSTN.
-Service providers are disponibilizing access to their SIP infrastrucures through WebSockets.
-Frameworks like JSSip or sipML5 are used on client side to encode SIP messages, making SIP accessible to web based applications.
+  \ac{SIP} is used in \ac{VoIP} applications duo its compatibility with PSTN.
+Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
+Frameworks like JSSip or sipML5 are used on client side to encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
 
@@
 
-  HTTP works on a request followed by a server response.
-Sometimes it's required that some informations are obtained in real time, as we saw, some NAT's don't suport callbacks from servers, one technique to overcome this problem is polling.
+  \ac{HTTP} works on a request followed by a server response.
+Sometimes it's required that some informations are obtained in real time, as we saw, some \ac{NAT}'s don't suport callbacks from servers, one technique to overcome this problem is polling.
 
@@
 
-  XMPP's bi-directional communication is achieved through BOSH, which basically consists on long polling.
+  \ac{XMPP}s bi-directional communication is achieved through \ac{BOSH}, which basically consists on long polling.
 
@@
   Since the early days of video technology, one of the problems that raised with it consisted of how to add more information onto it without generating multiple versions.
-Some implementations like \cite{embedded} added hypermedia information to empty space present on MPEG frames in order to provide interactive television, the MPEG coder and decoder were changed in order to handle hypermedia content.
+Some implementations like \cite{embedded} added hypermedia information to empty space present on \ac{MPEG} frames in order to provide interactive television, the \ac{MPEG} coder and decoder were changed in order to handle hypermedia content.
 
@@
 
-  Amongst multiple formats for subtitles, SAMI, SRT and SUB are used by video players that support them. Although those formats have styling available, they are quite limited to text.
+  Amongst multiple formats for subtitles, \ac{SAMI}, and \ac{SRT} are used by video players that support them. Although those formats have styling available, they are quite limited to text.
 
@@
 
-  HyVAL\cite{hyval} is an XML based language that was proposed for modeling composition, synchronization and interaction of video, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
+  HyVAL\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of video, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
 This approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
 
-  SMIL\cite{smil} was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
-With SMIL it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
+  \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
+With \ac{SMIL} it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
 
-  In order to create a multimedia rich hypercall, SMIL fits our goals, but it lacks on browser compatibility.
-Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of SMIL 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plugin.
+  In order to create a multimedia rich hypercall, \ac{SMIL} fits our goals, but it lacks on browser compatibility.
+Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \ac{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plugin.
 
-  SmillingWeb \cite{smillingweb} attempts to implement SMIL 3.0 with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
-But SmillingWeb isn't fully implemented yet and their scheduler engine loads the SMIL file only once, which could raise problems when leading with SMIL changes in real time.
+  SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
+But SmillingWeb isn't fully implemented yet and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes in real time.
 
-  SVG is a format that incorporates the animation feature of SMIL.
-Currently SVG allow us to add movement and animate attributes of elements.
-When embedded on HTML, it allows dynamic changes to inner content in real time, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
+  \ac{SVG} is a format that incorporates the animation feature of \ac{SMIL}.
+Currently \ac{SVG} allow us to add movement and animate attributes of elements.
+When embedded on \ac{HTML}, it allows dynamic changes to inner content in real time, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
 
-  Video functionalities are already embedded in HTML5, like SVG it is also possible to bind javascript functions for different kinds of events over videos.
+  Video functionalities are already embedded in \ac{HTML}5, like \ac{SVG} it is also possible to bind javascript functions for different kinds of events over videos.
 
-  By using techonologies that relies only on web standards (CSS, HTML5, Javascript, SVG), it's possible to raise communications to a new level.
-For example, with API's like WebGL, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
+  By using techonologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, Javascript and \ac{SVG}, it's possible to raise communications to a new level.
+For example, with \ac{API}s like WebGL, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
 Another example would be a collaborative spreadsheet using WebRTC.
@@
 
-  {\color{orange}
+  {\color{yellow}
   Collaboration is the basis for joint value creation.

### 2015-04-03 tese/template_project_latex.tex
[ADD] added more one citation, text about polling, long polling and advantages of websockets

@@
 \newacro{SDP}{Session Description Protocol}
+\newacro{WebRTC}{Web Real-Time Communication}
+\newacro{PSTN}{Public Switched Telephone Network}
 
@@
 The need to build a global comunications network in an era when almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
-IPv4 limits the number of public addresses in such a way that today is scarse \cite{ipv4}.
+\ac{IPv4} limits the number of public addresses in such a way that today is scarse \cite{ipv4}.
 One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same \ac{IP} address but different port, this technique is also known as \ac{NAT}.
@@
 
-When connection is established, either in a direct or indirect way (via \ac{TURN} servers), WebRTC cames to simplify how audio and video are transmited.
+When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} cames to simplify how audio and video are transmited.
 
-Compared to Skype, WebRTC allows web browser real time communications without installing any aditional application or plugin.
+Compared to Skype, \ac{WebRTC} allows web browser real time communications without installing any aditional application or plugin.
 
@@
 
-  WebRTC doesn't implement signaling, different applications may need different protocols, there is no single answer that fits all problems.
+  Signaling is the most important proccess for applications to exchange connection information about peers and servers, their capabilities and metadata.
+
+  \ac{WebRTC} doesn't implement signaling, different applications may need different protocols, there is no single answer that fits all problems.
 Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, WebSockets, Socket.io or by implementing a custom protocol.
 
-  WebRTC uses \ac{SDP} to define peer connection properties, types of supported media, codecs, protocols used and network information.
+  \ac{WebRTC} uses \ac{SDP} to define peer connection properties, types of supported media, codecs, protocols used and network information.
 
-  %https://bloggeek.me/siganling-protocol-webrtc/
-  {\color{red}[SIP]}
-
-  \ac{SIP} can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
-For instance, it's possible to interoperate with \ac{SIP}, mobile and fixed phones.
-
-  \ac{SIP} is used in \ac{VoIP} applications duo its compatibility with PSTN.
-Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
-Frameworks like JSSip or sipML5 are used on client side to encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
-
-  {\color{red}[XMPP]}
-
-  \ac{HTTP} works on a request followed by a server response.
+  One of signaling requisites is bi-directional communication over \ac{HTTP}.
+\ac{HTTP} works on a request followed by a server response, by other words, follows a unidirectional communication.
 Sometimes it's required that some informations are obtained in real time, as we saw, some \ac{NAT}'s don't suport callbacks from servers, one technique to overcome this problem is polling.
@@
 Long polling returns answers faster than simple polling.
+Both with simple polling and long polling, requests are sent with \ac{HTTP} headers, which are data overkill, specialy for small messages.
 
-  \ac{XMPP}s bi-directional communication is achieved through \ac{BOSH}, which basically consists on long polling.
+  {\color{blue}Comparison of the WebSocket protocol with the traditional method like Polling and HTTP long Polling shows the benefits it offers in terms of latency and network utilization. }
+
+  {\color{blue}With polling, the browser sends HTTP requests at regular intervals and immediately receives a response.
+This technique was the first attempt for the browser to deliver real-time information.
+Obviously, it looks like a good solution if the exact interval of message delivery is known, because you can synchronize the client request to occur only when information is available on the server.
+However, real-time data is often not that predictable, making unnecessary requests inevitable and as a result, many connections are opened and closed needlessly in low-message-rate situations.}
+
+  {\color{blue}With long-polling, the browser sends a request to the server and the server keeps the request open for a set period.
+If a notification is received within that period, a response containing the message is sent to the client.
+If a notification is not received within the set time period, the server sends a response to terminate the open request.
+It is important to understand, however, that when you have a high message volume, long-polling does not provide any substantial performance improvements over traditional polling.
+Here with long polling server doesn't send empty response it holds the request until the message is available or timeout expires.}
+
+  {\color{blue}The WebSocket protocol provides full-duplex, bidirectional communication channel that operates through a single socket over the web and can help build scalable real time web application.
+Once the connection is upgraded to WebSocket, messages can flow from the server to the browser the moment they arrive.}
+
+   {\color{red}[BOSH/COMET]}
+
+  %https://bloggeek.me/siganling-protocol-webrtc/
+  {\color{red}[SIP]}
+
+  \ac{SIP} can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
+For instance, it's possible to interoperate web communications with \ac{SIP}, mobile and fixed phones.
+
+  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}.
+Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
+Frameworks like jsSIP, QoffeeSIP and sipML5 are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.
+
+  {\color{red}[XMPP]}
+
+  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH}, which basically consists on long polling.
 
@@
 
-  {\color{yellow}
-  Collaboration is the basis for joint value creation.
-With increasing complexity of tasks to conduct in business, science, and administration, efficient team collaboration has become a crucial success factor.
+  {\color{blue} Collaboration is the basis for joint value creation.
+With increasing complexity of tasks to conduct in business, science, and administration, efficient team collaboration has become a crucial success factor.}
 
-  Software tools for collaboration typically support the exchange of data and centralized storage of information, for instance in shared workspaces.
+  {\color{blue}Software tools for collaboration typically support the exchange of data and centralized storage of information, for instance in shared workspaces.
 Only in recent years, technology became available allowing real-time collaboration via the web-browser.
-This enables team partners to edit a document concurrently and let others see changes instantly.
-  }
+This enables team partners to edit a document concurrently and let others see changes instantly.}
 

### 2015-04-04 tese/template_project_latex.tex
[ADD] talking about bosh, no reference made yet, although the paper is there

@@
 \usepackage{acronym}
-\newacro{TMN}{This Means Nothing}
-\newacro{VoIP}{Voice Over IP}
-\newacro{NAT}{Network Address Translation}
+\newacro{API}{Application Programming Interface}
+\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
+\newacro{CSS}{Cascading Style Sheets}
+\newacro{HTML}{HyperText Markup Language}
+\newacro{HTTP}{Hypertext Transfer Protocol}
 \newacro{ICE}{Interactive Connectivity Establishment}
-\newacro{STUN}{Session Traversal Utilities for NAT}
-\newacro{TURN}{Traversal Using Relays around NAT}
-\newacro{SMIL}{Synchronized Multimedia Integration Language}
-\newacro{XML}{eXtensible Markup Language}
-\newacro{HTML}{HyperText Markup Language}
-\newacro{SIP}{Session Initiation Protocol}
-\newacro{SVG}{Scalable Vector Graphics}
-\newacro{HTTP}{Hypertext Transfer Protocol}
-\newacro{API}{Application Programming Interface}
-\newacro{CSS}{Cascading Style Sheets}
 \newacro{IPv4}{Internet Protocol Version 4}
@@
 \newacro{IP}{Internet Protocol}
-\newacro{XMPP}{Extensible Messaging and Presence Protocol}
-\newacro{SMS}{Short Message Service}
-\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
 \newacro{MPEG}{Moving Picture Experts Group}
-\newacro{SRT}{SubRip Text}
+\newacro{NAT}{Network Address Translation}
+\newacro{PSTN}{Public Switched Telephone Network}
 \newacro{SAMI}{Synchronized Accessible Media Interchange}
 \newacro{SDP}{Session Description Protocol}
+\newacro{SIP}{Session Initiation Protocol}
+\newacro{SMIL}{Synchronized Multimedia Integration Language}
+\newacro{SMS}{Short Message Service}
+\newacro{SRT}{SubRip Text}
+\newacro{STUN}{Session Traversal Utilities for NAT}
+\newacro{SVG}{Scalable Vector Graphics}
+\newacro{TCP}{Transmission Control Protocol}
+\newacro{TMN}{This Means Nothing}
+\newacro{TURN}{Traversal Using Relays around NAT}
+\newacro{VoIP}{Voice Over IP}
 \newacro{WebRTC}{Web Real-Time Communication}
-\newacro{PSTN}{Public Switched Telephone Network}
-
+\newacro{XML}{eXtensible Markup Language}
+\newacro{XMPP}{Extensible Messaging and Presence Protocol}
 \overfullrule=2cm
@@
   Polling consists on sending periodic messages that the server responds imediately with empty content or fresh information.
-Simple polling wastes network bandwith and energy.
-Long polling consists on making the server wait until there is fresh information or expiring after some time.
-Long polling returns answers faster than simple polling.
-Both with simple polling and long polling, requests are sent with \ac{HTTP} headers, which are data overkill, specialy for small messages.
+Because real time communications are unpredictable, if the time between periodic requests is short, most of time the server will return empty results wasting network bandwith and energy.
+On the other hand, if the time between periodic requests is large, newer messages may arrive later.
 
-  {\color{blue}Comparison of the WebSocket protocol with the traditional method like Polling and HTTP long Polling shows the benefits it offers in terms of latency and network utilization. }
+  A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the receival, the client makes another request.
+This results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which are data overkill, specialy for small messages.
 
-  {\color{blue}With polling, the browser sends HTTP requests at regular intervals and immediately receives a response.
-This technique was the first attempt for the browser to deliver real-time information.
-Obviously, it looks like a good solution if the exact interval of message delivery is known, because you can synchronize the client request to occur only when information is available on the server.
-However, real-time data is often not that predictable, making unnecessary requests inevitable and as a result, many connections are opened and closed needlessly in low-message-rate situations.}
+  WebSockets provide bidirectional communications trough a full-duplex socket channel.
+WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
 
-  {\color{blue}With long-polling, the browser sends a request to the server and the server keeps the request open for a set period.
-If a notification is received within that period, a response containing the message is sent to the client.
-If a notification is not received within the set time period, the server sends a response to terminate the open request.
-It is important to understand, however, that when you have a high message volume, long-polling does not provide any substantial performance improvements over traditional polling.
-Here with long polling server doesn't send empty response it holds the request until the message is available or timeout expires.}
-
-  {\color{blue}The WebSocket protocol provides full-duplex, bidirectional communication channel that operates through a single socket over the web and can help build scalable real time web application.
-Once the connection is upgraded to WebSocket, messages can flow from the server to the browser the moment they arrive.}
-
-   {\color{red}[BOSH/COMET]}
+  \ac{BOSH} is a technique based on long polling, that uses two socket connections and allow sending client messages to server while a previous request is holded.
 

### 2015-04-04 tese/template_project_latex.tex
[ADD] added more stuff about XMPP

@@
 WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
+WebSockets may not be avalailable on every web browser, frameworks like \textit{socket.io} and \textit{SockJS} uses \ac{HTTP} when there is no support for WebSockets.
 
@@
 
-  %https://bloggeek.me/siganling-protocol-webrtc/
-  {\color{red}[SIP]}
+  {\color{blue} \ac{SIP} is a communications protocol for signaling and controlling multimedia communication sessions.
+The most common applications of SIP are in Internet telephony for voice and video calls, as well as instant messaging all over \ac{IP} networks.}
 
-  \ac{SIP} can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
+  \ac{SIP} with WebSockets can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
 For instance, it's possible to interoperate web communications with \ac{SIP}, mobile and fixed phones.
@@
 Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
-Frameworks like jsSIP, QoffeeSIP and sipML5 are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.
+Frameworks like \textit{jsSIP}, \textit{QoffeeSIP} and \textit{sipML5} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.
 
-  {\color{red}[XMPP]}
+  {\color{red}[Jitsi]}
+
+  {\color{blue} \ac{XMPP} is an open technology for real-time communication, which powers a wide range of applications including instant messaging, presence, multi-party chat, voice and video calls, collaboration, lightweight middleware, content syndication, and generalized routing of XML data.
+The technology pages provide more information about the various XMPP "building blocks".
+Several books about Jabber/XMPP technologies are available, as well.}
 

### 2015-04-05 tese/template_project_latex.tex
[ADD] added a very important reference, must read as soon as possible

@@
   A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the receival, the client makes another request.
-This results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which are data overkill, specialy for small messages.
+This results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which adds data overhead, specialy for small messages.
 

### 2015-04-06 tese/template_project_latex.tex
[ADD] almost finished XMPP

--- 2015-04-04 tese/template_project_latex.tex
@@
 \newacro{PSTN}{Public Switched Telephone Network}
+\newacro{RFC}{Request For Comments}
 \newacro{SAMI}{Synchronized Accessible Media Interchange}
@@
 \newacro{SIP}{Session Initiation Protocol}
+\newacro{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
 \newacro{SMIL}{Synchronized Multimedia Integration Language}
@@
 \newacro{WebRTC}{Web Real-Time Communication}
+\newacro{XEP}{XMPP Extensions}
 \newacro{XML}{eXtensible Markup Language}
@@
 WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
+WebSockets may not be avalailable on every web browser, frameworks like \textit{socket.io}\footnote{\url{http://socket.io/}} and \textit{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.
 
@@
 
-  %https://bloggeek.me/siganling-protocol-webrtc/
-  {\color{red}[SIP]}
+  \ac{SIP} is protocol used for negotiation, creation, modification and finalization of communication sessions between users.
+\ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as signaling protocol.
+The advantage of \ac{SIP} is the ability to make video and voice calls applications over \ac{IP} networks.
 
-  \ac{SIP} can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
-For instance, it's possible to interoperate web communications with \ac{SIP}, mobile and fixed phones.
+  The work group \ac{SIMPLE} proposed the creation of \ac{SIP} extensions, namely presence information (rfc5263\footnote{\url{http://tools.ietf.org/html/rfc5263}}) and instant messaging (rfc3428\footnote{\url{http://tools.ietf.org/html/rfc3428}}).
 
@@
 Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
-Frameworks like jsSIP, QoffeeSIP and sipML5 are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.
+Frameworks like \textit{jsSIP}\footnote{\url{http://jssip.net/}}, \textit{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \textit{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.
 
-  {\color{red}[XMPP]}
+  \ac{SIP} with WebSockets can be used as a signaling method for WebRTC applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
+For instance, it's possible to interoperate web communications with \ac{SIP} networks, mobile and fixed phones.
 
-  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH}, which basically consists on long polling.
+  \ac{XMPP} was initially developed for instant messaging (Jabber\footnote{\url{http://jabber.org/}}). It is nowadays an open technology for real-time communications.
+
+  \ac{XMPP} messages are \ac{XML} based, which are attractive for applications that need structured messages and rich hypermedia.
+\ac{XMPP} advantage is the addition of extensions, for example XEP-0096\footnote{\url{http://xmpp.org/extensions/xep-0096.html}}, which adds file transfer capabilities between two entities and XEP-0045\footnote{\url{http://xmpp.org/extensions/xep-0045.html}} which enables multi-user chat.
+
+  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH} (XEP-0206\footnote{\url{http://xmpp.org/extensions/xep-0206.html}}), which basically consists on long polling.
+This kind of communication is also possible through WebSockets (rfc7395\footnote{\url{http://tools.ietf.org/html/rfc7395}}).
+
+  Amongst multiple XMPP servers softwares are: ejabberd\footnote{\url{http://jabberd.im/}}, Metronome\footnote{\url{http://lightwitch.org/metronome}}, Openfire\footnote{\url{http://igniterealtime.org/projects/openfire/}} and Prosody\footnote{\url{http://prosody.im/}}.
+Ejabberd is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
 

### 2015-04-07 tese/section01.tex
[RFT] added a file per section

--- 2015-04-04 tese/template_project_latex.tex
@@
-\documentclass{llncs}
-\RequirePackage[english]{babel} % Se fizerem o texto em ingles
-\usepackage[english]{babel}   
-\usepackage[latin1]{inputenc}
-%\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-
-\usepackage[fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{graphicx}  % allows for indexgeneration
-\usepackage{subfig}  % allows for indexgeneration
-\usepackage{color}  % allows for indexgeneration
-\usepackage{url}
-\usepackage{hyperref}
-
-\usepackage{acronym}
-\newacro{API}{Application Programming Interface}
-\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
-\newacro{CSS}{Cascading Style Sheets}
-\newacro{HTML}{HyperText Markup Language}
-\newacro{HTTP}{Hypertext Transfer Protocol}
-\newacro{ICE}{Interactive Connectivity Establishment}
-\newacro{IPv4}{Internet Protocol Version 4}
-\newacro{IPv6}{Internet Protocol Version 6}
-\newacro{IP}{Internet Protocol}
-\newacro{MPEG}{Moving Picture Experts Group}
-\newacro{NAT}{Network Address Translation}
-\newacro{PSTN}{Public Switched Telephone Network}
-\newacro{SAMI}{Synchronized Accessible Media Interchange}
-\newacro{SDP}{Session Description Protocol}
-\newacro{SIP}{Session Initiation Protocol}
-\newacro{SMIL}{Synchronized Multimedia Integration Language}
-\newacro{SMS}{Short Message Service}
-\newacro{SRT}{SubRip Text}
-\newacro{STUN}{Session Traversal Utilities for NAT}
-\newacro{SVG}{Scalable Vector Graphics}
-\newacro{TCP}{Transmission Control Protocol}
-\newacro{TMN}{This Means Nothing}
-\newacro{TURN}{Traversal Using Relays around NAT}
-\newacro{VoIP}{Voice Over IP}
-\newacro{WebRTC}{Web Real-Time Communication}
-\newacro{XML}{eXtensible Markup Language}
-\newacro{XMPP}{Extensible Messaging and Presence Protocol}
-\overfullrule=2cm
-
-\begin{document}
-\pagestyle{plain}
-\mainmatter              % start of the contributions
-
-\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC}
-\author{%
-	Henrique Lopes Rocha \\
-	email1: henrique.rocha@ist.utl.pt \\
-}
-\institute{Instituto Superior Técnico}
-
-\maketitle              % typeset the title of the contribution
-
-\begin{abstract}
-    In here put your abstract.
-Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
-Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.
-Neque porro quisquam est, qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem.
-\keywords{keyworkd1, keyworkd2, keyworkd3}
-\end{abstract}
-
-\section{Introduction}
 The need to build a global comunications network in an era when almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
@@
 Compared to Skype, \ac{WebRTC} allows web browser real time communications without installing any aditional application or plugin.
-
-\section{\color{red}Signaling}
-
-  Signaling is the most important proccess for applications to exchange connection information about peers and servers, their capabilities and metadata.
-
-  \ac{WebRTC} doesn't implement signaling, different applications may need different protocols, there is no single answer that fits all problems.
-Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, WebSockets, Socket.io or by implementing a custom protocol.
-
-  \ac{WebRTC} uses \ac{SDP} to define peer connection properties, types of supported media, codecs, protocols used and network information.
-
-  One of signaling requisites is bi-directional communication over \ac{HTTP}.
-\ac{HTTP} works on a request followed by a server response, by other words, follows a unidirectional communication.
-Sometimes it's required that some informations are obtained in real time, as we saw, some \ac{NAT}'s don't suport callbacks from servers, one technique to overcome this problem is polling.
-
-  Polling consists on sending periodic messages that the server responds imediately with empty content or fresh information.
-Because real time communications are unpredictable, if the time between periodic requests is short, most of time the server will return empty results wasting network bandwith and energy.
-On the other hand, if the time between periodic requests is large, newer messages may arrive later.
-
-  A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the receival, the client makes another request.
-This results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which are data overkill, specialy for small messages.
-
-  WebSockets provide bidirectional communications trough a full-duplex socket channel.
-WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
-WebSockets may not be avalailable on every web browser, frameworks like \textit{socket.io} and \textit{SockJS} uses \ac{HTTP} when there is no support for WebSockets.
-
-  \ac{BOSH} is a technique based on long polling, that uses two socket connections and allow sending client messages to server while a previous request is holded.
-
-  {\color{blue} \ac{SIP} is a communications protocol for signaling and controlling multimedia communication sessions.
-The most common applications of SIP are in Internet telephony for voice and video calls, as well as instant messaging all over \ac{IP} networks.}
-
-  \ac{SIP} with WebSockets can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
-For instance, it's possible to interoperate web communications with \ac{SIP}, mobile and fixed phones.
-
-  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}.
-Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
-Frameworks like \textit{jsSIP}, \textit{QoffeeSIP} and \textit{sipML5} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.
-
-  {\color{red}[Jitsi]}
-
-  {\color{blue} \ac{XMPP} is an open technology for real-time communication, which powers a wide range of applications including instant messaging, presence, multi-party chat, voice and video calls, collaboration, lightweight middleware, content syndication, and generalized routing of XML data.
-The technology pages provide more information about the various XMPP "building blocks".
-Several books about Jabber/XMPP technologies are available, as well.}
-
-  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH}, which basically consists on long polling.
-
-\section{\color{red}Hypermedia}
-
-  Since the early days of video technology, one of the problems that raised with it consisted of how to add more information onto it without generating multiple versions.
-Some implementations like \cite{embedded} added hypermedia information to empty space present on \ac{MPEG} frames in order to provide interactive television, the \ac{MPEG} coder and decoder were changed in order to handle hypermedia content.
-
-  The need to translate movies, raised the problem whether it is apropriate to change the original video or audio.
-For example subtitles should be an entity independent from the video, in order to be personalized or replaced easily.
-
-  Amongst multiple formats for subtitles, \ac{SAMI}, and \ac{SRT} are used by video players that support them. Although those formats have styling available, they are quite limited to text.
-
-  Hypervideo is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
-An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific point in time.
-HyperCafe \cite{hypercafe} was an experimental project to expose hypervideo concepts that consisted on an interactive film by switching between conversations inside a cafe.
-
-  Detail-on-demand is a subset o hypervideo that allow us to obtain aditional information about something that apears along the video, like obtaining information about a painting that appears in a particular segment.
-Hyper-Hitchcock\cite{hitchcock} is an editor and player of detail-on-demand video.
-
-  In order to navigate through a dynamic video, one must be aware of time synchronization and the multiple time flows, it's important that all time, causality and behavior rules are well defined.
-
-  HyVAL\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of video, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
-This approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
-
-  \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
-With \ac{SMIL} it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
-
-  In order to create a multimedia rich hypercall, \ac{SMIL} fits our goals, but it lacks on browser compatibility.
-Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \ac{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plugin.
-
-  SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
-But SmillingWeb isn't fully implemented yet and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes in real time.
-
-  \ac{SVG} is a format that incorporates the animation feature of \ac{SMIL}.
-Currently \ac{SVG} allow us to add movement and animate attributes of elements.
-When embedded on \ac{HTML}, it allows dynamic changes to inner content in real time, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
-
-  Video functionalities are already embedded in \ac{HTML}5, like \ac{SVG} it is also possible to bind javascript functions for different kinds of events over videos.
-
-  By using techonologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, Javascript and \ac{SVG}, it's possible to raise communications to a new level.
-For example, with \ac{API}s like WebGL, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
-Another example would be a collaborative spreadsheet using WebRTC.
-With this, hypercalls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
-
-  In this project our goal is to enrich hypercalls with no limits, every user should be free to choose how it wants to be contacted and it wants to share its contents.
-
-  In order to give users a personalized communication channel, each user must have a personal web page where its available plugins could be downloaded from other peers, after that they can talk in the same language whatever it is.
-
-  {\color{blue} Collaboration is the basis for joint value creation.
-With increasing complexity of tasks to conduct in business, science, and administration, efficient team collaboration has become a crucial success factor.}
-
-  {\color{blue}Software tools for collaboration typically support the exchange of data and centralized storage of information, for instance in shared workspaces.
-Only in recent years, technology became available allowing real-time collaboration via the web-browser.
-This enables team partners to edit a document concurrently and let others see changes instantly.}
-
-  {\color{red}[Local Storage]}
-
-  {\color{red}[Video Recording]}
-
-  {\color{red}[Video Playback - Past and Present]}
-
-\subsection{Context}   % English
-\subsection{Problem Statement / Solution Statement} % English
-\subsection{Thesis Contributions} % English
-\subsection{Article Structure} % English
-% \section{Proposed Architecture} % English
-\subsection{Methodology} % Section
-\subsection{Planned Schedule} % English
-\section{Conclusions} % English
-\subsection{Summary} % English
-\section{Conclusions} % English
-\bibliographystyle{splncs03}
-\bibliography{references}
-
-\end{document}

### 2015-04-07 tese/section02.tex
[RFT] added a file per section

--- 2015-04-06 tese/template_project_latex.tex
@@
-\documentclass{llncs}
-\RequirePackage[english]{babel} % Se fizerem o texto em ingles
-\usepackage[english]{babel}   
-\usepackage[latin1]{inputenc}
-%\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-
-\usepackage[fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{graphicx}  % allows for indexgeneration
-\usepackage{subfig}  % allows for indexgeneration
-\usepackage{color}  % allows for indexgeneration
-\usepackage{url}
-\usepackage{hyperref}
-
-\usepackage{acronym}
-\newacro{API}{Application Programming Interface}
-\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
-\newacro{CSS}{Cascading Style Sheets}
-\newacro{HTML}{HyperText Markup Language}
-\newacro{HTTP}{Hypertext Transfer Protocol}
-\newacro{ICE}{Interactive Connectivity Establishment}
-\newacro{IPv4}{Internet Protocol Version 4}
-\newacro{IPv6}{Internet Protocol Version 6}
-\newacro{IP}{Internet Protocol}
-\newacro{MPEG}{Moving Picture Experts Group}
-\newacro{NAT}{Network Address Translation}
-\newacro{PSTN}{Public Switched Telephone Network}
-\newacro{RFC}{Request For Comments}
-\newacro{SAMI}{Synchronized Accessible Media Interchange}
-\newacro{SDP}{Session Description Protocol}
-\newacro{SIP}{Session Initiation Protocol}
-\newacro{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
-\newacro{SMIL}{Synchronized Multimedia Integration Language}
-\newacro{SMS}{Short Message Service}
-\newacro{SRT}{SubRip Text}
-\newacro{STUN}{Session Traversal Utilities for NAT}
-\newacro{SVG}{Scalable Vector Graphics}
-\newacro{TCP}{Transmission Control Protocol}
-\newacro{TMN}{This Means Nothing}
-\newacro{TURN}{Traversal Using Relays around NAT}
-\newacro{VoIP}{Voice Over IP}
-\newacro{WebRTC}{Web Real-Time Communication}
-\newacro{XEP}{XMPP Extensions}
-\newacro{XML}{eXtensible Markup Language}
-\newacro{XMPP}{Extensible Messaging and Presence Protocol}
-\overfullrule=2cm
-
-\begin{document}
-\pagestyle{plain}
-\mainmatter              % start of the contributions
-
-\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC}
-\author{%
-	Henrique Lopes Rocha \\
-	email1: henrique.rocha@ist.utl.pt \\
-}
-\institute{Instituto Superior Técnico}
-
-\maketitle              % typeset the title of the contribution
-
-\begin{abstract}
-    In here put your abstract.
-Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
-Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.
-Neque porro quisquam est, qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem.
-\keywords{keyworkd1, keyworkd2, keyworkd3}
-\end{abstract}
-
-\section{Introduction}
-The need to build a global comunications network in an era when almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
-\ac{IPv4} limits the number of public addresses in such a way that today is scarse \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same \ac{IP} address but different port, this technique is also known as \ac{NAT}.
-
-Initially \ac{NAT} offered an alternative for address exhaustion and a false sensation of better security, asymmetric \ac{NAT} became a vulgar configuration on the web.
-As a direct result, problems started to appear, the amount of ports that \ac{IP} disponibilizes is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follows this model to be implemented ineffectively.
-
-Applications based on multimedia and file sharing were one of the most strained by \ac{NAT}.
-Those kind applications require real time communication in order to achieve the best performance.
-\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overpass \ac{NAT}, although, none of those can establish direct connections on multiple level \ac{NAT}s.
-
-Most of client-server applications aren't affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints.
-Clearly this type of communication requires a more expensive infrastructure and, in most cases, more network usage, leading to a worse quality of service.
-The requirements of video communication makes this kind of model unsuitable.
-
-On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a P2P network relay to find the best peer but, after that, the behavior is much like client-server.
-Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}.
-{\color{red} [reference to \ac{NAT} types]}.
-\ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.
-
-When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} cames to simplify how audio and video are transmited.
-
-Compared to Skype, \ac{WebRTC} allows web browser real time communications without installing any aditional application or plugin.
-
-\section{\color{red}Signaling}
-
 Signaling is the most important proccess for applications to exchange connection information about peers and servers, their capabilities and metadata.
@@
 Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
-Frameworks like \textit{jsSIP}\footnote{\url{http://jssip.net/}}, \textit{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \textit{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.
+Frameworks like \textit{jsSIP}\footnote{\url{http://jssip.net/}}, \textit{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \textit{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
 
@@
 Ejabberd is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
-
-\section{\color{red}Hypermedia}
-
-  Since the early days of video technology, one of the problems that raised with it consisted of how to add more information onto it without generating multiple versions.
-Some implementations like \cite{embedded} added hypermedia information to empty space present on \ac{MPEG} frames in order to provide interactive television, the \ac{MPEG} coder and decoder were changed in order to handle hypermedia content.
-
-  The need to translate movies, raised the problem whether it is apropriate to change the original video or audio.
-For example subtitles should be an entity independent from the video, in order to be personalized or replaced easily.
-
-  Amongst multiple formats for subtitles, \ac{SAMI}, and \ac{SRT} are used by video players that support them. Although those formats have styling available, they are quite limited to text.
-
-  Hypervideo is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
-An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific point in time.
-HyperCafe \cite{hypercafe} was an experimental project to expose hypervideo concepts that consisted on an interactive film by switching between conversations inside a cafe.
-
-  Detail-on-demand is a subset o hypervideo that allow us to obtain aditional information about something that apears along the video, like obtaining information about a painting that appears in a particular segment.
-Hyper-Hitchcock\cite{hitchcock} is an editor and player of detail-on-demand video.
-
-  In order to navigate through a dynamic video, one must be aware of time synchronization and the multiple time flows, it's important that all time, causality and behavior rules are well defined.
-
-  HyVAL\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of video, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
-This approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
-
-  \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
-With \ac{SMIL} it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
-
-  In order to create a multimedia rich hypercall, \ac{SMIL} fits our goals, but it lacks on browser compatibility.
-Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \ac{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plugin.
-
-  SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
-But SmillingWeb isn't fully implemented yet and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes in real time.
-
-  \ac{SVG} is a format that incorporates the animation feature of \ac{SMIL}.
-Currently \ac{SVG} allow us to add movement and animate attributes of elements.
-When embedded on \ac{HTML}, it allows dynamic changes to inner content in real time, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
-
-  Video functionalities are already embedded in \ac{HTML}5, like \ac{SVG} it is also possible to bind javascript functions for different kinds of events over videos.
-
-  By using techonologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, Javascript and \ac{SVG}, it's possible to raise communications to a new level.
-For example, with \ac{API}s like WebGL, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
-Another example would be a collaborative spreadsheet using WebRTC.
-With this, hypercalls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
-
-  In this project our goal is to enrich hypercalls with no limits, every user should be free to choose how it wants to be contacted and it wants to share its contents.
-
-  In order to give users a personalized communication channel, each user must have a personal web page where its available plugins could be downloaded from other peers, after that they can talk in the same language whatever it is.
-
-  {\color{blue} Collaboration is the basis for joint value creation.
-With increasing complexity of tasks to conduct in business, science, and administration, efficient team collaboration has become a crucial success factor.}
-
-  {\color{blue}Software tools for collaboration typically support the exchange of data and centralized storage of information, for instance in shared workspaces.
-Only in recent years, technology became available allowing real-time collaboration via the web-browser.
-This enables team partners to edit a document concurrently and let others see changes instantly.}
-
-  {\color{red}[Local Storage]}
-
-  {\color{red}[Video Recording]}
-
-  {\color{red}[Video Playback - Past and Present]}
-
-\subsection{Context}   % English
-\subsection{Problem Statement / Solution Statement} % English
-\subsection{Thesis Contributions} % English
-\subsection{Article Structure} % English
-% \section{Proposed Architecture} % English
-\subsection{Methodology} % Section
-\subsection{Planned Schedule} % English
-\section{Conclusions} % English
-\subsection{Summary} % English
-\section{Conclusions} % English
-\bibliographystyle{splncs03}
-\bibliography{references}
-
-\end{document}

### 2015-04-07 tese/section03.tex
[RFT] added a file per section

--- 2015-04-04 tese/template_project_latex.tex
@@
-\documentclass{llncs}
-\RequirePackage[english]{babel} % Se fizerem o texto em ingles
-\usepackage[english]{babel}   
-\usepackage[latin1]{inputenc}
-%\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-
-\usepackage[fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{graphicx}  % allows for indexgeneration
-\usepackage{subfig}  % allows for indexgeneration
-\usepackage{color}  % allows for indexgeneration
-\usepackage{url}
-\usepackage{hyperref}
-
-\usepackage{acronym}
-\newacro{API}{Application Programming Interface}
-\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
-\newacro{CSS}{Cascading Style Sheets}
-\newacro{HTML}{HyperText Markup Language}
-\newacro{HTTP}{Hypertext Transfer Protocol}
-\newacro{ICE}{Interactive Connectivity Establishment}
-\newacro{IPv4}{Internet Protocol Version 4}
-\newacro{IPv6}{Internet Protocol Version 6}
-\newacro{IP}{Internet Protocol}
-\newacro{MPEG}{Moving Picture Experts Group}
-\newacro{NAT}{Network Address Translation}
-\newacro{PSTN}{Public Switched Telephone Network}
-\newacro{SAMI}{Synchronized Accessible Media Interchange}
-\newacro{SDP}{Session Description Protocol}
-\newacro{SIP}{Session Initiation Protocol}
-\newacro{SMIL}{Synchronized Multimedia Integration Language}
-\newacro{SMS}{Short Message Service}
-\newacro{SRT}{SubRip Text}
-\newacro{STUN}{Session Traversal Utilities for NAT}
-\newacro{SVG}{Scalable Vector Graphics}
-\newacro{TCP}{Transmission Control Protocol}
-\newacro{TMN}{This Means Nothing}
-\newacro{TURN}{Traversal Using Relays around NAT}
-\newacro{VoIP}{Voice Over IP}
-\newacro{WebRTC}{Web Real-Time Communication}
-\newacro{XML}{eXtensible Markup Language}
-\newacro{XMPP}{Extensible Messaging and Presence Protocol}
-\overfullrule=2cm
-
-\begin{document}
-\pagestyle{plain}
-\mainmatter              % start of the contributions
-
-\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC}
-\author{%
-	Henrique Lopes Rocha \\
-	email1: henrique.rocha@ist.utl.pt \\
-}
-\institute{Instituto Superior Técnico}
-
-\maketitle              % typeset the title of the contribution
-
-\begin{abstract}
-    In here put your abstract.
-Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
-Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.
-Neque porro quisquam est, qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem.
-\keywords{keyworkd1, keyworkd2, keyworkd3}
-\end{abstract}
-
-\section{Introduction}
-The need to build a global comunications network in an era when almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
-\ac{IPv4} limits the number of public addresses in such a way that today is scarse \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same \ac{IP} address but different port, this technique is also known as \ac{NAT}.
-
-Initially \ac{NAT} offered an alternative for address exhaustion and a false sensation of better security, asymmetric \ac{NAT} became a vulgar configuration on the web.
-As a direct result, problems started to appear, the amount of ports that \ac{IP} disponibilizes is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follows this model to be implemented ineffectively.
-
-Applications based on multimedia and file sharing were one of the most strained by \ac{NAT}.
-Those kind applications require real time communication in order to achieve the best performance.
-\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overpass \ac{NAT}, although, none of those can establish direct connections on multiple level \ac{NAT}s.
-
-Most of client-server applications aren't affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints.
-Clearly this type of communication requires a more expensive infrastructure and, in most cases, more network usage, leading to a worse quality of service.
-The requirements of video communication makes this kind of model unsuitable.
-
-On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a P2P network relay to find the best peer but, after that, the behavior is much like client-server.
-Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}.
-{\color{red} [reference to \ac{NAT} types]}.
-\ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.
-
-When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} cames to simplify how audio and video are transmited.
-
-Compared to Skype, \ac{WebRTC} allows web browser real time communications without installing any aditional application or plugin.
-
-\section{\color{red}Signaling}
-
-  Signaling is the most important proccess for applications to exchange connection information about peers and servers, their capabilities and metadata.
-
-  \ac{WebRTC} doesn't implement signaling, different applications may need different protocols, there is no single answer that fits all problems.
-Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, WebSockets, Socket.io or by implementing a custom protocol.
-
-  \ac{WebRTC} uses \ac{SDP} to define peer connection properties, types of supported media, codecs, protocols used and network information.
-
-  One of signaling requisites is bi-directional communication over \ac{HTTP}.
-\ac{HTTP} works on a request followed by a server response, by other words, follows a unidirectional communication.
-Sometimes it's required that some informations are obtained in real time, as we saw, some \ac{NAT}'s don't suport callbacks from servers, one technique to overcome this problem is polling.
-
-  Polling consists on sending periodic messages that the server responds imediately with empty content or fresh information.
-Because real time communications are unpredictable, if the time between periodic requests is short, most of time the server will return empty results wasting network bandwith and energy.
-On the other hand, if the time between periodic requests is large, newer messages may arrive later.
-
-  A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the receival, the client makes another request.
-This results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which are data overkill, specialy for small messages.
-
-  WebSockets provide bidirectional communications trough a full-duplex socket channel.
-WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
-WebSockets may not be avalailable on every web browser, frameworks like \textit{socket.io} and \textit{SockJS} uses \ac{HTTP} when there is no support for WebSockets.
-
-  \ac{BOSH} is a technique based on long polling, that uses two socket connections and allow sending client messages to server while a previous request is holded.
-
-  {\color{blue} \ac{SIP} is a communications protocol for signaling and controlling multimedia communication sessions.
-The most common applications of SIP are in Internet telephony for voice and video calls, as well as instant messaging all over \ac{IP} networks.}
-
-  \ac{SIP} with WebSockets can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
-For instance, it's possible to interoperate web communications with \ac{SIP}, mobile and fixed phones.
-
-  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}.
-Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
-Frameworks like \textit{jsSIP}, \textit{QoffeeSIP} and \textit{sipML5} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.
-
-  {\color{red}[Jitsi]}
-
-  {\color{blue} \ac{XMPP} is an open technology for real-time communication, which powers a wide range of applications including instant messaging, presence, multi-party chat, voice and video calls, collaboration, lightweight middleware, content syndication, and generalized routing of XML data.
-The technology pages provide more information about the various XMPP "building blocks".
-Several books about Jabber/XMPP technologies are available, as well.}
-
-  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH}, which basically consists on long polling.
-
-\section{\color{red}Hypermedia}
-
   Since the early days of video technology, one of the problems that raised with it consisted of how to add more information onto it without generating multiple versions.
@@
   By using techonologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, Javascript and \ac{SVG}, it's possible to raise communications to a new level.
-For example, with \ac{API}s like WebGL, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
+For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
 Another example would be a collaborative spreadsheet using WebRTC.
@@
   In order to give users a personalized communication channel, each user must have a personal web page where its available plugins could be downloaded from other peers, after that they can talk in the same language whatever it is.
-
-  {\color{blue} Collaboration is the basis for joint value creation.
-With increasing complexity of tasks to conduct in business, science, and administration, efficient team collaboration has become a crucial success factor.}
-
-  {\color{blue}Software tools for collaboration typically support the exchange of data and centralized storage of information, for instance in shared workspaces.
-Only in recent years, technology became available allowing real-time collaboration via the web-browser.
-This enables team partners to edit a document concurrently and let others see changes instantly.}
-
-  {\color{red}[Local Storage]}
-
-  {\color{red}[Video Recording]}
-
-  {\color{red}[Video Playback - Past and Present]}
-
-\subsection{Context}   % English
-\subsection{Problem Statement / Solution Statement} % English
-\subsection{Thesis Contributions} % English
-\subsection{Article Structure} % English
-% \section{Proposed Architecture} % English
-\subsection{Methodology} % Section
-\subsection{Planned Schedule} % English
-\section{Conclusions} % English
-\subsection{Summary} % English
-\section{Conclusions} % English
-\bibliographystyle{splncs03}
-\bibliography{references}
-
-\end{document}

### 2015-04-07 tese/section04.tex
[RFT] added a file per section

  Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity. Technologies like this are appearing along this days, but they couln't be possible years ago because technology was limited or unavailable. Although todays technology is limited on some aspects, we are doing progress in order to improve the web ecosystem, by creating standards and migrating to newer technologies.

  Our first concern on real time collaboration applications, besides the communication itself, is the data storage and representation. Because most browsers are recommended to limit local storage to at least five megabytes per origin, storing multimedia content is not a viable solution.

  If, for instance, one whants to rewind a real time video, recordings will be needed from who is streaming the video.

  {\color{red}[Video Recording]}

  {\color{red}[Video Playback - Past and Present]}

### 2015-04-07 tese/template_project_latex.tex
[RFT] added a file per section

--- 2015-04-04 tese/template_project_latex.tex
@@
 \section{Introduction}
-The need to build a global comunications network in an era when almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
-\ac{IPv4} limits the number of public addresses in such a way that today is scarse \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same \ac{IP} address but different port, this technique is also known as \ac{NAT}.
-
-Initially \ac{NAT} offered an alternative for address exhaustion and a false sensation of better security, asymmetric \ac{NAT} became a vulgar configuration on the web.
-As a direct result, problems started to appear, the amount of ports that \ac{IP} disponibilizes is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follows this model to be implemented ineffectively.
-
-Applications based on multimedia and file sharing were one of the most strained by \ac{NAT}.
-Those kind applications require real time communication in order to achieve the best performance.
-\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overpass \ac{NAT}, although, none of those can establish direct connections on multiple level \ac{NAT}s.
-
-Most of client-server applications aren't affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints.
-Clearly this type of communication requires a more expensive infrastructure and, in most cases, more network usage, leading to a worse quality of service.
-The requirements of video communication makes this kind of model unsuitable.
-
-On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a P2P network relay to find the best peer but, after that, the behavior is much like client-server.
-Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}.
-{\color{red} [reference to \ac{NAT} types]}.
-\ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.
-
-When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} cames to simplify how audio and video are transmited.
-
-Compared to Skype, \ac{WebRTC} allows web browser real time communications without installing any aditional application or plugin.
-
+  \input{section01.tex}
 \section{\color{red}Signaling}
-
-  Signaling is the most important proccess for applications to exchange connection information about peers and servers, their capabilities and metadata.
-
-  \ac{WebRTC} doesn't implement signaling, different applications may need different protocols, there is no single answer that fits all problems.
-Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, WebSockets, Socket.io or by implementing a custom protocol.
-
-  \ac{WebRTC} uses \ac{SDP} to define peer connection properties, types of supported media, codecs, protocols used and network information.
-
-  One of signaling requisites is bi-directional communication over \ac{HTTP}.
-\ac{HTTP} works on a request followed by a server response, by other words, follows a unidirectional communication.
-Sometimes it's required that some informations are obtained in real time, as we saw, some \ac{NAT}'s don't suport callbacks from servers, one technique to overcome this problem is polling.
-
-  Polling consists on sending periodic messages that the server responds imediately with empty content or fresh information.
-Because real time communications are unpredictable, if the time between periodic requests is short, most of time the server will return empty results wasting network bandwith and energy.
-On the other hand, if the time between periodic requests is large, newer messages may arrive later.
-
-  A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the receival, the client makes another request.
-This results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which are data overkill, specialy for small messages.
-
-  WebSockets provide bidirectional communications trough a full-duplex socket channel.
-WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
-WebSockets may not be avalailable on every web browser, frameworks like \textit{socket.io} and \textit{SockJS} uses \ac{HTTP} when there is no support for WebSockets.
-
-  \ac{BOSH} is a technique based on long polling, that uses two socket connections and allow sending client messages to server while a previous request is holded.
-
-  {\color{blue} \ac{SIP} is a communications protocol for signaling and controlling multimedia communication sessions.
-The most common applications of SIP are in Internet telephony for voice and video calls, as well as instant messaging all over \ac{IP} networks.}
-
-  \ac{SIP} with WebSockets can be used as a signaling method, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
-For instance, it's possible to interoperate web communications with \ac{SIP}, mobile and fixed phones.
-
-  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}.
-Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
-Frameworks like \textit{jsSIP}, \textit{QoffeeSIP} and \textit{sipML5} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications universe.
-
-  {\color{red}[Jitsi]}
-
-  {\color{blue} \ac{XMPP} is an open technology for real-time communication, which powers a wide range of applications including instant messaging, presence, multi-party chat, voice and video calls, collaboration, lightweight middleware, content syndication, and generalized routing of XML data.
-The technology pages provide more information about the various XMPP "building blocks".
-Several books about Jabber/XMPP technologies are available, as well.}
-
-  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH}, which basically consists on long polling.
-
+  \input{section02.tex}
 \section{\color{red}Hypermedia}
-
-  Since the early days of video technology, one of the problems that raised with it consisted of how to add more information onto it without generating multiple versions.
-Some implementations like \cite{embedded} added hypermedia information to empty space present on \ac{MPEG} frames in order to provide interactive television, the \ac{MPEG} coder and decoder were changed in order to handle hypermedia content.
-
-  The need to translate movies, raised the problem whether it is apropriate to change the original video or audio.
-For example subtitles should be an entity independent from the video, in order to be personalized or replaced easily.
-
-  Amongst multiple formats for subtitles, \ac{SAMI}, and \ac{SRT} are used by video players that support them. Although those formats have styling available, they are quite limited to text.
-
-  Hypervideo is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
-An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific point in time.
-HyperCafe \cite{hypercafe} was an experimental project to expose hypervideo concepts that consisted on an interactive film by switching between conversations inside a cafe.
-
-  Detail-on-demand is a subset o hypervideo that allow us to obtain aditional information about something that apears along the video, like obtaining information about a painting that appears in a particular segment.
-Hyper-Hitchcock\cite{hitchcock} is an editor and player of detail-on-demand video.
-
-  In order to navigate through a dynamic video, one must be aware of time synchronization and the multiple time flows, it's important that all time, causality and behavior rules are well defined.
-
-  HyVAL\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of video, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
-This approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
-
-  \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
-With \ac{SMIL} it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
-
-  In order to create a multimedia rich hypercall, \ac{SMIL} fits our goals, but it lacks on browser compatibility.
-Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \ac{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plugin.
-
-  SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
-But SmillingWeb isn't fully implemented yet and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes in real time.
-
-  \ac{SVG} is a format that incorporates the animation feature of \ac{SMIL}.
-Currently \ac{SVG} allow us to add movement and animate attributes of elements.
-When embedded on \ac{HTML}, it allows dynamic changes to inner content in real time, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
-
-  Video functionalities are already embedded in \ac{HTML}5, like \ac{SVG} it is also possible to bind javascript functions for different kinds of events over videos.
-
-  By using techonologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, Javascript and \ac{SVG}, it's possible to raise communications to a new level.
-For example, with \ac{API}s like WebGL, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
-Another example would be a collaborative spreadsheet using WebRTC.
-With this, hypercalls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
-
-  In this project our goal is to enrich hypercalls with no limits, every user should be free to choose how it wants to be contacted and it wants to share its contents.
-
-  In order to give users a personalized communication channel, each user must have a personal web page where its available plugins could be downloaded from other peers, after that they can talk in the same language whatever it is.
-
-  {\color{blue} Collaboration is the basis for joint value creation.
-With increasing complexity of tasks to conduct in business, science, and administration, efficient team collaboration has become a crucial success factor.}
-
-  {\color{blue}Software tools for collaboration typically support the exchange of data and centralized storage of information, for instance in shared workspaces.
-Only in recent years, technology became available allowing real-time collaboration via the web-browser.
-This enables team partners to edit a document concurrently and let others see changes instantly.}
-
-  {\color{red}[Local Storage]}
-
-  {\color{red}[Video Recording]}
-
-  {\color{red}[Video Playback - Past and Present]}
+  \input{section03.tex}
+\section{\color{red}Applications}
+  \input{section04.tex}
 

### 2015-04-07 tese/section01.tex
[ADD] added more references, a simple paragraph about skype

@@
 
-Compared to Skype, \ac{WebRTC} allows web browser real time communications without installing any aditional application or plugin.
+{\color{blue}Skype is a stand-alone application which is compatible with all major operating systems (Windows, Linux, MacOS) and mobile operating systems (Android, iOS, Windows).
+This enables voice, IM and video communication.
+It's main advantage is that it has a huge installation base (approx 299 million users).}
+
+Compared to Skype, \ac{WebRTC} is an open source technology that allows web browser real time communications without installing any aditional application or plugin.

### 2015-04-07 tese/section02.tex
[ADD] added more references, a simple paragraph about skype

@@
 
-  WebSockets provide bidirectional communications trough a full-duplex socket channel.
+  The WebSocket protocol (rfc6455\footnote{\url{http://tools.ietf.org/html/rfc6455}}) provides bidirectional communications trough a full-duplex socket channel.
 WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
@@
 
-  \ac{SIP} is protocol used for negotiation, creation, modification and finalization of communication sessions between users.
+  \ac{SIP} (rfc3261\footnote{\url{http://tools.ietf.org/html/rfc3261}}) is protocol used for negotiation, creation, modification and finalization of communication sessions between users.
 \ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as signaling protocol.

### 2015-04-11 tese/section01.tex
[ADD] review of NAT section

@@
 
-Initially \ac{NAT} offered an alternative for address exhaustion and a false sensation of better security, asymmetric \ac{NAT} became a vulgar configuration on the web.
+Initially \ac{NAT} offered an alternative for address exhaustion and a minimal sensation of security.
+There are four types of \ac{NAT}, "Full Cone NAT", "Restricted Cone NAT", "Port Restricted Cone NAT", "Symmetric NAT".
+
+Full Cone \ac{NAT} maps public each \ac{IP} address and port to a private \ac{IP} address and port, any external host can communicate with private hosts throught their mapped public address and port.
+This represents the least restrictive type of \ac{NAT} and as we will later, the unique type of \ac{NAT} that enables real time communications from point to point.
+
+Restricted Cone \ac{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host.
+With this type of \ac{NAT}, the private client can be contacted from any port of the same external host.
+
+Port Restricted Cone \ac{NAT} works in the same way as Restricted Cone \ac{NAT} but only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.
+
+Symmetric NAT maps different ports for each connection, as we will see later, this represents a problem on real time communications.
+
+Asymmetric \ac{NAT} became a vulgar configuration on the web.
 As a direct result, problems started to appear, the amount of ports that \ac{IP} disponibilizes is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follows this model to be implemented ineffectively.
@@
 
+\ac{STUN} servers are quite simple, they receive requests from \ac{NAT}ed clients, the source address of a request is the public address that \ac{NAT} mapped to the client, \ac{STUN} servers will reply the mapped public address to the client, so it knows its public \ac{IP} address and port.
+Because Symmetric \ac{NAT} changes \ac{IP} port for each different connection, \ac{STUN} servers will reply the \ac{IP} address and port of their connection, which will be useless to clients connections, that's why Symmetric \ac{NAT} represents a problem for real time communications.
+
+On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a P2P network relay to find the best peer but, after that, the behavior is much like client-server.
+Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}.
+\ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.
+
 Most of client-server applications aren't affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints.
@@
 The requirements of video communication makes this kind of model unsuitable.
-
-On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a P2P network relay to find the best peer but, after that, the behavior is much like client-server.
-Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}.
-{\color{red} [reference to \ac{NAT} types]}.
-\ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.
 

### 2015-04-12 tese/section01.tex
[ADD] added more details on SMIL

@@
 
-On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a P2P network relay to find the best peer but, after that, the behavior is much like client-server.
+On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a \ac{P2P} network relay to find the best peer but, after that, the behavior is much like client-server.
 Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}.
@@
 
-When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} cames to simplify how audio and video are transmited.
+When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} cames to simplify how audio and video are transmited through web browsers.
 
-{\color{blue}Skype is a stand-alone application which is compatible with all major operating systems (Windows, Linux, MacOS) and mobile operating systems (Android, iOS, Windows).
-This enables voice, IM and video communication.
-It's main advantage is that it has a huge installation base (approx 299 million users).}
-
-Compared to Skype, \ac{WebRTC} is an open source technology that allows web browser real time communications without installing any aditional application or plugin.
+Skype is an application that allows video, voice and instant messaging communication, its main strength is the amount of users that are using it nowadays.
+But compared to Skype, \ac{WebRTC} is an open source technology that allows web browser real time communications without installing any aditional application or plugin.

### 2015-04-12 tese/section02.tex
[ADD] added more details on SMIL

@@
 
-  \ac{WebRTC} doesn't implement signaling, different applications may need different protocols, there is no single answer that fits all problems.
+  \ac{WebRTC} doesn't implement signaling, different applications may require different protocols, there is no single answer that fits all problems.
 Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, WebSockets, Socket.io or by implementing a custom protocol.
 
-  \ac{WebRTC} uses \ac{SDP} to define peer connection properties, types of supported media, codecs, protocols used and network information.
+  \ac{WebRTC} uses \ac{SDP} (rfc4566\footnote{\url{http://tools.ietf.org/html/rfc4566}}) to define peer connection properties such as types of supported media, codecs, protocols used and network information.
+An \ac{SDP} offer describes to other peers the expected type of communication and its details, such as used transport protocols, codecs, security and other.
 

### 2015-04-12 tese/section03.tex
[ADD] added more details on SMIL

@@
 
-  \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia, for instance, it could be used to overlay subtitles on films.
+  \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films.
 With \ac{SMIL} it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
+
+  \ac{SMIL} is \ac{XML} based and defines twelve modules: \textit{Animation}, \textit{Content Control}, \textit{Layout}, \textit{Linking}, \textit{Media Objects}, \textit{SmilText}, \textit{Metainformation}, \textit{Structure}, \textit{Timing}, \textit{Time Manipulations}, \textit{State} and \textit{Transitions}.
+
+\begin{itemize}
+
+  \item \textbf{Animation} module contains elements and attributes that define a time based mechanism for composing the effects of animations.
+For example, this module can changes on \ac{XML} or \ac{CSS} attributes like colour and dimensions.
+
+  \item \textbf{Content Control} module contains elements and attributes that provide optimized alternatives for content delivery.
+For example, it could be used to change audio language in function of user's nationality, for videos with multiple audio channels.
+
+  \item \textbf{Layout} module contains elements and attributes for colouring and positioning media content, another layout mechanisms are also possible, such as \ac{CSS}.
+
+  \item \textbf{Linking} module contains elements and attributes for navigatonal hyperlinking. Navigation can be triggered by events or user interaction.
+
+  \item \textbf{Media Object} module contains elements and attributes for referencing rendering behaviour of external multimedia or control objects.
+
+  \item \textbf{SmilText} module contains elements and attributes that defines and controls timed text. For example this module could be used to create labels and captions.
+
+  \item \textbf{Metainformation} module contains elements and attributes that allows describing the \ac{SMIL} document.
+For example, this module could be used to define a movie details such as category, director, writters and cast.
+
+  \item \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content.
+This module defines a \textit{head} element that contains non temporal behaviour information defined by  \textit{Metainformation}, \textit{Layout} and \textit{Content Control} modules.
+This module also defines defines the \textit{body} element, where all temporal related module information is contained.
+
+  \item \textbf{Timing} module is the most important module on \ac{SMIL} specification, because of its complexity, it is divided into seventeen submodules for coordination and synchroniztion of media over time.
+The three main elements are \textit{seq}, \textit{excl} and \textit{par}, respectively, they play child elements in sequence, one at a time and all at the same time.
+
+  \item \textbf{Time Manipulations} module adds time behaviour attributes to \ac{SMIL} elements, such as speed, rate or time.
+
+  \item \textbf{State} module defines attributes that defines the state of \ac{SMIL} elements, such as element visibility, current element time, amount of repeated loops, playing state and many others.
+
+  \item \textbf{Transitions} module defines attributes and elements that defines transitions across multiple \ac{SMIL} elements according to \textit{Timing} module.
+
+\end{itemize}
+
+{\color{blue} The Document Object Model (DOM) is a cross-platform and language-independent convention for representing and interacting with objects in HTML, XHTML, and XML documents.[1] The nodes of every document are organized in a tree structure, called the DOM tree.
+Objects in the DOM tree may be addressed and manipulated by using methods on the objects.
+The public interface of a DOM is specified in its application programming interface (API).}
+
+{\color{blue} Broadly defined, the SMIL DOM is an Application Programming Interface (API) for SMIL documents and XML/HTML documents that integrate SMIL functionality.
+It defines the logical structure of documents and the way a document is accessed and manipulated.
+This is described more completely in "What is the Document Object Model".}
+
+  \ac{SMIL} components are used for integrating timing into XHTML and into SVG.
 

### 2015-04-12 tese/section03.tex
added more stuff about HyVAL (not too much), tried to add more about smilling web and found that Adobe Flash must be referenced

@@
 
-  HyVAL\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of video, the structure of HyVAL is derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
-This approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
+  HyVAL\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia.
+HyVAL defines defines video structure, internal video and external media objects.
+
+  HyVAL video structure objects defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically, this approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
+External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
 
@@
 
-{\color{blue} The Document Object Model (DOM) is a cross-platform and language-independent convention for representing and interacting with objects in HTML, XHTML, and XML documents.[1] The nodes of every document are organized in a tree structure, called the DOM tree.
-Objects in the DOM tree may be addressed and manipulated by using methods on the objects.
-The public interface of a DOM is specified in its application programming interface (API).}
+The \ac{DOM} is a standard \ac{API} that allows easy management of documents that are organized in a tree structure, by providing \ac{CRUD} operations over its elements and their attributes.
+\ac{DOM} makes it easy to interoperate between imperative and declarative programming languages.
 
-{\color{blue} Broadly defined, the SMIL DOM is an Application Programming Interface (API) for SMIL documents and XML/HTML documents that integrate SMIL functionality.
-It defines the logical structure of documents and the way a document is accessed and manipulated.
-This is described more completely in "What is the Document Object Model".}
+Like \ac{DOM}, \ac{SMIL} \ac{DOM} is an \ac{API} for \ac{SMIL} documents.
+Allowing \ac{CRUD} operations over \ac{SMIL} documents is an important feature for extending \ac{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \textit{javascript} functions.
 
-  \ac{SMIL} components are used for integrating timing into XHTML and into SVG.
+  \ac{SMIL}'s modules are used to synchronize and animate \ac{XHTML} and \ac{SVG} elements.
 
@@
 
-  SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
+{\color{red} [more on smilling web would be nice]}
+
+  SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 cross platform video player with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
 But SmillingWeb isn't fully implemented yet and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes in real time.
 
-  \ac{SVG} is a format that incorporates the animation feature of \ac{SMIL}.
-Currently \ac{SVG} allow us to add movement and animate attributes of elements.
-When embedded on \ac{HTML}, it allows dynamic changes to inner content in real time, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
+{\color{red} [flash is a must reference]}
+
+  \ac{SVG} is an \ac{XML} based format that incorporates the animation module of \ac{SMIL}.
+Currently \ac{SVG} allows to add movement and animate attributes of elements.
+When embedded on \ac{HTML}, it allows dynamic changes to inner content in real time through \ac{DOM} \ac{API}, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
 

### 2015-04-23 tese/section03.tex
added more pdfs, references to silverlight and flash

@@
 
-{\color{red} [flash is a must reference]}
-
   \ac{SVG} is an \ac{XML} based format that incorporates the animation module of \ac{SMIL}.
@@
 
+{\color{red} [svg is more than this]}
+
   Video functionalities are already embedded in \ac{HTML}5, like \ac{SVG} it is also possible to bind javascript functions for different kinds of events over videos.
+
+  Back in 1995, \textit{flash}\footnote{\url{flash site}} was developed for web-based animations, introducing video support in 2002, flash started to grow after that.
+Concurrency players, at that time, were focused on playing video and audio, flash had vector graphics and they were focused on streaming \textit{on-demand} video across multiple platforms.
+VP6 was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on internet connection latency.
+
+  Adobe Flash was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
+
+  \ac{RTMP} is a \ac{TCP} based protocol used to streaming audio, video and data between \ac{FMS} and flash player.
+A bidirectional connection is established between the two to in order allow real time coomunications.
+A flash player can stream a webcam video to \ac{FMS} according to \ac{RTMP} and another flash player can request a video stream to \ac{FMS} that can be pre-recorded stream, live stream or data.
+Multiple \ac{FMS} servers can be chained together in order to increase capacity and handle more streams simultaneosly.
+
+  \ac{FMS} can stream video and audio to one or more subscribers by sending a separate copy for each subscriber.
+With \ac{RTMFP} is possbile to stream video between flash players, allowing a publisher breaking a stream into pieces that can be cooperatively distributed in a P2P mesh, \ac{RTMFP} uses \ac{UDP} to speed packet delivery, although is not reliable, it is well suited for video streaming, like \ac{WebRTC} flash players also need to apply techninques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
+
+  Although \ac{HTML}5, javascript, \ac{CSS} and \ac{WebRTC} are implementing some functionalities of flash, it doesn't mean that flash will be replaced, instead of that both technologies can used to develop rich internet applications.
+It is also important to note that \ac{HTML}5 is more compatible with mobile devices than adobe flash.
+
+  Like Flash, Microsoft Silverlight is a cross browser plugin and platform that is used to develop rich internet applications.
+It supports vector graphics, animation and video.
+Compared to flash, which uses ActionScript, Silverlight applications can use languages like C\#, VisualBasic and \ac{XAML}.
+Silverlight uses a technique called Smooth Streaming from IIS Media Service that consists on delivering video in real time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.
 

### 2015-04-23 tese/template_project_latex.tex
[ADD] added more details on SMIL
added more stuff about HyVAL (not too much), tried to add more about smilling web and found that Adobe Flash must be referenced
added more pdfs, references to silverlight and flash

@@
 \newacro{CSS}{Cascading Style Sheets}
+\newacro{CPU}{Central Processing Unit}
+\newacro{CRUD}{Create, Read, Update and Delete}
+\newacro{DOM}{Document Object Model}
+\newacro{FMS}{Flash Media Server}
 \newacro{HTML}{HyperText Markup Language}
@@
 \newacro{NAT}{Network Address Translation}
+\newacro{P2P}{Peer-to-peer}
 \newacro{PSTN}{Public Switched Telephone Network}
+\newacro{RTMP}{Real Time Messaging Protocol}
+\newacro{RTMFP}{Real-Time Media Flow Protocol}
 \newacro{RFC}{Request For Comments}
@@
 \newacro{TURN}{Traversal Using Relays around NAT}
+\newacro{UDP}{User Datagram Protocol}
 \newacro{VoIP}{Voice Over IP}
 \newacro{WebRTC}{Web Real-Time Communication}
+\newacro{XAML}{eXtensible Application Markup Language}
 \newacro{XEP}{XMPP Extensions}
@@
 \newacro{XMPP}{Extensible Messaging and Presence Protocol}
+\newacro{XHTML}{eXtensible Hypertext Markup Language}
 \overfullrule=2cm
@@
 
-\title{Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC}
+\title{Comunicações Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC}
 \author{%

### 2015-04-25 tese/section03.tex
more one half page

@@
 
+  \ac{HTML} is a markup language based on \ac{XML} that is used for creating web pages.
+\ac{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages.
+Languages like \ac{CSS} and Javascript are tipically combined to \ac{HTML} for improving the interaction and appearance of a web page.
+
+  \ac{CSS} idea is to separate the structure of an \ac{XML} document from its appearance.
+\ac{CSS} defines styles for \ac{XML} tags based on their name, class or identifier, besides static styling it supports animations and transitions leading to a more dynamic content.
+
+  \textit{Javascript} is an imperative object oriented language based on ECMAScript.
+It is used mainly on client-side and executed by a web browser.
+Javascript has it own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of a pre-installed plugin.
+
+  \textit{Javascript} has compatibility issues amongst different web browsers, leading to different behaviours.
+To solve that problem, there are libraries written in javascript, namely \textit{jQuery}, that implements the same funcionality for multiple browsers, solving most of incompatibility issues.
+
 {\color{red} [more on smilling web would be nice]}
 
-  SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 cross platform video player with javascript and jQuery, which doesn't need to be installed and shouldn't have incompatibility issues.
+  SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 cross platform video player with javascript and \textit{jQuery}, which, unlike \cite{ambulant}, doesn't need to be installed and shouldn't have incompatibility issues.
 But SmillingWeb isn't fully implemented yet and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes in real time.
+
+  With the emergence of \ac{HTML}5, tags like \textit{video}, \textit{audio} and \textit{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format.
+Another important tag is \textit{canvas} that allows to draw graphics with javascript on a rectangle within a web page.
 
@@
 Currently \ac{SVG} allows to add movement and animate attributes of elements.
-When embedded on \ac{HTML}, it allows dynamic changes to inner content in real time through \ac{DOM} \ac{API}, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
+When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real time through \ac{DOM} \ac{API}, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
 
@@
 
-  Although \ac{HTML}5, javascript, \ac{CSS} and \ac{WebRTC} are implementing some functionalities of flash, it doesn't mean that flash will be replaced, instead of that both technologies can used to develop rich internet applications.
+  Although \ac{HTML}5, javascript, \ac{CSS} and \ac{WebRTC} are implementing some functionalities of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich internet applications.
 It is also important to note that \ac{HTML}5 is more compatible with mobile devices than adobe flash.

### 2015-04-25 tese/section03.tex
organized some bibliography

@@
   \ac{CSS} idea is to separate the structure of an \ac{XML} document from its appearance.
-\ac{CSS} defines styles for \ac{XML} tags based on their name, class or identifier, besides static styling it supports animations and transitions leading to a more dynamic content.
+\ac{CSS} defines styles for \ac{XML} tags based on their name, class, identifier or position, besides static styling it also supports animations and transitions leading to a more dynamic content.
 
@@
 It is used mainly on client-side and executed by a web browser.
-Javascript has it own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of a pre-installed plugin.
+Javascript has it own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plugins.
 
@@
 
-{\color{red} [more on smilling web would be nice]}
-
   SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 cross platform video player with javascript and \textit{jQuery}, which, unlike \cite{ambulant}, doesn't need to be installed and shouldn't have incompatibility issues.
 But SmillingWeb isn't fully implemented yet and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes in real time.
+
+{\color{red} [more on smilling web would be nice]}
 

### 2015-04-26 curriculo/cv.tex
[UPD] updated CV
added more on webrtc and smilling web

@@
     {\color{red} $\varheartsuit$} Java
-	 (Android, PlayFramework, LWJGL, MapReduce), C, C++ (OpenGL), C\# (XNA, Windows Forms, WPF, lpsolve), VB, Python, Scala, Scheme, HTML, CSS, JavaScript (jQuery, AJAX, WebGL), SQL, PHP, GLSL, HLSL.  
+	 (Android, PlayFramework, LWJGL, MapReduce), C, C++ (OpenGL), C\# (XNA, Windows Forms, WPF, lpsolve), VB, Python, Scala, Scheme, HTML, CSS (Bootstrap), JavaScript (jQuery, AJAX, WebGL, WebRTC), PHP (Wordpress), GLSL, HLSL.
    \section{databases}
-	MySQL, PostgreSQL, MongoDB, DynamoDB.
+	DynamoDB, MongoDB, MySQL, Neo4j, PostgreSQL, SQLite.
    \section{soft skills}
@@
   \entry
+    {2015}
+    {Bomb Raider}
+    {\href{http://play.google.com/store/apps/details?id=pt.ist.bombraider}{Google Play}}
+    {Bomb Raider is a simple singleplayer and multiplayer game where the goal is destroying all enemies in the map.}
+
+  \entry
     {2013}

### 2015-04-26 tese/section01.tex
added more on webrtc and smilling web

@@
 
-Skype is an application that allows video, voice and instant messaging communication, its main strength is the amount of users that are using it nowadays.
-But compared to Skype, \ac{WebRTC} is an open source technology that allows web browser real time communications without installing any aditional application or plugin.
+\ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser real time communications without installing any aditional application or plugin.
+
+\ac{WebRTC} uses \ac{UDP} for transporting data, which provides lower latencies than \ac{TCP}, but is not reliable and packet order and integrity are not assured.
+\ac{SCTP} and \ac{SRTP} are used for streaming data, providing a mechanism for congestion control and partial reliable delivery over \ac{UDP}.
+All transferred audio, data,  and video must be encrypted with \ac{DTLS} which provides the same security guarantess as \ac{TLS}.
+
+Skype is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.
+But compared to Skype, \ac{WebRTC} applications don't need to be pre-installed.
+
+Google Hangouts is video conference web application, in the past in order to use hangouts on a web browser a plugin was needed to be installed, nowadays hangouts is using \ac{WebRTC}.

### 2015-04-26 tese/section03.tex
added more on webrtc and smilling web

@@
 
-  SmillingWeb \cite{smillingweb} attempts to implement \ac{SMIL} 3.0 cross platform video player with javascript and \textit{jQuery}, which, unlike \cite{ambulant}, doesn't need to be installed and shouldn't have incompatibility issues.
-But SmillingWeb isn't fully implemented yet and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes in real time.
+  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \textit{JavaScript} and \textit{jQuery} which, unlike \cite{ambulant}, doesn't need to be installed and shouldn't have incompatibility issues.
 
-{\color{red} [more on smilling web would be nice]}
+  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3, they take into account unsuportted web browsers though Modernizr\footnote{modernizr url}, a simple JavaScript library that may require plugins if new features aren't supported.
+
+  But SmillingWeb just implements a subset of \ac{SMIL} 3.0 and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes due to real time communications.
+
+  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to low latency netwoks leading to experience pauses during playback.
 

### 2015-04-26 tese/section01.tex
more on webrtc

@@
 
+\ac{WebRTC} defines three main \ac{API}s: GetUserMedia, PeerConnection and DataChannel.
+
+\begin{itemize}
+  \item \textbf{GetUserMedia} allows from the browser to access to camera, microphone and device screen.
+
+  \item \textbf{PeerConnection} aquires connection data and negotiates with peers.
+
+  \item \textbf{DataChannel} allows to send whatever type of data to other peers.
+\end{itemize}
+
 \ac{WebRTC} uses \ac{UDP} for transporting data, which provides lower latencies than \ac{TCP}, but is not reliable and packet order and integrity are not assured.

### 2015-04-26 tese/section03.tex
more on webrtc

@@
 
-{\color{red} [svg is more than this]}
-
-  Video functionalities are already embedded in \ac{HTML}5, like \ac{SVG} it is also possible to bind javascript functions for different kinds of events over videos.
+  Video and audio functionalities are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind javascript functions for different kinds of events over video and audio elements.
 

### 2015-04-27 tese/section04.tex
more one half page
more on webrtc
added more bibliography

@@
 
-  {\color{red}[Video Recording]}
+  \ac{RTP}\footnote{rfc3550} is used for streaming audio and video over \ac{IP}, the multimedia content is transported on the payload of \ac{RTP} messages, \ac{RTP} contains headers for payload indentification.
+\ac{RTP} is independent from its payload type, allowing to transport any kind of encoded multimedia.
+A sequence number is used for sorting received packets.
+
+  \ac{RTP} allows to change its requirements and add extensions to it with profiles, one of the most used is the \ac{RTP} profile for audio and video \footnote{rfc3551} which lists the payload encodings and compression algorithms.
+This profile also assigns a name to each encoding which may be used other protocols like \ac{SDP}.
+
+  \ac{RTP} recorders are independent of payload encoding, they don't decode \ac{RTP} packets, they record packets instead, allowing to record all video and audio formats.
 
   {\color{red}[Video Playback - Past and Present]}
+
+  {\color{red}[Identity]}
+
+\quote{There is strong growth in the deployment of devices that integrate regular Web technologies such as HTML, CSS, and SVG, coupled with various device APIs.}

### 2015-04-29 tese/section04.tex
interactive RTP

@@
 
-  {\color{red}[Video Playback - Past and Present]}
+\begin{figure}[H]
+	\begin{center}
+		\centering
+		\includegraphics[width=0.35\textwidth]{figures/media_types.png}
+
+	\caption{Media Types}
+	\end{center}
+\end{figure}
+
+	Media Types can be described as discrete or continuos and interactive or non-interactive.
+For example an image is non-interctive and discrete, for instance, a video is continuos and non-interactive.
+A simple chat with just text is interactive and discrete.
+An animation that changes in function of user behaviour is interactive and continuous.
+
+	Streaming protocols like \ac {RTP} were designed for continuous and non-interactive media types, such as audio and video.
+Discrete and non-interactive media don't need to streamed through \ac{RTP} because they don't change with.
+For example if an image appears in a specific time interval, just the \ac{HTML} or javascript that will reference the image must be streamed, the image itself is then transfered through \ac{HTTP}.
+
+	In order to play any kind of stream, a player for interactive stream is needed downloading an environment, decoding the \ac{RTP} payload to determine the state and display it to the user.
+Streaming interactive media like the combination of \ac{HTML}, \ac{CSS} and JavaScript requires more than interpreting the code, a streamed user interface may contain an internal state that is not shown on code.
+
+	Everytime an event is proccessed on one of the endpoints, both sender and receivers state must stay synchronized, otherwise events may behave differently.
+
+	To achieve synchronization of interactive data most packets have three types: State, Delta-State and Event.
+State packet defines environment's complete state.
+Delta-State packets tranports just the piece of state that changed.
+Event packets informs that an event ocurred over the interactive media.
+
+	An \ac{RTP} recorder can have two operation modes, recording or playback.
+Traditional \ac{RTP} players can do random access, in contrast, interactive \ac{RTP} players must restore the environment and context at a given time.
+The environment is the initial state, so we can call it a non-interactive descrete media and handle it over \ac{HTTP}.
+After the receiver has received the environment, it should calculate the state at the given time.
+
+	If the \ac{RTP} recorder controls the correct data to send to the receivers, it cannot be a simple \ac{RTP} recorder as it must compute the state or delta-state to send.
+Therefore, if the receiver receives all recorded packets, it can calculate the current state from a nearest complete state.
+Having too much complete states, results on more precise random accesses but there is the storage space payback.
+On the other hand if there are few complete states recorded followed by delta-states, the recorded stream will occupy less storage space, but random accesses will be less granular.
+
+	By recording and streaming the interactive media's complete state periodically, it is possible to restore the media state even if messages are lost.
+
+	With such an interactive \ac{RTP} recorder it is then possible to record, play, fast forward, fast rewind, stop and jump to random positions.
+
+  \cite{interactive_stream} proposed an \ac{RTP} profile for real-time transmission of interactive media.
+This new profile reuses much of video and audio profile implementation, integrating the interactive component.
+
+  {\color{red}[Model View Controller (Suited for interactive RTP)]}
 

### 2015-04-30 tese/section01.tex
added more bibliography
jitsi stuff

@@
 \ac{SCTP} and \ac{SRTP} are used for streaming data, providing a mechanism for congestion control and partial reliable delivery over \ac{UDP}.
-All transferred audio, data,  and video must be encrypted with \ac{DTLS} which provides the same security guarantess as \ac{TLS}.
+All transferred audio, data and video must be encrypted with \ac{DTLS} which provides the same security guarantess as \ac{TLS}.
 
@@
 Google Hangouts is video conference web application, in the past in order to use hangouts on a web browser a plugin was needed to be installed, nowadays hangouts is using \ac{WebRTC}.
+
+Jitsi Meet\footnote{jitsi meet} is a \ac{WebRTC} collaborative application that uses Jitsi Videobridge for high quality and scalable video conferences and supports shared document editing.
+Jitsi Videobridge is a server that enables multiparty video calls.

### 2015-04-30 tese/section04.tex
jitsi stuff

@@
 This profile also assigns a name to each encoding which may be used other protocols like \ac{SDP}.
+
+  {\color{red}[DTLS and SRTP]}
 
@@
 
+	Multiparty video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants.
+Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
+
+	Jitsi Video Bridge \footnote{jitsi VB} receives one stream from every participant on a conference, either from a jitsi client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
+Although all the participants need to download all the streams from Jitsi Video Bridge Server, typically download rates are much bigger than upload rates, making this solution more feasible.
+
+	Jitsi Video Bridge uses \ac{XMPP} as a signaling protocol and its colibri extension (XEP-0340) to reserve channels for video transmission.
+Although this choice for signalling protocol, Jitsi Video Bridge also supports \ac{SIP} and \ac{Jingle}.
+
   {\color{red}[Model View Controller (Suited for interactive RTP)]}

### 2015-05-02 tese/section04.tex
added more on interactive recording

@@
 
-	Media Types can be described as discrete or continuos and interactive or non-interactive.
-For example an image is non-interctive and discrete, for instance, a video is continuos and non-interactive.
-A simple chat with just text is interactive and discrete.
+	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuos, the second one describes it as interactive or non-interactive.
+A discrete media is caracterized by its type not depending from time, as continuous media depending from it.
+Interactive media is caracterized by its state being changed by external events such as user interactions.
+
+	For example, an image is non-interctive and discrete, for instance, a video is continuos and non-interactive.
+A simple colaborative editor with just text is interactive and discrete.
 An animation that changes in function of user behaviour is interactive and continuous.
@@
 
-	In order to play any kind of stream, a player for interactive stream is needed downloading an environment, decoding the \ac{RTP} payload to determine the state and display it to the user.
+	In order to play any kind of stream, a player for interactive stream is needed for downloading an environment, decoding the \ac{RTP} payload to determine the state and display it to the user.
 Streaming interactive media like the combination of \ac{HTML}, \ac{CSS} and JavaScript requires more than interpreting the code, a streamed user interface may contain an internal state that is not shown on code.
@@
 
-	To achieve synchronization of interactive data most packets have three types: State, Delta-State and Event.
+	To achieve synchronization of interactive data most packets have three types: \textit{State}, \textit{Delta-State} and \textit{Event}.
 State packet defines environment's complete state.
-Delta-State packets tranports just the piece of state that changed.
+Delta-State packets transports just the piece of state that changed.
 Event packets informs that an event ocurred over the interactive media.
@@
 Therefore, if the receiver receives all recorded packets, it can calculate the current state from a nearest complete state.
-Having too much complete states, results on more precise random accesses but there is the storage space payback.
-On the other hand if there are few complete states recorded followed by delta-states, the recorded stream will occupy less storage space, but random accesses will be less granular.
+Streaming too much complete states, results on more precise random accesses but the trade-off is the higher bandwidth usage and used storage space on the recording server.
+On the other hand, if there are fewer complete states recorded followed by delta-states, the recorded stream will occupy less storage space, but random accesses will be less granular.
 
 	By recording and streaming the interactive media's complete state periodically, it is possible to restore the media state even if messages are lost.
+
+	In order to synchronize an interactive application state amongst participants, the needed objects to synchronize must be serializable and sent to other participants.
+
+  {\color{red}[Model View Controller (Suited for interactive RTP)]}
 
@@
 This new profile reuses much of video and audio profile implementation, integrating the interactive component.
+\cite{interactive_record} explained how to record interactive video with this new profile.
 
@@
 
-  {\color{red}[Model View Controller (Suited for interactive RTP)]}
-
   {\color{red}[Identity]}

### 2015-05-03 tese/section01.tex
14 pages, DTLS, TLS, SRTP, SCTP

@@
 
+\ac{TLS} doesn't support independent record decryption, for that it requires a reliable transport channel, tipically \ac{TCP}.
+The decryption of a record depends on the previous record, which for unreliable transport protocols like \ac{UDP} may represent a problem, either due to packet loss or different receival order.
+
+\ac{DTLS} is similar to \ac{TLS} but on top of \ac{UDP}, the main difference is the inclusion of a sequence number per packet that is used for packet re-ordering on receival and protects from duplicated packets.
+If a packet sequence number is less then the expected sequence number, the packet is discarded.
+If a packet sequence number is greater than the expected sequence number, the packet may be enqueued or discarded.
+By knowing the sequence of messages that are sent and received in \ac{TLS}, timers are used for packet retransmission avoiding ackowledgement messages.
+
+\ac{WebRTC}'s \textit{Data Channel} is built on top of \ac{SCTP}, which is secured by \ac{DTLS}.
+A \textit{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication.
+Each data channel direction can be tweaked for reliable or unreliable transmission, the same can be done for order delivery.
+A priority can also be defined for improving the quality of service for a particular stream compared to another.
+
 Skype is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.

### 2015-05-03 tese/section04.tex
14 pages, DTLS, TLS, SRTP, SCTP

@@
 
+\ac{RTCP} is used for controlling \ac{RTP} multimedia streams, it provides bandwidth statistics and control information that can be used for changing the quality of the stream in real time.
+
   \ac{RTP} allows to change its requirements and add extensions to it with profiles, one of the most used is the \ac{RTP} profile for audio and video \footnote{rfc3551} which lists the payload encodings and compression algorithms.
@@
 
-  {\color{red}[DTLS and SRTP]}
+  Another profile for \ac{RTP} is defined by \ac{SRTP}, which provides encryption, authentication and replay protection of \ac {RTP} traffic. The analogous secure protocol for \ac{RTCP} is \ac{SRTCP}.
 
-  \ac{RTP} recorders are independent of payload encoding, they don't decode \ac{RTP} packets, they record packets instead, allowing to record all video and audio formats.
+  Both \ac{SRTP} and \ac{SRTCP} uses \ac{AES} by default, which is symmetric-key algorithm for data encryption.
+Each packet is encrypted using a distinct keystream, otherwise using a single keystream with \ac{AES} on \ac{CBC} would make impossible to recover from packet loss.
+
+  wo keystream generators for \ac{AES} were defined: \textit{Segmented Integer Counter Mode} and \textit{f8-mode}.
+If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those keystream generators and it's fixed for each packet.
+
+  \ac{RTP} recorders are independent of payload encoding, they don't decode \ac{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.
 

### 2015-05-05 tese/section01.tex
english syntax for section 1, webrtc stack figure

--- 2015-04-30 tese/section01.tex
@@
-The need to build a global comunications network in an era when almost nobody had access to it, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
-\ac{IPv4} limits the number of public addresses in such a way that today is scarse \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same \ac{IP} address but different port, this technique is also known as \ac{NAT}.
+The need to build a global communications network in an age when almost nobody had access to technology and the number of future users was unpredictable, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
+\ac{IPv4} limits the number of public addresses in such a way that today is scarce \cite{ipv4}.
+One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same \ac{IP} address with a different port, this technique is also known as \ac{NAT}.
 
-Initially \ac{NAT} offered an alternative for address exhaustion and a minimal sensation of security.
-There are four types of \ac{NAT}, "Full Cone NAT", "Restricted Cone NAT", "Port Restricted Cone NAT", "Symmetric NAT".
+Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security.
+There are four types of \ac{NAT} implementations\footnote{rfc3489}, \textit{Full Cone NAT}, \textit{Restricted Cone NAT}, \textit{Port Restricted Cone NAT}, \textit{Symmetric NAT}.
 
-Full Cone \ac{NAT} maps public each \ac{IP} address and port to a private \ac{IP} address and port, any external host can communicate with private hosts throught their mapped public address and port.
+\textit{Full Cone} \ac{NAT} maps public each \ac{IP} address and port to a private \ac{IP} address and port, any external host can communicate with private hosts through their mapped public address and port.
 This represents the least restrictive type of \ac{NAT} and as we will later, the unique type of \ac{NAT} that enables real time communications from point to point.
 
-Restricted Cone \ac{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host.
+\textit{Restricted Cone} \ac{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host.
 With this type of \ac{NAT}, the private client can be contacted from any port of the same external host.
 
-Port Restricted Cone \ac{NAT} works in the same way as Restricted Cone \ac{NAT} but only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.
+\textit{Port Restricted Cone} \ac{NAT} works in the same way as Restricted Cone \ac{NAT}, but it only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.
 
-Symmetric NAT maps different ports for each connection, as we will see later, this represents a problem on real time communications.
+\textit{Symmetric} NAT maps different ports for each connection, as we will see later, this type of \ac{NAT} represents a problem on real time communications.
 
-Asymmetric \ac{NAT} became a vulgar configuration on the web.
-As a direct result, problems started to appear, the amount of ports that \ac{IP} disponibilizes is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follows this model to be implemented ineffectively.
+\textit{Asymmetric} \ac{NAT} became a vulgar configuration on the web.
+As a direct result, problems started to appear, the amount of ports that \ac{IP} makes available is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follow this model to be implemented ineffectively.
 
-Applications based on multimedia and file sharing were one of the most strained by \ac{NAT}.
+Applications based on multimedia and file sharing have been one of the most strained by \ac{NAT}.
 Those kind applications require real time communication in order to achieve the best performance.
@@
 
-\ac{STUN} servers are quite simple, they receive requests from \ac{NAT}ed clients, the source address of a request is the public address that \ac{NAT} mapped to the client, \ac{STUN} servers will reply the mapped public address to the client, so it knows its public \ac{IP} address and port.
-Because Symmetric \ac{NAT} changes \ac{IP} port for each different connection, \ac{STUN} servers will reply the \ac{IP} address and port of their connection, which will be useless to clients connections, that's why Symmetric \ac{NAT} represents a problem for real time communications.
+\ac{STUN} servers are quite simple, they receive requests from \ac{NAT}ed clients, the source address of a request is the public address that \ac{NAT} mapped to the client, \ac{STUN} servers will then reply the mapped public address to the client, so it knows its associated public \ac{IP} address and port.
+Symmetric \ac{NAT} changes \ac{IP} port for each different connection, for that reason \ac{STUN} servers will reply the \ac{IP} address and port of their connection, which will be useless to clients connections, that's why Symmetric \ac{NAT} represents a problem for real time communications.
 
-On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a \ac{P2P} network relay to find the best peer but, after that, the behavior is much like client-server.
+On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a \ac{P2P} network relay to find the best peer, but after that the behavior is much like client-server.
 Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}.
@@
 
-When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} cames to simplify how audio and video are transmited through web browsers.
+When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} came to simplify how audio and video are transmitted through web browsers.
 
-\ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser real time communications without installing any aditional application or plugin.
+\ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser real time communications without installing any additional application or plug-in.
+
+\begin{figure}[H]
+	\begin{center}
+		\centering
+		\includegraphics[width=0.9\textwidth]{figures/webrtc_stack.png}
+
+	\caption{WebRTC protocol Stack}
+	\end{center}
+\end{figure}
 
@@
 \ac{SCTP} and \ac{SRTP} are used for streaming data, providing a mechanism for congestion control and partial reliable delivery over \ac{UDP}.
-All transferred audio, data and video must be encrypted with \ac{DTLS} which provides the same security guarantess as \ac{TLS}.
+All transferred audio, data and video must be encrypted with \ac{DTLS} which provides the same security guarantees as \ac{TLS}.
 
-Skype is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.
-But compared to Skype, \ac{WebRTC} applications don't need to be pre-installed.
+\ac{TLS} doesn't support independent record decryption, for that it requires a reliable transport channel, typically \ac{TCP}.
+The decryption of a record depends on the previous record, which for unreliable transport protocols like \ac{UDP} may represent a problem, either due to packet loss or different reception order.
 
-Google Hangouts is video conference web application, in the past in order to use hangouts on a web browser a plugin was needed to be installed, nowadays hangouts is using \ac{WebRTC}.
+\ac{DTLS} is similar to \ac{TLS}, but on top of \ac{UDP}, the main difference is the inclusion of a sequence number per packet that is used for packet re-ordering on reception and protects from duplicated packets.
+If a packet sequence number is less than the expected sequence number the packet is discarded.
+If a packet sequence number is greater than the expected sequence number the packet may be enqueued or discarded.
+By knowing the sequence of messages that are sent and received in \ac{TLS}, timers are used for packet retransmission avoiding acknowledgement messages.
+
+\ac{WebRTC}'s \textit{Data Channel} is built on top of \ac{SCTP}, which is secured by \ac{DTLS}.
+A \textit{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication.
+Each data channel direction can be tweaked for reliable or unreliable transmission, the same can be done for order delivery.
+A priority can also be defined for improving the quality of service for a particular stream compared to another.
+
+\textit{Skype}\footnote{skype's site} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.
+But compared to \textit{Skype}, \ac{WebRTC} applications don't need to be pre-installed.
+
+\textit{Google Hangouts}\footnote{Hangouts site} is a video conference web application, in the past in order to use \textit{hangouts} on a web browser a plug-in was needed to be installed, nowadays hangouts is using \ac{WebRTC}.
 

### 2015-05-05 tese/section02.tex
added more on interactive recording
english syntax for section 1, webrtc stack figure
almost reviewed section 2

@@
-Signaling is the most important proccess for applications to exchange connection information about peers and servers, their capabilities and metadata.
+  Signaling is the most important process for applications to exchange connection information about peers and servers, their capabilities and meta-data.
 
   \ac{WebRTC} doesn't implement signaling, different applications may require different protocols, there is no single answer that fits all problems.
-Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, WebSockets, Socket.io or by implementing a custom protocol.
+Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, \textit{WebSockets}, \textit{Socket.io} or by implementing a custom protocol.
 
@@
 
-  Polling consists on sending periodic messages that the server responds imediately with empty content or fresh information.
-Because real time communications are unpredictable, if the time between periodic requests is short, most of time the server will return empty results wasting network bandwith and energy.
+  Polling consists on sending periodic messages that the server responds immediately with empty content or fresh information.
+Real time communications are unpredictable, if the time between periodic requests is short, most of the time the server will return empty results wasting network bandwidth and energy.
 On the other hand, if the time between periodic requests is large, newer messages may arrive later.
 
-  A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the receival, the client makes another request.
-This results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which adds data overhead, specialy for small messages.
+  A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the message receipt, the client makes another request.
+Long polling technique results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which adds data overhead, especially for short messages.
 
-  The WebSocket protocol (rfc6455\footnote{\url{http://tools.ietf.org/html/rfc6455}}) provides bidirectional communications trough a full-duplex socket channel.
+  The WebSocket protocol (rfc6455\footnote{\url{http://tools.ietf.org/html/rfc6455}}) provides bidirectional communications over a full-duplex socket channel.
 WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
-WebSockets may not be avalailable on every web browser, frameworks like \textit{socket.io}\footnote{\url{http://socket.io/}} and \textit{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.
+WebSockets may not be available on every web browser, frameworks like \textit{socket.io}\footnote{\url{http://socket.io/}} and \textit{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.
 
-  \ac{BOSH} is a technique based on long polling, that uses two socket connections and allow sending client messages to server while a previous request is holded.
+  \ac{BOSH} is a technique based on long polling that uses two socket connections and allow sending client messages to server while a previous request is held.
+
+{\color{magenta} [BOSH, you can do it!]}
 
@@
 \ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as signaling protocol.
-The advantage of \ac{SIP} is the ability to make video and voice calls applications over \ac{IP} networks.
+The advantage of \ac{SIP} is the ability to make video and voice call's applications over \ac{IP} networks.
 
-  The work group \ac{SIMPLE} proposed the creation of \ac{SIP} extensions, namely presence information (rfc5263\footnote{\url{http://tools.ietf.org/html/rfc5263}}) and instant messaging (rfc3428\footnote{\url{http://tools.ietf.org/html/rfc3428}}).
+  The working group \ac{SIMPLE}\footnote{\url{https://datatracker.ietf.org/wg/simple/documents/}} proposed the creation of \ac{SIP} extensions, namely presence information (rfc5263\footnote{\url{http://tools.ietf.org/html/rfc5263}}) and instant messaging (rfc3428\footnote{\url{http://tools.ietf.org/html/rfc3428}}).
 
   \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}.
-Service providers are disponibilizing access to their \ac{SIP} infrastrucures through WebSockets.
+Service providers making their \ac{SIP} infrastructures available through WebSockets.
 Frameworks like \textit{jsSIP}\footnote{\url{http://jssip.net/}}, \textit{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \textit{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
 
-  \ac{SIP} with WebSockets can be used as a signaling method for WebRTC applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
-For instance, it's possible to interoperate web communications with \ac{SIP} networks, mobile and fixed phones.
+  \ac{SIP} with \textit{WebSockets} can be used as a signaling method for \ac{WebRTC} applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
+For instance, it's possible to inter-operate web communications with \ac{SIP} networks, mobile and fixed phones.
+
+\textbf{\color{red}==================================}
 
@@
 Ejabberd is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
+
+  {\color{red}[Sig-o-fly]}

### 2015-05-05 tese/section01.tex
almost reviewed section 2
quase a roçar as 16 paginas

@@
 One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same \ac{IP} address with a different port, this technique is also known as \ac{NAT}.
+
+\begin{figure}[H]
+	\begin{center}
+		\centering
+		\includegraphics[width=0.6\textwidth]{figures/nat.png}
+
+	\caption{Network Address Translation}
+	\end{center}
+\end{figure}
 
@@
 
-\ac{WebRTC} defines three main \ac{API}s: GetUserMedia, PeerConnection and DataChannel.
+{\color{magenta} [check the coverage of this stack]}
+
+\ac{WebRTC} defines three main \ac{API}s: MediaStream, PeerConnection and DataChannel.
 
 \begin{itemize}
-  \item \textbf{GetUserMedia} allows from the browser to access to camera, microphone and device screen.
+  \item \textbf{MediaStream} allows from the browser to access to camera, microphone and device screen.
 
@@
 
-\ac{WebRTC}'s \textit{Data Channel} is built on top of \ac{SCTP}, which is secured by \ac{DTLS}.
+\begin{figure}[H]
+	\begin{center}
+		\centering
+		\includegraphics[width=0.6\textwidth]{figures/basic_protocols.png}
+
+	\caption{Overview of transport protocols}
+	\end{center}
+\end{figure}
+
+\ac{WebRTC}'s \textit{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}.
+\ac{DTLS} encapsulation provides confidentiality, authentication and integrity over the transfered data.
 A \textit{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication.
-Each data channel direction can be tweaked for reliable or unreliable transmission, the same can be done for order delivery.
-A priority can also be defined for improving the quality of service for a particular stream compared to another.
+Each data channel direction can be tweaked for reliable or unreliable transmission, the same can be done for order delivery and priority which can also be defined for improving the quality of service for a particular stream compared to another.
+
+\ac{WebRTC}'s \textit{MediaStream} is built on top of \ac{SRTP}, which requires an external mechanism for key exchange.
+\ac{DTLS} keys are negotiated on handshake in order to achieve a secure connection.
+The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining communications are done through \ac{UDP} without using \ac{DTLS}.
 

### 2015-05-06 tese/section01.tex
added references, almost concluding syntax checking for section 3

@@
+\subsection{Early days of the Internet and its remaining flaws}
+
 The need to build a global communications network in an age when almost nobody had access to technology and the number of future users was unpredictable, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
@@
 \begin{figure}[H]
-	\begin{center}
 	\centering
 	\includegraphics[width=0.6\textwidth]{figures/nat.png}
-
 	\caption{Network Address Translation}
-	\end{center}
 \end{figure}
@@
 Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security.
-There are four types of \ac{NAT} implementations\footnote{rfc3489}, \textit{Full Cone NAT}, \textit{Restricted Cone NAT}, \textit{Port Restricted Cone NAT}, \textit{Symmetric NAT}.
+There are four types of \ac{NAT} implementations\cite{rfc3489}, \textit{Full Cone NAT}, \textit{Restricted Cone NAT}, \textit{Port Restricted Cone NAT}, \textit{Symmetric NAT}.
 
@@
 
+\subsection{Real time communications}
+
 \ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser real time communications without installing any additional application or plug-in.
@@
 \begin{figure}[H]
-	\begin{center}
 	\centering
 	\includegraphics[width=0.9\textwidth]{figures/webrtc_stack.png}
-
 	\caption{WebRTC protocol Stack}
-	\end{center}
 \end{figure}
-
-{\color{magenta} [check the coverage of this stack]}
 
@@
 
-  \item \textbf{PeerConnection} aquires connection data and negotiates with peers.
+  \item \textbf{PeerConnection} acquires connection data and negotiates with peers.
 
@@
 \ac{SCTP} and \ac{SRTP} are used for streaming data, providing a mechanism for congestion control and partial reliable delivery over \ac{UDP}.
-All transferred audio, data and video must be encrypted with \ac{DTLS} which provides the same security guarantees as \ac{TLS}.
+All transferred audio, data and video must be encrypted with \ac{DTLS} symmetric keys.
+\ac{DTLS} provides the same security guarantees as \ac{TLS}.
 
@@
 If a packet sequence number is greater than the expected sequence number the packet may be enqueued or discarded.
-By knowing the sequence of messages that are sent and received in \ac{TLS}, timers are used for packet retransmission avoiding acknowledgement messages.
+By knowing the sequence of messages that are sent and received in \ac{TLS}, timers are used for packet retransmission avoiding acknowledgment messages.
 
 \begin{figure}[H]
-	\begin{center}
 	\centering
 	\includegraphics[width=0.6\textwidth]{figures/basic_protocols.png}
-
 	\caption{Overview of transport protocols}
-	\end{center}
 \end{figure}
@@
 
-\textit{Skype}\footnote{skype's site} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.
+\textit{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.
 But compared to \textit{Skype}, \ac{WebRTC} applications don't need to be pre-installed.
 
-\textit{Google Hangouts}\footnote{Hangouts site} is a video conference web application, in the past in order to use \textit{hangouts} on a web browser a plug-in was needed to be installed, nowadays hangouts is using \ac{WebRTC}.
+\textit{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is a video conference web application, in the past in order to use \textit{hangouts} on a web browser a plug-in was needed to be installed, nowadays hangouts is using \ac{WebRTC}.
 
-Jitsi Meet\footnote{jitsi meet} is a \ac{WebRTC} collaborative application that uses Jitsi Videobridge for high quality and scalable video conferences and supports shared document editing.
+Jitsi Meet\footnote{\url{http://jitsi.org/Projects/JitsiMeet}} is a \ac{WebRTC} collaborative application that uses Jitsi Videobridge for high quality and scalable video conferences and supports shared document editing.
 Jitsi Videobridge is a server that enables multiparty video calls.

### 2015-05-06 tese/section02.tex
quase a roçar as 16 paginas
added references, almost concluding syntax checking for section 3

@@
+\subsection{Signaling, meet and get to know}
+
   Signaling is the most important process for applications to exchange connection information about peers and servers, their capabilities and meta-data.
@@
 
-  \ac{WebRTC} uses \ac{SDP} (rfc4566\footnote{\url{http://tools.ietf.org/html/rfc4566}}) to define peer connection properties such as types of supported media, codecs, protocols used and network information.
+  \ac{WebRTC} uses \ac{SDP} \cite{rfc4566} to define peer connection properties such as types of supported media, codecs, protocols used and network information.
 An \ac{SDP} offer describes to other peers the expected type of communication and its details, such as used transport protocols, codecs, security and other.
@@
 
-  The WebSocket protocol (rfc6455\footnote{\url{http://tools.ietf.org/html/rfc6455}}) provides bidirectional communications over a full-duplex socket channel.
-WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to websocket type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
+  The WebSocket protocol \cite{rfc6455} provides bidirectional communications over a full-duplex socket channel.
+WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to \textit{WebSocket} type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
 WebSockets may not be available on every web browser, frameworks like \textit{socket.io}\footnote{\url{http://socket.io/}} and \textit{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.
 
-  \ac{BOSH} is a technique based on long polling that uses two socket connections and allow sending client messages to server while a previous request is held.
+  \ac{BOSH}\cite{xep0124} is a technique based on long polling that uses two socket connections and allow sending client messages to server while a previous request is held.
 
-{\color{magenta} [BOSH, you can do it!]}
+  \ac{BOSH} specification assumes that a connection manager is implemented to handle \ac{HTTP} connections.
+This connection manager is basically a translator from \ac{HTTP} to raw message so the server may think this communication is performed over \ac{TCP}.
 
-  \ac{SIP} (rfc3261\footnote{\url{http://tools.ietf.org/html/rfc3261}}) is protocol used for negotiation, creation, modification and finalization of communication sessions between users.
+  When the connection manager holds for a response for too long it responds with an empty body, this technique prevents an \ac{HTTP} session from expiring when the client is waiting for a response, thus expanding the session time.
+Expiring sessions can be expensive due to the overhead of establishing new connections, it worsens even more when \ac{HTTP} is used over \ac{SSL}.
+
+  If the server is holding a request, it maintains a second connection to receive more requests from the same client.
+The request on hold returns immediately with a possible empty body leaving its socket free, while the second connection serves the polling loop.
+The exchange of roles of those two connections allow to pull data from multiple contexts instead of being locked in just one.
+
+  If the client has data to send while a request is still open, it establishes a second socket connection to the connection manager to send a new request.
+The connection manager immediately responds to the previously held request (possibly with no data) and holds open this new request.
+This results in the connections switching roles.
+
+  \ac{SIP} \cite{rfc3261} is protocol used for negotiation, creation, modification and finalization of communication sessions between users.
 \ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as signaling protocol.
@@
 
-  The working group \ac{SIMPLE}\footnote{\url{https://datatracker.ietf.org/wg/simple/documents/}} proposed the creation of \ac{SIP} extensions, namely presence information (rfc5263\footnote{\url{http://tools.ietf.org/html/rfc5263}}) and instant messaging (rfc3428\footnote{\url{http://tools.ietf.org/html/rfc3428}}).
+  The working group \ac{SIMPLE}\footnote{\url{https://datatracker.ietf.org/wg/simple/documents/}} proposed the creation of \ac{SIP} extensions, namely presence information \cite{rfc5263} and instant messaging \cite{rfc3428}.
 
@@
 
+  \ac{XMPP} was initially developed for instant messaging and presence (Jabber\footnote{\url{http://jabber.org/}}).
+It is nowadays an open technology for standardized, decentralized, secure and extensible real-time communications.
+
+  \ac{XMPP} messages are \ac{XML} based, which are attractive for applications that need structured messages and rich hypermedia.
+\ac{XMPP} advantage is the addition of extensions, for example \cite{xep0096}, which adds file transfer capabilities between two entities and \cite{xep0045} which enables multi-user chat.
+
+  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH} \cite{xep0206}, which basically consists on long polling.
+This kind of communication is also possible through WebSockets \cite{rfc7395}.
+
+  Amongst multiple XMPP server solutions are: \textit{ejabberd}\footnote{\url{http://jabberd.im/}}, \textit{Metronome}\footnote{\url{http://lightwitch.org/metronome}}, \textit{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}} and \textit{Prosody}\footnote{\url{http://prosody.im/}}.
+\textit{Ejabberd} is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
+
 \textbf{\color{red}==================================}
 
-  \ac{XMPP} was initially developed for instant messaging (Jabber\footnote{\url{http://jabber.org/}}). It is nowadays an open technology for real-time communications.
-
-  \ac{XMPP} messages are \ac{XML} based, which are attractive for applications that need structured messages and rich hypermedia.
-\ac{XMPP} advantage is the addition of extensions, for example XEP-0096\footnote{\url{http://xmpp.org/extensions/xep-0096.html}}, which adds file transfer capabilities between two entities and XEP-0045\footnote{\url{http://xmpp.org/extensions/xep-0045.html}} which enables multi-user chat.
-
-  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH} (XEP-0206\footnote{\url{http://xmpp.org/extensions/xep-0206.html}}), which basically consists on long polling.
-This kind of communication is also possible through WebSockets (rfc7395\footnote{\url{http://tools.ietf.org/html/rfc7395}}).
-
-  Amongst multiple XMPP servers softwares are: ejabberd\footnote{\url{http://jabberd.im/}}, Metronome\footnote{\url{http://lightwitch.org/metronome}}, Openfire\footnote{\url{http://igniterealtime.org/projects/openfire/}} and Prosody\footnote{\url{http://prosody.im/}}.
-Ejabberd is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
-
   {\color{red}[Sig-o-fly]}

### 2015-05-06 tese/section03.tex
added more bibliography
added references, almost concluding syntax checking for section 3

@@
+\subsection{Hypermedia, more than words, more than images}
+
   Since the early days of video technology, one of the problems that raised with it consisted of how to add more information onto it without generating multiple versions.
-Some implementations like \cite{embedded} added hypermedia information to empty space present on \ac{MPEG} frames in order to provide interactive television, the \ac{MPEG} coder and decoder were changed in order to handle hypermedia content.
+Some implementations like \cite{embedded} added hypermedia information to empty space present on \ac{MPEG} frames in order to provide interactive television, the \ac{MPEG} encoder and decoder were changed in order to handle hypermedia content.
 
-  The need to translate movies, raised the problem whether it is apropriate to change the original video or audio.
-For example subtitles should be an entity independent from the video, in order to be personalized or replaced easily.
+  The need to translate movies, raised the problem whether it is appropriate to change the original video or audio.
+For example, subtitles should be an entity independent from the video, in order to be personalized or replaced easily.
 
@@
 
-  Hypervideo is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
-An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific point in time.
-HyperCafe \cite{hypercafe} was an experimental project to expose hypervideo concepts that consisted on an interactive film by switching between conversations inside a cafe.
+  Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
+An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time.
+\textit{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted on an interactive film by switching between conversations inside a cafe.
 
-  Detail-on-demand is a subset o hypervideo that allow us to obtain aditional information about something that apears along the video, like obtaining information about a painting that appears in a particular segment.
-Hyper-Hitchcock\cite{hitchcock} is an editor and player of detail-on-demand video.
+  Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment.
+\textit{Hyper-Hitchcock}\cite{hitchcock} is an editor and player of detail-on-demand video.
 
@@
 
-  HyVAL\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia.
+  \textit{HyVAL}\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia.
 HyVAL defines defines video structure, internal video and external media objects.
 
-  HyVAL video structure objects defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically, this approach is quite limitative if we want to apply hypervideo concepts to videos that don't follow this structure.
+  \textit{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically, this approach is quite restrictive if we want to apply hyper-video concepts to videos that don't follow this structure.
 External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
@@
   \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films.
-With \ac{SMIL} it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyperlinks, amongst multiple other functionalities.
+With \ac{SMIL} it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, amongst multiple other features.
 
-  \ac{SMIL} is \ac{XML} based and defines twelve modules: \textit{Animation}, \textit{Content Control}, \textit{Layout}, \textit{Linking}, \textit{Media Objects}, \textit{SmilText}, \textit{Metainformation}, \textit{Structure}, \textit{Timing}, \textit{Time Manipulations}, \textit{State} and \textit{Transitions}.
+  \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \textit{Animation}, \textit{Content Control}, \textit{Layout}, \textit{Linking}, \textit{Media Objects}, \textit{SmilText}, \textit{Metainformation}, \textit{Structure}, \textit{Timing}, \textit{Time Manipulations}, \textit{State} and \textit{Transitions}.
 
@@
 
-  \item \textbf{Linking} module contains elements and attributes for navigatonal hyperlinking. Navigation can be triggered by events or user interaction.
+  \item \textbf{Linking} module contains elements and attributes for navigational hyperlinking. Navigation can be triggered by events or user interaction.
 
@@
 
-  \item \textbf{SmilText} module contains elements and attributes that defines and controls timed text. For example this module could be used to create labels and captions.
+  \item \textbf{SmilText} module contains elements and attributes that defines and controls timed text. For example, this module could be used to create labels and captions.
 
   \item \textbf{Metainformation} module contains elements and attributes that allows describing the \ac{SMIL} document.
-For example, this module could be used to define a movie details such as category, director, writters and cast.
+For example, this module could be used to define a movie details such as category, director, writers and cast.
 
   \item \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content.
-This module defines a \textit{head} element that contains non temporal behaviour information defined by  \textit{Metainformation}, \textit{Layout} and \textit{Content Control} modules.
+This module defines a \textit{head} element that contains non temporal behavior information defined by  \textit{Metainformation}, \textit{Layout} and \textit{Content Control} modules.
 This module also defines defines the \textit{body} element, where all temporal related module information is contained.
 
-  \item \textbf{Timing} module is the most important module on \ac{SMIL} specification, because of its complexity, it is divided into seventeen submodules for coordination and synchroniztion of media over time.
+  \item \textbf{Timing} module is the most important module on \ac{SMIL} specification, due to its complexity, it is divided into seventeen sub-modules for coordination and synchronization of media over time.
 The three main elements are \textit{seq}, \textit{excl} and \textit{par}, respectively, they play child elements in sequence, one at a time and all at the same time.
 
-  \item \textbf{Time Manipulations} module adds time behaviour attributes to \ac{SMIL} elements, such as speed, rate or time.
+  \item \textbf{Time Manipulations} module adds time behavior attributes to \ac{SMIL} elements, such as speed, rate or time.
 
@@
 
-  \item \textbf{Transitions} module defines attributes and elements that defines transitions across multiple \ac{SMIL} elements according to \textit{Timing} module.
+  \item \textbf{Transitions} module defines attributes and elements for transitions across multiple \ac{SMIL} elements according to \textit{Timing} module.
 
@@
 The \ac{DOM} is a standard \ac{API} that allows easy management of documents that are organized in a tree structure, by providing \ac{CRUD} operations over its elements and their attributes.
-\ac{DOM} makes it easy to interoperate between imperative and declarative programming languages.
+\ac{DOM} makes it easy to inter-operate between imperative and declarative programming languages.
 
 Like \ac{DOM}, \ac{SMIL} \ac{DOM} is an \ac{API} for \ac{SMIL} documents.
-Allowing \ac{CRUD} operations over \ac{SMIL} documents is an important feature for extending \ac{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \textit{javascript} functions.
+Allowing \ac{CRUD} operations over \ac{SMIL} documents is an important feature for extending \ac{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \textit{JavaScript} functions.
 
@@
 
-  In order to create a multimedia rich hypercall, \ac{SMIL} fits our goals, but it lacks on browser compatibility.
-Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \ac{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plugin.
+  In order to create a multimedia rich hyper-call, \ac{SMIL} fits our goals, but it lacks on browser compatibility.
+Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \ac{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plug-in.
 
@@
 \ac{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages.
-Languages like \ac{CSS} and Javascript are tipically combined to \ac{HTML} for improving the interaction and appearance of a web page.
+Languages like \ac{CSS} and \textit{JavaScript} are typically combined to \ac{HTML} for improving the interaction and appearance of a web page.
 
   \ac{CSS} idea is to separate the structure of an \ac{XML} document from its appearance.
-\ac{CSS} defines styles for \ac{XML} tags based on their name, class, identifier or position, besides static styling it also supports animations and transitions leading to a more dynamic content.
+\ac{CSS} defines styles for \ac{XML} tags based on their name, class, identifier or position, besides static styling \ac{CSS} also supports animation and transitions leading to a more dynamic content.
 
-  \textit{Javascript} is an imperative object oriented language based on ECMAScript.
+  \textit{JavaScript} is an imperative object-oriented language based on \textit{ECMAScript}.
 It is used mainly on client-side and executed by a web browser.
-Javascript has it own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plugins.
+\textit{JavaScript} has its own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plug-ins.
 
-  \textit{Javascript} has compatibility issues amongst different web browsers, leading to different behaviours.
-To solve that problem, there are libraries written in javascript, namely \textit{jQuery}, that implements the same funcionality for multiple browsers, solving most of incompatibility issues.
+  \textit{JavaScript} has compatibility issues amongst different web browsers, leading to different behaviors.
+To solve that problem, there are libraries written in \textit{JavasSript}, namely \textit{jQuery} which implements the same functionality for multiple browsers, solving most of incompatibility issues.
 
@@
 
-  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3, they take into account unsuportted web browsers though Modernizr\footnote{modernizr url}, a simple JavaScript library that may require plugins if new features aren't supported.
+  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3, they take into account unsupported web browsers though \textit{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \textit{JavaScript} library that may require plug-ins if new features aren't supported.
 
@@
 
-  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to low latency netwoks leading to experience pauses during playback.
+\textbf{\color{red}==================================}
+
+  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to low latency networks leading to experience pauses during playback.
 
   With the emergence of \ac{HTML}5, tags like \textit{video}, \textit{audio} and \textit{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format.
-Another important tag is \textit{canvas} that allows to draw graphics with javascript on a rectangle within a web page.
+Another important tag is \textit{canvas} that allows to draw graphics with \textit{JavaScript} on a rectangle within a web page.
 
@@
 Currently \ac{SVG} allows to add movement and animate attributes of elements.
-When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real time through \ac{DOM} \ac{API}, besides that, it also allows to call javascript functions on events such as animation end, mouse over and mouse click.
+When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real time through \ac{DOM} \ac{API}, besides that, it also allows to call \textit{JavaScript} functions on events such as animation end, mouse over and mouse click.
 
-  Video and audio functionalities are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind javascript functions for different kinds of events over video and audio elements.
+  Video and audio functionalities are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind \textit{JavaScript} functions for different kinds of events over video and audio elements.
 
@@
 Concurrency players, at that time, were focused on playing video and audio, flash had vector graphics and they were focused on streaming \textit{on-demand} video across multiple platforms.
-VP6 was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on internet connection latency.
+VP6 was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on \textit{Internet} connection latency.
 
@@
   \ac{RTMP} is a \ac{TCP} based protocol used to streaming audio, video and data between \ac{FMS} and flash player.
-A bidirectional connection is established between the two to in order allow real time coomunications.
+A bidirectional connection is established between the two to in order allow real time communications.
 A flash player can stream a webcam video to \ac{FMS} according to \ac{RTMP} and another flash player can request a video stream to \ac{FMS} that can be pre-recorded stream, live stream or data.
-Multiple \ac{FMS} servers can be chained together in order to increase capacity and handle more streams simultaneosly.
+Multiple \ac{FMS} servers can be chained together in order to increase capacity and handle more streams simultaneously.
 
   \ac{FMS} can stream video and audio to one or more subscribers by sending a separate copy for each subscriber.
-With \ac{RTMFP} is possbile to stream video between flash players, allowing a publisher breaking a stream into pieces that can be cooperatively distributed in a P2P mesh, \ac{RTMFP} uses \ac{UDP} to speed packet delivery, although is not reliable, it is well suited for video streaming, like \ac{WebRTC} flash players also need to apply techninques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
+With \ac{RTMFP} is possbile to stream video between flash players, allowing a publisher breaking a stream into pieces that can be cooperatively distributed in a P2P mesh, \ac{RTMFP} uses \ac{UDP} to speed packet delivery, although is not reliable, it is well suited for video streaming, like \ac{WebRTC} flash players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
 
-  Although \ac{HTML}5, javascript, \ac{CSS} and \ac{WebRTC} are implementing some functionalities of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich internet applications.
+  Although \ac{HTML}5, \textit{JavaScript}, \ac{CSS} and \ac{WebRTC} are implementing some functionalities of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich \textit{Internet} applications.
 It is also important to note that \ac{HTML}5 is more compatible with mobile devices than adobe flash.
 
-  Like Flash, Microsoft Silverlight is a cross browser plugin and platform that is used to develop rich internet applications.
+  Like \textit{Flash}, \textit{Microsoft Silverlight} is a cross browser plug-in and platform that is used to develop rich \textit{Internet} applications.
 It supports vector graphics, animation and video.
-Compared to flash, which uses ActionScript, Silverlight applications can use languages like C\#, VisualBasic and \ac{XAML}.
-Silverlight uses a technique called Smooth Streaming from IIS Media Service that consists on delivering video in real time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.
+Compared to flash, which uses \textit{ActionScript}, \textit{Silverlight} applications can use languages like C\#, \textit{VisualBasic} and \ac{XAML}.
+\textit{Silverlight} uses a technique called \textit{Smooth Streaming} from \textit{IIS Media Service} that consists on delivering video in real time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.
 
-  By using techonologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, Javascript and \ac{SVG}, it's possible to raise communications to a new level.
-For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hypercall.
-Another example would be a collaborative spreadsheet using WebRTC.
-With this, hypercalls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
+  By using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \textit{JavaScript} and \ac{SVG}, it's possible to raise communications to a new level.
+For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.
+Another example would be a collaborative spreadsheet using \ac{WebRTC}.
+With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
 
-  In this project our goal is to enrich hypercalls with no limits, every user should be free to choose how it wants to be contacted and it wants to share its contents.
+  In this project our goal is to enrich hyper-calls with no limits, every user should be free to choose how it wants to be contacted and it wants to share its contents.
 
-  In order to give users a personalized communication channel, each user must have a personal web page where its available plugins could be downloaded from other peers, after that they can talk in the same language whatever it is.
+  In order to give users a personalized communication channel, each user must have a personal web page where its available plug-ins could be downloaded from other peers, after that they can talk in the same language whatever it is.

### 2015-05-06 tese/section04.tex
english syntax for section 1, webrtc stack figure
quase a roçar as 16 paginas
added references, almost concluding syntax checking for section 3

@@
+\subsection{Extending collaboration tools with time manipulation}
+
  	Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity.
-Technologies like this are appearing along this days, but they couln't be possible years ago because technology was limited or unavailable.
+Technologies like this are appearing along this days, but they couldn't be possible years ago because technology was limited or unavailable.
 Although todays technology is limited on some aspects, we are doing progress in order to improve the web ecosystem, by creating standards and migrating to newer technologies.
@@
 
-	If, for instance, one whants to rewind a real time video, recordings will be needed from who is streaming the video.
+	If, for instance, one wants to rewind a real time video, recordings will be needed from who is streaming the video.
 
-  \ac{RTP}\footnote{rfc3550} is used for streaming audio and video over \ac{IP}, the multimedia content is transported on the payload of \ac{RTP} messages, \ac{RTP} contains headers for payload indentification.
+  \ac{RTP}\cite{rfc3550} is used for streaming audio and video over \ac{IP}, the multimedia content is transported on the payload of \ac{RTP} messages, \ac{RTP} contains headers for payload identification.
 \ac{RTP} is independent from its payload type, allowing to transport any kind of encoded multimedia.
@@
 
-  \ac{RTP} allows to change its requirements and add extensions to it with profiles, one of the most used is the \ac{RTP} profile for audio and video \footnote{rfc3551} which lists the payload encodings and compression algorithms.
+  \ac{RTP} allows to change its requirements and add extensions to it with profiles, one of the most used is the \ac{RTP} profile for audio and video \cite{rfc3551} which lists the payload encodings and compression algorithms.
 This profile also assigns a name to each encoding which may be used other protocols like \ac{SDP}.
@@
   Both \ac{SRTP} and \ac{SRTCP} uses \ac{AES} by default, which is symmetric-key algorithm for data encryption.
-Each packet is encrypted using a distinct keystream, otherwise using a single keystream with \ac{AES} on \ac{CBC} would make impossible to recover from packet loss.
+Each packet is encrypted using a distinct key-stream, otherwise using a single key-stream with \ac{AES} on \ac{CBC} would make impossible to recover from packet loss.
 
-  wo keystream generators for \ac{AES} were defined: \textit{Segmented Integer Counter Mode} and \textit{f8-mode}.
-If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those keystream generators and it's fixed for each packet.
+  Two key-stream generators for \ac{AES} were defined: \textit{Segmented Integer Counter Mode} and \textit{f8-mode}.
+If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those key-stream generators and it's fixed for each packet.
 
@@
 \begin{figure}[H]
-	\begin{center}
 	\centering
-		\includegraphics[width=0.35\textwidth]{figures/media_types.png}
-
+	\includegraphics[width=0.45\textwidth]{figures/media_types.png}
 	\caption{Media Types}
-	\end{center}
 \end{figure}
 
-	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuos, the second one describes it as interactive or non-interactive.
-A discrete media is caracterized by its type not depending from time, as continuous media depending from it.
-Interactive media is caracterized by its state being changed by external events such as user interactions.
+	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuous, the second one describes it as interactive or non-interactive.
+A discrete media is characterized by its type not depending from time, as continuous media depending from it.
+Interactive media is characterized by its state being changed by external events such as user interactions.
 
-	For example, an image is non-interctive and discrete, for instance, a video is continuos and non-interactive.
-A simple colaborative editor with just text is interactive and discrete.
-An animation that changes in function of user behaviour is interactive and continuous.
+	For example, an image is non-interactive and discrete, for instance, a video is continuous and non-interactive.
+A simple collaborative editor with just text is interactive and discrete.
+An animation that changes in function of user behavior is interactive and continuous.
 
@@
 Discrete and non-interactive media don't need to streamed through \ac{RTP} because they don't change with.
-For example if an image appears in a specific time interval, just the \ac{HTML} or javascript that will reference the image must be streamed, the image itself is then transfered through \ac{HTTP}.
+For example if an image appears in a specific time interval, just the \ac{HTML} or \textit{JavaScript} that will reference the image must be streamed, the image itself is then transfered through \ac{HTTP}.
 
@@
 
-	Everytime an event is proccessed on one of the endpoints, both sender and receivers state must stay synchronized, otherwise events may behave differently.
+	Every time an event is processed on one of the endpoints, both sender and receivers state must stay synchronized, otherwise events may behave differently.
 
@@
 Delta-State packets transports just the piece of state that changed.
-Event packets informs that an event ocurred over the interactive media.
+Event packets informs that an event occurred over the interactive media.
 
@@
 Traditional \ac{RTP} players can do random access, in contrast, interactive \ac{RTP} players must restore the environment and context at a given time.
-The environment is the initial state, so we can call it a non-interactive descrete media and handle it over \ac{HTTP}.
+The environment is the initial state, so we can call it a non-interactive discrete media and handle it over \ac{HTTP}.
 After the receiver has received the environment, it should calculate the state at the given time.
@@
 
-	Multiparty video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants.
+	Multi-party video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants.
 Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
 
-	Jitsi Video Bridge \footnote{jitsi VB} receives one stream from every participant on a conference, either from a jitsi client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
+	Jitsi Video Bridge \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}} receives one stream from every participant on a conference, either from a jitsi client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
 Although all the participants need to download all the streams from Jitsi Video Bridge Server, typically download rates are much bigger than upload rates, making this solution more feasible.
 
-	Jitsi Video Bridge uses \ac{XMPP} as a signaling protocol and its colibri extension (XEP-0340) to reserve channels for video transmission.
-Although this choice for signalling protocol, Jitsi Video Bridge also supports \ac{SIP} and \ac{Jingle}.
+	Jitsi Video Bridge uses \ac{XMPP} as a signaling protocol and its colibri extension \cite{xep0340} to reserve channels for video transmission.
+Although this choice for signaling protocol, Jitsi Video Bridge also supports \ac{SIP}.
 

### 2015-05-06 tese/section03.tex
done syntax review, I may talk about sig-o-fly and identity stuff

@@
 
-\textbf{\color{red}==================================}
-
   Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to low latency networks leading to experience pauses during playback.
@@
   \ac{SVG} is an \ac{XML} based format that incorporates the animation module of \ac{SMIL}.
-Currently \ac{SVG} allows to add movement and animate attributes of elements.
-When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real time through \ac{DOM} \ac{API}, besides that, it also allows to call \textit{JavaScript} functions on events such as animation end, mouse over and mouse click.
+Currently, \ac{SVG} allows to add movement and animate attributes of elements.
+When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real-time through \ac{DOM} \ac{API}, besides that, it also allows to call \textit{JavaScript} functions on events such as animation end, mouse over and mouse click.
 
-  Video and audio functionalities are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind \textit{JavaScript} functions for different kinds of events over video and audio elements.
+  Video and audio features are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind \textit{JavaScript} functions for different kinds of events over video and audio elements.
 
-  Back in 1995, \textit{flash}\footnote{\url{flash site}} was developed for web-based animations, introducing video support in 2002, flash started to grow after that.
-Concurrency players, at that time, were focused on playing video and audio, flash had vector graphics and they were focused on streaming \textit{on-demand} video across multiple platforms.
-VP6 was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on \textit{Internet} connection latency.
+  Back in 1995, \textit{flash}\footnote{\url{http://www.adobe.com/products/flash.html}} was developed for web-based animations, introducing video support in 2002, \textit{flash} started to grow after that.
+Concurrency players, at that time, were focused on playing video and audio, \textit{flash} had vector graphics and they were focused on streaming \textit{on-demand} video across multiple platforms.
+\textit{VP6} was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on \textit{Internet} connection latency.
 
-  Adobe Flash was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
+  \textit{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
 
@@
   \ac{FMS} can stream video and audio to one or more subscribers by sending a separate copy for each subscriber.
-With \ac{RTMFP} is possbile to stream video between flash players, allowing a publisher breaking a stream into pieces that can be cooperatively distributed in a P2P mesh, \ac{RTMFP} uses \ac{UDP} to speed packet delivery, although is not reliable, it is well suited for video streaming, like \ac{WebRTC} flash players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
+With \ac{RTMFP} is possible to stream video between flash players, allowing a publisher breaking a stream into pieces that can be cooperatively distributed in a P2P mesh, \ac{RTMFP} uses \ac{UDP} to speed packet delivery, although is not reliable, it is well suited for video streaming, like \ac{WebRTC} flash players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
 
-  Although \ac{HTML}5, \textit{JavaScript}, \ac{CSS} and \ac{WebRTC} are implementing some functionalities of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich \textit{Internet} applications.
+  Although \ac{HTML}5, \textit{JavaScript}, \ac{CSS} and \ac{WebRTC} are implementing some features of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich \textit{Internet} applications.
 It is also important to note that \ac{HTML}5 is more compatible with mobile devices than adobe flash.
 
-  Like \textit{Flash}, \textit{Microsoft Silverlight} is a cross browser plug-in and platform that is used to develop rich \textit{Internet} applications.
+  Like \textit{Flash}, \textit{Microsoft Silverlight}\footnote{\url{http://www.microsoft.com/silverlight/}} is a cross browser plug-in and platform that is used to develop rich \textit{Internet} applications.
 It supports vector graphics, animation and video.
 Compared to flash, which uses \textit{ActionScript}, \textit{Silverlight} applications can use languages like C\#, \textit{VisualBasic} and \ac{XAML}.
-\textit{Silverlight} uses a technique called \textit{Smooth Streaming} from \textit{IIS Media Service} that consists on delivering video in real time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.
+\textit{Silverlight} uses a technique called \textit{Smooth Streaming} from \textit{IIS Media Service} that consists on delivering video in real-time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.
 
-  By using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \textit{JavaScript} and \ac{SVG}, it's possible to raise communications to a new level.
+  Using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \textit{JavaScript} and \ac{SVG}, will make possible to raise communications to a new level.
 For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.
@@
 
-  In order to give users a personalized communication channel, each user must have a personal web page where its available plug-ins could be downloaded from other peers, after that they can talk in the same language whatever it is.
+  In order to give users a personalized communication channel, each user must have a personal web page where its available plug-ins could be downloaded from other peers, after that, they can talk in the same language whatever it is.

### 2015-05-06 tese/section04.tex
done syntax review, I may talk about sig-o-fly and identity stuff

@@
 Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity.
-Technologies like this are appearing along this days, but they couldn't be possible years ago because technology was limited or unavailable.
-Although todays technology is limited on some aspects, we are doing progress in order to improve the web ecosystem, by creating standards and migrating to newer technologies.
+Technologies like these are appearing along these days, but they couldn't be possible years ago because technology was limited or unavailable.
+Although today's technology is limited on some aspects, we are doing progress in order to improve the web ecosystem, by creating standards and migrating to newer technologies.
 
  Our first concern on real time collaboration applications, besides the communication itself, is the data storage and representation.
-Because most browsers are recommended to limit local storage to at least five megabytes per origin, storing multimedia content is not a viable solution.
+Storing multimedia content is not a viable solution because most browsers are recommended to limit local storage to at least five megabytes per origin.
 
-	If, for instance, one wants to rewind a real time video, recordings will be needed from who is streaming the video.
+	If, for instance, one wants to rewind a real-time video, recordings will be needed from whom is streaming the video.
 
@@
 
-\ac{RTCP} is used for controlling \ac{RTP} multimedia streams, it provides bandwidth statistics and control information that can be used for changing the quality of the stream in real time.
+\ac{RTCP} is used for controlling \ac{RTP} multimedia streams, it provides bandwidth statistics and control information that can be used for changing the quality of the stream in real-time.
 
-  \ac{RTP} allows to change its requirements and add extensions to it with profiles, one of the most used is the \ac{RTP} profile for audio and video \cite{rfc3551} which lists the payload encodings and compression algorithms.
+  \ac{RTP} allows to change its requirements and add extensions to it with profiles, one of the most used is the \ac{RTP} profile for audio and video \cite{rfc3551} which lists the payload encoding and compression algorithms.
 This profile also assigns a name to each encoding which may be used other protocols like \ac{SDP}.
@@
 
-  Both \ac{SRTP} and \ac{SRTCP} uses \ac{AES} by default, which is symmetric-key algorithm for data encryption.
+  Both \ac{SRTP} and \ac{SRTCP} use \ac{AES} by default, which is symmetric-key algorithm for data encryption.
 Each packet is encrypted using a distinct key-stream, otherwise using a single key-stream with \ac{AES} on \ac{CBC} would make impossible to recover from packet loss.
@@
 	Streaming protocols like \ac {RTP} were designed for continuous and non-interactive media types, such as audio and video.
-Discrete and non-interactive media don't need to streamed through \ac{RTP} because they don't change with.
-For example if an image appears in a specific time interval, just the \ac{HTML} or \textit{JavaScript} that will reference the image must be streamed, the image itself is then transfered through \ac{HTTP}.
+Discrete and non-interactive media don't need to be streamed through \ac{RTP} because they don't change with.
+For example, if an image appears in a specific time interval, just the \ac{HTML} or \textit{JavaScript} that will reference the image must be streamed, the image itself is then transferred through \ac{HTTP}.
 
 	In order to play any kind of stream, a player for interactive stream is needed for downloading an environment, decoding the \ac{RTP} payload to determine the state and display it to the user.
-Streaming interactive media like the combination of \ac{HTML}, \ac{CSS} and JavaScript requires more than interpreting the code, a streamed user interface may contain an internal state that is not shown on code.
+Streaming interactive media like the combination of \ac{HTML}, \ac{CSS} and JavaScript require more than interpreting the code, a streamed user interface may contain an internal state that is not shown on code.
 
@@
 	To achieve synchronization of interactive data most packets have three types: \textit{State}, \textit{Delta-State} and \textit{Event}.
-State packet defines environment's complete state.
+State packet defines the environment complete state.
 Delta-State packets transports just the piece of state that changed.
@@
 	If the \ac{RTP} recorder controls the correct data to send to the receivers, it cannot be a simple \ac{RTP} recorder as it must compute the state or delta-state to send.
-Therefore, if the receiver receives all recorded packets, it can calculate the current state from a nearest complete state.
-Streaming too much complete states, results on more precise random accesses but the trade-off is the higher bandwidth usage and used storage space on the recording server.
+Therefore, if the receiver receives all recorded packets, it can calculate the current state from the previous complete state.
+Streaming too much complete states, results on more precise random accesses, but the trade-off is the higher bandwidth usage and used storage space on the recording server.
 On the other hand, if there are fewer complete states recorded followed by delta-states, the recorded stream will occupy less storage space, but random accesses will be less granular.
 
-	By recording and streaming the interactive media's complete state periodically, it is possible to restore the media state even if messages are lost.
+	It is possible to restore the media state even if messages are lost by recording and streaming the interactive media's complete state periodically.
 
@@
 
-  {\color{red}[Model View Controller (Suited for interactive RTP)]}
+  {\color{fade}[Model View Controller (Suited for interactive RTP)]}
 
@@
 
-	Jitsi Video Bridge \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}} receives one stream from every participant on a conference, either from a jitsi client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
-Although all the participants need to download all the streams from Jitsi Video Bridge Server, typically download rates are much bigger than upload rates, making this solution more feasible.
+	\textit{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}} receives one stream from every participant on a conference, either from a \textit{jitsi} client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
+Although all the participants need to download all the streams from \textit{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
 
-	Jitsi Video Bridge uses \ac{XMPP} as a signaling protocol and its colibri extension \cite{xep0340} to reserve channels for video transmission.
-Although this choice for signaling protocol, Jitsi Video Bridge also supports \ac{SIP}.
+	\textit{Jitsi Video Bridge} uses \ac{XMPP} as a signaling protocol and its \textit{colibri} extension \cite{xep0340} to reserve channels for video transmission.
+Although this choice for signaling protocol, \textit{Jitsi Video Bridge} also supports \ac{SIP}.
 
-  {\color{red}[Identity]}
-
-\quote{There is strong growth in the deployment of devices that integrate regular Web technologies such as HTML, CSS, and SVG, coupled with various device APIs.}
+  {\color{fade}[Identity is that important?]}

### 2015-05-06 tese/section02.tex
added sig-o-fly, a little bit running

@@
 
-\textbf{\color{red}==================================}
+  A more interesting approach for signaling would be \ac{SigOfly}\cite{sigofly} which allows inter-domain real-time communications.
+\ac{SigOfly} achieves inter-domain communication by making use of the \textit{Identity Providers} of each peer.
 
-  {\color{red}[Sig-o-fly]}
+  The caller entity downloads a page with all the code need, also known as messaging stub, to communicate with the called party, this code contains an implementation of the signaling protocol used in order to communicate to the called peer.
+If the called party domain is being overused it is possible to switch the caller and called parties role, after that the called entity downloads the stub code from the caller domain instead.
+
+  \ac{SigOfly} is an approach very flexible because participants on a video call are not tied to just one type o signaling implementation.
+Another important aspect of  \ac{SigOfly} is the ability to perform multi-party conversations either through a \textit{Mesh Topology} or a \textit{Multipoint Control Unit}.

### 2015-05-06 tese/section01.tex
corrected some minor syntax errors

@@
 
-Jitsi Meet\footnote{\url{http://jitsi.org/Projects/JitsiMeet}} is a \ac{WebRTC} collaborative application that uses Jitsi Videobridge for high quality and scalable video conferences and supports shared document editing.
-Jitsi Videobridge is a server that enables multiparty video calls.
+\textit{Jitsi Meet}\footnote{\url{http://jitsi.org/Projects/JitsiMeet}} is a \ac{WebRTC} collaborative application that uses \textit{Jitsi Videobridge} for high quality and scalable video conferences and supports shared document editing.
+\textit{Jitsi Videobridge} is a server that enables multi-party video calls.

### 2015-05-06 tese/section02.tex
corrected some minor syntax errors

@@
 \ac{HTTP} works on a request followed by a server response, by other words, follows a unidirectional communication.
-Sometimes it's required that some informations are obtained in real time, as we saw, some \ac{NAT}'s don't suport callbacks from servers, one technique to overcome this problem is polling.
+Sometimes it's required that some informations are obtained in real time, as we saw, some \ac{NAT}'s don't support callbacks from servers, one technique to overcome this problem is polling.
 

### 2015-05-06 tese/section03.tex
corrected some minor syntax errors

@@
   \item \textbf{Animation} module contains elements and attributes that define a time based mechanism for composing the effects of animations.
-For example, this module can changes on \ac{XML} or \ac{CSS} attributes like colour and dimensions.
+For example, this module can changes on \ac{XML} or \ac{CSS} attributes like color and dimensions.
 
@@
 
-  \item \textbf{Layout} module contains elements and attributes for colouring and positioning media content, another layout mechanisms are also possible, such as \ac{CSS}.
+  \item \textbf{Layout} module contains elements and attributes for coloring and positioning media content, another layout mechanisms are also possible, such as \ac{CSS}.
 
@@
 
-  \item \textbf{Media Object} module contains elements and attributes for referencing rendering behaviour of external multimedia or control objects.
+  \item \textbf{Media Object} module contains elements and attributes for referencing rendering behavior of external multimedia or control objects.
 

### 2015-05-15 curriculo/cv.tex
[ADD] added more on Wave, I must talk about goodow and jupiter

@@
     {\color{red} $\varheartsuit$} Java
-	 (Android, PlayFramework, LWJGL, MapReduce), C, C++ (OpenGL), C\# (XNA, Windows Forms, WPF, lpsolve), VB, Python, Scala, Scheme, HTML, CSS (Bootstrap), JavaScript (jQuery, AJAX, WebGL, WebRTC), PHP (Wordpress), GLSL, HLSL.
+	 (Android, PlayFramework, LWJGL, MapReduce, GWT, AWS), C, C++ (OpenGL), C\# (XNA, Windows Forms, WPF, lpsolve), VB, Python, Scala, Scheme, HTML, CSS (Bootstrap), JavaScript (jQuery, AJAX, WebGL, WebRTC), PHP (Wordpress), GLSL, HLSL.
    \section{databases}

### 2015-05-15 tese/section04.tex
added sig-o-fly, a little bit running
[ADD] added more on Wave, I must talk about goodow and jupiter

@@
 
-  Two key-stream generators for \ac{AES} were defined: \textit{Segmented Integer Counter Mode} and \textit{f8-mode}.
+  Two key-stream generators for \ac{AES} were defined: \emph{Segmented Integer Counter Mode} and \emph{f8-mode}.
 If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those key-stream generators and it's fixed for each packet.
@@
 Discrete and non-interactive media don't need to be streamed through \ac{RTP} because they don't change with.
-For example, if an image appears in a specific time interval, just the \ac{HTML} or \textit{JavaScript} that will reference the image must be streamed, the image itself is then transferred through \ac{HTTP}.
+For example, if an image appears in a specific time interval, just the \ac{HTML} or \emph{JavaScript} that will reference the image must be streamed, the image itself is then transferred through \ac{HTTP}.
 
@@
 
-	To achieve synchronization of interactive data most packets have three types: \textit{State}, \textit{Delta-State} and \textit{Event}.
+	To achieve synchronization of interactive data most packets have three types: \emph{State}, \emph{Delta-State} and \emph{Event}.
 State packet defines the environment complete state.
@@
 
-	\textit{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}} receives one stream from every participant on a conference, either from a \textit{jitsi} client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
-Although all the participants need to download all the streams from \textit{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
+	\emph{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}} receives one stream from every participant on a conference, either from a \emph{jitsi} client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
+Although all the participants need to download all the streams from \emph{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
 
-	\textit{Jitsi Video Bridge} uses \ac{XMPP} as a signaling protocol and its \textit{colibri} extension \cite{xep0340} to reserve channels for video transmission.
-Although this choice for signaling protocol, \textit{Jitsi Video Bridge} also supports \ac{SIP}.
+	\emph{Jitsi Video Bridge} uses \ac{XMPP} as a signaling protocol and its \emph{colibri} extension \cite{xep0340} to reserve channels for video transmission.
+Although this choice for signaling protocol, \emph{Jitsi Video Bridge} also supports \ac{SIP}.
 
-  {\color{fade}[Identity is that important?]}
+	\ac{OT} technology was originally developed for consistency maintenance and concurrency control over distributed objects, \ac{OT} algorithms are mainly used in collaborative applications such as distributed document edition.
+
+	\emph{Google Wave} was a distributed collaboration platform based on \emph{Jupiter}\cite{jupiter} that adopted \ac{OT} techniques, other Google products such as Google Docs are using this type of technology.
+In 2010 Google stopped the development of Google Wave and released the main components as Open Source code to Apache, the project is currently known as \emph{Apache Wave} and the reference implementation is named as \emph{Wave in a Box}.
+
+	\emph{ShareJS}\footnote{\url{http://sharejs.org/}} is an \ac{OT} \emph{JavaScript} library, developed by the ex \emph{Google Wave} engineer Joseph Gentle, for collaborative text and \ac{JSON} documents edition in real-time.
+
+	\emph{TogetherJS}\footnote{\url{http://togetherjs.com/}} is a JavaScript library that uses \ac{WebRTC} for collaborative web applications.
+It uses \ac{JSON} messages for \ac{OT} concurrency control but it does not provide storage.
+
+	{\color{blue} Goodow [12], is a recent FLOSS framework copying the Google Drive Real-Time API with additional clients for Android and iOS, while providing its own server implementation.}

### 2015-05-16 tese/arch01.tex
[ADD] goodow, [START] architecture

Taking into account the goals of this project and all the technology presented so far. Our proposal is the development of a web application that provides communication and collaboration in real-time.

The requirements for our web application are:

\begin{itemize}
 \item Enrich calls with any kink of multimedia in real-time.
 \item Ability perform tasks on real-time collaborative environment.
 \item Ability to record and playback interactive multimedia.
\end{itemize}

The infrastructure is composed by: Ice Server, Web Server, Stream Server, Signaling Server and Test Clients.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.55\textwidth]{figures/arch.png}
	\caption{System Architecture}
\end{figure}

Although the core infrastructure is important and crucial for the implementation of our web application, there is no preference for a specific software running on the web server. An important requirement for the Web Server is the WebSockets support, LAMP server was considered, although the chosen technology for web development is Play Framework with Java, which has a more evident separation between Model, View and Controller components.

For the signaling server there is also no preference between \ac{SIP} and \ac{XMPP}, as both support presence information feature. SigOFly could be adopted as well but using it would be irrelevant for the final result of this project. For sake of simplicity and easy deployment the chosen platform chosen is the Openfire \ac{XMPP} server.

The streaming server will be without any doubt \emph{Jitsi Videobridge} as \emph{Jitsi} team is working with \ac{WebRTC} and their code is open source.

The \ac{ICE} server is not required to be on our infrastructure as a public \ac{ICE} server could be used, if \ac{TURN} is used the network speed could drop due to resource share amongst other users. An \ac{ICE} server will be installed in order to prevent the influence of other users on our application.

On the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} will be installed as web browsers. Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Strophe}, \emph{Modernizr} and \emph{TogetherJS} will be downloaded from the Web Server and executed on the client side.

\emph{Modernizr} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.

\emph{Bootstrap} will be used to make user interface more appellative and responsive. With \emph{bootstrap} it's quite easy to develop applications that adapt to mobile devices with different screen sizes.

\emph{Strophe} library will be essential to communicate with \ac{XMPP} server and clients.

\emph{TogetherJS} will be used for the collaborative component of our web application, other libraries were considered but, as said before, \emph{TogetherJS} does not implement object storage making the object storage implementation free.

### 2015-05-16 tese/state01.tex
[ADD] goodow, [START] architecture

--- 2015-05-06 tese/section01.tex
@@
 Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security.
-There are four types of \ac{NAT} implementations\cite{rfc3489}, \textit{Full Cone NAT}, \textit{Restricted Cone NAT}, \textit{Port Restricted Cone NAT}, \textit{Symmetric NAT}.
+There are four types of \ac{NAT} implementations\cite{rfc3489}, \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.
 
-\textit{Full Cone} \ac{NAT} maps public each \ac{IP} address and port to a private \ac{IP} address and port, any external host can communicate with private hosts through their mapped public address and port.
+\emph{Full Cone} \ac{NAT} maps public each \ac{IP} address and port to a private \ac{IP} address and port, any external host can communicate with private hosts through their mapped public address and port.
 This represents the least restrictive type of \ac{NAT} and as we will later, the unique type of \ac{NAT} that enables real time communications from point to point.
 
-\textit{Restricted Cone} \ac{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host.
+\emph{Restricted Cone} \ac{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host.
 With this type of \ac{NAT}, the private client can be contacted from any port of the same external host.
 
-\textit{Port Restricted Cone} \ac{NAT} works in the same way as Restricted Cone \ac{NAT}, but it only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.
+\emph{Port Restricted Cone} \ac{NAT} works in the same way as Restricted Cone \ac{NAT}, but it only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.
 
-\textit{Symmetric} NAT maps different ports for each connection, as we will see later, this type of \ac{NAT} represents a problem on real time communications.
+\emph{Symmetric} NAT maps different ports for each connection, as we will see later, this type of \ac{NAT} represents a problem on real time communications.
 
-\textit{Asymmetric} \ac{NAT} became a vulgar configuration on the web.
+\emph{Asymmetric} \ac{NAT} became a vulgar configuration on the web.
 As a direct result, problems started to appear, the amount of ports that \ac{IP} makes available is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follow this model to be implemented ineffectively.
@@
 On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a \ac{P2P} network relay to find the best peer, but after that the behavior is much like client-server.
-Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \textit{full cone}.
+Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \emph{full cone}.
 \ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.
@@
 
-\ac{WebRTC}'s \textit{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}.
+\ac{WebRTC}'s \emph{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}.
 \ac{DTLS} encapsulation provides confidentiality, authentication and integrity over the transfered data.
-A \textit{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication.
+A \emph{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication.
 Each data channel direction can be tweaked for reliable or unreliable transmission, the same can be done for order delivery and priority which can also be defined for improving the quality of service for a particular stream compared to another.
 
-\ac{WebRTC}'s \textit{MediaStream} is built on top of \ac{SRTP}, which requires an external mechanism for key exchange.
+\ac{WebRTC}'s \emph{MediaStream} is built on top of \ac{SRTP}, which requires an external mechanism for key exchange.
 \ac{DTLS} keys are negotiated on handshake in order to achieve a secure connection.
@@
 
-\textit{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.
-But compared to \textit{Skype}, \ac{WebRTC} applications don't need to be pre-installed.
+\emph{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.
+But compared to \emph{Skype}, \ac{WebRTC} applications don't need to be pre-installed.
 
-\textit{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is a video conference web application, in the past in order to use \textit{hangouts} on a web browser a plug-in was needed to be installed, nowadays hangouts is using \ac{WebRTC}.
+\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is a video conference web application, in the past in order to use \emph{hangouts} on a web browser a plug-in was needed to be installed, nowadays hangouts is using \ac{WebRTC}.
 
-Jitsi Meet\footnote{\url{http://jitsi.org/Projects/JitsiMeet}} is a \ac{WebRTC} collaborative application that uses Jitsi Videobridge for high quality and scalable video conferences and supports shared document editing.
-Jitsi Videobridge is a server that enables multiparty video calls.
+\emph{Jitsi Meet}\footnote{\url{http://jitsi.org/Projects/JitsiMeet}} is a \ac{WebRTC} collaborative application that uses \emph{Jitsi Videobridge} for high quality and scalable video conferences and supports shared document editing.
+\emph{Jitsi Videobridge} is a server that enables multi-party video calls.

### 2015-05-16 tese/state02.tex
[ADD] goodow, [START] architecture

--- 2015-05-06 tese/section02.tex
@@
   \ac{WebRTC} doesn't implement signaling, different applications may require different protocols, there is no single answer that fits all problems.
-Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, \textit{WebSockets}, \textit{Socket.io} or by implementing a custom protocol.
+Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, \emph{WebSockets}, \emph{Socket.io} or by implementing a custom protocol.
 
@@
   The WebSocket protocol \cite{rfc6455} provides bidirectional communications over a full-duplex socket channel.
-WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to \textit{WebSocket} type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
-WebSockets may not be available on every web browser, frameworks like \textit{socket.io}\footnote{\url{http://socket.io/}} and \textit{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.
+WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
+WebSockets may not be available on every web browser, frameworks like \emph{socket.io}\footnote{\url{http://socket.io/}} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.
 
@@
 Service providers making their \ac{SIP} infrastructures available through WebSockets.
-Frameworks like \textit{jsSIP}\footnote{\url{http://jssip.net/}}, \textit{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \textit{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
+Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \emph{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
 
-  \ac{SIP} with \textit{WebSockets} can be used as a signaling method for \ac{WebRTC} applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
+  \ac{SIP} with \emph{WebSockets} can be used as a signaling method for \ac{WebRTC} applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
 For instance, it's possible to inter-operate web communications with \ac{SIP} networks, mobile and fixed phones.
@@
 
-  Amongst multiple XMPP server solutions are: \textit{ejabberd}\footnote{\url{http://jabberd.im/}}, \textit{Metronome}\footnote{\url{http://lightwitch.org/metronome}}, \textit{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}} and \textit{Prosody}\footnote{\url{http://prosody.im/}}.
-\textit{Ejabberd} is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
+  Amongst multiple XMPP server solutions are: \emph{ejabberd}\footnote{\url{http://jabberd.im/}}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}} and \emph{Prosody}\footnote{\url{http://prosody.im/}}.
+\emph{Ejabberd} is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
 
   A more interesting approach for signaling would be \ac{SigOfly}\cite{sigofly} which allows inter-domain real-time communications.
-\ac{SigOfly} achieves inter-domain communication by making use of the \textit{Identity Providers} of each peer.
+\ac{SigOfly} achieves inter-domain communication by making use of the \emph{Identity Providers} of each peer.
 
@@
   \ac{SigOfly} is an approach very flexible because participants on a video call are not tied to just one type o signaling implementation.
-Another important aspect of  \ac{SigOfly} is the ability to perform multi-party conversations either through a \textit{Mesh Topology} or a \textit{Multipoint Control Unit}.
+Another important aspect of  \ac{SigOfly} is the ability to perform multi-party conversations either through a \emph{Mesh Topology} or a \emph{Multipoint Control Unit}.

### 2015-05-16 tese/state03.tex
[ADD] goodow, [START] architecture

--- 2015-05-06 tese/section03.tex
@@
 An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time.
-\textit{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted on an interactive film by switching between conversations inside a cafe.
+\emph{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted on an interactive film by switching between conversations inside a cafe.
 
   Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment.
-\textit{Hyper-Hitchcock}\cite{hitchcock} is an editor and player of detail-on-demand video.
+\emph{Hyper-Hitchcock}\cite{hitchcock} is an editor and player of detail-on-demand video.
 
@@
 
-  \textit{HyVAL}\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia.
+  \emph{HyVAL}\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia.
 HyVAL defines defines video structure, internal video and external media objects.
 
-  \textit{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically, this approach is quite restrictive if we want to apply hyper-video concepts to videos that don't follow this structure.
+  \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically, this approach is quite restrictive if we want to apply hyper-video concepts to videos that don't follow this structure.
 External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
@@
 
-  \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \textit{Animation}, \textit{Content Control}, \textit{Layout}, \textit{Linking}, \textit{Media Objects}, \textit{SmilText}, \textit{Metainformation}, \textit{Structure}, \textit{Timing}, \textit{Time Manipulations}, \textit{State} and \textit{Transitions}.
+  \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Metainformation}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.
 
@@
   \item \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content.
-This module defines a \textit{head} element that contains non temporal behavior information defined by  \textit{Metainformation}, \textit{Layout} and \textit{Content Control} modules.
-This module also defines defines the \textit{body} element, where all temporal related module information is contained.
+This module defines a \emph{head} element that contains non temporal behavior information defined by  \emph{Metainformation}, \emph{Layout} and \emph{Content Control} modules.
+This module also defines defines the \emph{body} element, where all temporal related module information is contained.
 
   \item \textbf{Timing} module is the most important module on \ac{SMIL} specification, due to its complexity, it is divided into seventeen sub-modules for coordination and synchronization of media over time.
-The three main elements are \textit{seq}, \textit{excl} and \textit{par}, respectively, they play child elements in sequence, one at a time and all at the same time.
+The three main elements are \emph{seq}, \emph{excl} and \emph{par}, respectively, they play child elements in sequence, one at a time and all at the same time.
 
@@
 
-  \item \textbf{Transitions} module defines attributes and elements for transitions across multiple \ac{SMIL} elements according to \textit{Timing} module.
+  \item \textbf{Transitions} module defines attributes and elements for transitions across multiple \ac{SMIL} elements according to \emph{Timing} module.
 
@@
 Like \ac{DOM}, \ac{SMIL} \ac{DOM} is an \ac{API} for \ac{SMIL} documents.
-Allowing \ac{CRUD} operations over \ac{SMIL} documents is an important feature for extending \ac{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \textit{JavaScript} functions.
+Allowing \ac{CRUD} operations over \ac{SMIL} documents is an important feature for extending \ac{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \emph{JavaScript} functions.
 
@@
 \ac{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages.
-Languages like \ac{CSS} and \textit{JavaScript} are typically combined to \ac{HTML} for improving the interaction and appearance of a web page.
+Languages like \ac{CSS} and \emph{JavaScript} are typically combined to \ac{HTML} for improving the interaction and appearance of a web page.
 
@@
 
-  \textit{JavaScript} is an imperative object-oriented language based on \textit{ECMAScript}.
+  \emph{JavaScript} is an imperative object-oriented language based on \emph{ECMAScript}.
 It is used mainly on client-side and executed by a web browser.
-\textit{JavaScript} has its own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plug-ins.
+\emph{JavaScript} has its own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plug-ins.
 
-  \textit{JavaScript} has compatibility issues amongst different web browsers, leading to different behaviors.
-To solve that problem, there are libraries written in \textit{JavasSript}, namely \textit{jQuery} which implements the same functionality for multiple browsers, solving most of incompatibility issues.
+  \emph{JavaScript} has compatibility issues amongst different web browsers, leading to different behaviors.
+To solve that problem, there are libraries written in \emph{JavasSript}, namely \emph{jQuery} which implements the same functionality for multiple browsers, solving most of incompatibility issues.
 
-  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \textit{JavaScript} and \textit{jQuery} which, unlike \cite{ambulant}, doesn't need to be installed and shouldn't have incompatibility issues.
+  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, doesn't need to be installed and shouldn't have incompatibility issues.
 
-  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3, they take into account unsupported web browsers though \textit{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \textit{JavaScript} library that may require plug-ins if new features aren't supported.
+  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3, they take into account unsupported web browsers though \emph{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.
 
@@
 
-  With the emergence of \ac{HTML}5, tags like \textit{video}, \textit{audio} and \textit{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format.
-Another important tag is \textit{canvas} that allows to draw graphics with \textit{JavaScript} on a rectangle within a web page.
+  With the emergence of \ac{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format.
+Another important tag is \emph{canvas} that allows to draw graphics with \emph{JavaScript} on a rectangle within a web page.
 
@@
 Currently, \ac{SVG} allows to add movement and animate attributes of elements.
-When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real-time through \ac{DOM} \ac{API}, besides that, it also allows to call \textit{JavaScript} functions on events such as animation end, mouse over and mouse click.
+When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real-time through \ac{DOM} \ac{API}, besides that, it also allows to call \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.
 
-  Video and audio features are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind \textit{JavaScript} functions for different kinds of events over video and audio elements.
+  Video and audio features are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind \emph{JavaScript} functions for different kinds of events over video and audio elements.
 
-  Back in 1995, \textit{flash}\footnote{\url{http://www.adobe.com/products/flash.html}} was developed for web-based animations, introducing video support in 2002, \textit{flash} started to grow after that.
-Concurrency players, at that time, were focused on playing video and audio, \textit{flash} had vector graphics and they were focused on streaming \textit{on-demand} video across multiple platforms.
-\textit{VP6} was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on \textit{Internet} connection latency.
+  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}} was developed for web-based animations, introducing video support in 2002, \emph{flash} started to grow after that.
+Concurrency players, at that time, were focused on playing video and audio, \emph{flash} had vector graphics and they were focused on streaming \emph{on-demand} video across multiple platforms.
+\emph{VP6} was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on \emph{Internet} connection latency.
 
-  \textit{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
+  \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
 
@@
 
-  Although \ac{HTML}5, \textit{JavaScript}, \ac{CSS} and \ac{WebRTC} are implementing some features of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich \textit{Internet} applications.
+  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} are implementing some features of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich \emph{Internet} applications.
 It is also important to note that \ac{HTML}5 is more compatible with mobile devices than adobe flash.
 
-  Like \textit{Flash}, \textit{Microsoft Silverlight}\footnote{\url{http://www.microsoft.com/silverlight/}} is a cross browser plug-in and platform that is used to develop rich \textit{Internet} applications.
+  Like \emph{Flash}, \emph{Microsoft Silverlight}\footnote{\url{http://www.microsoft.com/silverlight/}} is a cross browser plug-in and platform that is used to develop rich \emph{Internet} applications.
 It supports vector graphics, animation and video.
-Compared to flash, which uses \textit{ActionScript}, \textit{Silverlight} applications can use languages like C\#, \textit{VisualBasic} and \ac{XAML}.
-\textit{Silverlight} uses a technique called \textit{Smooth Streaming} from \textit{IIS Media Service} that consists on delivering video in real-time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.
+Compared to flash, which uses \emph{ActionScript}, \emph{Silverlight} applications can use languages like C\#, \emph{VisualBasic} and \ac{XAML}.
+\emph{Silverlight} uses a technique called \emph{Smooth Streaming} from \emph{IIS Media Service} that consists on delivering video in real-time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.
 
-  Using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \textit{JavaScript} and \ac{SVG}, will make possible to raise communications to a new level.
+  Using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \emph{JavaScript} and \ac{SVG}, will make possible to raise communications to a new level.
 For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.

### 2015-05-16 tese/state04.tex
[ADD] goodow, [START] architecture

--- 2015-05-15 tese/section04.tex
@@
 
-	{\color{blue} Goodow [12], is a recent FLOSS framework copying the Google Drive Real-Time API with additional clients for Android and iOS, while providing its own server implementation.}
+	\emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}} is a collaborative framework with its own server implementation, it supports four types of collaborative elements: String, Lists, Maps and Custom objects.

### 2015-05-26 tese/arch01.tex
[ADD] final structure, ignore portuguese

@@
 
+\subsection{Infrastructure}
+
 The infrastructure is composed by: Ice Server, Web Server, Stream Server, Signaling Server and Test Clients.
@@
 Although the core infrastructure is important and crucial for the implementation of our web application, there is no preference for a specific software running on the web server.
-An important requirement for the Web Server is the WebSockets support, LAMP server was considered, although the chosen technology for web development is Play Framework with Java, which has a more evident separation between Model, View and Controller components.
+An important requirement for the Web Server is the WebSockets support, \emph{LAMP} server was considered, although the chosen technology for web development is \emph{Play Framework} with \emph{Java}, which has a more evident separation between \emph{Model}, \emph{View} and \emph{Controller} components.
 
 For the signaling server there is also no preference between \ac{SIP} and \ac{XMPP}, as both support presence information feature.
-SigOFly could be adopted as well but using it would be irrelevant for the final result of this project.
-For sake of simplicity and easy deployment the chosen platform chosen is the Openfire \ac{XMPP} server.
+\ac{SigOfly} could be adopted as well but using it would be irrelevant for the final result of this project.
+For sake of simplicity and easy deployment the chosen platform chosen is the \emph{Ejabberd} \ac{XMPP} application server since it implements \cite{xep0206} which is crucial for web applications, \emph{Ejabberd}, as said before, is also the \ac{XMPP} server that implements more extensions.
+For simplicity, \emph{Ejabberd} will be configured as a single node instead a member of cluster.
 
-The streaming server will be without any doubt \emph{Jitsi Videobridge} as \emph{Jitsi} team is working with \ac{WebRTC} and their code is open source.
+The streaming server will be \emph{Jitsi Videobridge} as \emph{Jitsi} team is working with \ac{WebRTC} and their code is open source.
 
@@
 
+\subsection{Software}
+
 On the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} will be installed as web browsers.
 Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Strophe}, \emph{Modernizr} and \emph{TogetherJS} will be downloaded from the Web Server and executed on the client side.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.95\textwidth]{figures/apparch.png}
+	\caption{App Architecture}
+\end{figure}
 

### 2015-05-26 tese/intro.tex
[ADD] final structure, ignore portuguese

\subsection{Context}   % English
 	Since the early days of human history, we tried to communicate over far locations, from smoke signals to letters delivered by messengers, real-time communications were limited or even nonexistent. Despite all the efforts made to improve communications, written communication could never replace face to face communication.

	With the advent of the telephone network, communications have taken a very important step for us to feel more connected with whom we communicate. Still, only the human voice was not enough, the invention of cameras and their consequent digitization were a huge step for real-time communications.

 	Although still nothing replace the physical contact with a person while we communicate, we are at a time where we can do more than just a visual and verbal communication.

	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronism between writers.

	Today, we can achieve more, it is possible to write a document collaboratively, correct spelling mistakes without wasting paper, restructure text at any moment, add a video to a newspaper article and more. Although much of what was said seem banal nowadays, none of this was possible before the computer's invention.

\subsection{Problem Statement} % English

	As communications technologies appeared, we adapted the way we communicate. This project aims to re-create how we communicate and make it a process more natural and easy to learn.

	For multiple reasons, some people tend to forget what they ear or see, either due to health problems or lack of sleep, a real-time system is not appropriate for people with short term memory loss, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.

	This project aims to extend audio, text and video communications in order to create more rich and collaborative interfaces, better content organization and time handling. All of these with help of only standard technologies like \ac{WebRTC}, any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.

\subsection{Thesis Contributions} % English

	A web application with an easy to learn user interface will be developed to accomplish our goals. Our application, unnamed yet, is supposed to run on most web browsers that are compatible with JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.

	All the problems faced during the development and limitations will be reported on the essay so that a future project better then ours can be easily and better developed.

\subsection{Article Structure} % English

	The second section is dedicated to related work, which is divided into five subsections. The first subsection describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. The second subsection describes the \ac{WebRTC} technology and the protocols needed to implement our project. The third subsection addresses the signaling component which is not defined on \ac{WebRTC} specifications. The fourth subsection presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. The  fifth and last subsection explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.

	In the third section we propose an architecture in order to develop our Web Application including all the need infrastructure and software.

	The fourth section describes our work methodology and what is our plan in order to solve our problem.

	The fifth section presents the summary and conclusions of our work.

### 2015-05-26 tese/method.tex
[ADD] final structure, ignore portuguese

\subsection{Scientific Methodology} % Section
Nesta seccao devem explicar como pensam abordar a vossa de tese, tendo em conta o problema, as contribuicoes que se propoem dar, a arquitectura definida.

Caso ja tenham iniciado a implementacao, ao ler esta seccao devemos saber, por exemplo:
\begin{itemize}
\item como procederam as escolhas das ferramentas, linguagens de programacao, ambientes de desenvolvimento, HW.
\item como efectuaram a  sua validacao e que problema ja encontraram
\item como prevem desenvolver a arquitectura proposta: por onde vao comecar, como vao integrar os componentes, que dificuldades esperam encontrar e que estrategias planeam para as ultrapassar
\item como pensam testar a solucao e validar a vossa ideia, que processo vao utilizar durante o desenvolvimento para facilitar a realizacao de testes e a recolha de resultados.
\end{itemize}

Caso ainda nao tenham iniciado a implementacao, ao ler esta seccao devemos saber, por exemplo:
\begin{itemize}
\item se pensam usar um simulador ou fazer um prototipo
\item se vai usar um simulador se ja sabem qual o, ou como pensam escolhe-lo
\item se vao fazer um prototipo, que equipamentos/linguagens de programação vão escolher/usar 
\item como pensam testar a solucao e validar a vossa ideia
\end{itemize}

### 2015-05-26 tese/project.tex
added more bibliography
interactive RTP
14 pages, DTLS, TLS, SRTP, SCTP
english syntax for section 1, webrtc stack figure
added references, almost concluding syntax checking for section 3
done syntax review, I may talk about sig-o-fly and identity stuff
added sig-o-fly, a little bit running
removed overfull hbox
[ADD] added more on Wave, I must talk about goodow and jupiter
[ADD] goodow, [START] architecture
[ADD] final structure, ignore portuguese

--- 2015-04-23 tese/template_project_latex.tex
@@
 \usepackage[english]{babel}   
-\usepackage[latin1]{inputenc}
-%\usepackage[utf8x]{inputenc}
+%\usepackage[latin1]{inputenc}
+\usepackage[utf8x]{inputenc}
 \usepackage[T1]{fontenc}
@@
 \usepackage{makeidx}  % allows for indexgeneration
-\usepackage{graphicx}  % allows for indexgeneration
 \usepackage{subfig}  % allows for indexgeneration
@@
 \usepackage{hyperref}
+\usepackage{float}
+\usepackage{mathtools}
+\usepackage{acronym}
+\usepackage{pgfgantt}
+\usepackage{adjustbox}
+\usepackage{rotating}
+\usepackage[graphicx]{realboxes}
+\definecolor{fade}{rgb}{0.8,0.8,0.8}
 
-\usepackage{acronym}
+\newacro{AES}{Advanced Encryption Standard}
 \newacro{API}{Application Programming Interface}
 \newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
-\newacro{CSS}{Cascading Style Sheets}
+\newacro{CBC}{Cipher Block Chaining}
 \newacro{CPU}{Central Processing Unit}
 \newacro{CRUD}{Create, Read, Update and Delete}
+\newacro{CSS}{Cascading Style Sheets}
 \newacro{DOM}{Document Object Model}
+\newacro{DTLS}{Datagram Transport Layer Security}
 \newacro{FMS}{Flash Media Server}
@@
 \newacro{IP}{Internet Protocol}
+\newacro{JSON}{JavaScript Object Notation}
 \newacro{MPEG}{Moving Picture Experts Group}
 \newacro{NAT}{Network Address Translation}
+\newacro{OT}{Operational transformation}
 \newacro{P2P}{Peer-to-peer}
 \newacro{PSTN}{Public Switched Telephone Network}
+\newacro{RFC}{Request For Comments}
+\newacro{RTCP}{Real Time Control Protocol}
+\newacro{RTMFP}{Real-Time Media Flow Protocol}
 \newacro{RTMP}{Real Time Messaging Protocol}
-\newacro{RTMFP}{Real-Time Media Flow Protocol}
-\newacro{RFC}{Request For Comments}
+\newacro{RTP}{Real-time Transport Protocol}
 \newacro{SAMI}{Synchronized Accessible Media Interchange}
+\newacro{SCTP}{Stream Control Transmission Protocol}
 \newacro{SDP}{Session Description Protocol}
+\newacro{SigOfly}{Signaling-On-the-fly}
+\newacro{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
 \newacro{SIP}{Session Initiation Protocol}
-\newacro{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
 \newacro{SMIL}{Synchronized Multimedia Integration Language}
 \newacro{SMS}{Short Message Service}
+\newacro{SRTCP}{Secure RTCP}
+\newacro{SRTP}{Secure Real-time Transport Protocol}
 \newacro{SRT}{SubRip Text}
+\newacro{SSL}{Secure Sockets Layer}
 \newacro{STUN}{Session Traversal Utilities for NAT}
@@
 \newacro{TCP}{Transmission Control Protocol}
+\newacro{TLS}{Transport Layer Security}
 \newacro{TMN}{This Means Nothing}
@@
 \newacro{WebRTC}{Web Real-Time Communication}
+\newacro{WebVTT}{Web Video Text Tracks}
 \newacro{XAML}{eXtensible Application Markup Language}
 \newacro{XEP}{XMPP Extensions}
+\newacro{XHTML}{eXtensible Hypertext Markup Language}
 \newacro{XML}{eXtensible Markup Language}
 \newacro{XMPP}{Extensible Messaging and Presence Protocol}
-\newacro{XHTML}{eXtensible Hypertext Markup Language}
-\overfullrule=2cm
+
+%\overfullrule=2cm
+
+\newganttchartelement*{mymilestone}{
+mymilestone/.style={
+shape=isosceles triangle,
+inner sep=0pt,
+draw=cyan,
+top color=white,
+bottom color=cyan!50
+},
+mymilestone incomplete/.style={
+/pgfgantt/mymilestone,
+draw=yellow,
+bottom color=yellow!50
+},
+mymilestone label font=\slshape,
+mymilestone left shift=0pt,
+mymilestone right shift=0pt
+}
+
+\newgantttimeslotformat{stardate}{%
+\def\decomposestardate##1.##2\relax{%
+\def\stardateyear{##1}\def\stardateday{##2}%
+}%
+\decomposestardate#1\relax%
+\pgfcalendardatetojulian{\stardateyear-01-01}{#2}%
+\advance#2 by-1\relax%
+\advance#2 by\stardateday\relax%
+}
 
@@
 \begin{abstract}
-    In here put your abstract.
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
+Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
+Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
+Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+
 Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
 Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.
-Neque porro quisquam est, qui dolorem ipsum quia dolor sit amet, consectetur, adipisci velit, sed quia non numquam eius modi tempora incidunt ut labore et dolore magnam aliquam quaerat voluptatem.
+
 \keywords{keyworkd1, keyworkd2, keyworkd3}
@@
 \section{Introduction}
-  \input{section01.tex}
-\section{\color{red}Signaling}
-  \input{section02.tex}
-\section{\color{red}Hypermedia}
-  \input{section03.tex}
-\section{\color{red}Applications}
-  \input{section04.tex}
+  \input{intro.tex}
 
-\subsection{Context}   % English
-\subsection{Problem Statement / Solution Statement} % English
-\subsection{Thesis Contributions} % English
-\subsection{Article Structure} % English
-% \section{Proposed Architecture} % English
-\subsection{Methodology} % Section
+\section{Related Work}
+  \input{state01.tex}
+  \input{state02.tex}
+  \input{state03.tex}
+  \input{state04.tex}
+
+\section{Proposed Architecture} % English
+  \input{arch01.tex}
+
+\section{Methodology} % Section
+
+		\input{method.tex}
+
 	\subsection{Planned Schedule} % English
+		\input{schd01.tex}
+
 \section{Conclusions} % English
+
 \subsection{Summary} % English
-\section{Conclusions} % English
+Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as contribuições definidas e a forma como estas se materializaram
+
+\subsection{Conclusions} % English
+Nesta secção devem ser retiradas conclusões do trabalho realizado, em face dos resultados obtidos \cite{exemplo}.
+
 \bibliographystyle{splncs03}

### 2015-05-26 tese/schd01.tex
[ADD] final structure, ignore portuguese

\tikzset{every picture/.style={xscale=0.51,yscale=0.51,transform shape}}

\begin{ganttchart}[vgrid, hgrid]{1}{36}
\gantttitle{2015}{28}
\gantttitle{2016}{8} \\
\gantttitle{Jun}{4}
\gantttitle{Jul}{4}
\gantttitle{Aug}{4}
\gantttitle{Sep}{4}
\gantttitle{Oct}{4}
\gantttitle{Nov}{4}
\gantttitle{Dec}{4}
\gantttitle{Jan}{4}
\gantttitle{Feb}{4} \\
\gantttitlelist{1,...,36}{1}\\
%First Group
\ganttgroup{Thesis}{1}{10} \\
\ganttbar{Task 1}{2}{3} \\
\ganttbar{Task 2}{4}{10} \\
\ganttbar{Task 3}{9}{12}\\
%\ganttlink{elem0}{elem1}
\ganttgroup{Resume}{1}{10} \\

%\ganttmilestone{Milestone 1}{11}
%Second Group
\ganttgroup{Implementation}{1}{1}\ganttgroup{}{4}{8}\ganttgroup{}{11}{26}\ganttgroup{}{29}{30} \\
\ganttbar{ICE configuration}{1}{1} \\
\ganttbar{XMPP configuration}{1}{1} \\
%\ganttlink{elem4}{elem5}
%\ganttmilestone{Milestone 1}{11}
%Third Group
\ganttbar{WebRTC basics}{4}{4} \\
\ganttbar{Video \& Audio Communication}{4}{5} \\
\ganttbar{Non-Interactive Record \& Playback}{6}{8} \\
\ganttbar{Interactive Record \& Playback}{11}{14}\\
\ganttmilestone{First Prototype}{18}\\
\ganttbar{Collaborative Environment}{15}{18}\\
\ganttmilestone{Second Prototype}{22} \\
\ganttbar{Interface Improvement I}{19}{22} \\
\ganttbar{Usability Tests I}{23}{23} \\
\ganttmilestone{Third Prototype}{25} \\
\ganttbar{Interface Improvement II}{24}{25} \\
\ganttbar{Usability Tests II}{26}{26} \\
\ganttbar{Performance Tests}{29}{30} \\
\ganttbar{Performance Improvements}{30}{30} \\
\ganttmilestone{Final Prototype}{30}

%\ganttlink{elem8}{elem9}

\ganttlink{elem1}{elem2}
\ganttlink{elem2}{elem3}
\ganttlink{elem9}{elem11}
\ganttlink{elem10}{elem11}
\ganttlink[link type=F-S]{elem12}{elem13}
\ganttlink{elem13}{elem14}
\ganttlink[link type=F-S]{elem14}{elem16}
\ganttlink[link type=F-S]{elem16}{elem18}
\ganttlink[link type=F-S]{elem18}{elem19}
\ganttlink[link type=F-S]{elem19}{elem21}
\ganttlink[link type=F-S]{elem21}{elem22}
\end{ganttchart}

### 2015-05-27 tese/intro.tex
need conclusions and abstract

@@
 \subsection{Context}   % English
- 	Since the early days of human history, we tried to communicate over far locations, from smoke signals to letters delivered by messengers, real-time communications were limited or even nonexistent.
+ 	Since the early days of Human History, we tried to communicate over far locations, from smoke signals to letters delivered by messengers, real-time communications were limited or even nonexistent.
 Despite all the efforts made to improve communications, written communication could never replace face to face communication.

### 2015-05-27 tese/method.tex
need conclusions and abstract

@@
 \subsection{Scientific Methodology} % Section
-Nesta seccao devem explicar como pensam abordar a vossa de tese, tendo em conta o problema, as contribuicoes que se propoem dar, a arquitectura definida.
 
-Caso ja tenham iniciado a implementacao, ao ler esta seccao devemos saber, por exemplo:
-\begin{itemize}
-\item como procederam as escolhas das ferramentas, linguagens de programacao, ambientes de desenvolvimento, HW.
-\item como efectuaram a  sua validacao e que problema ja encontraram
-\item como prevem desenvolver a arquitectura proposta: por onde vao comecar, como vao integrar os componentes, que dificuldades esperam encontrar e que estrategias planeam para as ultrapassar
-\item como pensam testar a solucao e validar a vossa ideia, que processo vao utilizar durante o desenvolvimento para facilitar a realizacao de testes e a recolha de resultados.
-\end{itemize}
+The most important task in order to implement our project is the infrastructure, without it our project wouldn't work.
+The basic infrastructure will be our first concern, namely the Web Server configuration, ICE and signaling servers.
 
-Caso ainda nao tenham iniciado a implementacao, ao ler esta seccao devemos saber, por exemplo:
-\begin{itemize}
-\item se pensam usar um simulador ou fazer um prototipo
-\item se vai usar um simulador se ja sabem qual o, ou como pensam escolhe-lo
-\item se vao fazer um prototipo, que equipamentos/linguagens de programação vão escolher/usar 
-\item como pensam testar a solucao e validar a vossa ideia
-\end{itemize}
+The Web development can only start after the basic infrastructure configuration. When leading with streams, the infrastructure will be complemented with stream servers.
+
+There will be four prototypes, the first one should implement most of streaming and collaborative functionalities.
+The second prototype will have the same functionalities but a more friendly user interface which will be tested and reviewed by potential users.
+The third prototype will be an improvement based on the users feedback.
+In the final prototype we will improve our system performance.
+
+Briefly, our web application will be tested with real users and unit tests, the system in general will be validated with benchmarks such as analyzing the its behavior and delays by growing the amount of users and data.

### 2015-05-29 tese/arch01.tex
[ADD] modules, changed titles

@@
 
-\subsection{Infrastructure}
+\subsection{Modules}
+	Our project is divided into five modules.
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.55\textwidth]{figures/archs.png}
+	\caption{Project Modules}
+\end{figure}
+
+ The Web Server sends Web Pages, containing the other modules information, to the client, after this, the client contacts all the other modules.
+The ICE server will be used for \ac{NAT} traversal.
+The Stream Server will be used to record streams for further playback.
+The signaling server will be responsible for user and calls management.
+The web server will provide libraries to the client in order to interact with the other modules.
+
+\subsection{Implementation Proposal}
 The infrastructure is composed by: Ice Server, Web Server, Stream Server, Signaling Server and Test Clients.
@@
 For sake of simplicity and easy deployment the chosen platform chosen is the \emph{Ejabberd} \ac{XMPP} application server since it implements \cite{xep0206} which is crucial for web applications, \emph{Ejabberd}, as said before, is also the \ac{XMPP} server that implements more extensions.
-For simplicity, \emph{Ejabberd} will be configured as a single node instead a member of cluster.
+Initially, for simplicity, \emph{Ejabberd} will be configured as a single node instead a member of cluster.
 
@@
 An \ac{ICE} server will be installed in order to prevent the influence of other users on our application.
-
-\subsection{Software}
 

### 2015-05-29 tese/intro.tex
[ADD] modules, changed titles

@@
 	With the advent of the telephone network, communications have taken a very important step for us to feel more connected with whom we communicate.
-Still, only the human voice was not enough, the invention of cameras and their consequent digitization were a huge step for real-time communications.
+Still, only the human voice was not enough, the invention of cameras and consequent video digitization were a huge step for real-time communications.
 
@@
 
-\subsection{Thesis Contributions} % English
+\subsection{Thesis Goals} % English
 
-	A web application with an easy to learn user interface will be developed to accomplish our goals.
+	A web application with an easy to learn user interface will be developed to accomplish solve our problem.
 Our application, unnamed yet, is supposed to run on most web browsers that are compatible with JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
@@
 
-\subsection{Article Structure} % English
+\subsection{Document Structure} % English
 
-	The second section is dedicated to related work, which is divided into five subsections.
-The first subsection describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints.
-The second subsection describes the \ac{WebRTC} technology and the protocols needed to implement our project.
-The third subsection addresses the signaling component which is not defined on \ac{WebRTC} specifications.
+	Section \ref{related} is dedicated to related work, which is divided into five subsections.
+
+	Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints.
+
+	Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project.
+
+	Section \ref{signaling} addresses the signaling component which is not defined on \ac{WebRTC} specifications.
 The fourth subsection presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity.
-The  fifth and last subsection explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
 
-	In the third section we propose an architecture in order to develop our Web Application including all the need infrastructure and software.
+	Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
 
-	The fourth section describes our work methodology and what is our plan in order to solve our problem.
+	In section \ref{arch} we propose an architecture in order to develop our Web Application including all the need infrastructure and software.
 
-	The fifth section presents the summary and conclusions of our work.
+	Section \ref{meth} describes our work methodology and what is our plan in order to solve our problem.
+
+	Section \ref{concl} presents the summary and conclusions of our work.

### 2015-06-01 tese/intro.tex
fixed some layout issues

@@
 
- 	Although still nothing replace the physical contact with a person while we communicate, we are at a time where we can do more than just a visual and verbal communication.
-
 	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronism between writers.
@@
 
+	As Martin Geddes\cite{geddes} states, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
+The concept of structured voice and video synchronized with hypermedia is called hypervoice.
+
 \subsection{Problem Statement} % English
 
-	As communications technologies appeared, we adapted the way we communicate. This project aims to re-create how we communicate and make it a process more natural and easy to learn.
+	As communications technologies appeared, we adapted the way we communicate.
+This project don't aims to replace our video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
 
@@
 
-	Section \ref{signaling} addresses the signaling component which is not defined on \ac{WebRTC} specifications.
-The fourth subsection presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity.
+	Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications.
+
+	Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity.
 

### 2015-06-01 tese/state02.tex
[ADD] modules, changed titles
fixed some layout issues

@@
-\subsection{Signaling, meet and get to know}
+\subsection{Signaling, meet and get to know}\label{signaling}
 
@@
 
-  The WebSocket protocol \cite{rfc6455} provides bidirectional communications over a full-duplex socket channel.
+  The WebSocket protocol \cite{rfc6455} allows bidirectional communications over a full-duplex socket channel.
 WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
@@
 
-  Amongst multiple XMPP server solutions are: \emph{ejabberd}\footnote{\url{http://jabberd.im/}}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}} and \emph{Prosody}\footnote{\url{http://prosody.im/}}.
-\emph{Ejabberd} is the one that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
+  Today, multiple XMPP server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}} and \emph{Prosody}\footnote{\url{http://prosody.im/}}.
+\emph{Ejabberd} is the server that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
 
-  A more interesting approach for signaling would be \ac{SigOfly}\cite{sigofly} which allows inter-domain real-time communications.
-\ac{SigOfly} achieves inter-domain communication by making use of the \emph{Identity Providers} of each peer.
+  Another interesting approach for signaling would be \ac{SigOfly} \cite{sigofly} which allows inter-domain real-time communications.
+\ac{SigOfly} provides inter-domain communication by making use of the \emph{Identity Providers} of each peer.
 

### 2015-06-01 tese/state04.tex
[ADD] modules, changed titles
fixed some layout issues

@@
-\subsection{Extending collaboration tools with time manipulation}
+\subsection{Extending collaboration tools with time manipulation}\label{collab}
 
@@
 
-	Streaming protocols like \ac {RTP} were designed for continuous and non-interactive media types, such as audio and video.
+	Streaming protocols like \ac {RTP} and \ac{RTSP} were designed for continuous and non-interactive media types, such as audio and video.
 Discrete and non-interactive media don't need to be streamed through \ac{RTP} because they don't change with.
@@
 
-  {\color{fade}[Model View Controller (Suited for interactive RTP)]}
+	\cite{interactive_stream} concluded that the ability to extract the objects state in order to synchronize them and the ability intercept events in order to control remote objects, are realized using the \ac{MVC} concept.
+This concept separates three components within an application.
+The \emph{Model} represents the information itself.
+The \emph{View} component shows the \emph{Model} to the user in a suitable and interactive way.
+The \emph{Controller} represents an action from the user to the \emph {Model}.
 
-	With such an interactive \ac{RTP} recorder it is then possible to record, play, fast forward, fast rewind, stop and jump to random positions.
+	Using \ac{MVC} concept will make possible to implement an interactive \ac{RTP} application that records, play, fast forward, fast rewind, stop and jump to random positions.
 
-  \cite{interactive_stream} proposed an \ac{RTP} profile for real-time transmission of interactive media.
+  \cite{interactive_stream} also proposed an \ac{RTP} profile for real-time transmission of interactive media.
 This new profile reuses much of video and audio profile implementation, integrating the interactive component.

### 2015-06-01 tese/intro.tex
improved intro

@@
 
-	As Martin Geddes\cite{geddes} states, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
+	As \cite{geddes} states, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
 The concept of structured voice and video synchronized with hypermedia is called hypervoice.
@@
 	As communications technologies appeared, we adapted the way we communicate.
-This project don't aims to replace our video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
+This project don't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
 
-	For multiple reasons, some people tend to forget what they ear or see, either due to health problems or lack of sleep, a real-time system is not appropriate for people with short term memory loss, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
+	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see, either due to health problems or lack of sleep, a real-time system is not appropriate for people with short term memory loss, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
+Real-time communication applications can make a difference on business, education and health sectors by providing tools to understand how we learn and work together.
 
-	This project aims to extend audio, text and video communications in order to create more rich and collaborative interfaces, better content organization and time handling.
+	This project aims to extend audio, text and video communications in order to create rich and collaborative interfaces with better content organization and time handling.
 All of these with help of only standard technologies like \ac{WebRTC}, any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.

### 2015-06-01 tese/intro.tex
Reviewed introduction.

--- 2015-05-29 tese/intro.tex
@@
 \subsection{Context}   % English
- 	Since the early days of Human History, we tried to communicate over far locations, from smoke signals to letters delivered by messengers, real-time communications were limited or even nonexistent.
+Since the early days of Human History, we tried to communicate over far locations, from smoke signals to letters delivered by messengers.
+Real-time communications were limited or even nonexistent.
 Despite all the efforts made to improve communications, written communication could never replace face to face communication.
+	With the advent of the telephone network, communications have taken a very important step for us to feel more connected with whom we communicate.
+Still, only the human voice was not enough, and the invention of cameras and consequent video digitization were a huge step for real-time communications.
 
-	With the advent of the telephone network, communications have taken a very important step for us to feel more connected with whom we communicate.
-Still, only the human voice was not enough, the invention of cameras and consequent video digitization were a huge step for real-time communications.
-
- 	Although still nothing replace the physical contact with a person while we communicate, we are at a time where we can do more than just a visual and verbal communication.
-
+ 	Although nothing replaces the physical contact with a person while we communicate, we are at a time where we can do more than just  visual and verbal communication.
 	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronism between writers.
@@
 	Today, we can achieve more, it is possible to write a document collaboratively, correct spelling mistakes without wasting paper, restructure text at any moment, add a video to a newspaper article and more.
-Although much of what was said seem banal nowadays, none of this was possible before the computer's invention.
+Although much of what was said seems banal nowadays, none of this was possible before the computer's invention.
 
@@
 	As communications technologies appeared, we adapted the way we communicate. This project aims to re-create how we communicate and make it a process more natural and easy to learn. 
+%RP objectivo demasiado ambicioso.
 
-	For multiple reasons, some people tend to forget what they ear or see, either due to health problems or lack of sleep, a real-time system is not appropriate for people with short term memory loss, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
+	For multiple reasons, some people tend to forget what they ear or see, either due to health problems or lack of sleep,
+       %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
+        a real-time system is not appropriate for people with short term memory loss, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
+        %RP estes dois ultimos pontos são melhores.
+Não fales em pessoas com problemas de memoria mas em termos de lidar com grandes quantidades de informação, realizarmos comunicações em grupo (ainda nãofalaste diss) e nem sempre ser possível estar presente.
+        %RP também há outros mercados, como online learning, projectos de longa duração que vão evoluindo e onde é preciso revisitar o que aconteceu antes.
+Fins histórios.
+Necessidade de manter registos devido a requisitos legais/registar história.
 
 	This project aims to extend audio, text and video communications in order to create more rich and collaborative interfaces, better content organization and time handling.
-All of these with help of only standard technologies like \ac{WebRTC}, any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
+All of these resorting only to standard technologies like \ac{WebRTC} any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
+        %RP esta última frase está confusa.
+
+        %Estás muito parco nos objectivos do projecto.
+Acho que falta dizer que é um sistema para complementar/subsitutir actuais sistemas de comunicação audio/video real-time, adicionando-lhe a capacidade de: adicionar conteúdo posteriomente (e.g.
+anotações) por forma a aumentar o seu valor; permitir acesso posterior (mesmo que apenas 5 minutos depois para quem chega atrasado); permitir navegar na comunicação (hypermedia, saltos, fast-forward, rewind).
 
@@
 
-	A web application with an easy to learn user interface will be developed to accomplish solve our problem.
-Our application, unnamed yet, is supposed to run on most web browsers that are compatible with JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
+A web application with an easy to learn user interface will be developed to accomplish solving our problem.
+Our application, unnamed yet, is targeted at web browsers that are compatible with JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
+%RP podes elaborar um pouco.
+Objectivos são: realizar o levantamento do estado da arte no espaço do problema para fundamentar uma solução; apresentar uma solução e desenhar uma arquitectura para dar resposta ao problema apresentado; construir protótipo funcional; desenhar e realizar uma avaliação/testes que provem a adequação e validade dos conceitos desenvolvidos (elaborar); escrever um artigo.
 
-	All the problems faced during the development and limitations will be reported on the essay so that a future project better then ours can be easily and better developed.
+	All the problems faced during the development and limitations will be reported on the thesis so that a future project better then ours can be easily and better developed.
+        %RP esta última frase não é muito necessária
 
 \subsection{Document Structure} % English
+This document is structured as follows. Section \ref{related} presents the related work.
 
-	Section \ref{related} is dedicated to related work, which is divided into five subsections.
+%, which is divided into five subsections. 
+%	Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
+%	Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
+%	Section \ref{signaling} addresses the signaling component which is not defined on \ac{WebRTC} specifications.
+The fourth subsection presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity.
+%	Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
+%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.
 
-	Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints.
-
-	Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project.
-
-	Section \ref{signaling} addresses the signaling component which is not defined on \ac{WebRTC} specifications.
-The fourth subsection presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity.
-
-	Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
-
-	In section \ref{arch} we propose an architecture in order to develop our Web Application including all the need infrastructure and software.
-
-	Section \ref{meth} describes our work methodology and what is our plan in order to solve our problem.
-
+	In Section \ref{arch}, we propose an architecture for an Web Application that fulfills the goals of this thesis, including all the need infrastructure and software.
+	Section \ref{meth} describes our work methodology and our plan for implementing and validating our proposed architecture.
 	Section \ref{concl} presents the summary and conclusions of our work.

### 2015-06-01 tese/intro.tex

@@
 
- 	Although nothing replaces the physical contact with a person while we communicate, we are at a time where we can do more than just  visual and verbal communication.
 	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronism between writers.
-
 	Today, we can achieve more, it is possible to write a document collaboratively, correct spelling mistakes without wasting paper, restructure text at any moment, add a video to a newspaper article and more.
@@
 
+	As \cite{geddes} states, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
+The concept of structured voice and video synchronized with hypermedia is called hypervoice.
+        %RP a citação do geddes faz sentido? Parece estar a falar de voz gerada por computador e não de video conferencia onde o que ouvimos são outras pessoas!
+
 \subsection{Problem Statement} % English
 
-	As communications technologies appeared, we adapted the way we communicate. This project aims to re-create how we communicate and make it a process more natural and easy to learn. 
-%RP objectivo demasiado ambicioso.
+	As communications technologies appeared, we adapted the way we communicate.
+This project don't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
 
-	For multiple reasons, some people tend to forget what they ear or see, either due to health problems or lack of sleep,
+	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see, either due to health problems or lack of sleep.
                %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
-        a real-time system is not appropriate for people with short term memory loss, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
+        A real-time system is not appropriate for people with short term memory loss, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
 %RP estes dois ultimos pontos são melhores.
 Não fales em pessoas com problemas de memoria mas em termos de lidar com grandes quantidades de informação, realizarmos comunicações em grupo (ainda nãofalaste diss) e nem sempre ser possível estar presente.
+        Real-time communication applications can make a difference on business, education and health sectors by providing tools to understand how we learn and work together.
         %RP também há outros mercados, como online learning, projectos de longa duração que vão evoluindo e onde é preciso revisitar o que aconteceu antes.
@@
 
-	This project aims to extend audio, text and video communications in order to create more rich and collaborative interfaces, better content organization and time handling.
-All of these resorting only to standard technologies like \ac{WebRTC} any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
+	This project aims to extend audio, text and video communications in order to create rich and collaborative interfaces with better content organization and time handling.
+All of these with help of only standard technologies like \ac{WebRTC}, any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
+
                 %RP esta última frase está confusa.
@@
 
-%, which is divided into five subsections. 
 %	Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
 %	Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
-%	Section \ref{signaling} addresses the signaling component which is not defined on \ac{WebRTC} specifications.
-The fourth subsection presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity.
+%	Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications. 
+%	Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
 %	Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.

### 2015-06-01 tese/state01.tex
[ADD] modules, changed titles
Reviewed first part of the SoTA

@@
-\subsection{Early days of the Internet and its remaining flaws}
+\subsection{Early days of the Internet and its remaining flaws}\label{early}
 
-The need to build a global communications network in an age when almost nobody had access to technology and the number of future users was unpredictable, caused that some protocols weren't suitable for a huge increase on the amount of publicly known users.
-\ac{IPv4} limits the number of public addresses in such a way that today is scarce \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible to redirect messages to members of its group through their private addresses, each member of the private network is identified publicly by the same \ac{IP} address with a different port, this technique is also known as \ac{NAT}.
+The need to build a global communications network in an age when almost nobody had access to that technology and the number of future users was unpredictable, lead to some protocols not being suitable for the huge growth of users that followed.
+\ac{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}.
+One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each member of the private network is identified publicly by the same \ac{IP} address with a different port.
+%RP o porto não identifica uma máquina (membro da rede privada) mas uma ligação.
+This technique is known as \ac{NAT}.
 
@@
 \end{figure}
+%RP na figura devias indicar que uns tem ips publicos e outros privados. Alternativamente, colocavas metade da nat box dentro da núvem para indicar que esta pertence à Internet e os pc não.
 
 Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security.
-There are four types of \ac{NAT} implementations\cite{rfc3489}, \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.
+%RP initially? E hoje?
+There are four types of \ac{NAT} implementations\cite{rfc3489}: \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.
+%RP apresentas 5. falas em asymmetric. É um erro?
 
-\emph{Full Cone} \ac{NAT} maps public each \ac{IP} address and port to a private \ac{IP} address and port, any external host can communicate with private hosts through their mapped public address and port.
+\emph{Full Cone} \ac{NAT} maps each public \ac{IP} address and port to a private \ac{IP} address and port.
+Any external host can communicate with private hosts through their mapped public address and port.
 This represents the least restrictive type of \ac{NAT} and as we will later, the unique type of \ac{NAT} that enables real time communications from point to point.
@@
 
-\emph{Asymmetric} \ac{NAT} became a vulgar configuration on the web.
-As a direct result, problems started to appear, the amount of ports that \ac{IP} makes available is also small compared to our current needs, worse than that, \ac{NAT} also difficult end-to-end communication, forcing most of applications that follow this model to be implemented ineffectively.
+\emph{Asymmetric} \ac{NAT} became a common configuration on the Internet.
+As a direct result, problems started to appear: the amount of ports that \ac{IP} makes available is also small compared to our current needs; worse than that, \ac{NAT} also difficults end-to-end communication, forcing most applications that follow this model to be implemented ineffectively.
 
-Applications based on multimedia and file sharing have been one of the most strained by \ac{NAT}.
-Those kind applications require real time communication in order to achieve the best performance.
-\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overpass \ac{NAT}, although, none of those can establish direct connections on multiple level \ac{NAT}s.
+%RP explicar que o nat funciona num cenário em que a máquina atrás do nat é apenas um consumidor de informação, trabalhando em modo cliente do modelo cliente-servidor. Em P2P isto falha.
 
-\ac{STUN} servers are quite simple, they receive requests from \ac{NAT}ed clients, the source address of a request is the public address that \ac{NAT} mapped to the client, \ac{STUN} servers will then reply the mapped public address to the client, so it knows its associated public \ac{IP} address and port.
-Symmetric \ac{NAT} changes \ac{IP} port for each different connection, for that reason \ac{STUN} servers will reply the \ac{IP} address and port of their connection, which will be useless to clients connections, that's why Symmetric \ac{NAT} represents a problem for real time communications.
+Applications based on multimedia and file sharing have been one of the most strained by \ac{NAT}. Those kind of applications require real time communication in order to achieve the best performance.
+%RP file sharing não precisa de real-time
+\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}, although, none of those can establish direct connections on multiple level \ac{NAT}s.
+%RP explicar o que são multiple level pois não falaste disto antes? São ``nested''?
 
-On the other hand, \ac{TURN} uses public servers to redirect traffic between private endpoints, it may use a \ac{P2P} network relay to find the best peer, but after that the behavior is much like client-server.
+\ac{STUN} servers are quite simple.
+They receive requests from \ac{NAT}ed clients, with the source address of a request being the public address that \ac{NAT} mapped to the client.
+\ac{STUN} servers will then reply to the client, providing the mapped public address, so it knows its associated public \ac{IP} address and port.
+Symmetric \ac{NAT} changes \ac{IP} port for each different connection, for that reason, when the \ac{STUN} servers reply with the \ac{IP} address and port of their connection, it will be useless for clients to use in other connection connections.
+That is why Symmetric \ac{NAT} represents a problem for real time communications.
+%RP os problemas que relatas têm a ver com P2P ou real-time?
+
+\ac{TURN} uses public servers to relay traffic between private endpoints.
+It may use a \ac{P2P} network relay to find the best peer, but after that, the behavior is much like client-server.
 Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \emph{full cone}.
 \ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.
+%RP ainda não explicaste o que é o ICE
+%RP explicar que para usar TURN as aplicações têm de estar cientes disso.
 
@@
 Clearly this type of communication requires a more expensive infrastructure and, in most cases, more network usage, leading to a worse quality of service.
-The requirements of video communication makes this kind of model unsuitable.
+The requirements of real-time video communication makes this kind of model unsuitable.
+%RP quando falas em infraestrutura mais dispendiosa para o video, está a falar da internet ou ainda estamos a falar de usar ICE?
 
@@
 
-\subsection{Real time communications}
+\subsection{Real time communications}\label{rtc}
 
-\ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser real time communications without installing any additional application or plug-in.
+\ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser based real time communications without installing any additional application or plug-in.
+%RP indicar que agora também já temos webrtc fora do browser devido ao seu sucesso e ao facto de reunir tecnologia state of the art?
 
@@
 \begin{itemize}
-  \item \textbf{MediaStream} allows from the browser to access to camera, microphone and device screen.
+  \item \textbf{MediaStream} allows the browser to access the camera, microphone and the device's screen.
 
   \item \textbf{PeerConnection} acquires connection data and negotiates with peers.
-
-  \item \textbf{DataChannel} allows to send whatever type of data to other peers.
+%RP establishes and negociates a connection with a peer for transmitting real-time video or audio?
+  \item \textbf{DataChannel} provides a channel for exchanging arbitrary data with other peers.
 \end{itemize}
 
-\ac{WebRTC} uses \ac{UDP} for transporting data, which provides lower latencies than \ac{TCP}, but is not reliable and packet order and integrity are not assured.
+\ac{WebRTC} uses \ac{UDP} for transporting data, which provides lower latencies than \ac{TCP}, but is not reliable and does not assure packet order and integrity.
 \ac{SCTP} and \ac{SRTP} are used for streaming data, providing a mechanism for congestion control and partial reliable delivery over \ac{UDP}.
@@
 
-\ac{DTLS} is similar to \ac{TLS}, but on top of \ac{UDP}, the main difference is the inclusion of a sequence number per packet that is used for packet re-ordering on reception and protects from duplicated packets.
+\ac{DTLS} is similar to \ac{TLS}, but is used on top of \ac{UDP}.
+The main difference is the inclusion of a sequence number per packet that is used for packet re-ordering on reception and protects from duplicated packets.
 If a packet sequence number is less than the expected sequence number the packet is discarded.
@@
 By knowing the sequence of messages that are sent and received in \ac{TLS}, timers are used for packet retransmission avoiding acknowledgment messages.
+%RP TLS?
 
@@
 \ac{WebRTC}'s \emph{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}.
-\ac{DTLS} encapsulation provides confidentiality, authentication and integrity over the transfered data.
+\ac{DTLS} encapsulation provides confidentiality, authentication and integrity to the transfered data.
 A \emph{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication.
-Each data channel direction can be tweaked for reliable or unreliable transmission, the same can be done for order delivery and priority which can also be defined for improving the quality of service for a particular stream compared to another.
+Each data channel direction can be configured for reliable or unreliable transmission, the same can be done for order delivery and priority.
+which can also be defined for improving the quality of service of a particular stream over the others.
 
@@
 The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining communications are done through \ac{UDP} without using \ac{DTLS}.
+%RP quais remaining?  A última frase é confusa
 
+\ac{WebRTC} aims to provide a standard platform for real-time audio and video on the Web. It arrives at a time when several proprietary products are well established.
 \emph{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.
@@
 
-\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is a video conference web application, in the past in order to use \emph{hangouts} on a web browser a plug-in was needed to be installed, nowadays hangouts is using \ac{WebRTC}.
+\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is another popular video conference web application.
+In the past, in order to use \emph{hangouts} on a web browser a plug-in had to be installed, nowadays hangouts is using \ac{WebRTC}.
 
@@
 \emph{Jitsi Videobridge} is a server that enables multi-party video calls.
+
+%RP a descricção do skype, hangouts e jistsi cai um pouco do céu.
+Não há uma transição e não é feita uma análise.
+Apenas são indicados.
+Não há comparação entre eles nem com o que pode ser feito por webrtc.

### 2015-06-01 tese/state02.tex
Reviewed second part of the SoTA

@@
-\subsection{Signaling, meet and get to know}\label{signaling}
+\subsection{Signaling: meet and get to know}
+\label{signaling}
 
-  Signaling is the most important process for applications to exchange connection information about peers and servers, their capabilities and meta-data.
-
-  \ac{WebRTC} doesn't implement signaling, different applications may require different protocols, there is no single answer that fits all problems.
-Amongst multiple options, signaling can be done by using \ac{SIP}, \ac{XMPP}, \emph{WebSockets}, \emph{Socket.io} or by implementing a custom protocol.
+  Signaling is the process by which applications exchange connection information about peers and servers, their capabilities and meta-data.
+  \ac{WebRTC} doesn't implement signaling, as different applications may require different protocols, and there is no single answer that fits all problems.
+  Multiple options are available for signaling, which can be performed using \ac{SIP}, \ac{XMPP}, \emph{WebSockets}, \emph{Socket.io} or by implementing a custom protocol.
 
@@
 
-  One of signaling requisites is bi-directional communication over \ac{HTTP}.
-\ac{HTTP} works on a request followed by a server response, by other words, follows a unidirectional communication.
-Sometimes it's required that some informations are obtained in real time, as we saw, some \ac{NAT}'s don't support callbacks from servers, one technique to overcome this problem is polling.
+  One of \ac{WebRTC} signaling's requisites is bi-directional communication over \ac{HTTP}.
+%RP acrescentei o webrtc pois isto não é um requisito geral de signaling. era o que querias dizer?
+  \ac{HTTP} uses a request-response paradigm, where a request is sent by the client, followed by a server response.
+  %in other words, it follows an unidirectional communication.
+%RP de certeza que queres dizer que o http é unidirecional?
+  Sometimes it is required that some information be obtained in real time, but we saw, some \ac{NAT}'s don't support callbacks from servers, preventing them from notifying clients as soon as an event occures.
+One technique to overcome this problem is polling.
+  %RP num documento formal não se usam contrações como it's.
+  %RP o teu ingles é parco em artigos. Faltam muitos the, of, it, etc.
 
-  Polling consists on sending periodic messages that the server responds immediately with empty content or fresh information.
+Polling consists on sending periodic messages to which the server responds immediately with empty content or fresh information.
 Real time communications are unpredictable, if the time between periodic requests is short, most of the time the server will return empty results wasting network bandwidth and energy.
-On the other hand, if the time between periodic requests is large, newer messages may arrive later.
+On the other hand, if the time between periodic requests is large, newer messages may arrive too late.
+%RP  Real time communications are unpredictable. Normalmente falamos de audio e video, que são previsíveis. Estás a referir-te às comunicações em si ou ao signaling e estabelecimento?
 
-  A technique called long polling consists on making the server hold the request until there is fresh information or expiring after some time, after the message receipt, the client makes another request.
-Long polling technique results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which adds data overhead, especially for short messages.
+A technique called long polling consists on making the server hold the request until there is fresh information or expiring it after some time.
+As soon as it receives the reply, the client makes another request.
+Long polling technique results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which add data overhead, especially for short messages.
 
-  The WebSocket protocol \cite{rfc6455} allows bidirectional communications over a full-duplex socket channel.
-WebSocket handshake phase specifies an \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, the remainder messages are done without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
-WebSockets may not be available on every web browser, frameworks like \emph{socket.io}\footnote{\url{http://socket.io/}} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}} uses \ac{HTTP} when there is no support for WebSockets.
+The WebSocket protocol allows bidirectional communications over a full-duplex socket channel \cite{rfc6455}.
+%RP coloca as citações no final da frase. Pode haver excepções onde a citação apenas cobre parte da frase ou precisas de colocar várias a dizer coisas diferentes.
+WebSocket handshake phase specifies a \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, but the remainder messages are exchanged without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
+WebSockets may not be available on every web browser, frameworks like \emph{socket.io}\footnote{\url{http://socket.io/}} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}} fall back to using \ac{HTTP} when there is no support for WebSockets.
 
-  \ac{BOSH}\cite{xep0124} is a technique based on long polling that uses two socket connections and allow sending client messages to server while a previous request is held.
-
-  \ac{BOSH} specification assumes that a connection manager is implemented to handle \ac{HTTP} connections.
-This connection manager is basically a translator from \ac{HTTP} to raw message so the server may think this communication is performed over \ac{TCP}.
-
-  When the connection manager holds for a response for too long it responds with an empty body, this technique prevents an \ac{HTTP} session from expiring when the client is waiting for a response, thus expanding the session time.
-Expiring sessions can be expensive due to the overhead of establishing new connections, it worsens even more when \ac{HTTP} is used over \ac{SSL}.
+  \ac{BOSH}\cite{xep0124} is a technique based on long polling that uses two socket connections and allows sending client messages to the server while a previous request is held.
+  The \ac{BOSH} specification assumes that a connection manager is implemented to handle \ac{HTTP} connections.
+This connection manager is basically a translator from \ac{HTTP} to raw message so that the server may be implemented as if this communication is performed over \ac{TCP}.
+  When the connection manager holds for a response for too long, it responds with an empty body, this technique prevents an \ac{HTTP} session from expiring when the client is waiting for a response, thus expanding the session time.
+Expiring sessions can be expensive due to the overhead of establishing new connections, which is even worse when \ac{HTTP} is used over \ac{SSL}.
 
@@
 The exchange of roles of those two connections allow to pull data from multiple contexts instead of being locked in just one.
-
   If the client has data to send while a request is still open, it establishes a second socket connection to the connection manager to send a new request.
@@
 This results in the connections switching roles.
+  %RP estas duas frases dizem o mesmo da perspectiva do cliente e do servidor. Não consegues simplificar numa só?
 
-  \ac{SIP} \cite{rfc3261} is protocol used for negotiation, creation, modification and finalization of communication sessions between users.
-\ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as signaling protocol.
+  \ac{SIP} \cite{rfc3261} is a protocol used for negotiation, creation, modification and finalization of communication sessions between users.
+\ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as a signaling protocol.
 The advantage of \ac{SIP} is the ability to make video and voice call's applications over \ac{IP} networks.
@@
 
-  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with \ac{PSTN}.
+  \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with the \ac{PSTN}.
   Service providers making their \ac{SIP} infrastructures available through WebSockets.
-Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \emph{sipML5}\footnote{\url{http://sipml5.org/}} are used on client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
-
+%RP a frase anterior é estranha
+  Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \emph{sipML5}\footnote{\url{http://sipml5.org/}} are used on the client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
+%RP quando usas footnotes para indicar sites deves acrescentar ``last accessed (data)''. Isto é importante pois os sites são documentos efémeros que podem dizer uma coisa diferente mais tarde.
   \ac{SIP} with \emph{WebSockets} can be used as a signaling method for \ac{WebRTC} applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
@@
 It is nowadays an open technology for standardized, decentralized, secure and extensible real-time communications.
-
-  \ac{XMPP} messages are \ac{XML} based, which are attractive for applications that need structured messages and rich hypermedia.
-\ac{XMPP} advantage is the addition of extensions, for example \cite{xep0096}, which adds file transfer capabilities between two entities and \cite{xep0045} which enables multi-user chat.
-
-  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH} \cite{xep0206}, which basically consists on long polling.
+  \ac{XMPP} messages are \ac{XML} based, which is attractive for applications that need structured messages and rich hypermedia.
+Another advantage of \ac{XMPP} is the addition of extensions, for example \cite{xep0096}, which adds file transfer capabilities between two entities and \cite{xep0045} which enables multi-user chat.
+  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH} \cite{xep0206}.
+  %RP falta um `` bi-directional communication OVER HTTP is achieved through''?
+  %RP BOSH já foi explicado antes pelo que não vale a pena voltar a falar em long-pooling
   This kind of communication is also possible through WebSockets \cite{rfc7395}.
-
   Today, multiple XMPP server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}} and \emph{Prosody}\footnote{\url{http://prosody.im/}}.
 \emph{Ejabberd} is the server that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
+  %RP fazes muitos paragrafos. Um paragrafo por tecnologia é normalmente suficiente se não forem muito grandes.
 
-  Another interesting approach for signaling would be \ac{SigOfly} \cite{sigofly} which allows inter-domain real-time communications.
+  Another interesting approach for signaling would be \ac{SigOfly} which allows inter-domain real-time communications, while abstracting the protocol used~\cite{sigofly}.
+%RP o ~ antes do cite faz com que não haja um espaço mas não haja uma mudança de linha entre a última palavra e a citação
   \ac{SigOfly} provides inter-domain communication by making use of the \emph{Identity Providers} of each peer. 
-
-  The caller entity downloads a page with all the code need, also known as messaging stub, to communicate with the called party, this code contains an implementation of the signaling protocol used in order to communicate to the called peer.
-If the called party domain is being overused it is possible to switch the caller and called parties role, after that the called entity downloads the stub code from the caller domain instead.
-
-  \ac{SigOfly} is an approach very flexible because participants on a video call are not tied to just one type o signaling implementation.
+  The caller entity downloads a page with all the code needed, also known as messaging stub, to communicate with the called party.
+  This code contains an implementation of the signaling protocol used in order to communicate to the called peer.
+If the called party domain is being overused, it is possible to switch the caller and called parties role, after that the called entity downloads the stub code from the caller domain instead.
+  \ac{SigOfly} is an approach very flexible because participants on a video call are not tied to just one type of signaling implementation.
 Another important aspect of \ac{SigOfly} is the ability to perform multi-party conversations either through a \emph{Mesh Topology} or a \emph{Multipoint Control Unit}.
+%RP os outros não permitem multi-party?

### 2015-06-01 tese/intro.tex
[FIX] middle of SotA Section 1

@@
 
-	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see, either due to health problems or lack of sleep.
+	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see.
                %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
-        A real-time system is not appropriate for people with short term memory loss, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
+        A real-time system is an huge source of information that requires much attention from its users, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
 %RP estes dois ultimos pontos são melhores.
 Não fales em pessoas com problemas de memoria mas em termos de lidar com grandes quantidades de informação, realizarmos comunicações em grupo (ainda nãofalaste diss) e nem sempre ser possível estar presente.
-        Real-time communication applications can make a difference on business, education and health sectors by providing tools to understand how we learn and work together.
+        Real-time communication applications can make a difference on business, education and health sectors by providing tools for teaching and learning online, teamworking and socializing.
         %RP também há outros mercados, como online learning, projectos de longa duração que vão evoluindo e onde é preciso revisitar o que aconteceu antes.
@@
 
-	This project aims to extend audio, text and video communications in order to create rich and collaborative interfaces with better content organization and time handling.
-All of these with help of only standard technologies like \ac{WebRTC}, any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
+	This project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
+creating annotations for improving content search) in order to increase its value.
+Another important aspect of this project is the ability to navigate in time and space by rewind communications, fast-forward, jump to certain points and choose another path.
 
@@
 A web application with an easy to learn user interface will be developed to accomplish solving our problem.
-Our application, unnamed yet, is targeted at web browsers that are compatible with JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
+Our application, unnamed yet, is targeted at web browsers that are compatible with only standard technologies like JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
+Any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
+
 %RP podes elaborar um pouco.
 Objectivos são: realizar o levantamento do estado da arte no espaço do problema para fundamentar uma solução; apresentar uma solução e desenhar uma arquitectura para dar resposta ao problema apresentado; construir protótipo funcional; desenhar e realizar uma avaliação/testes que provem a adequação e validade dos conceitos desenvolvidos (elaborar); escrever um artigo.
+
+Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solves our problem either completely or just a part of it.
+
+We will present an architecture that can meet our goals, implement the respective prototype and test it with real users, unitary tests and benchmarks.
+
+As \cite{geddes} said, the quality of the interaction worsens as the number of users increase.
+In our testing phase we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.
 
@@
 This document is structured as follows. Section \ref{related} presents the related work.
-
 %	Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
@@
 %RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.   
-
 	In Section \ref{arch}, we propose an architecture for an Web Application that fulfills the goals of this thesis, including all the need infrastructure and software.

### 2015-06-01 tese/project.tex
[ADD] modules, changed titles
fixed some layout issues
Reviewed introduction.
[FIX] middle of SotA Section 1

@@
 \usepackage{rotating}
+\usepackage{dirtytalk}
 \usepackage[graphicx]{realboxes}
@@
 \newacro{MPEG}{Moving Picture Experts Group}
+\newacro{MVC}{Model-View-Controller}
 \newacro{NAT}{Network Address Translation}
@@
 \newacro{RTCP}{Real Time Control Protocol}
+\newacro{RTSP}{Real Time Streaming Protocol}
 \newacro{RTMFP}{Real-Time Media Flow Protocol}
@@
 
-\title{Comunicações Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC}
+\title{Hyper-linked Communications: WebRTC enabled asynchronous collaboration}
 \author{%
@@
 \institute{Instituto Superior Técnico}
-
+%RP acrescenta número de aluno, nome do mestrado e ``cadeira de projecto''
 \maketitle              % typeset the title of the contribution
@@
 
-\keywords{keyworkd1, keyworkd2, keyworkd3}
+\keywords{WebRTC, asynchronous, communications, collaboration}
 \end{abstract}
@@
 
-\section{Related Work}
+\section{Related Work}\label{related}
+ This section is structured as follows.
+ Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
+ Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
+ Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications. 
+ Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
+ Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
+%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.
+
   \input{state01.tex}
@@
 
-\section{Proposed Architecture} % English
+\section{Proposed Architecture}\label{arch} % English
   \input{arch01.tex}
 
-\section{Methodology} % Section
+\section{Methodology}\label{meth} % Section
 
@@
 
-\section{Conclusions} % English
+\section{Conclusions}\label{concl} % English
 

### 2015-06-01 tese/state01.tex
[FIX] middle of SotA Section 1

@@
 \ac{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each member of the private network is identified publicly by the same \ac{IP} address with a different port.
+One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \ac{IP} address with a different port.
 %RP o porto não identifica uma máquina (membro da rede privada) mas uma ligação.
@@
 
-Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security.
+Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security, although its current wide usage, \ac{NAT}s are exposing their weaknesses to the application layer.
 %RP initially? E hoje?
@@
 
-\emph{Asymmetric} \ac{NAT} became a common configuration on the Internet.
+\emph{Full Cone}, \emph{Restricted Cone} and \emph{Port Restricted Cone} \ac{NAT}s became the common configurations on the Internet.
 As a direct result, problems started to appear: the amount of ports that \ac{IP} makes available is also small compared to our current needs; worse than that, \ac{NAT} also difficults end-to-end communication, forcing most applications that follow this model to be implemented ineffectively.
+
+Applications behind a \ac{NAT} are prevented to receive incoming connections from the public network, which forces them to behave as a client of a client-server model.
 
@@
 
-Applications based on multimedia and file sharing have been one of the most strained by \ac{NAT}. Those kind of applications require real time communication in order to achieve the best performance.
+Applications based on multimedia and peer-to-peer file sharing have been one of the most strained by \ac{NAT}.
+Those kind of applications require real time communication in order to achieve the best performance.
 %RP file sharing não precisa de real-time
 \ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}, although, none of those can establish direct connections on multiple level \ac{NAT}s.
+%HR XXX I was here
 %RP explicar o que são multiple level pois não falaste disto antes? São ``nested''?

### 2015-06-02 tese/state03.tex
[ADD] modules, changed titles
Reviewed third section of the SoTA

@@
-\subsection{Hypermedia, more than words, more than images}
+\subsection{Hypermedia: more than words, more than images}
+\label{hypermedia}
 
-  Since the early days of video technology, one of the problems that raised with it consisted of how to add more information onto it without generating multiple versions.
-Some implementations like \cite{embedded} added hypermedia information to empty space present on \ac{MPEG} frames in order to provide interactive television, the \ac{MPEG} encoder and decoder were changed in order to handle hypermedia content.
+Since the early days of video technology, one of the problems raised consisted on how to add more information onto video without generating multiple versions.
+%RP depois disto falta uma introdução a explicar que precisas de controlar a apresentação e execução de diversos conteúdos e que para isso vais precisar de uma tecnologia que permite sincronizar e controlar.
+Esta secção examina isso.
 
+Some \ac{MPEG} implementations, like \cite{embedded}, added hypermedia information to empty space present on \ac{MPEG} frames in order to provide interactive television, modifying the \ac{MPEG} encoder and decoder in order to handle hypermedia content.
+%RP esta última frase, com detalhes de implementação fica algo deslocada antes de apresentares os exemplos de hypermedia. Talvez deslocar para mais tarde.
+Hypermedia is a concept that holds the promise of future technology and features but it is also already present in our daily lives.
+%RP por exemplo na publicidade no youtube
+
+  Subtitles are an example of information that might be required.
   The need to translate movies, raised the problem whether it is appropriate to change the original video or audio.
@@
 
-  Amongst multiple formats for subtitles, \ac{SAMI}, and \ac{SRT} are used by video players that support them. Although those formats have styling available, they are quite limited to text.
+  \ac{SAMI}, and \ac{SRT} are two of the multiple formats for subtitles commonly supported by video players. Although those formats have styling available, they are quite limited to text.
 
   Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
-An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time.
-\emph{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted on an interactive film by switching between conversations inside a cafe.
+An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time in a video or audio track.
+\emph{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe.
 
@@
   In order to navigate through a dynamic video, one must be aware of time synchronization and the multiple time flows, it's important that all time, causality and behavior rules are well defined.
+  %RP one must be aware? O utilizador? Ou queres dizer que é preciso definir as relações? O que falas a seguir é sobre linguagens para isso.
 
@@
 HyVAL defines defines video structure, internal video and external media objects.
-
-  \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically, this approach is quite restrictive if we want to apply hyper-video concepts to videos that don't follow this structure.
+  \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically.
+  This approach is quite restrictive if we want to apply hyper-video concepts to videos that don't follow this structure.
 External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
+  %RP não percebo bem a última frase. O que é primary video?
 
   \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films.
-With \ac{SMIL} it's possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, amongst multiple other features.
-
+With \ac{SMIL} it is possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
+  %RP synchronize multiple sections of video? vários videos apresentados em simultâneo?
   \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Metainformation}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.
@@
 
-  \item \textbf{Animation} module contains elements and attributes that define a time based mechanism for composing the effects of animations.
-For example, this module can changes on \ac{XML} or \ac{CSS} attributes like color and dimensions.
+  \item The \textbf{Animation} module contains elements and attributes that define a time based mechanism for composing the effects of animations.
+For example, this module can perform changes on \ac{XML} or \ac{CSS} attributes like color and dimensions.
 
-  \item \textbf{Content Control} module contains elements and attributes that provide optimized alternatives for content delivery.
+  \item The \textbf{Content Control} module contains elements and attributes that provide optimized alternatives for content delivery.
 For example, it could be used to change audio language in function of user's nationality, for videos with multiple audio channels.
 
-  \item \textbf{Layout} module contains elements and attributes for coloring and positioning media content, another layout mechanisms are also possible, such as \ac{CSS}.
+  \item The \textbf{Layout} module contains elements and attributes for coloring and positioning media content. Other layout mechanisms are also possible, such as \ac{CSS}.
 
-  \item \textbf{Linking} module contains elements and attributes for navigational hyperlinking. Navigation can be triggered by events or user interaction.
+  \item The\textbf{Linking} module contains elements and attributes for navigational hyperlinking. Navigation can be triggered by events or user interaction.
 
-  \item \textbf{Media Object} module contains elements and attributes for referencing rendering behavior of external multimedia or control objects.
+  \item The \textbf{Media Object} module contains elements and attributes for referencing rendering behavior of external multimedia or control objects.
 
-  \item \textbf{SmilText} module contains elements and attributes that defines and controls timed text. For example, this module could be used to create labels and captions.
+  \item The \textbf{SmilText} module contains elements and attributes that define and control timed text. For example, this module could be used to create labels and captions.
 
-  \item \textbf{Metainformation} module contains elements and attributes that allows describing the \ac{SMIL} document.
-For example, this module could be used to define a movie details such as category, director, writers and cast.
+  \item The \textbf{Metainformation} module contains elements and attributes that allows describing the \ac{SMIL} document.
+For example, this module could be used to define movie details such as category, director, writers and cast.
 
-  \item \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content.
+  \item The \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content.
 This module defines a \emph{head} element that contains non temporal behavior information defined by  \emph{Metainformation}, \emph{Layout} and \emph{Content Control} modules.
-This module also defines defines the \emph{body} element, where all temporal related module information is contained.
+This module also defines the \emph{body} element, where all temporal related module information is contained.
 
-  \item \textbf{Timing} module is the most important module on \ac{SMIL} specification, due to its complexity, it is divided into seventeen sub-modules for coordination and synchronization of media over time.
-The three main elements are \emph{seq}, \emph{excl} and \emph{par}, respectively, they play child elements in sequence, one at a time and all at the same time.
+  \item The \textbf{Timing} module is the most important module on \ac{SMIL} specification.
+Due to its complexity, it is divided into seventeen sub-modules for coordination and synchronization of media over time.
+The three main elements are \emph{seq}, \emph{excl} and \emph{par}, that, respectively, play child elements in sequence, one at a time and all at the same time.
 
-  \item \textbf{Time Manipulations} module adds time behavior attributes to \ac{SMIL} elements, such as speed, rate or time.
+  \item The \textbf{Time Manipulations} module adds time behavior attributes to \ac{SMIL} elements, such as speed, rate or time.
 
-  \item \textbf{State} module defines attributes that defines the state of \ac{SMIL} elements, such as element visibility, current element time, amount of repeated loops, playing state and many others.
+  \item The \textbf{State} module defines attributes that define the state of \ac{SMIL} elements, such as element visibility, current element time, amount of repeated loops, playing state and many others.
 
-  \item \textbf{Transitions} module defines attributes and elements for transitions across multiple \ac{SMIL} elements according to \emph{Timing} module.
+  \item The \textbf{Transitions} module defines attributes and elements for transitions across multiple \ac{SMIL} elements according to the \emph{Timing} module.
 
@@
 \ac{DOM} makes it easy to inter-operate between imperative and declarative programming languages.
+%RP não arranjas uma referência para o dom?
 
@@
   \ac{SMIL}'s modules are used to synchronize and animate \ac{XHTML} and \ac{SVG} elements.
+  %RP ``are used'' ou ``can be used''?
 
-  In order to create a multimedia rich hyper-call, \ac{SMIL} fits our goals, but it lacks on browser compatibility.
+  \ac{SMIL} fits our goals for creating a multimedia rich hyper-call, but it lacks on browser compatibility.
 Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \ac{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plug-in.
 
+  %RP fazes uma mudanças de tema muito bruscas. Podes usar \subsubsection ou \paragraph para dividir isto melhor. E uma frase entre temas a servir de ``cola'' ou ``separação'' também não fica mal
   \ac{HTML} is a markup language based on \ac{XML} that is used for creating web pages.
 \ac{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages.
-Languages like \ac{CSS} and \emph{JavaScript} are typically combined to \ac{HTML} for improving the interaction and appearance of a web page.
+Languages like \ac{CSS} and \emph{JavaScript} are typically combined with \ac{HTML} for improving the interaction and appearance of a web page.
 
-  \ac{CSS} idea is to separate the structure of an \ac{XML} document from its appearance.
-\ac{CSS} defines styles for \ac{XML} tags based on their name, class, identifier or position, besides static styling \ac{CSS} also supports animation and transitions leading to a more dynamic content.
+  \ac{CSS}'s goal is to separate the structure of an \ac{XML} document from its appearance. \ac{CSS} defines styles for \ac{XML} tags based on their name, class, identifier or position.
+  Besides static styling \ac{CSS} also supports animation and transitions leading to more dynamic content.
+  %RP Tens muitas alturas em que não começas frases novas. Usas virgulas para separar conteúdo que não faz sentido ficar na mesma frase pois começa com um novo sujeito seguido de verbo.
 
@@
 
-  \emph{JavaScript} has compatibility issues amongst different web browsers, leading to different behaviors.
-To solve that problem, there are libraries written in \emph{JavasSript}, namely \emph{jQuery} which implements the same functionality for multiple browsers, solving most of incompatibility issues.
+  \emph{JavaScript} has compatibility issues among the different web browsers, leading to different behaviors.
+To solve that problem, there are libraries written in \emph{JavasSript}, namely \emph{jQuery}, that implements the same functionality for multiple browsers, masking most of the incompatibility issues.
 
-  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, doesn't need to be installed and shouldn't have incompatibility issues.
-
-  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3, they take into account unsupported web browsers though \emph{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.
-
-  But SmillingWeb just implements a subset of \ac{SMIL} 3.0 and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when leading with \ac{SMIL} changes due to real time communications.
-
-  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to low latency networks leading to experience pauses during playback.
+  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, doesn't require a plug-in to be installed and shouldn't have incompatibility issues.
+  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3.
+  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.
+  But SmillingWeb just implements a subset of \ac{SMIL} 3.0 and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when dealing with \ac{SMIL} changes due to real time communications.
+  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to low latency networks leading to  pauses during playback.
+%RP explicar melhor. Não vejo como low-latency pode ser mau
 
   With the emergence of \ac{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format.
-Another important tag is \emph{canvas} that allows to draw graphics with \emph{JavaScript} on a rectangle within a web page.
+Another important tag is \emph{canvas} that allows drawing graphics with \emph{JavaScript} on a rectangle within a web page.
+
+  %RP esta parte está um pouco confusa.
+Primeiro falas de smil, depois de html, depois de smil novamente.
+Devias agrupar tudo.
+Devias agrupar as soluçoes por tecnologia.
+Se preciso, primeiro fala de html, css, etc, e depois de smil e das implementação de smil (incluindo as que são implementadas em html e js).
 
   \ac{SVG} is an \ac{XML} based format that incorporates the animation module of \ac{SMIL}.
-Currently, \ac{SVG} allows to add movement and animate attributes of elements.
-When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real-time through \ac{DOM} \ac{API}, besides that, it also allows to call \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.
+Currently, \ac{SVG} allows adding movement and animating attributes of elements.
+When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real-time through the \ac{DOM} \ac{API}.
+Besides that, it also allows calling \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.
 
   Video and audio features are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind \emph{JavaScript} functions for different kinds of events over video and audio elements.
+%RP já tinhas falado de html. Junta tudo.
 
-  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}} was developed for web-based animations, introducing video support in 2002, \emph{flash} started to grow after that.
-Concurrency players, at that time, were focused on playing video and audio, \emph{flash} had vector graphics and they were focused on streaming \emph{on-demand} video across multiple platforms.
-\emph{VP6} was their choice on video codecs, optimizing for half of video size for the same quality and providing adjusted video quality based on \emph{Internet} connection latency.
+  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}} was developed for web-based animations.
+Introducing video support in 2002, \emph{flash} started to grow after that.
+Competitor's players, at that time were focused on playing video and audio, while \emph{flash} had vector graphics and focused on streaming \emph{on-demand} video across multiple platforms.
+\emph{VP6} was their choice of video codec, providing half the video size (in Byte) for the same quality and providing video quality adjusted to \emph{Internet} connection latency.
+  \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
+  %RP ``was'' quando? Em 2010? Ontem?
 
-  \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
-
-  \ac{RTMP} is a \ac{TCP} based protocol used to streaming audio, video and data between \ac{FMS} and flash player.
-A bidirectional connection is established between the two to in order allow real time communications.
-A flash player can stream a webcam video to \ac{FMS} according to \ac{RTMP} and another flash player can request a video stream to \ac{FMS} that can be pre-recorded stream, live stream or data.
+  \ac{RTMP} is a \ac{TCP} based protocol used for streaming audio, video and data between a \ac{FMS} and flash players.
+A bidirectional connection is established between the two in order to allow real time communications.
+A flash player can stream a webcam video to a \ac{FMS} using \ac{RTMP} or it can request a video stream to \ac{FMS} that can either be a pre-recorded stream, live stream or data.
 Multiple \ac{FMS} servers can be chained together in order to increase capacity and handle more streams simultaneously.
+  %RP chained ou usados em paralelo?
 
   \ac{FMS} can stream video and audio to one or more subscribers by sending a separate copy for each subscriber.
-With \ac{RTMFP} is possible to stream video between flash players, allowing a publisher breaking a stream into pieces that can be cooperatively distributed in a P2P mesh, \ac{RTMFP} uses \ac{UDP} to speed packet delivery, although is not reliable, it is well suited for video streaming, like \ac{WebRTC} flash players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
+With \ac{RTMFP} it is possible to stream video directly between flash players, allowing a publisher to break up a stream into pieces that can be cooperatively distributed in a P2P mesh.
+\ac{RTMFP} uses \ac{UDP} to speed packet delivery, which although it is not reliable, is well suited for video streaming.
+Like \ac{WebRTC}, flash players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
+  %RP as vezes fazes \emph{flash} e outras vezes não. Convém ser coerente.
 
-  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} are implementing some features of flash, it doesn't mean that flash will be replaced, instead of that both technologies can be used to develop rich \emph{Internet} applications.
-It is also important to note that \ac{HTML}5 is more compatible with mobile devices than adobe flash.
+  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} implement some of flash's features, it doesn't mean that flash will be replaced soon.
+  Instead, both technologies can be used to develop rich \emph{Internet} applications.
+  It is also important to note that \ac{HTML}5 is better supported in mobile devices than Adobe flash.
 
@@
 Compared to flash, which uses \emph{ActionScript}, \emph{Silverlight} applications can use languages like C\#, \emph{VisualBasic} and \ac{XAML}.
-\emph{Silverlight} uses a technique called \emph{Smooth Streaming} from \emph{IIS Media Service} that consists on delivering video in real-time with adjusted quality in function of bandwidth changing and \ac{CPU} usage.
+\emph{Silverlight} uses a technique called \emph{Smooth Streaming} from \emph{IIS Media Service} that consists on delivering video in real-time with adjusted quality in function of bandwidth variations and \ac{CPU} usage.
 
-  Using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \emph{JavaScript} and \ac{SVG}, will make possible to raise communications to a new level.
+  Using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \emph{JavaScript} and \ac{SVG}, will make it possible to raise communications to a new level.
+  %RP não concordo com frase.
+O que o uso de tecnologias standard faz é garantir que a app é suportada em vários dispositivos.
+Também podes dizer as essas tecnologias já permitem desenvolver a applicação que tens em mente, mas isso é outra coisa.
+  %RP evita ser tão ``vendedor''. Não digas coisas como ``to a new level'' mas cinge-te a afirmaçãos objectivas.
   For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.
@@
 With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
+  %RP porque escolheste falar aogra de webgl se não falas antes?
 
-  In this project our goal is to enrich hyper-calls with no limits, every user should be free to choose how it wants to be contacted and it wants to share its contents.
+  In this project our goal is to enrich hyper-calls with no limits, every user should be free to choose how it wants to be contacted and how he wants to share its contents.
+  %RP O ``no limits'' é muito vago. Tens de ser mais objectivo.
 
   In order to give users a personalized communication channel, each user must have a personal web page where its available plug-ins could be downloaded from other peers, after that, they can talk in the same language whatever it is.
+%RP acho que aqui já estás a falar de detalhes de implementação que o leitor não tem possibilidade de perceber antes de descreveres as funcionalidades de aplicação e a sua arquitectura.

### 2015-06-02 tese/state01.tex
possible end of SotA 1

@@
 
-\emph{Full Cone}, \emph{Restricted Cone} and \emph{Port Restricted Cone} \ac{NAT}s became the common configurations on the Internet.
+\emph{Non-Symmetric} \ac{NAT}s became the common configurations on the Internet.
 As a direct result, problems started to appear: the amount of ports that \ac{IP} makes available is also small compared to our current needs; worse than that, \ac{NAT} also difficults end-to-end communication, forcing most applications that follow this model to be implemented ineffectively.
@@
 Those kind of applications require real time communication in order to achieve the best performance.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.25\textwidth]{figures/multinat.png}
+	\caption{Multiple level \ac{NAT}s}
+\end{figure}
+
 %RP file sharing não precisa de real-time
@@
 Symmetric \ac{NAT} changes \ac{IP} port for each different connection, for that reason, when the \ac{STUN} servers reply with the \ac{IP} address and port of their connection, it will be useless for clients to use in other connection connections.
-That is why Symmetric \ac{NAT} represents a problem for real time communications.
+That is why Symmetric \ac{NAT} represents a problem for peer-to-peer communications.
 %RP os problemas que relatas têm a ver com P2P ou real-time?
@@
 Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \emph{full cone}.
-\ac{ICE} uses \ac{STUN} when it's possible and \ac{TURN} otherwise.
+\ac{ICE} is a technique that uses \ac{STUN} when direct communications are possible and \ac{TURN} while a direct communication isn't possible.
 %RP ainda não explicaste o que é o ICE
@@
 Most of client-server applications aren't affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints.
-Clearly this type of communication requires a more expensive infrastructure and, in most cases, more network usage, leading to a worse quality of service.
+Clearly \ac{TURN} requires a more expensive relaying infrastructure and, in most cases, more network usage, leading to a worse quality of service.
 The requirements of real-time video communication makes this kind of model unsuitable.
@@
 \ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser based real time communications without installing any additional application or plug-in.
+
+Some operating systems such as \emph{Android}, \emph{iOS}, \emph{Linux}, \emph{OSX} and \emph{Windows} implements native \ac{WebRTC} libraries, extending the usage of \ac{WebRTC} to applications outside the web browser.
+This native support can help to implement applications that record video and audio streams for further playback.
+
 %RP indicar que agora também já temos webrtc fora do browser devido ao seu sucesso e ao facto de reunir tecnologia state of the art?
@@
 If a packet sequence number is greater than the expected sequence number the packet may be enqueued or discarded.
-By knowing the sequence of messages that are sent and received in \ac{TLS}, timers are used for packet retransmission avoiding acknowledgment messages.
+By knowing the sequence of messages that are sent and received in \ac{DTLS}, timers are used for packet retransmission avoiding acknowledgment messages.
 %RP TLS?
@@
 \ac{DTLS} keys are negotiated on handshake in order to achieve a secure connection.
-The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining communications are done through \ac{UDP} without using \ac{DTLS}.
+The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining \ac{SRTP} communications are done through \ac{UDP} without using \ac{DTLS}.
 %RP quais remaining?  A última frase é confusa
@@
 \ac{WebRTC} aims to provide a standard platform for real-time audio and video on the Web. It arrives at a time when several proprietary products are well established.
-\emph{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength is the amount of users that are using it nowadays.
+\emph{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength are the amount of users that are using it nowadays and the ability to perform voice calls to the \ac{PSTN}.
 But compared to \emph{Skype}, \ac{WebRTC} applications don't need to be pre-installed.
@@
 In the past, in order to use \emph{hangouts} on a web browser a plug-in had to be installed, nowadays hangouts is using \ac{WebRTC}.
+\emph{Google Hangouts} supports viewing videos on \emph{youtube} synchronously, drawing collaboratively, creating music and playing multi-player games.
+These applications are implemented with \emph{Adobe Flash}.
 
 \emph{Jitsi Meet}\footnote{\url{http://jitsi.org/Projects/JitsiMeet}} is a \ac{WebRTC} collaborative application that uses \emph{Jitsi Videobridge} for high quality and scalable video conferences and supports shared document editing.
+\emph{Jitsi Meet} allows a great amount of users in the same conversation by identifying the current most active participant users and, by consequence, reducing the video and audio quality for all the other users.
 \emph{Jitsi Videobridge} is a server that enables multi-party video calls.

### 2015-06-02 tese/state04.tex
Finished reviewing the state of the art

@@
-\subsection{Extending collaboration tools with time manipulation}\label{collab}
+\subsection{Extending collaboration tools with time manipulation}
+\label{collab}
 
 Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity.
-Technologies like these are appearing along these days, but they couldn't be possible years ago because technology was limited or unavailable.
-Although today's technology is limited on some aspects, we are doing progress in order to improve the web ecosystem, by creating standards and migrating to newer technologies.
-
- Our first concern on real time collaboration applications, besides the communication itself, is the data storage and representation.
-Storing multimedia content is not a viable solution because most browsers are recommended to limit local storage to at least five megabytes per origin.
+Technologies like these are appearing along these days, but they were not be possible a few years ago because technology was limited or unavailable.
+Although today's technology is still limited on some aspects, we are doing progress in order to improve the web ecosystem, by creating standards and migrating to newer technologies.
+%RP aqui dizes ``we are doing progress''. Parece que é parte do teu trabalho e que tens algo a ver com isso. Muda para frase impessoal
 
 	If, for instance, one wants to rewind a real-time video, recordings will be needed from whom is streaming the video. 
+ Our first concern on real time collaboration applications, besides the communication itself, is the data storage and representation.
+Storing multimedia content is not a viable solution because most browsers recommend limiting local storage to at most five megabytes per origin.
 
-  \ac{RTP}\cite{rfc3550} is used for streaming audio and video over \ac{IP}, the multimedia content is transported on the payload of \ac{RTP} messages, \ac{RTP} contains headers for payload identification.
-\ac{RTP} is independent from its payload type, allowing to transport any kind of encoded multimedia.
+ \ac{RTP}\cite{rfc3550} is used for streaming audio and video over \ac{IP}.
+ Multimedia content is transported on the payload of \ac{RTP} messages, that contain headers for payload identification.
+\ac{RTP} is independent from its payload type, allowing it to transport any kind of encoded multimedia.
 A sequence number is used for sorting received packets.
@@
 
-  \ac{RTP} allows to change its requirements and add extensions to it with profiles, one of the most used is the \ac{RTP} profile for audio and video \cite{rfc3551} which lists the payload encoding and compression algorithms.
-This profile also assigns a name to each encoding which may be used other protocols like \ac{SDP}.
+  \ac{RTP} allows to change its requirements and add extensions to it with profiles.
+One of the most used ones is the \ac{RTP} profile for audio and video \cite{rfc3551}, which lists the payload encoding and compression algorithms.
+This profile also assigns a name to each encoding which may be used with other protocols like \ac{SDP}.
+  Another profile for \ac{RTP} is defined by \ac{SRTP}, which provides encryption, authentication and replay protection for \ac {RTP} traffic.
+The analogous secure protocol for \ac{RTCP} is \ac{SRTCP}.
 
-  Another profile for \ac{RTP} is defined by \ac{SRTP}, which provides encryption, authentication and replay protection of \ac {RTP} traffic. The analogous secure protocol for \ac{RTCP} is \ac{SRTCP}.
-
-  Both \ac{SRTP} and \ac{SRTCP} use \ac{AES} by default, which is symmetric-key algorithm for data encryption.
-Each packet is encrypted using a distinct key-stream, otherwise using a single key-stream with \ac{AES} on \ac{CBC} would make impossible to recover from packet loss.
+  Both \ac{SRTP} and \ac{SRTCP} use \ac{AES} by default, which is a symmetric-key algorithm for data encryption.
+Each packet is encrypted using a distinct key-stream, as otherwise, using a single key-stream with \ac{AES} on \ac{CBC} would make it impossible to recover from packet loss.
 
   Two key-stream generators for \ac{AES} were defined: \emph{Segmented Integer Counter Mode} and \emph{f8-mode}.
-If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those key-stream generators and it's fixed for each packet.
+If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those key-stream generators and it is fixed for each packet.
+  %RP isto está tudo no RFC3550? Se não, faltam muitas citações.
 
   \ac{RTP} recorders are independent of payload encoding, they don't decode \ac{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.
+
+  %RP estavas a falar de gravar video e passaste a falar de protocolos de streaming. Convém explicar o racional para esta abordagem. Começa a secção a explicar sobre o que vais falar e porquê.
+  %RP e usa subsecções...
 
@@
 	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuous, the second one describes it as interactive or non-interactive.
-A discrete media is characterized by its type not depending from time, as continuous media depending from it.
+A discrete media is characterized by not depending from time, and continuous media as depending on it.
 Interactive media is characterized by its state being changed by external events such as user interactions.
 
-	For example, an image is non-interactive and discrete, for instance, a video is continuous and non-interactive.
+	For example, an image is non-interactive and discrete, while a video is continuous and non-interactive.
 A simple collaborative editor with just text is interactive and discrete.
@@
 	Streaming protocols like \ac {RTP} and \ac{RTSP} were designed for continuous and non-interactive media types, such as audio and video.
-Discrete and non-interactive media don't need to be streamed through \ac{RTP} because they don't change with.
+Discrete and non-interactive media don't need to be streamed through \ac{RTP} because they don't change with time.
 For example, if an image appears in a specific time interval, just the \ac{HTML} or \emph{JavaScript} that will reference the image must be streamed, the image itself is then transferred through \ac{HTTP}.
@@
 	In order to play any kind of stream, a player for interactive stream is needed for downloading an environment, decoding the \ac{RTP} payload to determine the state and display it to the user.
-Streaming interactive media like the combination of \ac{HTML}, \ac{CSS} and JavaScript require more than interpreting the code, a streamed user interface may contain an internal state that is not shown on code.
+Streaming interactive media like a combination of \ac{HTML}, \ac{CSS} and JavaScript requires more than interpreting the code: a streamed user interface may contain an internal state that is not shown on code.
+        %RP não percebi
 
 	Every time an event is processed on one of the endpoints, both sender and receivers state must stay synchronized, otherwise events may behave differently.
-
-	To achieve synchronization of interactive data most packets have three types: \emph{State}, \emph{Delta-State} and \emph{Event}.
+	To achieve synchronization of interactive data, most packets have three types: \emph{State}, \emph{Delta-State} and \emph{Event}.
 State packet defines the environment complete state.
@@
 Event packets informs that an event occurred over the interactive media.
+%RP referencia?
 
@@
 Therefore, if the receiver receives all recorded packets, it can calculate the current state from the previous complete state.
-Streaming too much complete states, results on more precise random accesses, but the trade-off is the higher bandwidth usage and used storage space on the recording server.
+Streaming too many complete states results on more precise random accesses, but the trade-off is the higher bandwidth usage and used storage space on the recording server.
 On the other hand, if there are fewer complete states recorded followed by delta-states, the recorded stream will occupy less storage space, but random accesses will be less granular.
-
+        %RP aqui já falas em recording servers, que é um conceito que ainda não apresentaste.
+        %RP isto não parece um estado da arte. Não há referências e é tudo ``we''. pelo que parece ser apenas a tua opinião.
+        %RP esta secção precisa de ser revista em profuncidade.
+Os paragrafos fazem sentido mas não há continuidade entre eles.
+Parece que vais saltando entre ideias e conceitos.
+Tens de ser melhor organizado, dividido em sub-secções que devem começar com uma explicação do problema que vai ser discutido.
 	It is possible to restore the media state even if messages are lost by recording and streaming the interactive media's complete state periodically.
@@
 
-	\cite{interactive_stream} concluded that the ability to extract the objects state in order to synchronize them and the ability intercept events in order to control remote objects, are realized using the \ac{MVC} concept.
+	\cite{interactive_stream} concluded that the ability to extract the objects state in order to synchronize them and the ability to intercept events in order to control remote objects are realized using the \ac{MVC} concept.
+%RP are realized ou ``can be realized''?
         This concept separates three components within an application.
@@
 
-	Using \ac{MVC} concept will make possible to implement an interactive \ac{RTP} application that records, play, fast forward, fast rewind, stop and jump to random positions.
+	Using the \ac{MVC} concept will it make possible to implement an interactive \ac{RTP} application that records, play, fast forward, fast rewind, stop and jump to random positions.
+        %RP porquê RTP aplication? Ainda não disseste porque tem de ser RTP! Não era webrtc?
 
@@
 \cite{interactive_record} explained how to record interactive video with this new profile.
+        %RP não uses a citação como sujetio das frases. O trabalho não tem um nome?
 
+        %RP lá vais tu mudar de tema da conversa sem explicar porquê....
+        Even though one on one calls are common, there are occasions when several people take part of the same video call.
 	Multi-party video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants.
 Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
+        In this scenario, a more efficient alternative to peer-to-peer is the use of a Multipoint control unit.
 
 	\emph{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}} receives one stream from every participant on a conference, either from a \emph{jitsi} client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
-Although all the participants need to download all the streams from \emph{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
+Although all the participants still need to download all the streams from the \emph{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
+	\emph{Jitsi Video Bridge} uses \ac{XMPP} as a signaling protocol and its \emph{colibri} extension \cite{xep0340} to reserve channels for video transmission.
+Despite this choice for signaling protocol, \emph{Jitsi Video Bridge} also supports \ac{SIP}.
 
-	\emph{Jitsi Video Bridge} uses \ac{XMPP} as a signaling protocol and its \emph{colibri} extension \cite{xep0340} to reserve channels for video transmission.
-Although this choice for signaling protocol, \emph{Jitsi Video Bridge} also supports \ac{SIP}.
-
+	\emph{Google Wave} was a distributed collaboration platform based on \emph{Jupiter}\cite{jupiter} that adopted \ac{OT} techniques.
         	\ac{OT} technology was originally developed for consistency maintenance and concurrency control over distributed objects, \ac{OT} algorithms are mainly used in collaborative applications such as distributed document edition.
-
-	\emph{Google Wave} was a distributed collaboration platform based on \emph{Jupiter}\cite{jupiter} that adopted \ac{OT} techniques, other Google products such as Google Docs are using this type of technology.
+                Other Google products, such as Google Docs, are also using this type of technology.
 In 2010 Google stopped the development of Google Wave and released the main components as Open Source code to Apache, the project is currently known as \emph{Apache Wave} and the reference implementation is named as \emph{Wave in a Box}.
+                %RP acho que o google wave merecia uma descrição mais detalhada. Mas não era aqui. Vê o comentário no final
 
@@
 	\emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}} is a collaborative framework with its own server implementation, it supports four types of collaborative elements: String, Lists, Maps and Custom objects.
+        %RP estes 3 últimos parágrafos estão desligados do resto. O que pretendes apresentar?
+
+        %RP falta falar um pouco das ferramentas de colaboração já existentes? Wave, Hangouts, skype, etc? Acho que isto merecia uma secção à parte, mesmo que mais curta.

### 2015-06-02 tese/state02.tex
possbile end of SotA 2

@@
 %RP acrescentei o webrtc pois isto não é um requisito geral de signaling. era o que querias dizer?
+%HR sim é isso
   \ac{HTTP} uses a request-response paradigm, where a request is sent by the client, followed by a server response.
@@
 %RP de certeza que queres dizer que o http é unidirecional? 
-  Sometimes it is required that some information be obtained in real time, but we saw, some \ac{NAT}'s don't support callbacks from servers, preventing them from notifying clients as soon as an event occures.
+%HR era mais o facto de o servidor não poder comunicar com o cliente sem que o cliente comunique primeiro (tem que haver uma iniciativa por parte do cliente)
+  Sometimes it is required that some information be obtained in real time, but we saw, some \ac{NAT}'s do not support callbacks from servers, preventing them from notifying clients as soon as an event occures.
 One technique to overcome this problem is polling.
@@
 Polling consists on sending periodic messages to which the server responds immediately with empty content or fresh information.
-Real time communications are unpredictable, if the time between periodic requests is short, most of the time the server will return empty results wasting network bandwidth and energy.
+Text and presence messages are unpredictable, if the time between periodic requests is short, most of the time the server will return empty results wasting network bandwidth and energy.
 On the other hand, if the time between periodic requests is large, newer messages may arrive too late.
@@
 The exchange of roles of those two connections allow to pull data from multiple contexts instead of being locked in just one.
-  If the client has data to send while a request is still open, it establishes a second socket connection to the connection manager to send a new request.
-The connection manager immediately responds to the previously held request (possibly with no data) and holds open this new request.
-This results in the connections switching roles.
   %RP estas duas frases dizem o mesmo da perspectiva do cliente e do servidor. Não consegues simplificar numa só?
@@
   \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with the \ac{PSTN}.
-  Service providers making their \ac{SIP} infrastructures available through WebSockets.
+  Service providers are making their \ac{SIP} infrastructures available through WebSockets.
 %RP a frase anterior é estranha
@@
 Another advantage of \ac{XMPP} is the addition of extensions, for example \cite{xep0096}, which adds file transfer capabilities between two entities and \cite{xep0045} which enables multi-user chat.
-  \ac{XMPP}'s bi-directional communication is achieved through \ac{BOSH} \cite{xep0206}.
+  \ac{XMPP}'s bi-directional communication over \ac{HTTP} is achieved through \ac{BOSH} \cite{xep0206}.
   %RP falta um `` bi-directional communication OVER HTTP is achieved through''?
@@
 %RP os outros não permitem multi-party?
+%HR adicionei ao Hangouts e Skype, quanto ao XMPP já digo que existe uma extensão que permite multi-user chat

### 2015-06-02 tese/arch01.tex
Reviewed architecture.

@@
 \begin{itemize}
- \item Enrich calls with any kink of multimedia in real-time.
- \item Ability perform tasks on real-time collaborative environment.
+\item Enrich calls with any kind of multimedia in real-time.
+  %RP any kind não é demasiado ambicioso?
+ \item Ability to perform tasks on real-time collaborative environment.
  \item Ability to record and playback interactive multimedia.
@@
 
+%RP ainda não detalhaste as funcionalidades que queres incluir na aplicação. Até aqui tem sido tudo muito vago. convém agora datalhares as funcioalidades que vais implementar.
 \subsection{Modules}
-	Our project is divided into five modules.
+	Our system is divided into five modules.
 
@@
 	\includegraphics[width=0.55\textwidth]{figures/archs.png}
-	\caption{Project Modules}
+	\caption{System Modules}
+        \label{fig:modules}
 \end{figure}
+%RP as figuras têm de ser apresentadas. Figure~\ref{fig:modules} presents ...
 
@@
 The web server will provide libraries to the client in order to interact with the other modules.
+ %RP não queres falar da comunicação com outros peers?
+ %RP há conversas p2p ou todos os clientes ligam ao stream server? Acho que podes dar mais detalhes nesta fase
+
+ %RP antes de passar para a implementação deves falar em maior detalhe destes componentes. Como interagem entre si. Qual a informação que cada um guarda troca com os outros.
+ %RP quais os requisitos de tecnologia para cada componente.
+ %RP precisas de fornecer muitos mais detalhes. A arquitectura deverá ter várias páginas. A tua ainda só tem meia e passas logo para a implementação do protótipo.
 
@@
 Although the core infrastructure is important and crucial for the implementation of our web application, there is no preference for a specific software running on the web server.
-An important requirement for the Web Server is the WebSockets support, \emph{LAMP} server was considered, although the chosen technology for web development is \emph{Play Framework} with \emph{Java}, which has a more evident separation between \emph{Model}, \emph{View} and \emph{Controller} components.
+An important requirement for the Web Server is the WebSockets support, The \emph{LAMP} stack was considered, although the chosen technology for web development is the \emph{Play Framework} with \emph{Java}, which has a more evident separation between \emph{Model}, \emph{View} and \emph{Controller} components.
 
 For the signaling server there is also no preference between \ac{SIP} and \ac{XMPP}, as both support presence information feature.
-\ac{SigOfly} could be adopted as well but using it would be irrelevant for the final result of this project.
-For sake of simplicity and easy deployment the chosen platform chosen is the \emph{Ejabberd} \ac{XMPP} application server since it implements \cite{xep0206} which is crucial for web applications, \emph{Ejabberd}, as said before, is also the \ac{XMPP} server that implements more extensions.
-Initially, for simplicity, \emph{Ejabberd} will be configured as a single node instead a member of cluster.
+\ac{SigOfly} could be adopted as well, but using it would be irrelevant for the final result of this project.
+For the sake of simplicity and easy deployment, the chosen platform is the \emph{Ejabberd} \ac{XMPP} application server, since it implements \cite{xep0206} which is crucial for web applications.
+\emph{Ejabberd}, as said before, is also the \ac{XMPP} server that implements more extensions.
+Initially, for simplicity, \emph{Ejabberd} will be configured as a single node instead a member of a cluster.
 
-The streaming server will be \emph{Jitsi Videobridge} as \emph{Jitsi} team is working with \ac{WebRTC} and their code is open source.
+The streaming server will be the \emph{Jitsi Videobridge} as the \emph{Jitsi} team is working with \ac{WebRTC} and their code is open source.
 
-The \ac{ICE} server is not required to be on our infrastructure as a public \ac{ICE} server could be used, if \ac{TURN} is used the network speed could drop due to resource share amongst other users.
+The \ac{ICE} server is not required to be on our infrastructure, as a public \ac{ICE} server could be used.
+If \ac{TURN} is used, the network speed could drop due to resource sharing with other users.
 An \ac{ICE} server will be installed in order to prevent the influence of other users on our application.
@@
 \end{figure}
+%RP apresentar figura
 
 \emph{Modernizr} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.
+\emph{Bootstrap} will be used to make the user interface more appellative and responsive.
+With \emph{Bootstrap} it is quite easy to develop applications that adapt to mobile devices with different screen sizes.
+The \emph{Strophe} library will be essential to communicate with the \ac{XMPP} server and other clients.
+%RP era este o sentido da frase? Ou entre cliente e servidor apenas?
+\emph{TogetherJS} will be used for the collaborative component of our web application.
+Other libraries were considered but, as said before, \emph{TogetherJS} does not implement object storage making the object storage implementation free.
+%RP relê a frase. É mesmo isto que queres dizer? ``does not implement''
 
-\emph{Bootstrap} will be used to make user interface more appellative and responsive.
-With \emph{bootstrap} it's quite easy to develop applications that adapt to mobile devices with different screen sizes.
-
-\emph{Strophe} library will be essential to communicate with \ac{XMPP} server and clients.
-
-\emph{TogetherJS} will be used for the collaborative component of our web application, other libraries were considered but, as said before, \emph{TogetherJS} does not implement object storage making the object storage implementation free.
+%RP já não vais usar smil? Como fazes a sincronização. Explica, sff.

### 2015-06-02 tese/state01.tex

@@
 \ac{WebRTC} aims to provide a standard platform for real-time audio and video on the Web. It arrives at a time when several proprietary products are well established.
-\emph{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength are the amount of users that are using it nowadays and the ability to perform voice calls to the \ac{PSTN}.
+\emph{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice, instant messaging and multi-party communication over proprietary protocols, its main strength are the amount of users that are using it nowadays and the ability to perform voice calls to the \ac{PSTN}.
 But compared to \emph{Skype}, \ac{WebRTC} applications don't need to be pre-installed.
 
-\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is another popular video conference web application. 
+\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is another popular video multi-party conference web application. 
 In the past, in order to use \emph{hangouts} on a web browser a plug-in had to be installed, nowadays hangouts is using \ac{WebRTC}.

### 2015-06-02 tese/state03.tex
middle o SotA 3, starting to feel tired ...

@@
 Since the early days of video technology, one of the problems raised consisted on how to add more information onto video without generating multiple versions.
+This section examines technologies that allows different ways to present multimedia content in such a way that it change based on synchronization amongst other multimedia elements or user interaction.
+
+\emph{Hypertext} is a type of text that provides links to texts or other types of content, these links are known by \emph{hyperlinks}.
+\emph{Hypermedia} is an evolution of \emph{hypertext}, it includes audio, images, text and video.
+
+\emph{Hypermedia} concept brings the possibility to organize and overlay multimedia elements into a nonlinear linear structure.
+
+In the beginning of analog video technology, the navigation over it was quite limited to simple operations such as play, stop, rewind and fast-forward.
+As video started to be digitalized, new operations over video emerged, such as random jumps and chapter navigation through interactive menus present on \ac{DVD}.
+
 %RP depois disto falta uma introdução a explicar que precisas de controlar a apresentação e execução de diversos conteúdos e que para isso vais precisar de uma tecnologia que permite sincronizar e controlar.
@@
 
-  In order to navigate through a dynamic video, one must be aware of time synchronization and the multiple time flows, it's important that all time, causality and behavior rules are well defined.
+  In order to navigate through a dynamic video, we must be aware of time synchronization and the multiple time flows, it ils important that all time, causality and behavior rules are well defined.
   %RP one must be aware? O utilizador? Ou queres dizer que é preciso definir as relações? O que falas a seguir é sobre linguagens para isso.
@@
 HyVAL defines defines video structure, internal video and external media objects.
+  \emph{HyVAL} uses a primary video stream, around which all other elements are organized and synchronized.
   \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically. 
@@
   \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films.
-With \ac{SMIL} it is possible to synchronize multiple sections of video, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
+With \ac{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
   %RP synchronize multiple sections of video? vários videos apresentados em simultâneo?
+
   \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Metainformation}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.
@@
 The \ac{DOM} is a standard \ac{API} that allows easy management of documents that are organized in a tree structure, by providing \ac{CRUD} operations over its elements and their attributes.
-\ac{DOM} makes it easy to inter-operate between imperative and declarative programming languages.
+\ac{DOM} makes it easy to inter-operate between imperative and declarative programming languages\cite{dom}.
 %RP não arranjas uma referência para o dom?
@@
 
-  \ac{SMIL}'s modules are used to synchronize and animate \ac{XHTML} and \ac{SVG} elements.
+  \ac{SMIL}'s modules can be used to synchronize and animate \ac{XHTML} and \ac{SVG} elements.
   %RP ``are used'' ou ``can be used''?
@@
 Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \ac{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plug-in.
-
+% XXX
   %RP fazes uma mudanças de tema muito bruscas. Podes usar \subsubsection ou \paragraph para dividir isto melhor. E uma frase entre temas a servir de ``cola'' ou ``separação'' também não fica mal

### 2015-06-02 tese/state03.tex
almost end of SotA 3, starting to feel really tired ...

@@
 
-  \item The\textbf{Linking} module contains elements and attributes for navigational hyperlinking. Navigation can be triggered by events or user interaction.
+  \item The\textbf{Linking} module contains elements and attributes for navigational \emph{hyperlinking}. Navigation can be triggered by events or user interaction.
 
@@
 Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \ac{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plug-in.
+
+  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, doesn't require a plug-in to be installed and shouldn't have incompatibility issues.
+  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3.
+  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.
+  But SmillingWeb just implements a subset of \ac{SMIL} 3.0 and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when dealing with \ac{SMIL} changes due to real time communications.
+  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to high latency networks leading to  pauses during playback.
+
 % XXX
   %RP fazes uma mudanças de tema muito bruscas. Podes usar \subsubsection ou \paragraph para dividir isto melhor. E uma frase entre temas a servir de ``cola'' ou ``separação'' também não fica mal
-  \ac{HTML} is a markup language based on \ac{XML} that is used for creating web pages.
+
+  An alternative to \ac{SMIL}'s \emph{Layout} module is to use \ac{HTML} which is a markup language based on \ac{XML} that is used for creating web pages.
 \ac{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages.
-Languages like \ac{CSS} and \emph{JavaScript} are typically combined with \ac{HTML} for improving the interaction and appearance of a web page.
+Languages like \ac{CSS} and \emph{JavaScript} are typically used along with \ac{HTML} for improving the interaction and appearance of a web page.
 
@@
 
-  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, doesn't require a plug-in to be installed and shouldn't have incompatibility issues.
-  SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3.
-  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.
-  But SmillingWeb just implements a subset of \ac{SMIL} 3.0 and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when dealing with \ac{SMIL} changes due to real time communications.
-  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to low latency networks leading to  pauses during playback.
 %RP explicar melhor. Não vejo como low-latency pode ser mau
-
+  %HR troquei-me todo
   With the emergence of \ac{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format.
@@
 
-  Video and audio features are already possible with \ac{HTML}5, like \ac{SVG} it is also possible to bind \emph{JavaScript} functions for different kinds of events over video and audio elements.
-%RP já tinhas falado de html. Junta tudo.
-
   Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}} was developed for web-based animations.
@@
 \emph{VP6} was their choice of video codec, providing half the video size (in Byte) for the same quality and providing video quality adjusted to \emph{Internet} connection latency.
-  \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
+  In 2010, \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
   %RP ``was'' quando? Em 2010? Ontem?
+  %HR exacto.
 
-  \ac{RTMP} is a \ac{TCP} based protocol used for streaming audio, video and data between a \ac{FMS} and flash players.
+  \ac{RTMP} is a \ac{TCP} based protocol used for streaming audio, video and data between a \ac{FMS} and \emph{flash} players.
 A bidirectional connection is established between the two in order to allow real time communications.
-A flash player can stream a webcam video to a \ac{FMS} using \ac{RTMP} or it can request a video stream to \ac{FMS} that can either be a pre-recorded stream, live stream or data.
-Multiple \ac{FMS} servers can be chained together in order to increase capacity and handle more streams simultaneously.
+A \emph{flash} player can stream a webcam video to a \ac{FMS} using \ac{RTMP} or it can request a video stream to \ac{FMS} that can either be a pre-recorded stream, live stream or data.
+Multiple \ac{FMS} servers can be used in parallel to increase capacity and handle more streams simultaneously.
   %RP chained ou usados em paralelo?
@@
   \ac{FMS} can stream video and audio to one or more subscribers by sending a separate copy for each subscriber.
-With \ac{RTMFP} it is possible to stream video directly between flash players, allowing a publisher to break up a stream into pieces that can be cooperatively distributed in a P2P mesh.
+With \ac{RTMFP} it is possible to stream video directly between \emph{flash} players, allowing a publisher to break up a stream into pieces that can be cooperatively distributed in a P2P mesh.
 \ac{RTMFP} uses \ac{UDP} to speed packet delivery, which although it is not reliable, is well suited for video streaming.
-Like \ac{WebRTC}, flash players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
+Like \ac{WebRTC}, \emph{flash} players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
   %RP as vezes fazes \emph{flash} e outras vezes não. Convém ser coerente.
 
-  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} implement some of flash's features, it doesn't mean that flash will be replaced soon.
+  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} implement some of \emph{flash}'s features, it doesn't mean that \emph{flash} will be replaced soon.
   Instead, both technologies can be used to develop rich \emph{Internet} applications.
-  It is also important to note that \ac{HTML}5 is better supported in mobile devices than Adobe flash.
+  It is also important to note that \ac{HTML}5 is better supported in mobile devices than \emph{Adobe Flash}.
 
@@
 It supports vector graphics, animation and video.
-Compared to flash, which uses \emph{ActionScript}, \emph{Silverlight} applications can use languages like C\#, \emph{VisualBasic} and \ac{XAML}.
+Compared to \emph{flash}, which uses \emph{ActionScript}, \emph{Silverlight} applications can use languages like C\#, \emph{VisualBasic} and \ac{XAML}.
 \emph{Silverlight} uses a technique called \emph{Smooth Streaming} from \emph{IIS Media Service} that consists on delivering video in real-time with adjusted quality in function of bandwidth variations and \ac{CPU} usage.
 
-  Using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \emph{JavaScript} and \ac{SVG}, will make it possible to raise communications to a new level.
+  Using technologies that relies only on web standards, like \ac{CSS}, \ac{HTML}5, \emph{JavaScript} and \ac{SVG}, will make possible to develop an application that solves our problem with the advantage of being compatible with a greater amount of web browsers.
   %RP não concordo com frase.
@@
 With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
+ %XXX
   %RP porque escolheste falar aogra de webgl se não falas antes?

### 2015-06-02 tese/state03.tex
end of SotA 3, gonna sleep

@@
 
+  %RP esta parte está um pouco confusa.
+Primeiro falas de smil, depois de html, depois de smil novamente.
+Devias agrupar tudo.
+Devias agrupar as soluçoes por tecnologia.
+Se preciso, primeiro fala de html, css, etc, e depois de smil e das implementação de smil (incluindo as que são implementadas em html e js).
+
   \ac{CSS}'s goal is to separate the structure of an \ac{XML} document from its appearance. \ac{CSS} defines styles for \ac{XML} tags based on their name, class, identifier or position.
@@
 
-  %RP esta parte está um pouco confusa.
-Primeiro falas de smil, depois de html, depois de smil novamente.
-Devias agrupar tudo.
-Devias agrupar as soluçoes por tecnologia.
-Se preciso, primeiro fala de html, css, etc, e depois de smil e das implementação de smil (incluindo as que são implementadas em html e js).
+  For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.
+Another example would be a collaborative spreadsheet using \ac{WebRTC}.
+With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
+ %XXX
+  %RP porque escolheste falar aogra de webgl se não falas antes?
+%HR já troquei, o WebGL usa o canvas do HTML5, talvez seja este o local mais adequado
 
@@
   %RP evita ser tão ``vendedor''. Não digas coisas como ``to a new level'' mas cinge-te a afirmaçãos objectivas.
-  For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.
-Another example would be a collaborative spreadsheet using \ac{WebRTC}.
-With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
- %XXX
-  %RP porque escolheste falar aogra de webgl se não falas antes?
 
-  In this project our goal is to enrich hyper-calls with no limits, every user should be free to choose how it wants to be contacted and how he wants to share its contents.
+  %In this project our goal is to enrich hyper-calls with no limits, every user should be free to choose how it wants to be contacted and how he wants to share its contents.
   %RP O ``no limits'' é muito vago. Tens de ser mais objectivo.
+  %HR tanto esta como a seguinte frase foram escritas praticamente no inicio do estado de arte, nem faz sentido estarem aqui.
 
-  In order to give users a personalized communication channel, each user must have a personal web page where its available plug-ins could be downloaded from other peers, after that, they can talk in the same language whatever it is.
+  %In order to give users a personalized communication channel, each user must have a personal web page where its available plug-ins could be downloaded from other peers, after that, they can talk in the same language whatever it is.
 %RP acho que aqui já estás a falar de detalhes de implementação que o leitor não tem possibilidade de perceber antes de descreveres as funcionalidades de aplicação e a sua arquitectura.

### 2015-06-02 tese/arch01.tex
added abstract

@@
 \begin{itemize}
-\item Enrich calls with any kind of multimedia in real-time.
-  %RP any kind não é demasiado ambicioso?
+ \item Text, Audio and Video communication using \ac{WebRTC}.
+ \item Enrich communications by overlaying multiple media types.
  \item Ability to perform tasks on real-time collaborative environment.
  \item Ability to record and playback interactive multimedia.
+ \item File sharing and collaborative edition.
 \end{itemize}
@@
 \subsection{Modules}
-	Our system is divided into five modules.
 
@@
 %RP as figuras têm de ser apresentadas. Figure~\ref{fig:modules} presents ...
+Figure~\ref{fig:modules} presents the structure of our system which is divided into five modules.
 
  The Web Server sends Web Pages, containing the other modules information, to the client, after this, the client contacts all the other modules.
-The ICE server will be used for \ac{NAT} traversal.
+The \ac{ICE} server will be used for \ac{NAT} traversal.
 The Stream Server will be used to record streams for further playback.
-The signaling server will be responsible for user and calls management.
+The signaling server will be responsible for user and calls management and presence information.
 The web server will provide libraries to the client in order to interact with the other modules.
+
+Clients can discover chat rooms and other clients by querying the signaling server, they can create rooms for multi-party communication which is achieved by using \ac{WebRTC}'s \emph{MediaStream} and \emph{DataChannel}, namely for audio, video and text communication.
+
+Chat rooms can be public or private.
+Public chat rooms are moderated by a group of clients which initially is formed by the room creator, this type of room has no access restrictions by default, but that can be changed by its moderators.
+Private chat rooms will be visible for a list of clients or can be accessed by clients that have a link for that chat room.
+
  %RP não queres falar da comunicação com outros peers?
@@
 With \emph{Bootstrap} it is quite easy to develop applications that adapt to mobile devices with different screen sizes.
-The \emph{Strophe} library will be essential to communicate with the \ac{XMPP} server and other clients.
+The \emph{Strophe} library will be essential to communicate with the \ac{XMPP} server.
 %RP era este o sentido da frase? Ou entre cliente e servidor apenas?
+
 \emph{TogetherJS} will be used for the collaborative component of our web application.
-Other libraries were considered but, as said before, \emph{TogetherJS} does not implement object storage making the object storage implementation free.
+Other libraries were considered but, as said before, \emph{TogetherJS} does not provide object storage giving us the ability to choose how to store collaborative objects.
 %RP relê a frase. É mesmo isto que queres dizer? ``does not implement''
 
+The synchronization between multimedia elements will be performed trough \emph{JavasSript} events.
 %RP já não vais usar smil? Como fazes a sincronização. Explica, sff.

### 2015-06-02 tese/project.tex
middle o SotA 3, starting to feel tired ...
added abstract

@@
 \newacro{DTLS}{Datagram Transport Layer Security}
+\newacro{DVD}{Digital Versatile Discs}
 \newacro{FMS}{Flash Media Server}
@@
 \author{%
-	Henrique Lopes Rocha \\
-	email1: henrique.rocha@ist.utl.pt \\
+	Henrique Duarte Lopes Rocha \\
+	email1: henrique.rocha@tecnico.ulisboa.pt \\
+  email2: hdlopesrocha91@gmail.com \\
+  student number: 68621
 }
-\institute{Instituto Superior Técnico}
+\institute{Master Degree in Telecommunications and Informatics Engineering \\ Instituto Superior Técnico \\ Universidade de Lisboa}
 %RP acrescenta número de aluno, nome do mestrado e ``cadeira de projecto''
@@
 \begin{abstract}
+Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, for communication and collaboration services.
+This paradigm allows to synchronize and structure the communications content, providing a way to integrate social media content and collaborative tools into voice and video calls.
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
-Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
-Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur.
-Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.
+Voice and image together can express emotions and expose creativity like no other medium can. With hypermedia concepts we can add more value to voice and video communications.
 
-Sed ut perspiciatis unde omnis iste natus error sit voluptatem accusantium doloremque laudantium, totam rem aperiam, eaque ipsa quae ab illo inventore veritatis et quasi architecto beatae vitae dicta sunt explicabo.
-Nemo enim ipsam voluptatem quia voluptas sit aspernatur aut odit aut fugit, sed quia consequuntur magni dolores eos qui ratione voluptatem sequi nesciunt.
+WebRTC technology, allows real-time communications between web browsers without being forced to install and use additional applications or plug-ins.
+The nature of web browser applications already follow the hypermedia concept, which makes WebRTC the ideal technology to apply the hyper-linked communications concepts.
+
+The native support of WebRTC in operating systems extends its usage to outside the web browser, allowing to explore functionalities that browsers are poor to support such as video recording and massive information storage.
+
+Our goal in this project is to develop an application that applies Hyper-linked communications concept using WebRTC.
 

### 2015-06-02 tese/project.tex
commit befor dinner

@@
 	Henrique Duarte Lopes Rocha \\
-	email1: henrique.rocha@tecnico.ulisboa.pt \\
-  email2: hdlopesrocha91@gmail.com \\
+	email: henrique.rocha@tecnico.ulisboa.pt \\
   student number: 68621
@@
 
-\subsection{Summary} % English
-Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as contribuições definidas e a forma como estas se materializaram
+%\subsection{Summary} % English
+%Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as contribuições definidas e a forma como estas se materializaram
 
-\subsection{Conclusions} % English
-Nesta secção devem ser retiradas conclusões do trabalho realizado, em face dos resultados obtidos \cite{exemplo}.
+%\subsection{Conclusions} % English
+%Nesta secção devem ser retiradas conclusões do trabalho realizado, em face dos resultados obtidos \cite{exemplo}.
 

### 2015-06-02 tese/arch01.tex

@@
  \item Enrich communications by overlaying multiple media types.
+ \item Create annotations over video and voice content.
+ \item Search and navigate through annotations.
+ \item Structure presentation of the content.
  \item Ability to perform tasks on real-time collaborative environment.
@@
 	\centering
-	\includegraphics[width=0.55\textwidth]{figures/archs.png}
+	\includegraphics[width=0.6\textwidth]{figures/archs.png}
 	\caption{System Modules}
@@
 
+Each room will be associated to multiple Media-Types, interactive and non-interactive, discrete and continuous.
+This media types can be structured with chains of events.
+Some components of the user interface may not be synchronized to all participants, for example an help or suggestion window.
+
  %RP não queres falar da comunicação com outros peers?
@@
 	\centering
-	\includegraphics[width=0.55\textwidth]{figures/arch.png}
-	\caption{System Architecture}
+	\includegraphics[width=0.6\textwidth]{figures/arch.png}
+	\caption{System Infrastructure}
 \end{figure}
@@
 	\caption{App Architecture}
+	        \label{fig:apparch}
+
 \end{figure}
 %RP apresentar figura
-
+Figure~\ref{fig:apparch} presents the application architecture and the underlying technologies.
 \emph{Modernizr} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.
@@
 Other libraries were considered but, as said before, \emph{TogetherJS} does not provide object storage giving us the ability to choose how to store collaborative objects.
+We assume that \emph{JavaScript} objects are enough for implementing our application but \emph{TogetherJS} provides the flexibility to change our storage model.
 %RP relê a frase. É mesmo isto que queres dizer? ``does not implement''
 
-The synchronization between multimedia elements will be performed trough \emph{JavasSript} events.
+The synchronization between multimedia elements will be performed trough chains of \emph{JavasSript} events, other animations can be implemented with \ac{SVG} embedded on \ac{HTML}.
 %RP já não vais usar smil? Como fazes a sincronização. Explica, sff.

### 2015-06-02 tese/state04.tex

@@
 Technologies like these are appearing along these days, but they were not be possible a few years ago because technology was limited or unavailable.
-Although today's technology is still limited on some aspects, we are doing progress in order to improve the web ecosystem, by creating standards and migrating to newer technologies.
+Although today's technology is still limited on some aspects, progress is being done in order to improve the web ecosystem, by creating standards and migrating to newer technologies.
 %RP aqui dizes ``we are doing progress''. Parece que é parte do teu trabalho e que tens algo a ver com isso. Muda para frase impessoal

### 2015-06-02 tese/project.tex
Added a few sentences to the abstract.

@@
 \begin{abstract}
-Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, for communication and collaboration services.
-This paradigm allows to synchronize and structure the communications content, providing a way to integrate social media content and collaborative tools into voice and video calls.
+The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to communication and collaboration services.
+This paradigm allows to synchronize, structure and navigate communication content, providing a way to integrate social media content and collaborative tools into voice and video calls.
+Voice and image together can express emotions and expose creativity like no other medium can. With hypermedia concepts, we can add more value to voice and video communications.
 
-Voice and image together can express emotions and expose creativity like no other medium can. With hypermedia concepts we can add more value to voice and video communications.
-
-WebRTC technology, allows real-time communications between web browsers without being forced to install and use additional applications or plug-ins.
+WebRTC technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
 The nature of web browser applications already follow the hypermedia concept, which makes WebRTC the ideal technology to apply the hyper-linked communications concepts.
-
 The native support of WebRTC in operating systems extends its usage to outside the web browser, allowing to explore functionalities that browsers are poor to support such as video recording and massive information storage.
 
-Our goal in this project is to develop an application that applies Hyper-linked communications concept using WebRTC.
+Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable colaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed latter.
+This application will target the web browser platform, resorting to WebRTC.
+
+In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose an architecture to implemente an hyper-linked communication application based on WebRTC and provide a plan for its implementation and evaluation.
 

### 2015-06-02 tese/project.tex
started conclusions

--- 2015-06-02 tese/project.tex
@@
 \section{Conclusions}\label{concl} % English
+  The development of WebRTC by the World Wide Web Consortium (W3C) allows web browser-to-browser audio, video and data communication.
+WebRTC is enabling new usage scenarios for communication and collaboration applications.
+However, the nature of such scenarios lead to significant challenges, such as the need to record interactive audio and video streams.
+
+  This project proposes a solution that uses WebRTC for creating an application that provides audio, data, text and data real-time communications.
+Theses communications will be enriched using hypermedia concepts, the possibility to record and playback interactive content and the integration of collaborative tools.
 

### 2015-06-02 tese/method.tex
[ADD] modules, changed titles

@@
-\subsection{Scientific Methodology} % Section
+%RP falta introdução à secção a explicar o que vai ser discutido aqui.
+
+%RP texto que está nos 3 parágrafos seugintes ficava bem no planned schedule. Talvez esse fique melhor antes do evaluation methodology.
 
@@
 
-Briefly, our web application will be tested with real users and unit tests, the system in general will be validated with benchmarks such as analyzing the its behavior and delays by growing the amount of users and data.
+\subsection{Evaluation Methodology} % Section
+
+Our web application will be tested with real users, performance benchmarks and unit tests.
+
+%RP elaborar o tipo de testes que vais fazer com os utilizadores: pedir para fazer tarefas e cronometrar o tempo que demoram.
+preencher questionários, decobrirem sozinhos como fazer algo, fazer após explicações, etc.
+
+%RP detalhar os testes de desempenho de que falas.
+The system in general will be validated with benchmarks such as analyzing its behavior and delays by growing the amount of users and data.

### 2015-06-02 tese/project.tex
SotA 4 restrucure

@@
 
+\setcounter{secnumdepth}{3}
 \newacro{AES}{Advanced Encryption Standard}
@@
   This project proposes a solution that uses WebRTC for creating an application that provides audio, data, text and data real-time communications.
-Theses communications will be enriched using hypermedia concepts, the possibility to record and playback interactive content and the integration of collaborative tools.
+Theses communications will be enriched using hypermedia concepts, the possibility to record and playback its content and the integration of collaborative tools.
 

### 2015-06-02 tese/state01.tex
commit befor dinner
SotA 4 restrucure

--- 2015-06-02 tese/state01.tex
@@
 	\centering
-	\includegraphics[width=0.25\textwidth]{figures/multinat.png}
+	\includegraphics[width=0.30\textwidth]{figures/multinat.png}
 	\caption{Multiple level \ac{NAT}s}
@@
 \ac{WebRTC} aims to provide a standard platform for real-time audio and video on the Web. It arrives at a time when several proprietary products are well established.
-\emph{Skype}\footnote{\url{http://www.skype.com/}} is an application that allows video, voice and instant messaging communication over proprietary protocols, its main strength are the amount of users that are using it nowadays and the ability to perform voice calls to the \ac{PSTN}.
+\emph{Skype}\footnote{\url{http://www.skype.com/}(accessed June 1, 2015).} is an application that allows video, voice, instant messaging and multi-party communication over proprietary protocols, its main strength are the amount of users that are using it nowadays and the ability to perform voice calls to the \ac{PSTN}.
 But compared to \emph{Skype}, \ac{WebRTC} applications don't need to be pre-installed.
 
-\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}} is another popular video conference web application. 
+\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}(accessed June 1, 2015).} is another popular video multi-party conference web application. 
 In the past, in order to use \emph{hangouts} on a web browser a plug-in had to be installed, nowadays hangouts is using \ac{WebRTC}.
@@
 
-\emph{Jitsi Meet}\footnote{\url{http://jitsi.org/Projects/JitsiMeet}} is a \ac{WebRTC} collaborative application that uses \emph{Jitsi Videobridge} for high quality and scalable video conferences and supports shared document editing.
+\emph{Jitsi Meet}\footnote{\url{http://jitsi.org/Projects/JitsiMeet}(accessed June 1, 2015).} is a \ac{WebRTC} collaborative application that uses \emph{Jitsi Videobridge} for high quality and scalable video conferences and supports shared document editing.
 \emph{Jitsi Meet} allows a great amount of users in the same conversation by identifying the current most active participant users and, by consequence, reducing the video and audio quality for all the other users.

### 2015-06-02 tese/state02.tex
SotA 4 restrucure

@@
 WebSocket handshake phase specifies a \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, but the remainder messages are exchanged without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
-WebSockets may not be available on every web browser, frameworks like \emph{socket.io}\footnote{\url{http://socket.io/}} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}} fall back to using \ac{HTTP} when there is no support for WebSockets.
+WebSockets may not be available on every web browser, frameworks like \emph{socket.io}\footnote{\url{http://socket.io/}(accessed June 1, 2015).} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}(accessed June 1, 2015).} fall back to using \ac{HTTP} when there is no support for WebSockets.
 
@@
 
-  The working group \ac{SIMPLE}\footnote{\url{https://datatracker.ietf.org/wg/simple/documents/}} proposed the creation of \ac{SIP} extensions, namely presence information \cite{rfc5263} and instant messaging \cite{rfc3428}.
+  The working group \ac{SIMPLE}\footnote{\url{https://datatracker.ietf.org/wg/simple/documents/}(accessed June 1, 2015).} proposed the creation of \ac{SIP} extensions, namely presence information \cite{rfc5263} and instant messaging \cite{rfc3428}.
 
@@
 %RP a frase anterior é estranha
-  Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}} and \emph{sipML5}\footnote{\url{http://sipml5.org/}} are used on the client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
+  Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}(accessed June 1, 2015).}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}(accessed June 1, 2015).} and \emph{sipML5}\footnote{\url{http://sipml5.org/}} are used on the client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
 %RP quando usas footnotes para indicar sites deves acrescentar ``last accessed (data)''. Isto é importante pois os sites são documentos efémeros que podem dizer uma coisa diferente mais tarde.
@@
 
-  \ac{XMPP} was initially developed for instant messaging and presence (Jabber\footnote{\url{http://jabber.org/}}).
+  \ac{XMPP} was initially developed for instant messaging and presence (Jabber\footnote{\url{http://jabber.org/}(accessed June 1, 2015).}).
 It is nowadays an open technology for standardized, decentralized, secure and extensible real-time communications.
@@
   This kind of communication is also possible through WebSockets \cite{rfc7395}.
-  Today, multiple XMPP server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}} and \emph{Prosody}\footnote{\url{http://prosody.im/}}.
-\emph{Ejabberd} is the server that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}}.
+  Today, multiple XMPP server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}(accessed June 1, 2015).}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}(accessed June 1, 2015).}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}(accessed June 1, 2015).} and \emph{Prosody}\footnote{\url{http://prosody.im/}(accessed June 1, 2015).}.
+\emph{Ejabberd} is the server that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}(accessed June 1, 2015).}.
   %RP fazes muitos paragrafos. Um paragrafo por tecnologia é normalmente suficiente se não forem muito grandes.

### 2015-06-02 tese/state04.tex
SotA 4 restrucure

@@
 Storing multimedia content is not a viable solution because most browsers recommend limiting local storage to at most five megabytes per origin.
+
+\subsubsection{Streaming and Recording}~\\
+
+ 	In order to provide a way to record and playback streams, additional servers will be required to process and record the large amount of data generated by audio and video streaming.
 
@@
   %RP isto está tudo no RFC3550? Se não, faltam muitas citações.
+  %HR esse RFC é grande
 
   \ac{RTP} recorders are independent of payload encoding, they don't decode \ac{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.
+
+\subsubsection{Media Types}~\\
 
@@
 
-	In order to play any kind of stream, a player for interactive stream is needed for downloading an environment, decoding the \ac{RTP} payload to determine the state and display it to the user.
+	In order to play a stream, a player must be prepared to interpret the stream content.
+For interactive stream the player must download an environment, decode the \ac{RTP} payload to determine the state and display it to the user.
 Streaming interactive media like a combination of \ac{HTML}, \ac{CSS} and JavaScript requires more than interpreting the code: a streamed user interface may contain an internal state that is not shown on code.
         %RP não percebi
+
+ \subsubsection{Recording and Streaming Interactive Media}~\\
+
+    Mauve et al., 1999 proposed an \ac{RTP} profile for real-time transmission of interactive media\cite{interactive_stream}.
+This new profile reuses much of video and audio profile implementation, integrating the interactive component.
+Hilt et al., 1999 explained how to record interactive video with this new profile \cite{interactive_record}.
+        %RP não uses a citação como sujetio das frases. O trabalho não tem um nome?
 
@@
 %RP referencia?
+%HR vem dos senhores de cima  (Mauve e Hilt), isto estava mal organizado
 
@@
 
-	\cite{interactive_stream} concluded that the ability to extract the objects state in order to synchronize them and the ability to intercept events in order to control remote objects are realized using the \ac{MVC} concept.
+	\cite{interactive_stream} concluded that the ability to extract the objects state in order to synchronize them and the ability to intercept events in order to control remote objects can be realized using the \ac{MVC} concept.
 %RP are realized ou ``can be realized''?
@@
 
-	Using the \ac{MVC} concept will it make possible to implement an interactive \ac{RTP} application that records, play, fast forward, fast rewind, stop and jump to random positions.
+	Using the \ac{MVC} concept will it make possible to implement an interactive \ac{WebRTC} application that records, play, fast forward, fast rewind, stop and jump to random positions.
         %RP porquê RTP aplication? Ainda não disseste porque tem de ser RTP! Não era webrtc?
-
-        \cite{interactive_stream} also proposed an \ac{RTP} profile for real-time transmission of interactive media.
-This new profile reuses much of video and audio profile implementation, integrating the interactive component.
-\cite{interactive_record} explained how to record interactive video with this new profile.
-        %RP não uses a citação como sujetio das frases. O trabalho não tem um nome?
 
@@
 Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
-        In this scenario, a more efficient alternative to peer-to-peer is the use of a Multipoint control unit.
+        In this scenario, a more efficient alternative to peer-to-peer is the use of a Multipoint Control Unit (MCU).
 
@@
 Despite this choice for signaling protocol, \emph{Jitsi Video Bridge} also supports \ac{SIP}.
+
+  \subsubsection{Collaborative Environment}~\\
 

### 2015-06-02 tese/project.tex
Reviewed the conclusions

@@
 WebRTC is enabling new usage scenarios for communication and collaboration applications.
-However, the nature of such scenarios lead to significant challenges, such as the need to record interactive audio and video streams.
+This and other technologies allow the Hyper-link concept to be applied to a new media: real-time audio and video.
 
@@
 Theses communications will be enriched using hypermedia concepts, the possibility to record and playback its content and the integration of collaborative tools.
+
+  A prototype will be implemented in order to validate these concepts.
+In order to implement it, significant challenges, such as the need to record interactive audio and video streams, will have to be overcome, making this a challenging project.
 

### 2015-06-07 tese/intro.tex
possible end of SotA 1
[ADD] added first beamer presentation

@@
 
-	As \cite{geddes} states, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
+	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
 The concept of structured voice and video synchronized with hypermedia is called hypervoice.
@@
 
-As \cite{geddes} said, the quality of the interaction worsens as the number of users increase.
-In our testing phase we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.
+According to Martin Geddes, the quality of the interaction worsens as the number of users increase\cite{geddes}.
+In our testing phases we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.
 

### 2015-06-07 tese/project.tex
[ADD] added first beamer presentation

@@
 
-WebRTC technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
-The nature of web browser applications already follow the hypermedia concept, which makes WebRTC the ideal technology to apply the hyper-linked communications concepts.
-The native support of WebRTC in operating systems extends its usage to outside the web browser, allowing to explore functionalities that browsers are poor to support such as video recording and massive information storage.
+\ac{WebRTC} technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
+The nature of web browser applications already follow the hypermedia concept, which makes \ac{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
+The native support of \ac{WebRTC} in operating systems extends its usage to outside the web browser, allowing to explore functionalities that browsers are poor to support such as video recording and massive information storage.
 
-Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable colaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed latter.
-This application will target the web browser platform, resorting to WebRTC.
+Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed latter.
+This application will target the web browser platform, resorting to \ac{WebRTC}.
 
-In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose an architecture to implemente an hyper-linked communication application based on WebRTC and provide a plan for its implementation and evaluation.
+In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose an architecture to implement an hyper-linked communication application based on \ac{WebRTC} and provide a plan for its implementation and evaluation.
 
@@
 \section{Conclusions}\label{concl} % English
-The development of WebRTC by the World Wide Web Consortium (W3C) allows web browser-to-browser audio, video and data communication.
-WebRTC is enabling new usage scenarios for communication and collaboration applications.
+The development of \ac{WebRTC} by the World Wide Web Consortium (W3C) allows web browser-to-browser audio, video and data communication.
+\ac{WebRTC} is enabling new usage scenarios for communication and collaboration applications.
 This and other technologies allow the Hyper-link concept to be applied to a new media: real-time audio and video.
 
-  This project proposes a solution that uses WebRTC for creating an application that provides audio, data, text and data real-time communications.
+  This project proposes a solution that uses \ac{WebRTC} for creating an application that provides audio, data, text and data real-time communications.
 Theses communications will be enriched using hypermedia concepts, the possibility to record and playback its content and the integration of collaborative tools.

### 2015-06-07 tese/schd01.tex
need conclusions and abstract
[ADD] modules, changed titles
Reviewed methodology
[ADD] added first beamer presentation

@@
-\tikzset{every picture/.style={xscale=0.51,yscale=0.51,transform shape}}
+The most important task in order to implement our project is the infrastructure, without it our project wouldn't work.
+The basic infrastructure will be our first concern, namely the Web Server configuration, ICE and signaling servers.
+
+The Web development can only start after the basic infrastructure configuration. When leading with streams, the infrastructure will be complemented with stream servers.
+
+There will be four prototypes, the first one should implement most of streaming and collaborative functionalities.
+The second prototype will have the same functionalities but a more friendly user interface which will be tested and reviewed by potential users.
+The third prototype will be an improvement based on the users feedback.
+In the final prototype we will improve our system performance.
+
+\tikzset{every picture/.style={xscale=0.49,yscale=0.55,transform shape}}
 
@@
 %First Group
-\ganttgroup{Thesis}{1}{10} \\
-\ganttbar{Task 1}{2}{3} \\
-\ganttbar{Task 2}{4}{10} \\
-\ganttbar{Task 3}{9}{12}\\
+\ganttgroup{Thesis}{1}{35} \\
+\ganttbar{Related work evaluation}{1}{30}\\
+\ganttbar{Thesis writing}{19}{35} \\
+\ganttbar{Paper writing}{25}{35} \\
+\ganttmilestone{Delivery}{35}{35} \\
 %\ganttlink{elem0}{elem1}
-\ganttgroup{Resume}{1}{10} \\
+%\ganttgroup{Resume}{1}{10} \\
 
@@
 \ganttgroup{Implementation}{1}{1}\ganttgroup{}{4}{8}\ganttgroup{}{11}{26}\ganttgroup{}{29}{30} \\
-\ganttbar{ICE configuration}{1}{1} \\
-\ganttbar{XMPP configuration}{1}{1} \\
+\ganttbar{Basic configuration}{1}{1} \\
 %\ganttlink{elem4}{elem5}
@@
 
-\ganttlink{elem1}{elem2}
-\ganttlink{elem2}{elem3}
-\ganttlink{elem9}{elem11}
-\ganttlink{elem10}{elem11}
-\ganttlink[link type=F-S]{elem12}{elem13}
-\ganttlink{elem13}{elem14}
-\ganttlink[link type=F-S]{elem14}{elem16}
-\ganttlink[link type=F-S]{elem16}{elem18}
-\ganttlink[link type=F-S]{elem18}{elem19}
-\ganttlink[link type=F-S]{elem19}{elem21}
-\ganttlink[link type=F-S]{elem21}{elem22}
+\ganttlink{elem9}{elem10}
+\ganttlink[link type=F-S]{elem11}{elem12}
+\ganttlink{elem12}{elem13}
+\ganttlink[link type=F-S]{elem13}{elem15}
+\ganttlink[link type=F-S]{elem15}{elem17}
+\ganttlink[link type=F-S]{elem17}{elem18}
+\ganttlink[link type=F-S]{elem18}{elem20}
+\ganttlink[link type=F-S]{elem20}{elem21}
 \end{ganttchart}
+
+%RP o que estás duas semanas a fazer no final?
+Porque não entregas após a escrita.
+Se é tempo para eu rever, então deveria haver outra fase para corrigires.
+Penso que podemos considerar que isso fica tudo dentro do ``writting''.
+Se quiseres aumenta.
+
+%RP troca ``essay writting'' por ``thesis writting''

### 2015-06-07 tese/state02.tex
[ADD] added first beamer presentation

@@
 %RP a frase anterior é estranha
-  Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}(accessed June 1, 2015).}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}(accessed June 1, 2015).} and \emph{sipML5}\footnote{\url{http://sipml5.org/}} are used on the client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
+  Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}(accessed June 1, 2015).}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}(accessed June 1, 2015).} and \emph{sipML5}\footnote{\url{http://sipml5.org/}(accessed June 1, 2015).} are used on the client side to parse and encode \ac{SIP} messages, making \ac{SIP} accessible to web based applications.
 %RP quando usas footnotes para indicar sites deves acrescentar ``last accessed (data)''. Isto é importante pois os sites são documentos efémeros que podem dizer uma coisa diferente mais tarde.

### 2015-06-28 implementation_research/Ejabberd Resource Constraint: Users are not allowed to register accounts so fast | Sid___files/a_002.js
[ADD] register + login

@@ -1,0 +1,1 @@
+/**/skimlinksApplyHandlers({"country":"PT","sessid":"bd10f2249861afee148632498892277f","guid":"492594c6b114f0cdf04c0b628625ca94","merchant_domains":[]});

### 2015-06-28 implementation_research/xmpp - Ejabberd and in-band registration - Stack Overflow_files/ados.js

@@ -1,0 +1,34 @@
+var html__72501 = '\x3ca\x20href\x3d\x22http\x3a\x2f\x2fengine\x2eadzerk\x2enet\x2fr\x3fe\x3deyJhdiI6NDE0LCJhdCI6NCwiYnQiOjAsImNtIjo5MTgyOCwiY2giOjExNzgsImNyIjo3MzgwMjksImRpIjoiODkwYTgzZDMzOWNhNGQ5ODkxMTBlOWIyYWY3MWRkZmMiLCJkbSI6MSwiZmMiOjc2OTA3NywiZmwiOjE2NTI5MywiaXAiOiI4NS4xMzkuNS4yNDAiLCJrdyI6InhtcHAsZWphYmJlcmQiLCJudyI6MjIsInBjIjowLCJwciI6NzY1LCJydCI6MSwicmYiOiJodHRwOi8vd3d3Lmdvb2dsZS5wdC91cmw_c2E9dCZyY3Q9aiZxPSZlc3JjPXMmc291cmNlPXdlYiZjZD0xJnZlZD0wQ0NBUUZqQUEmdXJsPWh0dHA6Ly9zdGFja292ZXJmbG93LmNvbS9xdWVzdGlvbnMvNTczMjE3My9lamFiYmVyZC1hbmQtaW4tYmFuZC1yZWdpc3RyYXRpb24mZWk9RTdxT1ZjaXZOb2E3VWJlZXI5QUcmdXNnPUFGUWpDTkdyekROYWJhMV81ZnNPdzRXR2M1OWdyV0FNdlEmc2lnMj1IaVF2aVhBNkExTW1ZbW1ZZFp3TnRBJmJ2bT1idi45Njc4MzQwNSxkLmQyNCIsInN0Ijo4Mjc3LCJ1ayI6InVlMS1jM2Y1ZmIxMzY0ZTE0MzhkYWM5OTk4ZWQxYzJiMGU3MSIsInpuIjo0MywidHMiOjE0MzU0MTcxMTgwMDcsImJmIjp0cnVlLCJwbiI6ImFkemVyazIxMjY4Mjk1MjkiLCJ1ciI6Imh0dHA6Ly9wdC5zdGFja292ZXJmbG93LmNvbS8_dXRtX3NvdXJjZT1zdGFja292ZXJmbG93LmNvbSZ1dG1fbWVkaXVtPWFkJnV0bV9jYW1wYWlnbj1wdH
[…]

### 2015-06-28 implementation_research/xmpp - Ejabberd and in-band registration - Stack Overflow_files/full-anon.js

@@ -1,0 +1,3 @@
+"use strict";function moveScroller(){var e=$("#scroller").width(),t=function(){var t=$(window).scrollTop(),n=$("#scroller-anchor").offset().top,i=$("#scroller");t>n?i.height()>$(window).height()?i.css({"position":"fixed","top":"","bottom":"0px","width":e}):i.css({"position":"fixed","top":"0px","bottom":"","width":e}):n>=t&&i.css({"position":"relative","top":"","bottom":""})};$(window).scroll(t).resize(t),t()}function sanitizeAndSplitTags(e,t){e=$.trim(e).replace(/([A-Za-z0-9])\+(?=[A-Za-z0-9])/g,"$1 ");for(var n=e.split(/[\s|,;]+/),i=[],o=0;o<n.length;o++){var a=n[o].toLowerCase();StackExchange.options.site.nonAsciiTags||(a=StackExchange.helpers.noDiacritics(a)),a=a.replace(/_/g,"-");var s="[^a-z0-9.#+"+(t?"*":"")+(StackExchange.options.site.nonAsciiTags?"0-9,A-Z,_,a-z,ª
[…]

### 2015-06-28 implementation_research/xmpp - Ejabberd and in-band registration - Stack Overflow_files/post-validation.js

@@ -1,0 +1,1 @@
+StackExchange.postValidation=function(){function e(e,t,n,i){var r=e.find('input[type="submit"]:visible'),a=r.length&&r.is(":enabled");a&&r.attr("disabled",!0),o(e,i),s(e,t,n,i),c(e),u(e),d(e);var f=function(){1!=t||e.find(C).length?(l(e,i),a&&r.attr("disabled",!1)):setTimeout(f,250)};f()}function t(t,i,o,s,l){e(t,i,s,o);var c,u=function(e){if(e.success)if(l)l(e);else{var n=window.location.href.split("#")[0],r=e.redirectTo.split("#")[0];0==r.indexOf("/")&&(r=window.location.protocol+"//"+window.location.hostname+r),c=!0,window.location=e.redirectTo,n.toLowerCase()==r.toLowerCase()&&window.location.reload(!0)}else e.captchaHtml?e.nocaptcha?StackExchange.nocaptcha.init(e.captchaHtml,u):StackExchange.captcha.init(e.captchaHtml,u):e.errors?(t.find("input[name=priorAttemptCount]").val(function(e,t){return(+t+1||0).toString()}),p(e.errors,t,i,o,e.warnings)):t.find('input[type="submit"]:visible').parent().showErrorMessage(e.message)};t.submit(function(){if(t.find("#answer-from-ask").is(":checked"))return!0;var e=t.find(E);if("[Edit removed during grace period]"==$.trim(e.val()))return g(e,["Comment reserved for system use. Please use an appropriate comment."],f()),!1;a(),StackExchange.navPrevention&&StackExchange.navPrevention.stop();var i=t.find('input[type="submit"]:visible');if(i.parent().addSpinner(),StackExchange.helpers.disableSubmitButton(t),StackExchange.options.site.enableNewTagCreationWarning){var s=t.find(C).parent().find("input#tagnames"),l=s.prop("defaultValue");if(s.val()!==l)return $.ajax({"type":"GET","url":"/posts/new-tags-warning","dataType":"json","data":{"tags":s.val()},"success":function(e){e.showWarning?i.loadPopup({"html":e.html,"dontShow":!0,"prepend":!0,"loaded":function(e){n(e,t,c,o,u)}}):r(t,o,c,u)}}),!1}return setTimeout(function(){r(t,o,c,u)},0),!1})}function n(e,t,n,a,o){e.bind("popupClose",function(){i(t,n)}),e.find(".submit-post").
[…]

### 2016-01-10 draft/draft.tex
[ADD] thesis draft begin (XMPP stuff)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Arsclassica Article
% LaTeX Template
% Version 1.1 (10/6/14)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
]{scrartcl}

\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether

\title{Draft for master thesis} % The article title

\author{Henrique Rocha} % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block

\date{} % An optional date to appear under the author(s)

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Print the title/author/date block

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------

\section{Backend}

\subsection{XMPP Approach}

One of the advantages that XMPP based applications offer is the possibility to send messages across different servers, an XMPP user is identified by its \textit{Jabber ID} which is definined by the pair username and server name in the form \textit{username@server}.

The user profile is represented in terms of Data Forms (XEP-0004) protocol and managed (registration and update) through \textit{Info/Query} (IQ) [rfc3920/IQ Semantics] requests to the XMPP server which responds with an IQ response containing information to retrieve the registration fields needed to fill its profile (eg username, password, telephone).

An important aspect of this model is the ability to define a user as a set of attributes that can change dinamically in order to fit the IM applications needs.

The connection to the XMPP server could be done through the same web server as the user is connected, but this web server would handle a lot of connections and that would have a great impact on the overall system performance. Our proposed solution distributes the XMPP connections among users relieving this extra work to the web browsers.

The first working prototype of our solution implemented the user registration and authentication components through \textit{Strophe.js}. The next step on implementation would be the access to the user profile and consequentally all the other features needed for our solution. Typically a web application consists in multiple web pages which are navigated by the users. Each time a webpage is accessed, either from a new context or a transition from a previous page, its context is cleared except for local storage and cookies, the javascript context is cleared, including the XMPP connections performed by strophe.js, an automatic mechanism is required for avoiding the user implicit reauthentication.

One solution for reauthentication can be achieved by storing the user's JabberID and password on local storage and everytime a page is accessed the authentication is performed without the users knlowledge. Clearly this solution represents security flaws, the local storage can be easily accessed (HOW? explain!) locally or by performing a cross site scripting attack and reveal all the local storage, the same problem arises if the credentials are stored on cookies.

An alternative solution is known as Session Attachment (see https://metajack.im/2008/10/03/getting-attached-to-strophe/), which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} is passed to strophe.js in order to re-connect to the same stream on XMPP server. Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request which is worthless if a user mantains more than one tab opened, for example for multiple conversations at the same time.

Another alternative would be the development of our solution in a single web page, which would increse drastically the complexity of our web application.

We know that an \textit{XMPP} based application could simplify our work just by using functionalities that XMPP already implemment, but from all the \textit{XMPP} features, we don't need all of them, in fact we just need a subset of XMPP features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.

\subsection{Our approach}

In order to implement a solution without the \textit{XMPP}'s limitations that we presented on the previous section, we decided to implement our own backend.

One important aspect that we should take into account is the scalability of our solution. All the information that we need to store relative to the users and chat rooms should be stored on a scalable database.

As our database grows, we may need to distribue the load among multiple servers, which can be achieved on SQL servers by spliting tables among servers, but this could be a challenge operation when doing relations between tables. For this reason we opted for using a NoSQL database. Many of NoSQL databases are built having the scalability functionality from start.

One option for a NoSQL dabase is MongoDB, there are others but this one seems to have popularity, it is free and well documented. There are no strong reasons for excluding options like Apache Cassandra or others.

The authentication problem that XMPP raised when implementing a client on a web browser is easily solved by using cookies to identify authenticated users. A cookie is mantained as long as is specified by its expiration date and the user only needs to reauthenticate when it logs out or the cookie expires, changing from page to page does not raises any reauthentication problem.

On our solution the actions that are performed by clients are directly sent to web servers that control the validity of the actions and performs the respective changes to the database. This logic on the XMPP approach would be distributed among XMPP servers and not the webservers, so now our webservers will increase their complexity by implementing the features we needed from XMPP.

\subsubsection{Users}

\subsubsection{Groups}

\subsubsection{Permissions}

\subsubsection{Signaling}

\end{document}

### 2016-01-10 tese1/intro.tex
[ADD] thesis draft begin (XMPP stuff)

--- 2015-06-07 tese/intro.tex
@@
 
-	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
+	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice's capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
 The concept of structured voice and video synchronized with hypermedia is called hypervoice.

### 2016-01-10 tese1/method.tex
[ADD] thesis draft begin (XMPP stuff)

%RP falta introdução à secção a explicar o que vai ser discutido aqui.
%RP texto que está nos 3 parágrafos seugintes ficava bem no planned schedule. Talvez esse fique melhor antes do evaluation methodology.
This section presents how we plan to implement and test our solution.

\subsection{Evaluation Methodology} % Section
	Our solution will be evaluated with several experiments, either qualitative and quantitative.

Our web application will be tested with real users by asking them to perform some tasks, count the respective spent time and note their difficulties. After these tests, it will be suggested to users to fill out questionnaires.

The results of user tests be used to improve the following prototype iterations, it is expected to notice both qualitative and quantitative improvements.

%RP elaborar o tipo de testes que vais fazer com os utilizadores: pedir para fazer tarefas e cronometrar o tempo que demoram. preencher questionários, decobrirem sozinhos como fazer algo, fazer após explicações, etc.

%RP detalhar os testes de desempenho de que falas.

The system in general will be validated with benchmarks that will focus on application delay when performing heavy tasks such as sharing a file and editing it in collaboration with other users. The amount of users will gradually increase in order to understand how much the performance degrades.

### 2016-01-10 tese1/state04.tex
[ADD] thesis draft begin (XMPP stuff)

--- 2015-06-02 tese/state04.tex
@@
 
+Section \ref{recstream} presents the RTP protocol how it can be used to record media streams.
+Section \ref{mediatype} describes the media types and what types of media should be streamed.
+Section\ref{intrecord} addresses how an interactive media can be recorded.
+Section \ref{collabenv} presents the collaborative environment and libraries to synchronize distributed object among users.
+
+\subsubsection{Streaming and Recording}\label{recstream}~\\
 	If, for instance, one wants to rewind a real-time video, recordings will be needed from whom is streaming the video. 
@@
 Storing multimedia content is not a viable solution because most browsers recommend limiting local storage to at most five megabytes per origin.
-
-\subsubsection{Streaming and Recording}~\\
 
@@
 
-\subsubsection{Media Types}~\\
+        %RP lá vais tu mudar de tema da conversa sem explicar porquê....
+        Even though one on one calls are common, there are occasions when several people take part of the same video call.
+	Multi-party video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants.
+Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
+        In this scenario, a more efficient alternative to peer-to-peer is the use of a Multipoint Control Unit (MCU).
 
-  %RP estavas a falar de gravar video e passaste a falar de protocolos de streaming. Convém explicar o racional para esta abordagem. Começa a secção a explicar sobre o que vais falar e porquê.
-  %RP e usa subsecções...
+	\emph{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}(accessed June 2, 2015).} receives one stream from every participant on a conference, either from a \emph{jitsi} client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
+Although all the participants still need to download all the streams from the \emph{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
+	\emph{Jitsi Video Bridge} uses \ac{XMPP} as a signaling protocol and its \emph{colibri} extension \cite{xep0340} to reserve channels for video transmission.
+Despite this choice for signaling protocol, \emph{Jitsi Video Bridge} also supports \ac{SIP}.
 
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.45\textwidth]{figures/media_types.png}
-	\caption{Media Types}
-\end{figure}
-
+\subsubsection{Media Types}\label{mediatype}~\\
 	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuous, the second one describes it as interactive or non-interactive.
@@
 Interactive media is characterized by its state being changed by external events such as user interactions.
-
+  %RP estavas a falar de gravar video e passaste a falar de protocolos de streaming. Convém explicar o racional para esta abordagem. Começa a secção a explicar sobre o que vais falar e porquê.
+  %RP e usa subsecções...
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.4\textwidth]{figures/media_types.png}
+	\caption{Media Types}
+\end{figure}
 	For example, an image is non-interactive and discrete, while a video is continuous and non-interactive.
@@
 
- \subsubsection{Recording and Streaming Interactive Media}~\\
-
-    Mauve et al., 1999 proposed an \ac{RTP} profile for real-time transmission of interactive media\cite{interactive_stream}.
+ \subsubsection{Recording and Streaming Interactive Media}\label{intrecord}~\\
+    Mauve et al.
+proposed an \ac{RTP} profile for real-time transmission of interactive media\cite{interactive_stream}.
 This new profile reuses much of video and audio profile implementation, integrating the interactive component.
-Hilt et al., 1999 explained how to record interactive video with this new profile \cite{interactive_record}.
+Hilt et al.
+explained how to record interactive video with this new profile \cite{interactive_record}.
         %RP não uses a citação como sujetio das frases. O trabalho não tem um nome?
@@
 
-	\cite{interactive_stream} concluded that the ability to extract the objects state in order to synchronize them and the ability to intercept events in order to control remote objects can be realized using the \ac{MVC} concept.
+	Mauve et al.
+concluded that the ability to extract the objects state in order to synchronize them and the ability to intercept events in order to control remote objects can be realized using the \ac{MVC} concept\cite{interactive_stream}.
 %RP are realized ou ``can be realized''?
@@
 
-        %RP lá vais tu mudar de tema da conversa sem explicar porquê....
-        Even though one on one calls are common, there are occasions when several people take part of the same video call.
-	Multi-party video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants.
-Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
-        In this scenario, a more efficient alternative to peer-to-peer is the use of a Multipoint Control Unit (MCU).
-
-	\emph{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}} receives one stream from every participant on a conference, either from a \emph{jitsi} client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
-Although all the participants still need to download all the streams from the \emph{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
-	\emph{Jitsi Video Bridge} uses \ac{XMPP} as a signaling protocol and its \emph{colibri} extension \cite{xep0340} to reserve channels for video transmission.
-Despite this choice for signaling protocol, \emph{Jitsi Video Bridge} also supports \ac{SIP}.
-
-  \subsubsection{Collaborative Environment}~\\
-
+  \subsubsection{Collaborative Environment}\label{collabenv}~\\
 	\emph{Google Wave} was a distributed collaboration platform based on \emph{Jupiter}\cite{jupiter} that adopted \ac{OT} techniques.
@@
 
-	\emph{ShareJS}\footnote{\url{http://sharejs.org/}} is an \ac{OT} \emph{JavaScript} library, developed by the ex \emph{Google Wave} engineer Joseph Gentle, for collaborative text and \ac{JSON} documents edition in real-time.
+	\emph{ShareJS}\footnote{\url{http://sharejs.org/}(accessed June 2, 2015).} is an \ac{OT} \emph{JavaScript} library, developed by the ex \emph{Google Wave} engineer Joseph Gentle, for collaborative text and \ac{JSON} documents edition in real-time.
 
-	\emph{TogetherJS}\footnote{\url{http://togetherjs.com/}} is a JavaScript library that uses \ac{WebRTC} for collaborative web applications.
+	\emph{TogetherJS}\footnote{\url{http://togetherjs.com/}(accessed June 2, 2015).} is a JavaScript library that uses \ac{WebRTC} for collaborative web applications.
 It uses \ac{JSON} messages for \ac{OT} concurrency control but it does not provide storage.
 
-	\emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}} is a collaborative framework with its own server implementation, it supports four types of collaborative elements: String, Lists, Maps and Custom objects.
+	\emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}(accessed June 2, 2015).} is a collaborative framework with its own server implementation, it supports four types of collaborative elements: String, Lists, Maps and Custom objects.
         %RP estes 3 últimos parágrafos estão desligados do resto. O que pretendes apresentar?

### 2016-01-10 project/httpserver/app/dtos/KeyValue.java

@@ -1,0 +1,29 @@
+package dtos;
+
+public class KeyValue<T> {
+	private String key;
+	private T value;
+	public String getKey() {
+		return key;
+	}
+
+	public void setKey(String key) {
+		this.key = key;
+	}
+
+	public T getValue() {
+		return value;
+	}
+
+	public void setValue(T value) {
+		this.value = value;
+	}
+
+	public KeyValue(String key, T value) {
+		super();
+		this.key = key;
+		this.value = value;
+	}
+
+	
+}

### 2016-01-10 project/httpserver/app/models/Tag.java

@@ -3,6 +3,7 @@ package models;
 import com.mongodb.client.FindIterable;
 import com.mongodb.client.MongoCollection;
 import com.mongodb.client.MongoCursor;
+import dtos.KeyValue;
 import org.bson.Document;
 import org.bson.types.ObjectId;
 import services.Service;
@@ -17,125 +18,151 @@ import java.util.regex.Pattern;
  */
 public class Tag {
 
+    private List<String> search;
+    private ObjectId id = null;
+    private String type;
+    private ObjectId owner = null;
+    private Document filter= null;
 
-
-
-        private List<String> value;
-        private ObjectId id = null;
-
-    public Tag(ObjectId owner, List<String> value) {
+    public Tag(ObjectId owner, List<String> search, String type,Document filter) {
         this.owner = owner;
-        this.value = value;
+        this.search = search;
+        this.type = type;
+        this.filter = filter;
     }
-
 
     public ObjectId getOwner() {
-            return owner;
-        }
-
-        private ObjectId owner = null;
-
-
-        public Object getValue() {
-            return value;
-        }
-
-
-        private static MongoCollection<Document> collection;
-
-        public static MongoCollection<Document> getCollection() {
-            if(collection==null)
-                collection = Service.getDatabase().getCollection(Tag.class.getName());
-            return collection;
-        }
-
-        private static Tag load(Document doc) {
-            Tag user = new Tag();
-            user.id = doc.getObjectId("_id");
-            user.owner = doc.getObjectId("owner");
-            user.value = (List<String>) doc.get("value");
-            return user;
-        }
-
-        public Tag save() {
-            Document doc = new Document();
-
-            doc.put("value", value);
-            doc.put("owner", owner);
-
-            if(id==null){
-                getCollection().insertOne(doc);
-                id = doc.getObjectId("_id");
-            }
-            else {
-                doc.put("_id", id);
-                getCollection().findOneAndReplace(new Document("_id",id),doc);
-            }
-            return this;
-        }
-
-        public ObjectId getId() {
-            return id;
-        }
-
-
-        public static Tag findById(ObjectId id) {
-            Document doc = new Document("_id", id);
-            FindIterable<Document> iter = getCollection().find(doc);
-            doc = iter.first();
-            return doc != null ? load(doc) : null;
-        }
-
-        private Tag() {
-
-        }
-
-
-
-        public static List<Tag> listByOwner(ObjectId id) {
[…]

### 2016-01-21 curriculo/cv.tex
[ADD] local video\n[FIX] mic off on leave

@@
 \addbibresource{bibliography.bib}
-
 \begin{document}
@@
     ~
-    \href{mailto:hdlopesrocha91@gmail.com}{hdlopesrocha91\\@gmail.com}
+    \href{mailto:hdlopesrocha91@gmail.com}{hdlopesrocha91@gmail.com}
     \href{http://web.ist.utl.pt/ist168621}{WebSite}
@@
     {Instituto Superior Técnico - Taguspark}
-    {Telecommunications and Informatics Engineering}
-
+    {Telecommunications and Informatics Engineering (Expected Grade: 16 in 20)}
    \entry
@@
     {Instituto Superior Técnico - Taguspark}
-    {Communication Networks Engineering}
+    {Communication Networks Engineering (Final Grade: 13.2 in 20)}
 
@@
 
-%\section{publications}
-
-%Put your publications here!
-
-%%% This piece of code has been commented by Karol Kozioł due to biblatex errors. 
-% 
-%\printbibsection{article}{article in peer-reviewed journal}
-%\begin{refsection}
-%  \nocite{*}
-%  \printbibliography[sorting=chronological, type=inproceedings, title={international peer-reviewed conferences/proceedings}, notkeyword={france}, heading=subbibliography]
-%\end{refsection}
-%\begin{refsection}
-%  \nocite{*}
-%  \printbibliography[sorting=chronological, type=inproceedings, title={local peer-reviewed conferences/proceedings}, keyword={france}, heading=subbibliography]
-%\end{refsection}
-%\printbibsection{misc}{other publications}
-%\printbibsection{report}{research reports}
+\newpage
+\section{grades}
+\subsection{Bachelor - Communication Networks Engineering}
+\input{grades_bachelor.tex}
+\subsection{Master - Telecommunications and Informatics Engineering}
+\input{grades_master.tex}
 

### 2016-01-21 curriculo/grades_master.tex
[ADD] local video\n[FIX] mic off on leave

\textit{Mobile, Peer-To-Peer and Cloud Applications}
	\begin{itemize}
		\item Cloud Computing - 15
		\item Mobile Computing - 15
		\item Peer-to-Peer Systems and Overlay Networks - 15
	\end{itemize}
\textit{Entrepreneurial Leadership, Organization, Economy and Management}
	\begin{itemize}
		\item Decision Support Models - 14
	\end{itemize}
\textit{Common Branch}
	\begin{itemize}
		\item Forensics Cyber-Security - 16
		\item Middleware for Distributed Internet Applications - 14
		\item Telecommunications and Positioning Satellite Systems - 17
		\item Performance Evaluation and Dimensioning of Networks and Systems - 15
		\item Web Data Extraction and Analysis - 16
	\end{itemize}
\textit{Master Thesis}
	\begin{itemize}
		\item Master Project in Telecommunications and Computer Engineering - 17
	\end{itemize}

\textit{Elective Courses}
	\begin{itemize}
		\item Services Engineering - 17
		\item Database Administration and Tuning - 15
	\end{itemize}

### 2016-01-21 project/httpserver/app/main/UserSession.java

@@ -151,7 +151,7 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                 }
             });
 
-            ep.setStunServerAddress("74.125.206.127");
+            ep.setStunServerAddress("64.233.184.127");
             ep.setStunServerPort(19302);
             endPoints.put(name, ep);
         }

### 2016-01-21 project/httpserver/public/this/js/hyper.socket.js

@@ -9,10 +9,7 @@ var local_none = {
 	'offerToReceiveVideo':true 
 };
 
-var local_user = {
-	"audio" : true,
-	"video" : true
-};
+
 
 var screen_user = {
 	video: {
@@ -57,7 +54,7 @@ var Kurento = new (function() {
     var contentArrivedCallback = null;
     var setTimeCallback = null;
     var removedUserCallback= null;
-
+    var localVideoCallback=null;
 
 
 	this.microphoneState = true;
@@ -241,7 +238,7 @@ var Kurento = new (function() {
 		Kurento.peerConnection[name] = pc;
 	}
 	
-	this.start = function(groupId,mode,kscb,npcb,nvcb,mvcb,nrcb,nmcb,tacb,cacb,trcb,stcb,rucb) {
+	this.start = function(groupId,mode,kscb,npcb,nvcb,mvcb,nrcb,nmcb,tacb,cacb,trcb,stcb,rucb,lvcb) {
 		newParticipantsCallback = npcb;
 		newVideoCallback = nvcb;
 		mixerVideoCallback = mvcb;
@@ -252,6 +249,7 @@ var Kurento = new (function() {
 		talkReceivedCallback = trcb;
 		setTimeCallback = stcb;
 		removedUserCallback = rucb;
+		localVideoCallback = lvcb;
 		if ("WebSocket" in window) {
 			Kurento.webSocket = new WebSocket(wsurl("/ws/room/" + groupId));
 		
@@ -334,7 +332,9 @@ var Kurento = new (function() {
 
 				
 				if(mode==1){
-					navigator.mediaDevices.getUserMedia({ audio: true, video: true }).then(function(stream) {					
+					navigator.mediaDevices.getUserMedia({ audio: true, video: true }).then(function(stream) {
+						localVideoCallback(window.URL.createObjectURL(stream));
+
 						Kurento.peerConnection["main"].addStream(stream);
 						Kurento.peerConnection["main"].createOffer(function (lsd) {		
 							console.log("createOfferToSendReceive",lsd);
@@ -355,8 +355,9 @@ var Kurento = new (function() {
 				
 				// XXX [CLIENT_OFFER_01] XXX
 				else if(mode==0 ){
-					navigator.getUserMedia(local_user, function(stream) {
-						primaryStream = stream;
+					navigator.getUserMedia({"audio" : true,	"video" : true }, function(stream) {
+						localVideoCallback(window.URL.createObjectURL(stream));
[…]

### 2016-02-21 project/stats/draft.tex
[ADD] stats folder

\documentclass[
10pt, % Main document font size
a4paper, % Paper type, use 'letterpaper' for US Letter paper
oneside, % One page layout (no page indentation)
]{scrartcl}

\usepackage{amsmath} 
\usepackage{adjustbox}
\usepackage{textcomp}
\usepackage{environ}
\usepackage{tikz}
\usepackage{graphics}
\usepackage{pgfplots} 
\pgfplotsset{compat=1.9}

\title{Draft for master thesis} % The article title

\author{Henrique Rocha} 
\date{} 
\begin{document}

	\maketitle % Print the title/author/date block

	%----------------------------------------------------------------------------------------
	%	INTRODUCTION
	%----------------------------------------------------------------------------------------

	\section{Test}

\subsection{Usability Test}

	\subsubsection{Task 1 (Video)}
		Register User, add the coordinator as a friend, create a room, add the coordinator to the room wait for the coordinator to join the room and change for the individual view.

	\subsubsection{Task 2 (Tags)}
		The coordinator creates previously a room with existing content, namely, tags and subtitles.
		The user task is to enter in the public room and search by "Cauchy Rule" get an url for that time and send it back to the other users.
		Next the user should create a tag on exact arrival time (when he entered the room).

	\subsubsection{Task 3 (Content)}
		The user must create a subtitle with 10s duration at the time of arrival with the content "I've just arrived" and search for the previously created tag

	\subsubsection{Task 4 (Colaborative)}

\subsection{Performance Test}

	\begin{center}
	\input{points3.tex}
	\end{center}

	\begin{center}
	\input{points2.tex}
	\end{center}

	\begin{center}
	\input{points.tex}
	\end{center}
\end{document}

### 2016-03-02 curriculo/cv.tex
[ADD] update cv

@@
 \addbibresource{bibliography.bib}
+\usepackage{listings}
+\usepackage{color}
+
+\definecolor{pblue}{rgb}{0.13,0.13,1}
+\definecolor{pgreen}{rgb}{0,0.5,0}
+\definecolor{pred}{rgb}{0.9,0,0}
+\definecolor{pgrey}{rgb}{0.46,0.45,0.48}
+
+\lstset{language=Java,
+  showspaces=false,
+  showtabs=false,
+  breaklines=true,
+  showstringspaces=false,
+  breakatwhitespace=true,
+  commentstyle=\color{pgreen},
+  keywordstyle=\color{pblue},
+  stringstyle=\color{pred},
+  basicstyle=\ttfamily,
+  moredelim=[il][\textcolor{pgrey}]{$$},
+  moredelim=[is][\textcolor{pgrey}]{\%\%}{\%\%}
+}
+
 \begin{document}
@@
     {\color{red} $\varheartsuit$} Java
-	 (Android, PlayFramework, LWJGL, MapReduce, GWT, AWS), C, C++ (OpenGL), C\# (XNA, Windows Forms, WPF, lpsolve), VB, Python, Scala, Scheme, HTML, CSS (Bootstrap), JavaScript (jQuery, AJAX, WebGL, WebRTC), PHP (Wordpress), GLSL, HLSL.
+	 (Android, PlayFramework, LWJGL, MapReduce, GWT, AWS, JavaFX, Kurento Media Server, Jetty and Jersey), C, C++ (OpenGL), C\# (XNA, Windows Forms, WPF, lpsolve), VB, Python, Scala, Scheme, HTML, CSS (Bootstrap), JavaScript (jQuery, AJAX, WebGL, WebRTC), PHP (Wordpress), GLSL, HLSL.
    \section{databases}
@@
 \begin{entrylist}
+
   \entry
-    {11 2014\\(ongoing)}
-    {Web Development}
-    {Bullray}
-    {\emph{Development of an architecture for a chat application. Experience with Android, PlayFramework and MongoDB.}}
+    {01 2016\\(ongoing)}
+    {Applications Development}
+    {Bullray-CIT\\1000€/month (part time)}
+    {\emph{Development of a user interface for a bank's security device that is synchronized with servers.
+Microcontroller sensors information receival through \underline{USB} (\underline{org.usb4java}), user interface with \underline{JavaFX} and network camera discovery through nmap (using \underline{org.nmap4j}) } }
+
+ \entry
+    {11 2015\\(2 months)}
+    {Android Development}
+    {Bullray-CIT\\1000€/month (part time)}
+    {\emph{Optimization and bug resolution of an \underline{Android} application that uploads GPS, camera and microphone data in background.}}
+
+\entry
+    {08 2015\\(standby)}
+    {Systems Development}
+    {Bullray-CIT\\1000€/month (part time)}
+    {\emph{Development and configuration of a streaming server that supports receiving and recording video from authenticated users through \underline{ffmpeg}.
+I tried \underline{FFServer}, \underline{nginx with rtmp module} and \underline{Wowza Streaming Server}, the final decision was \underline{Kurento Media Server} with an \underline{RTP} endpoint for receiving video from network cameras and a \underline{WebRTC} endpoint for delivering the video to web browsers.
+}}
+
+\entry
+    {03 2015\\(ongoing)}
+    {Master Thesis}
+    {IST}
+    {\emph{Development of an interactive multi user video chat that supports recording, content overlay and a collaborative text editor.
+Used \underline{Kurento Media Server} for mixing streams into a single one and detecting QR codes.
+Experience with \underline{ot.js} for operation transformations (collaborative component).
+Other libraries used, \underline{jquery.qrcode}, \underline{vis.js}, \underline{bootstrap-typeahead}, \underline{codemirror.js} and \underline{adapter.js} }.
+Little experience with \underline{strophe.js} for communication between web browsers and \underline{XMPP} servers.
+}
+
+  \entry
+    {02 2015\\(ongoing)}
+    {Full-Stack Web Development}
+    {Bullray-CIT\\1000€/month (part time)}
+    {\emph{Development of a solution for a generic incident handling server, support for real time monitoring through \underline{WebSockets}.
+Experience with \underline{PlayFramework}, \underline{MongoDB} (with the official Java Driver), \underline{Maven}, \underline{jQuery} and \underline{Bootstrap}.}}
+
+  \entry
+    {11 2014\\(4 months)}
+    {Full-Stack Web Development}
+    {Bullray-CIT\\1000€/month (part time)}
+    {\emph{Development of an architecture for a chat application.
+Experience with \underline{Android}, \underline{PlayFramework}, \underline{MongoDB} (with Morphia), \underline{jQuery} and \underline{Bootstrap}.
+Message synchronization through HTTP long polling.}}
   \entry
@@
     {Scientific Initiation Scholarship}
-    {INESC}
-    {\emph{Frontend Development for Provide Results's Bimk project. Experience with PlayFramework, Revit API and GIT.}}
+    {INESC\\385€/month (part time)}
+    {\emph{Frontend Development for Provide Results's Bimk project. Experience with \underline{PlayFramework}, \underline{Revit API} and \underline{GIT}.}}
   \entry
@@
     {Scientific Initiation Scholarship}
-    {CEG-IST}
-    {\emph{Programming Decision Analysis Software (MACBETH). Development of an algorithm to suggest judgements based on past decisions using \emph{lpsolve}.}}
+    {CEG-IST\\385€/month (part time)}
+    {\emph{Programming a Decision Analysis Software (MACBETH) with \underline{WPF}. Development of an algorithm to suggest judgements based on past decisions using \underline{lpsolve}.}}
   \entry
@@
     {LCG}
-    {\emph{Frontend development for Decision Analysis Software using \emph{ASP.NET}. }}
+    {\emph{Frontend development for a Decision Analysis Software using \underline{ASP.NET}. }}
   \entry
@@
     {Garbage Collector}
-    {Jovens em Movimento, Oeiras}
+    {Jovens em Movimento, Oeiras\\120€/month(full time)}
     {Cleaning and maintenance of public spaces.}
@@
 \end{entrylist}
+%\newpage
 
@@
     {\href{http://play.google.com/store/apps/details?id=pt.ist.bombraider}{Google Play}}
-    {Bomb Raider is a simple singleplayer and multiplayer game where the goal is destroying all enemies in the map.}
+    {Bomb Raider is a simple singleplayer and multiplayer game where the goal is destroying all enemies in the map.
+I used \underline{OpenGLES} and \underline{WiFi-Direct} to develop this \underline{Android} game.}
 
@@
     {\href{http://www.windowsphone.com/en-us/store/app/infinity-edge/05516463-3a89-4351-996e-62e5b4519aeb}{windows marketplace}}
-    {First Person Shooter with spaceships.}
+    {First Person Shooter with spaceships made on my free time. I had experience with \underline{XNA} and \underline{HLSL}.}
 
@@
     {\href{http://web.ist.utl.pt/ist168621/?page=8}{Personal Project}}
-    {3D environment generator and Real Time Strategy game engine.}
+    {3D environment generator and Real Time Strategy game engine using \underline{OpenGL}.}
 
@@
 \section{interests}
-    3D Studio Max, Photoshop, Game Development, Web Design, Network Security, Jogging and Yoga.
+    3D Studio Max, Photoshop, Game Development, Web Design, Network Security, Jogging, Karate and Yoga.
 
-\newpage
-\section{grades}
-\subsection{Bachelor - Communication Networks Engineering}
-\input{grades_bachelor.tex}
-\subsection{Master - Telecommunications and Informatics Engineering}
-\input{grades_master.tex}
+	My mother said I should have a girlfriend...
 
+%\newpage
+%\section{grades}
+%\subsection{Bachelor - Communication Networks Engineering}
+%\input{grades_bachelor.tex}
+%\subsection{Master - Telecommunications and Informatics Engineering}
+%\input{grades_master.tex}
+
+\begin{lstlisting}
+System.out.println("... but I love programming.");
+return 0; // ends gracefully
+\end{lstlisting}
 \end{document}

### 2016-03-02 project/httpserver/app/main/UserSession.java

@@ -18,8 +18,6 @@ import services.ListMessagesService;
 
 import java.io.Closeable;
 import java.io.IOException;
-import java.security.MessageDigest;
-import java.security.NoSuchAlgorithmException;
 import java.util.*;
 
 
@@ -80,6 +78,9 @@ public class UserSession implements Closeable, Comparable<UserSession> {
      * @param out the out
      * @throws ServiceException the service exception
      */
+    private final EventListener<CodeFoundEvent> codeListener;
+    private ListenerSubscription codeSubscritption;
+
     public UserSession(final User user, final Room room, WebSocket.Out<String> out) throws ServiceException {
         this.out = out;
         this.user = user;
@@ -112,25 +113,33 @@ public class UserSession implements Closeable, Comparable<UserSession> {
         endPoint.setStunServerAddress("64.233.184.127");
         endPoint.setStunServerPort(19302);
 
-        qrCodeFilter = new ZBarFilter.Builder(room.getMediaPipeline()).build();
+        qrCodeFilter = new ZBarFilter.Builder(endPoint.getMediaPipeline()).build();
 
-        qrCodeFilter.addCodeFoundListener(new EventListener<CodeFoundEvent>() {
+
+
+
+       codeListener = new EventListener<CodeFoundEvent>() {
             @Override
             public void onEvent(CodeFoundEvent event) {
+                System.out.println(".");
                 final String content = event.getValue();
                 final String hash = Tools.md5(content);
                 if(hash!=null) {
+                    boolean containsQR = false;
                     synchronized (currentQRCodes) {
-                        if (!currentQRCodes.contains(hash)) {
-                            currentQRCodes.add(hash);
-                            JSONObject msg = new JSONObject();
-                            msg.put("id", "qrCode");
-                            msg.put("data", content);
-                            msg.put("hash", hash);
-                            room.sendMessage(msg.toString());
-                        }
+                        containsQR = currentQRCodes.contains(hash);
                     }
-                    System.out.println("=============================" + event.getCodeType() + "|" + event.getValue());
+                    if (!containsQR) {
+                        synchronized (currentQRCodes) {
+                            currentQRCodes.add(hash);
+                        }
+                        JSONObject msg = new JSONObject();
+                        msg.put("id", "qrCode");
+                        msg.put("data", content);
+                        msg.put("hash", hash);
+                        room.sendMessage(msg.toString());
+                    }
+
                     new Thread(new Runnable() {
                         @Override
[…]

### 2016-03-03 curriculo/cv.tex
[ADD] added sample video for testing

@@
   \entry
+    {201?}
+    {My personal website}
+    {\href{http://play.google.com/store/apps/details?id=pt.ist.bombraider}{henrique.rocha.pt}}
+    {I like PlayFramework, but I still like more to learn different technologies, that's why I did this website using ???, just for fun.}
+
+  \entry
     {2015}
@@
 
-	My mother said I should have a girlfriend...
+%	My mother said I should have a girlfriend...
 
@@
 \begin{lstlisting}
-System.out.println("... but I love programming.");
-return 0; // ends gracefully
+return 0;
 \end{lstlisting}

### 2016-03-03 project/httpserver/app/controllers/Application.java

@@ -2,19 +2,20 @@ package controllers;
 
 import dtos.AttributeDto;
 import exceptions.ServiceException;
-import models.Group;
-import models.Membership;
-import models.Relation;
-import models.User;
+import models.*;
 import org.bson.Document;
 import org.bson.types.ObjectId;
+import org.bson.types.Symbol;
+import play.api.Play;
 import play.mvc.Controller;
 import play.mvc.Result;
 import services.*;
 
 import java.io.File;
+import java.net.URL;
 import java.util.ArrayList;
 import java.util.Arrays;
+import java.util.Date;
 import java.util.List;
 
 
@@ -117,6 +118,24 @@ public class Application extends Controller {
 
         AddGroupMemberService joinService = new AddGroupMemberService(user1.getId().toString(), group0.getId().toString(), user2.getId().toString());
         joinService.execute();
+
+        Date end = new Date();
+        Date start = new Date(end.getTime()-120000);
+
+
+        Interval interval = new CreateIntervalService(group0.getId().toString(), start).execute();
+        interval.setEnd(end);
+
+        interval.save();
+        URL url = this.getClass().getClassLoader().getResource("video.mp4");
+
+        Recording rec = new Recording(group0.getId(),start);
+        rec.setEnd(end);
+        rec.setUrl(group0.getId().toString(),url.getFile());
+        rec.save();
+
+        new CreateTimeTagService(group0.getId().toString(),start,"Vetores (alguns exemplos)").execute();
+
         session().clear();
         return redirect("/");
     }

### 2016-03-03 project/httpserver/app/controllers/WSController.java

@@ -248,8 +248,7 @@ public class WSController extends Controller {
                                     try {
                                         Date time = Tools.FORMAT.parse(args.getString("time"));
                                         String title = args.getString("title");
-                                        String content = args.getString("content");
-                                        CreateTimeTagService service = new CreateTimeTagService(groupId, time, title, content);
+                                        CreateTimeTagService service = new CreateTimeTagService(groupId, time, title);
                                         TimeTag tag = service.execute();
                                         JSONObject msg = new JSONObject();
                                         msg.put("id", "tag");
@@ -281,12 +280,10 @@ public class WSController extends Controller {
                                         CreateHyperContentService service = new CreateHyperContentService(userId,
                                                 groupId, start, end, content);
                                         service.execute();
+                                        room.sendContents();
                                     } catch (ServiceException | ParseException e) {
                                         e.printStackTrace();
                                     }
-                                    // System.out.println("content");
-                                    room.sendContents();
-
                                 }
                                 break;
                                 case "deleteContent": {

### 2016-03-03 project/httpserver/app/models/TimeTag.java

@@ -28,10 +28,7 @@ public class TimeTag {
     
     /** The title. */
     private String title;
-    
-    /** The content. */
-    private String content;
-    
+
     /** The gid. */
     private ObjectId id, gid = null;
 
@@ -48,13 +45,11 @@ public class TimeTag {
      * @param gid the gid
      * @param time the time
      * @param title the title
-     * @param content the content
      */
-    public TimeTag(ObjectId gid, Date time, String title, String content) {
+    public TimeTag(ObjectId gid, Date time, String title) {
         this.time = time;
         this.gid = gid;
         this.title = title;
-        this.content = content;
     }
 
     /**
@@ -80,7 +75,6 @@ public class TimeTag {
         rec.time = doc.getDate("time");
         rec.gid = doc.getObjectId("gid");
         rec.title = doc.getString("title");
-        rec.content = doc.getString("content");
         return rec;
     }
 
@@ -154,7 +148,6 @@ public class TimeTag {
         doc.put("gid", gid);
         doc.put("time", time);
         doc.put("title", title);
-        doc.put("content", content);
 
         if (id == null)
             getCollection().insertOne(doc);
@@ -201,7 +194,6 @@ public class TimeTag {
         JSONObject obj = new JSONObject();
         obj.put("id", id.toString());
         obj.put("title", title);
-        obj.put("content", content);
         obj.put("time", Tools.FORMAT.format(time));
         return obj;
     }

### 2016-03-03 project/httpserver/public/this/js/hyper.socket.js

@@ -142,12 +142,11 @@ var HyperWebSocket = new (function() {
 		}));
 	}
 
-	this.createTag = function(time,title,content){
+	this.createTag = function(time,title){
 		ws.send(JSON.stringify({
 			id : "createTag",
 			time: time,
-			title: title,
-			content: content
+			title: title
 		}));
 	}
 	

### 2016-03-12 tese1/01_abstract.tex
[WR] 1st part corrections

--- 2015-06-07 tese/project.tex
@@
-\documentclass{llncs}
-\RequirePackage[english]{babel} % Se fizerem o texto em ingles
-\usepackage[english]{babel}   
-%\usepackage[latin1]{inputenc}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-
-\usepackage[fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{subfig}  % allows for indexgeneration
-\usepackage{color}  % allows for indexgeneration
-\usepackage{url}
-\usepackage{hyperref}
-\usepackage{float}
-\usepackage{mathtools}
-\usepackage{acronym}
-\usepackage{pgfgantt}
-\usepackage{adjustbox}
-\usepackage{rotating}
-\usepackage{dirtytalk}
-\usepackage[graphicx]{realboxes}
-\definecolor{fade}{rgb}{0.8,0.8,0.8}
-
-\setcounter{secnumdepth}{3}
-\newacro{AES}{Advanced Encryption Standard}
-\newacro{API}{Application Programming Interface}
-\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
-\newacro{CBC}{Cipher Block Chaining}
-\newacro{CPU}{Central Processing Unit}
-\newacro{CRUD}{Create, Read, Update and Delete}
-\newacro{CSS}{Cascading Style Sheets}
-\newacro{DOM}{Document Object Model}
-\newacro{DTLS}{Datagram Transport Layer Security}
-\newacro{DVD}{Digital Versatile Discs}
-\newacro{FMS}{Flash Media Server}
-\newacro{HTML}{HyperText Markup Language}
-\newacro{HTTP}{Hypertext Transfer Protocol}
-\newacro{ICE}{Interactive Connectivity Establishment}
-\newacro{IPv4}{Internet Protocol Version 4}
-\newacro{IPv6}{Internet Protocol Version 6}
-\newacro{IP}{Internet Protocol}
-\newacro{JSON}{JavaScript Object Notation}
-\newacro{MPEG}{Moving Picture Experts Group}
-\newacro{MVC}{Model-View-Controller}
-\newacro{NAT}{Network Address Translation}
-\newacro{OT}{Operational transformation}
-\newacro{P2P}{Peer-to-peer}
-\newacro{PSTN}{Public Switched Telephone Network}
-\newacro{RFC}{Request For Comments}
-\newacro{RTCP}{Real Time Control Protocol}
-\newacro{RTSP}{Real Time Streaming Protocol}
-\newacro{RTMFP}{Real-Time Media Flow Protocol}
-\newacro{RTMP}{Real Time Messaging Protocol}
-\newacro{RTP}{Real-time Transport Protocol}
-\newacro{SAMI}{Synchronized Accessible Media Interchange}
-\newacro{SCTP}{Stream Control Transmission Protocol}
-\newacro{SDP}{Session Description Protocol}
-\newacro{SigOfly}{Signaling-On-the-fly}
-\newacro{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
-\newacro{SIP}{Session Initiation Protocol}
-\newacro{SMIL}{Synchronized Multimedia Integration Language}
-\newacro{SMS}{Short Message Service}
-\newacro{SRTCP}{Secure RTCP}
-\newacro{SRTP}{Secure Real-time Transport Protocol}
-\newacro{SRT}{SubRip Text}
-\newacro{SSL}{Secure Sockets Layer}
-\newacro{STUN}{Session Traversal Utilities for NAT}
-\newacro{SVG}{Scalable Vector Graphics}
-\newacro{TCP}{Transmission Control Protocol}
-\newacro{TLS}{Transport Layer Security}
-\newacro{TMN}{This Means Nothing}
-\newacro{TURN}{Traversal Using Relays around NAT}
-\newacro{UDP}{User Datagram Protocol}
-\newacro{VoIP}{Voice Over IP}
-\newacro{WebRTC}{Web Real-Time Communication}
-\newacro{WebVTT}{Web Video Text Tracks}
-\newacro{XAML}{eXtensible Application Markup Language}
-\newacro{XEP}{XMPP Extensions}
-\newacro{XHTML}{eXtensible Hypertext Markup Language}
-\newacro{XML}{eXtensible Markup Language}
-\newacro{XMPP}{Extensible Messaging and Presence Protocol}
-
-%\overfullrule=2cm
-
-\newganttchartelement*{mymilestone}{
-mymilestone/.style={
-shape=isosceles triangle,
-inner sep=0pt,
-draw=cyan,
-top color=white,
-bottom color=cyan!50
-},
-mymilestone incomplete/.style={
-/pgfgantt/mymilestone,
-draw=yellow,
-bottom color=yellow!50
-},
-mymilestone label font=\slshape,
-mymilestone left shift=0pt,
-mymilestone right shift=0pt
-}
-
-\newgantttimeslotformat{stardate}{%
-\def\decomposestardate##1.##2\relax{%
-\def\stardateyear{##1}\def\stardateday{##2}%
-}%
-\decomposestardate#1\relax%
-\pgfcalendardatetojulian{\stardateyear-01-01}{#2}%
-\advance#2 by-1\relax%
-\advance#2 by\stardateday\relax%
-}
-
-\begin{document}
-\pagestyle{plain}
-\mainmatter              % start of the contributions
-
-\title{Hyper-linked Communications: WebRTC enabled asynchronous collaboration}
-\author{%
-	Henrique Duarte Lopes Rocha \\
-	email: henrique.rocha@tecnico.ulisboa.pt \\
-  student number: 68621
-}
-\institute{Master Degree in Telecommunications and Informatics Engineering \\ Instituto Superior Técnico \\ Universidade de Lisboa}
-%RP acrescenta número de aluno, nome do mestrado e ``cadeira de projecto''
-\maketitle              % typeset the title of the contribution
-
 \begin{abstract}
@@
 The nature of web browser applications already follow the hypermedia concept, which makes \ac{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
-The native support of \ac{WebRTC} in operating systems extends its usage to outside the web browser, allowing to explore functionalities that browsers are poor to support such as video recording and massive information storage.
+{\color{blue}The web browser platform provides an abstraction layer that makes possible to create applications that run indenpendently from the operating system.}
+The native support of \ac{WebRTC} in operating systems extends its usage to outside the web browser, allowing to explore functionalities that web browsers are poor to support such as video recording and massive information storage.
 
-Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed latter.
+Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed later.
 This application will target the web browser platform, resorting to \ac{WebRTC}.
@@
 \end{abstract}
-
-\section{Introduction}
-  \input{intro.tex}
-
-\section{Related Work}\label{related}
- This section is structured as follows.
- Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
- Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
- Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications. 
- Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
- Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
-%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.
-
-  \input{state01.tex}
-  \input{state02.tex}
-  \input{state03.tex}
-  \input{state04.tex}
-
-\section{Proposed Architecture}\label{arch} % English
-  \input{arch01.tex}
-
-\section{Methodology}\label{meth} % Section
-
-		\input{method.tex}
-
-	\subsection{Planned Schedule} % English
-		\input{schd01.tex}
-
-\section{Conclusions}\label{concl} % English
-The development of \ac{WebRTC} by the World Wide Web Consortium (W3C) allows web browser-to-browser audio, video and data communication.
-\ac{WebRTC} is enabling new usage scenarios for communication and collaboration applications.
-This and other technologies allow the Hyper-link concept to be applied to a new media: real-time audio and video.
-
-  This project proposes a solution that uses \ac{WebRTC} for creating an application that provides audio, data, text and data real-time communications.
-Theses communications will be enriched using hypermedia concepts, the possibility to record and playback its content and the integration of collaborative tools.
-
-  A prototype will be implemented in order to validate these concepts.
-In order to implement it, significant challenges, such as the need to record interactive audio and video streams, will have to be overcome, making this a challenging project.
-
-%\subsection{Summary} % English
-%Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as contribuições definidas e a forma como estas se materializaram
-
-%\subsection{Conclusions} % English
-%Nesta secção devem ser retiradas conclusões do trabalho realizado, em face dos resultados obtidos \cite{exemplo}.
-
-\bibliographystyle{splncs03}
-\bibliography{references}
-
-\end{document}

### 2016-03-12 tese1/02_introduction.tex
[WR] 1st part corrections

--- 2015-06-01 tese/intro.tex
@@
+\section{Introduction}
+
 \subsection{Context}   % English
@@
 
-	As \cite{geddes} states, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
-The concept of structured voice and video synchronized with hypermedia is called hypervoice.
+	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice's capacity to conveying rich and complex social and emotional meaning} , although nothing replaces the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
+The concept of structured voice and video synchronized with hypermedia is called hypervoice.{\color{red}\textbf{[cite!]}}
         %RP a citação do geddes faz sentido? Parece estar a falar de voz gerada por computador e não de video conferencia onde o que ouvimos são outras pessoas!
-
-\subsection{Problem Statement} % English
-
-	As communications technologies appeared, we adapted the way we communicate.
-This project don't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
-
-	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see.
-               %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
-        A real-time system is an huge source of information that requires much attention from its users, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
-%RP estes dois ultimos pontos são melhores.
-Não fales em pessoas com problemas de memoria mas em termos de lidar com grandes quantidades de informação, realizarmos comunicações em grupo (ainda nãofalaste diss) e nem sempre ser possível estar presente.
-        Real-time communication applications can make a difference on business, education and health sectors by providing tools for teaching and learning online, teamworking and socializing.
-        %RP também há outros mercados, como online learning, projectos de longa duração que vão evoluindo e onde é preciso revisitar o que aconteceu antes.
-Fins histórios.
-Necessidade de manter registos devido a requisitos legais/registar história.
-
-	This project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
-creating annotations for improving content search) in order to increase its value.
-Another important aspect of this project is the ability to navigate in time and space by rewind communications, fast-forward, jump to certain points and choose another path.
-
-                %RP esta última frase está confusa.
-
-        %Estás muito parco nos objectivos do projecto.
-Acho que falta dizer que é um sistema para complementar/subsitutir actuais sistemas de comunicação audio/video real-time, adicionando-lhe a capacidade de: adicionar conteúdo posteriomente (e.g.
-anotações) por forma a aumentar o seu valor; permitir acesso posterior (mesmo que apenas 5 minutos depois para quem chega atrasado); permitir navegar na comunicação (hypermedia, saltos, fast-forward, rewind).
-
-\subsection{Thesis Goals} % English
-
-A web application with an easy to learn user interface will be developed to accomplish solving our problem.
-Our application, unnamed yet, is targeted at web browsers that are compatible with only standard technologies like JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
-Any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
-
-%RP podes elaborar um pouco.
-Objectivos são: realizar o levantamento do estado da arte no espaço do problema para fundamentar uma solução; apresentar uma solução e desenhar uma arquitectura para dar resposta ao problema apresentado; construir protótipo funcional; desenhar e realizar uma avaliação/testes que provem a adequação e validade dos conceitos desenvolvidos (elaborar); escrever um artigo.
-
-Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solves our problem either completely or just a part of it.
-
-We will present an architecture that can meet our goals, implement the respective prototype and test it with real users, unitary tests and benchmarks.
-
-As \cite{geddes} said, the quality of the interaction worsens as the number of users increase.
-In our testing phase we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.
-
-	All the problems faced during the development and limitations will be reported on the thesis so that a future project better then ours can be easily and better developed.
-        %RP esta última frase não é muito necessária
-
-\subsection{Document Structure} % English
-This document is structured as follows. Section \ref{related} presents the related work.
-%	Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
-%	Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
-%	Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications. 
-%	Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
-%	Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
-%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.   
-	In Section \ref{arch}, we propose an architecture for an Web Application that fulfills the goals of this thesis, including all the need infrastructure and software.
-	Section \ref{meth} describes our work methodology and our plan for implementing and validating our proposed architecture.
-	Section \ref{concl} presents the summary and conclusions of our work.

### 2016-03-12 tese1/03_problem.tex
[WR] 1st part corrections

--- 2015-06-01 tese/intro.tex
@@
-\subsection{Context}   % English
-Since the early days of Human History, we tried to communicate over far locations, from smoke signals to letters delivered by messengers.
-Real-time communications were limited or even nonexistent.
-Despite all the efforts made to improve communications, written communication could never replace face to face communication.
-With the advent of the telephone network, communications have taken a very important step for us to feel more connected with whom we communicate.
-Still, only the human voice was not enough, and the invention of cameras and consequent video digitization were a huge step for real-time communications.
-
-	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronism between writers.
-	Today, we can achieve more, it is possible to write a document collaboratively, correct spelling mistakes without wasting paper, restructure text at any moment, add a video to a newspaper article and more.
-Although much of what was said seems banal nowadays, none of this was possible before the computer's invention.
-
-	As \cite{geddes} states, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
-The concept of structured voice and video synchronized with hypermedia is called hypervoice.
-        %RP a citação do geddes faz sentido? Parece estar a falar de voz gerada por computador e não de video conferencia onde o que ouvimos são outras pessoas!
-
 \subsection{Problem Statement} % English
@@
 	As communications technologies appeared, we adapted the way we communicate.
-This project don't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
+This project doesn't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
+
+	{\color{blue}With the advent of WebRTC, it became possible to develop video conference web apllications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.}
+
+    Real-time communication applications can make a difference on business, education and health sectors by providing tools for teaching and learning online, teamworking and socializing.
 
@@
                %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
-        A real-time system is an huge source of information that requires much attention from its users, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
+        A real-time system is a huge source of information that requires much attention from its users.
+An application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
 %RP estes dois ultimos pontos são melhores.
 Não fales em pessoas com problemas de memoria mas em termos de lidar com grandes quantidades de informação, realizarmos comunicações em grupo (ainda nãofalaste diss) e nem sempre ser possível estar presente.
-        Real-time communication applications can make a difference on business, education and health sectors by providing tools for teaching and learning online, teamworking and socializing.
         %RP também há outros mercados, como online learning, projectos de longa duração que vão evoluindo e onde é preciso revisitar o que aconteceu antes.
@@
 Necessidade de manter registos devido a requisitos legais/registar história.
-
-	This project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
-creating annotations for improving content search) in order to increase its value.
-Another important aspect of this project is the ability to navigate in time and space by rewind communications, fast-forward, jump to certain points and choose another path.
 
@@
 anotações) por forma a aumentar o seu valor; permitir acesso posterior (mesmo que apenas 5 minutos depois para quem chega atrasado); permitir navegar na comunicação (hypermedia, saltos, fast-forward, rewind).
-
-\subsection{Thesis Goals} % English
-
-A web application with an easy to learn user interface will be developed to accomplish solving our problem.
-Our application, unnamed yet, is targeted at web browsers that are compatible with only standard technologies like JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
-Any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
-
-%RP podes elaborar um pouco.
-Objectivos são: realizar o levantamento do estado da arte no espaço do problema para fundamentar uma solução; apresentar uma solução e desenhar uma arquitectura para dar resposta ao problema apresentado; construir protótipo funcional; desenhar e realizar uma avaliação/testes que provem a adequação e validade dos conceitos desenvolvidos (elaborar); escrever um artigo.
-
-Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solves our problem either completely or just a part of it.
-
-We will present an architecture that can meet our goals, implement the respective prototype and test it with real users, unitary tests and benchmarks.
-
-As \cite{geddes} said, the quality of the interaction worsens as the number of users increase.
-In our testing phase we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.
-
-	All the problems faced during the development and limitations will be reported on the thesis so that a future project better then ours can be easily and better developed.
-        %RP esta última frase não é muito necessária
-
-\subsection{Document Structure} % English
-This document is structured as follows. Section \ref{related} presents the related work.
-%	Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
-%	Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
-%	Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications. 
-%	Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
-%	Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
-%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.   
-	In Section \ref{arch}, we propose an architecture for an Web Application that fulfills the goals of this thesis, including all the need infrastructure and software.
-	Section \ref{meth} describes our work methodology and our plan for implementing and validating our proposed architecture.
-	Section \ref{concl} presents the summary and conclusions of our work.

### 2016-03-12 tese1/04_goals.tex
[WR] 1st part corrections

--- 2016-01-10 tese1/intro.tex
@@
-\subsection{Context}   % English
-Since the early days of Human History, we tried to communicate over far locations, from smoke signals to letters delivered by messengers.
-Real-time communications were limited or even nonexistent.
-Despite all the efforts made to improve communications, written communication could never replace face to face communication.
-With the advent of the telephone network, communications have taken a very important step for us to feel more connected with whom we communicate.
-Still, only the human voice was not enough, and the invention of cameras and consequent video digitization were a huge step for real-time communications.
+\subsection{Thesis Goals} % English
 
-	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronism between writers.
-	Today, we can achieve more, it is possible to write a document collaboratively, correct spelling mistakes without wasting paper, restructure text at any moment, add a video to a newspaper article and more.
-Although much of what was said seems banal nowadays, none of this was possible before the computer's invention.
-
-	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice's capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
-The concept of structured voice and video synchronized with hypermedia is called hypervoice.
-        %RP a citação do geddes faz sentido? Parece estar a falar de voz gerada por computador e não de video conferencia onde o que ouvimos são outras pessoas!
-
-\subsection{Problem Statement} % English
-
-	As communications technologies appeared, we adapted the way we communicate.
-This project don't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
-
-	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see.
-               %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
-        A real-time system is an huge source of information that requires much attention from its users, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
-%RP estes dois ultimos pontos são melhores.
-Não fales em pessoas com problemas de memoria mas em termos de lidar com grandes quantidades de informação, realizarmos comunicações em grupo (ainda nãofalaste diss) e nem sempre ser possível estar presente.
-        Real-time communication applications can make a difference on business, education and health sectors by providing tools for teaching and learning online, teamworking and socializing.
-        %RP também há outros mercados, como online learning, projectos de longa duração que vão evoluindo e onde é preciso revisitar o que aconteceu antes.
-Fins histórios.
-Necessidade de manter registos devido a requisitos legais/registar história.
-
-	This project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
+	Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
 creating annotations for improving content search) in order to increase its value.
-Another important aspect of this project is the ability to navigate in time and space by rewind communications, fast-forward, jump to certain points and choose another path.
-
-                %RP esta última frase está confusa.
-
-        %Estás muito parco nos objectivos do projecto.
-Acho que falta dizer que é um sistema para complementar/subsitutir actuais sistemas de comunicação audio/video real-time, adicionando-lhe a capacidade de: adicionar conteúdo posteriomente (e.g.
-anotações) por forma a aumentar o seu valor; permitir acesso posterior (mesmo que apenas 5 minutos depois para quem chega atrasado); permitir navegar na comunicação (hypermedia, saltos, fast-forward, rewind).
-
-\subsection{Thesis Goals} % English
+Another important goal of this project is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.
 
@@
 
-Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solves our problem either completely or just a part of it.
+Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solve our problem either completely or just a part of it, not excluding solutions that may appear during the implementation of our project.
 
@@
         %RP esta última frase não é muito necessária
-
-\subsection{Document Structure} % English
-This document is structured as follows. Section \ref{related} presents the related work.
-%	Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
-%	Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
-%	Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications. 
-%	Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
-%	Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
-%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.   
-	In Section \ref{arch}, we propose an architecture for an Web Application that fulfills the goals of this thesis, including all the need infrastructure and software.
-	Section \ref{meth} describes our work methodology and our plan for implementing and validating our proposed architecture.
-	Section \ref{concl} presents the summary and conclusions of our work.

### 2016-03-12 tese1/05_structure.tex
[WR] 1st part corrections

--- 2015-06-01 tese/intro.tex
@@
-\subsection{Context}   % English
-Since the early days of Human History, we tried to communicate over far locations, from smoke signals to letters delivered by messengers.
-Real-time communications were limited or even nonexistent.
-Despite all the efforts made to improve communications, written communication could never replace face to face communication.
-With the advent of the telephone network, communications have taken a very important step for us to feel more connected with whom we communicate.
-Still, only the human voice was not enough, and the invention of cameras and consequent video digitization were a huge step for real-time communications.
-
-	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronism between writers.
-	Today, we can achieve more, it is possible to write a document collaboratively, correct spelling mistakes without wasting paper, restructure text at any moment, add a video to a newspaper article and more.
-Although much of what was said seems banal nowadays, none of this was possible before the computer's invention.
-
-	As \cite{geddes} states, \say{No computer in our lifetimes will ever rival a human voice’s capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
-The concept of structured voice and video synchronized with hypermedia is called hypervoice.
-        %RP a citação do geddes faz sentido? Parece estar a falar de voz gerada por computador e não de video conferencia onde o que ouvimos são outras pessoas!
-
-\subsection{Problem Statement} % English
-
-	As communications technologies appeared, we adapted the way we communicate.
-This project don't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
-
-	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see.
-               %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
-        A real-time system is an huge source of information that requires much attention from its users, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
-%RP estes dois ultimos pontos são melhores.
-Não fales em pessoas com problemas de memoria mas em termos de lidar com grandes quantidades de informação, realizarmos comunicações em grupo (ainda nãofalaste diss) e nem sempre ser possível estar presente.
-        Real-time communication applications can make a difference on business, education and health sectors by providing tools for teaching and learning online, teamworking and socializing.
-        %RP também há outros mercados, como online learning, projectos de longa duração que vão evoluindo e onde é preciso revisitar o que aconteceu antes.
-Fins histórios.
-Necessidade de manter registos devido a requisitos legais/registar história.
-
-	This project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
-creating annotations for improving content search) in order to increase its value.
-Another important aspect of this project is the ability to navigate in time and space by rewind communications, fast-forward, jump to certain points and choose another path.
-
-                %RP esta última frase está confusa.
-
-        %Estás muito parco nos objectivos do projecto.
-Acho que falta dizer que é um sistema para complementar/subsitutir actuais sistemas de comunicação audio/video real-time, adicionando-lhe a capacidade de: adicionar conteúdo posteriomente (e.g.
-anotações) por forma a aumentar o seu valor; permitir acesso posterior (mesmo que apenas 5 minutos depois para quem chega atrasado); permitir navegar na comunicação (hypermedia, saltos, fast-forward, rewind).
-
-\subsection{Thesis Goals} % English
-
-A web application with an easy to learn user interface will be developed to accomplish solving our problem.
-Our application, unnamed yet, is targeted at web browsers that are compatible with only standard technologies like JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
-Any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
-
-%RP podes elaborar um pouco.
-Objectivos são: realizar o levantamento do estado da arte no espaço do problema para fundamentar uma solução; apresentar uma solução e desenhar uma arquitectura para dar resposta ao problema apresentado; construir protótipo funcional; desenhar e realizar uma avaliação/testes que provem a adequação e validade dos conceitos desenvolvidos (elaborar); escrever um artigo.
-
-Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solves our problem either completely or just a part of it.
-
-We will present an architecture that can meet our goals, implement the respective prototype and test it with real users, unitary tests and benchmarks.
-
-As \cite{geddes} said, the quality of the interaction worsens as the number of users increase.
-In our testing phase we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.
-
-	All the problems faced during the development and limitations will be reported on the thesis so that a future project better then ours can be easily and better developed.
-        %RP esta última frase não é muito necessária
-
 \subsection{Document Structure} % English

### 2016-03-12 tese1/06_related_work.tex
[WR] 1st part corrections

--- 2015-06-02 tese/project.tex
@@
-\documentclass{llncs}
-\RequirePackage[english]{babel} % Se fizerem o texto em ingles
-\usepackage[english]{babel}   
-%\usepackage[latin1]{inputenc}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
-
-\usepackage[fleqn]{amsmath}
-\usepackage{amssymb}
-\usepackage{makeidx}  % allows for indexgeneration
-\usepackage{subfig}  % allows for indexgeneration
-\usepackage{color}  % allows for indexgeneration
-\usepackage{url}
-\usepackage{hyperref}
-\usepackage{float}
-\usepackage{mathtools}
-\usepackage{acronym}
-\usepackage{pgfgantt}
-\usepackage{adjustbox}
-\usepackage{rotating}
-\usepackage{dirtytalk}
-\usepackage[graphicx]{realboxes}
-\definecolor{fade}{rgb}{0.8,0.8,0.8}
-
-\setcounter{secnumdepth}{3}
-\newacro{AES}{Advanced Encryption Standard}
-\newacro{API}{Application Programming Interface}
-\newacro{BOSH}{Bidirectional-streams Over Synchronous HTTP}
-\newacro{CBC}{Cipher Block Chaining}
-\newacro{CPU}{Central Processing Unit}
-\newacro{CRUD}{Create, Read, Update and Delete}
-\newacro{CSS}{Cascading Style Sheets}
-\newacro{DOM}{Document Object Model}
-\newacro{DTLS}{Datagram Transport Layer Security}
-\newacro{DVD}{Digital Versatile Discs}
-\newacro{FMS}{Flash Media Server}
-\newacro{HTML}{HyperText Markup Language}
-\newacro{HTTP}{Hypertext Transfer Protocol}
-\newacro{ICE}{Interactive Connectivity Establishment}
-\newacro{IPv4}{Internet Protocol Version 4}
-\newacro{IPv6}{Internet Protocol Version 6}
-\newacro{IP}{Internet Protocol}
-\newacro{JSON}{JavaScript Object Notation}
-\newacro{MPEG}{Moving Picture Experts Group}
-\newacro{MVC}{Model-View-Controller}
-\newacro{NAT}{Network Address Translation}
-\newacro{OT}{Operational transformation}
-\newacro{P2P}{Peer-to-peer}
-\newacro{PSTN}{Public Switched Telephone Network}
-\newacro{RFC}{Request For Comments}
-\newacro{RTCP}{Real Time Control Protocol}
-\newacro{RTSP}{Real Time Streaming Protocol}
-\newacro{RTMFP}{Real-Time Media Flow Protocol}
-\newacro{RTMP}{Real Time Messaging Protocol}
-\newacro{RTP}{Real-time Transport Protocol}
-\newacro{SAMI}{Synchronized Accessible Media Interchange}
-\newacro{SCTP}{Stream Control Transmission Protocol}
-\newacro{SDP}{Session Description Protocol}
-\newacro{SigOfly}{Signaling-On-the-fly}
-\newacro{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
-\newacro{SIP}{Session Initiation Protocol}
-\newacro{SMIL}{Synchronized Multimedia Integration Language}
-\newacro{SMS}{Short Message Service}
-\newacro{SRTCP}{Secure RTCP}
-\newacro{SRTP}{Secure Real-time Transport Protocol}
-\newacro{SRT}{SubRip Text}
-\newacro{SSL}{Secure Sockets Layer}
-\newacro{STUN}{Session Traversal Utilities for NAT}
-\newacro{SVG}{Scalable Vector Graphics}
-\newacro{TCP}{Transmission Control Protocol}
-\newacro{TLS}{Transport Layer Security}
-\newacro{TMN}{This Means Nothing}
-\newacro{TURN}{Traversal Using Relays around NAT}
-\newacro{UDP}{User Datagram Protocol}
-\newacro{VoIP}{Voice Over IP}
-\newacro{WebRTC}{Web Real-Time Communication}
-\newacro{WebVTT}{Web Video Text Tracks}
-\newacro{XAML}{eXtensible Application Markup Language}
-\newacro{XEP}{XMPP Extensions}
-\newacro{XHTML}{eXtensible Hypertext Markup Language}
-\newacro{XML}{eXtensible Markup Language}
-\newacro{XMPP}{Extensible Messaging and Presence Protocol}
-
-%\overfullrule=2cm
-
-\newganttchartelement*{mymilestone}{
-mymilestone/.style={
-shape=isosceles triangle,
-inner sep=0pt,
-draw=cyan,
-top color=white,
-bottom color=cyan!50
-},
-mymilestone incomplete/.style={
-/pgfgantt/mymilestone,
-draw=yellow,
-bottom color=yellow!50
-},
-mymilestone label font=\slshape,
-mymilestone left shift=0pt,
-mymilestone right shift=0pt
-}
-
-\newgantttimeslotformat{stardate}{%
-\def\decomposestardate##1.##2\relax{%
-\def\stardateyear{##1}\def\stardateday{##2}%
-}%
-\decomposestardate#1\relax%
-\pgfcalendardatetojulian{\stardateyear-01-01}{#2}%
-\advance#2 by-1\relax%
-\advance#2 by\stardateday\relax%
-}
-
-\begin{document}
-\pagestyle{plain}
-\mainmatter              % start of the contributions
-
-\title{Hyper-linked Communications: WebRTC enabled asynchronous collaboration}
-\author{%
-	Henrique Duarte Lopes Rocha \\
-	email: henrique.rocha@tecnico.ulisboa.pt \\
-  student number: 68621
-}
-\institute{Master Degree in Telecommunications and Informatics Engineering \\ Instituto Superior Técnico \\ Universidade de Lisboa}
-%RP acrescenta número de aluno, nome do mestrado e ``cadeira de projecto''
-\maketitle              % typeset the title of the contribution
-
-\begin{abstract}
-The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to communication and collaboration services.
-This paradigm allows to synchronize, structure and navigate communication content, providing a way to integrate social media content and collaborative tools into voice and video calls.
-Voice and image together can express emotions and expose creativity like no other medium can. With hypermedia concepts, we can add more value to voice and video communications.
-
-WebRTC technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
-The nature of web browser applications already follow the hypermedia concept, which makes WebRTC the ideal technology to apply the hyper-linked communications concepts.
-The native support of WebRTC in operating systems extends its usage to outside the web browser, allowing to explore functionalities that browsers are poor to support such as video recording and massive information storage.
-
-Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable colaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed latter.
-This application will target the web browser platform, resorting to WebRTC.
-
-In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose an architecture to implemente an hyper-linked communication application based on WebRTC and provide a plan for its implementation and evaluation.
-
-\keywords{WebRTC, asynchronous, communications, collaboration}
-\end{abstract}
-
-\section{Introduction}
-  \input{intro.tex}
-
 \section{Related Work}\label{related}
@@
 %RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.
-
-  \input{state01.tex}
-  \input{state02.tex}
-  \input{state03.tex}
-  \input{state04.tex}
-
-\section{Proposed Architecture}\label{arch} % English
-  \input{arch01.tex}
-
-\section{Methodology}\label{meth} % Section
-
-		\input{method.tex}
-
-	\subsection{Planned Schedule} % English
-		\input{schd01.tex}
-
-\section{Conclusions}\label{concl} % English
-  The development of WebRTC by the World Wide Web Consortium (W3C) allows web browser-to-browser audio, video and data communication.
-WebRTC is enabling new usage scenarios for communication and collaboration applications.
-However, the nature of such scenarios lead to significant challenges, such as the need to record interactive audio and video streams.
-
-  This project proposes a solution that uses WebRTC for creating an application that provides audio, data, text and data real-time communications.
-Theses communications will be enriched using hypermedia concepts, the possibility to record and playback its content and the integration of collaborative tools.
-
-%\subsection{Summary} % English
-%Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as contribuições definidas e a forma como estas se materializaram
-
-%\subsection{Conclusions} % English
-%Nesta secção devem ser retiradas conclusões do trabalho realizado, em face dos resultados obtidos \cite{exemplo}.
-
-\bibliographystyle{splncs03}
-\bibliography{references}
-
-\end{document}

### 2016-03-12 tese1/07_related_work_nat.tex
[WR] 1st part corrections

--- 2015-06-02 tese/state01.tex
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{figures/nat.png}
+	\includegraphics[width=0.4\textwidth]{figures/nat.png}
 	\caption{Network Address Translation}
@@
 
-Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security, although its current wide usage, \ac{NAT}s are exposing their weaknesses to the application layer.
+Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security, although its current wide usage, \ac{NAT}s are exposing their weaknesses to the application layer, {\color{blue}namely on applications that require direct communications between two private networks.}
 %RP initially? E hoje?
@@
 Any external host can communicate with private hosts through their mapped public address and port.
-This represents the least restrictive type of \ac{NAT} and as we will later, the unique type of \ac{NAT} that enables real time communications from point to point.
+This represents the least restrictive type of \ac{NAT} and, as we will see later, the unique type of \ac{NAT} that enables real time communications from point to point.
 
@@
 
-\emph{Symmetric} NAT maps different ports for each connection, as we will see later, this type of \ac{NAT} represents a problem on real time communications.
+\emph{Symmetric} NAT maps different ports for each connection. As we will see later, this type of \ac{NAT} represents a problem on real time communications.
 
@@
 
-Applications behind a \ac{NAT} are prevented to receive incoming connections from the public network, which forces them to behave as a client of a client-server model.
+{\color{blue}Unless the router that performs as \ac{NAT} has forwarding rules to every desired ports of each user}, applications behind a \ac{NAT} are prevented from receiving incoming connections from the public network, which forces them to behave as a client of a client-server model.
 
@@
 %RP file sharing não precisa de real-time
-\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}, although, none of those can establish direct connections on multiple level \ac{NAT}s.
+\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}. None of those can establish direct connections on multiple level \ac{NAT}s.
 %HR XXX I was here
@@
 When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} came to simplify how audio and video are transmitted through web browsers.
-
-\subsection{Real time communications}\label{rtc}
-
-\ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser based real time communications without installing any additional application or plug-in.
-
-Some operating systems such as \emph{Android}, \emph{iOS}, \emph{Linux}, \emph{OSX} and \emph{Windows} implements native \ac{WebRTC} libraries, extending the usage of \ac{WebRTC} to applications outside the web browser.
-This native support can help to implement applications that record video and audio streams for further playback.
-
-%RP indicar que agora também já temos webrtc fora do browser devido ao seu sucesso e ao facto de reunir tecnologia state of the art?
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.9\textwidth]{figures/webrtc_stack.png}
-	\caption{WebRTC protocol Stack}
-\end{figure}
-
-\ac{WebRTC} defines three main \ac{API}s: MediaStream, PeerConnection and DataChannel.
-
-\begin{itemize}
-  \item \textbf{MediaStream} allows the browser to access the camera, microphone and the device's screen.
-
-  \item \textbf{PeerConnection} acquires connection data and negotiates with peers.
-%RP establishes and negociates a connection with a peer for transmitting real-time video or audio?
-  \item \textbf{DataChannel} provides a channel for exchanging arbitrary data with other peers.
-\end{itemize}
-
-\ac{WebRTC} uses \ac{UDP} for transporting data, which provides lower latencies than \ac{TCP}, but is not reliable and does not assure packet order and integrity.
-\ac{SCTP} and \ac{SRTP} are used for streaming data, providing a mechanism for congestion control and partial reliable delivery over \ac{UDP}.
-All transferred audio, data and video must be encrypted with \ac{DTLS} symmetric keys.
-\ac{DTLS} provides the same security guarantees as \ac{TLS}.
-
-\ac{TLS} doesn't support independent record decryption, for that it requires a reliable transport channel, typically \ac{TCP}.
-The decryption of a record depends on the previous record, which for unreliable transport protocols like \ac{UDP} may represent a problem, either due to packet loss or different reception order.
-
-\ac{DTLS} is similar to \ac{TLS}, but is used on top of \ac{UDP}.
-The main difference is the inclusion of a sequence number per packet that is used for packet re-ordering on reception and protects from duplicated packets.
-If a packet sequence number is less than the expected sequence number the packet is discarded.
-If a packet sequence number is greater than the expected sequence number the packet may be enqueued or discarded.
-By knowing the sequence of messages that are sent and received in \ac{DTLS}, timers are used for packet retransmission avoiding acknowledgment messages.
-%RP TLS?
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.6\textwidth]{figures/basic_protocols.png}
-	\caption{Overview of transport protocols}
-\end{figure}
-
-\ac{WebRTC}'s \emph{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}.
-\ac{DTLS} encapsulation provides confidentiality, authentication and integrity to the transfered data.
-A \emph{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication.
-Each data channel direction can be configured for reliable or unreliable transmission, the same can be done for order delivery and priority.
-which can also be defined for improving the quality of service of a particular stream over the others.
-
-\ac{WebRTC}'s \emph{MediaStream} is built on top of \ac{SRTP}, which requires an external mechanism for key exchange.
-\ac{DTLS} keys are negotiated on handshake in order to achieve a secure connection.
-The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining \ac{SRTP} communications are done through \ac{UDP} without using \ac{DTLS}.
-%RP quais remaining?  A última frase é confusa
-
-\ac{WebRTC} aims to provide a standard platform for real-time audio and video on the Web. It arrives at a time when several proprietary products are well established.
-\emph{Skype}\footnote{\url{http://www.skype.com/}(accessed June 1, 2015).} is an application that allows video, voice, instant messaging and multi-party communication over proprietary protocols, its main strength are the amount of users that are using it nowadays and the ability to perform voice calls to the \ac{PSTN}.
-But compared to \emph{Skype}, \ac{WebRTC} applications don't need to be pre-installed.
-
-\emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}(accessed June 1, 2015).} is another popular video multi-party conference web application. 
-In the past, in order to use \emph{hangouts} on a web browser a plug-in had to be installed, nowadays hangouts is using \ac{WebRTC}.
-\emph{Google Hangouts} supports viewing videos on \emph{youtube} synchronously, drawing collaboratively, creating music and playing multi-player games.
-These applications are implemented with \emph{Adobe Flash}.
-
-\emph{Jitsi Meet}\footnote{\url{http://jitsi.org/Projects/JitsiMeet}(accessed June 1, 2015).} is a \ac{WebRTC} collaborative application that uses \emph{Jitsi Videobridge} for high quality and scalable video conferences and supports shared document editing.
-\emph{Jitsi Meet} allows a great amount of users in the same conversation by identifying the current most active participant users and, by consequence, reducing the video and audio quality for all the other users.
-\emph{Jitsi Videobridge} is a server that enables multi-party video calls.
-
-%RP a descricção do skype, hangouts e jistsi cai um pouco do céu.
-Não há uma transição e não é feita uma análise.
-Apenas são indicados.
-Não há comparação entre eles nem com o que pode ser feito por webrtc.

### 2016-03-12 tese1/08_related_work_rtc.tex
[WR] 1st part corrections

--- 2015-06-02 tese/state01.tex
@@
-\subsection{Early days of the Internet and its remaining flaws}\label{early}
-
-The need to build a global communications network in an age when almost nobody had access to that technology and the number of future users was unpredictable, lead to some protocols not being suitable for the huge growth of users that followed.
-\ac{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \ac{IP} address with a different port.
-%RP o porto não identifica uma máquina (membro da rede privada) mas uma ligação.
-This technique is known as \ac{NAT}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.6\textwidth]{figures/nat.png}
-	\caption{Network Address Translation}
-\end{figure}
-%RP na figura devias indicar que uns tem ips publicos e outros privados. Alternativamente, colocavas metade da nat box dentro da núvem para indicar que esta pertence à Internet e os pc não.
-
-Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security, although its current wide usage, \ac{NAT}s are exposing their weaknesses to the application layer.
-%RP initially? E hoje?
-There are four types of \ac{NAT} implementations\cite{rfc3489}: \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.
-%RP apresentas 5. falas em asymmetric. É um erro?
-
-\emph{Full Cone} \ac{NAT} maps each public \ac{IP} address and port to a private \ac{IP} address and port.
-Any external host can communicate with private hosts through their mapped public address and port.
-This represents the least restrictive type of \ac{NAT} and as we will later, the unique type of \ac{NAT} that enables real time communications from point to point.
-
-\emph{Restricted Cone} \ac{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host.
-With this type of \ac{NAT}, the private client can be contacted from any port of the same external host.
-
-\emph{Port Restricted Cone} \ac{NAT} works in the same way as Restricted Cone \ac{NAT}, but it only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.
-
-\emph{Symmetric} NAT maps different ports for each connection, as we will see later, this type of \ac{NAT} represents a problem on real time communications.
-
-\emph{Non-Symmetric} \ac{NAT}s became the common configurations on the Internet.
-As a direct result, problems started to appear: the amount of ports that \ac{IP} makes available is also small compared to our current needs; worse than that, \ac{NAT} also difficults end-to-end communication, forcing most applications that follow this model to be implemented ineffectively.
-
-Applications behind a \ac{NAT} are prevented to receive incoming connections from the public network, which forces them to behave as a client of a client-server model.
-
-%RP explicar que o nat funciona num cenário em que a máquina atrás do nat é apenas um consumidor de informação, trabalhando em modo cliente do modelo cliente-servidor. Em P2P isto falha.
-
-Applications based on multimedia and peer-to-peer file sharing have been one of the most strained by \ac{NAT}.
-Those kind of applications require real time communication in order to achieve the best performance.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.30\textwidth]{figures/multinat.png}
-	\caption{Multiple level \ac{NAT}s}
-\end{figure}
-
-%RP file sharing não precisa de real-time
-\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}, although, none of those can establish direct connections on multiple level \ac{NAT}s.
-%HR XXX I was here
-%RP explicar o que são multiple level pois não falaste disto antes? São ``nested''?
-
-\ac{STUN} servers are quite simple.
-They receive requests from \ac{NAT}ed clients, with the source address of a request being the public address that \ac{NAT} mapped to the client.
-\ac{STUN} servers will then reply to the client, providing the mapped public address, so it knows its associated public \ac{IP} address and port.
-Symmetric \ac{NAT} changes \ac{IP} port for each different connection, for that reason, when the \ac{STUN} servers reply with the \ac{IP} address and port of their connection, it will be useless for clients to use in other connection connections.
-That is why Symmetric \ac{NAT} represents a problem for peer-to-peer communications.
-%RP os problemas que relatas têm a ver com P2P ou real-time?
-
-\ac{TURN} uses public servers to relay traffic between private endpoints.
-It may use a \ac{P2P} network relay to find the best peer, but after that, the behavior is much like client-server.
-Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \emph{full cone}.
-\ac{ICE} is a technique that uses \ac{STUN} when direct communications are possible and \ac{TURN} while a direct communication isn't possible.
-%RP ainda não explicaste o que é o ICE
-%RP explicar que para usar TURN as aplicações têm de estar cientes disso.
-
-Most of client-server applications aren't affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints.
-Clearly \ac{TURN} requires a more expensive relaying infrastructure and, in most cases, more network usage, leading to a worse quality of service.
-The requirements of real-time video communication makes this kind of model unsuitable.
-%RP quando falas em infraestrutura mais dispendiosa para o video, está a falar da internet ou ainda estamos a falar de usar ICE?
-
-When connection is established, either in a direct or indirect way (via \ac{TURN} servers), \ac{WebRTC} came to simplify how audio and video are transmitted through web browsers.
-
 \subsection{Real time communications}\label{rtc}
@@
 
-Some operating systems such as \emph{Android}, \emph{iOS}, \emph{Linux}, \emph{OSX} and \emph{Windows} implements native \ac{WebRTC} libraries, extending the usage of \ac{WebRTC} to applications outside the web browser.
+Some operating systems such as \emph{Android}, \emph{iOS}, \emph{Linux}, \emph{OSX} and \emph{Windows} implement native \ac{WebRTC} libraries, extending the usage of \ac{WebRTC} to applications outside the web browser.
 This native support can help to implement applications that record video and audio streams for further playback.
@@
 	\centering
-	\includegraphics[width=0.9\textwidth]{figures/webrtc_stack.png}
+	\includegraphics[width=0.8\textwidth]{figures/webrtc_stack.png}
 	\caption{WebRTC protocol Stack}
@@
 
-\ac{TLS} doesn't support independent record decryption, for that it requires a reliable transport channel, typically \ac{TCP}.
-The decryption of a record depends on the previous record, which for unreliable transport protocols like \ac{UDP} may represent a problem, either due to packet loss or different reception order.
+\ac{TLS} doesn't support independent packet decryption\cite{rfc6347}, for that it requires a reliable transport channel, typically \ac{TCP}.
+The decryption of a packet depends on the previous packet, which for unreliable transport protocols like \ac{UDP} may represent a problem, either due to packet loss or different reception order.
 

### 2016-03-12 tese1/10_related_work_hypermedia.tex
[WR] 1st part corrections

--- 2015-06-02 tese/state03.tex
@@
   SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3.
-  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.
+  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}(accessed June 2, 2015).}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.
   But SmillingWeb just implements a subset of \ac{SMIL} 3.0 and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when dealing with \ac{SMIL} changes due to real time communications.
@@
 
-  For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.
+  For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}(accessed June 2, 2015).}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.
 Another example would be a collaborative spreadsheet using \ac{WebRTC}.
@@
 
-  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}} was developed for web-based animations.
+  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}(accessed June 2, 2015).} was developed for web-based animations.
 Introducing video support in 2002, \emph{flash} started to grow after that.
@@
 
-  Like \emph{Flash}, \emph{Microsoft Silverlight}\footnote{\url{http://www.microsoft.com/silverlight/}} is a cross browser plug-in and platform that is used to develop rich \emph{Internet} applications.
+  Like \emph{Flash}, \emph{Microsoft Silverlight}\footnote{\url{http://www.microsoft.com/silverlight/}(accessed June 2, 2015).} is a cross browser plug-in and platform that is used to develop rich \emph{Internet} applications.
 It supports vector graphics, animation and video.

### 2016-03-12 tese1/11_related_work_collaboration.tex
[WR] 1st part corrections

--- 2016-01-10 tese1/state04.tex
@@
 	\emph{ShareJS}\footnote{\url{http://sharejs.org/}(accessed June 2, 2015).} is an \ac{OT} \emph{JavaScript} library, developed by the ex \emph{Google Wave} engineer Joseph Gentle, for collaborative text and \ac{JSON} documents edition in real-time.
+{\color{blue} It uses \emph{ShareDB} \footnote{\url{https://github.com/share/sharedb}(accessed March 10, 2016)} for its backend and data model, which supports simple integration with any database.
+One of \emph{ShareDB}'s integration is over \emph{MongoDB}\footnote{\url{https://github.com/share/sharedb-mongo}(accessed March 10, 2016)}}.
 
@@
 It uses \ac{JSON} messages for \ac{OT} concurrency control but it does not provide storage.
+{\color{blue}\emph{TogetherJS} uses its own servers for the signaling phase and it supports microphone and mouse sharing between users.}
 
@@
 
+	{\color{blue}\emph{OT.js}\footnote{\url{http://operational-transformation.github.io}(accessed March 10, 2016)} is a \emph{JavaScript} library that only implements operation transformations on client side over plain text.
+An implication of implementing just the client side is the extra effort that is necessary to implement content's persistent storage, besides this drawback this library is very flexible because it's not tied to a specific database.}
+
+{\color{green} For ease of development we decided to use the same backend for \ac{OT} since our web server relies on \emph{Play Framework} with Java and \emph{ShareJS} relies on \emph{NodeJS}\footnote{\url{https://nodejs.org/en/(accessed March 10, 2016)}} }
+
+{\color{green} \emph{TogetherJS} requires a simple server (also known as hub) that echoes messages between clients, all the synchronization work is performed by the clients.
+Altough the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}, we would need read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}
+
         %RP falta falar um pouco das ferramentas de colaboração já existentes? Wave, Hangouts, skype, etc? Acho que isto merecia uma secção à parte, mesmo que mais curta.

### 2016-03-12 tese1/XX_implementation_model.tex
[WR] 1st part corrections

The model is the most important component of our solution, a badly designed model can imply serious deficulties when implementing new features that are not part of the plans, sometimes we had to redesign the model in order to support new features.

\subsubsection {Generic model for common operations}

For designing our model we have taken into account generic programming techniques. We observed that operations like searching for an object was quite repeated amongst different types of objects.

Our first decision for our model in order to avoid repeated code, was the isolation of the object's attributes from themselves, so we could apply the search operation to a set of attributes indepently from the object type. To this generic set of properties we call \textbf{Data} and each object of this type has a reference to the owner, which is a unique identification number.

The identication number by itself is not sufficient to identify an object, objects from different types can have the same identification number. In order to solve this problem, when an object is created, its correspondent \textbf{Data} must contain the owner's object type.

When an attribute is created it must be specified if the attribute is searchable, otherwise it could be simple to search for attributes that could reveal sensible information about it. For example if we consider that a user could have a health related attribute, searching by a disease would reveal which users could suffer from a certain disease, the leak of that kind of information could, for instance, change the agreement between users and health insurance companies.

Another important attribute specification is the owner identifiability, which tells us if the attribute identifies the object. This specification let us create abstract authentication services, for example a user can login into our system by providing any attribute that identifies himself, for example the e-mail but others are possible like the username or cellphone number.

Not less important, attributes can specify aggregations of objects. For example, the user role is an aggregator property that within users it allows the identifaction of each one is administrator. This aggregation specification is independent from the object type, so it's possible to search for the administrative role and return users and groups of users that contains that property.

Our model supports defining read and write permissions for each attribute, for example an attribute can be set to public read access or private write access to a set of entities. Implicitly if an entity can write an attribute it can also read it.

\subsubsection{User model}
\subsubsection{Relation model}
\subsubsection{Group model}
\subsubsection{Membership model}
\subsubsection{File storage}

\subsubsection{Message model}
\subsubsection{Time Tag model}
\subsubsection{Hyper Content model}
\subsubsection{Collaborative Content model}
\subsubsection{Recording model}
\subsubsection{Interval model}

### 2016-03-12 tese1/project.tex
[WR] 1st part corrections

--- 2015-06-07 tese/project.tex
@@
 
-\begin{abstract}
-The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to communication and collaboration services.
-This paradigm allows to synchronize, structure and navigate communication content, providing a way to integrate social media content and collaborative tools into voice and video calls.
-Voice and image together can express emotions and expose creativity like no other medium can. With hypermedia concepts, we can add more value to voice and video communications.
-
-\ac{WebRTC} technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
-The nature of web browser applications already follow the hypermedia concept, which makes \ac{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
-The native support of \ac{WebRTC} in operating systems extends its usage to outside the web browser, allowing to explore functionalities that browsers are poor to support such as video recording and massive information storage.
-
-Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed latter.
-This application will target the web browser platform, resorting to \ac{WebRTC}.
-
-In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose an architecture to implement an hyper-linked communication application based on \ac{WebRTC} and provide a plan for its implementation and evaluation.
-
-\keywords{WebRTC, asynchronous, communications, collaboration}
-\end{abstract}
-
-\section{Introduction}
-  \input{intro.tex}
-
-\section{Related Work}\label{related}
- This section is structured as follows.
- Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
- Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
- Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications. 
- Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
- Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
-%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.
-
-  \input{state01.tex}
-  \input{state02.tex}
-  \input{state03.tex}
-  \input{state04.tex}
+\input{01_abstract.tex}
+\input{02_introduction.tex}
+\input{03_problem.tex}
+\input{04_goals.tex}
+\input{05_structure.tex}
+\input{06_related_work.tex}
+\input{07_related_work_nat.tex}
+\input{08_related_work_rtc.tex}
+\input{09_related_work_signaling.tex}
+\input{10_related_work_hypermedia.tex}
+\input{11_related_work_collaboration.tex}
+%\input{XX_implementation_model.tex}
 
@@
 
-%\subsection{Summary} % English
-%Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as contribuições definidas e a forma como estas se materializaram
-
-%\subsection{Conclusions} % English
-%Nesta secção devem ser retiradas conclusões do trabalho realizado, em face dos resultados obtidos \cite{exemplo}.
-
 \bibliographystyle{splncs03}

### 2016-03-12 project/httpserver/app/controllers/Application.java

@@ -1,17 +1,15 @@
 package controllers;
 
 import dtos.AttributeDto;
+import dtos.PermissionDto;
 import exceptions.ServiceException;
 import models.*;
 import org.bson.Document;
 import org.bson.types.ObjectId;
-import org.bson.types.Symbol;
-import play.api.Play;
 import play.mvc.Controller;
 import play.mvc.Result;
 import services.*;
 
-import java.io.File;
 import java.net.URL;
 import java.util.ArrayList;
 import java.util.Arrays;
@@ -57,34 +55,34 @@ public class Application extends Controller {
         User user1, user2, user3, user4;
         {
             List<AttributeDto> attributes = new ArrayList<AttributeDto>();
-            attributes.add(new AttributeDto("email", "hdlopesrocha", AttributeDto.Access.READ, AttributeDto.Visibility.PUBLIC, true, true, false));
-            attributes.add(new AttributeDto("name", "Henrique Rocha", AttributeDto.Access.READ, AttributeDto.Visibility.PUBLIC, false, true, false));
-            attributes.add(new AttributeDto("photo", "/assets/images/user1.jpeg", AttributeDto.Access.READ, AttributeDto.Visibility.PUBLIC, false, false, false));
-            user1 = new RegisterUserService("qazokm", attributes).execute();
+            attributes.add(new AttributeDto("email", "hdlopesrocha", true, true, false));
+            attributes.add(new AttributeDto("name", "Henrique Rocha", false, true, false));
+            attributes.add(new AttributeDto("photo", "/assets/images/user1.jpeg", false, false, false));
+            user1 = new RegisterUserService("qazokm",new ArrayList<PermissionDto>() ,attributes).execute();
         }
 
         {
             List<AttributeDto> attributes = new ArrayList<AttributeDto>();
-            attributes.add(new AttributeDto("email", "nbhatt", AttributeDto.Access.READ, AttributeDto.Visibility.PUBLIC, true, true, false));
-            attributes.add(new AttributeDto("name", "Nikhil Bhatt", AttributeDto.Access.READ, AttributeDto.Visibility.PUBLIC, false, true, false));
-            attributes.add(new AttributeDto("photo", "/assets/images/user2.jpeg", AttributeDto.Access.READ, AttributeDto.Visibility.PUBLIC, false, false, false));
-            user2 = new RegisterUserService("qazokm", attributes).execute();
+            attributes.add(new AttributeDto("email", "nbhatt", true, true, false));
[…]

### 2016-03-12 project/httpserver/app/dtos/PermissionDto.java

@@ -1,0 +1,46 @@
+package dtos;
+
+import java.util.List;
+import java.util.Set;
+
+/**
+ * Created by hdlopesrocha on 07-12-2015.
+ */
+public class PermissionDto {
+	
+	/** The key. */
+	private String key;
+	private Set<String> readSet;	// empty means all can read
+	private Set<String> writeSet;	// empty means no one can write
+	
+	/**
+	 * Instantiates a new attribute dto.
+	 *
+	 * @param key
+	 *            the key
+	 */
+	public PermissionDto(String key, Set<String> readSet,Set<String> writeSet) {
+		this.key = key;
+		this.readSet = readSet;
+		this.writeSet = writeSet;
+	}
+
+	/**
+	 * Gets the key.
+	 *
+	 * @return the key
+	 */
+	public String getKey() {
+		return key;
+	}
+
+	public Set<String> getReadSet() {
+		return readSet;
+	}
+
+	public Set<String> getWriteSet() {
+		return writeSet;
+	}
+
+
+}

### 2016-03-12 project/httpserver/app/main/KurentoManager.java

@@ -7,7 +7,6 @@ import org.kurento.client.MediaPipeline;
 import org.kurento.repository.RepositoryClient;
 import org.kurento.repository.RepositoryClientProvider;
 
-
 import java.util.Collections;
 import java.util.Map;
 import java.util.concurrent.ConcurrentHashMap;

### 2016-03-12 project/httpserver/app/main/Room.java

@@ -17,8 +17,6 @@ import javax.annotation.PreDestroy;
 import java.io.Closeable;
 import java.io.IOException;
 import java.util.*;
-import java.util.concurrent.ConcurrentHashMap;
-import java.util.concurrent.ConcurrentMap;
 
 
 /**
@@ -44,6 +42,12 @@ public class Room implements Closeable {
     
     /** The hub port. */
     private HubPort hubPort;
+
+    public Object getOperationLock() {
+        return operationLock;
+    }
+
+    private Object operationLock = new Object();
 
     /**
      * Instantiates a new room.

### 2016-03-12 project/httpserver/app/main/Tools.java

@@ -13,7 +13,6 @@ import java.time.ZonedDateTime;
 import java.util.Date;
 import java.util.Random;
 import java.util.TimeZone;
-import java.util.regex.Matcher;
 import java.util.regex.Pattern;
 
 

### 2016-03-12 project/httpserver/app/main/UserSession.java

@@ -473,20 +473,14 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                             msg.put("id", "setTime");
                             msg.put("time", Tools.FORMAT.format(rec.getStart()));
                             timeOffset = new Date().getTime()- rec.getStart().getTime();
-
                             sendMessage(msg.toString());
-
-
-
                         }
                     }
                 }
             } else {
                 System.out.println("No video here!");
             }
-
         } catch (ServiceException e) {
-            // TODO Auto-generated catch block
             e.printStackTrace();
         }
 

### 2016-03-12 project/httpserver/app/models/Permission.java

@@ -4,22 +4,45 @@ import com.mongodb.client.FindIterable;
 import com.mongodb.client.MongoCollection;
 import com.mongodb.client.MongoCursor;
 import dtos.AttributeDto;
+import dtos.PermissionDto;
 import org.bson.Document;
 import org.bson.types.ObjectId;
 import services.Service;
 
-import java.util.ArrayList;
-import java.util.List;
+import java.util.*;
 
 
 /**
- * Created by hdlopesrocha on 06-01-2016.
+ * The Class Permission.
  */
 public class Permission {
 
-    /** The collection. */
-    private static MongoCollection<Document> collection;
-    
+
+    public static class Entry {
+        private Set<ObjectId> readSet;	// empty means all can read
+        private Set<ObjectId> writeSet;	// empty means no one can write
+
+        public Entry(List<ObjectId> readSet, List<ObjectId> writeSet) {
+            this.readSet = new HashSet<>();
+            this.writeSet = new HashSet<>();
+            this.readSet.addAll(readSet);
+            this.writeSet.addAll(writeSet);
+        }
+
+        public Entry(Set<ObjectId> readSet, Set<ObjectId> writeSet) {
+            this.readSet = readSet;
+            this.writeSet = writeSet;
+        }
+
+        public Set<ObjectId> getReadSet() {
+            return readSet;
+        }
+
+        public Set<ObjectId> getWriteSet() {
+            return writeSet;
+        }
+    }
+
     /** The id. */
     private ObjectId id = null;
     
@@ -27,33 +50,53 @@ public class Permission {
     private ObjectId owner = null;
     
     /** The data. */
-    private Document data = null;
+    private Map<String, Entry> permissions = null;
+
+
 
     /**
      * Instantiates a new permission.
      *
      * @param owner the owner
+     * @param permissions the permissions
      * @param attributes the attributes
      */
-    public Permission(ObjectId owner, List<AttributeDto> attributes) {
-        this.data = new Document();
+    public Permission(ObjectId owner, Map<String,Entry> permissions, List<AttributeDto> attributes) {
+		List<String> invalidPermisssions = new ArrayList<String>();
+        this.permissions = new TreeMap<String, Entry>();
 
-        for (AttributeDto attr : attributes) {
-            if (attr.getVisibility().equals(AttributeDto.Visibility.PRIVATE)) {
-                List<ObjectId> array = new ArrayList<ObjectId>();
-                array.add(owner);
-                data.append(attr.getKey(), array);
-            }
-        }
+		
[…]

### 2016-03-12 project/httpserver/app/services/CreateGroupService.java

@@ -1,11 +1,12 @@
 package services;
 
 import dtos.AttributeDto;
+import dtos.PermissionDto;
 import exceptions.ServiceException;
 import models.*;
 import org.bson.types.ObjectId;
 
-import java.util.List;
+import java.util.*;
 
 
 /**
@@ -22,6 +23,7 @@ public class CreateGroupService extends Service<Group> {
     /** The visibility. */
     private final Group.Visibility visibility;
 
+    private final List<PermissionDto> permissions;
     /**
      * Instantiates a new creates the group service.
      *
@@ -29,10 +31,11 @@ public class CreateGroupService extends Service<Group> {
      * @param visibility the visibility
      * @param attributes the attributes
      */
-    public CreateGroupService(String uid, Group.Visibility visibility, List<AttributeDto> attributes) {
+    public CreateGroupService(String uid, Group.Visibility visibility, List<PermissionDto> permissions, List<AttributeDto> attributes) {
         this.caller = new ObjectId(uid);
         this.attributes = attributes;
         this.visibility = visibility;
+        this.permissions = permissions;
     }
 
     /*
@@ -46,10 +49,26 @@ public class CreateGroupService extends Service<Group> {
         group.save();
         Membership membership = new Membership(caller, group.getId());
         membership.save();
-        attributes.add(new AttributeDto("type", Group.class.getName(), AttributeDto.Access.READ, AttributeDto.Visibility.PUBLIC, false, false, true));
+        attributes.add(new AttributeDto("type", Group.class.getName(), false, false, true));
+        permissions.add(new PermissionDto("type", new HashSet<String>(),new HashSet<String>()));
         new Data(group.getId(), attributes).save();
         new Search(group.getId(), attributes).save();
-        new Permission(group.getId(), attributes).save();
+
+        Map<String,Permission.Entry> realPermissions = new TreeMap<>();
+        for(PermissionDto p : permissions){
+            Set<ObjectId> readSet = new HashSet<>();
+            Set<ObjectId> writeSet = new HashSet<>();
+            for(String str : p.getReadSet()){
+                readSet.add(new ObjectId(str));
+            }
+            for(String str : p.getWriteSet()){
+                writeSet.add(new ObjectId(str));
+            }
+            realPermissions.put(p.getKey(), new Permission.Entry(readSet,writeSet));
+
+        }
+
+        new Permission(group.getId(),realPermissions, attributes).save();
         return group;
     }
 

### 2016-03-12 project/httpserver/app/services/ListOwnerAttributesService.java

@@ -1,6 +1,7 @@
 package services;
 
 
+import dtos.PermissionDto;
 import exceptions.ServiceException;
 import models.Data;
 import models.Permission;
@@ -8,7 +9,8 @@ import org.bson.Document;
 import org.bson.types.ObjectId;
 
 import java.util.List;
-
+import java.util.Map;
+import java.util.Set;
 
 
 /**
@@ -56,11 +58,11 @@ public class ListOwnerAttributesService extends Service<Document> {
             doc = (Document) doc.get("data");
 
             if (permission != null) {
-                Document dPerm = permission.getData();
-                for (String key : dPerm.keySet()) {
-                    List<ObjectId> aPerms = (List<ObjectId>) dPerm.get(key);
+                Map<String, Permission.Entry> dPerm = permission.getData();
+                for (Map.Entry<String, Permission.Entry> perm : dPerm.entrySet()) {
+                    Set<ObjectId> aPerms = perm.getValue().getReadSet();
                     if (!aPerms.contains(caller)) {
-                        doc.remove(key);
+                        doc.remove(perm.getKey());
                     }
                 }
 

### 2016-03-12 project/httpserver/app/services/RegisterUserService.java

@@ -2,14 +2,15 @@ package services;
 
 
 import dtos.AttributeDto;
+import dtos.PermissionDto;
 import exceptions.ServiceException;
 import models.Data;
 import models.Permission;
 import models.Search;
 import models.User;
+import org.bson.types.ObjectId;
 
-import java.util.List;
-
+import java.util.*;
 
 
 /**
@@ -22,16 +23,17 @@ public class RegisterUserService extends Service<User> {
     
     /** The attributes. */
     private List<AttributeDto> attributes;
-
+private List<PermissionDto> permissions;
     /**
      * Instantiates a new register user service.
      *
      * @param password the password
      * @param attributes the attributes
      */
-    public RegisterUserService(final String password, List<AttributeDto> attributes) {
+    public RegisterUserService(final String password, List<PermissionDto> permissions, List<AttributeDto> attributes) {
         this.password = password;
         this.attributes = attributes;
+        this.permissions = permissions;
     }
 
 
@@ -42,12 +44,28 @@ public class RegisterUserService extends Service<User> {
      */
     @Override
     public synchronized User dispatch() throws ServiceException {
-        attributes.add(new AttributeDto("type", User.class.getName(), AttributeDto.Access.READ, AttributeDto.Visibility.PUBLIC, false, false, true));
-
+        attributes.add(new AttributeDto("type", User.class.getName(), false, false, true));
+        permissions.add(new PermissionDto("type",new HashSet<String>(),new HashSet<String>()));
         User user = new User(password);
         user.save();
         new Search(user.getId(), attributes).save();
-        new Permission(user.getId(), attributes).save();
+
+
+        Map<String,Permission.Entry> realPermissions = new TreeMap<>();
+        for(PermissionDto p : permissions){
+            Set<ObjectId> readSet = new HashSet<>();
+            Set<ObjectId> writeSet = new HashSet<>();
+            for(String str : p.getReadSet()){
+                readSet.add(new ObjectId(str));
+            }
+            for(String str : p.getWriteSet()){
+                writeSet.add(new ObjectId(str));
+            }
+            realPermissions.put(p.getKey(), new Permission.Entry(readSet,writeSet));
+
+        }
+
+        new Permission(user.getId(),realPermissions, attributes).save();
         new Data(user.getId(), attributes).save();
 
         return user;

### 2016-03-13 tese1/XX_arch.tex
[ADD] model, minor code refactoring

--- 2016-01-10 draft/draft.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Arsclassica Article
-% LaTeX Template
-% Version 1.1 (10/6/14)
-%
-% This template has been downloaded from:
-% http://www.LaTeXTemplates.com
-%
-% Original author:
-% Lorenzo Pantieri (http://www.lorenzopantieri.net) with extensive modifications by:
-% Vel (vel@latextemplates.com)
-%
-% License:
-% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%----------------------------------------------------------------------------------------
-%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
-
-\documentclass[
-10pt, % Main document font size
-a4paper, % Paper type, use 'letterpaper' for US Letter paper
-oneside, % One page layout (no page indentation)
-]{scrartcl}
-
-\hyphenation{Fortran hy-phen-ation} % Specify custom hyphenation points in words with dashes where you would like hyphenation to occur, or alternatively, don't put any dashes in a word to stop hyphenation altogether
-
-\title{Draft for master thesis} % The article title
-
-\author{Henrique Rocha} % The article author(s) - author affiliations need to be specified in the AUTHOR AFFILIATIONS block
-
-\date{} % An optional date to appear under the author(s)
-
-%----------------------------------------------------------------------------------------
-
-\begin{document}
-
-\maketitle % Print the title/author/date block
-
-%----------------------------------------------------------------------------------------
-%	INTRODUCTION
-%----------------------------------------------------------------------------------------
+---------------------------------------------------------------
 
@@
 
-An alternative solution is known as Session Attachment (see https://metajack.im/2008/10/03/getting-attached-to-strophe/), which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} is passed to strophe.js in order to re-connect to the same stream on XMPP server.
+An alternative solution is known as Session Attachment \footnote{\url{https://metajack.im/2008/10/03/getting-attached-to-strophe/}}, which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} is passed to strophe.js in order to re-connect to the same stream on XMPP server.
 Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request which is worthless if a user mantains more than one tab opened, for example for multiple conversations at the same time.
@@
 This logic on the XMPP approach would be distributed among XMPP servers and not the webservers, so now our webservers will increase their complexity by implementing the features we needed from XMPP.
-
-\subsubsection{Users}
-
-\subsubsection{Groups}
-
-\subsubsection{Permissions}
-
-\subsubsection{Signaling}
-
-\end{document}

### 2016-03-13 tese1/XX_implementation_model.tex
[ADD] model, minor code refactoring

@@
+\subsection{Model}
 The model is the most important component of our solution, a badly designed model can imply serious deficulties when implementing new features that are not part of the plans, sometimes we had to redesign the model in order to support new features.
 
-\subsubsection {Generic model for common operations}
+\subsubsection {Generic model}
 
@@
 Our first decision for our model in order to avoid repeated code, was the isolation of the object's attributes from themselves, so we could apply the search operation to a set of attributes indepently from the object type.
-To this generic set of properties we call \textbf{Data} and each object of this type has a reference to the owner, which is a unique identification number.
+To this generic set of properties we call data and each object of this type has a reference to the owner, which is a unique identification number.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_data}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.35\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_data_permission}
+    \end{subfigure}
+    \caption{Generic data model}
+\end{figure}
 
 The identication number by itself is not sufficient to identify an object, objects from different types can have the same identification number.
-In order to solve this problem, when an object is created, its correspondent \textbf{Data} must contain the owner's object type.
+In order to solve this problem, when an object is created, its correspondent data must contain the owner's object type.
 
@@
 \subsubsection{User model}
+
+The user model is not tied to the user attributes, the information maintained in this model is just used for authentication purposes.
+Passwords are not stored in plaintext, instead we apply hashing and salting techniques \cite{password} in order to make it harder to decode the password by an attacker.
+We use \emph{SHA-1} and a random salt per user with 32 characters long.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[b]{0.25\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_user}
+    \end{subfigure}
+    \caption{User model}
+\end{figure}
+
 \subsubsection{Relation model}
+
+A relation between two entities $e_1$ and $e_2$ is represented by the pair $e_1\rightarrow e_2$, where $e_1$ is the source and $e_2$ is the target.
+This relation is said bi-directional if and only if it also exists the relation $e_2\rightarrow e_1$.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_relation}
+    \end{subfigure}
+    \caption{Relation model}
+\end{figure}
+
+A user can only interact with users that maintain a relationship with him or with users within the same group.
+In order to validate a relationship, both users must agree on that relationship, by other words it must exist a bi-directional relation between both users.
+
 \subsubsection{Group model}
-\subsubsection{Membership model}
-\subsubsection{File storage}
+
+A group can be public or private.
+If the group is public then it is visible to all users that mantain a relationship with a member of this group.
+If the group is private then it is only visible to their members.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_group}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_group_membership}
+    \end{subfigure}
+    \caption{Group model}
+\end{figure}
+
+The group membership is a special case of relation, where the target entity is always a group.
+
+When a group is created, a group membership is automatically assigned by its creator.
+
+Entities that have a membership with a group can create more memberships by sharing an invite token or by specifying new group members.
 
 \subsubsection{Message model}
-\subsubsection{Time Tag model}
-\subsubsection{Hyper Content model}
+
+A message is composed by its content, sequence number, time of creation and source and target identification numbers.
+The message's target can reference any of object but our application is only covering messages to groups.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_message}
+    \end{subfigure}
+    \caption{Message model}
+\end{figure}
+
+\subsubsection{Hyper content model}
+
+During a group conversation it is possible to tag points in time for making it easy to access that time either by searching or sharing with other users.
+
+A time tag contains a title, the correspondent group identification number and the time itself.
+
+The hyper content is used to synchronize content among users during a conversation.
+Every hyper content must have a start and ending time, the correspondent group identification number and the content itself in the form of text.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_hyper_content}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_time_tag}
+    \end{subfigure}    
+    \caption{Hyper content model}
+\end{figure}
+
 \subsubsection{Collaborative Content model}
+
+Within a conversation, users can write documents collaboratively. Each document has a content and a reference to the correspondent group.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_collaborative_content}
+    \end{subfigure}
+    \caption{Collaborative content model}
+\end{figure}
+
 \subsubsection{Recording model}
-\subsubsection{Interval model}
+
+During a conversation, users may allow sharing their web cameras, by doing so their video is stored in recording chunks.
+Each chunk represents an interval of time  $T=[ \,c^{start},c^{end}[ \,$, it contains a reference to a group and a mapping between entity that was recorded and the correspondent video url.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_recording_chunk}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.3\textwidth}
+    	\includegraphics[width=\textwidth]{figures/model_recording_interval}
+    \end{subfigure}
+    \caption{Recording model}
+\end{figure}
+
+A set of chunks $S=[c_1,c_2,\ldots,c_n]$ is said continuous if $\forall c_i\in S, \exists c_j \in S$ where $j\neq i$ and $c_i^{start} = c_j^{end} \vee c_i^{end} = c_j^{start}$.
+
+A recording interval represents a continuous set of recording chunks.

### 2016-03-13 project/httpserver/app/controllers/Application.java

@@ -121,7 +121,7 @@ public class Application extends Controller {
         Date start = new Date(end.getTime()-835000);
 
 
-        Interval interval = new CreateIntervalService(group0.getId().toString(), start).execute();
+        RecordingInterval interval = new CreateIntervalService(group0.getId().toString(), start).execute();
         interval.setEnd(end);
         interval.save();
 
@@ -130,7 +130,7 @@ public class Application extends Controller {
             String formatted = String.format("%02d", i);
             URL url = this.getClass().getClassLoader().getResource("video"+formatted+".mp4");
             Date da = new Date(start.getTime()+i*10000);
-            Recording rec = new Recording(group0.getId(), da);
+            RecordingChunk rec = new RecordingChunk(group0.getId(), da);
             rec.setEnd(new Date(da.getTime()+10000));
             rec.setUrl(group0.getId().toString(), url.getFile());
             rec.save();
@@ -207,13 +207,13 @@ public class Application extends Controller {
             GetGroupInviteService service = new GetGroupInviteService(session("uid"), groupId);
             String token = service.execute();
             if (group != null && user != null) {
-                Membership membership = Membership.findByUserGroup(user.getId(), group.getId());
+                GroupMembership membership = GroupMembership.findByUserGroup(user.getId(), group.getId());
                 boolean isPublic = false;
 
                 if (group.getVisibility().equals(Group.Visibility.PUBLIC)) {
                     isPublic = true;
                     if (membership == null) {
-                        new Membership(user.getId(), group.getId()).save();
+                        new GroupMembership(user.getId(), group.getId()).save();
                     }
                 }
 

### 2016-03-13 project/httpserver/app/main/Room.java

@@ -35,7 +35,7 @@ public class Room implements Closeable {
     private final Group group;
     
     /** The interval. */
-    private Interval interval;
+    private RecordingInterval interval;
     
     /** The composite. */
     private Hub composite = null;
@@ -128,7 +128,7 @@ public class Room implements Closeable {
                     start = new Date();
                 }
 
-                Recording rec = new Recording(group.getId(),start);
+                RecordingChunk rec = new RecordingChunk(group.getId(),start);
                 new Recorder(hubPort, duration, new Recorder.RecorderHandler() {
 
                     @Override
@@ -247,7 +247,7 @@ public class Room implements Closeable {
                         new JSONArray().put(myProfile.put("online", true)));
 
                 ListGroupMembersService service = new ListGroupMembersService(userId, getGroupId());
-                for (KeyValuePair<Membership, User> m : service.execute()) {
+                for (KeyValuePair<GroupMembership, User> m : service.execute()) {
                     UserSession otherSession = null;
                     for (UserSession os : participants) {
                         if (os.getUser().getId().equals(m.getValue().getId())) {

### 2016-03-13 project/httpserver/app/models/User.java

@@ -25,6 +25,9 @@ public class User implements Comparable<User> {
     /** The id. */
     private ObjectId id = null;
 
+    private static MessageDigest messageDigest = null;
+
+
     /**
      * Instantiates a new user.
      */
@@ -97,6 +100,13 @@ public class User implements Comparable<User> {
         return null;
     }
 
+    private static MessageDigest getMessageDigest() throws NoSuchAlgorithmException {
+        if(messageDigest==null){
+            messageDigest = MessageDigest.getInstance("SHA1");
+        }
+        return messageDigest;
+    }
+
     /**
      * Sha1.
      *
@@ -104,10 +114,10 @@ public class User implements Comparable<User> {
      * @return the string
      * @throws NoSuchAlgorithmException the no such algorithm exception
      */
-    private static String sha1(final String input)
+    private static String sha1(final byte [] input)
             throws NoSuchAlgorithmException {
-        final MessageDigest mDigest = MessageDigest.getInstance("SHA1");
-        final byte[] result = mDigest.digest(input.getBytes());
+        final MessageDigest mDigest = getMessageDigest();
+        final byte[] result = mDigest.digest(input);
         final StringBuffer sb = new StringBuffer();
         for (int i = 0; i < result.length; i++) {
             sb.append(Integer.toString((result[i] & 0xff) + 0x100, 16)
@@ -126,7 +136,7 @@ public class User implements Comparable<User> {
      */
     private static String getHash(final String password, final String salt) {
         try {
-            return sha1(salt + password);
+            return sha1((salt + password).getBytes());
         } catch (NoSuchAlgorithmException e) {
             e.printStackTrace();
         }

### 2016-03-13 project/httpserver/app/services/GetCurrentRecordingService.java

@@ -2,7 +2,7 @@ package services;
 
 import com.mongodb.client.FindIterable;
 import exceptions.BadRequestException;
-import models.Recording;
+import models.RecordingChunk;
 import models.User;
 import org.bson.Document;
 import org.bson.types.ObjectId;
@@ -14,7 +14,7 @@ import java.util.Date;
 /**
  * The Class GetCurrentRecordingService.
  */
-public class GetCurrentRecordingService extends Service<Recording> {
+public class GetCurrentRecordingService extends Service<RecordingChunk> {
 
     /** The caller. */
     private final User caller;
@@ -45,13 +45,13 @@ public class GetCurrentRecordingService extends Service<Recording> {
      * @see services.Service#dispatch()
      */
     @Override
-    public Recording dispatch() throws BadRequestException {
+    public RecordingChunk dispatch() throws BadRequestException {
 
-        FindIterable<Document> iter = Recording.getCollection().find(new Document("gid", groupId).append("end", new Document("$gte", time)).append("start", new Document("$lt", time)));
+        FindIterable<Document> iter = RecordingChunk.getCollection().find(new Document("gid", groupId).append("end", new Document("$gte", time)).append("start", new Document("$lt", time)));
 
         Document first = iter.first();
 
-        return first != null ? Recording.load(first) : null;
+        return first != null ? RecordingChunk.load(first) : null;
     }
 
     /*

### 2016-03-13 project/httpserver/app/services/JoinGroupInviteService.java

@@ -2,7 +2,7 @@ package services;
 
 import exceptions.ServiceException;
 import models.Group;
-import models.Membership;
+import models.GroupMembership;
 import org.bson.types.ObjectId;
 
 
@@ -47,8 +47,8 @@ public class JoinGroupInviteService extends Service<Boolean> {
         if (group != null) {
             ans = group.matchInvite(token);
             if (ans) {
-                if (Membership.findByUserGroup(caller, groupId) == null) {
-                    new Membership(caller, groupId).save();
+                if (GroupMembership.findByUserGroup(caller, groupId) == null) {
+                    new GroupMembership(caller, groupId).save();
                 }
             }
         }

### 2016-03-13 project/httpserver/app/services/ListOwnerAttributesService.java

@@ -1,10 +1,9 @@
 package services;
 
 
-import dtos.PermissionDto;
 import exceptions.ServiceException;
 import models.Data;
-import models.Permission;
+import models.DataPermission;
 import org.bson.Document;
 import org.bson.types.ObjectId;
 
@@ -51,15 +50,15 @@ public class ListOwnerAttributesService extends Service<Document> {
      */
     @Override
     public Document dispatch() throws ServiceException {
-        Permission permission = Permission.findByOwner(user);
+        DataPermission permission = DataPermission.findByOwner(user);
         Document doc = (Document) Data.findByOwner(user, projection);
 
         if (doc != null) {
             doc = (Document) doc.get("data");
 
             if (permission != null) {
-                Map<String, Permission.Entry> dPerm = permission.getData();
-                for (Map.Entry<String, Permission.Entry> perm : dPerm.entrySet()) {
+                Map<String, DataPermission.Entry> dPerm = permission.getData();
+                for (Map.Entry<String, DataPermission.Entry> perm : dPerm.entrySet()) {
                     Set<ObjectId> aPerms = perm.getValue().getReadSet();
                     if (!aPerms.contains(caller)) {
                         doc.remove(perm.getKey());

### 2016-03-13 project/httpserver/app/services/RemoveGroupMemberService.java

@@ -1,7 +1,7 @@
 package services;
 
 import models.Group;
-import models.Membership;
+import models.GroupMembership;
 import models.User;
 import org.bson.types.ObjectId;
 
@@ -41,11 +41,11 @@ public class RemoveGroupMemberService extends Service<Void> {
      */
     @Override
     public Void dispatch() {
-        Membership m = Membership.findByUserGroup(member.getId(), group.getId());
+        GroupMembership m = GroupMembership.findByUserGroup(member.getId(), group.getId());
         if (m != null) {
             m.delete();
         }
-        if (Membership.listByGroup(group.getId()).size() == 0) {
+        if (GroupMembership.listByGroup(group.getId()).size() == 0) {
             group.delete();
         }
         return null;
@@ -60,7 +60,7 @@ public class RemoveGroupMemberService extends Service<Void> {
     public boolean canExecute() {
         boolean ret = user != null && group != null && member != null;
         if (ret) {
-            return Membership.findByUserGroup(user.getId(), group.getId()) != null;
+            return GroupMembership.findByUserGroup(user.getId(), group.getId()) != null;
         }
 
         return false;

### 2016-03-13 project/httpserver/app/services/SearchDataService.java

@@ -1,11 +1,11 @@
 package services;
 
-import dtos.KeyValue;
-import models.Search;
-import org.bson.types.ObjectId;
-
 import java.util.ArrayList;
 import java.util.List;
+
+import dtos.KeyValue;
+import models.Data;
+import org.bson.types.ObjectId;
 
 
 
@@ -14,92 +14,85 @@ import java.util.List;
  */
 public class SearchDataService extends Service<List<String>> {
 
-    /** The caller. */
-    private ObjectId caller;
-    
-    /** The offset. */
-    private Integer limit = null, offset = null;
-    
-    /** The count. */
-    private Long count = null;
-    
-    /** The total. */
-    private Long total = null;
-    
-    /** The search. */
-    private String search = null;
-    
-    /** The filters. */
-    private List<List<KeyValue<String>>> filters = null;
+	/** The caller. */
+	private ObjectId caller;
+	
+	/** The offset. */
+	private Integer limit=null, offset = null;
+	
+	/** The count. */
+	private Long count =null;
+	
+	/** The total. */
+	private Long total =null;
+	
+	/** The search. */
+	private String search =null;
+	
+	/** The filters. */
+	private List<List<KeyValue<String>>> filters=null;
 
-    /**
-     * Instantiates a new search data service.
-     *
-     * @param callerId the caller id
-     * @param offset the offset
-     * @param limit the limit
-     * @param search the search
-     * @param filters the filters
-     */
-    public SearchDataService(String callerId, Integer offset, Integer limit, String search, List<List<KeyValue<String>>> filters) {
-        this.offset = offset;
-        this.limit = limit;
-        this.search = search;
-        this.caller = new ObjectId(callerId);
-        this.filters = filters;
+	/**
+	 * Instantiates a new search data service.
+	 *
+	 * @param callerId the caller id
+	 * @param offset the offset
+	 * @param limit the limit
+	 * @param search the search
+	 * @param filters the filters
+	 */
+	public SearchDataService(String callerId, Integer offset, Integer limit, String search, List<List<KeyValue<String>>> filters) {
+		this.offset = offset;
+		this.limit = limit;
+		this.search = search;
+		this.caller = new ObjectId(callerId);
+		this.filters = filters;
+	}
 
+	/*
+	 * (non-Javadoc)
+	 *
+	 * @see services.Service#dispatch()
+	 */
+	@Override
[…]

### 2016-03-13 project/httpserver/app/services/SearchGroupCandidatesService.java

@@ -53,7 +53,7 @@ public class SearchGroupCandidatesService extends Service<JSONArray> {
         Set<ObjectId> relations = new HashSet<ObjectId>();
 
         for (Relation u : new ListRelationsService(user.toString()).execute()) {
-            if (Membership.findByUserGroup(u.getTo(), group) == null) {
+            if (GroupMembership.findByUserGroup(u.getTo(), group) == null) {
                 relations.add(u.getTo());
             }
         }
@@ -64,7 +64,7 @@ public class SearchGroupCandidatesService extends Service<JSONArray> {
         List<List<KeyValue<String>>> filters = new ArrayList<>();
         filters.add(filter);
 
-        for (Search m : Search.search(query, null, null, filters)) {
+        for (Data m : Data.search(query, null, null, filters)) {
             if (relations.contains(m.getOwner())) {
                 Document doc =Data.findByOwner(m.getOwner(),null);
                 doc = (Document) doc.get("data");

### 2016-03-13 project/httpserver/app/services/SetAttributesService.java

@@ -1,57 +1,55 @@
 package services;
 
+import java.util.List;
 
 import dtos.AttributeDto;
 import exceptions.ServiceException;
-import models.Search;
+import models.Data;
 import org.bson.types.ObjectId;
-
-import java.util.List;
-
 
 
 /**
  * The Class SetAttributesService.
  */
-public class SetAttributesService extends Service<Search> {
+public class SetAttributesService extends Service<Data> {
 
-    /** The owner. */
-    private ObjectId owner;
-    
-    /** The attributes. */
-    private List<AttributeDto> attributes;
+	/** The owner. */
+	private ObjectId owner;
+	
+	/** The attributes. */
+	private List<AttributeDto> attributes;
 
-    /**
-     * Instantiates a new sets the attributes service.
-     *
-     * @param owner the owner
-     * @param attributes the attributes
-     */
-    public SetAttributesService(final String owner, List<AttributeDto> attributes) {
-        this.owner = owner != null ? new ObjectId(owner) : null;
-        this.attributes = attributes;
-    }
+	/**
+	 * Instantiates a new sets the attributes service.
+	 *
+	 * @param owner the owner
+	 * @param attributes the attributes
+	 */
+	public SetAttributesService(final String owner, List<AttributeDto> attributes) {
+		this.owner = owner !=null ? new ObjectId(owner) : null;
+		this.attributes = attributes;
+	}
 
-    /*
-     * (non-Javadoc)
-     *
-     * @see services.Service#dispatch()
-     */
-    @Override
-    public synchronized Search dispatch() throws ServiceException {
-        Search.deleteByOwner(owner);
-        Search data = new Search(owner, attributes).save();
-        return data;
-    }
+	/*
+	 * (non-Javadoc)
+	 *
+	 * @see services.Service#dispatch()
+	 */
+	@Override
+	public synchronized Data dispatch() throws ServiceException {
+		Data.deleteByOwner(owner);
+		Data data = new Data(owner,attributes).save();
+		return data;
+	}
 
-    /*
-     * (non-Javadoc)
-     *
-     * @see services.Service#canExecute()
-     */
-    @Override
-    public boolean canExecute() {
-        return owner != null;
-    }
+	/*
+	 * (non-Javadoc)
+	 * 
+	 * @see services.Service#canExecute()
+	 */
+	@Override
+	public boolean canExecute() {
+		return owner != null;
+	}
 
 }

### 2016-03-13 tese1/XX_implementation_model.tex
[ADD] model, minor code refactoring

@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.35\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_data}
@@
 
-Not less important, attributes can specify aggregations of objects.
+% Not less important, attributes can specify aggregations of objects.
 For example, the user role is an aggregator property that within users it allows the identifaction of each one is administrator.
 This aggregation specification is independent from the object type, so it's possible to search for the administrative role and return users and groups of users that contains that property.
+<= [currently this operation is possible but it is not specified]
 
-Our model supports defining read and write permissions for each attribute, for example an attribute can be set to public read access or private write access to a set of entities.
+Not less important, our model supports defining read and write permissions for each attribute, for example an attribute can be set to public read access or private write access to a set of entities.
 Implicitly if an entity can write an attribute it can also read it.
+
+The permission object is a mask of the data object, for each existing attribute in the data object it may exist the same attribute on the permission object specifying what entities can read or write it.
 
@@
     \centering
-    \begin{subfigure}[b]{0.25\textwidth}
+    \begin{subfigure}[b]{0.20\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_user}
@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.25\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_relation}
@@
 
-A user can only interact with users that maintain a relationship with him or with users within the same group.
-In order to validate a relationship, both users must agree on that relationship, by other words it must exist a bi-directional relation between both users.
+A user can only interact with friends or with group members.
+In order to validate a friendship, both users must agree on that friendship, by other words it must exist a bi-directional relation between both users.
 
@@
 A group can be public or private.
-If the group is public then it is visible to all users that mantain a relationship with a member of this group.
+If the group is public then it is visible to all users that mantain a friendship with a member of this group.
 If the group is private then it is only visible to their members.
@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.30\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_group}
     \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.25\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_group_membership}
@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.25\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_message}
@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.25\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_hyper_content}
     \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.25\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_time_tag}
@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.25\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_collaborative_content}
@@
     \centering
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.25\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_recording_chunk}
     \end{subfigure}
-    \begin{subfigure}[b]{0.3\textwidth}
+    \begin{subfigure}[b]{0.25\textwidth}
     	\includegraphics[width=\textwidth]{figures/model_recording_interval}

### 2016-03-13 project/httpserver/app/models/Data.java

@@ -7,7 +7,6 @@ import java.util.regex.Pattern;
 import dtos.AttributeDto;
 import dtos.KeyValue;
 import org.bson.Document;
-import org.bson.json.JsonParseException;
 import org.bson.types.ObjectId;
 import com.mongodb.client.FindIterable;
 import com.mongodb.client.MongoCollection;
@@ -23,13 +22,16 @@ public class Data {
     private ObjectId id = null;
 
     /** The data. */
-    private Document data = null;
+    private Document properties = null;
 
     /** The owner. */
     private ObjectId owner = null;
 
     /** The search. */
-    private List<String> search;
+    private List<String> searchableValues;
+
+    private List<String> identifiableKeys;
+
 
     /**
      * Instantiates a new data.
@@ -41,20 +43,25 @@ public class Data {
      */
     public Data(ObjectId owner, List<AttributeDto> attributes) {
         this.id = getIdFromOwner(owner);
-        this.data = new Document();
-
-        this.search = new ArrayList<String>();
+        this.properties = new Document();
+        this.searchableValues = new ArrayList<String>();
+        this.identifiableKeys = new ArrayList<String>();
 
         for (AttributeDto attr : attributes) {
             if (attr.isSearchable()) {
-                search.add(attr.getValue().toString().toLowerCase());
+                searchableValues.add(attr.getValue().toString().toLowerCase());
+            }
+            if (attr.isIdentifiable()) {
+                identifiableKeys.add(attr.getKey());
             }
         }
+
+
 
         for (AttributeDto attr : attributes) {
             Object value = attr.getValue();
 
-            data.append(attr.getKey(), value);
+            properties.append(attr.getKey(), value);
         }
         this.owner = owner;
     }
@@ -101,21 +108,30 @@ public class Data {
         user.id = doc.getObjectId("_id");
 
         user.owner = doc.getObjectId("owner");
-        user.data = (Document) doc.get("data");
-        user.search = (List<String>) doc.get("search");
+        user.properties = (Document) doc.get("data");
+        user.searchableValues = (List<String>) doc.get("search");
+        user.identifiableKeys = (List<String>) doc.get("iks");
+
         return user;
     }
 
+    public boolean isIdentifier(String key) {
+        return identifiableKeys.contains(key);
+    }
+
+
     /**
      * Save.
+
      *
      * @return the data
      */
     public Data save() {
         Document doc = new Document();
-        doc.put("data", data);
+        doc.put("data", properties);
[…]

### 2016-03-14 tese1/XX_implementation_signaling.tex
[RFT] improved protocol readability

\subsection{Signaling Protocol}

Our signaling protocol consists in sending JSON formated messages over WebSockets

### 2016-03-14 project/httpserver/app/controllers/Application.java

@@ -37,9 +37,6 @@ public class Application extends Controller {
     }
 
 
-    public Result test() {
-        return ok(views.html.test.render());
-    }
 
 
     /**

### 2016-03-14 project/httpserver/app/main/Room.java

@@ -243,7 +243,7 @@ public class Room implements Closeable {
                 JSONObject myProfile = new JSONObject(attributes1.toJson());
                 myProfile.put("id", userId);
 
-                final JSONObject myAdvertise = new JSONObject().put("id", "participants").put("data",
+                final JSONObject myAdvertise = new JSONObject().put("cmd", "participants").put("data",
                         new JSONArray().put(myProfile.put("online", true)));
 
                 ListGroupMembersService service = new ListGroupMembersService(userId, getGroupId());
@@ -265,7 +265,7 @@ public class Room implements Closeable {
                     }
                     otherUsers.put(otherProfile);
                 }
-                final JSONObject currentParticipants = new JSONObject().put("id", "participants").put("data", otherUsers);
+                final JSONObject currentParticipants = new JSONObject().put("cmd", "participants").put("data", otherUsers);
                 participant.sendMessage(currentParticipants.toString());
 
                 return participant;
@@ -291,7 +291,7 @@ public class Room implements Closeable {
                 JSONObject result = new JSONObject(attributes.toJson());
                 result.put("id", uid);
 
-                final JSONObject myAdvertise = new JSONObject().put("id", "participants").put("data",
+                final JSONObject myAdvertise = new JSONObject().put("cmd", "participants").put("data",
                         new JSONArray().put(result.put("online", false)));
                 sendMessage(myAdvertise.toString());
             } catch (ServiceException e) {

### 2016-03-14 project/httpserver/app/main/UserSession.java

@@ -94,7 +94,7 @@ public class UserSession implements Closeable, Comparable<UserSession> {
         endPoint.addOnIceCandidateListener(new EventListener<OnIceCandidateEvent>() {
             public void onEvent(OnIceCandidateEvent event) {
                 JSONObject response = new JSONObject();
-                response.put("id", "iceCandidate");
+                response.put("cmd", "iceCandidate");
                 response.put("data", new JSONObject(JsonUtils.toJson(event.getCandidate())));
 
                 try {
@@ -138,7 +138,7 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                             currentQRCodes.put(hash,startTime);
                         }
                         JSONObject msg = new JSONObject();
-                        msg.put("id", "qrCode");
+                        msg.put("cmd", "qrCode");
                         msg.put("data", content);
                         msg.put("hash", hash);
                         room.sendMessage(msg.toString());
@@ -161,7 +161,7 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                                     startTime = currentQRCodes.remove(hash);
                                 }
                                 JSONObject msg = new JSONObject();
-                                msg.put("id", "qrCode");
+                                msg.put("cmd", "qrCode");
                                 msg.put("hash", hash);
                                 room.sendMessage(msg.toString());
 
@@ -263,7 +263,7 @@ public class UserSession implements Closeable, Comparable<UserSession> {
         }
 
         JSONObject jRoot = new JSONObject();
-        jRoot.put("id", "content");
+        jRoot.put("cmd", "content");
         jRoot.put("data", jArr);
         jRoot.put("more", hasMore);
         return jRoot.toString();
@@ -470,7 +470,7 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                         else {
                             System.out.println("not seekable!");
                             JSONObject msg = new JSONObject();
-                            msg.put("id", "setTime");
+                            msg.put("cmd", "setTime");
                             msg.put("time", Tools.FORMAT.format(rec.getStart()));
                             timeOffset = new Date().getTime()- rec.getStart().getTime();
                             sendMessage(msg.toString());
@@ -543,7 +543,7 @@ public class UserSession implements Closeable, Comparable<UserSession> {
         // XXX [CLIENT_OFFER_06] XXX
[…]

### 2016-03-14 project/httpserver/app/services/ListOwnerAttributesService.java

@@ -7,6 +7,7 @@ import models.DataPermission;
 import org.bson.Document;
 import org.bson.types.ObjectId;
 
+import java.util.ArrayList;
 import java.util.List;
 import java.util.Map;
 import java.util.Set;
@@ -57,7 +58,7 @@ public class ListOwnerAttributesService extends Service<Document> {
             doc = (Document) doc.get("data");
 
             if (permission != null) {
-                for(String key : doc.keySet()){
+                for(String key : new ArrayList<>(doc.keySet())){
                     if(!permission.hasReadPermission(caller,key)){
                         doc.remove(key);
                     }

### 2016-03-15 tese1/09_related_work_signaling.tex
[ADD] organizing arch

--- 2015-06-07 tese/state02.tex
@@
   This kind of communication is also possible through WebSockets \cite{rfc7395}.
-  Today, multiple XMPP server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}(accessed June 1, 2015).}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}(accessed June 1, 2015).}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}(accessed June 1, 2015).} and \emph{Prosody}\footnote{\url{http://prosody.im/}(accessed June 1, 2015).}.
+  Today, multiple \ac{XMPP} server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}(accessed June 1, 2015).}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}(accessed June 1, 2015).}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}(accessed June 1, 2015).} and \emph{Prosody}\footnote{\url{http://prosody.im/}(accessed June 1, 2015).}.
 \emph{Ejabberd} is the server that implements more \ac{RFC} specifications and \ac{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}(accessed June 1, 2015).}.
   %RP fazes muitos paragrafos. Um paragrafo por tecnologia é normalmente suficiente se não forem muito grandes.
+
+{\color{blue}
+
+    One of the advantages that \ac{XMPP} based applications offer is the possibility to send messages across different servers, an \ac{XMPP} user is identified by its \textit{Jabber ID} which is definined by the pair username and server name in the form \textit{username@server}.
+
+    The user profile is represented in terms of Data Forms\cite{xep0004} protocol and managed (registration and update) through \textit{Info/Query} (IQ) \cite{rfc3920} requests to the \ac{XMPP} server which responds with an IQ response containing information to retrieve the registration fields needed to fill its profile (eg username, password, telephone).
+
+    An important aspect of this model is the ability to define a user as a set of attributes that can change dinamically in order to fit the IM applications needs.
+
+    The connection to the \ac{XMPP} server can be done through the same web server as the user is connected, but this web server would handle a lot of connections and that would have a great impact on the overall system performance.
+
+    Typically a web application consists in multiple web pages which are navigated by users.
+Each time a webpage is accessed, either from a new context or a transition from a previous page, its context is cleared except for local storage and cookies, the javascript context is cleared, including the \ac{XMPP} connections performed by \emph{strophe.js}, an automatic mechanism is required for avoiding the user implicit reauthentication.
+
+    One solution for reauthentication can be achieved by storing the user's JabberID and password on local storage and everytime a page is accessed the authentication is performed without the users knlowledge.
+Clearly this solution represents security flaws, the local storage can be easily accessed (HOW?
+explain!) locally or by performing a cross site scripting attack and reveal all the local storage, the same problem arises if the credentials are stored on cookies.
+
+    An alternative solution is known as Session Attachment \footnote{\url{https://metajack.im/2008/10/03/getting-attached-to-strophe/}}, which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} is passed to strophe.js in order to re-connect to the same stream on \ac{XMPP} server.
+Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request which is worthless if a user mantains more than one tab opened, for example for multiple conversations at the same time.
+
+    Another alternative would be the development of our solution in a single web page, which would increse drastically the complexity of our web application.
+
+    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that XMPP already implemment, but from all the \ac{XMPP} features, we don't need all of them, in fact we just need a subset of \ac{XMPP} features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.
+}
 

### 2016-03-15 tese1/XX_arch.tex
[ADD] organizing arch

@@
 \section{Backend}
-
-\subsection{XMPP Approach}
-
-One of the advantages that XMPP based applications offer is the possibility to send messages across different servers, an XMPP user is identified by its \textit{Jabber ID} which is definined by the pair username and server name in the form \textit{username@server}.
-
-The user profile is represented in terms of Data Forms (XEP-0004) protocol and managed (registration and update) through \textit{Info/Query} (IQ) [rfc3920/IQ Semantics] requests to the XMPP server which responds with an IQ response containing information to retrieve the registration fields needed to fill its profile (eg username, password, telephone).
-
-An important aspect of this model is the ability to define a user as a set of attributes that can change dinamically in order to fit the IM applications needs.
-
-The connection to the XMPP server could be done through the same web server as the user is connected, but this web server would handle a lot of connections and that would have a great impact on the overall system performance.
-Our proposed solution distributes the XMPP connections among users relieving this extra work to the web browsers.
-
-The first working prototype of our solution implemented the user registration and authentication components through \textit{Strophe.js}.
-The next step on implementation would be the access to the user profile and consequentally all the other features needed for our solution.
-Typically a web application consists in multiple web pages which are navigated by the users.
-Each time a webpage is accessed, either from a new context or a transition from a previous page, its context is cleared except for local storage and cookies, the javascript context is cleared, including the XMPP connections performed by strophe.js, an automatic mechanism is required for avoiding the user implicit reauthentication.
-
-One solution for reauthentication can be achieved by storing the user's JabberID and password on local storage and everytime a page is accessed the authentication is performed without the users knlowledge.
-Clearly this solution represents security flaws, the local storage can be easily accessed (HOW?
-explain!) locally or by performing a cross site scripting attack and reveal all the local storage, the same problem arises if the credentials are stored on cookies.
-
-An alternative solution is known as Session Attachment \footnote{\url{https://metajack.im/2008/10/03/getting-attached-to-strophe/}}, which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} is passed to strophe.js in order to re-connect to the same stream on XMPP server.
-Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request which is worthless if a user mantains more than one tab opened, for example for multiple conversations at the same time.
-
-Another alternative would be the development of our solution in a single web page, which would increse drastically the complexity of our web application.
-
-We know that an \textit{XMPP} based application could simplify our work just by using functionalities that XMPP already implemment, but from all the \textit{XMPP} features, we don't need all of them, in fact we just need a subset of XMPP features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.
 

### 2016-03-15 tese1/XX_implementation_signaling.tex
[ADD] organizing arch

@@
 
-Our signaling protocol consists in sending JSON formated messages over WebSockets
+Our signaling protocol consists in send and receiving JSON formated messages over WebSockets by both the server and client.
+
+\begin{lstlisting}[caption={General structure of our WebSocket messages},language=json]
+{
+	"cmd":<cmd>,
+	"data":<data>
+}
+\end{lstlisting}
+
+When a user enters in a group conference, after the page is completely loaded it is created a WebSocket that mantains a connection with our web servers. 
+Before creating the web socket, we must identify the user and check if he has permissions to participate in the conference.
+This is done by retrieving the session id from the cookie provided by the user through the \ac{HTTP} headers.
+It's important to save de user identification before the socket is created because after the the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455} the \ac{HTTP} context is lost.
+
+When the connection is established between the server and the client, a \emph{PeerConnection} is created on the client and imediately after an \ac{WebRTC} endpoint on the server specifying a possible set of \ac{ICE} servers to connect.
+
+The user is asked if he whants to share its camera and microphone, share screen or just receive streams from the server.
+If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2015).} may ask to install a pluggin if the browser doesn't support screen sharing.
+
+If the user decides to share either from camera of screen, \emph{getUserMedia} is called with the correspondent contraints in order to obtain a local stream.
+
+\begin{lstlisting}[caption={Media constraints},language=JavaScript]
+var screenShareConstraints = {	
+	"video": {
+		"mediaSource": "window" || "screen"
+	}, 
+	"audio": false
+};
+var cameraMicrophoneConstraints = {
+	"audio":true, 
+	"video":true 
+};
+var receiveOnlyConstraints = {
+	"offerToReceiveAudio":true,
+	"offerToReceiveVideo":true
+};
+\end{lstlisting}
+
+A popup is raised in order to ask the user to give permission to share those resources.
+If the resources are shared successfully the user creates an offer, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through WebSockets.
+If the resources are not shared or the user specified to receive stream only, an offer is created specifying the contraints for receiving only video and audio.
+
+The server receive and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint.
+Then a \emph{local session description} is created on the server and sent back to the client, after that the server trys to gather \ac{ICE} candidates.
+
+The client receives the server answer, sets the \emph{remote session description} and gets the ice candidates from the \ac{ICE} server.
+
+After a while both the server and client receive the \ac{ICE} candidates that allow the client to connect directly to \ac{KMS} and vice-versa.
+The candidates are received at the client which sets them to its \emph{PeerConnection}.
+The same is done on the server which receives the \ac{ICE} candidates from the client and propagates them to \ac{KMS}.
+
+The media session is established, the server starts to record any received stream and the client created an \ac{URL} correspondent to the stream location.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}[b]{\textwidth}
+    	\includegraphics[width=\textwidth]{figures/signaling}
+    \end{subfigure}
+    \caption{Signaling sequence diagram}
+\end{figure}

### 2016-03-15 project/httpserver/app/main/UserSession.java

@@ -530,16 +530,16 @@ public class UserSession implements Closeable, Comparable<UserSession> {
     /**
      * Process offer.
      *
-     * @param description the description
+     * @param rsd the description
      */
-    public void processOffer(String description) {
-        receiveOnly = description.contains("a=recvonly");
+    public void processOffer(String rsd) {
+        receiveOnly = rsd.contains("a=recvonly");
 
         // XXX [CLIENT_ICE_04] XXX
         // XXX [CLIENT_OFFER_04] XXX
         // XXX [CLIENT_OFFER_05] XXX
 
-        String lsd = endPoint.processOffer(description);
+        String lsd = endPoint.processOffer(rsd);
         // XXX [CLIENT_OFFER_06] XXX
         JSONObject data = new JSONObject().put("sdp", lsd).put("type", "answer");
 

### 2016-03-15 project/httpserver/public/this/js/screen.js

@@ -1,603 +1,0 @@
-// Last time updated at Sep 23, 2015, 08:32:23
-// Latest file can be found here: https://cdn.webrtc-experiment.com/screen.js
-// Muaz Khan     - https://github.com/muaz-khan
-// MIT License   - https://www.webrtc-experiment.com/licence/
-// Documentation - https://github.com/muaz-khan/WebRTC-Experiment/tree/master/screen-sharing
-(function() {
-
-    // a middle-agent between public API and the Signaler object
-    window.Screen = function(channel) {
-        var signaler, self = this;
-        this.channel = channel || location.href.replace(/\/|:|#|%|\.|\[|\]/g, '');
-
-        // get alerted for each new meeting
-        this.onscreen = function(screen) {
-            if (self.detectedRoom) return;
-            self.detectedRoom = true;
-
-            self.view(screen);
-        };
-
-        function initSignaler(roomid) {
-            signaler = new Signaler(self, (roomid && roomid.length) || self.channel);
-        }
-
-        function captureUserMedia(callback, extensionAvailable) {
-            getScreenId(function(error, sourceId, screen_constraints) {
-                console.log('screen_constraints', JSON.stringify(screen_constraints, null, '\t'));
-                navigator.getUserMedia(screen_constraints, function(stream) {
-                    stream.onended = function() {
-                        if (self.onuserleft) self.onuserleft('self');
-                    };
-
-                    self.stream = stream;
-
-                    var video = document.createElement('video');
-                    video.id = 'self';
-                    video[isFirefox ? 'mozSrcObject' : 'src'] = isFirefox ? stream : window.webkitURL.createObjectURL(stream);
-                    video.autoplay = true;
-                    video.controls = true;
-                    video.play();
-
-                    self.onaddstream({
-                        video: video,
-                        stream: stream,
-                        userid: 'self',
-                        type: 'local'
-                    });
-
-                    callback(stream);
-                }, function(error) {
-                    if (isChrome && location.protocol === 'http:') {
-                        alert('You\'re not testing it on SSL origin (HTTPS domain) otherwise you didn\'t enable --allow-http-screen-capture command-line flag on canary.');
-                    } else if (isChrome) {
-                        alert('Screen capturing is either denied or not supported. Please install chrome extension for screen capturing or run chrome with command-line flag: --enable-usermedia-screen-capturing');
-                    } else if (isFirefox) {
-                        alert(Firefox_Screen_Capturing_Warning);
-                    }
-
-                    console.error(e);
-                });
-            });
[…]

### 2016-03-15 tese1/11_related_work_collaboration.tex
[ADD] together JS more info

@@
 
-	\emph{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}(accessed June 2, 2015).} receives one stream from every participant on a conference, either from a \emph{jitsi} client or a \ac{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
+	\emph{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}(accessed June 2, 2015).} receives one stream from every participant on a conference, either from a \emph{jitsi} client or a \emph{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
 Although all the participants still need to download all the streams from the \emph{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
@@
 Despite this choice for signaling protocol, \emph{Jitsi Video Bridge} also supports \ac{SIP}.
+
+	{\color{blue}
+	\ac{KMS} supports transcoding, group communications, recording, mixing, broadcasting, aplying filters, image and sound analysis, for example it allows face and qrcode detection.
+\ac{KMS} functionalites are exposed through \emph{JSON-RPC} over \emph{WebSocket}, there are three clients available: JavaScript Client for web browsers, Java Client for Java EE Servers and JavaScript Cleint for Node.js servers.
+	\ac{KMS} suports streaming over \emph{WebRTC}, \ac{HTTP} and \ac{RTP} endpoints.
+Another important component of \ac{KMS} is \emph{Kurento Repository} which suports recording and playing directly from \emph{MongoDB}, that is important for providing a scalable media storage.
+Unlinke \emph{Jitsi Video Bridge}, \ac{KMS} does not enforce a specific signaling protocol.
+	}
 
@@
 It uses \ac{JSON} messages for \ac{OT} concurrency control but it does not provide storage.
-{\color{blue}\emph{TogetherJS} uses its own servers for the signaling phase and it supports microphone and mouse sharing between users.}
+{\color{blue}\emph{TogetherJS} uses its own servers for the signaling phase and it supports microphone and mouse sharing between users.
+It requires a simple server (also known as hub) that echoes messages between clients, all the synchronization work is performed by the clients.
+Altough the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}, if we would not choose \emph{NodeJS} for server implementation we would need read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}
 
@@
 
-{\color{green} \emph{TogetherJS} requires a simple server (also known as hub) that echoes messages between clients, all the synchronization work is performed by the clients.
-Altough the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}, we would need read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}
-
         %RP falta falar um pouco das ferramentas de colaboração já existentes? Wave, Hangouts, skype, etc? Acho que isto merecia uma secção à parte, mesmo que mais curta.

### 2016-03-18 tese1/08_related_work_rtc.tex
[ADD] together JS more info
[RFT] latex tables\n[ADD]streaming ok

@@
 
-\begin{figure}[H]
+\newcolumntype{Y}{>{\centering\arraybackslash}X}
+%\renewcommand{\arraystretch}{2}
+
+\begin{table}[]
 \centering
-	\includegraphics[width=0.8\textwidth]{figures/webrtc_stack.png}
+\label{my-label}
+    \begin{tabularx}{\textwidth}{*8{Y}@{}m{0pt}@{}} 
+    	\cline{5-8}
+		&&&\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{~~~~~~MediaStream~~~~~~} & \multicolumn{2}{c|}{~~~~~~DataChannel~~~~~~} 										&\\[12pt]
+		\hline 
+		\multicolumn{1}{|c|}{XHR} & \multicolumn{1}{c|}{SSE} & \multicolumn{2}{c|}{WebSocket} & \multicolumn{2}{c|}{SRTP} & \multicolumn{2}{c|}{SCTP} 				&\\[12pt]
+		\cline{1-3}\cline{6-8} 
+		\multicolumn{3}{|c|}{HTTP 1.X/2.0} & \multicolumn{1}{c|}{~~~~~~~~~~~~~~~~~~} & \multicolumn{1}{c}{~~~~~~~~~~~~~~~~} & \multicolumn{3}{|c|}{Session (DTLS)} 	&\\[12pt]
+		\hline
+		\multicolumn{4}{|c|}{Session (TLS) - optional} & \multicolumn{4}{c|}{ICE, STUN, TURN} 																		&\\[12pt]
+		\hline
+		\multicolumn{4}{|c|}{Transport (TCP)} & \multicolumn{4}{c|}{Transport (UDP)} 																				&\\[12pt]
+		\hline
+		\multicolumn{8}{|c|}{Network (IP)}  																														&\\[12pt]
+		\hline
+	\end{tabularx}
 	\caption{WebRTC protocol Stack}
-\end{figure}
+
+\end{table}
 
@@
 %RP TLS?
-
-\begin{figure}[H]
+%\def\tabularxcolumn#1{m{#1}}
+\begin{table}[]
 \centering
-	\includegraphics[width=0.6\textwidth]{figures/basic_protocols.png}
+    \begin{tabularx}{\textwidth}{X|Y|Y|Y|@{}m{0pt}@{}}
+\cline{2-4}
+                                                     & \textbf{TCP}  & \textbf{UDP}     & \textbf{SCTP}    &\\[12pt]\hline
+\multicolumn{1}{|c|}{\textbf{Reliability}}           & reliable      & unreliable       & configurable     &\\[12pt]\hline
+\multicolumn{1}{|c|}{\textbf{Delivery}}              & ordered       & unordered        & configurable     &\\[12pt]\hline
+\multicolumn{1}{|c|}{\textbf{Transmission}}          & byte-oriented & message-oriented & message-oriented &\\[12pt]\hline
+\multicolumn{1}{|c|}{\textbf{Flow control}}          & yes           & no               & yes              &\\[12pt]\hline
+\multicolumn{1}{|c|}{\textbf{Congestion control}} 	 & yes           & no               & yes              &\\[12pt]\hline
+\end{tabularx}
 	\caption{Overview of transport protocols}
-\end{figure}
+\end{table}
 

### 2016-03-18 tese1/XX_implementation_recording.tex
[RFT] latex tables\n[ADD]streaming ok

\subsection{Stream Recording}
	In this section we describe our approaches for the stream recording implementation, namely recording on the client side and server side.

\subsubsection{Client side recording}
	In order to record used shared streams, our first approach consisted on recording video and audio streams into blobs of limited duration. Because each user was recording directly from their shared local resources, we achieved the best stream quality.

	After recording, each block was uploaded to our server through \ac{HTTP} and saved into the filesystem, the block metadata was created and inserted on our database. Each block metadata contained the file's location for the recorded block, the starting date, duration, user identification and group identification. When the file was completely saved the metadata was advertised to the remaining users. This metadata was simply used to refresh the user interface, the metadata was completely discard after that because an huge amount of small blocks metadata would use more and more memory as time went by.

	The process to play a video was quite simple, the user specified which user and date was intended to play, the server calculated the intersection between the requested date and the block bounds and returned the file to the client.

	Altough this idea was fairly simple, we couldn't achieve seamless sequential block switching. Downloading the video file always toke a noticeable amount of time. To solve this problem while we were playing a block, the next one was downloading in parallel so when the current block finished playing we would have an available block to play.

	Block switching became more acceptable, but switching the url always produced a flash. We solved this problem by having two layers and set the url to the back layer, the front layer freezed with the last frame, then we changed the back layer to front.

	After we implemented our recording solution using this approach, we tested localy and remotely. For remote connections we observed a fairly high bandwidth usage mainly because blocks were both sent and received at maximum quality.

\subsubsection{Server side recording to filesystem}

	Before recording any type of stream we had to analyse the user media offer in order to check if a video as realy being received by \ac{KMS}, otherwise if we would not verify the user's offer, the recording video would be black.

	The streaming content received on \ac{KMS} was already compressed due to \ac{WebRTC}'s exchanged \ac{QoS} metrics data. As a direct consequence our recorder solution used on most cases less disk space per block, but would never use more storage than client side recording. \ac{KMS} allows recording files using \ac{WebM} and \ac{mp4} containers.

	With server side recording, the user would maintain always the same stream url even if it is playing realtime video or reproducing recorded video. When a user desires to play recorded video, a websocket message is sent specifying the time and the intended user id, the group identification is not sent because it is already associated to the websocket. The server performs the same calculations in order to find a block that intersects the requested time, plays it and when finished the next part is automatically played wihtout the user intervention.

	We observed differences in image quality when switching parts, that was even noticeable if we set a short block duration. We also noticed a small gap on audio when switching blocks but it was acceptable and speech recognition was not very affected.

	Back when we were implementing our solution, \ac{KMS} had not support for seeking videos, which meant that blocks would would always start playing from their beggining. This lead to a playing time error that could be at most half the duration of a block {\color{red}[time figure with t/2 maximum error]}.

	For playing video with an higher velocity we used \emph{ffmpeg}\footnote{\url{https://www.ffmpeg.org/(accessed: 17 March 2016)}} to convert the block into a new video with the desired velocity and seek time. Because the media duration is known, when the video started to being convert the headers located at the begining of the file were already written and that made it possible to stream while converting.

	{\color{red} [TALK ABOUT COMPLEXITY OF DECODE VS DECODE+MANIPULATE+ENCODE]}

	Although we implemented a solution that worked, we imediately noticed that \emph{ffmpeg} would take some time to initialize and that lead to pauses between switching parts.

	Later the \emph{Kurento} team released a version with support for seeking videos but we had to suspend the implementation of the fast forwarding feature as currently \ac{KMS} is not supporting that.

	We could implement fast forwarding without real time conversion by creating multiple versions of the same video with different velocities after the recording of a block. When a user needed to play he would also need to specify one of the available velocities. We did not followed this approach has it would require a bigger disk space usage.

\subsubsection{Server side recording to database}

	One of our concerns during the development of our solution was the storage scalability. Saving files directly into filesystem would require an extra effort to distribute and replicate files among servers. For that purpose \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB}.

	One of the features that \emph{Kurento Repository} provides is the ability to play directly from the database without having to download the entire file to \ac{KMS}. The same is true for recording, but because the file headers are in the beggining and the file is written until is stops, the headers don't contain the necessary information for seeking the file.

	Although we gain with scalability with this approach, we lose access over the file for changing it to fit our needs, namely for using \emph{ffmpeg} or other video manipulation tool.

	Although we did not implemented recorded file seeking, that could be achieved by waiting for full file recording and then proceed to database insertion with the correct headers. Another approach would be the specification of the file duration before recording so the correct file headers could be written a priori. Both approaches were not possible to implement using just the \emph{Kurento} clients, we would need change the source code of \emph{Kurento} in order to add those new features.

### 2016-03-18 tese1/XX_implementation_signaling.tex
[RFT] latex tables\n[ADD]streaming ok

@@
 
-Our signaling protocol consists in send and receiving JSON formated messages over WebSockets by both the server and client.
+The signaling protocol could be implemented using \ac{HTTP} messages but they would transport extra information such as \ac{HTTP} headers and would follow a request response signaling mechnism which would not be the best option, as multiple ice candidates can arrive at any time.
+For that reason bi-directional communications were prefered.
+
+Our signaling protocol consists in send and receiving JSON formated messages over WebSockets by both the server and the client.
 
@@
 Before creating the web socket, we must identify the user and check if he has permissions to participate in the conference.
-This is done by retrieving the session id from the cookie provided by the user through the \ac{HTTP} headers.
+The user identification is done by retrieving the session id from the cookie provided by the user through the \ac{HTTP} headers then we retrieve all the information need from database in order to check if the user has permissions to join the conference room.
 It's important to save de user identification before the socket is created because after the the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455} the \ac{HTTP} context is lost.
@@
 
+The \emph{local session description} contains the session identifier, codecs, containers, transport protocols and ports used per media type.
+The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} doesn't need to mix, record and analyse streams coming from the user \ac{WebRTC} endpoint.
+
 The server receive and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint.
@@
 
-The media session is established, the server starts to record any received stream and the client created an \ac{URL} correspondent to the stream location.
+An \ac{ICE} candidate contains an \ac{IP}, port, used transport protocol and an attribute named \emph{sdpMLineIndex} that is used for mapping to the \emph{remote session description} media type.
+
+Both intervenients test the conectivity of each \ac{ICE} candidate.
+When a connection is established the user and server start to interchange stream data but other \ac{ICE} candidates may arrive with better connections, when that happens the connection changes seamlessly.
+
+With the media session established, the server starts to record any received stream and the client creates an \ac{URL} correspondent to the stream location.
 
@@
     \centering
-    \begin{subfigure}[b]{\textwidth}
+    \begin{subfigure}[b]{\linewidth}
     	\includegraphics[width=\textwidth]{figures/signaling}

### 2016-03-18 project/httpserver/app/main/Room.java

@@ -145,7 +145,7 @@ public class Room implements Closeable {
                         array.put(Tools.FORMAT.format(interval.getEnd()));
 
                         JSONObject msg = new JSONObject();
-                        msg.put("id", "rec");
+                        msg.put("cmd", "rec");
                         msg.put(interval.getId().toString(), array);
                         sendMessage(msg.toString());
 

### 2016-03-18 project/httpserver/public/adapter/adapter.js

@@ -1,812 +1,178 @@
-/*! adapterjs - v0.13.0 - 2016-01-08 */
-
-// Adapter's interface.
-var AdapterJS = AdapterJS || {};
-
-// Browserify compatibility
-if(typeof exports !== 'undefined') {
-  module.exports = AdapterJS;
-}
-
-AdapterJS.options = AdapterJS.options || {};
-
-// uncomment to get virtual webcams
-// AdapterJS.options.getAllCams = true;
-
-// uncomment to prevent the install prompt when the plugin in not yet installed
-// AdapterJS.options.hidePluginInstallPrompt = true;
-
-// AdapterJS version
-AdapterJS.VERSION = '0.13.0';
-
-// This function will be called when the WebRTC API is ready to be used
-// Whether it is the native implementation (Chrome, Firefox, Opera) or
-// the plugin
-// You may Override this function to synchronise the start of your application
-// with the WebRTC API being ready.
-// If you decide not to override use this synchronisation, it may result in
-// an extensive CPU usage on the plugin start (once per tab loaded)
-// Params:
-//    - isUsingPlugin: true is the WebRTC plugin is being used, false otherwise
-//
-AdapterJS.onwebrtcready = AdapterJS.onwebrtcready || function(isUsingPlugin) {
-  // The WebRTC API is ready.
-  // Override me and do whatever you want here
-};
-
-// New interface to store multiple callbacks, private
-AdapterJS._onwebrtcreadies = [];
-
-// Sets a callback function to be called when the WebRTC interface is ready.
-// The first argument is the function to callback.\
-// Throws an error if the first argument is not a function
-AdapterJS.webRTCReady = function (callback) {
-  if (typeof callback !== 'function') {
-    throw new Error('Callback provided is not a function');
-  }
-
-  if (true === AdapterJS.onwebrtcreadyDone) {
-    // All WebRTC interfaces are ready, just call the callback
-    callback(null !== AdapterJS.WebRTCPlugin.plugin);
-  } else {
-    // will be triggered automatically when your browser/plugin is ready.
-    AdapterJS._onwebrtcreadies.push(callback);
-  }
-};
-
-// Plugin namespace
-AdapterJS.WebRTCPlugin = AdapterJS.WebRTCPlugin || {};
-
[…]

### 2016-03-18 project/httpserver/public/adapter/adapter.screenshare.js

@@ -1,4 +1,4 @@
-/*! adapterjs - v0.13.0 - 2016-01-08 */
+/*! adapterjs - v0.13.1 - 2016-03-15 */
 
 // Adapter's interface.
 var AdapterJS = AdapterJS || {};
@@ -11,13 +11,13 @@ if(typeof exports !== 'undefined') {
 AdapterJS.options = AdapterJS.options || {};
 
 // uncomment to get virtual webcams
- AdapterJS.options.getAllCams = true;
+// AdapterJS.options.getAllCams = true;
 
 // uncomment to prevent the install prompt when the plugin in not yet installed
- AdapterJS.options.hidePluginInstallPrompt = true;
+// AdapterJS.options.hidePluginInstallPrompt = true;
 
 // AdapterJS version
-AdapterJS.VERSION = '0.13.0';
+AdapterJS.VERSION = '0.13.1';
 
 // This function will be called when the WebRTC API is ready to be used
 // Whether it is the native implementation (Chrome, Firefox, Opera) or
@@ -58,6 +58,7 @@ AdapterJS.webRTCReady = function (callback) {
 AdapterJS.WebRTCPlugin = AdapterJS.WebRTCPlugin || {};
 
 // The object to store plugin information
+/* jshint ignore:start */
 AdapterJS.WebRTCPlugin.pluginInfo = {
   prefix : 'Tem',
   plugName : 'TemWebRTCPlugin',
@@ -74,6 +75,7 @@ if(!!navigator.platform.match(/^Mac/i)) {
 else if(!!navigator.platform.match(/^Win/i)) {
   AdapterJS.WebRTCPlugin.pluginInfo.downloadLink = 'http://bit.ly/1kkS4FN';
 }
+/* jshint ignore:end */
 
 AdapterJS.WebRTCPlugin.TAGS = {
   NONE  : 'none',
@@ -160,16 +162,14 @@ AdapterJS.WebRTCPlugin.callWhenPluginReady = null;
 // This function is the only private function that is not encapsulated to
 // allow the plugin method to be called.
 __TemWebRTCReady0 = function () {
-  webrtcDetectedVersion = AdapterJS.WebRTCPlugin.plugin.version;
-
   if (document.readyState === 'complete') {
     AdapterJS.WebRTCPlugin.pluginState = AdapterJS.WebRTCPlugin.PLUGIN_STATES.READY;
     AdapterJS.maybeThroughWebRTCReady();
   } else {
[…]

### 2016-03-18 project/httpserver/public/this/js/hyper.socket.js

@@ -306,16 +306,16 @@ var HyperWebSocket = new (function() {
 				
 				switch(cmd){
 					case 'iceCandidate':
-						// console.log(id,message);
+						console.log(cmd,message);
 						var candidate = new RTCIceCandidate(message.data);
 				
 						pc.addIceCandidate(candidate, function() {
-							// console.log(candidate);
+
 						}, logError);
 					break;
 					case 'answer':
 
-						// console.log(id, message);
+						console.log(cmd, message);
 						var rsd = new RTCSessionDescription(message.data);
 						// XXX [CLIENT_OFFER_08] XXX
 						pc.setRemoteDescription(rsd, function(){
@@ -410,13 +410,12 @@ var HyperWebSocket = new (function() {
 				kscb();
 
                 // XXX [CLIENT_ICE_01] XXX
-                pc = new RTCPeerConnection(/*{
-                    iceServers : [ {
-                        urls : "stun:stun.l.google.com:19302"
-                    }, {
-                        urls : "stun:23.21.150.121"
-                    } ]
-                }*/);
+                pc = new RTCPeerConnection({
+                    iceServers : [
+                        { urls : "stun:stun.l.google.com:19302" },
+                        { urls : "stun:23.21.150.121" }
+                    ]
+                });
 
                 // XXX [CLIENT_ICE_02] XXX
                 pc.onicecandidate = function(event) {

### 2016-03-19 curriculo/cv.tex
[]

--- 2016-03-02 curriculo/cv.tex
@@
     {Applications Development}
-    {Bullray-CIT\\1000€/month (part time)}
+    {Bullray-CIT}
     {\emph{Development of a user interface for a bank's security device that is synchronized with servers.
@@
     {Android Development}
-    {Bullray-CIT\\1000€/month (part time)}
+    {Bullray-CIT}
     {\emph{Optimization and bug resolution of an \underline{Android} application that uploads GPS, camera and microphone data in background.}}
@@
     {Systems Development}
-    {Bullray-CIT\\1000€/month (part time)}
+    {Bullray-CIT}
     {\emph{Development and configuration of a streaming server that supports receiving and recording video from authenticated users through \underline{ffmpeg}.
@@
     {Full-Stack Web Development}
-    {Bullray-CIT\\1000€/month (part time)}
+    {Bullray-CIT}
     {\emph{Development of a solution for a generic incident handling server, support for real time monitoring through \underline{WebSockets}.
@@
     {Full-Stack Web Development}
-    {Bullray-CIT\\1000€/month (part time)}
+    {Bullray-CIT}
     {\emph{Development of an architecture for a chat application.
@@
     {Scientific Initiation Scholarship}
-    {INESC\\385€/month (part time)}
+    {INESC}
     {\emph{Frontend Development for Provide Results's Bimk project. Experience with \underline{PlayFramework}, \underline{Revit API} and \underline{GIT}.}}
@@
     {Scientific Initiation Scholarship}
-    {CEG-IST\\385€/month (part time)}
+    {CEG-IST}
     {\emph{Programming a Decision Analysis Software (MACBETH) with \underline{WPF}. Development of an algorithm to suggest judgements based on past decisions using \underline{lpsolve}.}}
@@
     {Garbage Collector}
-    {Jovens em Movimento, Oeiras\\120€/month(full time)}
+    {Jovens em Movimento, Oeiras}
     {Cleaning and maintenance of public spaces.}
@@
 \begin{entrylist}
+%  \entry
+%    {201?}
+%    {My personal website}
+%    {\href{http://play.google.com/store/apps/details?id=pt.ist.bombraider}{henrique.rocha.pt}}
+%    {I like PlayFramework, but I still like more to learn different technologies, that's why I did this website using ???, just for fun.}
+
   \entry

### 2016-03-20 tese2/01_abstract.tex
[ADD] template2

--- 2016-03-12 tese1/01_abstract.tex
@@
-\begin{abstract}
 The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to communication and collaboration services.
@@
 
-\keywords{WebRTC, asynchronous, communications, collaboration}
-\end{abstract}
+\vspace{1cm}
+
+% TODO 4 to 6 keywords;
+\textbf{\Large Keywords:} WebRTC, asynchronous, communications, collaboration

### 2016-03-20 tese2/11_related_work_collaboration.tex
[ADD] template2

--- 2016-03-15 tese1/11_related_work_collaboration.tex
@@
-\subsection{Extending collaboration tools with time manipulation}
+\section{Extending collaboration tools with time manipulation}
 \label{collab}
@@
 
-\subsubsection{Streaming and Recording}\label{recstream}~\\
+\subsection{Streaming and Recording}\label{recstream}~\\
 	If, for instance, one wants to rewind a real-time video, recordings will be needed from whom is streaming the video. 
@@
 
-\subsubsection{Media Types}\label{mediatype}~\\
+\subsection{Media Types}\label{mediatype}~\\
 	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuous, the second one describes it as interactive or non-interactive.
@@
 
- \subsubsection{Recording and Streaming Interactive Media}\label{intrecord}~\\
+ \subsection{Recording and Streaming Interactive Media}\label{intrecord}~\\
     Mauve et al.
@@
 
-  \subsubsection{Collaborative Environment}\label{collabenv}~\\
+  \subsection{Collaborative Environment}\label{collabenv}~\\
 	\emph{Google Wave} was a distributed collaboration platform based on \emph{Jupiter}\cite{jupiter} that adopted \ac{OT} techniques.
@@
 
+\begin{table}[H]
+\centering
+
+\label{my-label}
+\begin{tabular}{|c|c|c|c|}
+\hline
+\textbf{Library} & \textbf{Own Server} & \textbf{Own Storage} & \textbf{Operations} \\ \hline
+ShareJS          & yes                 & yes                  & text+objects        \\ \hline
+TogetherJS       & yes                 & no                   & text+objects        \\ \hline
+Goodow           & yes                 & yes                  & text+objects        \\ \hline
+%Etherpad        & yes                 & yes                  & text (not sure)     \\ \hline
+OT.js            & no                  & no                   & text                \\ \hline
+\end{tabular}
+\caption{Comparision between Operational Transformation libraries}
+\end{table}
+
         %RP falta falar um pouco das ferramentas de colaboração já existentes? Wave, Hangouts, skype, etc? Acho que isto merecia uma secção à parte, mesmo que mais curta.

### 2016-03-20 tese2/13_implementation_model.tex
[ADD] template2

--- 2016-03-13 tese1/XX_implementation_model.tex
@@
-\subsection{Model}
+\section{Model}
 The model is the most important component of our solution, a badly designed model can imply serious deficulties when implementing new features that are not part of the plans, sometimes we had to redesign the model in order to support new features.
 
-\subsubsection {Generic model}
+\subsection {Generic model}
 
@@
 
-\begin{figure}[H]
+\begin{table}[H]
 \centering
-    \begin{subfigure}[b]{0.35\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_data}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.35\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_data_permission}
-    \end{subfigure}
+\label{my-label}
+    \begin{tabular}{cc}
+        \begin{tabular}{|ll|}
+            \hline
+            \multicolumn{2}{|c|}{\textbf{Data}}                             \\ \hline
+            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\medbullet$ owner           & ObjectId                         \\ 
+            $\medbullet$ properties      & Document                         \\ 
+            $\medcirc$ searchableValue & List\textless Text\textgreater     \\ 
+            $\medcirc$ identiableKeys  & List\textless Text\textgreater     \\ \hline
+        \end{tabular}
+        \begin{tabular}{|ll|}
+            \hline
+            \multicolumn{2}{|c|}{\textbf{DataPermission}}                   \\ \hline
+            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\medbullet$ owner           & ObjectId                         \\ 
+            $\medcirc$ readPermissions & Document                           \\ 
+            $\medcirc$ writePermissions  & Document                         \\ \hline
+        \end{tabular}    
+    \end{tabular}
 \caption{Generic data model}
-\end{figure}
+\end{table}
 
@@
 
-\subsubsection{User model}
+\subsection{User model}
 
@@
 
-\begin{figure}[H]
+\begin{table}[H]
 \centering
-    \begin{subfigure}[b]{0.20\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_user}
-    \end{subfigure}
+\label{my-label}
+    \begin{tabular}{|ll|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{User}}                             \\ \hline
+        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+        $\medbullet$ hash           & Text                         \\ 
+        $\medbullet$ salt      & Text                         \\ \hline
+    \end{tabular}
 \caption{User model}
-\end{figure}
+\end{table}
 
-\subsubsection{Relation model}
+\subsection{Relation model}
 
@@
 
-\begin{figure}[H]
+\begin{table}[H]
 \centering
-    \begin{subfigure}[b]{0.25\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_relation}
-    \end{subfigure}
+\label{my-label}
+    \begin{tabular}{|ll|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{Relation}}                             \\ \hline
+        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+        $\medbullet$ source           & ObjectId                         \\ 
+        $\medbullet$ target      & ObjectId                         \\ \hline
+    \end{tabular}
 \caption{Relation model}
-\end{figure}
+\end{table}
 
@@
 
-\subsubsection{Group model}
+\subsection{Group model}
 
@@
 
-\begin{figure}[H]
+\begin{table}[H]
 \centering
-    \begin{subfigure}[b]{0.30\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_group}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.25\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_group_membership}
-    \end{subfigure}
+\label{my-label}
+    \begin{tabular}{cc}
+        \begin{tabular}{|ll|}
+            \hline
+            \multicolumn{2}{|c|}{\textbf{Group}}                             \\ \hline
+            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\medcirc$ inviteToken           & Text                         \\ 
+            $\medbullet$ visibility      & Text                         \\ \hline
+        \end{tabular}
+
+        \begin{tabular}{|ll|}
+            \hline
+            \multicolumn{2}{|c|}{\textbf{GroupMembership}}                   \\ \hline
+            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\medbullet$ groupId           & ObjectId                         \\ 
+            $\medbullet$ userId & ObjectId                           \\ \hline
+        \end{tabular}    
+    \end{tabular}
 \caption{Group model}
-\end{figure}
+\end{table}
 
@@
 
-\subsubsection{Message model}
+\subsection{Message model}
 
@@
 
-\begin{figure}[H]
+\begin{table}[H]
 \centering
-    \begin{subfigure}[b]{0.25\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_message}
-    \end{subfigure}
+\label{my-label}
+    \begin{tabular}{|ll|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{Message}}                             \\ \hline
+        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+        $\medbullet$ source           & ObjectId                         \\ 
+        $\medbullet$ target      & ObjectId                         \\ 
+        $\medbullet$ sequence      & Number                         \\ 
+        $\medbullet$ content      & Text                         \\ \hline
+    \end{tabular}
 \caption{Message model}
-\end{figure}
+\end{table}
 
-\subsubsection{Hyper content model}
+\subsection{Hyper content model}
 
@@
 
-\begin{figure}[H]
+\begin{table}[H]
 \centering
-    \begin{subfigure}[b]{0.25\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_hyper_content}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.25\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_time_tag}
-    \end{subfigure}    
+\label{my-label}
+    \begin{tabular}{cc}
+        \begin{tabular}{|ll|}
+            \hline
+            \multicolumn{2}{|c|}{\textbf{HyperContent}}                             \\ \hline
+            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\medbullet$ groupId            & ObjectId                         \\ 
+            $\medbullet$ start              & Date                         \\ 
+            $\medbullet$ end                & Date                         \\ 
+            $\medbullet$ content            & Text                         \\ \hline
+        \end{tabular}
+
+        \begin{tabular}{|ll|}
+            \hline
+            \multicolumn{2}{|c|}{\textbf{TimeTag}}                   \\ \hline
+            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\medbullet$ groupId            & ObjectId                         \\ 
+            $\medbullet$ title & Text                           \\ 
+            $\medbullet$ time & Date                           \\ \hline
+
+        \end{tabular}    
+    \end{tabular}
     \caption{Hyper content model}
-\end{figure}
+\end{table}
 
-\subsubsection{Collaborative Content model}
+\subsection{Collaborative Content model}
 
@@
 
-\begin{figure}[H]
+\begin{table}[H]
 \centering
-    \begin{subfigure}[b]{0.25\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_collaborative_content}
-    \end{subfigure}
+\label{my-label}
+    \begin{tabular}{|ll|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{CollaborativeContent}}                             \\ \hline
+        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+        $\medbullet$ groupId            & ObjectId                         \\ 
+        $\medbullet$ content            & Text                         \\ \hline
+    \end{tabular}
     \caption{Collaborative content model}
-\end{figure}
+\end{table}
 
-\subsubsection{Recording model}
+\subsection{Recording model}
 
 During a conversation, users may allow sharing their web cameras, by doing so their video is stored in recording chunks.
-Each chunk represents an interval of time  $T=[ \,c^{start},c^{end}[ \,$, it contains a reference to a group and a mapping between entity that was recorded and the correspondent video url.
+Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains a reference to a group and a mapping between entity that was recorded and the correspondent video url.
 
-\begin{figure}[H]
+\begin{table}[H]
 \centering
-    \begin{subfigure}[b]{0.25\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_recording_chunk}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.25\textwidth}
-    	\includegraphics[width=\textwidth]{figures/model_recording_interval}
-    \end{subfigure}
+\label{my-label}
+    \begin{tabular}{cc}
+        \begin{tabular}{|ll|}
+            \hline
+            \multicolumn{2}{|c|}{\textbf{RecordingInterval}}                   \\ \hline
+            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\medbullet$ groupId            & ObjectId                         \\ 
+             $\medbullet$ start             & Date                         \\ 
+            $\medbullet$ end                & Date                         \\ \hline
+        \end{tabular}   
+        \begin{tabular}{|ll|}
+            \hline
+            \multicolumn{2}{|c|}{\textbf{RecordingChunk}}                             \\ \hline
+            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\medbullet$ groupId            & ObjectId                         \\ 
+            $\medbullet$ interval           & ObjectId                      \\ 
+            $\medbullet$ start              & Date                         \\ 
+            $\medbullet$ end                & Date                         \\ 
+            $\medbullet$ sequenceNumber     & Number                         \\ 
+            $\medbullet$ urls               & Document                       \\ \hline
+        \end{tabular}
+    \end{tabular}
     \caption{Recording model}
-\end{figure}
+\end{table}
 

### 2016-03-20 tese2/15_implementation_recording.tex
[ADD] template2

--- 2016-03-18 tese1/XX_implementation_recording.tex
@@
-\subsection{Stream Recording}
+\section{Stream Recording}
 	In this section we describe our approaches for the stream recording implementation, namely recording on the client side and server side.
 
-\subsubsection{Client side recording}
+\subsection{Client side recording}
 	In order to record used shared streams, our first approach consisted on recording video and audio streams into blobs of limited duration.
@@
 
-\subsubsection{Server side recording to filesystem}
+\subsection{Server side recording to filesystem}
 
@@
 As a direct consequence our recorder solution used on most cases less disk space per block, but would never use more storage than client side recording.
-\ac{KMS} allows recording files using \ac{WebM} and \ac{mp4} containers.
+\ac{KMS} allows recording files using \emph{webm} and \emph{mp4} containers.
 
@@
 	Back when we were implementing our solution, \ac{KMS} had not support for seeking videos, which meant that blocks would would always start playing from their beggining.
-This lead to a playing time error that could be at most half the duration of a block {\color{red}[time figure with t/2 maximum error]}.
+This lead to a theoric playing time error that can be at most half the duration of a block.
+In practice the maximum playing time error coincides with the block duration because we perform the intersection between the requested date and the block but we could decrease the error by half if we calculate the intersection with the requested time plus half the duration.
 
-	For playing video with an higher velocity we used \emph{ffmpeg}\footnote{\url{https://www.ffmpeg.org/(accessed: 17 March 2016)}} to convert the block into a new video with the desired velocity and seek time.
+	\begin{figure}
+			\centering
+
+	\begin{tikzpicture}[y=1cm, x=1cm, thick, font=\footnotesize]
+
+		\tikzset{
+		   brace_top/.style={
+		     decoration={brace},
+		     decorate
+		   },
+		   brace_bottom/.style={
+		     decoration={brace, mirror},
+		     decorate
+		   }
+		}
+
+		% time line hour
+		\draw[line width=1.2pt, ->, >=latex'](0,-3.0) -- coordinate (x axis) (9,-3.0) node[right] {time}; 
+		\foreach \x in {1,2,3,4,5} \draw (\x*1.5,-2.9) -- (\x*1.5,-3.1) node[below] {};
+
+		% top brace
+		\draw [brace_top] (1*1.5+0.03,-2.7) -- node [above, pos=0.5] {$ block_{n}$} 	(3*1.5-0.03,-2.7);
+		\draw [brace_top] (3*1.5+0.03,-2.7) -- node [above, pos=0.5] {$ block_{n+1}$} 	(5*1.5-0.03,-2.7);
+
+	    \draw  node[fill,circle,scale=0.6]at (2*1.5,-3.0)  {};%
+
+		% low brace period
+
+		\draw [brace_bottom] (1*1.5+0.03,-3.3) -- node [below, pos=0.5] {$e_{n}=\frac{d}{2}$} 		(2*1.5-0.03,-3.3);
+		\draw [brace_bottom] (2*1.5+0.03,-3.3) -- node [below, pos=0.5] {$e_{n+1}=\frac{d}{2}$} 	(3*1.5-0.03,-3.3);
+		\draw [brace_bottom] (3*1.5+0.03,-3.3) -- node [below, pos=0.5] {$d$} 						(5*1.5-0.03,-3.3);
+
+	\end{tikzpicture}
+		\caption{Theoretic maximum playing error}
+	\end{figure}
+
+	\begin{figure}
+			\centering
+	\begin{tikzpicture}[y=1cm, x=1cm, thick, font=\footnotesize]
+
+		\tikzset{
+		   brace_top/.style={
+		     decoration={brace},
+		     decorate
+		   },
+		   brace_bottom/.style={
+		     decoration={brace, mirror},
+		     decorate
+		   }
+		}
+
+		% time line hour
+		\draw[line width=1.2pt, ->, >=latex'](0,-3.0) -- coordinate (x axis) (9,-3.0) node[right] {time}; 
+		\foreach \x in {1,2,3,4,5} \draw (\x*1.5,-2.9) -- (\x*1.5,-3.1) node[below] {};
+
+		% top brace
+		\draw [brace_top] (1*1.5+0.03,-2.7) -- node [above, pos=0.5] {$ block_{n}$} 	(3*1.5-0.03,-2.7);
+		\draw [brace_top] (3*1.5+0.03,-2.7) -- node [above, pos=0.5] {$ block_{n+1}$} 	(5*1.5-0.03,-2.7);
+
+	    \draw  node[fill,circle,scale=0.6]at (3*1.5,-3.0)  {};%
+
+		% low brace period
+
+		\draw [brace_bottom] (1*1.5+0.03,-3.3) -- node [below, pos=0.5] {$e_{n}=d$} 	(3*1.5-0.03,-3.3);
+		\draw [brace_bottom] (3*1.5+0.03,-3.3) -- node [below, pos=0.5] {$d$} 		(5*1.5-0.03,-3.3);
+
+	\end{tikzpicture}
+		\caption{Practical maximum playing error}
+	\end{figure}
+
+	For playing video with an higher velocity we used \emph{ffmpeg}\footnote{\url{https://www.ffmpeg.org/}(accessed: 17 March 2016)} to convert the block into a new video with the desired velocity and seek time.
 Because the media duration is known, when the video started to being convert the headers located at the begining of the file were already written and that made it possible to stream while converting.
 
-	{\color{red} [TALK ABOUT COMPLEXITY OF DECODE VS DECODE+MANIPULATE+ENCODE]}
+	%{\color{red} [TALK ABOUT COMPLEXITY OF DECODE VS DECODE+MANIPULATE+ENCODE]}
 
@@
 
-\subsubsection{Server side recording to database}
+\subsection{Server side recording to database}
 

### 2016-03-20 tese2/abstract.tex
[ADD] template2

\chapter*{Abstract}

% DO NOT CHANGE THIS - Add entry in the table of contents as chapter
\addcontentsline{toc}{chapter}{Abstract}

\input{01_abstract.tex}

\cleardoublepage

### 2016-03-20 tese2/implementation.tex
[ADD] template2

\chapter{Implementation}
\label{chapter:implementation}

%\section{Implementation Options}
%Neste secção devem apresentar as opções de implementação que tinham ao vosso
%dispor, avaliá-las e justificar a escolha que realizaram. Isto pode englobar:
%\begin{itemize}
%    \item Simuladores
%    \item Linguagens e ambientes de programação
%    \item Sistemas operativos
%    \item Hardware
%\end{itemize}
%Fica sempre bem na avaliação colocar uma tabela com as características
%pretendidas e as que são satisfeitas pelas várias opções (tipo catálogo com as
%características dos automóveis). A escolha deve surgir naturalmente, com base na
%opção que tem mais cruzes\ldots

%\section{Architecture}
%Nesta secção devem explicar como implementaram a vossa solução, apresentando
%as simplificações que efectuaram, face ao modelo inicialmente previsto. As
%simplificações devem ser devidamente justificadas. Se for possível, devem indicar
%que estas não põem em causa as contribuições da tese.
%Podem ainda descrever os principais problemas que tiveram e a forma como os
%abordaram e resolveram.

%Se estiverem a usar um simulador devem:

%\begin{itemize}
%    \item explicar o funcionamento do simulador
%    \item explicar as alterações e modelos que desenvolveram no simulador e que permitem validar a vossa ideia
%\end{itemize}

%Se estiveram a desenvolver SW, sem simulador devem:
%\begin{itemize}
%    \item explicar os módulos, interfaces, estruturas de dados, etc\ldots
%\end{itemize}

%Sempre que possível, ilustrem a arquitectura com figuras que demonstrem a
%evolução face à arquitectura da secção anterior. Isto é, usem as figuras anteriores
%e façam as modificações necessárias à obtenção da arquitectura do protótipo.
%\ldots

\input{13_implementation_model.tex}
\input{14_implementation_signaling.tex}
\input{15_implementation_recording.tex}

### 2016-03-20 tese2/introduction.tex
[ADD] template2

--- 2016-01-10 tese1/intro.tex
@@
-\subsection{Context}   % English
+\chapter{Introduction}
+\label{chapter:introduction}
+%The introduction is certainly the most read section of any deliverable, and it largely determines the attitude of the reader/reviewer will have toward the work.
+Therefore, it is probably the most delicate part of the writing of a report.
+
+\section{Background}
+\label{section:background}
+%Nesta secção devem descrever a área em que a vossa tese se insere, de forma a contextualizar o problema que vão resolver. No âmbito da área devem identificar os principais problemas que existem.
+
 	Since the early days of Human History, we tried to communicate over far locations, from smoke signals to letters delivered by messengers.
@@
 
-	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice's capacity to conveying rich and complex social and emotional meaning} , although nothing replace the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
-The concept of structured voice and video synchronized with hypermedia is called hypervoice.
-        %RP a citação do geddes faz sentido? Parece estar a falar de voz gerada por computador e não de video conferencia onde o que ouvimos são outras pessoas!
+	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice's capacity to conveying rich and complex social and emotional meaning} , although nothing replaces the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
+The concept of structured voice and video synchronized with hypermedia is called hypervoice\cite{geddes}.
 
-\subsection{Problem Statement} % English
+\section{Proposed Solution}
+\label{section:proposed}
 
 	As communications technologies appeared, we adapted the way we communicate.
-This project don't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
+This project doesn't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
+
+	{\color{blue}With the advent of WebRTC, it became possible to develop video conference web apllications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.}
+
+    Real-time communication applications can make a difference on business, education and health sectors by providing tools for teaching and learning online, teamworking and socializing.
 
@@
                %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
-        A real-time system is an huge source of information that requires much attention from its users, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
-%RP estes dois ultimos pontos são melhores.
-Não fales em pessoas com problemas de memoria mas em termos de lidar com grandes quantidades de informação, realizarmos comunicações em grupo (ainda nãofalaste diss) e nem sempre ser possível estar presente.
-        Real-time communication applications can make a difference on business, education and health sectors by providing tools for teaching and learning online, teamworking and socializing.
-        %RP também há outros mercados, como online learning, projectos de longa duração que vão evoluindo e onde é preciso revisitar o que aconteceu antes.
-Fins histórios.
-Necessidade de manter registos devido a requisitos legais/registar história.
+	A real-time system is a huge source of information that requires much attention from its users.
+An application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
 
-	This project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
+\section{Thesis Contribution}
+\label{section:contribution}
+%Nesta secção devem identificar como é que a vossa solução vai contribuir para resolver o problema.
+
+	Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
 creating annotations for improving content search) in order to increase its value.
-Another important aspect of this project is the ability to navigate in time and space by rewind communications, fast-forward, jump to certain points and choose another path.
-
-                %RP esta última frase está confusa.
-
-        %Estás muito parco nos objectivos do projecto.
-Acho que falta dizer que é um sistema para complementar/subsitutir actuais sistemas de comunicação audio/video real-time, adicionando-lhe a capacidade de: adicionar conteúdo posteriomente (e.g.
-anotações) por forma a aumentar o seu valor; permitir acesso posterior (mesmo que apenas 5 minutos depois para quem chega atrasado); permitir navegar na comunicação (hypermedia, saltos, fast-forward, rewind).
-
-\subsection{Thesis Goals} % English
+Another important goal of this project is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.
 
@@
 
-%RP podes elaborar um pouco.
-Objectivos são: realizar o levantamento do estado da arte no espaço do problema para fundamentar uma solução; apresentar uma solução e desenhar uma arquitectura para dar resposta ao problema apresentado; construir protótipo funcional; desenhar e realizar uma avaliação/testes que provem a adequação e validade dos conceitos desenvolvidos (elaborar); escrever um artigo.
-
-Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solves our problem either completely or just a part of it.
+	Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solve our problem either completely or just a part of it, not excluding solutions that may appear during the implementation of our project.
 
@@
 	All the problems faced during the development and limitations will be reported on the thesis so that a future project better then ours can be easily and better developed.
-        %RP esta última frase não é muito necessária
 
-\subsection{Document Structure} % English
-This document is structured as follows. Section \ref{related} presents the related work.
-%	Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
-%	Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
-%	Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \ac{WebRTC} specifications. 
-%	Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
-%	Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
-%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.   
-	In Section \ref{arch}, we propose an architecture for an Web Application that fulfills the goals of this thesis, including all the need infrastructure and software.
-	Section \ref{meth} describes our work methodology and our plan for implementing and validating our proposed architecture.
-	Section \ref{concl} presents the summary and conclusions of our work.
+\section{Outline}
+
+This document is structured as follows:
+
+\begin{itemize}
+\item \textbf{Chapter \ref{chapter:introduction}} presents the motivation, background and proposed solution.
+\item \textbf{Chapter \ref{chapter:relatedwork}} describes the previous work in the field.
+\item \textbf{Chapter \ref{chapter:architecture}} describes the system requirements and the architecture for an Web Application that fulfills the goals of this thesis.
+\item \textbf{Chapter \ref{chapter:implementation}} describes the implementation of our Web Application and the technologies chosen.
+\item \textbf{Chapter \ref{chapter:evaluation}} describes the evaluation tests performed and the corresponding results.
+\item \textbf{Chapter \ref{chapter:conclusion}} summarizes the work developed and future work.
+\end{itemize}
+
+\cleardoublepage

### 2016-03-25 tese2/07_related_work_nat.tex
[ADD] template2
[ADD] change architecture

--- 2016-03-12 tese1/07_related_work_nat.tex
@@
-\subsection{Early days of the Internet and its remaining flaws}\label{early}
+\section{Early days of the Internet and its remaining flaws}\label{early}
 
-The need to build a global communications network in an age when almost nobody had access to that technology and the number of future users was unpredictable, lead to some protocols not being suitable for the huge growth of users that followed.
+The need to build a global communications network in an age when almost nobody had access to that technology and the unpredictability of the number of future users, lead to some protocols not being suitable for the explosive growth and proliferation of users that followed.
 \ac{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}.
-One way to overcome this problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \ac{IP} address with a different port.
+{\color{blue} Today we live in the \emph{Internet of Things} era, users and companies may have multiple devices that require internet connectivity, from personal computers, smart phones, tablets, traffic sensors, public vigillance systems, health monitoring tools and so on, all those devices combined cannot be publicly addressed by \ac{IPv4}.}
+
+One way to overcome the \ac{IPv4} address scarcity problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \ac{IP} address with a different port.
 %RP o porto não identifica uma máquina (membro da rede privada) mas uma ligação.
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{figures/nat.png}
+	\includegraphics[width=0.7\textwidth]{figures/nat.pdf}
 	\caption{Network Address Translation}
@@
 	\centering
-	\includegraphics[width=0.30\textwidth]{figures/multinat.png}
-	\caption{Multiple level \ac{NAT}s}
+	\includegraphics[width=0.45\textwidth]{figures/multinat.pdf}
+	\caption{Multiple level NATs}
 \end{figure}
@@
 \ac{STUN} servers will then reply to the client, providing the mapped public address, so it knows its associated public \ac{IP} address and port.
-Symmetric \ac{NAT} changes \ac{IP} port for each different connection, for that reason, when the \ac{STUN} servers reply with the \ac{IP} address and port of their connection, it will be useless for clients to use in other connection connections.
+Symmetric \ac{NAT} changes \ac{IP} port for each different connection, for that reason, when the \ac{STUN} servers reply with the \ac{IP} address and port of their connection, it will be useless for clients to use them on other application connections.
 That is why Symmetric \ac{NAT} represents a problem for peer-to-peer communications.
@@
 
-\ac{TURN} uses public servers to relay traffic between private endpoints.
+On the other hand, \ac{TURN} uses public servers to relay traffic between private endpoints.
 It may use a \ac{P2P} network relay to find the best peer, but after that, the behavior is much like client-server.

### 2016-03-25 tese2/09_related_work_signaling.tex
[ADD] template2
[ADD] change architecture

--- 2016-03-15 tese1/09_related_work_signaling.tex
@@
-\subsection{Signaling: meet and get to know}
+\section{Signaling: meet and get to know}
 \label{signaling}
@@
   \ac{WebRTC} doesn't implement signaling, as different applications may require different protocols, and there is no single answer that fits all problems.
-  Multiple options are available for signaling, which can be performed using \ac{SIP}, \ac{XMPP}, \emph{WebSockets}, \emph{Socket.io} or by implementing a custom protocol.
+  Multiple options are available for signaling, which can be performed using \ac{SIP}, \ac{XMPP}, \emph{WebSockets}, \emph{Socket.io}\footnote{\url{http://socket.io/}(accessed June 1, 2015).} or by implementing a custom protocol.
+
+  \subsection{WebRTC}
 
@@
 
-  One of \ac{WebRTC} signaling's requisites is bi-directional communication over \ac{HTTP}.
+  One of \ac{WebRTC} signaling's requisites is bi-directional communication.
 %RP acrescentei o webrtc pois isto não é um requisito geral de signaling. era o que querias dizer?
@@
 As soon as it receives the reply, the client makes another request.
-Long polling technique results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which add data overhead, especially for short messages.
+Long polling technique results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \ac{HTTP} headers, which add data overhead and can be noticed especially for short sized messages.
 
-The WebSocket protocol allows bidirectional communications over a full-duplex socket channel \cite{rfc6455}.
+The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel \cite{rfc6455}, by other words it supports sending and receiving data simultaneously.
 %RP coloca as citações no final da frase. Pode haver excepções onde a citação apenas cobre parte da frase ou precisas de colocar várias a dizer coisas diferentes.
-WebSocket handshake phase specifies a \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, but the remainder messages are exchanged without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
-WebSockets may not be available on every web browser, frameworks like \emph{socket.io}\footnote{\url{http://socket.io/}(accessed June 1, 2015).} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}(accessed June 1, 2015).} fall back to using \ac{HTTP} when there is no support for WebSockets.
+ \emph{WebSocket} handshake phase specifies a \ac{HTTP} header in order to upgrade to \emph{WebSocket} type of communication, but the remainder messages are exchanged without \ac{HTTP} headers, which leads to much smaller messages and better network usage.
+WebSockets may not be available on every web browser, frameworks like \emph{socket.io} and \emph{SockJS}\footnote{\url{http://github.com/sockjs}(accessed June 1, 2015).} fall back to using \ac{HTTP} long polling when there is no support for \emph{WebSockets}.
 
@@
 
+  \subsection{Session Initiation Protocol}
+
   \ac{SIP} \cite{rfc3261} is a protocol used for negotiation, creation, modification and finalization of communication sessions between users.
 \ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as a signaling protocol.
-The advantage of \ac{SIP} is the ability to make video and voice call's applications over \ac{IP} networks.
+The advantage of \ac{SIP} is the ability to make video and voice calls between the telephone network and applications over \ac{IP} networks.
 
@@
   \ac{SIP} is used in \ac{VoIP} applications due to its compatibility with the \ac{PSTN}.
-  Service providers are making their \ac{SIP} infrastructures available through WebSockets.
+  Service providers are making their \ac{SIP} infrastructures available through WebSockets\footnote{\url{https://tools.ietf.org/html/draft-ietf-sipcore-sip-websocket-10} (accessed March 25, 2016)}.
 %RP a frase anterior é estranha
@@
 
+\subsection{Extensible Messaging and Presence Protocol}
+
   \ac{XMPP} was initially developed for instant messaging and presence (Jabber\footnote{\url{http://jabber.org/}(accessed June 1, 2015).}).
@@
   \ac{XMPP} messages are \ac{XML} based, which is attractive for applications that need structured messages and rich hypermedia.
-Another advantage of \ac{XMPP} is the addition of extensions, for example \cite{xep0096}, which adds file transfer capabilities between two entities and \cite{xep0045} which enables multi-user chat.
+Another advantage of \ac{XMPP} is the addition of extensions, for example \emph{XEP-0096}\cite{xep0096}, which adds file transfer capabilities between two entities and \emph{XEP-0045}\cite{xep0045} which enables multi-user chat.
   \ac{XMPP}'s bi-directional communication over \ac{HTTP} is achieved through \ac{BOSH} \cite{xep0206}.
@@
 
-    One of the advantages that \ac{XMPP} based applications offer is the possibility to send messages across different servers, an \ac{XMPP} user is identified by its \textit{Jabber ID} which is definined by the pair username and server name in the form \textit{username@server}.
+    One of the advantages that \ac{XMPP} based applications offer is the possibility to send messages across different servers.
+An \ac{XMPP} user is identified by its \textit{Jabber ID} which is definined by the pair username and server name in the form \textit{username@server}.
 
@@
 
-    An important aspect of this model is the ability to define a user as a set of attributes that can change dinamically in order to fit the IM applications needs.
+    An important aspect of this model is the ability to define a user as a set of attributes that can change dinamically in order to fit the \ac{IM} applications needs.
 
@@
 
-    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that XMPP already implemment, but from all the \ac{XMPP} features, we don't need all of them, in fact we just need a subset of \ac{XMPP} features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.
+    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that \ac{XMPP} already implemment, but from all the \ac{XMPP} features, we don't need all of them, in fact we just need a subset of \ac{XMPP} features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.
 }
+
+  \subsection{Signaling-On-the-fly}
 

### 2016-03-25 tese2/11_related_work_collaboration.tex
[ADD] change architecture

@@
 
-	\emph{ShareJS}\footnote{\url{http://sharejs.org/}(accessed June 2, 2015).} is an \ac{OT} \emph{JavaScript} library, developed by the ex \emph{Google Wave} engineer Joseph Gentle, for collaborative text and \ac{JSON} documents edition in real-time.
+    Among other platforms and libraries we present \emph{ShareJS}\footnote{\url{http://sharejs.org/}(accessed June 2, 2015).}, \emph{TogetherJS}\footnote{\url{http://togetherjs.com/}(accessed June 2, 2015).}, \emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}(accessed June 2, 2015).}, \emph{Etherpad Lite}\footnote{\url{https://github.com/ether/etherpad-lite}(Accessed 20 March 2016)} and \emph{otJS}\footnote{\url{http://operational-transformation.github.io}(accessed March 10, 2016)}.
+
+	\emph{ShareJS} is an \ac{OT} \emph{JavaScript} library, developed by the ex \emph{Google Wave} engineer Joseph Gentle, for collaborative text and \ac{JSON} documents edition in real-time.
 {\color{blue} It uses \emph{ShareDB} \footnote{\url{https://github.com/share/sharedb}(accessed March 10, 2016)} for its backend and data model, which supports simple integration with any database.
@@
 
-	\emph{TogetherJS}\footnote{\url{http://togetherjs.com/}(accessed June 2, 2015).} is a JavaScript library that uses \ac{WebRTC} for collaborative web applications.
+	\emph{TogetherJS} is a JavaScript library that uses \ac{WebRTC} for collaborative web applications.
 It uses \ac{JSON} messages for \ac{OT} concurrency control but it does not provide storage.
@@
 
-	\emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}(accessed June 2, 2015).} is a collaborative framework with its own server implementation, it supports four types of collaborative elements: String, Lists, Maps and Custom objects.
+	\emph{Goodow} is a collaborative framework with its own server implementation, it supports four types of collaborative elements: String, Lists, Maps and Custom objects.
         %RP estes 3 últimos parágrafos estão desligados do resto. O que pretendes apresentar?
 
-	{\color{blue}\emph{OT.js}\footnote{\url{http://operational-transformation.github.io}(accessed March 10, 2016)} is a \emph{JavaScript} library that only implements operation transformations on client side over plain text.
-An implication of implementing just the client side is the extra effort that is necessary to implement content's persistent storage, besides this drawback this library is very flexible because it's not tied to a specific database.}
+	{\color{blue}\emph{Etherpad Lite} is a collaborative text editor implemented with \emph{NodeJS} that allows not only the text operations but also to associate authorships to pieces of text.
+It supports adding functionality trought plugins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example:
+		\begin{itemize}
+		\item exporting and importing document formats such as \emph{DOC}, \emph{PDF}, \emph{ODT} and \emph{DOCX}.
+		\item painting and drawing.
+		\item video and audio chat using WebRTC.
+		\item create slideshows.
+		\item spell checking.
+		\item text to speech.
+		\end{itemize}
+	}
 
-{\color{green} For ease of development we decided to use the same backend for \ac{OT} since our web server relies on \emph{Play Framework} with Java and \emph{ShareJS} relies on \emph{NodeJS}\footnote{\url{https://nodejs.org/en/(accessed March 10, 2016)}} }
+	{\color{blue}\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations on client side over plain text.
+An implication of implementing just the client side is the extra effort that is necessary to implement content's persistent storage, besides this drawback this library is very flexible because it's not tied to a specific database or server side technology.}
 
@@
 Goodow           & yes                 & yes                  & text+objects        \\ \hline
-%Etherpad        & yes                 & yes                  & text (not sure)     \\ \hline
+Etherpad Lite    & yes                 & yes                  & extendable 			    \\ \hline
 OT.js            & no                  & no                   & text                \\ \hline

### 2016-03-25 tese2/14_implementation_signaling.tex
[ADD] template2
[ADD] change architecture

--- 2016-03-18 tese1/XX_implementation_signaling.tex
@@
-\subsection{Signaling Protocol}
+\section{Signaling Protocol}
 
 The signaling protocol could be implemented using \ac{HTTP} messages but they would transport extra information such as \ac{HTTP} headers and would follow a request response signaling mechnism which would not be the best option, as multiple ice candidates can arrive at any time.
-For that reason bi-directional communications were prefered.
+Using a \ac{REST} \ac{API} over \ac{HTTP} for signaling and allowing other methods to be called at the same time would lead to open multiple \ac{TCP} connections at the same time.
+
+Instead if we create a \emph{WebSocket} \ac{API} we only need one \ac{TCP} connection and at the same provide bi-directional communications without additional headers.
 
@@
     \centering
-    \begin{subfigure}[b]{\linewidth}
-    	\includegraphics[width=\textwidth]{figures/signaling}
+    \begin{subfigure}{}
+    	\includegraphics[width=1.0\textwidth]{figures/signaling}
     \end{subfigure}

### 2016-03-25 tese2/16_implementation_content.tex
[ADD] change architecture

\section{Hyper Content}
	In this section we describe the algorithm we created to show synchronized interactive content to clients.

	\subsection{Synchronization}

	In our solution content is represented as simple text that can contain \ac{HTML}, \ac{CSS} or even \emph{JavaScript}, this content is displayed on defined intervals of time.

	Multiple contents can be displayed at the same time, in order to achieve that, we define layers below the video with the same size. Each layer is associated to the content.

	We have taken into account that the amount of content tends to grow with time and the user should only have access to a subset of content instead all of them, which would be very ineficcient.

	The content to display for each user depends on the user's position on the timeline, this position is given by an offset between the current time and the navigated time. If the user is watchiing the content in realtime this offset is constant so there is no need to synchronize this value.

	Each content is divided into two components, the \emph{start} and the \emph{end} which are sent to users. The \emph{start} component contains a timestamp, the content identification number and the content itself in form of text. The \emph{end} component only contains the timestamp and the content identifcation number.

	The content to return is given by the union of two content subsets:
	\begin{itemize}
		\item The intersection of the content interval and user's time.
		\item A subset of contents which starting time is immediately following the user's time.
	\end{itemize}

	The second subset is used for predicting which content the user will watch and avoid requests during its events.

	The content description messages that are sent to each user contains the contant itself on the form of a list of events and an atrtibute that specifies if the server has more content to return.

	When a user enters the conference room, imediatelly after the \emph{WebSocket} creation the server sends him the current content.

	The user receives the content, sorts all components by time and creates a set of events. All the events before the user's time are rendered and removed from the set of events. A timer is schedulled for the first component on the event set and the process repeats while the set is not empty.

	If the set is empty, there two options, if the server contained more content a new request for content is made and the proccess starts from the beggining, namely the server sends the correpondent content again. If the server has no more content the process is stopped until it sends more.

	The users will receive their contents from the server on five different situations:

	\begin{itemize}
		\item Conference room entrance (advertised by server).
		\item Set of events empty and server has more content (client requested).
		\item Content is created (advertised by server).
		\item Content is removed (advertised by server).
		\item User navigates to different point in time (client requested).
	\end{itemize}

	\subsection{Content creation}

	In order to create content the user has the option to write simple movie captions without wrtting any code, otherwise, as mentioned before, it can write \ac{HTML}, \ac{CSS} and \emph{JavaScript}. The definition of the content's starting and ending time by the user it's not an easy task. Defining the content's time to appear in realtime would require a previous user plan. Otherwise the user could make a speech and add the content for later reproduction.

	In order to help the content creator to show and synchronize its content in realtime we allow the user to encode its content into \emph{QR codes} and show it to the camera in real time.

	\ac{KMS} lets registring event handlers for \emph{QR code} detection. The component that detects \emph{QR codes} on \ac{KMS} is called periodically and fires the handlers with the decoded content. This mechanism does not detect if the \emph{QR code} enters or leaves the screen. We had to implement our own mechnism for detecting those events.

	Each user session in the server maintains a map with the contents that are present on the screen. In order to aplly our algorithm more efficiently we calculate the content's hash through the \emph{md5} method.

	If the hash is not present in the map it means the \emph{QR code} was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in realtime. By doing an analogy to the \emph{Nyquist} theorem we know that if the same \emph{QR code} is not detected after a time bigger then two periods we can conclude that the \emph{QR code} leaved the screen and we add the correspondent content into our database.

	Our main content synchronization mechanism is time based but with some programming knowledge it is possible to insert \emph{JavaScript} code that fires events on user interaction. For example, after a teacher's lecture, it is possible to show a quiz to the users in order to understand what they learned and then submit the data to the server for further analysis.

	\subsection{Security Concerns}

	Our solution is flexible on what kind of interactions are possible to the users in real time but allowing users to write \emph{Javascript} that is executed on the other users browser would attackers to misuse their resources and access to critical information.

	We could solve this problem easily by escaping any \emph{script} tag present on the content, but we would sacrifice the kind of interactions that are possible.

	By not allowing javascript we would need to implement all actions a priori and fire them when a type of message is received. We decided to ignore the security vulnarabilities that are exposed by evaluating \emph{javascript} because in order to offer the same interactions we would have to do an exaustive functionaly requirements gathering.

	Another way to solve this problem, which is not within the goals of this thesis, is to analyse the javascript code and detect if it is malicious.

	\subsection{Tags creation}

	Tags are simpler way save points in time and share them with other users. When a user enters in a conference room the server sends all tags so they can be displayed directly into the user timeline. Each time a tag is created all users are advertised so they can update their interfaces.

	\subsection{Content Search}

	Users can easily search for tags and contents and travel to their correspondent times. In the case of hyper content after handling the result from database we ignore \ac{HTML} tags, extract the text with \emph{Jsoup}\footnote{\url{http://jsoup.org/} (Accessed 21 March 2016)} and apply the query again. We extract the text from \ac{HTML} because accidentatly searching for text contained in \ac{HTML} tags would lead to incorrect results.

### 2016-03-25 tese2/abstract.tex
[ADD] change architecture

@@
 
-\input{01_abstract.tex}
+The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to communication and collaboration services.
+This paradigm allows to synchronize, structure and navigate communication content, providing a way to integrate social media content and collaborative tools into voice and video calls.
+Voice and image together can express emotions and expose creativity like no other medium can. With hypermedia concepts, we can add more value to voice and video communications.
+
+\ac{WebRTC} technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
+The nature of web browser applications already follow the hypermedia concept, which makes \ac{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
+{\color{blue}The web browser platform provides an abstraction layer that makes possible to create applications that run indenpendently from the operating system.}
+The native support of \ac{WebRTC} in operating systems extends its usage to outside the web browser, allowing to explore functionalities that web browsers are poor to support such as video recording and massive information storage.
+
+Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed later.
+This application will target the web browser platform, resorting to \ac{WebRTC}.
+
+{\color{blue}
+In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose and implement an architecture for an hyper-linked communication application based on \ac{WebRTC} and provide the results for its implementation and evaluation.
+}
+
+\vspace{1cm}
+
+% TODO 4 to 6 keywords;
+\textbf{\Large Keywords:} WebRTC, asynchronous, communications, collaboration
 

### 2016-03-25 tese2/architecture.tex
[ADD] change architecture

\chapter{Architecture}
\label{chapter:architecture}
%Esta secção deve englobar:

%\begin{itemize}
%    \item Uma análise de requisitos, onde devem “abrir” portas para as principais
%        características da vossa arquitectura.
%    \item Uma descrição de alto nível da arquitectura do vosso sistema. Esta descrição
%        pode ser:
%        \begin{itemize}
%            \item A arquitectura de rede, explicando os principais componentes e as
%                funções que executam.
%            \item A arquitectura de SW da aplicação, explicando os principais
%                componentes e as funções que executam.
%        \end{itemize}
%    \item Descrições detalhadas de todos os componentes da vossa arquitectura.
%    \item As escolhas que efectuarem em termos de arquitectura global e de detalhe,
%        devem ser justificadas, face aos requisitos que identificaram.
%\end{itemize}

%Sempre que possível, ilustrem a arquitectura com figuras, que descrevam
%sucintamente o modelo como na Figure~\ref{fig:logoCNM}.

%\begin{figure}[!htb]
%  \centering
%  \includegraphics[width=0.25\textwidth]{Figures/logoCNM.png}
%  \caption[Caption for figure in TOC]{Caption for figure.}
%  \label{fig:logoCNM}
%\end{figure}

Taking into account the goals of this project and all the technology presented so far. Our proposal is the development of a web application that provides communication and collaboration in real-time.

The requirements for our web application are:

\begin{itemize}
 \item Text, Audio and Video communication using \ac{WebRTC}.
 \item Ability to record and playback interactive multimedia.
 \item Create annotations over video and voice content.
 \item Enrich communications by overlaying multiple media types.
 \item Search and navigate through annotations and contents.
 \item Ability to create and share time links among users.
 \item Mix multiple user streams into a single stream.
 \item Sound detection for showing the current speaker.
 \item Structure presentation of the content.
 \item Collaborative text edition.
 \item Easy interface for content creation and synchronization. 
\end{itemize}

	Clients can discover chat rooms and other clients by querying the signaling server, they can create rooms for multi-party communication which is achieved by using \ac{WebRTC}'s \emph{MediaStream} and \emph{DataChannel}, namely for audio, video and text communication.

	Chat rooms can be public or private. Public chat rooms are moderated by a group of clients which initially is formed by the room creator, this type of room has no access restrictions by default, but that can be changed by its moderators. Private chat rooms will be visible for a list of clients or can be accessed by clients that have a link for that chat room.

\section{Modules}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{figures/modules.pdf}
	\caption{System Modules}
        \label{fig:modules}
\end{figure}
%RP as figuras têm de ser apresentadas. Figure~\ref{fig:modules} presents ...
	Figure~\ref{fig:modules} presents the structure of our system which is divided into six modules:

\begin{itemize}

\item \underline{Application module} - responsible to provide information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client}.

\item \underline{Signaling module} - responsible for \emph{Client} and \emph{Stream} coordination.

 \item \underline{NAT Traversal module} - used by \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase which ends by establishing the connection between them.

 \item \underline{Stream module} - responsible to deliver and receive multimedia content from the \emph{Client}.

 \item \underline{Storage module} - responsible store all the model information and media recorded.

 \item \underline{Client module} - responsible for the interaction with the user.

\end{itemize}

% Each room will be associated to multiple Media-Types, interactive and non-interactive, discrete and continuous. This media types can be structured with chains of events. Some components of the user interface may not be synchronized to all participants, for example an help or suggestion window.

\section{Implementation Proposal}
The infrastructure is composed by: Ice Server, Web Server, Stream Server, Signaling Server and Test Clients.

\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{figures/infrastructure.pdf}
	\caption{System Infrastructure}
\end{figure}

In order to simplify our solution, we propose both the \emph{Application} and \emph{Signaling} modules implementation within the same server application so it could be easy to control which resources the client has permission to access.

We have taken into account the compatibility between the streaming server, database and the operation transformation solutions in order choose the appropriate framework to implement our web server.

Based on our study about signaling protocols we decided to exclude \ac{XMPP} due to the development difficulties that it exposes when using multiple web pages, namely the user re-authentication performed each time a page is loaded and the single tab usage limitation. By consequence we have not choosen \emph{Jitsi Video Brigde} as it uses \ac{XMPP}.

By excluding \emph{Jitsi Video Brigde} we have decided that our solution must use \ac{KMS}. Our web server could be implemented more easily with \emph{NodeJS} or \emph{Java} due to the fact \ac{KMS} provides clients for both technologies but others could be used as \ac{KMS} also exposes their \ac{API} via \emph{WebSockets}.

By default, \emph{Kurento Repository} is implemented over \emph{MongoDB}, for convenience our storage model will also use the same database.

For the web server we could choose \emph{NodeJS} or any \emph{Java} based web framework. One of our criteria for choosing the web framework is the ability to follow the \ac{MVC} paradigm which can help us to organize our code. Without any strong reason we have decided to implement our webserver with \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java}.

For the collaborative text editor, we have choosen \emph{OT.js} due to its indenpendence from server implementation and storage choice.

For the \emph{NAT Traversal} module an \ac{ICE} server is not required to be on our infrastructure, as a public \ac{STUN} server can be used for testing our solution.

On the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} should be installed as web browsers. Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Adapter.js}, \emph{OT.js} can be downloaded from the Web Server and executed on the client side.

\begin{table}[H]
\centering

    \begin{tabular}{cccccccc@{}m{0pt}@{}}
	\hline 
\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} &   \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[12pt]\hline
\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
\end{tabular}
	\caption{App Architecture}
	        \label{fig:apparch}
\end{table}

Figure~\ref{fig:apparch} presents the application architecture and the underlying technologies. \emph{Adapter.js} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.
\emph{Bootstrap} will be used to make the user interface more appellative and responsive. With \emph{Bootstrap} it is quite easy to develop applications that adapt to mobile devices with different screen sizes.

The synchronization between multimedia elements will be performed trough chains of \emph{JavasSript} events or by specifying the interval of time which time content must be visible. Other animations can be implemented with \ac{SVG} embedded on \ac{HTML}.

### 2016-03-25 tese2/frontCover.tex
[ADD] change architecture

\thispagestyle {empty}

% IST Logo
% parameters: bb=llx lly urx ury (bounding box), width=h_length, height=v_length, angle=angle, scale=factor, clip=true/false, draft=true/false. 
\newgeometry{left=2cm,top=2.5cm, right=2.5cm, bottom=0cm}
\includegraphics[width=5.0cm]{Logo.png}

\begin{center}
%
% Figure (Image or plot)
\vspace{0.3cm}
% TODO - Use of an image is optional
\hspace{-0.5cm}\includegraphics[height=50mm]{Figures/logo.png}
% TODO - If you don't use an image uncomment the next line
% \vspace{5.0cm}

% Title, author and degree
\vspace{0.8cm}
{\FontLb Hyper-linked Communications: WebRTC enabled asynchronous collaboration} \\
%\vspace{0.2cm}
%{\FontMn Subtítulo (facultativo)} \\
%\vspace{1.9cm}
\vspace{2.6cm}
{\FontMb Henrique Duarte Lopes Rocha} \\
\vspace{1.9cm}
{\FontLn Disserta\c{c}\~{a}o para a obten\c{c}\~{a}o de Grau de Mestre em} \\
\vspace{0.3cm}
{\FontLb Engenharia de Telecomunicações e Informática} \\
\vspace{1.9cm}
{\FontMb J\'{u}ri} \\
\vspace{0.3cm}
{\FontSn %
Presidente:        Nome do Presidente \\
Orientador:        Nome do Orientador \\
Co-orientador:     Nome do Co-orientador \\
Vogais:            Nome do Vogal 1 \\
                   Nome do Vogal 2 \\
}
\vspace{1.5cm}

{\FontMb M\^{e}s e Ano} \\
%
\end{center}

\cleardoublepage

\restoregeometry

### 2016-03-25 project/httpserver/app/main/KurentoManager.java

@@ -75,6 +75,14 @@ public class KurentoManager {
             return null;
         }
         Room room = rooms.get(groupId);
+        try {
+            if (room != null) {
+                room.getMediaPipeline().getName();
+            }
+        }catch (Exception e){
+            room = null;
+        }
+
         if (room == null) {
             MediaPipeline mp = kurento.createMediaPipeline();
             mp.setName(groupId);

### 2016-03-25 project/httpserver/public/this/js/hyper.socket.js

@@ -77,12 +77,12 @@ var HyperWebSocket = new (function() {
 				if(perc>0.05){
 					if(!soundDetected){
 						soundDetected = true;
-						HyperWebSocket.talk(true);
+						this.talk(true);
 					}
 				}else {
 					if(soundDetected){
 						soundDetected = false;
-						HyperWebSocket.talk(false);
+						this.talk(false);
 					}	
 				}
 			}
@@ -90,7 +90,7 @@ var HyperWebSocket = new (function() {
 			if(soundProc>4){
 				soundProc = 0;
 			}
-		};
+		}.bind(this);
 		
 		microphone.connect(javascriptNode);
 		javascriptNode.connect(audioContext.destination);
@@ -463,7 +463,7 @@ var HyperWebSocket = new (function() {
                                 }));
                             }, logError);
                         }, logError,remote_constraints);
-                    }).catch(logError);
+                    }.bind(this)).catch(logError);
 				}
 				else if(mode==2){
 					pc.createOffer(function (lsd) {

### 2016-03-26 project/stats/false_cpu.tex
[ADD] performance tests with and without qrcodes

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Cpu Usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=0, xmax=800,
	    ymin=0, ymax=600,
	    xtick={0,100,...,2000},
	    ytick={0,100,...,2000},
	    extra x ticks={10,70,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_7_mongod.csv};
		\addlegendentry{Mongod}

		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_67_kurento-media-server.csv};
		\addlegendentry{Kurento Media Server}

		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_35_kurento-repo.csv};
		\addlegendentry{Kurento Repository}

		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_233_bash.csv};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}

### 2016-03-26 project/stats/false_mem.tex
[ADD] performance tests with and without qrcodes

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Mem Usage},
	    xlabel={Time [seconds]},
	    ylabel={Memory [MB]},
	    xmin=0, xmax=800,
	    ymin=0, ymax=2000,
	    xtick={0,100,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={10,70,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/res_7_mongod.csv};
		\addlegendentry{Mongod}

		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/res_67_kurento-media-server.csv};
		\addlegendentry{Kurento Media Server}

		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/res_35_kurento-repo.csv};
		\addlegendentry{Kurento Repository}

		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-25,y index=1] {false/res_233_bash.csv};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}

### 2016-03-26 project/stats/false_net.tex
[ADD] performance tests with and without qrcodes

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Net Usage},
	    xlabel={Time [seconds]},
	    ylabel={NET [kB/s]},
	    xmin=0, xmax=800,
	    ymin=0, ymax=1000,
	    xtick={0,100,...,2000},
	    ytick={0,100,...,2000},
	    extra x ticks={10,70,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/sent_eth0.csv};
	\addlegendentry{Sent(eth0)}

	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/recv_eth0.csv};
	\addlegendentry{Recv(eth0)}

	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/recv_lo.csv};
	\addlegendentry{Sent/Recv(lo)}

	\end{axis}
\end{tikzpicture}

### 2016-03-26 project/stats/true_cpu.tex
[ADD] performance tests with and without qrcodes

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Cpu Usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=0, xmax=800,
	    ymin=0, ymax=1200,
	    xtick={0,100,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={10,70,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_7_mongod.csv};
		\addlegendentry{Mongod}

		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_67_kurento-media-server.csv};
		\addlegendentry{Kurento Media Server}

		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_35_kurento-repo.csv};
		\addlegendentry{Kurento Repository}

		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_233_bash.csv};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}

### 2016-03-26 project/stats/true_mem.tex
[ADD] performance tests with and without qrcodes

--- 2016-03-26 project/stats/false_mem.tex
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/res_7_mongod.csv};
+		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/res_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/res_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/res_35_kurento-repo.csv};
+		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/res_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-25,y index=1] {false/res_233_bash.csv};
+		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {true/res_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-26 project/stats/true_net.tex
[ADD] performance tests with and without qrcodes

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Net Usage},
	    xlabel={Time [seconds]},
	    ylabel={NET [kB/s]},
	    xmin=0, xmax=800,
	    ymin=0, ymax=1000,
	    xtick={0,100,...,1000},
	    ytick={0,100,...,1000},
	    extra x ticks={10,70,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/sent_eth0.csv};
	\addlegendentry{Sent(eth0)}

	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/recv_eth0.csv};
	\addlegendentry{Recv(eth0)}

	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/recv_lo.csv};
	\addlegendentry{Sent/Recv(lo)}

	\end{axis}
\end{tikzpicture}

### 2016-03-26 tese2/13_implementation_model.tex
[ADD] performance tests with and without qrcodes

@@
 
-During a group conversation it is possible to tag points in time for making it easy to access that time either by searching or sharing with other users.
+During a group conversation it is possible to create time annotations for making it easy to access that time either by searching or sharing with other users.
 
-A time tag contains a title, the correspondent group identification number and the time itself.
+A time annotation contains a title, the correspondent group identification number and the time itself.
 
@@
             \hline
-            \multicolumn{2}{|c|}{\textbf{TimeTag}}                   \\ \hline
+            \multicolumn{2}{|c|}{\textbf{TimeAnnotation}}                   \\ \hline
             $\Diamondblack$ \underline{id}  & ObjectId                      \\ 

### 2016-03-26 tese2/16_implementation_content.tex
[ADD] performance tests with and without qrcodes

@@
 
-	\subsection{Tags creation}
+	\subsection{Time annotations creation}
 
-	Tags are simpler way save points in time and share them with other users.
-When a user enters in a conference room the server sends all tags so they can be displayed directly into the user timeline.
-Each time a tag is created all users are advertised so they can update their interfaces.
+	Time annotations are simpler way save points in time and share them with other users.
+When a user enters in a conference room the server sends all annotations so they can be displayed directly into the user timeline.
+Each time annotation is created all users are advertised so they can update their interfaces.
 
@@
 
-	Users can easily search for tags and contents and travel to their correspondent times.
+	Users can easily search for annotations and contents and travel to their correspondent times.
 In the case of hyper content after handling the result from database we ignore \ac{HTML} tags, extract the text with \emph{Jsoup}\footnote{\url{http://jsoup.org/} (Accessed 21 March 2016)} and apply the query again.

### 2016-03-26 project/httpserver/app/controllers/WSController.java

@@ -7,7 +7,7 @@ import main.Tools;
 import main.UserSession;
 import models.RecordingInterval;
 import models.Message;
-import models.TimeTag;
+import models.TimeAnnotation;
 import models.User;
 import org.bson.Document;
 import org.bson.types.ObjectId;
@@ -66,8 +66,8 @@ public class WSController extends Controller {
 
                     try {
                         ListTimeTagsService service = new ListTimeTagsService(userId, groupId);
-                        List<TimeTag> tags = service.execute();
-                        for (TimeTag tag : tags) {
+                        List<TimeAnnotation> tags = service.execute();
+                        for (TimeAnnotation tag : tags) {
                             JSONObject msg = new JSONObject();
                             msg.put("cmd", "tag");
                             msg.put("data", tag.toJson());
@@ -251,7 +251,7 @@ public class WSController extends Controller {
                                         Date time = Tools.FORMAT.parse(args.getString("time"));
                                         String title = args.getString("title");
                                         CreateTimeTagService service = new CreateTimeTagService(groupId, time, title);
-                                        TimeTag tag = service.execute();
+                                        TimeAnnotation tag = service.execute();
                                         JSONObject msg = new JSONObject();
                                         msg.put("cmd", "tag");
                                         msg.put("data", tag.toJson());

### 2016-03-26 project/httpserver/app/services/DeleteTimeTagService.java

@@ -1,7 +1,7 @@
 package services;
 
 import exceptions.ServiceException;
-import models.TimeTag;
+import models.TimeAnnotation;
 import org.bson.types.ObjectId;
 
 
@@ -34,7 +34,7 @@ public class DeleteTimeTagService extends Service<Void> {
     public Void dispatch() throws ServiceException {
         ObjectId tid = ObjectId.isValid(tagId) ? new ObjectId(tagId):null;
         if(tid!=null) {
-            TimeTag tag = TimeTag.findById(tid);
+            TimeAnnotation tag = TimeAnnotation.findById(tid);
             tag.delete();
         }
         return null;

### 2016-03-26 project/httpserver/app/services/Service.java

@@ -19,10 +19,7 @@ public abstract class Service<T> {
     
     /** The files. */
     static GridFS files;
-    
-    /** The last oid. */
-    static long lastOid = 0L;
-    
+
     /** The db. */
     private static DB db;
     
@@ -40,7 +37,7 @@ public abstract class Service<T> {
      *
      * @return the client
      */
-    private static MongoClient getClient() {
+    private synchronized static MongoClient getClient() {
         if (client == null) {
             client = new MongoClient(ServerAddress.defaultHost());
         }
@@ -65,7 +62,7 @@ public abstract class Service<T> {
      * @return the db
      */
     @SuppressWarnings("deprecation")
-    private static DB getDB() {
+    private synchronized static DB getDB() {
         if (db == null) {
             db = Service.getClient().getDB(DB_NAME);
         }
@@ -77,7 +74,7 @@ public abstract class Service<T> {
      *
      * @return the database
      */
-    public static MongoDatabase getDatabase() {
+    public synchronized static MongoDatabase getDatabase() {
         if (database == null) {
             database = getClient().getDatabase(DB_NAME);
         }

### 2016-03-27 tese2/17_implementation_deploy.tex
[ADD] deployment and test description

\section{Solution deployment}
	\label{section:deployment}

	In this section we describe the hardware and software we have used to deploy our solution.

	\subsection{Hardware}

	The hardware was gently provided by INESC-ID\footnote{\url{http://www.inesc-id.pt/} (Accessed March 26, 2016)}, the specifications of the server that we used are specified on table \ref{table:hwspecs}.

	\begin{table}[H]
\centering
\caption{Hardware specifications}
\label{table:hwspecs}
\begin{tabular}{|l|l|}
\hline
Server & Supermicro SuperServer 6027R-72RF\footnotemark \\ \hline
CPU & 2 x Intel Xeon E5-2640V2, LGA 2011, 2.0GHz, 8C/16T \\ \hline
RAM & 8 x DDR3 REG16G-1600DDR3, 16GB, DDR3-1600, Registered ECC, memory \\ \hline
Network cards & 2 x Intel Corporation I350 Gigabit Network Connection (rev 01)  \\ \hline
Disks & \begin{tabular}[c]{@{}l@{}}2x SAMSUNG SSD 840 PRO 256GB SATA III (Drive 0 - RAID 1 - 237.486 GB)\\
4x WESTERN DIGITAL 3TB SATA III 64MB RED (Drive 1 - RAID 5 - 8.185TB)\end{tabular} \\ \hline
\end{tabular}
\end{table}

 \footnotetext{\url{http://www.supermicro.com/products/system/2U/6027/SYS-6027R-72RF.cfm} (Acessed March 26, 2016)}

\subsection{Operating System}

Below we present an overview of the operating system configurations of the machine we used for deployment.

	\begin{table}[H]
\centering
\caption{Operating system specifications}
\label{table:osspecs}
\begin{tabular}{|l|l|}
\hline
Operating System & Linux version 3.16.0-4-amd64 \\ \hline
Distribution & Debian GNU/Linux 8 (jessie)\footnotemark \\ \hline
Swap & 16GB (Drive 0)\\ \hline
Root & 40GB (Drive0)\\ \hline
EFI & 200MB (Drive 0)\\ \hline
Bcache & Remaining space (Drive 0 \& Drive 1) \\ \hline
\end{tabular}
\end{table}

\footnotetext{\url{https://www.debian.org/} (Accessed March 26, 2016)}

\subsection{Software}
	INESC-ID provided us a restricted linux account without administration permissions, which we prevented us the installation of our solution directly on the machine because we would need administrative priviledges to install all the software that our solution depends on. Although that limitation, they provided access to Docker\footnote{\url{https://www.docker.com/} (Accessed March 27, 2016)} which run on host operating system as an isolated process in userspace.

	Inside a Docker container we have the administrative priviledges to install all the software dependencies.

	We present, on table \ref{table:softspecs}, the software we have installed inside our docker container including the versions that are in use.

	\begin{table}[H]
\centering
\caption{Installed Software}
\label{table:softspecs}
\begin{tabular}{|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Name}} & \multicolumn{1}{c|}{\textbf{Version}}         \\ \hline
Ubuntu Server\footnotemark & 14.04 LTS   \\ \hline
Oracle Java & 1.8.0\_77   \\ \hline
MongoDB & 3.0.10   \\ \hline
Kurento Media Server & 6.4.0 \\ \hline
Kurento Repository & 6.3.1 \\ \hline
Python & 2.7.6 \\ \hline
\end{tabular}
\end{table}
 \footnotetext{\url{http://www.ubuntu.com/} (Accessed March 27, 2016)}

### 2016-03-27 tese2/evaluation.tex
[ADD] deployment and test description

\chapter{Evaluation}
\label{chapter:evaluation}
  In the previous chapter we described how we implemented our solution namely the data model that we used, the description of the signaling protocol, how we performed stream recording, how we overlayed interactive content to video and lastly how we deployed our solution.

  In this chapter we describe how we tested our solution, show and analyse the results in order to validate the contributions of this thesis.

\section{Tests Objectives}

  We have tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.

  We have also tested the performance of our solution by measuring the used resources. Those performance tests are crucial to ensure that our solution is in fact stable and users can use it timelessly without decreasing the quality of their experience.

\section{Tests Scenarios}

  In this section we describe both performance and usability test scenarios that we have applied.

  \subsection {Performance Tests}
      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects with a periodicity of one second; CPU, physical memory information realtive to each running process and network usage relative to each interface.

      We would like to collect network information relative to each process, which \emph{nethogs}\footnote{\url{https://raboof.github.io/nethogs/} (Accessed March 27, 2016)} provide but it could not capture the network usage of some processes. Although this would be useful for a deep analysis, we know which network interfaces are used to establish connections between processes as we show on table \ref{table:interfacemap}.

  \begin{table}[H]
\centering
\caption{Mapping of connections by network interface}
\label{table:interfacemap}
\begin{tabular}{|l|l|}
\hline
\multicolumn{1}{|c|}{\textbf{Interface}} & \multicolumn{1}{c|}{\textbf{Connection}}         \\ \hline
Loopback & Web Server $\leftrightarrow$ MongoDB  \\ \hline
Loopback & Web Server $\leftrightarrow$ Kurento Media Server  \\ \hline
Loopback & Kurento Repository $\leftrightarrow$ MongoDB  \\ \hline
Loopback & Kurento Repository $\leftrightarrow$ Kurento Media Server  \\ \hline
Ethernet & Web Server $\leftrightarrow$ Client\\ \hline
Ethernet & Kurento Media Server $\leftrightarrow$ Client\\ \hline
\end{tabular}
\end{table}

      The performance test scenario that we have defined consists on two phases, the first phase consists only on having users, with similar computer specifications, entering sequentially on the conference room, the second phase consists on the users leaving the conference room. Each event, joining and leaving, occurs with intervals of one minute.

  \subsection {Usability Tests}

      In order to the usability of our solution, we haver performed usability tests with the help of real users with different backgrounds and ages.

      We handled a guide to the users with five tasks to perform, the metrics we used for each task were: number of clicks, number of errors including a description and time spent.

      The tasks we asked to the user are:

      \begin{enumerate}
      \item Login into the system with the provided credentials and accept the received friendship request. Find the coordinator's brother and add him as a friend.

      \item Create a private conference room, enter and share your screen, add the coordinator to the conference room and chat with him. After that use the collaborative editor in order to write at the same time as the coordinator. Lastly save the editor and leave the conference room.

      \item Enter in the conference room correspondent to the third task as an observer and navigate to the specified annotation. Watch the video until a list of topics is overlayed in the video and choose one of them and leave the conference room.

      \item Enter in the conference room correspondent to the fourth task by sharing your camera and create a time annotaion in the instant of time when you entered, navigate to the current time, search for the annotation and leave the conference room.

      \item Enter in the conference room correspondent to the fourth task by sharing your camera and add a subtitle in the video, preview it, specify an interval of time and save it. Then show the provided \emph{QR code} to the camera and leave the leave the conference room.
      \end{enumerate}

      The goal of the first task consists on evaluating the interface for authentication and friendship management.

      The second task is used to make the users familiar with our tools for collaborative content edition.

      The goal of the third task is the demonstrate the navigation functionalities and synchronized interactive content superposition.

      The fourth task is used to analyse the behaviour and difficulties of the user when creating a time annotation.

      The goal of the fifth task is to create synchronized overlayed content, demonstrate the difficulties of such task and present an easier way to synchronize content.

\section{Test Results}
Nesta secção devem apresentar os resultados dos testes, quer sobre a forma
de tabelas, quer sobre a forma de gráficos. As tabelas e os gráficos devem ser
apresentados e depois analisados, detalhadamente.
\ldots

Here is an example of a table~\ref{table:simple}.

\begin{table}[!htb]
  \begin{center}
  \caption[Table caption shown in TOC]{Table caption}
    \begin{tabular}{|c|c|}
      \hline
      item 1 & item 2 \\
      \hline
      item 3 & item 4 \\
      \hline
    \end{tabular}
  \end{center}
  \label{table:simple}
\end{table}

### 2016-03-27 tese2/stats/test_full_features_cpu.tex
[ADD] deployment and test description

--- 2016-03-26 project/stats/true_cpu.tex
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_7_mongod.csv};
+		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/cpu_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_67_kurento-media-server.csv};
+		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/cpu_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_35_kurento-repo.csv};
+		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/cpu_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_233_bash.csv};
+		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/cpu_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-27 tese2/stats/test_full_features_mem.tex
[ADD] deployment and test description

--- 2016-03-26 project/stats/true_mem.tex
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/res_7_mongod.csv};
+		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/res_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/res_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/res_35_kurento-repo.csv};
+		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/res_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {true/res_233_bash.csv};
+		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/res_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-27 tese2/stats/test_full_features_net.tex
[ADD] deployment and test description

--- 2016-03-26 project/stats/true_net.tex
@@
 
-	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/sent_eth0.csv};
+	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/sent_eth0.csv};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/recv_eth0.csv};
+	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/recv_eth0.csv};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/recv_lo.csv};
+	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/recv_lo.csv};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-03-27 tese2/stats/test_without_qrcode_cpu.tex
[ADD] deployment and test description

--- 2016-03-26 project/stats/false_cpu.tex
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_7_mongod.csv};
+		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/cpu_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_67_kurento-media-server.csv};
+		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/cpu_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_35_kurento-repo.csv};
+		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/cpu_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_233_bash.csv};
+		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/cpu_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-27 tese2/stats/test_without_qrcode_mem.tex
[ADD] deployment and test description

--- 2016-03-26 project/stats/true_mem.tex
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/res_7_mongod.csv};
+		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/res_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/res_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/res_35_kurento-repo.csv};
+		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/res_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {true/res_233_bash.csv};
+		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/res_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-27 tese2/stats/test_without_qrcode_net.tex
[ADD] deployment and test description

--- 2016-03-26 project/stats/false_net.tex
@@
 
-	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/sent_eth0.csv};
+	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/sent_eth0.csv};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/recv_eth0.csv};
+	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/recv_eth0.csv};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/recv_lo.csv};
+	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {test_without_qrcode/recv_lo.csv};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-03-27 tese2/17_implementation_deploy.tex
[ADD] amazing explanation of network test results

@@
 
-Below we present an overview of the operating system configurations of the machine we used for deployment.
+On table \ref{table:osspecs} we present an overview of the operating system configurations of the machine we used for deployment.
 
@@
 
+	We could use one docker image for each component but in order to reduce the network usage we prefered to install all the components within the same image.
+At the same we provide an all-in-one easy to deploy solution as all the \ac{IP}s are local.
+
 	We present, on table \ref{table:softspecs}, the software we have installed inside our docker container including the versions that are in use.

### 2016-03-27 tese2/architecture.tex
[ADD] amazing explanation of network test results

@@
 	\centering
-	\includegraphics[width=0.7\textwidth]{figures/modules.pdf}
+	\includegraphics[width=0.8\textwidth]{figures/modules.pdf}
 	\caption{System Modules}
@@
 \centering
-
+	\caption{Application Architecture}
+	\label{table:apparch}
     \begin{tabular}{cccccccc@{}m{0pt}@{}}
@@
 \multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
-\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} &   \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[12pt]\hline
+\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{~~~~~ot.js~~~~~} & \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} & \multicolumn{1}{c|}{adapter.js} &   \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} &\\[12pt]\hline
 \multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
 \end{tabular}
-	\caption{App Architecture}
 	        \label{fig:apparch}

### 2016-03-27 tese2/evaluation.tex
[ADD] amazing explanation of network test results

@@
 Loopback & Web Server $\leftrightarrow$ Kurento Media Server  \\ \hline
-Loopback & Kurento Repository $\leftrightarrow$ MongoDB  \\ \hline
-Loopback & Kurento Repository $\leftrightarrow$ Kurento Media Server  \\ \hline
+Loopback & Kurento Media Server $\leftrightarrow$ Kurento Repository $\leftrightarrow$ MongoDB  \\ \hline
 Ethernet & Web Server $\leftrightarrow$ Client\\ \hline
@@
 
-      The performance test scenario that we have defined consists on two phases, the first phase consists only on having users, with similar computer specifications, entering sequentially on the conference room, the second phase consists on the users leaving the conference room.
-Each event, joining and leaving, occurs with intervals of one minute.
+      The performance test scenario that we have defined consists on two phases, the first phase consists only on having users, with similar computer and network specifications, entering sequentially on the conference room, the second phase consists on the users leaving the conference room.
+Each event, joining and leaving, occurs with intervals of one minute in total of thirteen minutes (780 seconds).
 
@@
 
-      The goal of the third task is the demonstrate the navigation functionalities and synchronized interactive content superposition.
+      The goal of the third task is to demonstrate the navigation functionalities and synchronized interactive content superposition.
 
@@
 \section{Test Results}
-Nesta secção devem apresentar os resultados dos testes, quer sobre a forma
-de tabelas, quer sobre a forma de gráficos. As tabelas e os gráficos devem ser
-apresentados e depois analisados, detalhadamente.
-\ldots
+  \subsection {Performance Tests}
 
-Here is an example of a table~\ref{table:simple}.
+      We are presenting in this sections the results of our performance tests after implementing our solution.
 
-\begin{table}[!htb]
+      Every time a user shares its camera or its screen in the context of a conference room, its offered video is sent to the server and independently from watching an individual stream or the mixed version the server just sends a unique stream back to the user.
+
+      From the media server prespective if there are $n$ clients connected each of them sending and receiving one stream, it is expected that server sends and receives also $n$ streams.
+By so we expect that the amount of network increases lineraly as users join a confrence room.
+
+      Figure \ref{fig:test_full_features_net} confirms our expectations, each vertical yellow line represents each event, the first seven events are users entering the conference room the next ones represent users leaving the conversation.
+
+\begin{figure}[H]
   \begin{center}
-  \caption[Table caption shown in TOC]{Table caption}
-    \begin{tabular}{|c|c|}
-      \hline
-      item 1 & item 2 \\
-      \hline
-      item 3 & item 4 \\
-      \hline
-    \end{tabular}
+    \input{stats/test_full_features_net.tex}
   \end{center}
-  \label{table:simple}
-\end{table}
+  \caption{Network usage after implementing all features}
+  \label{fig:test_full_features_net}
+\end{figure}
+
+      The blue peaks are caused by the signaling phase and web page downloads, including resources such as images, stylesheets and javascript files. 
+      The green peaks are caused by that being transfered between \ac{KMS} and \emph{MongoDB} through \emph{Kurento Repository} each peak occurs each time a block of video is recorded which in this case is every ten seconds.
+      The recordings are synchronized so all user and mixed blocks starts and ends at the same time, that's why the amount of work done every ten seconds acumulates and because this is poerformed locally the maximum transfer rate is limited by the performance of the memory as buffers are writen to buffers then to disks.
+      Client's sent data transfer rate has no significant peaks as signaling information contains few information.
+
+      Figure \ref{fig:summary_full_net} shows the average transmission rate per interval of consecutive events.
+
+\begin{figure}[H]
+  \begin{center}
+    \input{stats/summary_full_net.tex}
+  \end{center}
+  \caption{Average network usage per interval of events after implementing all features}
+  \label{fig:summary_full_net}
+\end{figure}
+
+If we consider all streams equal and an incoming stream uses $x$KB/s, with $n$ incoming streams the rate of data received at \ac{KMS} is $nx$ and the expected rate of data transfered to \emph{Kurento Repo} using localhost is $(n+1)x$.
+Due to the data being transfered also from \emph{Kurento Repo} to \emph{MongoDB} via localhost, the expected total rate of data transfered through localhost is $2((n+1)x) = (2n+2)x$, which explains the reason for the average amount of data transfered through localhost being more than twice the average amount of data received from users.
+
+With this results we conclude that if we want to scale our solution's storage using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \emph{MongoDB} should be installed in the same machine because the loopback interface can handle bigger transfer rates than the remaining network interfaces.
+For the same reason another possible configuration for scaling the storage is to install both \emph{Kurento Repository} and \ac{KMS} on the same machine.

### 2016-03-27 tese2/stats/summary_full_net.tex
[ADD] amazing explanation of network test results

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=8cm,
	    title={Network usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [kB/s]},
	    xmin=-10, xmax=790,
	    ymin=0, ymax=2000,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

	\addplot[color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_eth0.csv};
	\addlegendentry{Sent(eth0)}

	\addplot[color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_recv_eth0.csv};
	\addlegendentry{Recv(eth0)}

	\addplot[color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_lo.csv};
	\addlegendentry{Sent/Recv(lo)}

	\end{axis}
\end{tikzpicture}

### 2016-03-27 tese2/stats/test_full_features_cpu.tex
[ADD] amazing explanation of network test results

--- 2016-03-26 project/stats/true_cpu.tex
@@
 	    ylabel={CPU [percentage]},
-	    xmin=0, xmax=800,
+	    xmin=-10, xmax=790,
 	    ymin=0, ymax=1200,
-	    xtick={0,100,...,2000},
+	    xtick={0,60,...,2000},
 	    ytick={0,200,...,2000},
-	    extra x ticks={10,70,...,1000},
+	    extra x ticks={0,60,...,2000},
     	extra x tick labels={,,},
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_7_mongod.csv};
+		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_67_kurento-media-server.csv};
+		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_35_kurento-repo.csv};
+		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {true/cpu_233_bash.csv};
+		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_233_bash.csv};
 		\addlegendentry{Java}
@@
 	\end{axis}
+	\caption{Percentage of CPU used during the performance tests}
+	\label{fig:test_full_features_cpu}
 \end{tikzpicture}

### 2016-03-27 tese2/stats/test_full_features_mem.tex
[ADD] amazing explanation of network test results

--- 2016-03-26 project/stats/true_mem.tex
@@
 	    ylabel={Memory [MB]},
-	    xmin=0, xmax=800,
+	    xmin=-10, xmax=790,
 	    ymin=0, ymax=2000,
-	    xtick={0,100,...,2000},
+	    xtick={0,60,...,2000},
 	    ytick={0,200,...,2000},
-	    extra x ticks={10,70,...,1000},
+	    extra x ticks={0,60,...,2000},
     	extra x tick labels={,,},
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/res_7_mongod.csv};
+		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/res_35_kurento-repo.csv};
+		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {true/res_233_bash.csv};
+		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-27 tese2/stats/test_full_features_net.tex
[ADD] amazing explanation of network test results

--- 2016-03-26 project/stats/true_net.tex
@@
 	    width=\textwidth,
-	   	height=6cm,
-	    title={Net Usage},
+	   	height=8cm,
+	    title={Network usage},
 	    xlabel={Time [seconds]},
-	    ylabel={NET [kB/s]},
-	    xmin=0, xmax=800,
-	    ymin=0, ymax=1000,
-	    xtick={0,100,...,1000},
-	    ytick={0,100,...,1000},
-	    extra x ticks={10,70,...,1000},
+	    ylabel={Transfer rate [kB/s]},
+	    xmin=-10, xmax=790,
+	    ymin=0, ymax=2000,
+	    xtick={0,60,...,2000},
+	    ytick={0,200,...,2000},
+	    extra x ticks={0,60,...,2000},
     	extra x tick labels={,,},
@@
 
-	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/sent_eth0.csv};
+	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/sent_eth0.csv};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/recv_eth0.csv};
+	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/recv_eth0.csv};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/recv_lo.csv};
+	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/recv_lo.csv};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-03-27 tese2/stats/test_without_qrcode_cpu.tex
[ADD] amazing explanation of network test results

--- 2016-03-26 project/stats/false_cpu.tex
@@
 	    ylabel={CPU [percentage]},
-	    xmin=0, xmax=800,
+	    xmin=-10, xmax=790,
 	    ymin=0, ymax=600,
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_7_mongod.csv};
+		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_67_kurento-media-server.csv};
+		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_35_kurento-repo.csv};
+		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_233_bash.csv};
+		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-27 tese2/stats/test_without_qrcode_mem.tex
[ADD] amazing explanation of network test results

--- 2016-03-26 project/stats/true_mem.tex
@@
 	    ylabel={Memory [MB]},
-	    xmin=0, xmax=800,
+	    xmin=-10, xmax=790,
 	    ymin=0, ymax=2000,
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/res_7_mongod.csv};
+		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/res_35_kurento-repo.csv};
+		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-112,y index=1] {true/res_233_bash.csv};
+		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-27 tese2/stats/test_summary.py

@@ -1,0 +1,37 @@
+# top -b -d 1 > top.txt
+# ifstat -t -i wlan0
+import re
+import sys
+import json
+
+
+shift = int(sys.argv[1]) if len(sys.argv)>0 else 0 
+
+
+summary_time = None
+
+
+sum = 0
+len = 0
+
+for line in sys.stdin:
+	tokens = line.split()
+	time = long(tokens[0])
+	value = float(tokens[1])
+
+	if summary_time is None:
+		summary_time = time
+		
+	sum = sum + value
+	len = len + 1
+	
+	if (time - shift) % 60 == 0:
+		new_value = sum/len
+	
+		print str(summary_time) + " " + str(new_value)
+		print str(time) + " " + str(new_value)
+		sum = 0
+		len = 0
+			
+		summary_time = time
+

### 2016-03-28 tese2/14_implementation_signaling.tex
[ADD] performance tests, updated playFW, websocket simplification

@@
 
-Our signaling protocol consists in send and receiving JSON formated messages over WebSockets by both the server and the client.
+Our signaling protocol consists in send and receiving \ac{JSON} formated messages like listing \ref{lst:msgtype} over WebSockets by both the server and the client.
 
-\begin{lstlisting}[caption={General structure of our WebSocket messages},language=json]
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={General structure of our WebSocket messages},label={lst:msgtype},language=json]
 {
@@
 \end{lstlisting}
+\end{minipage}
 
@@
 If the user decides to share either from camera of screen, \emph{getUserMedia} is called with the correspondent contraints in order to obtain a local stream.
+Use the constraints presented on listing \ref{lst:constraints}.
 
-\begin{lstlisting}[caption={Media constraints},language=JavaScript]
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Media constraints},label={lst:constraints},language=JavaScript]
 var screenShareConstraints = {	
@@
 \end{lstlisting}
+\end{minipage}
 
 A popup is raised in order to ask the user to give permission to share those resources.
-If the resources are shared successfully the user creates an offer, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through WebSockets.
+If the resources are shared successfully the user creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through WebSockets.
 If the resources are not shared or the user specified to receive stream only, an offer is created specifying the contraints for receiving only video and audio.
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Offer created by client},label={lst:sig01},language=json]
+{
+	"cmd":"offer",
+	"data":{
+		"type":"offer",
+		"sdp":<sdp>	// this is huge
+	}
+}
+\end{lstlisting}
+\end{minipage}
 
@@
 The server receive and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint.
-Then a \emph{local session description} is created on the server and sent back to the client, after that the server trys to gather \ac{ICE} candidates.
+Then a \emph{local session description} like listing \ref{lst:sig02} is created on the server and sent back to the client, after that the server trys to gather \ac{ICE} candidates.
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Answer created by KMS},label={lst:sig02},language=json]
+{
+	"cmd":"answer",
+	"data":{
+		"type":"answer",
+		"sdp":<sdp>	// this is huge
+	}
+}
+\end{lstlisting}
+\end{minipage}
 
@@
 The candidates are received at the client which sets them to its \emph{PeerConnection}.
-The same is done on the server which receives the \ac{ICE} candidates from the client and propagates them to \ac{KMS}.
+The same is done on the server which receives the \ac{ICE} candidates like on listing \ref{lst:sig03} from the client and propagates them to \ac{KMS}.
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={ICE candidates sent by KMS and client},label={lst:sig03},language=json]
+{
+	"cmd":"iceCandidate",
+	"data":{
+		"sdpMLineIndex":0,
+		"candidate":"candidate:15 1 TCP 843056127 146.193.224.20 48828 typ srflx raddr 192.168.1.105 rport 48828 tcptype passive",
+		"sdpMid":"audio"
+	},
+}
+\end{lstlisting}
+\end{minipage}
 
@@
 
-\begin{figure}[H]
+\begin{figure}[!htb]
     \centering
     \begin{subfigure}{}
-    	\includegraphics[width=1.0\textwidth]{figures/signaling}
+    	\includegraphics[width=0.8\textwidth]{figures/signaling}
     \end{subfigure}

### 2016-03-28 tese2/16_implementation_content.tex
[ADD] performance tests, updated playFW, websocket simplification

@@
 
+	Listing \ref{lst:pseudo_render} presents the pseudo code for our content scheduler.
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Pseudo code for hyper content scheduler},label={lst:pseudo_render},language=JavaScript]
+function scheduleContent() {
+    var navigatedTime = getCurrentTime() - getUserOffset();
+    while ( hasLocalEvents() ) {
+        if ( firstEventIsOlderThanNavigatedTime() ) {
+            if ( firstEventIsStart() ) {
+            	addHtmlLayer(event);
+            } else {
+            	removeHtmlLayer(event);
+            }
+            removeFirstEvent();
+        } else {
+            break;
+        }
+    }
+    if ( hasNoLocalEvents() ) {
+    	return;	// nothing to do
+    } 
+    if ( hasNoStartEvents() && serverHasMoreContent() ) { 
+       	requestContentFromServer(navigatedTime);
+    } else {
+        waitUntilNextEvent();
+        scheduleContent();
+    }
+}
+\end{lstlisting}
+\end{minipage}
+
 	\subsection{Content creation}

### 2016-03-28 tese2/17_implementation_deploy.tex
[ADD] performance tests, updated playFW, websocket simplification

@@
 
-	\begin{table}[H]
+	\begin{table}[!htb]
 \centering
@@
 Oracle Java & 1.8.0\_77   \\ \hline
+Play Framework & 2.5.0  \\ \hline
 MongoDB & 3.0.10   \\ \hline

### 2016-03-28 tese2/evaluation.tex
[ADD] performance tests, updated playFW, websocket simplification

@@
 
-  \begin{table}[H]
+  \begin{table}[!htb]
 \centering
@@
 
-\begin{figure}[H]
+\begin{figure}[!htb]
   \begin{center}
@@
 For the same reason another possible configuration for scaling the storage is to install both \emph{Kurento Repository} and \ac{KMS} on the same machine.
+
+Figure \ref{fig:test_full_features_mem} shows the memory usage during our performance test case.
+Both \ac{JVM}, \emph{MongoDB} and \ac{KMS} performs their own memory management by holding and recycling objects when needed.
+The expected and observed behaviour of the memory usage is growth memory usage while the users are entering in the conference room and a memory usage stabilization aftwerwards.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \input{stats/test_full_features_mem.tex}
+  \end{center}
+  \caption{Memory usage after implementing all features}
+  \label{fig:test_full_features_mem}
+\end{figure}
+
+\emph{MongoDB} memory usage keeps increasing because it trys to fit part of the database on ram for fast read access.
+\emph{MongoDB} checkpoints data to disk every 60 seconds or when journal data exceeds 2GB\footnote{\url{https://docs.mongodb.org/manual/faq/storage/}(Accessed March 28, 2016)}, that explains the small memory usage peaks during our test case and stabilization at the end.
+
+Figure \ref{fig:test_full_features_cpu} shows the percentage of \ac{CPU} usage during our performance test case.
+As we can see, the amount of \ac{CPU} increases and decreases linearly in function of the amount of conference participants.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \input{stats/test_full_features_cpu.tex}
+  \end{center}
+  \caption{Percentage of CPU used during the performance tests}
+  \label{fig:test_full_features_cpu}
+\end{figure}
+
+Figure \ref{fig:test_full_features_cpu_zoom} is the same as figure \ref{fig:test_full_features_cpu} but zoomed over \emph{MongoDB}, \emph{Java} and \emph{Kurento Repository}.
+As we can see there are periodic proccessing peaks every ten seconds that are due to block recording.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \input{stats/test_full_features_cpu_zoom.tex}
+  \end{center}
+  \caption{Percentage of CPU used during the performance tests}
+  \label{fig:test_full_features_cpu_zoom}
+\end{figure}
+
+   Figure \ref{fig:summary_full_cpu} shows the average used \ac{CPU} per interval of consecutive events.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \input{stats/summary_full_cpu.tex}
+  \end{center}
+  \caption{Average percentage of CPU used per interval of events after implementing all features (zoomed)}
+  \label{fig:summary_full_cpu}
+\end{figure}
+
+  Just for testing purposes we have done the same performance tests without detecting \emph{QR codes} in order to understand how \ac{CPU} intensive is this task.
+Figure \ref{fig:test_without_qrcode_cpu} shows the results for the same experience but with \emph{QR code} detection disabled.
+
+  \begin{figure}[!htb]
+  \begin{center}
+    \input{stats/test_without_qrcode_cpu.tex}
+  \end{center}
+  \caption{Percentage of CPU used during the performance tests without QR code detection}
+  \label{fig:test_without_qrcode_cpu}
+\end{figure}
+
+We conclude that \emph{QR code} detection is a very intensive task, approximately doubling the amount of work performed by \ac{CPU}.
+Without this feature the network and memory usage had insignificate changes compared to \ac{CPU} usage.

### 2016-03-28 tese2/stats/summary_full_cpu.tex
[ADD] performance tests, updated playFW, websocket simplification

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={CPU usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [kB/s]},
	    xmin=-10, xmax=850,
	    ymin=0, ymax=1400,
	    xtick={0,60,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_mongo.csv};
		\addlegendentry{Mongod}

		\addplot[color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kms.csv};
		\addlegendentry{Kurento Media Server}

		\addplot[color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kr.csv};
		\addlegendentry{Kurento Repository}

		\addplot[color=magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_play.csv};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}

### 2016-03-28 tese2/stats/summary_full_net.tex
[ADD] performance tests, updated playFW, websocket simplification

@@
 	    width=\textwidth,
-	   	height=8cm,
+	   	height=6cm,
 	    title={Network usage},
@@
 	    ylabel={Transfer rate [kB/s]},
-	    xmin=-10, xmax=790,
-	    ymin=0, ymax=2000,
+	    xmin=-10, xmax=850,
+	    ymin=0, ymax=1400,
 	    xtick={0,60,...,2000},

### 2016-03-28 tese2/stats/test_full_features_cpu.tex
[ADD] performance tests, updated playFW, websocket simplification

@@
 	   	height=6cm,
-	    title={Cpu Usage},
+	    title={CPU usage},
 	    xlabel={Time [seconds]},
 	    ylabel={CPU [percentage]},
-	    xmin=-10, xmax=790,
+	    xmin=-10, xmax=850,
 	    ymin=0, ymax=1200,
@@
 	\end{axis}
-	\caption{Percentage of CPU used during the performance tests}
-	\label{fig:test_full_features_cpu}
+
 \end{tikzpicture}

### 2016-03-28 tese2/stats/test_full_features_cpu_zoom.tex
[ADD] performance tests, updated playFW, websocket simplification

--- 2016-03-28 tese2/stats/test_full_features_cpu.tex
@@
 	    xmin=-10, xmax=850,
-	    ymin=0, ymax=1200,
+	    ymin=0, ymax=100,
 	    xtick={0,60,...,2000},
-	    ytick={0,200,...,2000},
+	    ytick={0,20,...,2000},
 	    extra x ticks={0,60,...,2000},

### 2016-03-28 tese2/stats/test_full_features_mem.tex
[ADD] performance tests, updated playFW, websocket simplification

@@
 	   	height=6cm,
-	    title={Mem Usage},
+	    title={Memory Usage},
 	    xlabel={Time [seconds]},
 	    ylabel={Memory [MB]},
-	    xmin=-10, xmax=790,
-	    ymin=0, ymax=2000,
+	    xmin=-10, xmax=850,
+	    ymin=0, ymax=1800,
 	    xtick={0,60,...,2000},

### 2016-03-28 tese2/stats/test_without_qrcode_mem.tex
[ADD] performance tests, updated playFW, websocket simplification

@@
 	    ylabel={Memory [MB]},
-	    xmin=-10, xmax=790,
+	    xmin=-10, xmax=850,
 	    ymin=0, ymax=2000,

### 2016-03-28 project/httpserver/public/adapter/adapter.js

@@ -1,1710 +1,673 @@
-(function(f){if(typeof exports==="object"&&typeof module!=="undefined"){module.exports=f()}else if(typeof define==="function"&&define.amd){define([],f)}else{var g;if(typeof window!=="undefined"){g=window}else if(typeof global!=="undefined"){g=global}else if(typeof self!=="undefined"){g=self}else{g=this}g.adapter = f()}})(function(){var define,module,exports;return (function e(t,n,r){function s(o,u){if(!n[o]){if(!t[o]){var a=typeof require=="function"&&require;if(!u&&a)return a(o,!0);if(i)return i(o,!0);var f=new Error("Cannot find module '"+o+"'");throw f.code="MODULE_NOT_FOUND",f}var l=n[o]={exports:{}};t[o][0].call(l.exports,function(e){var n=t[o][1][e];return s(n?n:e)},l,l.exports,e,t,n,r)}return n[o].exports}var i=typeof require=="function"&&require;for(var o=0;o<r.length;o++)s(r[o]);return s})({1:[function(require,module,exports){
-/*
- *  Copyright (c) 2016 The WebRTC project authors. All Rights Reserved.
- *
- *  Use of this source code is governed by a BSD-style license
- *  that can be found in the LICENSE file in the root of the source
- *  tree.
- */
-'use strict';
+/*! adapterjs - v0.13.2 - 2016-03-18 */
 
-// Shimming starts here.
-(function() {
-  // Utils.
-  var logging = require('./utils').log;
-  var browserDetails = require('./utils').browserDetails;
-  // Export to the adapter global object visible in the browser.
-  module.exports.browserDetails = browserDetails;
-  module.exports.extractVersion = require('./utils').extractVersion;
-  module.exports.disableLog = require('./utils').disableLog;
+// Adapter's interface.
+var AdapterJS = AdapterJS || {};
 
-  // Uncomment if you do not want any logging at all including the switch
-  // statement below. Can also be turned off in the browser via
-  // adapter.disableLog(true) but then logging from the switch statement below
-  // will still appear.
-  //require('./utils').disableLog(true);
+// Browserify compatibility
+if(typeof exports !== 'undefined') {
+  module.exports = AdapterJS;
+}
 
-  // Warn if version is not supported regardless of browser.
-  // Min version can be set per browser in utils.js
-  if (browserDetails.version < browserDetails.minVersion) {
-    logging('Browser: ' + browserDetails.browser + ' Version: ' +
[…]

### 2016-03-28 project/httpserver/public/adapter/adapter.screenshare.js

@@ -1,4 +1,4 @@
-/*! adapterjs - v0.13.1 - 2016-03-15 */
+/*! adapterjs - v0.13.2 - 2016-03-18 */
 
 // Adapter's interface.
 var AdapterJS = AdapterJS || {};
@@ -17,7 +17,7 @@ AdapterJS.options = AdapterJS.options || {};
 // AdapterJS.options.hidePluginInstallPrompt = true;
 
 // AdapterJS version
-AdapterJS.VERSION = '0.13.1';
+AdapterJS.VERSION = '0.13.2';
 
 // This function will be called when the WebRTC API is ready to be used
 // Whether it is the native implementation (Chrome, Firefox, Opera) or
@@ -3167,7 +3167,10 @@ if ( navigator.mozGetUserMedia ||
 
     AdapterJS.getUserMedia = getUserMedia =
        window.getUserMedia = navigator.getUserMedia;
-    navigator.mediaDevices.getUserMedia = requestUserMedia;
+    if ( navigator.mediaDevices &&
+      typeof Promise !== 'undefined') {
+      navigator.mediaDevices.getUserMedia = requestUserMedia;
+    }
   }
 
   // For chrome, use an iframe to load the screensharing extension

### 2016-03-28 project/httpserver/public/this/js/group.content.js

@@ -59,11 +59,9 @@ function removeContent(key) {
 
 function renderContent() {
     var currentTime = timeline.getCurrentTime() - timeline.hyper_offset;
-
     while (localContent.length != 0) {
         var entry = localContent[0];
         var time = new Date(entry.time);
-
         if (time.getTime() <= currentTime) {
             localContent.shift();
             if (entry.type == "start") {
@@ -72,7 +70,6 @@ function renderContent() {
                 removeContent(entry.id);
             }
             console.log("render",entry,localContent);
-
         } else {
             break;
         }
@@ -84,18 +81,14 @@ function renderContent() {
             hasStarts = true;
             break;
         }
-
     }
     if (localContent.length == 0) {
         // no content, keep calm
-    } else if (!hasStarts) {
-        if(moreContent){
-            // all events started, check if others may start
-            HyperWebSocket.getContent();
-        }else {
-            // there are ends
-            scheduleRender(currentTime);
-        }
+        return;
+    }
+    if (!hasStarts && moreContent) {
+        // all events started, check if others may start
+        HyperWebSocket.getContent();
     } else {
         // wait for next event
         scheduleRender(currentTime);

### 2016-03-28 project/httpserver/public/this/js/hyper.socket.js

@@ -33,6 +33,7 @@ var HyperWebSocket = new (function() {
     // ================== UTILITIES ==================
     // ===============================================
 
+
     function logError(err) {
         console.log(err);
     }
@@ -295,114 +296,131 @@ var HyperWebSocket = new (function() {
 
 		if ("WebSocket" in window) {
 			ws = new WebSocket(wsurl("/ws/room/" + groupId));
-		
+
+            ws.callbacks = [];
+		    ws.onCmd = function(cmd,callback){
+                ws.callbacks.push({cmd:cmd,callback:callback});
+                return ws;
+		    };
+
 			ws.onerror = function() {
 				console.log("Error!");
 			};
-			
+
+			ws.onCmd('iceCandidate',function(message){
+	            console.log(message);
+				var candidate = new RTCIceCandidate(message.data);
+				pc.addIceCandidate(candidate, function() {
+
+				}, logError);
+			});
+
+		    ws.onCmd('answer',function(message){
+                console.log(message);
+                var rsd = new RTCSessionDescription(message.data);
+                // XXX [CLIENT_OFFER_08] XXX
+                pc.setRemoteDescription(rsd, function(){
+                    // console.log("setRemoteSessionDescription",rsd);
+                },logError);
+			});
+
+		    ws.onCmd('participants',function(message){
+	            if(participantPresenceHandler){
+    				for(var userId in message.data){
+					    participantPresenceHandler(message.data[userId]);
+					}
+				}
+			});
+
+		    ws.onCmd('removedUser',function(message){
+	            if(removedUserHandler){
+                    removedUserHandler(message.uid);
+                }
+			});
+
+	        ws.onCmd('content',function(message){
+                if(hyperContentArrivedHandler){
+                    hyperContentArrivedHandler(message.data,message.more);
+                }
+			});
+
+	        ws.onCmd('rec',function(message){
+                if(videoRecordingHandler){
+                    delete message.cmd;
+                    videoRecordingHandler(message);
+                }
+			});
+
+	        ws.onCmd('msg',function(message){
+                for(var i in message.data){
+                    var item = message.data[i];
+                    console.log("msg",item);
+                    if(messageArrivedHandler){
+                        messageArrivedHandler(item.source,item.time,item.text,item.name, item.id, item.seq);
+                    }
+                }
+			});
+
+	        ws.onCmd('tag',function(message){
+                if(timeTagArrivedHandler){
+                     var item = message.data;
+                     timeTagArrivedHandler(item.id,item.time,item.title);
+                }
+			});
+
[…]

### 2016-03-29 tese2/16_implementation_content.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)

@@
 	In this section we describe the algorithm we created to show synchronized interactive content to clients.
+
+	\subsection{Content creation}
+
+	In order to create content the user has the option to write simple movie captions without wrtting any code, otherwise, as mentioned before, it can write \ac{HTML}, \ac{CSS} and \emph{JavaScript}.
+The definition of the content's starting and ending time by the user it's not an easy task.
+Defining the content's time to appear in realtime would require a previous user plan.
+Otherwise the user could make a speech and add the content for later reproduction.
+
+	Figure \ref{fig:creation} shows the user interface for creating interactive hyper content manually.
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=0.6\textwidth]{figures/edition.png}
+		\caption{Manual content creation}
+		\label{fig:creation}
+	\end{figure}
+
+	Listing \ref{lst:createcontent} shows the structure of the content created by one user.
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Exampe of content created by one user},label={lst:createcontent},language=json]
+{
+	"cmd":"createContent",
+	"start":"2016-03-29T03:56:40.000Z",
+	"end":"2016-03-29T03:56:41.000Z",
+	"content":"<h1>Hello</h1>"
+}
+\end{lstlisting}
+\end{minipage}
+	In order to help the content creator to show and synchronize its content in realtime we allow the user to encode its content into \emph{QR codes} and show it to the camera in real time.
+
+	\ac{KMS} lets registring event handlers for \emph{QR code} detection.
+The component that detects \emph{QR codes} on \ac{KMS} is called periodically and fires the handlers with the decoded content.
+This mechanism does not detect if the \emph{QR code} enters or leaves the screen.
+We had to implement our own mechnism for detecting those events.
+
+	Each user session in the server maintains a map with the contents that are present on the screen.
+In order to aplly our algorithm more efficiently we calculate the content's hash through the \emph{md5} method.
+
+	If the hash is not present in the map it means the \emph{QR code} was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in realtime.
+By doing an analogy to the \emph{Nyquist} theorem we know that if the same \emph{QR code} is not detected after a time bigger then two periods we can conclude that the \emph{QR code} leaved the screen and we add the correspondent content into our database.
+Listing \ref{lst:pseudo_qrcode} shows the pseudo code for \emph{QR code} leaving detection.
+
+\begin{minipage}[!htb]{\linewidth}
+\begin{lstlisting}[caption={Pseudo code for QR code leaving detection},label={lst:pseudo_qrcode},language=JavaScript]
+var ongoingCodesMap = {};
+function onCodeFound(content) {
+	var startingTime = getCurrentTime() - getUserOffset();
+	var hash = md5(content);
+	if(ongoingCodesMap[hash] == null){
+		sendQRCodeToEditingUser(hash,content);
+	}
+	ongoingCodesMap[hash] = startingTime;
+	waitTwoSeconds(); // onCodeFound() may be called from other threads
+	var newestTime = ongoingCodesMap[hash]; 
+	if(newestTime == startingTime){
+		var endingTime = getCurrentTime() - getUserOffset();
+		ongoingCodesMap[hash] = null; 
+		sendQRCodeToEditingUser(hash,null); // remove from UI
+		insertContentIntoDatabase(startingTime,endingTime,content);
+		sendContentToUsers(startingTime,endingTime,content);
+	}
+}
+\end{lstlisting}
+\end{minipage}
+
+	Our main content synchronization mechanism is time based but with some programming knowledge it is possible to insert \emph{JavaScript} code that fires events on user interaction.
+For example, after a teacher's lecture, it is possible to show a quiz to the users in order to understand what they learned and then submit the data to the server for further analysis.
 
@@
 
+	Listing \ref{lst:sentcontent} shows the structure of the content sent to users.
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Exampe of content sent to users},label={lst:sentcontent},language=json]
+{
+	"cmd":"content",
+	"data":[{
+			"id":"56f9fd1aa986c615fab43d69",
+			"time":"2016-03-29T03:56:40.000Z",
+			"type":"start",
+			"content":"<h1>Hello</h1>"
+		},{
+			"id":"56f9fd1aa986c615fab43d69",
+			"time":"2016-03-29T03:56:41.000Z",
+			"type":"end"
+		}
+	],
+	"more":false
+}
+\end{lstlisting}
+\end{minipage}
+
 	When a user enters the conference room, imediatelly after the \emph{WebSocket} creation the server sends him the current content.
@@
 
-\begin{minipage}{\linewidth}
+\begin{minipage}[!htb]{\linewidth}
 \begin{lstlisting}[caption={Pseudo code for hyper content scheduler},label={lst:pseudo_render},language=JavaScript]
@@
 
-	\subsection{Content creation}
-
-	In order to create content the user has the option to write simple movie captions without wrtting any code, otherwise, as mentioned before, it can write \ac{HTML}, \ac{CSS} and \emph{JavaScript}.
-The definition of the content's starting and ending time by the user it's not an easy task.
-Defining the content's time to appear in realtime would require a previous user plan.
-Otherwise the user could make a speech and add the content for later reproduction.
-
-	In order to help the content creator to show and synchronize its content in realtime we allow the user to encode its content into \emph{QR codes} and show it to the camera in real time.
-
-	\ac{KMS} lets registring event handlers for \emph{QR code} detection.
-The component that detects \emph{QR codes} on \ac{KMS} is called periodically and fires the handlers with the decoded content.
-This mechanism does not detect if the \emph{QR code} enters or leaves the screen.
-We had to implement our own mechnism for detecting those events.
-
-	Each user session in the server maintains a map with the contents that are present on the screen.
-In order to aplly our algorithm more efficiently we calculate the content's hash through the \emph{md5} method.
-
-	If the hash is not present in the map it means the \emph{QR code} was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in realtime.
-By doing an analogy to the \emph{Nyquist} theorem we know that if the same \emph{QR code} is not detected after a time bigger then two periods we can conclude that the \emph{QR code} leaved the screen and we add the correspondent content into our database.
-
-	Our main content synchronization mechanism is time based but with some programming knowledge it is possible to insert \emph{JavaScript} code that fires events on user interaction.
-For example, after a teacher's lecture, it is possible to show a quiz to the users in order to understand what they learned and then submit the data to the server for further analysis.
-
 	\subsection{Security Concerns}
@@
 
+	\subsection{Time Manipulation}
+
+	We have used \emph{vis.js} to display out timeline.
+This library was created for content navigation through time, but it was not designed to be always moving automatically, we have created a background timer that performs the animation of moving the window of time bounds and user navigated time marker.
+
+	Figure \ref{fig:timeline} shows the the graphic appearance of our interactive timeline, in order to navigate through time the user must drag and drop the timeline horizontaly, when the user drops the timeline an event handler is called with the new user's time offset which will be used to send a message to the server in order to choose the correspondent content to display.
+
+	\begin{figure}[!htb]
+		\centering
+		\includegraphics[width=\textwidth]{figures/timeline.png}
+		\caption{Interactive timeline}
+		\label{fig:timeline}
+	\end{figure}
+
+	We have noticed the server time being different from client time and this created graphic inconsistencies such as showing existing recorded blocks of movie in the future.
+To solve this problem the server sends its time to client imediately after the websocket creation, we synchronize the timeline with the server and although it may exist a small error due to network transmission time the graphical error is negligible.
+
 	\subsection{Time annotations creation}
@@
 
+	Figure \ref{fig:annotation} shows the creation and placement of annotations on the timeline, to save the annotation the user must click on the floppy icon in order to save it and advertise the conference participants.
+
+	\begin{figure}[!htb]
+		\centering
+		\includegraphics[width=0.7\textwidth]{figures/annotation.png}
+		\caption{Creating a time annotation}
+		\label{fig:annotation}
+	\end{figure}
+
+	Besides the ability to create tags it is also possible to create time hyperlinks, see figure \ref{fig:timelink}, that can be sent to other users externaly so they can navigate directly to the content when entering in the conference room.
+
+	\begin{figure}[!htb]
+		\centering
+		\includegraphics[width=0.8\textwidth]{figures/timelink.png}
+		\caption{Time link creation}
+		\label{fig:timelink}
+	\end{figure}
+
 	\subsection{Content Search}
@@
 We extract the text from \ac{HTML} because accidentatly searching for text contained in \ac{HTML} tags would lead to incorrect results.
+
+	Figure \ref{fig:search} shows how the search results are displayed to the user. Each result entry contains an icon specifying the type of result (hyper content, time annotation, ...).
+
+\begin{figure}[!htb]
+\centering
+\begin{minipage}[b]{0.55\linewidth}
+\centering
+
+		\includegraphics[width=\textwidth]{figures/search.png}
+	      a) Result list
+\label{fig:minipage1}
+\end{minipage}
+\quad
+\begin{minipage}[b]{0.40\linewidth}
+		\centering
+
+		\includegraphics[width=\textwidth]{figures/search2.png}
+	       b) Selected result
+\label{fig:minipage2}
+\end{minipage}
+
+		\caption{Example of search results}
+		\label{fig:search}
+\end{figure}

### 2016-03-29 tese2/stats/summary_full_cpu.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)

@@
 
-		\addplot[color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_mongo.csv};
+		\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_mongo.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kms.csv};
+		\addplot[color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kms.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kr.csv};
+		\addplot[color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kr.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[color=magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_play.csv};
+		\addplot[color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_play.csv};
 		\addlegendentry{Java}

### 2016-03-29 tese2/stats/summary_full_net.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)

@@
 
-	\addplot[color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_eth0.csv};
+	\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_eth0.csv};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_recv_eth0.csv};
+	\addplot[color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_recv_eth0.csv};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_lo.csv};
+	\addplot[color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_lo.csv};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-03-29 tese2/stats/test_full_features_cpu.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)

@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_7_mongod.csv};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_67_kurento-media-server.csv};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_35_kurento-repo.csv};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_233_bash.csv};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-29 tese2/stats/test_full_features_mem.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)

@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_7_mongod.csv};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_35_kurento-repo.csv};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_233_bash.csv};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-29 tese2/stats/test_full_features_net.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)

--- 2016-03-26 project/stats/true_net.tex
@@
 	   	height=6cm,
-	    title={Net Usage},
+	    title={Network usage},
 	    xlabel={Time [seconds]},
-	    ylabel={NET [kB/s]},
-	    xmin=0, xmax=800,
-	    ymin=0, ymax=1000,
-	    xtick={0,100,...,1000},
-	    ytick={0,100,...,1000},
-	    extra x ticks={10,70,...,1000},
+	    ylabel={Transfer rate [kB/s]},
+	    xmin=-10, xmax=850,
+	    ymin=0, ymax=1400,
+	    xtick={0,60,...,2000},
+	    ytick={0,200,...,2000},
+	    extra x ticks={0,60,...,2000},
     	extra x tick labels={,,},
@@
 
-	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {true/sent_eth0.csv};
+	\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/sent_eth0.csv};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {true/recv_eth0.csv};
+	\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/recv_eth0.csv};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {true/recv_lo.csv};
+	\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/recv_lo.csv};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-03-29 tese2/stats/test_without_qrcode_cpu.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)

--- 2016-03-26 project/stats/false_cpu.tex
@@
 	   	height=6cm,
-	    title={Cpu Usage},
+	    title={CPU usage},
 	    xlabel={Time [seconds]},
 	    ylabel={CPU [percentage]},
-	    xmin=0, xmax=800,
+	    xmin=-10, xmax=850,
 	    ymin=0, ymax=600,
@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_7_mongod.csv};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_67_kurento-media-server.csv};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_35_kurento-repo.csv};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-25,y index=1] {false/cpu_233_bash.csv};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-29 tese2/stats/test_without_qrcode_mem.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)

@@
 
-		\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_7_mongod.csv};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_7_mongod.csv};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_67_kurento-media-server.csv};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_35_kurento-repo.csv};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_35_kurento-repo.csv};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=magenta] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_233_bash.csv};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_233_bash.csv};
 		\addlegendentry{Java}

### 2016-03-29 tese2/stats/test_without_qrcode_net.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)

--- 2016-03-26 project/stats/false_net.tex
@@
 	    width=\textwidth,
-	   	height=6cm,
-	    title={Net Usage},
+	   	height=8cm,
+	    title={Network usage},
 	    xlabel={Time [seconds]},
-	    ylabel={NET [kB/s]},
-	    xmin=0, xmax=800,
-	    ymin=0, ymax=1000,
+	    ylabel={Transfer rate [kB/s]},
+	    xmin=-10, xmax=850,
+	    ymin=0, ymax=2000,
 	    xtick={0,100,...,2000},
@@
 
-	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/sent_eth0.csv};
+	\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/sent_eth0.csv};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/recv_eth0.csv};
+	\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/recv_eth0.csv};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/recv_lo.csv};
+	\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/recv_lo.csv};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-03-30 project/httpserver/public/this/js/search.group.js
[RFT] ui improvements

@@ -1,51 +1,0 @@
-
-
-var SearchGeneric = new (function() {
-	
-	this.source = function(query,process){
-        $.get( "/api/search?query="+query, function( data ) {
-          	console.log(data);
-            process(data.data);
-        });
-	};
-	
-	this.displayText = function(item){
-		return item.name;
-	};
-	
-	this.highlighter = function(item){
-		if(item.type=="models.User"){
-			return "<i class='fa fa-user'></i> "  + item.name;
-		}else if(item.type=="models.Group"){
-			return "<i class='fa fa-group'></i> " + item.name;
-		}else {
-			return JSON.stringify(item);
-		}
-	};
-	
-	this.matcher = function(item){
-		return true;
-	};
-	
-	this.updater = function(item){
-		console.log(item);
-		if(item.type=="models.User"){
-			document.location = "/user/"+item.id;
-		}else if(item.type=="models.Group"){
-			document.location = "/group/"+item.id;
-		}
-		return item;
-	}
-
-	return this;
-})();
-
-
-
-
-	
-	
-	
-	
-	
-		

### 2016-03-31 tese2/evaluation.tex
[ADD] too much stuff, very tired

@@
 
-If we consider all streams equal and an incoming stream uses $x$KB/s, with $n$ incoming streams the rate of data received at \ac{KMS} is $nx$ and the expected rate of data transfered to \emph{Kurento Repo} using localhost is $(n+1)x$.
+If we consider all streams equal and an incoming stream uses $x$kbps, with $n$ incoming streams the rate of data received at \ac{KMS} is $nx$ and the expected rate of data transfered to \emph{Kurento Repo} using localhost is $(n+1)x$.
 Due to the data being transfered also from \emph{Kurento Repo} to \emph{MongoDB} via localhost, the expected total rate of data transfered through localhost is $2((n+1)x) = (2n+2)x$, which explains the reason for the average amount of data transfered through localhost being more than twice the average amount of data received from users.
@@
 
-\emph{MongoDB} memory usage keeps increasing because it trys to fit part of the database on ram for fast read access.
-\emph{MongoDB} checkpoints data to disk every 60 seconds or when journal data exceeds 2GB\footnote{\url{https://docs.mongodb.org/manual/faq/storage/}(Accessed March 28, 2016)}, that explains the small memory usage peaks during our test case and stabilization at the end.
+\emph{MongoDB} memory usage keeps increasing because it trys to fit part of the database on \ac{RAM} for fast read access.
+\emph{MongoDB} checkpoints data to disk every 60 seconds or when journal data exceeds 2GB\footnote{\url{https://docs.mongodb.org/manual/faq/storage/}(Accessed March 28, 2016)}, that explains the small memory usage peaks during our test case.
+When the conference room is empty no there are no video recordings, which explains the memory stabilization at the end.
 
@@
 Without this feature the network and memory usage had insignificate changes compared to \ac{CPU} usage.
+
+  \subsection {Usability Tests}
+
+In this section we present the results of our usability tests.
+The first time we tested our solution by providing tasks to users, we have observed that our solution was not perfect.
+We have faced usability problems during our tests, he had to improve our solution's usability and start the tests again.
+
+  \subsubsection {First phase}
+
+In the first testing phase we have performed just three tests with users and ceased for improvements.
+
+Both users that tested our solution had difficulties on the first task when finding new friendship requests and adding new friends, as friendship request were possible to access only from the top bar and new user profiles were only accessible through the search bar.
+In order to solve this problem we have added a menu for adding new users in the friends list and another near the friends list as it can be seen on figure \ref{fig:test_ui_01_02_03}.
+
+\begin{figure}[!htb]
+\centering
+\begin{minipage}[b]{0.2\linewidth}
+\centering
+
+    \includegraphics[width=\textwidth]{figures/test_ui_01.png}
+        a) Friend list path
+\label{fig:minipage1}
+\end{minipage}
+\begin{minipage}[b]{0.35\linewidth}
+    \centering
+
+    \includegraphics[width=\textwidth]{figures/test_ui_02.png}
+         b) Friend requests path
+\label{fig:minipage2}
+\end{minipage}
+\begin{minipage}[b]{0.2\linewidth}
+    \centering
+
+    \includegraphics[width=\textwidth]{figures/test_ui_03.png}
+         c) Search bar path
+\label{fig:minipage2}
+\end{minipage}
+
+    \caption{Multiple paths for finding people}
+    \label{fig:test_ui_01_02_03}
+\end{figure}
+
+On the second task one user clicked on the friends list for adding a friend to room and by consequence made him leave the conference room to the user profile.
+We solved this problem by showing a popup with the user profile and an additional button to add the user to the conference room as it can be seen on figure \ref{fig:test_ui_04}.
+
+\begin{figure}[!htb]
+\centering
+\begin{minipage}[b]{0.7\linewidth}
+\centering
+
+    \includegraphics[width=\textwidth]{figures/test_ui_04.png}
+\end{minipage}
+
+    \caption{Adding user to conference room from user profile}
+    \label{fig:test_ui_04}
+\end{figure}
+
+On the third task we noticed that both users were not expecting to search the content directly from the search bar, instead they used the timeline.
+
+On the fourth task both users didn't placed their annotations on wright place, we only provide one way to place annotations, which is by dragging them on the timeline. 
+After creating an annotation users also didn't understand that they had to save their changes on the timeline, we solved this problem by saving automatically each time the user changed any annotation, the old user interface can be seen on figure \ref{fig:test_ui_05}.
+
+\begin{figure}[!htb]
+\centering
+\begin{minipage}[b]{0.3\linewidth}
+\centering
+
+    \includegraphics[width=\textwidth]{figures/test_ui_05.png}
+\end{minipage}
+
+    \caption{Save button that was removed}
+    \label{fig:test_ui_05}
+\end{figure}
+
+Both users had success with fifth task but they revealed difficulties choosing the time interval for the hyper content, we are aware that this is a difficult task and that's why we have introduced the \emph{QR code} content creation mechanism which was very easy to learn.
+
+We have also gathered comments and suggestions after letting the users explore our system, extra improvements were made such as:
+
+\begin{itemize}
+\item Giving feedback to user when saving content and collaborative editors.
+\item Support file upload on our chat.
+\item Starting with our timeline more zoomed in order to give more sensation of time.
+\item Create content with starting time and duration insted ending time.
+\end{itemize}
+
+  \subsubsection {Second phase}

### 2016-03-31 tese2/stats/summary_full_cpu.tex
[ADD] too much stuff, very tired

@@
 	    xlabel={Time [seconds]},
-	    ylabel={Transfer rate [kB/s]},
+	    ylabel={CPU [percentage]},
 	    xmin=-10, xmax=850,
-	    ymin=0, ymax=1400,
+	    ymin=0, ymax=1200,
 	    xtick={0,60,...,2000},
@@
 
-		\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_mongo.csv};
+		\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_mongo.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kms.csv};
+		\addplot[color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kms.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kr.csv};
+		\addplot[color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kr.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_play.csv};
+		\addplot[color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_play.dat};
 		\addlegendentry{Java}

### 2016-03-31 tese2/stats/summary_full_net.tex
[ADD] too much stuff, very tired

@@
 	    xlabel={Time [seconds]},
-	    ylabel={Transfer rate [kB/s]},
+	    ylabel={Transfer rate [mbps]},
 	    xmin=-10, xmax=850,
-	    ymin=0, ymax=1400,
+	    ymin=0, ymax=10,
 	    xtick={0,60,...,2000},
-	    ytick={0,200,...,2000},
+	    ytick={0,1,...,100},
 	    extra x ticks={0,60,...,2000},
@@
 
-	\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_eth0.csv};
+	\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/summary_full_sent_eth0.dat};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_recv_eth0.csv};
+	\addplot[color=Red] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/summary_full_recv_eth0.dat};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_sent_lo.csv};
+	\addplot[color=Green] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/summary_full_sent_lo.dat};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-03-31 tese2/stats/test_full_features_cpu.tex
[ADD] too much stuff, very tired

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_7_mongod.csv};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_67_kurento-media-server.csv};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_67_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_35_kurento-repo.csv};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_233_bash.csv};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_233_bash.dat};
 		\addlegendentry{Java}

### 2016-03-31 tese2/stats/test_full_features_mem.tex
[ADD] too much stuff, very tired

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_7_mongod.csv};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}/1000000] {stats/test_full_features/res_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}/1000000] {stats/test_full_features/res_67_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_35_kurento-repo.csv};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}/1000000] {stats/test_full_features/res_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_233_bash.csv};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}/1000000] {stats/test_full_features/res_233_bash.dat};
 		\addlegendentry{Java}

### 2016-03-31 tese2/stats/test_full_features_net.tex
[ADD] too much stuff, very tired

--- 2016-03-27 tese2/stats/test_full_features_net.tex
@@
 	   	height=6cm,
-	    title={Net Usage},
+	    title={Network usage},
 	    xlabel={Time [seconds]},
-	    ylabel={NET [kB/s]},
-	    xmin=0, xmax=800,
-	    ymin=0, ymax=1000,
-	    xtick={0,100,...,1000},
-	    ytick={0,100,...,1000},
-	    extra x ticks={10,70,...,1000},
+	    ylabel={Transfer rate [mbps]},
+	    xmin=-10, xmax=850,
+	    ymin=0, ymax=25,
+	    xtick={0,60,...,2000},
+	    ytick={0,5,...,100},
+	    extra x ticks={0,60,...,2000},
     	extra x tick labels={,,},
@@
 
-	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/sent_eth0.csv};
+	\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/sent_eth0.dat};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/recv_eth0.csv};
+	\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/recv_eth0.dat};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-112,y index=1] {test_full_features/recv_lo.csv};
+	\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/recv_lo.dat};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-03-31 tese2/stats/test_without_qrcode_cpu.tex
[ADD] too much stuff, very tired

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_7_mongod.csv};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-35,y expr=\thisrowno{1}] {stats/test_without_qrcode/cpu_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_67_kurento-media-server.csv};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-35,y expr=\thisrowno{1}] {stats/test_without_qrcode/cpu_67_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_35_kurento-repo.csv};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-35,y expr=\thisrowno{1}] {stats/test_without_qrcode/cpu_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/cpu_233_bash.csv};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-35,y expr=\thisrowno{1}] {stats/test_without_qrcode/cpu_233_bash.dat};
 		\addlegendentry{Java}

### 2016-03-31 tese2/stats/test_without_qrcode_mem.tex
[ADD] too much stuff, very tired

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_7_mongod.csv};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_67_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_35_kurento-repo.csv};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_233_bash.csv};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_233_bash.dat};
 		\addlegendentry{Java}

### 2016-03-31 tese2/stats/test_without_qrcode_net.tex
[ADD] too much stuff, very tired

--- 2016-03-26 project/stats/false_net.tex
@@
 	    width=\textwidth,
-	   	height=6cm,
-	    title={Net Usage},
+	   	height=8cm,
+	    title={Network usage},
 	    xlabel={Time [seconds]},
-	    ylabel={NET [kB/s]},
-	    xmin=0, xmax=800,
-	    ymin=0, ymax=1000,
+	    ylabel={Transfer rate [mbps]},
+	    xmin=-10, xmax=850,
+	    ymin=0, ymax=24,
 	    xtick={0,100,...,2000},
-	    ytick={0,100,...,2000},
+	    ytick={0,5,...,100},
 	    extra x ticks={10,70,...,1000},
@@
 
-	\addplot[smooth,color=blue] table[x expr=\thisrowno{0}-25,y index=1] {false/sent_eth0.csv};
+	\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-35,y expr=\thisrowno{1}*8/1000000] {stats/test_without_qrcode/sent_eth0.dat};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=red] table[x expr=\thisrowno{0}-25,y index=1] {false/recv_eth0.csv};
+	\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-35,y expr=\thisrowno{1}*8/1000000] {stats/test_without_qrcode/recv_eth0.dat};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[smooth,color=green] table[x expr=\thisrowno{0}-25,y index=1] {false/recv_lo.csv};
+	\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-35,y expr=\thisrowno{1}*8/1000000] {stats/test_without_qrcode/recv_lo.dat};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-03-31 project/httpserver/app/controllers/Application.java

@@ -266,21 +266,6 @@ public class Application extends Controller {
         }
     }
 
-    /**
-     * User profile.
-     *
-     * @param userId the user id
-     * @return the result
-     */
-    public Result userProfile(String userId) {
-        if (Tools.userExists(session("uid"))) {
-            boolean from = Relation.findByEndpoint(new ObjectId(session("uid")), new ObjectId(userId)) != null;
-            boolean to = Relation.findByEndpoint(new ObjectId(userId), new ObjectId(session("uid"))) != null;
-            return ok(views.html.profile.render(userId, from, to));
-        } else {
-            return ok(views.html.sign.render());
-        }
-    }
 
     /**
      * Join.

### 2016-03-31 project/httpserver/app/controllers/Rest.java

@@ -50,25 +50,41 @@ public class Rest extends Controller {
         return forbidden();
     }
 
+    /**
+     * User profile.
+     *
+     * @param userId the user id
+     * @return the result
+     */
+    public Result relationState(String userId) {
+        if (Tools.userExists(session("uid"))) {
+            boolean from = Relation.findByEndpoint(new ObjectId(session("uid")), new ObjectId(userId)) != null;
+            boolean to = Relation.findByEndpoint(new ObjectId(userId), new ObjectId(session("uid"))) != null;
+            JSONObject ans = new JSONObject();
+            ans.put("from",from);
+            ans.put("to",to);
+
+            return ok(ans.toString()).as(APP_JSON);
+        } else {
+            return forbidden();
+        }
+    }
 
     public Result uploadFile() throws ServiceException {
         MultipartFormData form = request().body().asMultipartFormData();
         JSONObject ans = new JSONObject();
 
-            if(session("uid")!=null){
-                for(FilePart fp : form.getFiles()){
-                    KeyValueFile kvf = new KeyValueFile(fp.getKey(),fp.getFilename(),fp.getFile());
-                    UploadFileService service = new UploadFileService(kvf);
-                    String url = service.execute();
-                    JSONObject obj =  new JSONObject();
-                    obj.put("filename",kvf.getName());
-                    obj.put("url",url);
-                    ans.put(kvf.getKey(),obj);
-                }
-
-
+        if(session("uid")!=null){
+            for(FilePart fp : form.getFiles()){
+                KeyValueFile kvf = new KeyValueFile(fp.getKey(),fp.getFilename(),fp.getFile());
+                UploadFileService service = new UploadFileService(kvf);
+                String url = service.execute();
+                JSONObject obj =  new JSONObject();
+                obj.put("filename",kvf.getName());
+                obj.put("url",url);
+                ans.put(kvf.getKey(),obj);
             }
-
+        }
 
         return ok(ans.toString()).as(APP_JSON);
     }
@@ -604,6 +620,10 @@ public class Rest extends Controller {
         return ok(ret.getToken());
 
     }
+
+
+
+
 
     public Result search() throws ServiceException {
         Map<String, String[]> query = request().queryString();

### 2016-03-31 project/httpserver/public/this/js/group.handler.js

@@ -11,7 +11,7 @@ HyperWebSocket.setParticipantPresenceHandler(function (info) {
     var online = info.online;
     console.log("*",info);
     if (!participants[userId]) {
-        participants[userId] = true;
+        participants[userId] = {id:userId,name:name,photo:photo};
         var del_button = "<i style='color:red' onclick='removeMemberConfirmation(\""+ userId+ "\",\""+name+"\")' class='fa fa-times'></i>";
         var onclick = 'onclick=\'selectUser("'+userId+'")\'';
         var style='style="background-image:url('+photo+');height:40px;width:40px" class="pull-left media-object circular"';
@@ -109,14 +109,6 @@ HyperWebSocket.setOnLookAtTimeRequestHandler(function(time) {
     console.log("setTime",time);
     timeline.setTime(time);
     HyperWebSocket.getContent();
-});
-
-HyperWebSocket.setOnUserRemovedHandler(function (uid) {
-    $("#vidMember"+uid).remove();
-    participants[uid]=false;
-    if(uid=="@userId"){
-        document.location = "/";
-    }
 });
 
 HyperWebSocket.setOnLocalVideoAvailableHandler(function(streamUrl){

### 2016-03-31 project/httpserver/public/this/js/hyper.http.js

@@ -134,7 +134,15 @@ var HyperHttp = (function(){
             success: success
         });
 	}
-	
+
+	this.relationState = function(userId, success){
+        $.ajax({
+            url: "/api/relation/"+userId,
+            type: 'GET',
+            success: success
+        });
+    }
+
 	this.addRelation = function(userId, success){
 		$.ajax({
             url: "/api/relation/"+userId,

### 2016-03-31 project/httpserver/public/this/js/search.generic.js

@@ -30,7 +30,7 @@ var SearchGeneric = new (function() {
 	this.updater = function(item){
 		console.log(item);
 		if(item.type=="models.User"){
-			document.location = "/user/"+item.id;
+    		showUserProfile(item.id);
 		}else if(item.type=="models.Group"){
 			document.location = "/group/"+item.id;
 		}
@@ -56,50 +56,3 @@ var SearchPeople = new (function() {
 	return this;
 })();
 
-
-var SearchGroup = new (function() {
-
-	this.source = function(query,process){
-        $.get( "/api/search?query="+query, function( data ) {
-          	console.log(data);
-            process(data.data);
-        });
-	};
-
-	this.displayText = function(item){
-		return item.name;
-	};
-
-	this.highlighter = function(item){
-		if(item.type=="models.User"){
-			return "<i class='fa fa-user'></i> "  + item.name;
-		}else if(item.type=="models.Group"){
-			return "<i class='fa fa-group'></i> " + item.name;
-		}else {
-			return JSON.stringify(item);
-		}
-	};
-
-	this.matcher = function(item){
-		return true;
-	};
-
-	this.updater = function(item){
-		console.log(item);
-		if(item.type=="models.User"){
-			document.location = "/user/"+item.id;
-		}else if(item.type=="models.Group"){
-			document.location = "/group/"+item.id;
-		}
-		return item;
-	}
-
-	return this;
-})();
-
-
-
-
-
-	
-		

### 2016-03-31 project/stats/parser.py

@@ -17,8 +17,8 @@ for line in sys.stdin:
 			tipo = tokens[1]
 			if tipo == "NETW" and len(tokens) ==5:
 				inter = tokens[2]
-				recv = float(tokens[3])/1000
-				sent = float(tokens[4])/1000
+				recv = float(tokens[3])
+				sent = float(tokens[4])
 		
 				if inter not in data_netw:		
 					data_netw[inter] = {}
@@ -30,8 +30,8 @@ for line in sys.stdin:
 			elif tipo == "PROC" and len(tokens) ==7:
 				pid = int(tokens[2])
 				name = tokens[3]
-				res = float(tokens[4])/1000000
-				virt = float(tokens[5])/1000000
+				res = float(tokens[4])
+				virt = float(tokens[5])
 				cpu = float(tokens[6])
 			
 			
@@ -52,7 +52,7 @@ for field in ['cpu','virt','res']:
 		name = proc["name"]
 		
 		if len(proc['samples'])>10:
-			fo = open(field+"_"+str(p)+"_"+name+".csv", "wb")
+			fo = open(field+"_"+str(p)+"_"+name+".dat", "wb")
 			for sample in proc['samples']:
 				x = sample['time']
 				y = sample[field]
@@ -65,7 +65,7 @@ for field in ['sent','recv']:
 		proc = data_netw[p]
 
 		if len(proc['samples'])>10:
-			fo = open(field+"_"+p+".csv", "wb")
+			fo = open(field+"_"+p+".dat", "wb")
 			for sample in proc['samples']:
 				x = sample['time']
 				y = sample[field]

### 2016-04-02 tese2/11_related_work_collaboration.tex
[ADD] amazing explanation of network test results
[ADD] spell checking, explained private network component at architecture

@@
 Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
-        In this scenario, a more efficient alternative to peer-to-peer is the use of a Multipoint Control Unit (MCU).
+        In this scenario, a more efficient alternative to peer-to-peer is the use of a \ac{MCU}.
 
@@
 	{\color{blue}
-	\ac{KMS} supports transcoding, group communications, recording, mixing, broadcasting, aplying filters, image and sound analysis, for example it allows face and qrcode detection.
-\ac{KMS} functionalites are exposed through \emph{JSON-RPC} over \emph{WebSocket}, there are three clients available: JavaScript Client for web browsers, Java Client for Java EE Servers and JavaScript Cleint for Node.js servers.
-	\ac{KMS} suports streaming over \emph{WebRTC}, \ac{HTTP} and \ac{RTP} endpoints.
-Another important component of \ac{KMS} is \emph{Kurento Repository} which suports recording and playing directly from \emph{MongoDB}, that is important for providing a scalable media storage.
+	\ac{KMS} supports trans-coding, group communications, recording, mixing, broadcasting, applying filters, image and sound analysis, for example it allows face and qrcode detection.
+\ac{KMS} functionalities are exposed through \emph{JSON-RPC} over \emph{WebSocket}, there are three clients available: \emph{JavaScript} client for web browsers, \emph{Java} client for \emph{Java EE} servers and \emph{JavaScript} client for \emph{Node.js} servers.
+	\ac{KMS} supports streaming over \emph{WebRTC}, \ac{HTTP} and \ac{RTP} endpoints.
+Another important component of \ac{KMS} is \emph{Kurento Repository} which supports recording and playing directly from \emph{MongoDB}, that is important for providing a scalable media storage.
 Unlinke \emph{Jitsi Video Bridge}, \ac{KMS} does not enforce a specific signaling protocol.
@@
 It requires a simple server (also known as hub) that echoes messages between clients, all the synchronization work is performed by the clients.
-Altough the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}, if we would not choose \emph{NodeJS} for server implementation we would need read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}
+Although the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}, if we would not choose \emph{NodeJS} for server implementation we would need read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}
 
@@
 	{\color{blue}\emph{Etherpad Lite} is a collaborative text editor implemented with \emph{NodeJS} that allows not only the text operations but also to associate authorships to pieces of text.
-It supports adding functionality trought plugins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example:
+It supports adding functionality trough plug-ins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example:
 		\begin{itemize}
@@
 		\item video and audio chat using WebRTC.
-		\item create slideshows.
+		\item create sideshows.
 		\item spell checking.
@@
 \centering
-
+\caption{Comparision between Operational Transformation libraries}
 \label{my-label}
@@
 \end{tabular}
-\caption{Comparision between Operational Transformation libraries}
 \end{table}

### 2016-04-02 tese2/13_implementation_model.tex
[ADD] amazing explanation of network test results
[ADD] performance tests, updated playFW, websocket simplification
[ADD] spell checking, explained private network component at architecture

@@
 \section{Model}
-The model is the most important component of our solution, a badly designed model can imply serious deficulties when implementing new features that are not part of the plans, sometimes we had to redesign the model in order to support new features.
+The model is the most important component of our solution, a badly designed model can imply serious difficulties when implementing new features that are not part of the plans, sometimes we had to redesign the model in order to support new features.
 
@@
 
-Our first decision for our model in order to avoid repeated code, was the isolation of the object's attributes from themselves, so we could apply the search operation to a set of attributes indepently from the object type.
+Our first decision for our model in order to avoid repeated code, was the isolation of the object's attributes from themselves, so we could apply the search operation to a set of attributes independently from the object type.
 To this generic set of properties we call data and each object of this type has a reference to the owner, which is a unique identification number.
 
-\begin{table}[H]
+\begin{table}[!htb]
 \centering
+\caption{Generic data model}
 \label{my-label}
@@
     \end{tabular}
-\caption{Generic data model}
 \end{table}
@@
 Another important attribute specification is the owner identifiability, which tells us if the attribute identifies the object.
-This specification let us create abstract authentication services, for example a user can login into our system by providing any attribute that identifies himself, for example the e-mail but others are possible like the username or cellphone number.
+This specification let us create abstract authentication services, for example a user can login into our system by providing any attribute that identifies himself, for example the e-mail but others are possible like the user name or cellphone number.
 
@@
 The user model is not tied to the user attributes, the information maintained in this model is just used for authentication purposes.
-Passwords are not stored in plaintext, instead we apply hashing and salting techniques \cite{password} in order to make it harder to decode the password by an attacker.
+Passwords are not stored in plain text, instead we apply hashing and salting techniques \cite{password} in order to make it harder to decode the password by an attacker.
 We use \emph{SHA-1} and a random salt per user with 32 characters long.
 
-\begin{table}[H]
+\begin{table}[!htb]
 \centering
+\caption{User model}
 \label{my-label}
@@
     \end{tabular}
-\caption{User model}
 \end{table}
@@
 
-\begin{table}[H]
+\begin{table}[!htb]
 \centering
+\caption{Relation model}
 \label{my-label}
@@
     \end{tabular}
-\caption{Relation model}
 \end{table}
@@
 A group can be public or private.
-If the group is public then it is visible to all users that mantain a friendship with a member of this group.
+If the group is public then it is visible to all users that maintain a friendship with a member of this group.
 If the group is private then it is only visible to their members.
 
-\begin{table}[H]
+\begin{table}[!htb]
 \centering
+\caption{Group model}
 \label{my-label}
@@
     \end{tabular}
-\caption{Group model}
 \end{table}
@@
 
-\begin{table}[H]
+\begin{table}[!htb]
 \centering
+\caption{Message model}
 \label{my-label}
@@
     \end{tabular}
-\caption{Message model}
 \end{table}
@@
 
-\begin{table}[H]
+\begin{table}[!htb]
 \centering
+\caption{Hyper content model}
 \label{my-label}
@@
     \end{tabular}
-    \caption{Hyper content model}
 \end{table}
@@
 
-\begin{table}[H]
+\begin{table}[!htb]
 \centering
+    \caption{Collaborative content model}
     \label{my-label}
@@
     \end{tabular}
-    \caption{Collaborative content model}
 \end{table}
@@
 During a conversation, users may allow sharing their web cameras, by doing so their video is stored in recording chunks.
-Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains a reference to a group and a mapping between entity that was recorded and the correspondent video url.
+Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains a reference to a group and a mapping between entity that was recorded and the correspondent video \ac{URL}.
 
-\begin{table}[H]
+\begin{table}[!htb]
 \centering
+    \caption{Recording model}
     \label{my-label}
@@
     \end{tabular}
-    \caption{Recording model}
 \end{table}

### 2016-04-02 tese2/14_implementation_signaling.tex
[ADD] figuresfiguresfiguresfiguresfiguresfiguresfiguresfigures x-)
[ADD] spell checking, explained private network component at architecture

@@
 
-The signaling protocol could be implemented using \ac{HTTP} messages but they would transport extra information such as \ac{HTTP} headers and would follow a request response signaling mechnism which would not be the best option, as multiple ice candidates can arrive at any time.
+The signaling protocol plays a fundamental role on our solution, after the web server validates the user access,
+it allows the users to directly connect to \ac{KMS} which is placed in a private network and let the server and users negotiate media types and encoding information to use during the conversation.
+
+With respect to the signaling protocol, it could be implemented using \ac{HTTP} messages but they would transport extra information such as \ac{HTTP} headers and would follow a request-response signaling mechanism which would not be the best option, as multiple ice candidates can arrive at any time.
 Using a \ac{REST} \ac{API} over \ac{HTTP} for signaling and allowing other methods to be called at the same time would lead to open multiple \ac{TCP} connections at the same time.
@@
 Instead if we create a \emph{WebSocket} \ac{API} we only need one \ac{TCP} connection and at the same provide bi-directional communications without additional headers.
+Moreover, using \emph{WebSocket} allows the server to send messages to the client without the client having to request it, which can be slightly faster than using \ac{HTTP}.
 
@@
 
-When a user enters in a group conference, after the page is completely loaded it is created a WebSocket that mantains a connection with our web servers. 
+When a user enters in a group conference, after the page is completely loaded it is created a WebSocket that maintains a connection with our web servers. 
 Before creating the web socket, we must identify the user and check if he has permissions to participate in the conference.
 The user identification is done by retrieving the session id from the cookie provided by the user through the \ac{HTTP} headers then we retrieve all the information need from database in order to check if the user has permissions to join the conference room.
-It's important to save de user identification before the socket is created because after the the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455} the \ac{HTTP} context is lost.
+It's important to save the user identification before the socket is created because after the the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455} the \ac{HTTP} context is lost.
 
-When the connection is established between the server and the client, a \emph{PeerConnection} is created on the client and imediately after an \ac{WebRTC} endpoint on the server specifying a possible set of \ac{ICE} servers to connect.
+When the connection is established between the server and the client, a \emph{PeerConnection} is created on the client and immediately after an \ac{WebRTC} endpoint on the server specifying a possible set of \ac{ICE} servers to connect.
 
-The user is asked if he whants to share its camera and microphone, share screen or just receive streams from the server.
-If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2015).} may ask to install a pluggin if the browser doesn't support screen sharing.
+The user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server.
+If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2015).} may ask to install a plug-in if the browser doesn't support screen sharing.
 
-If the user decides to share either from camera of screen, \emph{getUserMedia} is called with the correspondent contraints in order to obtain a local stream.
+If the user decides to share either from camera of screen, \emph{getUserMedia} is called with the correspondent constraints in order to obtain a local stream.
 Use the constraints presented on listing \ref{lst:constraints}.
@@
 
-A popup is raised in order to ask the user to give permission to share those resources.
+A pop-up is raised in order to ask the user to give permission to share those resources.
 If the resources are shared successfully the user creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through WebSockets.
-If the resources are not shared or the user specified to receive stream only, an offer is created specifying the contraints for receiving only video and audio.
+If the resources are not shared or the user specified to receive stream only, an offer is created specifying the constraints for receiving only video and audio.
 
@@
 The \emph{local session description} contains the session identifier, codecs, containers, transport protocols and ports used per media type.
-The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} doesn't need to mix, record and analyse streams coming from the user \ac{WebRTC} endpoint.
+The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} doesn't need to mix, record and analyze streams coming from the user \ac{WebRTC} endpoint.
 
 The server receive and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint.
-Then a \emph{local session description} like listing \ref{lst:sig02} is created on the server and sent back to the client, after that the server trys to gather \ac{ICE} candidates.
+Then a \emph{local session description} like listing \ref{lst:sig02} is created on the server and sent back to the client, after that the server tries to gather \ac{ICE} candidates.
 
@@
 
-After a while both the server and client receive the \ac{ICE} candidates that allow the client to connect directly to \ac{KMS} and vice-versa.
+Subsequently, after a while both the server and client receive the \ac{ICE} candidates that allow the client to connect directly to \ac{KMS} and vice-versa.
 The candidates are received at the client which sets them to its \emph{PeerConnection}.
@@
 
-Both intervenients test the conectivity of each \ac{ICE} candidate.
+In addition both intervenients test the connectivity of each \ac{ICE} candidate.
 When a connection is established the user and server start to interchange stream data but other \ac{ICE} candidates may arrive with better connections, when that happens the connection changes seamlessly.
 
-With the media session established, the server starts to record any received stream and the client creates an \ac{URL} correspondent to the stream location.
+Having the media session established, the server starts to record any received stream and the client creates an \ac{URL} correspondent to the stream location.
 
@@
     \begin{subfigure}{}
-    	\includegraphics[width=0.8\textwidth]{figures/signaling}
+    	\includegraphics[width=0.9\textwidth]{figures/signaling}
     \end{subfigure}

### 2016-04-02 tese2/15_implementation_recording.tex
[ADD] change architecture
[ADD] spell checking, explained private network component at architecture

@@
 
-	After recording, each block was uploaded to our server through \ac{HTTP} and saved into the filesystem, the block metadata was created and inserted on our database.
-Each block metadata contained the file's location for the recorded block, the starting date, duration, user identification and group identification.
-When the file was completely saved the metadata was advertised to the remaining users.
-This metadata was simply used to refresh the user interface, the metadata was completely discard after that because an huge amount of small blocks metadata would use more and more memory as time went by.
+	After recording, each block was uploaded to our server through \ac{HTTP} and saved into the file system, the block meta-data was created and inserted on our database.
+Each block meta-data contained the file's location for the recorded block, the starting date, duration, user identification and group identification.
+When the file was completely saved the meta-data was advertised to the remaining users.
+This meta-data was simply used to refresh the user interface, the meta-data was completely discard after that because an huge amount of small blocks meta-data would use more and more memory as time went by.
 
@@
 
-	Altough this idea was fairly simple, we couldn't achieve seamless sequential block switching.
+	Although this idea was fairly simple, we couldn't achieve seamless sequential block switching.
 Downloading the video file always toke a noticeable amount of time.
@@
 
-	Block switching became more acceptable, but switching the url always produced a flash.
-We solved this problem by having two layers and set the url to the back layer, the front layer freezed with the last frame, then we changed the back layer to front.
+	Block switching became more acceptable, but switching the \ac{URL} always produced a flash.
+We solved this problem by having two layers and set the \ac{URL} to the back layer, the front layer frozen with the last frame, then we changed the back layer to front.
 
-	After we implemented our recording solution using this approach, we tested localy and remotely.
+	After we implemented our recording solution using this approach, we tested locally and remotely.
 For remote connections we observed a fairly high bandwidth usage mainly because blocks were both sent and received at maximum quality.
 
-\subsection{Server side recording to filesystem}
+\subsection{Server side recording to file system}
 
-	Before recording any type of stream we had to analyse the user media offer in order to check if a video as realy being received by \ac{KMS}, otherwise if we would not verify the user's offer, the recording video would be black.
+	Before recording any type of stream we had to analyze the user media offer in order to check if a video as really being received by \ac{KMS}, otherwise if we would not verify the user's offer, the recording video would be black.
 
-	The streaming content received on \ac{KMS} was already compressed due to \ac{WebRTC}'s exchanged \ac{QoS} metrics data.
+	The streaming content received on \ac{KMS} was already compressed due to \ac{WebRTC}'s exchanged quality of service metrics data.
 As a direct consequence our recorder solution used on most cases less disk space per block, but would never use more storage than client side recording.
@@
 
-	With server side recording, the user would maintain always the same stream url even if it is playing realtime video or reproducing recorded video.
-When a user desires to play recorded video, a websocket message is sent specifying the time and the intended user id, the group identification is not sent because it is already associated to the websocket.
-The server performs the same calculations in order to find a block that intersects the requested time, plays it and when finished the next part is automatically played wihtout the user intervention.
+	With server side recording, the user would maintain always the same stream \ac{URL} even if it is playing real-time video or reproducing recorded video.
+When a user desires to play recorded video, a \emph{webSocket} message is sent specifying the time and the intended user id, the group identification is not sent because it is already associated to the \emph{webSocket}.
+The server performs the same calculations in order to find a block that intersects the requested time, plays it and when finished the next part is automatically played without the user intervention.
 
@@
 
-	Back when we were implementing our solution, \ac{KMS} had not support for seeking videos, which meant that blocks would would always start playing from their beggining.
-This lead to a theoric playing time error that can be at most half the duration of a block.
+	Back when we were implementing our solution, \ac{KMS} had not support for seeking videos, which meant that blocks would would always start playing from their beginning.
+This lead to a theoretic playing time error that can be at most half the duration of a block.
 In practice the maximum playing time error coincides with the block duration because we perform the intersection between the requested date and the block but we could decrease the error by half if we calculate the intersection with the requested time plus half the duration.
@@
 	\end{tikzpicture}
-		\caption{Theoretic maximum playing error}
+		\caption{Theoretic maximum playing error $(e)$}
 	\end{figure}
@@
 	\end{tikzpicture}
-		\caption{Practical maximum playing error}
+		\caption{Practical maximum playing error $(e)$}
 	\end{figure}
@@
 	For playing video with an higher velocity we used \emph{ffmpeg}\footnote{\url{https://www.ffmpeg.org/}(accessed: 17 March 2016)} to convert the block into a new video with the desired velocity and seek time.
-Because the media duration is known, when the video started to being convert the headers located at the begining of the file were already written and that made it possible to stream while converting.
+Because the media duration is known, when the video started to being convert the headers located at the beginning of the file were already written and that made it possible to stream while converting.
 
@@
 
-	Although we implemented a solution that worked, we imediately noticed that \emph{ffmpeg} would take some time to initialize and that lead to pauses between switching parts.
+	Although we implemented a solution that worked, we immediately noticed that \emph{ffmpeg} would take some time to initialize and that lead to pauses between switching parts.
 
@@
 	One of our concerns during the development of our solution was the storage scalability.
-Saving files directly into filesystem would require an extra effort to distribute and replicate files among servers.
+Saving files directly into file system would require an extra effort to distribute and replicate files among servers.
 For that purpose \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB}.
@@
 	One of the features that \emph{Kurento Repository} provides is the ability to play directly from the database without having to download the entire file to \ac{KMS}.
-The same is true for recording, but because the file headers are in the beggining and the file is written until is stops, the headers don't contain the necessary information for seeking the file.
+The same is true for recording, but because the file headers are in the beginning and the file is written until is stops, the headers don't contain the necessary information for seeking the file.
 

### 2016-04-02 tese2/16_implementation_content.tex
[RFT] ui improvements
[ADD] spell checking, explained private network component at architecture

@@
 
-	In order to create content the user has the option to write simple movie captions without wrtting any code, otherwise, as mentioned before, it can write \ac{HTML}, \ac{CSS} and \emph{JavaScript}.
+	In order to create content the user has the option to write simple movie captions without writing any code, otherwise, as mentioned before, it can write \ac{HTML}, \ac{CSS} and \emph{JavaScript}.
 The definition of the content's starting and ending time by the user it's not an easy task.
-Defining the content's time to appear in realtime would require a previous user plan.
+Defining the content's time to appear in real-time would require a previous user plan.
 Otherwise the user could make a speech and add the content for later reproduction.
@@
 \end{minipage}
-	In order to help the content creator to show and synchronize its content in realtime we allow the user to encode its content into \emph{QR codes} and show it to the camera in real time.
+	In order to help the content creator to show and synchronize its content in real-time we allow the user to encode its content into \emph{QR codes} and show it to the camera in real time.
 
-	\ac{KMS} lets registring event handlers for \emph{QR code} detection.
+	\ac{KMS} lets registering event handlers for \emph{QR code} detection.
 The component that detects \emph{QR codes} on \ac{KMS} is called periodically and fires the handlers with the decoded content.
 This mechanism does not detect if the \emph{QR code} enters or leaves the screen.
-We had to implement our own mechnism for detecting those events.
+We had to implement our own mechanism for detecting those events.
 
 	Each user session in the server maintains a map with the contents that are present on the screen.
-In order to aplly our algorithm more efficiently we calculate the content's hash through the \emph{md5} method.
+In order to apply our algorithm more efficiently we calculate the content's hash through the \emph{md5} method.
 
-	If the hash is not present in the map it means the \emph{QR code} was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in realtime.
+	If the hash is not present in the map it means the \emph{QR code} was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real-time.
 By doing an analogy to the \emph{Nyquist} theorem we know that if the same \emph{QR code} is not detected after a time bigger then two periods we can conclude that the \emph{QR code} leaved the screen and we add the correspondent content into our database.
@@
 
-	Multiple contents can be displayed at the same time, in order to achieve that, we define layers below the video with the same size. Each layer is associated to the content.
+	Multiple contents can be displayed at the same time, in order to achieve that, we define layers above the video with the same size. Each layer is associated to the content.
 
-	We have taken into account that the amount of content tends to grow with time and the user should only have access to a subset of content instead all of them, which would be very ineficcient.
+	We have taken into account that the amount of content tends to grow with time and the user should only have access to a subset of content instead all of them, which would be very inefficient.
 
-	The content to display for each user depends on the user's position on the timeline, this position is given by an offset between the current time and the navigated time.
-If the user is watchiing the content in realtime this offset is constant so there is no need to synchronize this value.
+	The content to display for each user depends on the user's position on the time-line, this position is given by an offset between the current time and the navigated time.
+If the user is watching the content in real-time this offset is constant so there is no need to synchronize this value.
 
 	Each content is divided into two components, the \emph{start} and the \emph{end} which are sent to users.
-The \emph{start} component contains a timestamp, the content identification number and the content itself in form of text.
-The \emph{end} component only contains the timestamp and the content identifcation number.
+The \emph{start} component contains a time stamp, the content identification number and the content itself in form of text.
+The \emph{end} component only contains the time stamp and the content identification number.
 
@@
 
-	The content description messages that are sent to each user contains the contant itself on the form of a list of events and an atrtibute that specifies if the server has more content to return.
+	The content description messages that are sent to each user contains the constant itself on the form of a list of events and an attribute that specifies if the server has more content to return.
 
@@
 
-	When a user enters the conference room, imediatelly after the \emph{WebSocket} creation the server sends him the current content.
+	When a user enters the conference room, immediately after the \emph{WebSocket} creation the server sends him the current content.
 
@@
 All the events before the user's time are rendered and removed from the set of events.
-A timer is schedulled for the first component on the event set and the process repeats while the set is not empty.
+A timer is scheduled for the first component on the event set and the process repeats while the set is not empty.
 
-	If the set is empty, there two options, if the server contained more content a new request for content is made and the proccess starts from the beggining, namely the server sends the correpondent content again.
+	If the set is empty, there two options, if the server contained more content a new request for content is made and the process starts from the beginning, namely the server sends the correspondent content again.
 If the server has no more content the process is stopped until it sends more.
@@
 
-	Our solution is flexible on what kind of interactions are possible to the users in real time but allowing users to write \emph{Javascript} that is executed on the other users browser would attackers to misuse their resources and access to critical information.
+	Our solution is flexible on what kind of interactions are possible to the users in real time but allowing users to write \emph{JavaScript} that is executed on the other users browser would attackers to misuse their resources and access to critical information.
 
@@
 
-	By not allowing javascript we would need to implement all actions a priori and fire them when a type of message is received.
-We decided to ignore the security vulnarabilities that are exposed by evaluating \emph{javascript} because in order to offer the same interactions we would have to do an exaustive functionaly requirements gathering.
+	By not allowing \emph{JavaScript} we would need to implement all actions a priori and fire them when a type of message is received.
+We decided to ignore the security vulnerabilities that are exposed by evaluating \emph{JavaScript} because in order to offer the same interactions we would have to do an exhaustive functional requirements gathering.
 
-	Another way to solve this problem, which is not within the goals of this thesis, is to analyse the javascript code and detect if it is malicious.
+	Another way to solve this problem, which is not within the goals of this thesis, is to analyze the \emph{JavaScript} code and detect if it is malicious.
 
@@
 
-	We have used \emph{vis.js} to display out timeline.
+	We have used \emph{vis.js} to display out time-line.
 This library was created for content navigation through time, but it was not designed to be always moving automatically, we have created a background timer that performs the animation of moving the window of time bounds and user navigated time marker.
 
-	Figure \ref{fig:timeline} shows the the graphic appearance of our interactive timeline, in order to navigate through time the user must drag and drop the timeline horizontaly, when the user drops the timeline an event handler is called with the new user's time offset which will be used to send a message to the server in order to choose the correspondent content to display.
+	Figure \ref{fig:timeline} shows the the graphic appearance of our interactive time-line, in order to navigate through time the user must drag and drop the time-line horizontally, when the user drops the time-line an event handler is called with the new user's time offset which will be used to send a message to the server in order to choose the correspondent content to display.
 
@@
 		\includegraphics[width=\textwidth]{figures/timeline.png}
-		\caption{Interactive timeline}
+		\caption{Interactive time-line}
 		\label{fig:timeline}
@@
 	We have noticed the server time being different from client time and this created graphic inconsistencies such as showing existing recorded blocks of movie in the future.
-To solve this problem the server sends its time to client imediately after the websocket creation, we synchronize the timeline with the server and although it may exist a small error due to network transmission time the graphical error is negligible.
+To solve this problem the server sends its time to client immediately after the \emph{webSocket} creation, we synchronize the time-line with the server and although it may exist a small error due to network transmission time the graphical error is negligible.
 
@@
 	Time annotations are simpler way save points in time and share them with other users.
-When a user enters in a conference room the server sends all annotations so they can be displayed directly into the user timeline.
+When a user enters in a conference room the server sends all annotations so they can be displayed directly into the user time-line.
 Each time annotation is created all users are advertised so they can update their interfaces.
 
-	Figure \ref{fig:annotation} shows the creation and placement of annotations on the timeline, to save the annotation the user must click on the floppy icon in order to save it and advertise the conference participants.
+	Figure \ref{fig:annotation} shows the creation and placement of annotations on the time-line, to save the annotation the user must click on the floppy icon in order to save it and advertise the conference participants.
 
@@
 
-	Besides the ability to create tags it is also possible to create time hyperlinks, see figure \ref{fig:timelink}, that can be sent to other users externaly so they can navigate directly to the content when entering in the conference room.
+	Besides the ability to create tags it is also possible to create time hyper-links, see figure \ref{fig:timelink}, that can be sent to other users externally so they can navigate directly to the content when entering in the conference room.
 
@@
 In the case of hyper content after handling the result from database we ignore \ac{HTML} tags, extract the text with \emph{Jsoup}\footnote{\url{http://jsoup.org/} (Accessed 21 March 2016)} and apply the query again.
-We extract the text from \ac{HTML} because accidentatly searching for text contained in \ac{HTML} tags would lead to incorrect results.
+We extract the text from \ac{HTML} because accidentally searching for text contained in \ac{HTML} tags would lead to incorrect results.
 

### 2016-04-02 tese2/17_implementation_deploy.tex
[ADD] spell checking, explained private network component at architecture

@@
 \subsection{Software}
-	INESC-ID provided us a restricted linux account without administration permissions, which we prevented us the installation of our solution directly on the machine because we would need administrative priviledges to install all the software that our solution depends on.
-Although that limitation, they provided access to Docker\footnote{\url{https://www.docker.com/} (Accessed March 27, 2016)} which run on host operating system as an isolated process in userspace.
+	\emph{INESC-ID} provided us a restricted \emph{Linux} account without administration permissions, which we prevented us the installation of our solution directly on the machine because we would need administrative privileges to install all the software that our solution depends on.
+Although that limitation, they provided access to Docker\footnote{\url{https://www.docker.com/} (Accessed March 27, 2016)} which run on host operating system as an isolated process in user space.
 
-	Inside a Docker container we have the administrative priviledges to install all the software dependencies.
+	Inside a Docker container we have the administrative privileges to install all the software dependencies.
 
-	We could use one docker image for each component but in order to reduce the network usage we prefered to install all the components within the same image.
+	We could use one docker image for each component but in order to reduce the network usage we preferred to install all the components within the same image.
 At the same we provide an all-in-one easy to deploy solution as all the \ac{IP}s are local.

### 2016-04-02 tese2/architecture.tex
[ADD] spell checking, explained private network component at architecture

@@
 
-Taking into account the goals of this project and all the technology presented so far. Our proposal is the development of a web application that provides communication and collaboration in real-time.
+Taking into account the goals of this project and all the technology presented so far.
+Our proposal is the development of a web application that provides communication and collaboration features in real-time.
 
@@
 
-	Clients can discover chat rooms and other clients by querying the signaling server, they can create rooms for multi-party communication which is achieved by using \ac{WebRTC}'s \emph{MediaStream} and \emph{DataChannel}, namely for audio, video and text communication.
+	Moreover we allow clients to discover chat rooms and other clients by navigating on the web pages provided by our web server.
+In addition users can create rooms for multi-party audio and video communication communication which is achieved by using \ac{WebRTC}'s \emph{MediaStream} and \emph{PeerConnection}.
 
-	Chat rooms can be public or private.
-Public chat rooms are moderated by a group of clients which initially is formed by the room creator, this type of room has no access restrictions by default, but that can be changed by its moderators.
-Private chat rooms will be visible for a list of clients or can be accessed by clients that have a link for that chat room.
+	Furthermore, chat rooms can be public or private.
+Public chat rooms are moderated by a group of clients which initially is only the room creator, this type of room has no access restrictions by default, but that can be changed by its moderators.
+Private chat rooms will be visible for a defined list of clients or can be accessed by clients that have a link for that conference room.
 
@@
 
-\begin{figure}[H]
+	In this section we present the modules of our application.
+
+\begin{figure}[!htb]
 	\centering
@@
 \section{Implementation Proposal}
-The infrastructure is composed by: Ice Server, Web Server, Stream Server, Signaling Server and Test Clients.
+The infrastructure is composed by: web server, stream server, signaling server, database and video repository.
 
@@
 
-In order to simplify our solution, we propose both the \emph{Application} and \emph{Signaling} modules implementation within the same server application so it could be easy to control which resources the client has permission to access.
+In order to simplify our solution, we propose both the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules implementation within the same server application so it could be easy to deploy as a single image.
+To this set of modules we often call \emph{backend}.
 
-We have taken into account the compatibility between the streaming server, database and the operation transformation solutions in order choose the appropriate framework to implement our web server.
+Having established that the \emph{backend} modules are placed in the same machine, that helps controlling which resources the client has permission to access as those modules are seen as a private network.
 
-Based on our study about signaling protocols we decided to exclude \ac{XMPP} due to the development difficulties that it exposes when using multiple web pages, namely the user re-authentication performed each time a page is loaded and the single tab usage limitation.
-By consequence we have not choosen \emph{Jitsi Video Brigde} as it uses \ac{XMPP}.
+To qualify the above we provide public access to \ac{HTTP} server ports, maintaining the access to other components restricted through firewall rules.
+In relation to the database there is no directly access from outside, all the database information is accessed via our application server which validates the permissions of users when acting above our system.
 
-By excluding \emph{Jitsi Video Brigde} we have decided that our solution must use \ac{KMS}.
-Our web server could be implemented more easily with \emph{NodeJS} or \emph{Java} due to the fact \ac{KMS} provides clients for both technologies but others could be used as \ac{KMS} also exposes their \ac{API} via \emph{WebSockets}.
+On the other hand, the access to our streaming servers is also restricted but clients can connect to them after concluding the signaling phase.
+This signaling phase may or not proceed in function of the client access permissions, for example if a user is trying to access a private conference room that is not member of or has no invitation link, the signaling server rejects to start the signaling phase and the user cannot access to the streaming server.
+
+Furthermore, we have taken into account the compatibility between the streaming server, database and the operation transformation solutions in order choose the appropriate framework to implement our web server.
+
+As a result of our study about signaling protocols we have decided to exclude \ac{XMPP} due to the development difficulties that it exposes when using multiple web pages, namely the user re-authentication performed each time a page is loaded and the single tab usage limitation.
+By consequence we have not chosen \emph{Jitsi Video Brigde} as it uses \ac{XMPP}.
+
+Accordingly, by excluding \emph{Jitsi Video Brigde} we have decided that our solution must use \ac{KMS}.
+Our web server could be implemented easily with \emph{NodeJS} or \emph{Java} due to the fact \ac{KMS} provides clients for both technologies but others could be used as \ac{KMS} also exposes their \ac{API} via \emph{WebSockets}.
 
@@
 
-For the web server we could choose \emph{NodeJS} or any \emph{Java} based web framework.
+Due to the fact we are going to use \ac{KMS} as streaming server solution, we could choose \emph{NodeJS} or any \emph{Java} based web framework for implementing our web application server.
 One of our criteria for choosing the web framework is the ability to follow the \ac{MVC} paradigm which can help us to organize our code.
-Without any strong reason we have decided to implement our webserver with \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java}.
+Without any strong reason we have decided to implement our web server with \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java}.
 
-For the collaborative text editor, we have choosen \emph{OT.js} due to its indenpendence from server implementation and storage choice.
+Importantly, for the collaborative text editor, we have chosen \emph{OT.js} due to its server and storage implementation choice independence.
 
-For the \emph{NAT Traversal} module an \ac{ICE} server is not required to be on our infrastructure, as a public \ac{STUN} server can be used for testing our solution.
+For the \emph{NAT Traversal} module an \ac{ICE} server is not required to be on our infrastructure, as a public \ac{STUN} server can be used for testing our solution although we recognize that for a production environment we would need to maintain our own \ac{TURN} servers in order to ensure connectivity to all clients.
 
-On the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} should be installed as web browsers.
-Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Adapter.js}, \emph{OT.js} can be downloaded from the Web Server and executed on the client side.
+Not less important, on the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} should be installed as web browsers.
+Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Adapter.js}, \emph{OT.js} can be downloaded from the web server and executed on the client side.
 
@@
 
-Figure~\ref{fig:apparch} presents the application architecture and the underlying technologies.
+Figure~\ref{fig:apparch} presents the application architecture and the underlying technologies seen from the user's perspective.
 \emph{Adapter.js} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.

### 2016-04-02 tese2/evaluation.tex
[ADD] spell checking, explained private network component at architecture

@@
 
-  In this chapter we describe how we tested our solution, show and analyse the results in order to validate the contributions of this thesis.
+  In this chapter we describe how we tested our solution, show and analyze the results in order to validate the contributions of this thesis.
 
@@
 
-\section{Tests Scenarios}
+  \section {Performance Tests}
 
-  In this section we describe both performance and usability test scenarios that we have applied.
+     In this section we describe performance test scenarios that we have applied and their respective results.
 
-  \subsection {Performance Tests}
-      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects with a periodicity of one second; CPU, physical memory information realtive to each running process and network usage relative to each interface.
+    \subsection{Tests Scenarios}
+
+      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects with a periodicity of one second; CPU, physical memory information relative to each running process and network usage relative to each interface.
 
@@
 
-  \subsection {Usability Tests}
-
-      In order to the usability of our solution, we haver performed usability tests with the help of real users with different backgrounds and ages.
-
-      We handled a guide to the users with five tasks to perform, the metrics we used for each task were: number of clicks, number of errors including a description and time spent.
-
-      The tasks we asked to the user are:
-
-      \begin{enumerate}
-      \item Login into the system with the provided credentials and accept the received friendship request. Find the coordinator's brother and add him as a friend.
-
-      \item Create a private conference room, enter and share your screen, add the coordinator to the conference room and chat with him.
-After that use the collaborative editor in order to write at the same time as the coordinator.
-Lastly save the editor and leave the conference room.
-
-      \item Enter in the conference room correspondent to the third task as an observer and navigate to the specified annotation.
-Watch the video until a list of topics is overlayed in the video and choose one of them and leave the conference room.
-
-      \item Enter in the conference room correspondent to the fourth task by sharing your camera and create a time annotaion in the instant of time when you entered, navigate to the current time, search for the annotation and leave the conference room.
-
-      \item Enter in the conference room correspondent to the fourth task by sharing your camera and add a subtitle in the video, preview it, specify an interval of time and save it.
-Then show the provided \emph{QR code} to the camera and leave the leave the conference room.
-      \end{enumerate}
-
-      The goal of the first task consists on evaluating the interface for authentication and friendship management.
-
-      The second task is used to make the users familiar with our tools for collaborative content edition.
-
-      The goal of the third task is to demonstrate the navigation functionalities and synchronized interactive content superposition.
-
-      The fourth task is used to analyse the behaviour and difficulties of the user when creating a time annotation.
-
-      The goal of the fifth task is to create synchronized overlayed content, demonstrate the difficulties of such task and present an easier way to synchronize content.
-
-\section{Test Results}
-  \subsection {Performance Tests}
+\subsection{Test Results}
 
@@
       From the media server prespective if there are $n$ clients connected each of them sending and receiving one stream, it is expected that server sends and receives also $n$ streams.
-By so we expect that the amount of network increases lineraly as users join a confrence room.
+By so we expect that the amount of network increases linearly as users join a conference room.
 
@@
       The green peaks are caused by that being transfered between \ac{KMS} and \emph{MongoDB} through \emph{Kurento Repository} each peak occurs each time a block of video is recorded which in this case is every ten seconds.
-      The recordings are synchronized so all user and mixed blocks starts and ends at the same time, that's why the amount of work done every ten seconds acumulates and because this is poerformed locally the maximum transfer rate is limited by the performance of the memory as buffers are writen to buffers then to disks.
+      The recordings are synchronized so all user and mixed blocks starts and ends at the same time, that's why the amount of work done every ten seconds accumulates and because this is performed locally the maximum transfer rate is limited by the performance of the memory as buffers are written to buffers then to disks.
       Client's sent data transfer rate has no significant peaks as signaling information contains few information.
@@
 Both \ac{JVM}, \emph{MongoDB} and \ac{KMS} performs their own memory management by holding and recycling objects when needed.
-The expected and observed behaviour of the memory usage is growth memory usage while the users are entering in the conference room and a memory usage stabilization aftwerwards.
+The expected and observed behavior of the memory usage is growth memory usage while the users are entering in the conference room and a memory usage stabilization afterwards.
 
@@
 
-\emph{MongoDB} memory usage keeps increasing because it trys to fit part of the database on \ac{RAM} for fast read access.
+\emph{MongoDB} memory usage keeps increasing because it tries to fit part of the database on \ac{RAM} for fast read access.
 \emph{MongoDB} checkpoints data to disk every 60 seconds or when journal data exceeds 2GB\footnote{\url{https://docs.mongodb.org/manual/faq/storage/}(Accessed March 28, 2016)}, that explains the small memory usage peaks during our test case.
@@
 Figure \ref{fig:test_full_features_cpu_zoom} is the same as figure \ref{fig:test_full_features_cpu} but zoomed over \emph{MongoDB}, \emph{Java} and \emph{Kurento Repository}.
-As we can see there are periodic proccessing peaks every ten seconds that are due to block recording.
+As we can see there are periodic processing peaks every ten seconds that are due to block recording.
 
@@
 We conclude that \emph{QR code} detection is a very intensive task, approximately doubling the amount of work performed by \ac{CPU}.
-Without this feature the network and memory usage had insignificate changes compared to \ac{CPU} usage.
+Without this feature the network and memory usage had insignificant changes compared to \ac{CPU} usage.
 
-  \subsection {Usability Tests}
+\section {Usability Tests}
+     In this section we describe usability test scenarios that we have applied and their respective results.
+
+    \subsection{Tests Scenarios}
+
+      In order to the usability of our solution, we haver performed usability tests with the help of real users with different backgrounds and ages.
+
+      We handled a guide to the users with five tasks to perform, the metrics we used for each task were: number of clicks, number of errors including a description and time spent.
+
+      The tasks we asked to the user are:
+
+      \begin{enumerate}
+      \item Login into the system with the provided credentials and accept the received friendship request. Find the coordinator's brother and add him as a friend.
+
+      \item Create a private conference room, enter and share your screen, add the coordinator to the conference room and chat with him.
+After that use the collaborative editor in order to write at the same time as the coordinator.
+Lastly save the editor and leave the conference room.
+
+      \item Enter in the conference room correspondent to the third task as an observer and navigate to the specified annotation.
+Watch the video until a list of topics is overlayed in the video and choose one of them and leave the conference room.
+
+      \item Enter in the conference room correspondent to the fourth task by sharing your camera and create a time annotation in the instant of time when you entered, navigate to the current time, search for the annotation and leave the conference room.
+
+      \item Enter in the conference room correspondent to the fourth task by sharing your camera and add a subtitle in the video, preview it, specify an interval of time and save it.
+Then show the provided \emph{QR code} to the camera and leave the leave the conference room.
+      \end{enumerate}
+
+      The goal of the first task consists on evaluating the interface for authentication and friendship management.
+
+      The second task is used to make the users familiar with our tools for collaborative content edition.
+
+      The goal of the third task is to demonstrate the navigation functionalities and synchronized interactive content superposition.
+
+      The fourth task is used to analyse the behavior and difficulties of the user when creating a time annotation.
+
+      The goal of the fifth task is to create synchronized overlayed content, demonstrate the difficulties of such task and present an easier way to synchronize content.
+
+  \subsection{Test Results}
 
@@
 On the second task one user clicked on the friends list for adding a friend to room and by consequence made him leave the conference room to the user profile.
-We solved this problem by showing a popup with the user profile and an additional button to add the user to the conference room as it can be seen on figure \ref{fig:test_ui_04}.
+We solved this problem by showing a pop-up with the user profile and an additional button to add the user to the conference room as it can be seen on figure \ref{fig:test_ui_04}.
 
@@
 
-On the third task we noticed that both users were not expecting to search the content directly from the search bar, instead they used the timeline.
+On the third task we noticed that both users were not expecting to search the content directly from the search bar, instead they used the time-line.
 
-On the fourth task both users didn't placed their annotations on wright place, we only provide one way to place annotations, which is by dragging them on the timeline. 
-After creating an annotation users also didn't understand that they had to save their changes on the timeline, we solved this problem by saving automatically each time the user changed any annotation, the old user interface can be seen on figure \ref{fig:test_ui_05}.
+On the fourth task both users didn't placed their annotations on wright place, we only provide one way to place annotations, which is by dragging them on the time-line. 
+After creating an annotation users also didn't understand that they had to save their changes on the time-line, we solved this problem by saving automatically each time the user changed any annotation, the old user interface can be seen on figure \ref{fig:test_ui_05}.
 
@@
 \item Support file upload on our chat.
-\item Starting with our timeline more zoomed in order to give more sensation of time.
-\item Create content with starting time and duration insted ending time.
+\item Starting with our time-line more zoomed in order to give more sensation of time.
+\item Create content with starting time and duration instead ending time.
 \end{itemize}

### 2016-04-03 tese2/stats/summary_full_net.tex
[FIX] memory leak ready for new performance tests

@@
 
-	\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/summary_full_sent_eth0.dat};
+	\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/summary_sent_eth0.dat};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[color=Red] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/summary_full_recv_eth0.dat};
+	\addplot[color=Red] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/summary_recv_eth0.dat};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[color=Green] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/summary_full_sent_lo.dat};
+	\addplot[color=Green] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/summary_sent_lo.dat};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-04-03 tese2/stats/test_two_times_cpu.tex
[FIX] memory leak ready for new performance tests

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={CPU usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=-10, xmax=1690,
	    ymin=0, ymax=1200,
	    xtick={0,120,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-96,y index=1] {stats/test_two_times/cpu_7_mongod.dat};
		\addlegendentry{Mongod}

		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-96,y index=1] {stats/test_two_times/cpu_67_kurento-media-server.dat};
		\addlegendentry{Kurento Media Server}

		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-96,y index=1] {stats/test_two_times/cpu_35_kurento-repo.dat};
		\addlegendentry{Kurento Repository}

		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-96,y index=1] {stats/test_two_times/cpu_233_bash.dat};
		\addlegendentry{Java}

	\end{axis}

\end{tikzpicture}

### 2016-04-03 tese2/stats/test_two_times_mem.tex
[FIX] memory leak ready for new performance tests

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Memory Usage},
	    xlabel={Time [seconds]},
	    ylabel={Memory [MB]},
	    xmin=-10, xmax=1690,
	    ymin=0, ymax=1800,
	    xtick={0,120,...,2000},
	    ytick={0,200,...,2000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-96,y expr=\thisrowno{1}/1000000] {stats/test_two_times/res_7_mongod.dat};
		\addlegendentry{Mongod}

		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-96,y expr=\thisrowno{1}/1000000] {stats/test_two_times/res_67_kurento-media-server.dat};
		\addlegendentry{Kurento Media Server}

		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-96,y expr=\thisrowno{1}/1000000] {stats/test_two_times/res_35_kurento-repo.dat};
		\addlegendentry{Kurento Repository}

		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-96,y expr=\thisrowno{1}/1000000] {stats/test_two_times/res_233_bash.dat};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}

### 2016-04-03 project/httpserver/app/main/Recorder.java

@@ -1,5 +1,7 @@
 package main;
 
+import jdk.nashorn.internal.runtime.ECMAException;
+import org.kurento.client.Continuation;
 import org.kurento.client.MediaElement;
 import org.kurento.client.MediaProfileSpecType;
 import org.kurento.client.RecorderEndpoint;
@@ -10,15 +12,14 @@ import org.kurento.repository.service.pojo.RepositoryItemRecorder;
 import java.io.File;
 import java.io.IOException;
 import java.nio.file.Files;
-import java.util.Collections;
-import java.util.Date;
-import java.util.Map;
+import java.util.*;
 
 
 /**
  * The Class Recorder.
  */
 public class Recorder {
+
 
     private static boolean useRepo = true;
 
@@ -33,17 +34,18 @@ public class Recorder {
     private RecorderEndpoint recorder;
     private RepositoryItemRecorder item;
     private File tempFile;
+    private Timer timer = new Timer();
+    private RecorderHandler handler;
 
     public Recorder(MediaElement endPoint, int duration, RecorderHandler handler) {
+        this.handler = handler;
         new Thread(new Runnable() {
             @Override
             public void run() {
                 try {
                     synchronized (endPoint) {
-                        System.out.println("Recording...");
 
                         RecorderEndpoint.Builder builder;
-
                         if (useRepo) {
                             item = KurentoManager.repository.createRepositoryItem(Collections.emptyMap());
                             builder = new RecorderEndpoint.Builder(endPoint.getMediaPipeline(), item.getUrl());
@@ -51,48 +53,20 @@ public class Recorder {
                             tempFile = File.createTempFile("hyper", ".webm");
                             builder = new RecorderEndpoint.Builder(endPoint.getMediaPipeline(), "file://" + tempFile.getAbsolutePath());
                         }
-                        builder.stopOnEndOfStream().stopOnEndOfStream().withMediaProfile(MediaProfileSpecType.WEBM).build();
+
+
+                        builder.stopOnEndOfStream().withMediaProfile(MediaProfileSpecType.WEBM);
                         recorder = builder.build();
-
-
                         endPoint.connect(recorder);
                         recorder.record();
 
-                        try {
-                            Thread.sleep(duration);
-                        } catch (InterruptedException e) {
-                            e.printStackTrace();
-                        }
+                        timer.schedule(new TimerTask() {
+                            @Override
+                            public void run() {
+                                stop();
+                            }
+                        },duration);
 
-
-                        try {
-                            recorder.stop();
[…]

### 2016-04-03 project/httpserver/app/services/CreateIntervalService.java

@@ -21,8 +21,6 @@ public class CreateIntervalService extends Service<RecordingInterval> {
     /** The start. */
     private final Date start;
 
-    /** The inter. */
-    private RecordingInterval inter = null;
 
 
     /**
@@ -36,14 +34,6 @@ public class CreateIntervalService extends Service<RecordingInterval> {
         this.start = start;
     }
 
-    /**
-     * Gets the interval.
-     *
-     * @return the interval
-     */
-    public RecordingInterval getInterval() {
-        return inter;
-    }
 
     /*
      * (non-Javadoc)
@@ -53,7 +43,7 @@ public class CreateIntervalService extends Service<RecordingInterval> {
     @Override
     public RecordingInterval dispatch() throws ServiceException {
         synchronized (LOCK) {
-            inter = new RecordingInterval(groupId, start, start);
+            RecordingInterval inter = new RecordingInterval(groupId, start, start);
             inter.save();
             return inter;
         }

### 2016-04-03 project/ice-server/build.sh

@@ -1,0 +1,1 @@
+docker build -t hrocha/ice-server .

### 2016-04-07 tese2/evaluation.tex
[FIX] memory leak ready for new performance tests
[ADD] client side tests

@@
 
-      We are presenting in this sections the results of our performance tests after implementing our solution.
+      We are presenting in this section the results of our performance tests after implementing our solution.
+\subsubsection{Network usage}
 
@@
 
+On the other hand figure \ref{fig:test_client_net} shows the respective network usage on the client side during our test case.
+We observe that int he first seconds the client adjusts the video quality that sends to \ac{KMS}, whenever a new client enters in the conference we observer that \ac{KMS} decreases the video quality in order to instantaneously integrate a new user into the conference room, after a while \ac{KMS} realizes that its network can handle the increase of clients and sends the video with a better quality to every participant.
+When a user leaves the conference room \ac{KMS} has no need to decrease the participant's video quality as less network bandwidth will be used.
+
+  \begin{figure}[!htb]
+  \begin{center}
+    \input{stats/test_client_net.tex}
+  \end{center}
+  \caption{Client Network usage during our test case}
+  \label{fig:test_client_net}
+\end{figure}
+
+\subsubsection{Memory usage}
+
 Figure \ref{fig:test_full_features_mem} shows the memory usage during our performance test case.
@@
 When the conference room is empty no there are no video recordings, which explains the memory stabilization at the end.
+
+\emph{KMS} memory usage was increasing and in fact we did not expected that, we have done more tests and we have observed that by disabling the recorder feature memory usage kept linearly related to the amount of clients present on the room.
+Therefore, we have verified our implementation and we have concluded that nothing was wrong with our code, we have confirmed that every allocated resource was indeed released after the end of recording.
+As a result of our intensive tests and consecutive failures we suspected that the problem was not ours.
+
+Hence, in order to confirm our suspicions, we have decided to read the implementation of \ac{KMS} and we have found that \ac{KMS} was not releasing the memory if the recorder was stopped by us.
+Subsequently, we have changed the source code and submitted to \emph{Kurento} team through a \emph{GIT}'s \emph{"pull request"}.
+
+Figure \ref{fig:test_ram_fixed_mem} shows the memory usage during our performance test case taking into account the patch that we applied to \ac{KMS}.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \input{stats/test_ram_fixed_mem.tex}
+  \end{center}
+  \caption{Memory usage after fixing recorder memory leak}
+  \label{fig:test_ram_fixed_mem}
+\end{figure}
+
+\subsubsection{CPU usage}
 
@@
 Without this feature the network and memory usage had insignificant changes compared to \ac{CPU} usage.
+
+\subsubsection{Two consecutive test cases}
+
+We have also performed two consecutive test cases in order to understand if we the influence of the first test case over the second, as observed in figure \ref{fig:test_two_times_cpu}, there is no \ac{CPU} influence between two test cases.
+
+  \begin{figure}[!htb]
+  \begin{center}
+    \input{stats/test_two_times_cpu.tex}
+  \end{center}
+  \caption{Percentage of CPU used during two consecutive test cases}
+  \label{fig:test_two_times_cpu}
+\end{figure}
+
+On the other hand due to memory recycling techniques we can observe in figure \ref{fig:test_two_times_mem} that some of the memory that was allocated in the first test was reused in the second case.
+
+\begin{figure}[!htb]
+  \begin{center}
+    \input{stats/test_two_times_mem.tex}
+  \end{center}
+  \caption{Memory usage during two consecutive test cases}
+  \label{fig:test_two_times_mem}
+\end{figure}
+
+\subsubsection{One hour test case}
+
+   Figure \ref{fig:test_hour_mem} shows the memory used by our solution with seven clients simultaneously on the same conference room during one hour.
+As we can observe \ac{KMS} memory usage stabilizes.
+
+  \begin{figure}[!htb]
+  \begin{center}
+    \input{stats/test_hour_mem.tex}
+  \end{center}
+  \caption{Memory usage during one hour with seven clients}
+  \label{fig:test_hour_mem}
+\end{figure}
 

### 2016-04-07 tese2/stats/test_client_net.tex
[ADD] client side tests

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Network usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [kbps]},
	    xmin=-10, xmax=790,
	    ymin=0, ymax=1000,
	    xtick={0,60,...,2000},
	    ytick={0,100,...,5000},
	    extra x ticks={0,60,...,2000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

	\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-0,y expr=\thisrowno{1}/1000] {stats/client/send.dat};
	\addlegendentry{Sent(eth0)}

	\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-0,y expr=\thisrowno{1}/1000] {stats/client/recv.dat};
	\addlegendentry{Recv(eth0)}

	\end{axis}
\end{tikzpicture}

### 2016-04-07 tese2/stats/test_hour_cpu.tex
[ADD] client side tests

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={CPU usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=0, xmax=3600,
	    ymin=0, ymax=1000,
	    xtick={0,500,...,5000},
	    ytick={0,100,...,2000},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[smooth,color=Blue,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_8_mongod.dat};
		\addlegendentry{Mongod}

		\addplot[smooth,color=Red,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_91_kurento-media-server.dat};
		\addlegendentry{Kurento Media Server}

		\addplot[smooth,color=Green,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_36_kurento-repo.dat};
		\addlegendentry{Kurento Repository}

		\addplot[smooth,color=Magenta,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_228_bash.dat};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}

### 2016-04-07 tese2/stats/test_hour_mem.tex
[ADD] client side tests

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=6cm,
	    title={Memory Usage},
	    xlabel={Time [seconds]},
	    ylabel={Memory [MB]},
	    xmin=0, xmax=3600,
	    ymin=0, ymax=4000,
	    xtick={0,500,...,5000},
	    ytick={0,500,...,4000},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[smooth,color=Blue,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_8_mongod.dat};
		\addlegendentry{Mongod}

		\addplot[smooth,color=Red,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_91_kurento-media-server.dat};
		\addlegendentry{Kurento Media Server}

		\addplot[smooth,color=Green,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_36_kurento-repo.dat};
		\addlegendentry{Kurento Repository}

		\addplot[smooth,color=Magenta,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_228_bash.dat};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}

### 2016-04-07 tese2/stats/test_hour_net.tex
[ADD] client side tests

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=8cm,
	    title={Network usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [mbps]},
	    xmin=0, xmax=3600,
	    ymin=0, ymax=24,
	    xtick={0,500,...,5000},
	    ytick={0,5,...,100},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

	\addplot[smooth,color=Blue,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/sent_eth0.dat};
	\addlegendentry{Sent(eth0)}

	\addplot[smooth,color=Red,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/recv_eth0.dat};
	\addlegendentry{Recv(eth0)}

	\addplot[smooth,color=Green,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/recv_lo.dat};
	\addlegendentry{Sent/Recv(lo)}

	\end{axis}
\end{tikzpicture}

### 2016-04-07 tese2/stats/test_ram_fixed_mem.tex
[ADD] client side tests

--- 2016-03-29 tese2/stats/test_full_features_mem.tex
@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_7_mongod.csv};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_67_kurento-media-server.csv};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_67_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_35_kurento-repo.csv};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/res_233_bash.csv};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_233_bash.dat};
 		\addlegendentry{Java}

### 2016-04-07 tese2/stats/test_two_times_cpu.tex
[ADD] client side tests

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-96,y index=1] {stats/test_two_times/cpu_7_mongod.dat};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-96,y index=1] {stats/test_two_times/cpu_67_kurento-media-server.dat};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_68_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-96,y index=1] {stats/test_two_times/cpu_35_kurento-repo.dat};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-96,y index=1] {stats/test_two_times/cpu_233_bash.dat};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_234_bash.dat};
 		\addlegendentry{Java}

### 2016-04-07 tese2/stats/test_two_times_mem.tex
[ADD] client side tests

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-96,y expr=\thisrowno{1}/1000000] {stats/test_two_times/res_7_mongod.dat};
+		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-96,y expr=\thisrowno{1}/1000000] {stats/test_two_times/res_67_kurento-media-server.dat};
+		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_68_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-96,y expr=\thisrowno{1}/1000000] {stats/test_two_times/res_35_kurento-repo.dat};
+		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-96,y expr=\thisrowno{1}/1000000] {stats/test_two_times/res_233_bash.dat};
+		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_234_bash.dat};
 		\addlegendentry{Java}

### 2016-04-07 project/httpserver/app/controllers/WSController.java

@@ -5,10 +5,9 @@ import main.Global;
 import main.Room;
 import main.Tools;
 import main.UserSession;
-import models.RecordingInterval;
 import models.Message;
+import models.RecordingInterval;
 import models.TimeAnnotation;
-import models.User;
 import org.bson.Document;
 import org.bson.types.ObjectId;
 import org.json.JSONArray;
@@ -23,7 +22,9 @@ import services.*;
 
 import java.io.IOException;
 import java.text.ParseException;
-import java.util.*;
+import java.util.Date;
+import java.util.List;
+import java.util.UUID;
 
 public class WSController extends Controller {
 
@@ -39,10 +40,9 @@ public class WSController extends Controller {
             public void onReady(WebSocket.In<String> in, WebSocket.Out<String> out) {
                 System.out.println("Join room " + groupId);
                 Room room = Global.manager.getRoom(groupId);
-                User user = User.findById(new ObjectId(userId));
                 if (room != null) {
                     // Join room
-                    final UserSession userSession = room.join(user, out);
+                    final UserSession userSession = room.join(userId, out);
                     {
                         List<RecordingInterval> intervals = RecordingInterval.listByGroup(new ObjectId(groupId));
                         JSONObject msg = new JSONObject();
@@ -139,7 +139,7 @@ public class WSController extends Controller {
                                 case "createMessage": {
                                     String data = args.getString("data");
                                     CreateMessageService messageService = new CreateMessageService(
-                                            user.getId().toString(), groupId, data);
+                                            userId, groupId, data);
                                     try {
                                         Message message = messageService.execute();
                                         JSONArray messagesArray = new JSONArray();
@@ -228,7 +228,7 @@ public class WSController extends Controller {
                                     JSONObject msg = new JSONObject();
                                     msg.put("cmd", "talk");
                                     JSONObject data = new JSONObject();
-                                    data.put("uid", user.getId().toString());
+                                    data.put("uid", userId);
                                     data.put("value", value);
                                     msg.put("data", data);
                                     room.sendMessage(msg.toString());

### 2016-04-07 project/httpserver/app/models/DataPermission.java

@@ -8,7 +8,10 @@ import org.bson.Document;
 import org.bson.types.ObjectId;
 import services.Service;
 
-import java.util.*;
+import java.util.ArrayList;
+import java.util.List;
+import java.util.Map;
+import java.util.Set;
 
 
 /**

### 2016-04-07 project/httpserver/app/services/ListOwnerAttributesService.java

@@ -9,8 +9,6 @@ import org.bson.types.ObjectId;
 
 import java.util.ArrayList;
 import java.util.List;
-import java.util.Map;
-import java.util.Set;
 
 
 /**

### 2016-04-07 project/httpserver/app/services/SetAttributesService.java

@@ -1,11 +1,11 @@
 package services;
 
-import java.util.List;
-
 import dtos.AttributeDto;
 import exceptions.ServiceException;
 import models.Data;
 import org.bson.types.ObjectId;
+
+import java.util.List;
 
 
 /**

### 2016-04-07 project/httpserver/public/this/js/hyper.socket.js

@@ -1,5 +1,4 @@
 window.AudioContext = window.AudioContext || window.webkitAudioContext;
-navigator.getUserMedia = navigator.webkitGetUserMedia || navigator.mozGetUserMedia;
 
 var HyperWebSocket = new (function() {
     var primaryStream = null;

### 2016-04-08 tese2/13_implementation_model.tex
[ADD] speech detection

@@
 During a conversation, users may allow sharing their web cameras, by doing so their video is stored in recording chunks.
-Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains a reference to a group and a mapping between entity that was recorded and the correspondent video \ac{URL}.
+Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains a reference to a group and list of \ac{URL}s with the correspondent owner identification number and \emph{WebSocket} session id.
+
+In order to allow different devices within the same user, storing just the \ac{URL} with an associated user id is not enough as in this case it is needed one more parameter to differentiate the different \emph{webSocket} sessions opened by the same user.
+For this reason we had to associate a random session identification number to each \emph{webSocket}.
 

### 2016-04-08 tese2/16_implementation_content.tex
[ADD] speech detection

@@
 \end{figure}
+
+	\subsection{Switching views}
+
+		We provide a way for users to select the composite view of the conference room or the video shared by a particular user device as it can be seen on figure \ref{fig:devices}.
+
+		In order to maintain a list of available user devices that can be played at the navigated time if a user plays or ends playing a block of recorded video it obtains at the same time a list of all devices available for each user at that time.
+
+		On the other hand, if a user changes to real time, the list of devices is extracted from the set of \emph{webSockets} that are associated to the respective conference room.
+This list of devices is also sent to the users that are also in real time mode whenever a new user enters or leaves the conference room.
+
+	\begin{figure}[!htb]
+		\centering
+		\includegraphics[width=0.8\textwidth]{figures/devices.png}
+		\caption{Multiple devices per user}
+		\label{fig:devices}
+	\end{figure}
+
+		We also provide an automatic mechanism for switching the view to the current speaker.
+With respect to sound analysis, the sound samples are analyzed on the client side through the web audio \ac{API}.
+
+		Our speech detector is straightforward, we could perform a spectral analysis in order to understand if the analyzed sound contains frequencies in the range of the human voice but instead we just capture sound samples in real time and calculate the maximum sound amplitude.
+If a sound sample has an amplitude bigger than a factor of the maximum amplitude (we have used an empirical value of 10\%) we say that the user is speaking and therefore we send a message to the server if the speaking state has changed.
+Subsequently, the server receives the user speaking state and sends it to the other users so they can request a different view.

### 2016-04-08 tese2/evaluation.tex
[ADD] speech detection

@@
 Figure \ref{fig:test_full_features_cpu} shows the percentage of \ac{CPU} usage during our performance test case.
-As we can see, the amount of \ac{CPU} increases and decreases linearly in function of the amount of conference participants.
+Each 100\% represents one \ac{CPU} core, although that does not mean one \ac{CPU} is fully used, for example two cores at 60\% represent 120\% \ac{CPU} usage.
+As we can see, the percentage of \ac{CPU} used increases and decreases linearly in function of the amount of conference participants.
 

### 2016-04-08 tese2/stats/summary_full_cpu.tex
[ADD] speech detection

@@
 	    width=\textwidth,
-	   	height=6cm,
+	   	height=8cm,
 	    title={CPU usage},
@@
 
-		\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_mongo.dat};
+		\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/summary_cpu_mongo.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kms.dat};
+		\addplot[color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/summary_cpu_kms.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_kr.dat};
+		\addplot[color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/summary_cpu_kr.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/summary_full_cpu_play.dat};
+		\addplot[color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/summary_cpu_play.dat};
 		\addlegendentry{Java}

### 2016-04-08 tese2/stats/summary_full_net.tex
[ADD] speech detection

@@
 	    width=\textwidth,
-	   	height=6cm,
+	   	height=8cm,
 	    title={Network usage},

### 2016-04-08 tese2/stats/test_full_features_mem.tex
[ADD] speech detection

@@
 	    width=\textwidth,
-	   	height=6cm,
+	   	height=8cm,
 	    title={Memory Usage},

### 2016-04-08 tese2/stats/test_without_qrcode_mem.tex
[ADD] speech detection

@@
 	    width=\textwidth,
-	   	height=6cm,
+	   	height=8cm,
 	    title={Mem Usage},

### 2016-04-08 project/httpserver/app/main/UserSession.java

@@ -290,7 +290,7 @@ public class UserSession implements Closeable, Comparable<UserSession> {
         endPoint.release();
     }
 
-    public void sendChannels(){
+    public JSONObject sendChannels(){
         JSONArray channels = new JSONArray();
 
         if(timeOffset==0L) {
@@ -386,11 +386,15 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                 List<RecordingUrl> userRecs = rec.getUrl(userId !=null ? userId : room.getGroupId(),sessionId);
                 List<RecordingUrl> groupRecs = rec.getUrl(room.getGroupId(),sessionId);
 
-                if(userRecs.size()>0){
-                    ownerUrl =  userRecs.get(0).getUrl();
-                }
+
                 if(groupRecs.size()>0){
                     groupUrl = groupRecs.get(0).getUrl();
+                }
+
+                if(userRecs.size()>0){
+                    ownerUrl =  userRecs.get(0).getUrl();
+                }else {
+                    ownerUrl = groupUrl;
                 }
             }
 

### 2016-04-10 tese2/13_implementation_model.tex
[ADD] query next chunk while playing, [ADD] shedule next play if not the correct time

@@
 The model is the most important component of our solution, a badly designed model can imply serious difficulties when implementing new features that are not part of the plans, sometimes we had to redesign the model in order to support new features.
+
+\subsection {Schema representation}
+    \emph{MongoDB} has a slightly different terminology from relational databases.
+The first big difference is instead of having tables \emph{MongoDB} stores its objects on collections.
+The analogous data structure to the table row is a document.
+
+    Each \emph{MongoDB} document is represented by a \emph{JSON} object and, as a result, each document may have different attributes within a collection.
+Needless to say that following this approach we don't need to create the collections with a well predefined schema, in fact we don't to define it at all.
+Although, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure.
+
+    Similarly to relational databases, \emph{MongoDB} requires a primary key (typically typed as \emph{ObjectId} and named \emph{"\_id"}) for each document which is automatically assigned if not specified.
+
+    In order to define \emph{foreign keys}, we just store them as \emph{ObjectId}s if and only if the foreign keys point to documents within a unique collection, otherwise we need an additional attribute to specify which collection is the \emph{foreign key} pointing at.
+
+    In respect to attributes nullability, \emph{MongoDB} does not enforce a document's attribute to have a non nullable value, although, for sake of good functionality, we perform those constraints validation programatically and by so we also represent them in our schema.
+
+    An example of schema representation can be seen on table \ref{table:schema}.
+
+\begin{table}[!htb]
+\centering
+\caption{Schema representation}
+\label{table:schema}
+    \begin{tabular}{|ll|}
+        \hline
+        \multicolumn{2}{|c|}{\textbf{Collection name}}       \\ \hline
+        $\Diamondblack$ \underline{\_id (primary key)}  & ObjectId   \\ 
+        $\medbullet$ Not nullable property name           & Property type                    \\ 
+        $\medcirc$ Nullable property name      & Property type                         \\
+        $\medcirc$ Reference to document     & ObjectId                         \\
+        $\medbullet$ Embedded document      & Document                         \\ 
+        $\medbullet$ Embedded list      & List[Type]                         \\ \hline
+
+    \end{tabular}
+\end{table}
 
@@
             $\medbullet$ owner           & ObjectId                            \\ 
-            $\medbullet$ properties      & Document                         \\ 
-            $\medcirc$ searchableValue & List\textless Text\textgreater     \\ 
-            $\medcirc$ identiableKeys  & List\textless Text\textgreater     \\ \hline
+            $\medbullet$ properties      & List[Attribute] \\ 
+            $\medcirc$ searchableValues & List[Text]        \\ \hline
         \end{tabular}
@@
             \hline
-            \multicolumn{2}{|c|}{\textbf{DataPermission}}                   \\ \hline
-            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
-            $\medbullet$ owner           & ObjectId                         \\ 
-            $\medcirc$ readPermissions & Document                           \\ 
-            $\medcirc$ writePermissions  & Document                         \\ \hline
+            \multicolumn{2}{|c|}{\textbf{Attribute}}                           \\ \hline
+            $\medbullet$ key           & Text                                  \\ 
+            $\medcirc$ value           & Object                              \\ 
+            $\medbullet$ identifiable    & Boolean                             \\ 
+            $\medcirc$ readPermissions &  List[ObjectId]   \\ 
+            $\medcirc$ writePermissions  & List[ObjectId]  \\ \hline
         \end{tabular}    
@@
 
-The identication number by itself is not sufficient to identify an object, objects from different types can have the same identification number.
+The identification number by itself is not sufficient to identify an object, objects from different types can have the same identification number.
 In order to solve this problem, when an object is created, its correspondent data must contain the owner's object type.
 
-When an attribute is created it must be specified if the attribute is searchable, otherwise it could be simple to search for attributes that could reveal sensible information about it.
+Whenever an attribute is created, it must be specified the attributes name, its value, searcheability, identifiability and the set of object identification numbers that can read and write that attribute.
+
+In relation to our permission mechanism, if the read or write sets are not specified we assume the attribute is readable and writable by everyone, conversely if the read and write sets are empty, nobody is allowed to read or write the attribute.
+Implicitly if an entity can write an attribute it can also read it.
+
+In particular, if all attributes were searchable it could be simple to search for attributes that could reveal sensible information about an object.
 For example if we consider that a user could have a health related attribute, searching by a disease would reveal which users could suffer from a certain disease, the leak of that kind of information could, for instance, change the agreement between users and health insurance companies.
+For this reason only the specified attributes as searchable will be taken into account when performing keyword searches.
 
@@
 
+Not less important, in order to get an object properties efficiently, we have created an index over the \emph{owner} attribute.
+In the meantime , we have also created an index over the \emph{searchableValues} in order to improve the keyword search performance.
+
+In summary, with this model we can perform search and identification of any kind of objects, as we will see on the following models, the \emph{user} and \emph{group} models are using this generic model for storing their attributes.
 % Not less important, attributes can specify aggregations of objects.
@@
 
-Not less important, our model supports defining read and write permissions for each attribute, for example an attribute can be set to public read access or private write access to a set of entities.
-Implicitly if an entity can write an attribute it can also read it.
-
-The permission object is a mask of the data object, for each existing attribute in the data object it may exist the same attribute on the permission object specifying what entities can read or write it.
-
 \subsection{User model}
@@
 Passwords are not stored in plain text, instead we apply hashing and salting techniques \cite{password} in order to make it harder to decode the password by an attacker.
-We use \emph{SHA-1} and a random salt per user with 32 characters long.
+Accordingly, we use \emph{SHA-1} and a random salt per user with 32 characters long.
 

### 2016-04-10 project/httpserver/app/services/GetNextRecordingService.java

@@ -3,18 +3,16 @@ package services;
 import com.mongodb.client.FindIterable;
 import exceptions.BadRequestException;
 import models.RecordingChunk;
-import models.User;
 import org.bson.Document;
 import org.bson.types.ObjectId;
 
 import java.util.Date;
 
 
-
 /**
  * The Class GetCurrentRecordingService.
  */
-public class GetCurrentRecordingService extends Service<RecordingChunk> {
+public class GetNextRecordingService extends Service<RecordingChunk> {
 
     /** The caller. */
     private final ObjectId caller;
@@ -27,7 +25,7 @@ public class GetCurrentRecordingService extends Service<RecordingChunk> {
     private final ObjectId groupId;
 
     /** The time. */
-    private final Date time;
+    private final ObjectId recId;
     private final String sid;
 
 
@@ -36,16 +34,15 @@ public class GetCurrentRecordingService extends Service<RecordingChunk> {
      *
      * @param callerId the caller id
      * @param groupId the group id
-     * @param time the time
      */
-    public GetCurrentRecordingService(String callerId, String groupId, String owner, String sid, Date time) {
+    public GetNextRecordingService(String callerId, String groupId, String owner, String sid, ObjectId recId) {
         this.caller = new ObjectId(callerId);
         this.groupId = new ObjectId(groupId);
 
         System.out.println(owner+"|"+sid);
         this.owner = new ObjectId(owner);
         this.sid = sid;
-        this.time = time;
+        this.recId = recId;
     }
 
     /*
@@ -58,24 +55,12 @@ public class GetCurrentRecordingService extends Service<RecordingChunk> {
 
         RecordingChunk ans = getResult1();
 
-        if(ans==null){
-            // try group
-            ans = getResult2();
-        }
         return ans;
 
     }
 
     private RecordingChunk getResult1(){
-        Document query = new Document("gid", groupId).append("owner",owner).append("sid",sid).append("end", new Document("$gte", time));
-        FindIterable<Document> iter = RecordingChunk.getCollection().find(query).limit(1);
-        Document first = iter.first();
-        return first != null ? RecordingChunk.load(first) : null;
-    }
-
-
-    private RecordingChunk getResult2(){
-        Document query = new Document("gid", groupId).append("owner",groupId).append("end", new Document("$gte", time));
+        Document query = new Document("gid", groupId).append("owner",owner).append("sid",sid).append("_id", new Document("$gt", recId));
[…]

### 2016-04-10 tese2/13_implementation_model.tex
[ADD] thesis video play mechanisms

@@
 Our first decision for our model in order to avoid repeated code, was the isolation of the object's attributes from themselves, so we could apply the search operation to a set of attributes independently from the object type.
-To this generic set of properties we call data and each object of this type has a reference to the owner, which is a unique identification number.
+To this generic set of properties we call data (table \ref{table:generic}) and each object of this type has a reference to the owner, which is a unique identification number.
 
@@
 \caption{Generic data model}
-\label{my-label}
+\label{table:generic}
     \begin{tabular}{cc}
@@
 
-The user model is not tied to the user attributes, the information maintained in this model is just used for authentication purposes.
+The user model is not tied to the user attributes (table \ref{table:user}), the information maintained in this model is just used for authentication purposes.
 Passwords are not stored in plain text, instead we apply hashing and salting techniques \cite{password} in order to make it harder to decode the password by an attacker.
@@
 \caption{User model}
-\label{my-label}
+\label{table:user}
     \begin{tabular}{|ll|}
@@
 
-A relation between two entities $e_1$ and $e_2$ is represented by the pair $e_1\rightarrow e_2$, where $e_1$ is the source and $e_2$ is the target.
+A relation between two entities $e_1$ and $e_2$ is represented by the pair $e_1\rightarrow e_2$ (table \ref{table:relation}), where $e_1$ is the source and $e_2$ is the target.
 This relation is said bi-directional if and only if it also exists the relation $e_2\rightarrow e_1$.
@@
 \caption{Relation model}
-\label{my-label}
+\label{table:relation}
     \begin{tabular}{|ll|}
@@
 
-A group can be public or private.
+A group is composed by an \emph{id}, \emph{inviteToken} and a \emph{visibility} as shown on table \ref{table:group}.
+
+Moreover, a group can be public or private.
 If the group is public then it is visible to all users that maintain a friendship with a member of this group.
@@
 \caption{Group model}
-\label{my-label}
+\label{table:group}
     \begin{tabular}{cc}
@@
 
-A message is composed by its content, sequence number, time of creation and source and target identification numbers.
+A message is composed by its content, time of creation and source and target identification numbers (table \ref{table:message}).
 The message's target can reference any of object but our application is only covering messages to groups.
@@
 \caption{Message model}
-\label{my-label}
+\label{table:message}
     \begin{tabular}{|ll|}
@@
         $\medbullet$ target      & ObjectId                         \\ 
-        $\medbullet$ sequence      & Number                         \\ 
         $\medbullet$ content      & Text                         \\ \hline
@@
 During a conversation, users may allow sharing their web cameras, by doing so their video is stored in recording chunks.
-Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains a reference to a group and list of \ac{URL}s with the correspondent owner identification number and \emph{WebSocket} session id.
+Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains the media \ac{URL}, a reference to a group, an owner identification number and the correspondent \emph{WebSocket} session id.
 
@@
             $\medbullet$ groupId            & ObjectId                         \\ 
-            $\medbullet$ interval           & ObjectId                      \\ 
+            $\medbullet$ owner           & ObjectId                      \\ 
+            $\medbullet$ sessionId          & Text                       \\
             $\medbullet$ start              & Date                         \\ 
             $\medbullet$ end                & Date                         \\ 
-            $\medbullet$ sequenceNumber     & Number                         \\ 
-            $\medbullet$ urls               & Document                       \\ \hline
+            $\medbullet$ url               & Text                       \\ \hline
         \end{tabular}

### 2016-04-10 tese2/16_implementation_content.tex
[ADD] thesis video play mechanisms

@@
 
-	\subsection{Synchronization}
+	\subsection{Interactive media synchronization}
 
@@
 Subsequently, the server receives the user speaking state and sends it to the other users so they can request a different view.
+
+\subsection{Non interactive media synchronization}
+		By default each user receives the real time mixed video and audio of every users. As described on the previous section, users can switch the current view even for recorded media.
+
+		Whenever a user changes the current viewed time, our server calculates which chunks to reproduce.
+
+		In order to determine which chunks are going to be played we could find any chunk that matches the session id and intersects the given user time.
+Although this seems logical, we limit the size of the results to one and ignore the starting component of the recording chunk.
+As a consequence, we will obtain the recording chunk that ends after the playing time.
+
+		The obtained chunk may or not intersect the given time, if it intersects than we immediately start playing the chunk, otherwise we know that the chunks starting time is after the playing time and we schedule the player to reproduce after a time duration given by the difference of the playing time and chunk starting time.
+
+		In order to reduce the delay between switching parts we perform a query on the database for finding the next chunk to play while the selected media is playing.
+In order to find the next chunk we use the chunk that has a bigger id immediately after the playing chunk.
+
+		In order to allow the users to watch an individual image and ear the audio of everyone, we are using two players, one for audio and another for video.
+
+		If a user leaves a conference room, it is expected that the sequence of played blocks ends.
+When a block is not found for a given \emph{webSocket} session we try to find a block for the same user and if not possible we try to play a composite recorded block, lastly, if every attempt fails we wait for the next interval the contains media.

### 2016-04-10 tese2/13_implementation_model.tex
[ADD] indices until message

@@
         $\medbullet$ Embedded list      & List[Type]                \\ \hline
-
     \end{tabular}
@@
             \multicolumn{2}{|c|}{\textbf{Data}}            \\ \hline
-            $\Diamondblack$ \underline{id}  & ObjectId                         \\ 
+            $\Diamondblack$ \underline{\_id}  & ObjectId     \\ 
             $\medbullet$ owner           & ObjectId        \\ 
@@
         \multicolumn{2}{|c|}{\textbf{User}}         \\ \hline
-        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+        $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
         $\medbullet$ hash           & Text          \\ 
@@
         \multicolumn{2}{|c|}{\textbf{Relation}}     \\ \hline
-        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+        $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
         $\medbullet$ source           & ObjectId    \\ 
@@
 In order to validate a friendship, both users must agree on that friendship, by other words it must exist a bi-directional relation between both users.
+
+In order to improve performance of queries over \emph{Relation} collection we have created indexes on the \emph{source}, \emph{target} and also on the pair composed by both attributes.
 
@@
             \multicolumn{2}{|c|}{\textbf{Group}}        \\ \hline
-            $\Diamondblack$ \underline{id}  & ObjectId       \\ 
+            $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
             $\medcirc$ inviteToken           & Text     \\ 
@@
             \multicolumn{2}{|c|}{\textbf{GroupMembership}}  \\ \hline
-            $\Diamondblack$ \underline{id}  & ObjectId       \\ 
+            $\Diamondblack$ \underline{\_id}  & ObjectId      \\ 
             $\medbullet$ groupId            & ObjectId      \\ 
@@
 Entities that have a membership with a group can create more memberships by sharing an invite token or by specifying new group members.
+
+In order to improve performance of queries over \emph{GroupMembership} collection we have created indexes on the \emph{groupId}, \emph{userId} and also on the pair composed by both attributes.
 
@@
         \multicolumn{2}{|c|}{\textbf{Message}}      \\ \hline
-        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+        $\Diamondblack$ \underline{\_id}  & ObjectId  \\ 
         $\medbullet$ source           & ObjectId    \\ 
@@
 \end{table}
+
+In order to improve performance of queries over \emph{Relation} collection we have created indexes on the \emph{target} attribute and also on the pair composed by \emph{target} and \emph{\_id} attributes for finding and sorting the messages received by an entity more efficiently.
 
@@
             \multicolumn{2}{|c|}{\textbf{HyperContent}}  \\ \hline
-            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\Diamondblack$ \underline{\_id}  & ObjectId   \\ 
             $\medbullet$ groupId            & ObjectId   \\ 
@@
             \multicolumn{2}{|c|}{\textbf{TimeAnnotation}}  \\ \hline
-            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\Diamondblack$ \underline{\_id}  & ObjectId     \\ 
             $\medbullet$ groupId            & ObjectId     \\ 
@@
         \multicolumn{2}{|c|}{\textbf{CollaborativeContent}}  \\ \hline
-        $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+        $\Diamondblack$ \underline{\_id}  & ObjectId           \\ 
         $\medbullet$ groupId            & ObjectId           \\ 
@@
             \multicolumn{2}{|c|}{\textbf{RecordingInterval}}  \\ \hline
-            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\Diamondblack$ \underline{\_id}  & ObjectId        \\ 
             $\medbullet$ groupId            & ObjectId        \\ 
@@
             \multicolumn{2}{|c|}{\textbf{RecordingChunk}}  \\ \hline
-            $\Diamondblack$ \underline{id}  & ObjectId                      \\ 
+            $\Diamondblack$ \underline{\_id}  & ObjectId     \\ 
             $\medbullet$ groupId            & ObjectId     \\ 

### 2016-04-10 project/httpserver/app/main/UserSession.java

@@ -28,11 +28,8 @@ public class UserSession implements Closeable, Comparable<UserSession> {
 
     /** The out. */
     private final WebSocket.Out<String> out;
-    
     private final Room room;
-
     private final Object playerLock = new Object();
-
     private boolean play = true;
     private long timeOffset = 0L;
     private String playSid;
@@ -41,19 +38,12 @@ public class UserSession implements Closeable, Comparable<UserSession> {
     private Boolean hasAudio = true;
     private Boolean hasVideo = true;
     private UUID sid = UUID.randomUUID();
-
     private Recorder recorder;
     private ZBarFilter qrCodeFilter;
-
-    public String getUserId() {
-        return userId;
-    }
-
     private WebRtcEndpoint endPoint;
     private HubPort compositePort;
     private PlayerEndpoint playerVideo;
     private PlayerEndpoint playerAudio;
-
     private RecordingChunk nextVideo;
     private RecordingChunk nextAudio;
     private Timer scheduledPlayer;
@@ -443,19 +433,15 @@ public class UserSession implements Closeable, Comparable<UserSession> {
             if (playerAudio != null) {
                 playerAudio.release();
             }
-
             playerVideo = tempVideo;
             playerAudio = tempAudio;
 
 
             if (play) {
-
                 tempAudio.play();
                 tempVideo.play();
-
                 tempAudio.connect(endPoint, MediaType.AUDIO);
                 tempVideo.connect(endPoint, MediaType.VIDEO);
-
                 if(tempVideo.getVideoInfo().getIsSeekable() && tempAudio.getVideoInfo().getIsSeekable()) {
                     // this is the ideal behavior
                     tempVideo.setPosition(seekVideo);
@@ -470,7 +456,6 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                     timeOffset = new Date().getTime()- audioChunk.getStart().getTime();
                     sendMessage(msg.toString());
                 }
-
                 try {
                     // prepare next audio
                     if (nextAudio == null) {
@@ -547,7 +532,6 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                 System.out.println("Already playing!");
             }
         }
-
         sendChannels();
     }
 
@@ -701,6 +685,9 @@ public class UserSession implements Closeable, Comparable<UserSession> {
 
     public String getSid() {
         return sid.toString();
+    }
+    public String getUserId() {
+        return userId;
     }
 
 }

### 2016-04-10 project/httpserver/app/models/Message.java

@@ -61,8 +61,11 @@ public class Message {
      * @return the collection
      */
     private static MongoCollection<Document> getCollection() {
-        if (collection == null)
+        if (collection == null) {
             collection = Service.getDatabase().getCollection(Message.class.getName());
+            collection.createIndex(new Document("target",1));
+            collection.createIndex(new Document("target",1).append("_id",1));
+        }
         return collection;
     }
 
@@ -123,19 +126,15 @@ public class Message {
         Document doc = new Document();
         if (id != null)
             doc.put("_id", id);
-
         doc.put("target", target);
         doc.put("source", source);
         doc.put("time", time);
         doc.put("text", content);
-
         if (id == null) {
             getCollection().insertOne(doc);
         } else
             getCollection().replaceOne(new Document("_id", id), doc);
-
         id = doc.getObjectId("_id");
-
     }
 
     /**

### 2016-04-10 project/httpserver/app/models/Relation.java

@@ -35,10 +35,8 @@ public class Relation {
      * @param ti the ti
      */
     public Relation(ObjectId fi, ObjectId ti) {
-
         this.target = ti;
         this.source = fi;
-
     }
 
     /**
@@ -47,8 +45,12 @@ public class Relation {
      * @return the collection
      */
     private static MongoCollection<Document> getCollection() {
-        if (collection == null)
+        if (collection == null) {
             collection = Service.getDatabase().getCollection(Relation.class.getName());
+            collection.createIndex(new Document("s",1));
+            collection.createIndex(new Document("t",1));
+            collection.createIndex(new Document("s",1).append("t",1));
+        }
         return collection;
     }
 
@@ -61,9 +63,8 @@ public class Relation {
     private static Relation load(Document doc) {
         Relation user = new Relation();
         user.setId(doc.getObjectId("_id"));
-        user.setFrom(doc.getObjectId("fi"));
-        user.setTo(doc.getObjectId("ti"));
-
+        user.setFrom(doc.getObjectId("s"));
+        user.setTo(doc.getObjectId("t"));
         return user;
     }
 
@@ -75,7 +76,7 @@ public class Relation {
      * @return the relation
      */
     public static Relation findByEndpoint(ObjectId a, ObjectId b) {
-        Document doc = new Document("fi", a).append("ti", b);
+        Document doc = new Document("s", a).append("t", b);
         FindIterable<Document> iter = getCollection().find(doc);
         doc = iter.first();
         return doc != null ? load(doc) : null;
@@ -88,10 +89,8 @@ public class Relation {
      * @return the list
      */
     public static List<Relation> listFrom(ObjectId ep) {
-        Document doc = new Document("fi", ep);
+        Document doc = new Document("s", ep);
         FindIterable<Document> iter = getCollection().find(doc);
-        doc = iter.first();
-
         Iterator<Document> i = iter.iterator();
         List<Relation> ret = new ArrayList<Relation>();
         while (i.hasNext()) {
@@ -107,10 +106,8 @@ public class Relation {
      * @return the list
      */
     public static List<Relation> listTo(ObjectId ep) {
-        Document doc = new Document("ti", ep);
+        Document doc = new Document("t", ep);
         FindIterable<Document> iter = getCollection().find(doc);
-        doc = iter.first();
-
         Iterator<Document> i = iter.iterator();
         List<Relation> ret = new ArrayList<Relation>();
[…]

### 2016-04-11 tese2/13_implementation_model.tex
[ADD] chat stuff, collaborative editor

@@
 
-    In respect to attributes nullability, \emph{MongoDB} does not enforce a document's attribute to have a non nullable value, although, for sake of good functionality, we perform those constraints validation programatically and by so we also represent them in our schema.
+    In respect to attributes nullability, \emph{MongoDB} does not enforce a document's attribute to have a non nullable value, although, for sake of good functionality, we perform those constraints validation programmatically and by so we also represent them in our schema.
 
@@
 
-A time annotation contains a title, the correspondent group identification number and the time itself.
+A time annotation (table \ref{table:hyper}) contains a title, the correspondent group identification number and the time itself.
 
 The hyper content is used to synchronize content among users during a conversation.
-Every hyper content must have a start and ending time, the correspondent group identification number and the content itself in the form of text.
+Table \ref{table:hyper} shows that every hyper content must have a start and ending time, the correspondent group identification number and the content itself in the form of text.
+Beside those attributes, in order to perform queries over the \ac{HTML} contents with more precision, we have added an additional \emph{searchableContent} attribute that contains just the searchable content extracted from the content excluding the \ac{HTML} tags and parameters using \emph{Jsoup}\footnote{\url{http://jsoup.org/} (Accessed April 11, 2016)}.
 
@@
 \caption{Hyper content model}
-\label{my-label}
+\label{table:hyper}
     \begin{tabular}{cc}
@@
             $\medbullet$ end                & Date       \\ 
-            $\medbullet$ content            & Text       \\ \hline
+            $\medbullet$ content            & Text       \\ 
+            $\medbullet$ searchableContent  & Text       \\ \hline
         \end{tabular}
@@
 
+For example the content showed in listing \ref{lst:jsoup} would produce the searchable content "Click here!".
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Example of HTML content},label={lst:jsoup},language=JavaScript]
+<div class="subtitle">
+    <a href="/">Click here!</a>
+</div>
+\end{lstlisting}
+\end{minipage}
+
+In order to improve the performance of searching content and calculate interval intersections, we have created indexes over the following sequences of attributes:
+
+\begin{itemize}
+\item{\emph{\{groupId,start,end\}} $\rightarrow$ finding visible contents for a given time  within a group (intersections).}
+\item{\emph{\{groupId,start\}} $\rightarrow$  finding contents that starts after a given time within a group (for pre-loading).}
+\item{\emph{\{groupId,searchableContent\}} $\rightarrow$ searching contents by keywords within a group.}  
+\end{itemize}
+
 \subsection{Collaborative Content model}
 
-Within a conversation, users can write documents collaboratively. Each document has a content and a reference to the correspondent group.
+Within a conversation, users can write documents collaboratively. As we can see in table \ref{table:collaborative}, each document has a content and a reference to the correspondent group.
 
@@
     \caption{Collaborative content model}
-    \label{my-label}
+    \label{table:collaborative}
     \begin{tabular}{|ll|}
@@
         $\medbullet$ groupId            & ObjectId           \\ 
-        $\medbullet$ content            & Text               \\ \hline
+        $\medcirc$ content            & Text               \\ \hline
     \end{tabular}
 \end{table}
+
+In order to improve the performance of finding the group's content we have created and index over the \emph{groupId} attribute.
 
@@
 During a conversation, users may allow sharing their web cameras, by doing so their video is stored in recording chunks.
-Each chunk represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains the media \ac{URL}, a reference to a group, an owner identification number and the correspondent \emph{WebSocket} session id.
+Each chunk, as seen in table \ref{table:recording}, represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains the media \ac{URL}, a reference to a group, an owner identification number and the correspondent \emph{WebSocket} session id.
 
@@
     \caption{Recording model}
-    \label{my-label}
+    \label{table:recording}
     \begin{tabular}{cc}
@@
 A recording interval represents a continuous set of recording chunks.
+
+In order to improve the performance of searching recording chunks and calculate which chunks are are going to be reproduced, we have created indexes over the following sequences of attributes: 
+\begin{itemize}
+    \item{\emph{\{groupId,start,end\}} $\rightarrow$ finding all available chunks for a given time} 
+    \item{\emph{\{groupId,sessionId,end\}} $\rightarrow$ finding intersecting chunks with a given time within a session}
+    \item{\emph{\{groupId,owner,end\}} $\rightarrow$ finding intersecting chunks with a given time within a user (for a different session) or group}
+    \item{\emph{\{groupId,sessionId,\_id\}} $\rightarrow$ finding chunks that follow a given \emph{id} within a session}
+    \item{\emph{\{groupId,owner,\_id\}} $\rightarrow$ finding chunks that follow a given \emph{id} within a user or group}
+\end{itemize}
+
+Not less important we have also created an index over the \emph{RecordingInterval}'s \emph{groupId} attribute for listing all intervals withing a conference room.

### 2016-04-11 tese2/16_implementation_content.tex
[ADD] chat stuff, collaborative editor

@@
 When a block is not found for a given \emph{webSocket} session we try to find a block for the same user and if not possible we try to play a composite recorded block, lastly, if every attempt fails we wait for the next interval the contains media.
+
+\subsection{Chat}
+		Lastly and not less important the chat functionality was implemented using \emph{webSockets}, when a message is sent by a user, it is received by the server, which records it and sends to all the users.
+
+		In order to support receiving a huge amount of messages and not having a bottleneck when loading the conference room, we just load an amount of messages that fill the chat panel.
+When a user scrolls the chat to read old messages, we detect when the chat scroll bar reaches the limit and we request more older messages to the server automatically.
+
+		Moreover, when the clients receive messages, whether older or newer ones, their ids are used to place them in the correct position.
+By other words, older messages are placed in the top of the chat panel and newer ones in the bottom.
+
+		Besides that, we also support sending files to other users, which are stored on database and accessed through an \ac{URL}.
+
+\subsection{Collaborative editor}
+
+Our collaborative editor is a simple text editor that is synchronized with all participants within a conference room implemented using \emph{ot.js}.
+
+The state of our collaborative editor is not saved on database every time it changes. Instead the users just synchronize the editor content among them and save it a priori.
+
+If a user joins a conference room which editor was changed and the content differs from the version stored in the database, our server does not have the right content to deliver the user.
+As a consequence our server names a random participant to yield its collaborative content and delivers it to the joining user.
+
+All operations that we perform over the conference room's collaborative editor are performed in mutual exclusion so all updates are delivered in the same order to all participants.
+For example consider two write events $W_1$ and $W_2$ if they occur at the same time, our server will receive one of the events first, say $W_1$, and delivers is to every users including its writer.
+Only then the server delivers the $W_2$ content.
+With this serialization mechanism we can ensure the content is equal on every user editor.
+
+Accordingly to \emph{ot.js} there are three types of operations:
+\begin{itemize}
+\item{retain} $\rightarrow$ advance the cursor by a given number of positions.
+\item{insert} $\rightarrow$ write the given text in the current position.
+\item{delete} $\rightarrow$ delete a given number of characters forward.
+\end{itemize}
+
+In summary, our server just ensures mutual exclusion and delivers the messages among users without even parsing the messages content.

### 2016-04-11 project/httpserver/app/main/Tools.java

@@ -52,63 +52,6 @@ public class Tools {
         return builder.toString();
     }
 
-    private static Pattern HTML_PATTERN = Pattern.compile("<(\\w+)>.*?</\\1>");
-
-    public static String getTextFromHtml(String html){
-/*
-        Matcher m = HTML_PATTERN.matcher(html);
-        String text = "";
-        while (m.find()) {
-            text += m.group();
-        }
-        return text;
-*/
-
-        return Jsoup.parse(html).text();
-    }
-
-    /**
-     * Friendly time.
-     *
-     * @param time the time
-     * @return the string
-     */
-    static public String friendlyTime(String time) {
-        try {
-
-            if (time.charAt(time.length() - 1) != 'Z') {
-                time += "Z";
-            }
-
-
-            ZonedDateTime zdt = ZonedDateTime.parse(time);
-            LocalDateTime ldt = zdt.toLocalDateTime();
-
-
-            return String.format("%02d", ldt.getDayOfMonth()) + "/" + String.format("%02d", ldt.getMonthValue()) + "/" + String.format("%04d", ldt.getYear())
-                    + " " + String.format("%02d", ldt.getHour()) + ":" + String.format("%02d", ldt.getMinute()) + ":" + String.format("%02d", ldt.getSecond());
-        } catch (Exception e) {
-            e.printStackTrace();
-            // return "00/00/0000 00:00:00";
-        }
-        return time;
-    }
-
-    /**
-     * Gets the current time.
-     *
-     * @return the current time
-     */
-    public static String getCurrentTime() {
-        return FORMAT.format(new Date());
-        //return DateTimeFormatter.ofPattern("yyyy-MM-dd'T'HH:mm:ss'Z'").withZone(ZoneOffset.UTC).format(Instant.now());
-    }
-
-    public static int clamp(int min, int x, int max) {
-        return (x<min? min :(x>=max? max : x));
-
-    }
-
     public static String md5(String content) {
         try {
             final MessageDigest messageDigest = MessageDigest.getInstance("MD5");

### 2016-04-11 project/httpserver/app/models/CollaborativeContent.java

@@ -48,8 +48,10 @@ public class CollaborativeContent {
      * @return the collection
      */
     public static MongoCollection<Document> getCollection() {
-        if (collection == null)
+        if (collection == null) {
             collection = Service.getDatabase().getCollection(CollaborativeContent.class.getName());
+            collection.createIndex(new Document("gid",1));
+        }
         return collection;
     }
 

### 2016-04-11 project/httpserver/app/models/HyperContent.java

@@ -6,6 +6,7 @@ import main.Tools;
 import org.bson.Document;
 import org.bson.types.ObjectId;
 import org.json.JSONObject;
+import org.jsoup.Jsoup;
 import services.ListCurrentRecordingsService;
 import services.Service;
 
@@ -29,7 +30,9 @@ public class HyperContent {
     
     /** The content. */
     private String content = null;
-    
+
+    private String searchableContent;
+
     /** The group id. */
     private ObjectId id, groupId = null;
 
@@ -46,7 +49,15 @@ public class HyperContent {
         this.end = end;
         this.groupId = gid;
         this.content = content;
+        this.searchableContent = getTextFromHtml(content);
     }
+
+
+    private static String getTextFromHtml(String html){
+
+        return Jsoup.parse(html).text().toLowerCase();
+    }
+
 
     /**
      * Instantiates a new hyper content.
@@ -62,6 +73,9 @@ public class HyperContent {
     public static MongoCollection<Document> getCollection() {
         if (collection == null) {
             collection = Service.getDatabase().getCollection(HyperContent.class.getName());
+            collection.createIndex(new Document("gid",1).append("start",1).append("end",1));
+            collection.createIndex(new Document("gid",1).append("start",1));
+            collection.createIndex(new Document("gid",1).append("search","Text"));
         }
         return collection;
     }
@@ -79,6 +93,8 @@ public class HyperContent {
         content.end = doc.getDate("end");
         content.groupId = doc.getObjectId("gid");
         content.content = doc.getString("content");
+        content.searchableContent = doc.getString("search");
+
         return content;
     }
 
@@ -104,7 +120,7 @@ public class HyperContent {
      */
     public static List<HyperContent> search(ObjectId gid, String query) {
         Pattern regex = Pattern.compile(query, Pattern.CASE_INSENSITIVE);
-        Document doc = new Document("gid", gid).append("content", regex);
+        Document doc = new Document("gid", gid).append("search", regex);
         FindIterable<Document> iter = getCollection().find(doc);
         return deserialize(iter.iterator());
     }
@@ -121,6 +137,7 @@ public class HyperContent {
         doc.put("end", end);
         doc.put("gid", groupId);
         doc.put("content", content);
+        doc.put("search", searchableContent);
         if (id == null) {
             getCollection().insertOne(doc);
         }else {
@@ -226,4 +243,7 @@ public class HyperContent {
         return obj;
     }
 
+    public String getSearchableContent() {
+        return searchableContent;
+    }
 }

### 2016-04-11 project/httpserver/app/models/RecordingChunk.java

@@ -67,7 +67,11 @@ public class RecordingChunk {
     public static MongoCollection<Document> getCollection() {
         if (collection == null) {
             collection = Service.getDatabase().getCollection(RecordingChunk.class.getName());
-            collection.createIndex(new Document("gid",1));
+            collection.createIndex(new Document("gid",1).append("start",1).append("end",1));
+            collection.createIndex(new Document("gid",1).append("sid",1).append("end",1));
+            collection.createIndex(new Document("gid",1).append("owner",1).append("end",1));
+            collection.createIndex(new Document("gid",1).append("sid",1).append("_id",1));
+            collection.createIndex(new Document("gid",1).append("owner",1).append("_id",1));
         }
         return collection;
     }
@@ -90,51 +94,6 @@ public class RecordingChunk {
         return rec;
     }
 
-    /**
-     * Count by group.
-     *
-     * @param owner the owner
-     * @return the long
-     */
-    public static long countByGroup(ObjectId groupId , ObjectId owner, String sid) {
-        Document doc = new Document("gid", groupId).append("owner",owner).append("sid",sid);
-        return getCollection().count(doc);
-    }
-
-/*
-    public List<RecordingUrl> getUrls(){
-        List<RecordingUrl> ans = new ArrayList<RecordingUrl>();
-        for(Document doc : urls){
-            RecordingUrl ru = new RecordingUrl(doc);
-            ans.add(ru);
-        }
-
-        return ans;
-    }
-
-
-    public List<RecordingUrl> getUrl(String owner, String sid){
-        List<RecordingUrl> ans = new ArrayList<RecordingUrl>();
-        for(Document doc : urls){
-            RecordingUrl ru = new RecordingUrl(doc);
-            if(ru.getUid()!=null && ru.getUid().equals(owner)){
-                if(ru.getSid().equals(sid)) {
-                    ans.add(0, ru);
-                }else {
-                    ans.add(ru);
-                }
-            }
-        }
-
-        return ans;
-    }
-
-
-    public synchronized void setUrl(RecordingUrl ru){
-       urls.add(ru.toDocument());
-    }
-
-*/
 
 
     /**
@@ -155,23 +114,21 @@ public class RecordingChunk {
      */
     public synchronized void save() {
         Document doc = new Document();
-        if (id != null)
+        if (id != null) {
             doc.put("_id", id);
+        }
         doc.put("owner", owner);
         doc.put("gid", groupId);
         doc.put("start", start);
[…]

### 2016-04-11 project/httpserver/app/services/GetCurrentRecordingService.java

@@ -57,14 +57,19 @@ public class GetCurrentRecordingService extends Service<RecordingChunk> {
     public RecordingChunk dispatch() throws BadRequestException {
         RecordingChunk ans = getResult1();
         if(ans==null){
-            // try group
+            // try same user
             ans = getResult2();
+        }
+
+        if(ans==null){
+            // try group
+            ans = getResult3();
         }
         return ans;
     }
 
     private RecordingChunk getResult1(){
-        Document query = new Document("gid", groupId).append("owner",owner).append("sid",sid).append("end", new Document("$gte", time));
+        Document query = new Document("gid", groupId).append("sid",sid).append("end", new Document("$gte", time));
         FindIterable<Document> iter = RecordingChunk.getCollection().find(query).limit(1);
         Document first = iter.first();
         return first != null ? RecordingChunk.load(first) : null;
@@ -72,6 +77,13 @@ public class GetCurrentRecordingService extends Service<RecordingChunk> {
 
 
     private RecordingChunk getResult2(){
+        Document query = new Document("gid", groupId).append("owner",owner).append("end", new Document("$gte", time));
+        FindIterable<Document> iter = RecordingChunk.getCollection().find(query).limit(1);
+        Document first = iter.first();
+        return first != null ? RecordingChunk.load(first) : null;
+    }
+
+    private RecordingChunk getResult3(){
         Document query = new Document("gid", groupId).append("owner",groupId).append("end", new Document("$gte", time));
         FindIterable<Document> iter = RecordingChunk.getCollection().find(query).limit(1);
         Document first = iter.first();

### 2016-04-11 project/httpserver/app/services/GetNextRecordingService.java

@@ -54,19 +54,40 @@ public class GetNextRecordingService extends Service<RecordingChunk> {
     public RecordingChunk dispatch() throws BadRequestException {
 
         RecordingChunk ans = getResult1();
-
+        if(ans==null){
+            // try same user
+            ans = getResult2();
+        }
+        if(ans==null){
+            // try group
+            ans = getResult3();
+        }
         return ans;
 
     }
 
     private RecordingChunk getResult1(){
-        Document query = new Document("gid", groupId).append("owner",owner).append("sid",sid).append("_id", new Document("$gt", recId));
+        Document query = new Document("gid", groupId).append("sid",sid).append("_id", new Document("$gt", recId));
+        FindIterable<Document> iter = RecordingChunk.getCollection().find(query).limit(1);
+        Document first = iter.first();
+        return first != null ? RecordingChunk.load(first) : null;
+    }
+
+    private RecordingChunk getResult2(){
+        Document query = new Document("gid", groupId).append("owner",owner).append("_id", new Document("$gt", recId));
         FindIterable<Document> iter = RecordingChunk.getCollection().find(query).limit(1);
         Document first = iter.first();
         return first != null ? RecordingChunk.load(first) : null;
     }
 
 
+
+    private RecordingChunk getResult3(){
+        Document query = new Document("gid", groupId).append("owner",groupId).append("_id", new Document("$gt", recId));
+        FindIterable<Document> iter = RecordingChunk.getCollection().find(query).limit(1);
+        Document first = iter.first();
+        return first != null ? RecordingChunk.load(first) : null;
+    }
 
 
     /*

### 2016-04-11 project/httpserver/app/services/ListCurrentRecordingsService.java

@@ -47,7 +47,7 @@ public class ListCurrentRecordingsService extends Service<List<RecordingChunk>> 
     @Override
     public List<RecordingChunk> dispatch() throws BadRequestException {
         List<RecordingChunk> ans = new ArrayList<>();
-        Document query = new Document("gid", groupId).append("end", new Document("$gte", time)).append("start", new Document("$lt", time));
+        Document query = new Document("gid", groupId).append("start", new Document("$lt", time)).append("end", new Document("$gte", time));
 
 
         FindIterable<Document> findIter = RecordingChunk.getCollection().find(query);

### 2016-04-17 tese2/evaluation.tex
[ADD] user tests results

@@
 
-On the fourth task both users didn't placed their annotations on wright place, we only provide one way to place annotations, which is by dragging them on the time-line. 
+On the fourth task both users didn't placed their annotations on right place, we only provide one way to place annotations, which is by dragging them on the time-line. 
 After creating an annotation users also didn't understand that they had to save their changes on the time-line, we solved this problem by saving automatically each time the user changed any annotation, the old user interface can be seen on figure \ref{fig:test_ui_05}.
@@
   \subsubsection {Second phase}
+
+On the second phase of our user interface tests, in a general view, we have noticed great improvements on the learning time.
+
+From the data that we have collected, namely the task duration (figure \ref{fig:user_times}), number of clicks (figure \ref{fig:user_clicks}), number of errors (figure \ref{fig:user_errors}), task difficulty (figure \ref{fig:user_diffs}) and qualitative evaluation (figure \ref{fig:user_evals}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.
+
+As a result of both true average and variance being unknown and the usage of a relatively small amount of samples, we had to use \emph{t-distribution} to estimate our metrics confidence intervals.
+
+Accordingly, we have used the following formula to calculate the intervals for the expected value of the true average with 95\% confidence:
+
+$[\bar{x}-F^{-1}_{t_{(n-1)}}(1-\alpha/2)\times \frac{s}{\sqrt{n}}, \bar{x}+F^{-1}_{t_{(n-1)}}(1-\alpha/2)\times \frac{s}{\sqrt{n}}]$
+
+Where $\alpha=0.05$, $n$ is the number of samples (which may not coincide with the number of tests due to failed tasks) and $s$ is the sample standard deviation that is given by:
+
+$s=\sqrt{\frac{1}{n-1}\Sigma^{n}_{i=1}(x_i-\bar{x})^2}$
+
+\begin{figure}[!htb]
+\centering
+\begin{minipage}[b]{0.45\linewidth}
+  \begin{center}
+    \input{stats/user_times.tex}
+  \end{center}
+  \caption{Time spent per task}
+  \label{fig:user_times}
+\end{minipage}
+\begin{minipage}[b]{0.45\linewidth}
+   \begin{center}
+    \input{stats/user_clicks.tex}
+  \end{center}
+  \caption{Mouse clicks per task}
+  \label{fig:user_clicks}
+\end{minipage}
+\end{figure}
+
+Accordingly to figure \ref{fig:user_evals}, we can observe that most users had less difficulties with the first two tasks, which represents types of tasks that most users are familiar with.
+As soon as users had to navigate in time, manipulate annotations and create content (respectively \emph{task3}, \emph{task4} and \emph{task5}) we have observed that they revealed more difficulties.
+Most of those difficulties, based on the users feedback, were mainly due to those concepts not being familiar to them.
+
+\begin{figure}[!htb]
+\centering
+\begin{minipage}[b]{0.45\linewidth}
+  \begin{center}
+    \input{stats/user_errors.tex}
+  \end{center}
+  \caption{Errors per task}
+  \label{fig:user_errors}
+\end{minipage}
+\begin{minipage}[b]{0.45\linewidth}
+  \begin{center}
+    \input{stats/user_diffs.tex}
+  \end{center}
+  \caption{Difficulty per task}
+  \label{fig:user_diffs}
+\end{minipage}
+\end{figure}
+
+Although we had relatively bad results with some users, we have explained to the users that could not conclude the tasks or performed incorrectly the most efficient way to perform the requested tasks.
+Some users suggested to display more hints in order to achieve a faster learning but after all they were impressed and gave us a better evaluation (as seen on figure \ref{fig:user_evals}).
+
+\begin{figure}[!htb]
+  \begin{center}
+    \input{stats/user_evals.tex}
+  \end{center}
+  \caption{Solution evaluation}
+  \label{fig:user_evals}
+\end{figure}
+
+Most users gave us worse evaluations on our user interface layout and content editor, which was due to having a lot of tools present in the same web page and some of them were hidden due the screen size.
+In some cases users had to scroll down in order to find the tools they were looking for.
+
+Another weak aspect was our content editor, which in fact we recognize the difficulty to handle with, most due to the amount of information that is necessary to create a synchronized content (starting time, duration and content itself).
+Some users have suggested that the content should also be present on the time-line so they could be easily dragged and resized (on time).
+
+We are aware that placing content on the time-line will reduce our solutions performance, especially when there is a relatively big amount of content due to the content that is present on the time-line being loaded all in once.
+Although, we recognize that for some cases (relatively low amount of content), displaying the content on the time-line could not have a great impact on our solutions performance.

### 2016-04-17 tese2/stats/test_client_net.tex
[ADD] user tests results

@@
 
-	\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-0,y expr=\thisrowno{1}/1000] {stats/client/send.dat};
+	\addplot[color=Blue] table[x expr=\thisrowno{0}-0,y expr=\thisrowno{1}/1000] {stats/client/send.dat};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-0,y expr=\thisrowno{1}/1000] {stats/client/recv.dat};
+	\addplot[color=Red] table[x expr=\thisrowno{0}-0,y expr=\thisrowno{1}/1000] {stats/client/recv.dat};
 	\addlegendentry{Recv(eth0)}

### 2016-04-17 tese2/stats/test_full_features_cpu.tex
[ADD] user tests results

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_7_mongod.dat};
+		\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_67_kurento-media-server.dat};
+		\addplot[color=Red] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_67_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_35_kurento-repo.dat};
+		\addplot[color=Green] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_233_bash.dat};
+		\addplot[color=Magenta] table[x expr=\thisrowno{0}-122,y index=1] {stats/test_full_features/cpu_233_bash.dat};
 		\addlegendentry{Java}

### 2016-04-17 tese2/stats/test_ram_fixed_mem.tex
[ADD] user tests results

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_7_mongod.dat};
+		\addplot[color=Blue] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_67_kurento-media-server.dat};
+		\addplot[color=Red] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_67_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_35_kurento-repo.dat};
+		\addplot[color=Green] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_233_bash.dat};
+		\addplot[color=Magenta] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_233_bash.dat};
 		\addlegendentry{Java}

### 2016-04-17 tese2/stats/test_two_times_cpu.tex
[ADD] user tests results

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_7_mongod.dat};
+		\addplot[color=Blue] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_68_kurento-media-server.dat};
+		\addplot[color=Red] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_68_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_35_kurento-repo.dat};
+		\addplot[color=Green] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_234_bash.dat};
+		\addplot[color=Magenta] table[x expr=\thisrowno{0}-72,y index=1] {stats/test_two_times3/cpu_234_bash.dat};
 		\addlegendentry{Java}

### 2016-04-17 tese2/stats/test_two_times_mem.tex
[ADD] user tests results

@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_7_mongod.dat};
+		\addplot[color=Blue] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_68_kurento-media-server.dat};
+		\addplot[color=Red] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_68_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_35_kurento-repo.dat};
+		\addplot[color=Green] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_234_bash.dat};
+		\addplot[color=Magenta] table[x expr=\thisrowno{0}-72,y expr=\thisrowno{1}/1000000] {stats/test_two_times3/res_234_bash.dat};
 		\addlegendentry{Java}

### 2016-04-17 tese2/stats/test_without_qrcode_mem.tex
[ADD] user tests results

@@
 	    ytick={0,200,...,2000},
-	    extra x ticks={10,70,...,1000},
+	    extra x ticks={0,60,...,1000},
     	extra x tick labels={,,},
@@
 
-		\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_7_mongod.dat};
+		\addplot[color=Blue] table[x expr=\thisrowno{0}-31,y index=1] {stats/test_without_qrcode/res_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_67_kurento-media-server.dat};
+		\addplot[color=Red] table[x expr=\thisrowno{0}-31,y index=1] {stats/test_without_qrcode/res_67_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_35_kurento-repo.dat};
+		\addplot[color=Green] table[x expr=\thisrowno{0}-31,y index=1] {stats/test_without_qrcode/res_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta] table[x expr=\thisrowno{0}-35,y index=1] {stats/test_without_qrcode/res_233_bash.dat};
+		\addplot[color=Magenta] table[x expr=\thisrowno{0}-31,y index=1] {stats/test_without_qrcode/res_233_bash.dat};
 		\addlegendentry{Java}

### 2016-04-17 tese2/stats/user_evals.tex
[ADD] user tests results

\begin{tikzpicture}[scale=1.0],
\centering
\begin{axis}[
height=6cm,
width=14cm,  
  ymax=100,
  ymin=0,
  xmin=0.5,
  xmax=10.5,
  axis y line*=left,
  axis x line*=bottom,
  xticklabels={Sound quality,Video quality,Layout organization,Content editor,Collaborative editor,Time navigation, Annotation handling, Content search,Member management,Chat},
  xtick={1,...,10},
  ytick={0,20,...,200},
      ymajorgrids=true,
      xmajorgrids=true,
      grid style=dashed,
  x tick label style={rotate=45,anchor=east}]
\addplot+[only marks][error bars/.cd,y dir=both, y explicit]
coordinates {
(1,69) +- (8,8)
(2,50) +- (7,7)
(3,58) +- (11,11)
(4,54) +- (12,12)
(5,64) +- (12,12)
(6,75) +- (8,8)
(7,67) +- (8,8)
(8,66) +- (10,10)
(9,80) +- (8,8)
(10,78) +-(8,8)
};
\addplot[dashed] coordinates {(0,0) (5.5,0)};
\end{axis}
\end{tikzpicture}%

### 2016-04-17 tese2/stats/user_times.tex
[ADD] user tests results

\begin{tikzpicture}[scale=1.0],
\centering
\begin{axis}[
height=6cm,
width=7cm,  
  ymax=150,
  ymin=0,
  xmin=0.5,
  xmax=5.5,
  axis y line*=left,
  axis x line*=bottom,
  xticklabels={Task 1,Task 2,Task 3,Task 4,Task 5},
  xtick={1,...,5},
  ytick={0,20,...,200},
        ymajorgrids=true,
      xmajorgrids=true,
      grid style=dashed,
  x tick label style={rotate=45,anchor=east}]
\addplot+[only marks][error bars/.cd,y dir=both, y explicit]
coordinates {
(1,57.7) +- (12.827,12.827)
(2,120.44) +- (16.041,16.041)
(3,80.409) +- (15.067,15.067)
(4,99.91) +- (11.948,11.948)
(5,117.92) +- (16.31,16.31)
};
\addplot[dashed] coordinates {(0,0) (5.5,0)};
\end{axis}
\end{tikzpicture}%

### 2016-04-17 tese2/evaluation.tex
.

@@
 Accordingly, we have used the following formula to calculate the intervals for the expected value of the true average with 95\% confidence:
+\begin{equation}
+[\bar{x}-F^{-1}_{t_{(n-1)}}(1-\alpha/2)\times \frac{s}{\sqrt{n}}, \bar{x}+F^{-1}_{t_{(n-1)}}(1-\alpha/2)\times \frac{s}{\sqrt{n}}]
+\end{equation}
 
-$[\bar{x}-F^{-1}_{t_{(n-1)}}(1-\alpha/2)\times \frac{s}{\sqrt{n}}, \bar{x}+F^{-1}_{t_{(n-1)}}(1-\alpha/2)\times \frac{s}{\sqrt{n}}]$
-
-Where $\alpha=0.05$, $n$ is the number of samples (which may not coincide with the number of tests due to failed tasks) and $s$ is the sample standard deviation that is given by:
-
-$s=\sqrt{\frac{1}{n-1}\Sigma^{n}_{i=1}(x_i-\bar{x})^2}$
+Where $\alpha=0.05$, $n$ is the number of samples (which may not coincide with the number of tests due to failed tasks), $\bar{x}$ is the sample mean, $x_i$ is the sample value and $s$ is the sample standard deviation that is given by:
+\begin{equation}
+s=\sqrt{\frac{1}{n-1}\Sigma^{n}_{i=1}(x_i-\bar{x})^2}
+\end{equation}
 
@@
 
+[COMPARE WITH MINIMAL RESULTS (best clicks, best time, no errors)]
+
 Although we had relatively bad results with some users, we have explained to the users that could not conclude the tasks or performed incorrectly the most efficient way to perform the requested tasks.
@@
 Although, we recognize that for some cases (relatively low amount of content), displaying the content on the time-line could not have a great impact on our solutions performance.
+
+In conclusion, 100\% of our testers though that our solution was an innovation and 92.31\% recommended using our solution.

### 2016-04-23 tese2/evaluation.tex
[ADD] optimal ratio confidence intervals
[RFT] using gnuplot

@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_full_features_net.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_full_features_net.eps}
   \caption{Network usage after implementing all features}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/summary_full_net.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/summary_full_net.eps}
   \caption{Average network usage per interval of events after implementing all features}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_client_net.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_client_net.eps}
   \caption{Client Network usage during our test case}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_full_features_mem.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_full_features_mem.eps}
   \caption{Memory usage after implementing all features}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_ram_fixed_mem.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_ram_fixed_mem.eps}
   \caption{Memory usage after fixing recorder memory leak}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_full_features_cpu.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_full_features_cpu.eps}
   \caption{Percentage of CPU used during the performance tests}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_full_features_cpu_zoom.tex}
-  \end{center}
-  \caption{Percentage of CPU used during the performance tests}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_full_features_cpu_zoom.eps}
+  \caption{Percentage of CPU used during the performance tests (zoomed)}
   \label{fig:test_full_features_cpu_zoom}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/summary_full_cpu.tex}
-  \end{center}
-  \caption{Average percentage of CPU used per interval of events after implementing all features (zoomed)}
+  \centering
+  \includegraphics[width=\textwidth]{stats/summary_full_cpu.eps}
+  \caption{Average percentage of CPU used per interval of events after implementing all features}
   \label{fig:summary_full_cpu}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_without_qrcode_cpu.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_without_qrcode_cpu.eps}
   \caption{Percentage of CPU used during the performance tests without QR code detection}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_two_times_cpu.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_two_times_cpu.eps}
   \caption{Percentage of CPU used during two consecutive test cases}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_two_times_mem.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_two_times_mem.eps}
   \caption{Memory usage during two consecutive test cases}
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/test_hour_mem.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/test_hour_mem.eps}
   \caption{Memory usage during one hour with seven clients}
@@
 
+In order to measure the users learning speed, we have performed tests with experienced users in order to retrieve the optimal task duration and minimal task clicks.
+
+To this end, with regard to optimal task duration and minimal clicks, we obtained the values shown in table \ref{table:optimal}.
+
+\begin{table}[H]
+\centering
+\caption{Metrics for an experienced user}
+\label{table:optimal}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+\textbf{Task} & 1 & 2 & 3 & 4 & 5 \\ \hline
+\textbf{Time} & 20 & 35 & 30 & 25 & 35 \\ \hline
+\textbf{Clicks} & 7 & 10 & 5 & 9 & 7 \\ \hline
+\end{tabular}
+\end{table}
+
 [COMPARE WITH MINIMAL RESULTS (best clicks, best time, no errors)]

### 2016-04-23 tese2/stats/user_evals.tex
[RFT] using gnuplot

@@
 width=14cm,  
+  xlabel={Questions},
+  ylabel={User evaluation},
   ymax=100,

### 2016-04-23 tese2/stats/user_times.tex
[RFT] using gnuplot

@@
 width=7cm,  
+  xlabel={Tasks},
+  ylabel={Duration (seconds)},
   ymax=150,
@@
 \addplot[dashed] coordinates {(0,0) (5.5,0)};
+\addplot[only marks,mark=*,mark options={}] coordinates {
+(1,20) (2,35) (3,30) (4,25) (5,35)
+};
+
 \end{axis}

### 2016-04-23 project/httpserver/app/models/Message.java

@@ -11,6 +11,7 @@ import services.Service;
 import java.util.ArrayList;
 import java.util.Date;
 import java.util.List;
+import java.util.UUID;
 
 
 /**

### 2016-04-23 tese2/stats/replot.sh

@@ -1,0 +1,5 @@
+for f in *.plt
+do
+	echo "gnuplot < $f" 
+	gnuplot < $f > ${f%.plt}.eps
+done

### 2016-04-23 tese2/stats/old/test_full_features_mem.tex
[ADD] minor changes

--- 2016-04-17 tese2/stats/test_ram_fixed_mem.tex
@@
 
-		\addplot[color=Blue] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_7_mongod.dat};
+		\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}/1000000] {stats/test_full_features/res_7_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[color=Red] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_67_kurento-media-server.dat};
+		\addplot[color=Red] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}/1000000] {stats/test_full_features/res_67_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[color=Green] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_35_kurento-repo.dat};
+		\addplot[color=Green] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}/1000000] {stats/test_full_features/res_35_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[color=Magenta] table[x expr=\thisrowno{0}-68,y expr=\thisrowno{1}/1000000] {stats/test_ram_fixed/res_233_bash.dat};
+		\addplot[color=Magenta] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}/1000000] {stats/test_full_features/res_233_bash.dat};
 		\addlegendentry{Java}

### 2016-04-23 tese2/stats/old/test_full_features_net.tex
[ADD] minor changes

--- 2016-04-08 tese2/stats/test_full_features_net.tex
@@
 
-	\addplot[smooth,color=Blue] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/sent_eth0.dat};
+	\addplot[color=Blue] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/sent_eth0.dat};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=Red] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/recv_eth0.dat};
+	\addplot[color=Red] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/recv_eth0.dat};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[smooth,color=Green] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/recv_lo.dat};
+	\addplot[color=Green] table[x expr=\thisrowno{0}-122,y expr=\thisrowno{1}*8/1000000] {stats/test_full_features/recv_lo.dat};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-04-23 tese2/stats/old/test_hour_cpu.tex
[ADD] minor changes

--- 2016-04-08 tese2/stats/test_hour_cpu.tex
@@
 
-		\addplot[smooth,color=Blue,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_8_mongod.dat};
+		\addplot[color=Blue,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_8_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_91_kurento-media-server.dat};
+		\addplot[color=Red,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_91_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_36_kurento-repo.dat};
+		\addplot[color=Green,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_36_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_228_bash.dat};
+		\addplot[color=Magenta,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}] {stats/test_hour/cpu_228_bash.dat};
 		\addlegendentry{Java}

### 2016-04-23 tese2/stats/old/test_hour_mem.tex
[ADD] minor changes

--- 2016-04-08 tese2/stats/test_hour_mem.tex
@@
 
-		\addplot[smooth,color=Blue,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_8_mongod.dat};
+		\addplot[color=Blue,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_8_mongod.dat};
 		\addlegendentry{Mongod}
 
-		\addplot[smooth,color=Red,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_91_kurento-media-server.dat};
+		\addplot[color=Red,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_91_kurento-media-server.dat};
 		\addlegendentry{Kurento Media Server}
 
-		\addplot[smooth,color=Green,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_36_kurento-repo.dat};
+		\addplot[color=Green,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_36_kurento-repo.dat};
 		\addlegendentry{Kurento Repository}
 
-		\addplot[smooth,color=Magenta,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_228_bash.dat};
+		\addplot[color=Magenta,each nth point={5}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}/1000000] {stats/test_hour/res_228_bash.dat};
 		\addlegendentry{Java}

### 2016-04-23 tese2/stats/old/test_hour_net.tex
[ADD] minor changes

--- 2016-04-07 tese2/stats/test_hour_net.tex
@@
 
-	\addplot[smooth,color=Blue,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/sent_eth0.dat};
+	\addplot[color=Blue,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/sent_eth0.dat};
 	\addlegendentry{Sent(eth0)}
 
-	\addplot[smooth,color=Red,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/recv_eth0.dat};
+	\addplot[color=Red,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/recv_eth0.dat};
 	\addlegendentry{Recv(eth0)}
 
-	\addplot[smooth,color=Green,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/recv_lo.dat};
+	\addplot[color=Green,each nth point={10}] table[x expr=\thisrowno{0}-50,y expr=\thisrowno{1}*8/1000000] {stats/test_hour/recv_lo.dat};
 	\addlegendentry{Sent/Recv(lo)}

### 2016-04-23 tese2/stats/old/test_without_qrcode_cpu.tex
[ADD] minor changes

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=8cm,
	    title={CPU usage},
	    xlabel={Time [seconds]},
	    ylabel={CPU [percentage]},
	    xmin=-10, xmax=850,
	    ymin=0, ymax=600,
	    xtick={0,100,...,2000},
	    ytick={0,100,...,2000},
	    extra x ticks={0,60,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

		\addplot[color=Blue] table[x expr=\thisrowno{0}-31,y expr=\thisrowno{1}] {stats/test_without_qrcode/cpu_7_mongod.dat};
		\addlegendentry{Mongod}

		\addplot[color=Red] table[x expr=\thisrowno{0}-31,y expr=\thisrowno{1}] {stats/test_without_qrcode/cpu_67_kurento-media-server.dat};
		\addlegendentry{Kurento Media Server}

		\addplot[color=Green] table[x expr=\thisrowno{0}-31,y expr=\thisrowno{1}] {stats/test_without_qrcode/cpu_35_kurento-repo.dat};
		\addlegendentry{Kurento Repository}

		\addplot[color=Magenta] table[x expr=\thisrowno{0}-31,y expr=\thisrowno{1}] {stats/test_without_qrcode/cpu_233_bash.dat};
		\addlegendentry{Java}

	\end{axis}
\end{tikzpicture}

### 2016-04-23 tese2/stats/old/test_without_qrcode_net.tex
[ADD] minor changes

\begin{tikzpicture}
	\begin{axis}[
	    width=\textwidth,
	   	height=8cm,
	    title={Network usage},
	    xlabel={Time [seconds]},
	    ylabel={Transfer rate [mbps]},
	    xmin=-10, xmax=850,
	    ymin=0, ymax=24,
	    xtick={0,100,...,2000},
	    ytick={0,5,...,100},
	    extra x ticks={0,60,...,1000},
    	extra x tick labels={,,},
		extra tick style={
			grid=major,
			major grid style={yellow,thick,solid},
			tick label style={rotate=90,anchor=east}
		},
     legend cell align=left,
     legend style={legend pos=north west,font=\tiny},
	    ymajorgrids=true,
	    grid style=dashed,
	]

	\addplot[color=Blue] table[x expr=\thisrowno{0}-31,y expr=\thisrowno{1}*8/1000000] {stats/test_without_qrcode/sent_eth0.dat};
	\addlegendentry{Sent(eth0)}

	\addplot[color=Red] table[x expr=\thisrowno{0}-31,y expr=\thisrowno{1}*8/1000000] {stats/test_without_qrcode/recv_eth0.dat};
	\addlegendentry{Recv(eth0)}

	\addplot[color=Green] table[x expr=\thisrowno{0}-31,y expr=\thisrowno{1}*8/1000000] {stats/test_without_qrcode/recv_lo.dat};
	\addlegendentry{Sent/Recv(lo)}

	\end{axis}
\end{tikzpicture}

### 2016-04-26 tese2/16_implementation_content.tex
[ADD] starting conclusions

@@
 \subsection{Chat}
-		Lastly and not less important the chat functionality was implemented using \emph{webSockets}, when a message is sent by a user, it is received by the server, which records it and sends to all the users.
+		Not less important the chat functionality was implemented using \emph{webSockets}, when a message is sent by a user, it is received by the server, which records it and sends to all the users.
 

### 2016-04-26 tese2/conclusions.tex
[ADD] starting conclusions

\chapter{Conclusions}
\label{chapter:conclusion}

\section{Summary}
\label{section:summary}
Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as
contribuições definidas e a forma como estas se materializaram

In this thesis, our goal was the development of a web application using \emph{WebRTC} that complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time. Another important goal of this project is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.

Having this in mind, we started to analyze the technologies necessary to implement our solution. In a first glance, we have analyzed the problems that real time communication applications are facing when using the current Internet infrastructure. Next we have we have analyzed the main components of the \emph{WebRTC} stack. Furthermore, we have analyzed how we could implement the signalization component that \emph{WebRTC} does not not define. Hence, with the transport component being analyzed, we have studied how we could expose content in a synchronized and interactive way. Moreover, we have performed an analysis on the kinds of media types the particularities of each one when manipulating time. Lastly, on the state of the art context, we have analyzed existing frameworks and libraries that could help us implementing a collaborative environment.

Based on our evaluation of the state of the art, we have defined the architecture of our proposed solution.

Although we have used our own signaling protocol on our implementation, the first steps we have made during the development of our prototype were using \ac{XMPP}, we have discovered the drawbacks of this approach in practice and we moved our learnings to the state of the art. Besides starting our prototype using this approach, we have used \emph{strophe.js} for performing communications with \ac{XMPP} servers. We have found and fixed a bug present on \emph{strophe.js} library that silenced an error raised by duplicated user registrations.

Another important contribution we have made was a bug fix of a memory leak present on \ac{KMS} which we discovered when we were doing performance tests with our solution.

\section{Achievements}
\label{section:achievements}
	We have successfully implemented the basic functionalities of our prototype and spare some time for adding more valuable features such as the ability to create content by exposing \emph{QR codes} to the camera and lastly perform changes to the user interface in order to improve the quality of user's experience.

	The performance tests that we have executed showed that our system is stable and more importantly that our web server is lightweight and most of processing power is dedicated to the streaming server.

	Our usability tests show results that are considerably worse than the established optimal values due to our solution propose a different way to communicate that most people are not used to. Although we have obtained those results, in general our users gave us positive feedback and valuable advices which we have used to improve our system.

\section{Future Work}
\label{section:future}
Nesta secção devem identificar o trabalho futuro, sob duas perspectivas:
\begin{itemize}
\item o trabalho que resulta directamente dos problemas que a vossa proposta criou, ou não conseguiu resolver
\item o trabalho que resulta da evolução do sistemas
\end{itemize}

\cleardoublepage

### 2016-04-26 tese2/conclusions.tex
[ADD] conclusions

@@
 
+As a consequence of not using \ac{XMPP} for the signaling protocol, we have defined our own protocol based on \emph{WebSockets} for performing the stream between clients and \ac{KMS}.
+From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \emph{QR code} detection for creating hyper content.
+
 Another important contribution we have made was a bug fix of a memory leak present on \ac{KMS} which we discovered when we were doing performance tests with our solution.
@@
 \label{section:future}
-Nesta secção devem identificar o trabalho futuro, sob duas perspectivas:
-\begin{itemize}
-\item o trabalho que resulta directamente dos problemas que a vossa proposta criou, ou não conseguiu resolver
-\item o trabalho que resulta da evolução do sistemas
-\end{itemize}
+	Playing back video with a faster rate is not possible using the current version of \ac{KMS} which we expect the availability of that feature in a near future but we have proposed an alternative way to implement faster playback by using \emph{ffmpeg} to convert the video before playing it.
 
+	Although we have tested our solution in a powerful machine for the current time, our performance tests revealed that the streaming component uses a lot of resources.
+We left for a future work a deep analysis on the scalability of our system which we have proposed different approaches but we have not tested them.
+
+	Another aspect we could have tested was the performance of our solution when using \ac{TURN} servers for relaying the traffic that fails using \ac{STUN}.
+
+	Lastly, we have chosen functionality over security in respect to displaying content to users which lead to security flaws on our solution.
+Although we have not solved the security problems, we have proposed a solution which at the same time limits the flexibility of adding new functionalities to our prototype.
 \cleardoublepage

### 2016-04-28 tese2/conclusions.tex
[RFT] rolled back to previous Makefile

@@
 \label{section:summary}
-Neste secção deve-se fazer o resumo do trabalho efectuado, retomando a ideia, as
-contribuições definidas e a forma como estas se materializaram
-
 In this thesis, our goal was the development of a web application using \emph{WebRTC} that complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time.

### 2016-04-28 project/httpserver/app/controllers/Rest.java

@@ -613,7 +613,6 @@ public class Rest extends Controller {
         User ret = new RegisterUserService(password, attributes).execute();
         session("uid",ret.getId().toString());
         return ok(ret.getToken());
-
     }
 
 

### 2016-04-28 project/httpserver/app/controllers/WSController.java

@@ -127,211 +127,242 @@ public class WSController extends Controller {
 
                     // For each event received on the socket,
                     in.onMessage(new Callback<String>() {
-                        public void invoke(String event) {
-
-                            JSONObject args = new JSONObject(event);
-                            String cmd = args.getString("cmd");
-
-
-                            if(!"talk".equals(cmd))
-                                System.out.println("\nRECV: " + event);
-                            switch (cmd) {
-                                case "createMessage": {
-                                    String data = args.getString("data");
-                                    CreateMessageService messageService = new CreateMessageService(
-                                            userId, groupId, data);
-                                    try {
-                                        Message message = messageService.execute();
-                                        JSONArray messagesArray = new JSONArray();
-                                        JSONObject messageObj = message.toJsonObject();
-                                        messagesArray.put(messageObj);
-                                        JSONObject msg = new JSONObject();
-                                        msg.put("cmd", "msg");
-                                        msg.put("data", messagesArray);
-                                        room.sendMessage(msg.toString());
-
-                                    } catch (ServiceException e) {
-                                        e.printStackTrace();
-                                    }
-                                    // send msg
-                                }
-                                break;
-                                case "getMessages": {
-                                    int len = args.getInt("len");
-                                    Long ts = args.getLong("ts");
-                                    userSession.sendMessages(ts, len);
-                                }
-                                break;
-                                case "offer": {
-                                    JSONObject data = args.getJSONObject("data");
-                                    String rsd = data.getString("sdp");
-                                    userSession.processOffer(rsd);
-                                }
-                                break;
-                                case "iceCandidate": {
-                                    JSONObject jCand = args.getJSONObject("candidate");
-                                    IceCandidate candidate = new IceCandidate(jCand.getString("candidate"),
-                                            jCand.getString("sdpMid"), jCand.getInt("sdpMLineIndex"));
-                                    userSession.addCandidate(candidate);
-                                }
-                                break;
-                                case "setRealTime": {
-                                    new Thread(new Runnable() {
-                                        @Override
-                                        public void run() {
-                                            String owner = args.optString("uid",null);
-                                            String sessionId = args.optString("sid", null);
-                                            userSession.setRealTime(owner,sessionId);
-                                            userSession.sendMessage(userSession.getContent());
-                                        }
-                                    }).start();
-                                }
-                                break;
-                                case "removeUser": {
-                                    String uid = args.optString("uid", null);
[…]

### 2016-04-28 project/httpserver/app/main/Room.java

@@ -334,7 +334,7 @@ public class Room implements Closeable {
         return alternativeSession;
     }
 
-    public UserSession getUser(final UUID sid) {
+    public UserSession getUser(final String sid) {
         if(sid!=null) {
             synchronized (participants) {
                 for (UserSession session : participants) {

### 2016-04-28 project/httpserver/app/main/UserSession.java

@@ -60,6 +60,8 @@ public class UserSession implements Closeable, Comparable<UserSession> {
         this.room = room;
         playSid = "group";
         initEndpoint();
+
+        System.out.println("new UserSession ---> " + sid.toString());
     }
 
     private void initEndpoint(){

### 2016-04-28 project/httpserver/public/this/js/hyper.socket.js

@@ -26,13 +26,11 @@ var HyperWebSocket = new (function() {
     var local_none = {offerToReceiveAudio:true,offerToReceiveVideo:true};
     var screen_user = {video: {mediaSource: 'window' || 'screen'}, audio: false};
     var camera_user = {audio:true, video:true };
-
     var audio_constraints = { 'offerToReceiveAudio':true,'offerToReceiveVideo':true};
     var remote_constraints = { 'offerToReceiveAudio':true,'offerToReceiveVideo':true};
     // ===============================================
     // ================== UTILITIES ==================
     // ===============================================
-
 
     function logError(err) {
         console.log(err);
@@ -82,7 +80,7 @@ var HyperWebSocket = new (function() {
 						soundDetected = true;
 						this.talk(true);
 					}
-				}else {
+				} else {
 					if(soundDetected){
 						soundDetected = false;
 						this.talk(false);
@@ -189,7 +187,6 @@ var HyperWebSocket = new (function() {
 			uid:userId
 		}));
 	}
-	
 
     this.removeUserGroup = function(userId){
         ws.send(JSON.stringify({
@@ -197,8 +194,6 @@ var HyperWebSocket = new (function() {
             uid:userId
         }));
     }
-
-
 
 	this.receiveRealTime = function(userId,sessionId){
 		console.log("receiveRealTime",userId);
@@ -211,7 +206,6 @@ var HyperWebSocket = new (function() {
 		}
 		ws.send(JSON.stringify(msg));
 	}
-
 
 	this.receiveHistoric = function(userId,offset,sessionId){
 		console.log("receiveHistoric",userId);
@@ -227,6 +221,7 @@ var HyperWebSocket = new (function() {
 	}
 	
 	this.sendMessage = function(text){
+	    console.log("sendMessage",text);
 		ws.send(JSON.stringify({
 			cmd: "createMessage",
 			data: text
@@ -252,7 +247,6 @@ var HyperWebSocket = new (function() {
     this.setServerTimeSynchronizationHandler = function(handler){
         serverTimeHandler = handler;
     }
-
 
     this.setChannelsArrivedHandler = function (handler) {
         channelsArrivedHandler = handler;
@@ -539,7 +533,7 @@ var HyperWebSocket = new (function() {
 		
 			window.onbeforeunload = function() {
 				ws.close();
-			};
+			}.bind(this);
 		
 		} else {
 			console.log("no websocket support!");

### 2016-04-29 tese2/12_related_work_summary.tex
[ADD] chaper 2 summary

\section{Chapter Summary}
\label{related:summary}

In order to implement an hyper-linked communication solution, several design decisions are to be made. The \emph{Internet} structure, their protocols and browser capabilities are a key factors when implementing a solution that allows bi-directional communications, interactive media and collaboration environments.

Bi-directional communications between clients have different needs from request-response based communications between clients and servers due to the introduction of \ac{NAT}, this leads to the appearance of mechanisms such as \ac{STUN}, \ac{TURN} and \ac{ICE} in order to bypass the limits imposed by \ac{NAT}.

With those problems solved using techniques such as described, \ac{WebRTC} came to introduce an \ac{API} for video, audio and data communications through web browsers without using plug-ins.

By itself \ac{WebRTC} does not define how each user gets to know each other and how the information flows between users. For this reason we have studied the multiple ways we could implement this \emph{get-to-know} mechanism which is known by signaling protocol.

Therefore, with the communications established, we had to discuss the different types of media and what can be done with each kind in order to increase the value of communications among users.

### 2016-04-29 tese2/resumo.tex
[ADD] resumo (PT)

\chapter*{Resumo}

% DO NOT CHANGE THIS - Add entry in the table of contents as chapter
\addcontentsline{toc}{chapter}{Resumo}

O conceito de comunicações hiper-ligadas aplica muitos dos conceitos de hiper-media, bastante utilizados no conteúdo \emph{Web}, para serviços se comunicação e colaboração. Este paradigma permite sincronizar, estruturar e navegar sobre conteúdo proveniente de communicações, colocando ao dispor formas de integrar conteúdo de mídia social e ferramentas colaborativas em chamadas de voz e vídeo.
A voz e a imagem justas exprimem emoções e expoem a creatividade como nenhum outro tipo de media permite. Através dos conceitos de hiper-media, podemos adicionar mais valor às comunicações por voz e vídeo.

A tecnologia \ac{WebRTC} permite realizar comunicações em tempo real entre navegadores web sem a necessidade de instalar aplicações adicionais ou \emph{plug-ins}. A natureza das aplicações \emph{web} já usufruem dos conceitos de hiper-media, o que faz do \ac{WebRTC} a tecnologia ideal para aplicar o conceito de comunicações hiper-ligadas.
A plataforma providenciada por navegadores \emph{web} disponibiliza uma camada de abstração que torna possível correr aplicações independentemente do sistema operativo.
O suporte nativo do \ac{WebRTC} nos sistemas operativos extendem a sua utilização para fora do contexto dos navegadores \emph{web}, possibilitando explorar funcionalidades que os navegadores \emph{web} suportam de uma forma limitada como a gravação de vídeo e armazenamento de informação em massa.

O nosso objectivo neste projecto é o desenvolvimento de uma aplicação que utiliza o conceito de comunicações hiper-ligadas de forma a enriquecer ambientes colaborativos e comunicações audio e vídeo com vários tipos de media, que podem ser acedidos em tempo real ou mais tarde.
Esta aplicação será desenvolvida sobre a plataforma \emph{web}, recorrendo ao \ac{WebRTC}.

Neste documento, apresentamos o actual Estado de Arte nas comunicações hiper-ligadas e as tecnologias relacionadas, propomos e implementamos uma arquitectura para uma aplicação com comunicações hiper-ligadas baseada no \emph{WebRTC} e mostramos os resultados da implementação e avaliação.

\vspace{1cm}

% TODO 4 to 6 keywords;
\textbf{\Large Palavras-chave:} WebRTC, comunicações, asíncronas, colaboração

\cleardoublepage

### 2016-04-30 tese2/18_implementation_summary.tex
[ADD] chapter summary

\section{Chapter Summary}
\label{implementation:summary}

In this chapter we have described how we implemented the various components present on our architecture, the challenges that we have faced and the solutions we have found.

We started by defining the database model, which influenced directly our system's behavior and the extensibility to new functionalities.

Then, we have underlined the requirements of our signaling protocol and as consequence we have described, in detail, the protocol itself.

With the signaling protocol implemented we observed the first outcomes of using \ac{WebRTC}, the basic functionality we have implemented was an echo of the video and audio sent by users. However we have implemented other features such as switching to other user streams, recording and mixing multiple streams into a single one.

In respect to hyper-content, we have described how to create and search for content either being superimposed content to video and time annotations. In relation to displaying content to users we have also defined how we have synchronized the content to show and the security concerns of our choices.

Then, we have described how we have implemented our chat and collaborative editor.

Lastly, although we have tested our solution with our own development machine during the implementation, we have described how we have deployed our system for allowing a public access and perform tests in a real case environment.

### 2016-04-30 tese2/acknowledgements.tex
[ADD] chapter summary

\chapter*{\acknowledgments}

% Add entry in the table of contents as chapter
\addcontentsline{toc}{chapter}{\acknowledgments}

I would like to thank my dissertation supervisor Ricardo Pereira and co-supervisor Paulo Chainho for their active support and ideas that were crucial for the development of this thesis.

I should not forget all the people involved on the prototype's test phase, they not only tested but also gave valuable feedback that we used to improve the quality of our prototype.

Also, I would like to show my gratitude to the \emph{Kurento} team for developing the streaming server that I've used, their impressive help support and receptiveness to discuss new functionalities.

Not less important I should thank all IST employees that maintained the space I used during this journey.

Lastly, I should thank all the people that supported me psychologically and made me going my own way.

Thank you!

\cleardoublepage

### 2016-04-30 tese2/architecture.tex
[ADD] optimal ratio confidence intervals
[ADD] chapter summary

@@
 \end{tabular}
-	        \label{fig:apparch}
 \end{table}
 
-Figure~\ref{fig:apparch} presents the application architecture and the underlying technologies seen from the user's perspective.
+Table~\ref{table:apparch} presents the application architecture and the underlying technologies seen from the user's perspective.
 \emph{Adapter.js} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.
@@
 Other animations can be implemented with \ac{SVG} embedded on \ac{HTML}.
+
+\section{Chapter Summary}
+\label{architecture:summary}
+
+In this chapter we have shown the modules needed in order to implement our solution.
+
+Several technologies were taken into account for implementing each module, we have studied the pros and cons of each technology and decided the architecture for our solution.

### 2016-04-30 tese2/evaluation.tex
[RFT] gnuplot confidence intervals
[ADD] chapter summary

@@
 
+In order to measure the users learning speed, we have performed tests with experienced users in order to retrieve the optimal task duration and minimal task clicks.
+
+To this end, with regard to optimal task duration and minimal clicks, we obtained the values shown in table \ref{table:optimal}.
+
+\begin{table}[H]
+\centering
+\caption{Metrics for an experienced user}
+\label{table:optimal}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+\textbf{Task} & 1 & 2 & 3 & 4 & 5 \\ \hline
+\textbf{Duration (seconds)} & 20 & 35 & 30 & 25 & 35 \\ \hline
+\textbf{Number of clicks} & 7 & 10 & 5 & 9 & 7 \\ \hline
+\end{tabular}
+\end{table}
+
 From the data that we have collected, namely the task duration (figure \ref{fig:user_times}), number of clicks (figure \ref{fig:user_clicks}), number of errors (figure \ref{fig:user_errors}), task difficulty (figure \ref{fig:user_diffs}) and qualitative evaluation (figure \ref{fig:user_evals}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.
@@
 \centering
-\begin{minipage}[b]{0.45\linewidth}
-  \begin{center}
-    \input{stats/user_times.tex}
-  \end{center}
+\begin{minipage}{.5\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{stats/user_times.eps}
     \caption{Time spent per task}
     \label{fig:user_times}
-\end{minipage}
-\begin{minipage}[b]{0.45\linewidth}
-   \begin{center}
-    \input{stats/user_clicks.tex}
-  \end{center}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{stats/user_clicks.eps}
     \caption{Mouse clicks per task}
@@
 \centering
-\begin{minipage}[b]{0.45\linewidth}
-  \begin{center}
-    \input{stats/user_errors.tex}
-  \end{center}
+\begin{minipage}{.5\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{stats/user_errors.eps}
   \caption{Errors per task}
   \label{fig:user_errors}
-\end{minipage}
-\begin{minipage}[b]{0.45\linewidth}
-  \begin{center}
-    \input{stats/user_diffs.tex}
-  \end{center}
+\end{minipage}%
+\begin{minipage}{.5\textwidth}
+  \centering
+    \includegraphics[width=\textwidth]{stats/user_diffs.eps}
   \caption{Difficulty per task}
@@
 
-In order to measure the users learning speed, we have performed tests with experienced users in order to retrieve the optimal task duration and minimal task clicks.
-
-To this end, with regard to optimal task duration and minimal clicks, we obtained the values shown in table \ref{table:optimal}.
-
-\begin{table}[H]
-\centering
-\caption{Metrics for an experienced user}
-\label{table:optimal}
-\begin{tabular}{|c|c|c|c|c|c|}
-\hline
-\textbf{Task} & 1 & 2 & 3 & 4 & 5 \\ \hline
-\textbf{Time} & 20 & 35 & 30 & 25 & 35 \\ \hline
-\textbf{Clicks} & 7 & 10 & 5 & 9 & 7 \\ \hline
-\end{tabular}
-\end{table}
-
-[COMPARE WITH MINIMAL RESULTS (best clicks, best time, no errors)]
-
 Although we had relatively bad results with some users, we have explained to the users that could not conclude the tasks or performed incorrectly the most efficient way to perform the requested tasks.
@@
 \begin{figure}[!htb]
-  \begin{center}
-    \input{stats/user_evals.tex}
-  \end{center}
+  \centering
+  \includegraphics[width=\textwidth]{stats/user_evals.eps}
   \caption{Solution evaluation}
@@
 In conclusion, 100\% of our testers though that our solution was an innovation and 92.31\% recommended using our solution.
+
+\section{Chapter Summary}
+\label{evaluation:summary}
+
+In this chapter we describe the test scenarios that we have created in order to evaluate our system.
+
+The performance tests revealed unexpected results which were crucial to find and solve performance issues.
+Another important aspect of our performance test results, were the understanding of how data flows between our system modules.
+
+In respect to user interface tests, we have obtained valuable feedback that we used to improve our prototype's usability.
+
+In order to understand how good the results were for each performed task, we have defined optimal values for the different metrics.
+In general the results reveal a significant difference between observed values and optimal values which can be explained by the users lack of experience with systems such as we have created.

### 2016-05-04 tese2/introduction.tex
[ADD] introduction improvements,[FIX] notification counter fix

@@
 	With the advent of the telephone network, communications have taken a very important step for us to feel more connected with whom we communicate.
-Still, only the human voice was not enough, and the invention of cameras and consequent video digitization were a huge step for real-time communications.
+Still, only the human voice was not enough, and the invention of cameras and consequent video digitalization were a huge step for real-time communications.
 
 	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronism between writers.
-	Today, we can achieve more, it is possible to write a document collaboratively, correct spelling mistakes without wasting paper, restructure text at any moment, add a video to a newspaper article and more.
+	Today, we can achieve more, it is possible to write a document collaboratively, correct spelling mistakes without wasting physical resources, restructure text at any moment, add a video to a newspaper article and more.
 Although much of what was said seems banal nowadays, none of this was possible before the computer's invention.
@@
 	As communications technologies appeared, we adapted the way we communicate.
-This project doesn't aims to replace the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
+{\color{blue} The purpose of this project is not the replacement of} the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
 
-	{\color{blue}With the advent of WebRTC, it became possible to develop video conference web apllications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.}
+	{\color{blue}With the advent of WebRTC and its successive integration with web browsers, it became possible to develop video conference web applications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.}
 
-    Real-time communication applications can make a difference on business, education and health sectors by providing tools for teaching and learning online, teamworking and socializing.
+    Furthermore, real-time communication applications can make a significant difference on business, education and health sectors by providing tools for developing teaching and learning online, teamworking and socializing web applications.
 
@@
                %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
-	A real-time system is a huge source of information that requires much attention from its users.
-An application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
+	Therefore, a real-time system is a huge source of information that requires much attention from its users.
+In this context, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
+
+	{\color{blue}Another point to consider is making available collaborative tools for text edition, instant messaging and video superimposed content creation and visualization.}
 
@@
 	Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
-creating annotations for improving content search) in order to increase its value.
-Another important goal of this project is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.
+creating time annotations for improving content search) in order to increase its value.
+It is also important to highlight another goal of this project which is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.
 
@@
 
-	We will present an architecture that can meet our goals, implement the respective prototype and test it with real users, unitary tests and benchmarks.
+	We will present an architecture that can meet our goals, implement the respective prototype and test it with real users and performance benchmarks.
 

### 2016-05-05 tese2/abstract.tex
reviewed abstract

@@
 \ac{WebRTC} technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
-The nature of web browser applications already follow the hypermedia concept, which makes \ac{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
-{\color{blue}The web browser platform provides an abstraction layer that makes possible to create applications that run indenpendently from the operating system.}
-The native support of \ac{WebRTC} in operating systems extends its usage to outside the web browser, allowing to explore functionalities that web browsers are poor to support such as video recording and massive information storage.
+The nature of web browser applications already follows the hypermedia concept, which makes \ac{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
+{\color{blue}The web browser platform provides an abstraction layer that makes it possible to create applications that run independently from the operating system.}
+The native support for \ac{WebRTC} in operating systems extends its usage to outside the web browser, allowing for the exploration of functionalities for which web browsers provide poor support, such as video recording and massive information storage.
 
@@
 
+%RP acho que estás a vender muito mal o teu trabalho.
+O abstract é a primeira caisa que se lê e em função disso decidimos se lemos mais ou não.
+Tens aqui a oportunidade de chamar a atenção para algumas das funções inovadoras.
+O texto está muit genérico.
+podes falar em coisas concretas: video conferência, chat, edição colaborativo de documento, partilha de ficheiros, possibilidade de criar legendas, tags, etc.
+Possibilidade de navegar, voltar atrás, etc como no IPTV.
+Gestão de amigos, gerir salas, etc, etc.
+
 {\color{blue}
@@
 }
+
+%RP em vez de ``provide the results for its implementation and evaluation'', diz algo como: ``this work was evaluated using a panel of 20 users, who reported that they liked to use it and thought it to be extremelly inovative ...
+and if given the change, would use it ...''
 

### 2016-05-05 tese2/introduction.tex
Reviewed introduction

@@
 
-	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronism between writers.
-	Today, we can achieve more, it is possible to write a document collaboratively, correct spelling mistakes without wasting physical resources, restructure text at any moment, add a video to a newspaper article and more.
+	In the past, handwritten documents were limited to a writer per page at a time. Writing a book collaboratively was a difficult task due to synchronization between writers.
+	Today, we can achieve more.
+It is possible to write a document collaboratively, correct spelling mistakes without wasting physical resources, restructure text at any moment, add a video to a newspaper article and more.
 Although much of what was said seems banal nowadays, none of this was possible before the computer's invention.
 
-	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice's capacity to conveying rich and complex social and emotional meaning} , although nothing replaces the physical contact with a person while we communicate, we are at a time when we can do more than just a visual and verbal communication, hypermedia can be added to video and voice in order to extend its value.
+	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice's capacity to conveying rich and complex social and emotional meaning}.
+Although nothing replaces the physical contact with a person while we communicate, we are at a time when we can do more than just visual and verbal communication.
+Hypermedia can be added to video and voice in order to extend its value.
 The concept of structured voice and video synchronized with hypermedia is called hypervoice\cite{geddes}.
@@
 	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see.
-               %RP acho que toda a gente se esquece eventualmente. Pode não lhe parecer importante na altura. Não é preciso estar doente!
+        %RP Parece-me que até aqui ainda é background ou motivação. Não me parece que estejas já a descrever a solução proposta.
+
 	Therefore, a real-time system is a huge source of information that requires much attention from its users.
@@
 
+%RP Estás a misturar goals com contribuições. Nesta altura deves falar do que foi feito (na solução). Os objectivos podem ficar no inicio da arquitectura para a motivar e condicionar
+
 A web application with an easy to learn user interface will be developed to accomplish solving our problem.
@@
 
+%RP ``will be developed''!!! ``was''. Não podes estar a falar no futuro. Isto é passado. Estás a descrever o que fizeste.
+
 	Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solve our problem either completely or just a part of it, not excluding solutions that may appear during the implementation of our project.
@@
 	We will present an architecture that can meet our goals, implement the respective prototype and test it with real users and performance benchmarks.
+
+        %RP mais uma vez, passado e muito vago.
 
@@
 
+        %RP este capítulo precisa de uma revisão. A estrutura tem de ser:
+        %Motivação/background - descrever os problemas / desafios / oportunidades que motivam a tese
+        %Proposed solution - descrever por alto a solução para o problema, uma aplicação web com determinadas funcionalides (listando todas elas)
+        %contribution - as contribuições são: uma arquitectura, um interface utilizador para lidar com um conceito novo (andar no tempo), um protótipo funcional, teste com utilizadores e respectivas conclusões (avaliação experimental), documentação (tese).
+
 \section{Outline}
 
-This document is structured as follows:
+This rest of this document is structured as follows:
 
 \begin{itemize}
-\item \textbf{Chapter \ref{chapter:introduction}} presents the motivation, background and proposed solution.
+%\item \textbf{Chapter \ref{chapter:introduction}} presents the motivation, background and proposed solution.
 \item \textbf{Chapter \ref{chapter:relatedwork}} describes the previous work in the field.
@@
 \item \textbf{Chapter \ref{chapter:implementation}} describes the implementation of our Web Application and the technologies chosen.
-\item \textbf{Chapter \ref{chapter:evaluation}} describes the evaluation tests performed and the corresponding results.
-\item \textbf{Chapter \ref{chapter:conclusion}} summarizes the work developed and future work.
+\item \textbf{Chapter \ref{chapter:evaluation}} presents the evaluation tests performed and the corresponding results.
+\item \textbf{Chapter \ref{chapter:conclusion}} summarizes the work developed and proposes future work.
 \end{itemize}

### 2016-05-05 tese2/abstract.tex
[RFT] long abstract

@@
 
-Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed later.
-This application will target the web browser platform, resorting to \ac{WebRTC}.
+\sout{Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed later.}
+%This application will target the web browser platform, resorting to \ac{WebRTC}.
+
+{\color{red}Our goal in this project is to develop an application targeted to the web platform, resorting to \ac{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \emph{HTML5}, \emph{CSS3} and \emph{JavaScript} including continuous media types such as video, music and animations.}
+
+{\color{red}One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again or introduce hyper-content to it such as time annotations, interactive lists of topics and subtitles.
+In this context we also provide a simpler method for creating and synchronizing hyper-content using \emph{QR codes}.}
+
+{\color{red}In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links and files to conference participants.}
+
+{\color{red}Furthermore, another relevant feature is the possibility to compose multiple video streams shared by conference participants into a single one, which enables adding more users to conference rooms without impact on client's performance.
+Users can change to individual streams on demand or automatically to the talking users.}
 

### 2016-05-05 tese2/abstract.tex
[RFT] long abstract

@@
 
-\sout{Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed later.}
+%\sout{Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed later.}
 %This application will target the web browser platform, resorting to \ac{WebRTC}.
@@
 
-{\color{red}Furthermore, another relevant feature is the possibility to compose multiple video streams shared by conference participants into a single one, which enables adding more users to conference rooms without impact on client's performance.
+{\color{red}Furthermore, another relevant feature is the possibility to compose multiple video streams into a single one, which enables adding more users to conference rooms without impacting on clients performance.
 Users can change to individual streams on demand or automatically to the talking users.}

### 2016-05-06 tese2/07_related_work_nat.tex
Reviewed recent changes to state of the art

@@
 \ac{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}.
-{\color{blue} Today we live in the \emph{Internet of Things} era, users and companies may have multiple devices that require internet connectivity, from personal computers, smart phones, tablets, traffic sensors, public vigillance systems, health monitoring tools and so on, all those devices combined cannot be publicly addressed by \ac{IPv4}.}
+{\color{blue} Today we live in the \emph{Internet of Things} era.
+  Users and companies may have multiple devices that require Internet connectivity, such as personal computers, smart phones, tablets, traffic sensors, public vigillance systems, health monitoring tools and so on.
+All those devices combined cannot be publicly addressed by \ac{IPv4}.}
+%RP vigillance -> surveilance?
+%RP usa acrónimo IoT para Internet of Things
 
@@
 
-Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security, although its current wide usage, \ac{NAT}s are exposing their weaknesses to the application layer, {\color{blue}namely on applications that require direct communications between two private networks.}
-%RP initially? E hoje?
+Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security.
+However, due to their current wide usage, \ac{NAT}s weaknesses are being exposing at the application layer, {\color{blue}namely impacting applications that require direct communications between two private networks.}
+
 There are four types of \ac{NAT} implementations\cite{rfc3489}: \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.

### 2016-05-06 tese2/09_related_work_signaling.tex
Reviewed recent changes to state of the art

@@
   \ac{XMPP} messages are \ac{XML} based, which is attractive for applications that need structured messages and rich hypermedia.
-Another advantage of \ac{XMPP} is the addition of extensions, for example \emph{XEP-0096}\cite{xep0096}, which adds file transfer capabilities between two entities and \emph{XEP-0045}\cite{xep0045} which enables multi-user chat.
+Another advantage of \ac{XMPP} is the addition of extensions, for example \emph{XEP-0096}\cite{xep0096}, which adds file transfer capabilities between two entities, and \emph{XEP-0045}\cite{xep0045}, which enables multi-user chat.
   \ac{XMPP}'s bi-directional communication over \ac{HTTP} is achieved through \ac{BOSH} \cite{xep0206}.
@@
     One of the advantages that \ac{XMPP} based applications offer is the possibility to send messages across different servers.
-An \ac{XMPP} user is identified by its \textit{Jabber ID} which is definined by the pair username and server name in the form \textit{username@server}.
+An \ac{XMPP} user is identified by its \textit{Jabber ID}, which is defined by the pair username and server name in the form \textit{username@server}.
 
-    The user profile is represented in terms of Data Forms\cite{xep0004} protocol and managed (registration and update) through \textit{Info/Query} (IQ) \cite{rfc3920} requests to the \ac{XMPP} server which responds with an IQ response containing information to retrieve the registration fields needed to fill its profile (eg username, password, telephone).
+    The user profile is represented in terms of Data Forms\cite{xep0004} protocol and managed (registration and update) through \textit{Info/Query} (IQ) \cite{rfc3920} requests to the \ac{XMPP} server, which responds with an IQ response containing information to retrieve the registration fields needed to fill its profile (eg username, password, telephone).
+%RP frase demasiado longa. Torna-se confusa.
 
-    An important aspect of this model is the ability to define a user as a set of attributes that can change dinamically in order to fit the \ac{IM} applications needs.
+    An important aspect of this model is the ability to define a user as a set of attributes that can change dinamically in order to fit the \ac{IM} application's needs.
 
@@
     Typically a web application consists in multiple web pages which are navigated by users.
-Each time a webpage is accessed, either from a new context or a transition from a previous page, its context is cleared except for local storage and cookies, the javascript context is cleared, including the \ac{XMPP} connections performed by \emph{strophe.js}, an automatic mechanism is required for avoiding the user implicit reauthentication.
+Each time a webpage is accessed, either from a new context or a transition from a previous page, its context is cleared except for local storage and cookies.
+The javascript context is cleared, including the \ac{XMPP} connections performed by \emph{strophe.js}.
+As such, an automatic mechanism is required for avoiding the implicit user reauthentication.
+    One solution for reauthentication can be achieved by storing the user's JabberID and password on local storage and everytime a page is accessed the authentication is performed without the users knowledge.
+Clearly this solution represents security flaws as the local storage can be easily accessed locally or by performing a cross site scripting attack to reveal all the local storage.
+The same problem arises if the credentials are stored on cookies.
+    %RP tinhas este comentário (HOW? explain!)
+    %RP o strophe.js caiu aqui da paraquedas! Explicar o que é
 
-    One solution for reauthentication can be achieved by storing the user's JabberID and password on local storage and everytime a page is accessed the authentication is performed without the users knlowledge.
-Clearly this solution represents security flaws, the local storage can be easily accessed (HOW?
-explain!) locally or by performing a cross site scripting attack and reveal all the local storage, the same problem arises if the credentials are stored on cookies.
-
-    An alternative solution is known as Session Attachment \footnote{\url{https://metajack.im/2008/10/03/getting-attached-to-strophe/}}, which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} is passed to strophe.js in order to re-connect to the same stream on \ac{XMPP} server.
-Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request which is worthless if a user mantains more than one tab opened, for example for multiple conversations at the same time.
+    An alternative solution is known as Session Attachment \footnote{\url{https://metajack.im/2008/10/03/getting-attached-to-strophe/}}, which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} be passed to strophe.js in order to re-connect to the same stream on \ac{XMPP} server.
+Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request making it worthless if a user mantains more than one tab opened, for example for multiple conversations at the same time.
+    %RP either ou both?
+    %RP verifica se não alterei o sentido da frase
 
@@
 
-    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that \ac{XMPP} already implemment, but from all the \ac{XMPP} features, we don't need all of them, in fact we just need a subset of \ac{XMPP} features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.
+    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that \ac{XMPP} already implemment, but  we don't need all of them.
+In fact, we just need a subset of \ac{XMPP} features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.
 }

### 2016-05-06 tese2/12_related_work_summary.tex
Reviewed recent changes to state of the art

@@
 
-In order to implement an hyper-linked communication solution, several design decisions are to be made.
-The \emph{Internet} structure, their protocols and browser capabilities are a key factors when implementing a solution that allows bi-directional communications, interactive media and collaboration environments.
+In order to implement an hyper-linked communication solution, several design decisions had to be made.
+The limitations imposed by the \emph{Internet's} structure, its protocols and a browser's capabilities are key factors to consider when implementing a solution that allows bi-directional communications, interactive media and collaboration environment.
 
-Bi-directional communications between clients have different needs from request-response based communications between clients and servers due to the introduction of \ac{NAT}, this leads to the appearance of mechanisms such as \ac{STUN}, \ac{TURN} and \ac{ICE} in order to bypass the limits imposed by \ac{NAT}.
+Due to the use of \ac{NAT, bi-directional communications between clients have different needs from request-response based communications between clients and servers.
+This lead to the appearance of mechanisms such as \ac{STUN}, \ac{TURN} and \ac{ICE}, in order to bypass the limits imposed by \ac{NAT}.
 
-With those problems solved using techniques such as described, \ac{WebRTC} came to introduce an \ac{API} for video, audio and data communications through web browsers without using plug-ins.
+\ac{WebRTC} introduced an \ac{API} for video, audio and data communications through web browsers without using plug-ins.
+However, \ac{WebRTC} by itself does not define how users get to know each other nor how information flows between users.
+For this reason, we have studied the multiple ways we could implement this \emph{get-to-know} mechanism which is known as signaling protocol.
 
-By itself \ac{WebRTC} does not define how each user gets to know each other and how the information flows between users.
-For this reason we have studied the multiple ways we could implement this \emph{get-to-know} mechanism which is known by signaling protocol.
-
-Therefore, with the communications established, we had to discuss the different types of media and what can be done with each kind in order to increase the value of communications among users.
+With the communications establishment issue solved, we had to discuss the different types of media and what can be done with each kind in order to increase the value of communications among users.
+%RP falas de mais coisas: manipulação de tempo, gravação, edição colaborativa

### 2016-05-06 tese2/abstract.tex

@@
 
-\ac{WebRTC} technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
-The nature of web browser applications already follows the hypermedia concept, which makes \ac{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
+\emph{WebRTC} technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
+The nature of web browser applications already follows the hypermedia concept, which makes \emph{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
 {\color{blue}The web browser platform provides an abstraction layer that makes it possible to create applications that run independently from the operating system.}
-The native support for \ac{WebRTC} in operating systems extends its usage to outside the web browser, allowing for the exploration of functionalities for which web browsers provide poor support, such as video recording and massive information storage.
+The native support for \emph{WebRTC} in operating systems extends its usage to outside the web browser, allowing for the exploration of functionalities for which web browsers provide poor support, such as video recording and massive information storage.
 
@@
 
-{\color{red}Our goal in this project is to develop an application targeted to the web platform, resorting to \ac{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \emph{HTML5}, \emph{CSS3} and \emph{JavaScript} including continuous media types such as video, music and animations.}
+{\color{red}Our goal in this project is to develop an application targeted to the web platform, resorting to \emph{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types.}
 
@@
 {\color{blue}
-In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose and implement an architecture for an hyper-linked communication application based on \ac{WebRTC} and provide the results for its implementation and evaluation.
+In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose and implement an architecture for an hyper-linked communication application based on \emph{WebRTC} and provide the results for its implementation and evaluation.
 }

### 2016-05-06 tese2/07_related_work_nat.tex
[RFT] abstract and introduction rewrite

--- 2016-03-25 tese2/07_related_work_nat.tex
@@
 
-One way to overcome the \ac{IPv4} address scarcity problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \ac{IP} address with a different port.
+To this end, one way to overcome the \ac{IPv4} address scarcity problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \ac{IP} address with a different port.
 %RP o porto não identifica uma máquina (membro da rede privada) mas uma ligação.
@@
 %RP initially? E hoje?
-There are four types of \ac{NAT} implementations\cite{rfc3489}: \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.
+It is also important to highlight that it exists four types of \ac{NAT} implementations\cite{rfc3489}: \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.
 %RP apresentas 5. falas em asymmetric. É um erro?
@@
 
-Applications based on multimedia and peer-to-peer file sharing have been one of the most strained by \ac{NAT}.
-Those kind of applications require real time communication in order to achieve the best performance.
+Applications based on multimedia and peer-to-peer file sharing have been one of the most strained by the \ac{NAT} technique.
+It is also important to highlight that those kind of applications require real time communication in order to achieve the best performance.
 

### 2016-05-06 tese2/09_related_work_signaling.tex
[RFT] abstract and introduction rewrite

--- 2016-03-25 tese2/09_related_work_signaling.tex
@@
   Signaling is the process by which applications exchange connection information about peers and servers, their capabilities and meta-data.
-  \ac{WebRTC} doesn't implement signaling, as different applications may require different protocols, and there is no single answer that fits all problems.
-  Multiple options are available for signaling, which can be performed using \ac{SIP}, \ac{XMPP}, \emph{WebSockets}, \emph{Socket.io}\footnote{\url{http://socket.io/}(accessed June 1, 2015).} or by implementing a custom protocol.
+  In particular, \ac{WebRTC} does not implement signaling, as different applications may require different protocols and there is no single answer that fits all problems.
+  As a consequence, multiple options are available for filling the missing \ac{WebRTC}'s signaling component, which can be performed using \ac{SIP}, \ac{XMPP}, \emph{WebSockets}, \emph{Socket.io}\footnote{\url{http://socket.io/}(accessed June 1, 2015).} or by implementing a custom protocol.
 

### 2016-05-06 tese2/introduction.tex
[RFT] abstract and introduction rewrite

@@
 
-\section{Proposed Solution}
-\label{section:proposed}
-
 	As communications technologies appeared, we adapted the way we communicate.
@@
 
+        %RP Parece-me que até aqui ainda é background ou motivação. Não me parece que estejas já a descrever a solução proposta.
+
+\section{Proposed Solution}
+\label{section:proposed}
+
 	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see.
-        %RP Parece-me que até aqui ainda é background ou motivação. Não me parece que estejas já a descrever a solução proposta.
 
@@
 
-	{\color{blue}Another point to consider is making available collaborative tools for text edition, instant messaging and video superimposed content creation and visualization.}
+	{\color{red}Our goal in this project is to develop an application targeted to the web platform, resorting to \ac{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \emph{HTML5}, \emph{CSS3} and \emph{JavaScript} including continuous media types such as video, music and animations.}
+
+	{\color{red}One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again or introduce hyper-content to it such as time annotations, interactive lists of topics and subtitles.
+In this context we also provide a simpler method for creating and synchronizing hyper-content using \emph{QR codes}.}
+
+	{\color{red}In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links and files to conference participants.}
+
+	{\color{red}Furthermore, another relevant feature is the possibility to compose multiple video streams into a single one, which enables adding more users to conference rooms without impacting on clients performance.
+Users can change to individual streams on demand or automatically to the talking users.}
 
@@
 
-A web application with an easy to learn user interface will be developed to accomplish solving our problem.
+A web application with an easy to learn user interface was developed to accomplish solving our problem.
 Our application, unnamed yet, is targeted at web browsers that are compatible with only standard technologies like JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
-Any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
+Any additional plug-in was avoided, \emph{JavaScript} libraries were preferred as they can be downloaded on the fly.
 
@@
 
-	Although we propose an initial solution to solve our problem, we will make a continuous effort to survey the systems that solve our problem either completely or just a part of it, not excluding solutions that may appear during the implementation of our project.
-
-	We will present an architecture that can meet our goals, implement the respective prototype and test it with real users and performance benchmarks.
+	We have presented an architecture that can meet our goals, implemented the respective prototype and tested it with real users and performance benchmarks.
 
@@
 
-	All the problems faced during the development and limitations will be reported on the thesis so that a future project better then ours can be easily and better developed.
+	All the problems faced during the development and limitations were reported on the thesis so that a future project better then ours can be easily and better developed.
 

### 2016-05-06 tese2/07_related_work_nat.tex

--- 2016-05-06 tese2/07_related_work_nat.tex
@@
 There are four types of \ac{NAT} implementations\cite{rfc3489}: \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.
-%RP apresentas 5. falas em asymmetric. É um erro?
 

### 2016-05-06 tese2/11_related_work_collaboration.tex
[ADD] references to tables

@@
 Section \ref{mediatype} describes the media types and what types of media should be streamed.
-Section\ref{intrecord} addresses how an interactive media can be recorded.
+Section \ref{intrecord} addresses how an interactive media can be recorded.
 Section \ref{collabenv} presents the collaborative environment and libraries to synchronize distributed object among users.
@@
 	{\color{blue}
-	\ac{KMS} supports trans-coding, group communications, recording, mixing, broadcasting, applying filters, image and sound analysis, for example it allows face and qrcode detection.
-\ac{KMS} functionalities are exposed through \emph{JSON-RPC} over \emph{WebSocket}, there are three clients available: \emph{JavaScript} client for web browsers, \emph{Java} client for \emph{Java EE} servers and \emph{JavaScript} client for \emph{Node.js} servers.
+	  \ac{KMS} supports trans-coding, group communications, recording, mixing, broadcasting, applying filters and performing image and sound analysis (for example it allows face and QRcode detection).
+\ac{KMS} functionalities are exposed through \emph{JSON-RPC} over \emph{WebSocket}.
+          There are three clients available: \emph{JavaScript} client for web browsers, \emph{Java} client for \emph{Java EE} servers and \emph{JavaScript} client for \emph{Node.js} servers. 
 	\ac{KMS} supports streaming over \emph{WebRTC}, \ac{HTTP} and \ac{RTP} endpoints.
-Another important component of \ac{KMS} is \emph{Kurento Repository} which supports recording and playing directly from \emph{MongoDB}, that is important for providing a scalable media storage.
-Unlinke \emph{Jitsi Video Bridge}, \ac{KMS} does not enforce a specific signaling protocol.
+Another important component of \ac{KMS} is \emph{Kurento Repository}, which supports recording and playing directly from \emph{MongoDB}.
+That is important for providing a scalable media storage.
+Unlike \emph{Jitsi Video Bridge}, \ac{KMS} does not enforce a specific signaling protocol.
 	}
@@
   %RP e usa subsecções...
-\begin{figure}[H]
+\begin{figure}[!htb]
 	\centering
@@
 	\emph{Google Wave} was a distributed collaboration platform based on \emph{Jupiter}\cite{jupiter} that adopted \ac{OT} techniques.
-        	\ac{OT} technology was originally developed for consistency maintenance and concurrency control over distributed objects, \ac{OT} algorithms are mainly used in collaborative applications such as distributed document edition.
+        \ac{OT} technology was originally developed for consistency maintenance and concurrency control over distributed objects.
+        \ac{OT} algorithms are mainly used in collaborative applications such as distributed document edition.
     Other Google products, such as Google Docs, are also using this type of technology.
@@
 {\color{blue} It uses \emph{ShareDB} \footnote{\url{https://github.com/share/sharedb}(accessed March 10, 2016)} for its backend and data model, which supports simple integration with any database.
-One of \emph{ShareDB}'s integration is over \emph{MongoDB}\footnote{\url{https://github.com/share/sharedb-mongo}(accessed March 10, 2016)}}.
+One of \emph{ShareDB}'s integrations is over \emph{MongoDB}\footnote{\url{https://github.com/share/sharedb-mongo}(accessed March 10, 2016)}}.
 
@@
 {\color{blue}\emph{TogetherJS} uses its own servers for the signaling phase and it supports microphone and mouse sharing between users.
-It requires a simple server (also known as hub) that echoes messages between clients, all the synchronization work is performed by the clients.
-Although the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}, if we would not choose \emph{NodeJS} for server implementation we would need read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}
+It requires a simple server (also known as hub) that echoes messages between clients, with all the synchronization work being performed by the clients.
+Despite the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}.
+If we chose a server other than \emph{NodeJS} for our implementation, we would need to read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}
+        %RP terias de alterar o servidor onde corres o software (é o que está escrito) ou portar o hub para o novo servidor? São coisas muito diferentes.
 
@@
 
-	{\color{blue}\emph{Etherpad Lite} is a collaborative text editor implemented with \emph{NodeJS} that allows not only the text operations but also to associate authorships to pieces of text.
-It supports adding functionality trough plug-ins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example:
+	{\color{blue}\emph{Etherpad Lite} is a collaborative text editor implemented with \emph{NodeJS} that allows not only the basic text operations but also to associate authorships to pieces of text.
+It supports adding functionality through plug-ins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example:
 		\begin{itemize}
@@
 	}
+%RP acho que não vale a pena fazer o itemize. Podia ficar tudo na linha, assim: 1, 2, 3...
 
-	{\color{blue}\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations on client side over plain text.
-An implication of implementing just the client side is the extra effort that is necessary to implement content's persistent storage, besides this drawback this library is very flexible because it's not tied to a specific database or server side technology.}
+	{\color{blue}\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations over plain text on the client side.
+An implication of implementing just the client side is the extra effort that is necessary to implement content's persistent storage.
+Besides this drawback, this library is very flexible because it's not tied to a specific database or server side technology.}
 
-\begin{table}[H]
+	{\color{blue} In summary, table \ref{table:otcomparision} shows a comparision of the \ac{OT} libraries presented so far.}
+
+\begin{table}[!htb]
 \centering
 \caption{Comparision between Operational Transformation libraries}
-\label{my-label}
+\label{table:otcomparision}
 \begin{tabular}{|c|c|c|c|}

### 2016-05-06 tese2/12_related_work_summary.tex

@@
 
-Due to the use of \ac{NAT, bi-directional communications between clients have different needs from request-response based communications between clients and servers.
+Due to the use of \ac{NAT}, bi-directional communications between clients have different needs from request-response based communications between clients and servers.
 This lead to the appearance of mechanisms such as \ac{STUN}, \ac{TURN} and \ac{ICE}, in order to bypass the limits imposed by \ac{NAT}.

### 2016-05-06 tese2/architecture.tex

@@
 
+%RP isto pode ser muito expandido. Podes começar por descrever os goals (mais alto nível): audio e video conferência multi-party, possibilidade de rever o que foi dito antes, chat, etc etc.
+% Só depois traduzes isso em requisitos, tanto funcionais como não funcionais
+% Depois apresentas a arquitectura, fazendo um paralelo entre os módulo e escolhas realizadas e quais os requisitos que cada um deles satisfaz.
+
 The requirements for our web application are:
@@
 
+%RP capacidade de lidar com NAT
+
 	Moreover we allow clients to discover chat rooms and other clients by navigating on the web pages provided by our web server.
 In addition users can create rooms for multi-party audio and video communication communication which is achieved by using \ac{WebRTC}'s \emph{MediaStream} and \emph{PeerConnection}.
+        %RP falas em peerconnection, mas na figura 3.1 apenas há um cliente, o que obviamente impede mostrar a comunicação entre peers.
+        %RP acho que podes fazer outro diagrama com a comunicação entre os vários componentes, onde apareceriam vários clientes.
+A aí indicas o tipo de ligação entre cada par de componentes (HTTP, PeerConnectino, websocket, MediaStream ...)
 
 	Furthermore, chat rooms can be public or private.
-Public chat rooms are moderated by a group of clients which initially is only the room creator, this type of room has no access restrictions by default, but that can be changed by its moderators.
-Private chat rooms will be visible for a defined list of clients or can be accessed by clients that have a link for that conference room.
+Public chat rooms are moderated by a group of clients which initially consistes only of the room creator.
+This type of room has no access restrictions by default, but that can be changed by its moderators.
+Private chat rooms will only be visible to a defined list of clients or can be accessed by clients that have a link for that conference room.
 
 \section{Modules}
-
-	In this section we present the modules of our application.
+In this section we present the several modules that were designed in order com fulfill the set requirements.
+	Figure~\ref{fig:modules} presents the structure of our system which was divided into six modules:
 
@@
 \end{figure}
-%RP as figuras têm de ser apresentadas. Figure~\ref{fig:modules} presents ...
-	Figure~\ref{fig:modules} presents the structure of our system which is divided into six modules:
 
@@
 
-\item \underline{Application module} - responsible to provide information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client}.
+\item \underline{Application module} - responsible for providing information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client}.
 
@@
 
- \item \underline{Storage module} - responsible store all the model information and media recorded.
+ \item \underline{Storage module} - provides two main functionalities: store the model information and media recorded.
+This is the single module responsible for persistent storage.
+It stores user and communication data as well as all the data required among user sessions.
+It is also use to store all the communication streams, so that they can be viewed later.
 
@@
 \end{itemize}
+
+%RP Acho que podes escrever muito mais sobre cada componente. Isto está muito curto. Fala de algumas das funções que proporcionam. Vê o que fiz com o Storage Module
+
+%RP também podes fazer um diagrama com vários utilizadores e mostrar quais os streams que vão de e para cada um (mostrar a mistura de som, video, etc).
 
@@
 The infrastructure is composed by: web server, stream server, signaling server, database and video repository.
+%RP tens da fazer o paralelo entre isto e os componentes apresentados na secção anterior.
+O mesmo com a imagem, que parece não estar relacionada.
+Podes colocar os elementos na mesma posição, ou dentro das caixas da imagem 3.1
 
@@
 
-In order to simplify our solution, we propose both the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules implementation within the same server application so it could be easy to deploy as a single image.
-To this set of modules we often call \emph{backend}.
+In order to simplify our solution, we propose that the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules be implemented within the same server application, so it could be easy to deploy as a single image.
+To this set of modules, we often call \emph{backend}.
 
 Having established that the \emph{backend} modules are placed in the same machine, that helps controlling which resources the client has permission to access as those modules are seen as a private network.
+%RP na mesma ``server application'' ou ``machine''?
+São coisa diferentes!
+Apresentas vantages de instalar tudo na mesma máquina mas não falas das desvantagens.
+Menor escalabilidade parece-me ser logo uma delas.
+Na figura 3.2 aparecem várias máquinas!
 
 To qualify the above we provide public access to \ac{HTTP} server ports, maintaining the access to other components restricted through firewall rules.
-In relation to the database there is no directly access from outside, all the database information is accessed via our application server which validates the permissions of users when acting above our system.
+%RP qualify?
+%RP Imagem com firewall e estes portos? Mas isto já me parece muito detalhe de implementação.
+In relation to the database there is no directly access from the outside. All the database information is accessed via our application server which validates the permissions of users on our system.
 
-On the other hand, the access to our streaming servers is also restricted but clients can connect to them after concluding the signaling phase.
-This signaling phase may or not proceed in function of the client access permissions, for example if a user is trying to access a private conference room that is not member of or has no invitation link, the signaling server rejects to start the signaling phase and the user cannot access to the streaming server.
+On the other hand, the access to our streaming servers is also restricted, but clients can connect to them after concluding the signaling phase.
+This signaling phase may or not proceed in function of the client access permissions.
+For example, if a user is trying to access a private conference room that he is not a member of nor has an invitation link for, the signaling server refuses to start the signaling phase and the user cannot access the streaming server.
 
-Furthermore, we have taken into account the compatibility between the streaming server, database and the operation transformation solutions in order choose the appropriate framework to implement our web server.
+Furthermore, we have taken into account the compatibility between the streaming server, database and the operation transformation solutions, in order choose the appropriate framework to implement our web server.
+%RP esta secção devia começar aqui, a explicar a escolha das tecnologias/produtos a usar em cada componente.
+O que tens antes são pormenores que só faz sentido discutir no final.
+Tens de ir sempre do mais geral para o mais detalhado.
 
 As a result of our study about signaling protocols we have decided to exclude \ac{XMPP} due to the development difficulties that it exposes when using multiple web pages, namely the user re-authentication performed each time a page is loaded and the single tab usage limitation.
-By consequence we have not chosen \emph{Jitsi Video Brigde} as it uses \ac{XMPP}.
+As a consequence we have not chosen \emph{Jitsi Video Brigde} for stream processing as it uses \ac{XMPP}.
 
 Accordingly, by excluding \emph{Jitsi Video Brigde} we have decided that our solution must use \ac{KMS}.
-Our web server could be implemented easily with \emph{NodeJS} or \emph{Java} due to the fact \ac{KMS} provides clients for both technologies but others could be used as \ac{KMS} also exposes their \ac{API} via \emph{WebSockets}.
+Our web server could be implemented easily with \emph{NodeJS} or \emph{Java} due to the fact \ac{KMS} provides clients for both technologies.
+But others could also be used as \ac{KMS} also exposes their \ac{API} via \emph{WebSockets}.
+
+Due to the fact we are going to use \ac{KMS} as streaming server solution, we could choose \emph{NodeJS} or any \emph{Java} based web framework for implementing our web application server.
+One of our criteria for choosing the web framework is the ability to follow the \ac{MVC} paradigm which can help us to organize our code.
+Without any strong reason we have decided to implement our web server with the \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java}.
+%RP nem sequer por estares mais familiarizado com java ou a playframework? Ou porque querias aprender? Alguma razão há-de ter havido
 
@@
 
-Due to the fact we are going to use \ac{KMS} as streaming server solution, we could choose \emph{NodeJS} or any \emph{Java} based web framework for implementing our web application server.
-One of our criteria for choosing the web framework is the ability to follow the \ac{MVC} paradigm which can help us to organize our code.
-Without any strong reason we have decided to implement our web server with \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java}.
-
 Importantly, for the collaborative text editor, we have chosen \emph{OT.js} due to its server and storage implementation choice independence.
 
-For the \emph{NAT Traversal} module an \ac{ICE} server is not required to be on our infrastructure, as a public \ac{STUN} server can be used for testing our solution although we recognize that for a production environment we would need to maintain our own \ac{TURN} servers in order to ensure connectivity to all clients.
+For the \emph{NAT Traversal} module, an \ac{ICE} server is not required to be on our infrastructure, as a public \ac{STUN} server can be used for testing our solution.
+Nevertheless, we recognize that for a production environment we would need to maintain our own \ac{TURN} servers in order to ensure connectivity to all clients.
 
-Not less important, on the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} should be installed as web browsers.
-Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Adapter.js}, \emph{OT.js} can be downloaded from the web server and executed on the client side.
+Not less important, on the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} could be installed as web browsers.
+As such, both should be supported.
+Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Adapter.js}, \emph{OT.js} can be downloaded from the web server and executed on the client side using any of these two browsers.
+%RP relembrar que o IE e o Safari não suportam webrtc?
 
@@
 
+%RP acho interessante o que queres fazer aqui mas acho que está mal conseguido. A linha de baixo, devia ter coisas ao mesmo nível, o que não aconteçe.
+% Qual a razão para HTTP estar por baixo do jquery e user interface estar ao lado?
+% Sugiro colocar a applicação no meio e as várias interfaces (webrtc, websocket e http) uma de cada lado (esquerda, baixo e direita), indicando o que fica do outro lado (outro cliente, servidor web, servidor stream, etc)
+
 Table~\ref{table:apparch} presents the application architecture and the underlying technologies seen from the user's perspective.
@@
 With \emph{Bootstrap} it is quite easy to develop applications that adapt to mobile devices with different screen sizes.
+%RP bootstrap não aparece na figura!
 
-The synchronization between multimedia elements will be performed trough chains of \emph{JavasSript} events or by specifying the interval of time which time content must be visible.
+The synchronization between multimedia elements will be performed through chains of \emph{JavasSript} events or by specifying the interval of time which time content must be visible.
 Other animations can be implemented with \ac{SVG} embedded on \ac{HTML}.
+%RP Caiu do céu! Tens de explicar e ligar melhor a sequencia de texto
 

### 2016-05-06 tese2/07_related_work_nat.tex
[RFT] related work nat  minor improvements

@@
 \ac{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}.
-{\color{blue} Today we live in the \emph{Internet of Things} era.
-  Users and companies may have multiple devices that require Internet connectivity, such as personal computers, smart phones, tablets, traffic sensors, public vigillance systems, health monitoring tools and so on.
+{\color{blue} Today we live in the \ac{IoT} era.
+  Users and companies may have multiple devices that require Internet connectivity, such as personal computers, smart phones, tablets, traffic sensors, public surveilance systems, health monitoring tools and so on.
 All those devices combined cannot be publicly addressed by \ac{IPv4}.}
@@
 %RP o porto não identifica uma máquina (membro da rede privada) mas uma ligação.
-This technique is known as \ac{NAT}.
+This technique is known as \ac{NAT} (figure \ref{fig:nat}).
 
-\begin{figure}[H]
+\begin{figure}[!htb]
 	\centering
@@
 	\caption{Network Address Translation}
+	\label{fig:nat}
 \end{figure}
@@
 
-\begin{figure}[H]
+\begin{figure}[!htb]
 	\centering
@@
 	\caption{Multiple level NATs}
+	\label{fig:multinat}
 \end{figure}
@@
 %RP file sharing não precisa de real-time
-\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}. None of those can establish direct connections on multiple level \ac{NAT}s.
-%HR XXX I was here
+\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}. None of those can establish direct connections on multiple level \ac{NAT}s (figure \ref{fig:multinat}).
 %RP explicar o que são multiple level pois não falaste disto antes? São ``nested''?

### 2016-05-06 tese2/11_related_work_collaboration.tex
[RFT] state of art nearly closed

@@
 	{\color{blue}\emph{Etherpad Lite} is a collaborative text editor implemented with \emph{NodeJS} that allows not only the basic text operations but also to associate authorships to pieces of text.
-It supports adding functionality through plug-ins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example:
-		\begin{itemize}
-		\item exporting and importing document formats such as \emph{DOC}, \emph{PDF}, \emph{ODT} and \emph{DOCX}.
-		\item painting and drawing.
-		\item video and audio chat using WebRTC.
-		\item create sideshows.
-		\item spell checking.
-		\item text to speech.
-		\end{itemize}
+It supports adding functionality through plug-ins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example: exporting and importing document formats such as \emph{DOC}, \emph{PDF}, \emph{ODT} and \emph{DOCX}, painting and drawing, video and audio chat using WebRTC, create sideshows, spell checking and text to speech.
 	}

### 2016-05-06 tese2/09_related_work_signaling.tex
[FIX] reference to stroph.js

@@
   \ac{XMPP}'s bi-directional communication over \ac{HTTP} is achieved through \ac{BOSH} \cite{xep0206}.
-  %RP falta um `` bi-directional communication OVER HTTP is achieved through''?
-  %RP BOSH já foi explicado antes pelo que não vale a pena voltar a falar em long-pooling
   This kind of communication is also possible through WebSockets \cite{rfc7395}.
@@
 
-    An important aspect of this model is the ability to define a user as a set of attributes that can change dinamically in order to fit the \ac{IM} application's needs.
+    An important aspect of this model is the ability to define a user as a set of attributes that can change dynamically in order to fit the \ac{IM} application's needs.
 
     The connection to the \ac{XMPP} server can be done through the same web server as the user is connected, but this web server would handle a lot of connections and that would have a great impact on the overall system performance.
+On the other hand, users can directly connect directly to \ac{XMPP} servers from web applications using the \emph{JavaScript} library \emph{strophe.js} \footnote{\url{http://strophe.im/strophejs/} (accessed May 6, 2016).}
 
     Typically a web application consists in multiple web pages which are navigated by users.
-Each time a webpage is accessed, either from a new context or a transition from a previous page, its context is cleared except for local storage and cookies.
-The javascript context is cleared, including the \ac{XMPP} connections performed by \emph{strophe.js}.
+Each time a web page is accessed, either from a new context or a transition from a previous page, its context is cleared except for local storage and cookies.
+The \emph{JavaScript} context is cleared, including the \ac{XMPP} connections performed by \emph{strophe.js}.
 As such, an automatic mechanism is required for avoiding the implicit user reauthentication.
-    One solution for reauthentication can be achieved by storing the user's JabberID and password on local storage and everytime a page is accessed the authentication is performed without the users knowledge.
+    One solution for reauthentication can be achieved by storing the user's JabberID and password on local storage and every time a page is accessed the authentication is performed without the users knowledge.
 Clearly this solution represents security flaws as the local storage can be easily accessed locally or by performing a cross site scripting attack to reveal all the local storage.
@@
 
-    An alternative solution is known as Session Attachment \footnote{\url{https://metajack.im/2008/10/03/getting-attached-to-strophe/}}, which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} be passed to strophe.js in order to re-connect to the same stream on \ac{XMPP} server.
-Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request making it worthless if a user mantains more than one tab opened, for example for multiple conversations at the same time.
+    An alternative solution is known as Session Attachment \footnote{\url{https://metajack.im/2008/10/03/getting-attached-to-strophe/} (accessed May 6, 2016).}, which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} be passed to \emph{strophe.js} in order to re-connect to the same stream on \ac{XMPP} server.
+Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request making it worthless if a user maintains more than one tab opened, for example for multiple conversations at the same time.
     %RP either ou both?
@@
 
-    Another alternative would be the development of our solution in a single web page, which would increse drastically the complexity of our web application.
+    Another alternative would be the development of our solution in a single web page, which would increase drastically the complexity of our web application.
 
-    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that \ac{XMPP} already implemment, but  we don't need all of them.
+    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that \ac{XMPP} already implement, but  we don't need all of them.
 In fact, we just need a subset of \ac{XMPP} features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.

### 2016-05-06 tese2/07_related_work_nat.tex
[FIX] reference to stroph.js
[RFT] removed some blue text and some old RP comments

@@
 \ac{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}.
-{\color{blue} Today we live in the \ac{IoT} era.
+Today we live in the \ac{IoT} era.
   Users and companies may have multiple devices that require Internet connectivity, such as personal computers, smart phones, tablets, traffic sensors, public surveilance systems, health monitoring tools and so on.
-All those devices combined cannot be publicly addressed by \ac{IPv4}.}
+All those devices combined cannot be publicly addressed by \ac{IPv4}.
 %RP vigillance -> surveilance?
@@
 Initially \ac{NAT} offered an alternative to address exhaustion and a minimal sensation of security.
-However, due to their current wide usage, \ac{NAT}s weaknesses are being exposing at the application layer, {\color{blue}namely impacting applications that require direct communications between two private networks.}
+However, due to their current wide usage, \ac{NAT}s weaknesses are being exposing at the application layer, namely impacting applications that require direct communications between two private networks.
 
@@
 
-{\color{blue}Unless the router that performs as \ac{NAT} has forwarding rules to every desired ports of each user}, applications behind a \ac{NAT} are prevented from receiving incoming connections from the public network, which forces them to behave as a client of a client-server model.
-
-%RP explicar que o nat funciona num cenário em que a máquina atrás do nat é apenas um consumidor de informação, trabalhando em modo cliente do modelo cliente-servidor. Em P2P isto falha.
+Unless the router that performs as \ac{NAT} has forwarding rules to every desired ports of each user, applications behind a \ac{NAT} are prevented from receiving incoming connections from the public network, which forces them to behave as a client of a client-server model.
 
@@
 
-%RP file sharing não precisa de real-time
 \ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}. None of those can establish direct connections on multiple level \ac{NAT}s (figure \ref{fig:multinat}).
@@
 That is why Symmetric \ac{NAT} represents a problem for peer-to-peer communications.
-%RP os problemas que relatas têm a ver com P2P ou real-time?
 
@@
 \ac{ICE} is a technique that uses \ac{STUN} when direct communications are possible and \ac{TURN} while a direct communication isn't possible.
-%RP ainda não explicaste o que é o ICE
 %RP explicar que para usar TURN as aplicações têm de estar cientes disso.
@@
 The requirements of real-time video communication makes this kind of model unsuitable.
-%RP quando falas em infraestrutura mais dispendiosa para o video, está a falar da internet ou ainda estamos a falar de usar ICE?
 

### 2016-05-06 tese2/11_related_work_collaboration.tex
[RFT] removed some blue text and some old RP comments

@@
 
-	{\color{blue}
 	  \ac{KMS} supports trans-coding, group communications, recording, mixing, broadcasting, applying filters and performing image and sound analysis (for example it allows face and QRcode detection).
@@
 Unlike \emph{Jitsi Video Bridge}, \ac{KMS} does not enforce a specific signaling protocol.
-	}
 
@@
 	\emph{ShareJS} is an \ac{OT} \emph{JavaScript} library, developed by the ex \emph{Google Wave} engineer Joseph Gentle, for collaborative text and \ac{JSON} documents edition in real-time.
-{\color{blue} It uses \emph{ShareDB} \footnote{\url{https://github.com/share/sharedb}(accessed March 10, 2016)} for its backend and data model, which supports simple integration with any database.
-One of \emph{ShareDB}'s integrations is over \emph{MongoDB}\footnote{\url{https://github.com/share/sharedb-mongo}(accessed March 10, 2016)}}.
+It uses \emph{ShareDB} \footnote{\url{https://github.com/share/sharedb}(accessed March 10, 2016)} for its backend and data model, which supports simple integration with any database.
+One of \emph{ShareDB}'s integrations is over \emph{MongoDB}\footnote{\url{https://github.com/share/sharedb-mongo}(accessed March 10, 2016)}.
 
@@
 It uses \ac{JSON} messages for \ac{OT} concurrency control but it does not provide storage.
-{\color{blue}\emph{TogetherJS} uses its own servers for the signaling phase and it supports microphone and mouse sharing between users.
+\emph{TogetherJS} uses its own servers for the signaling phase and it supports microphone and mouse sharing between users.
 It requires a simple server (also known as hub) that echoes messages between clients, with all the synchronization work being performed by the clients.
 Despite the hub's reduced complexity, \emph{TogetherJS}'s server is implemented with \emph{NodeJS}.
-If we chose a server other than \emph{NodeJS} for our implementation, we would need to read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.}
+If we chose a server other than \emph{NodeJS} for our implementation, we would need to read the hub's source code and understand what changes are needed to perform on our web server in order to make it compatible with \emph{TogetherJS}.
         %RP terias de alterar o servidor onde corres o software (é o que está escrito) ou portar o hub para o novo servidor? São coisas muito diferentes.
@@
 
-	{\color{blue}\emph{Etherpad Lite} is a collaborative text editor implemented with \emph{NodeJS} that allows not only the basic text operations but also to associate authorships to pieces of text.
+	\emph{Etherpad Lite} is a collaborative text editor implemented with \emph{NodeJS} that allows not only the basic text operations but also to associate authorships to pieces of text.
 It supports adding functionality through plug-ins\footnote{\url{http://static.etherpad.org/plugins.html}(Accessed 20 March 2016)}, for example: exporting and importing document formats such as \emph{DOC}, \emph{PDF}, \emph{ODT} and \emph{DOCX}, painting and drawing, video and audio chat using WebRTC, create sideshows, spell checking and text to speech.
-	}
+
 %RP acho que não vale a pena fazer o itemize. Podia ficar tudo na linha, assim: 1, 2, 3...
 
-	{\color{blue}\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations over plain text on the client side.
+	\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations over plain text on the client side.
 An implication of implementing just the client side is the extra effort that is necessary to implement content's persistent storage.
-Besides this drawback, this library is very flexible because it's not tied to a specific database or server side technology.}
+Besides this drawback, this library is very flexible because it's not tied to a specific database or server side technology.
 
-	{\color{blue} In summary, table \ref{table:otcomparision} shows a comparision of the \ac{OT} libraries presented so far.}
+	In summary, table \ref{table:otcomparision} shows a comparision of the \ac{OT} libraries presented so far.
 

### 2016-05-07 tese2/13_implementation_model.tex
Reviewing data model

@@
-\section{Model}
-The model is the most important component of our solution, a badly designed model can imply serious difficulties when implementing new features that are not part of the plans, sometimes we had to redesign the model in order to support new features.
+%RP antes desta secção fala um introdução ao capítulo a explicar o que vai ser apresentado.
 
+\section{Data Model}
+The data model is a critical component of our solution, as a badly designed model can imply serious difficulties when implementing new features that are not part of the plans.
+During the course of this project, we had to redesign the model more than once in order to support new features.
+%RP mudei ``most important'' para critical
+
+%
 \subsection {Schema representation}
@@
     Each \emph{MongoDB} document is represented by a \emph{JSON} object and, as a result, each document may have different attributes within a collection.
-Needless to say that following this approach we don't need to create the collections with a well predefined schema, in fact we don't to define it at all.
+Needless to say that, by following this approach we don't need to create the collections with a well predefined schema.
+In fact we don't need to define it at all.
 Although, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure.
+    %RP a última frase é confusa. O que queres dizer? Reescreve
 
-    Similarly to relational databases, \emph{MongoDB} requires a primary key (typically typed as \emph{ObjectId} and named \emph{"\_id"}) for each document which is automatically assigned if not specified.
+    Similarly to relational databases, \emph{MongoDB} requires a primary key (typically of type \emph{ObjectId} and named \emph{"\_id"}) for each document, which is automatically assigned if not specified.
 
@@
 
-    In respect to attributes nullability, \emph{MongoDB} does not enforce a document's attribute to have a non nullable value, although, for sake of good functionality, we perform those constraints validation programmatically and by so we also represent them in our schema.
+    In respect to attributes nullability, \emph{MongoDB} does not enforce a document's attribute to have a non nullable value, although, for sake of good functionality, we perform those constraints validation programmatically and as such we also represent them in our schema.
+    %RP nullable ou null?
 
@@
 
-For designing our model we have taken into account generic programming techniques. We observed that operations like searching for an object was quite repeated amongst different types of objects.
+For designing our model we have taken into account generic programming techniques. We observed that operations like searching for an object were quite repeated across different types of objects.
 
-Our first decision for our model in order to avoid repeated code, was the isolation of the object's attributes from themselves, so we could apply the search operation to a set of attributes independently from the object type.
-To this generic set of properties we call data (table \ref{table:generic}) and each object of this type has a reference to the owner, which is a unique identification number.
+Our first decision for our model, in order to avoid repeated code, was the isolation of the object's attributes from themselves, so we could apply the search operation to a set of attributes independently of the object type.
+To this generic set of properties we call data (Table \ref{table:generic}) and each object of this type has a reference to the owner, which is a unique identification number.
 
@@
 
-The identification number by itself is not sufficient to identify an object, objects from different types can have the same identification number.
+The identification number by itself is not sufficient to identify an object, as objects from different types can have the same identification number.
 In order to solve this problem, when an object is created, its correspondent data must contain the owner's object type.
+%RP mas o _id é apresentado com sendo uma chave primária na figura 4.2, logo única!
 
-Whenever an attribute is created, it must be specified the attributes name, its value, searcheability, identifiability and the set of object identification numbers that can read and write that attribute.
+Whenever an attribute is created, its name, value and indentifiability must be specified. The set of objects that can read and write that attribute must also be defined. 
+In regard to our permission mechanism, if the read or write sets are not specified we assume the attribute is readable and writable by everyone.
+Conversely if the read and write sets are empty, nobody is allowed to read or write the attribute.
+Implicitly, if an entity can write an attribute, it can also read it.
 
-In relation to our permission mechanism, if the read or write sets are not specified we assume the attribute is readable and writable by everyone, conversely if the read and write sets are empty, nobody is allowed to read or write the attribute.
-Implicitly if an entity can write an attribute it can also read it.
-
-In particular, if all attributes were searchable it could be simple to search for attributes that could reveal sensible information about an object.
-For example if we consider that a user could have a health related attribute, searching by a disease would reveal which users could suffer from a certain disease, the leak of that kind of information could, for instance, change the agreement between users and health insurance companies.
+In particular, if all attributes were searchable, it could be simple to search for attributes that could reveal sensible information about an object.
+For example if we consider that a user could have a health related attribute, searching by a disease would reveal which users could suffer from a certain disease.
+The leak of that kind of information could, for instance, change the agreement between users and health insurance companies.
 For this reason only the specified attributes as searchable will be taken into account when performing keyword searches.
+%RP era melhor que o exemplo fosse relativo à tua aplicação e não algo abstracto como saúde.
+%RP Dás permissões de leitura e escrita a objectos. Normalmente é a utilizadores. Explicar relação. Assumo que sejam aqui referenciados apenas objectos que representam utilizadores
 
 Another important attribute specification is the owner identifiability, which tells us if the attribute identifies the object.
-This specification let us create abstract authentication services, for example a user can login into our system by providing any attribute that identifies himself, for example the e-mail but others are possible like the user name or cellphone number.
+This specification lets us create abstract authentication services.
+For example a user can login into our system by providing any attribute that identifies himself, e.g.
+the e-mail but others are possible like the user name or cellphone number.
 
-Not less important, in order to get an object properties efficiently, we have created an index over the \emph{owner} attribute.
-In the meantime , we have also created an index over the \emph{searchableValues} in order to improve the keyword search performance.
+Not less important, in order to get an object's properties efficiently, we have created an index over the \emph{owner} attribute.
+We have also created an index over the \emph{searchableValues} in order to improve the keyword search performance.
 
-In summary, with this model we can perform search and identification of any kind of objects, as we will see on the following models, the \emph{user} and \emph{group} models are using this generic model for storing their attributes.
+In summary, with this model we can perform search and identification of any kind of objects, as we will see on the following models.
+In particular, the \emph{user} and \emph{group} models are using this generic model for storing their attributes.
 % Not less important, attributes can specify aggregations of objects.
@@
 
-The user model is not tied to the user attributes (table \ref{table:user}), the information maintained in this model is just used for authentication purposes.
+The user model is not tied to the user attributes (Table \ref{table:user}), the information maintained in this model is just used for authentication purposes.
 Passwords are not stored in plain text, instead we apply hashing and salting techniques \cite{password} in order to make it harder to decode the password by an attacker.
-Accordingly, we use \emph{SHA-1} and a random salt per user with 32 characters long.
+Accordingly, we use \emph{SHA-1} and a random salt per user with 32 characters.
+%RP Table x, Figure y, etc é sempre maiusculas.
 
@@
 
-A relation between two entities $e_1$ and $e_2$ is represented by the pair $e_1\rightarrow e_2$ (table \ref{table:relation}), where $e_1$ is the source and $e_2$ is the target.
-This relation is said bi-directional if and only if it also exists the relation $e_2\rightarrow e_1$.
+A relation between two entities $e_1$ and $e_2$ is represented by the pair $e_1\rightarrow e_2$ (Table \ref{table:relation}), where $e_1$ is the source and $e_2$ is the target.
+This relation is said to be bi-directional if and only if it also exists the relation $e_2\rightarrow e_1$.
 
@@
 A user can only interact with friends or with group members.
-In order to validate a friendship, both users must agree on that friendship, by other words it must exist a bi-directional relation between both users.
+In order to validate a friendship, both users must agree on that friendship, in other words it, there must exist a bi-directional relation between both users.
 
-In order to improve performance of queries over \emph{Relation} collection we have created indexes on the \emph{source}, \emph{target} and also on the pair composed by both attributes.
+In order to improve the performance of queries over the \emph{Relation} collection we have created indexes on the \emph{source}, \emph{target} and also on the pair composed by both attributes.
 
@@
 A group is composed by an \emph{id}, \emph{inviteToken} and a \emph{visibility} as shown on table \ref{table:group}.
-
 Moreover, a group can be public or private.
-If the group is public then it is visible to all users that maintain a friendship with a member of this group.
-If the group is private then it is only visible to their members.
+If the group is public, then it is visible to all users that maintain a friendship with a member of this group.
+If the group is private, then it is only visible to its members.
 
@@
 The group membership is a special case of relation, where the target entity is always a group.
-
-When a group is created, a group membership is automatically assigned by its creator.
+When a group is created, a group membership is automatically assigned to its creator.
 
 Entities that have a membership with a group can create more memberships by sharing an invite token or by specifying new group members.
+%RP isto quer dizer que quem tem o token se pode juntar ao grupo? Explicar. O que impede alguém de o partihar com outros?
 

### 2016-05-07 tese2/abstract.tex

@@
 
-{\color{red}One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again or introduce hyper-content to it such as time annotations, interactive lists of topics and subtitles.
-In this context we also provide a simpler method for creating and synchronizing hyper-content using \emph{QR codes}.}
-
-{\color{red}In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links and files to conference participants.}
-
-{\color{red}Furthermore, another relevant feature is the possibility to compose multiple video streams into a single one, which enables adding more users to conference rooms without impacting on clients performance.
-Users can change to individual streams on demand or automatically to the talking users.}
+{\color{red}In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor, time annotations, a chat that supports sending files to conference participants and mechanism to superimpose hyper-content to video.}
 
@@
 {\color{blue}
-In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose and implement an architecture for an hyper-linked communication application based on \emph{WebRTC} and provide the results for its implementation and evaluation.
+In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose and implement an architecture for an hyper-linked communication application based on \emph{WebRTC}.
+This work was evaluated using a panel of 20 users, who reported that they liked to use it and thought it to be extremely innovative.
 }

### 2016-05-07 tese2/13_implementation_model.tex
Reviewed data model

@@
 
-%
+%RP devias começar com um overview dos dados que é necessário guardar para a aplicação
+%RP ao apresentar o modelo falas de coisas como mensagens para grupos, mas ainda não sabemos nada disso.
+Na arquitectura, devias falar da concepção, do modelo de comunicação.
+Que tens utilizadores, grupos, mensagens, chats, convites, etc...
+
 \subsection {Schema representation}
@@
 
+%RP antes de falares de cada modelo, que tal incluir uma figura com a relação entre os vários tipos de dados.
+Mesmo não sendo SQL, as relações existem conceptualmente e falas em chaves estrangeiras, pelo que tal figura faz sentido e ajuda a perceber.
+
 \subsection{User model}
@@
 
-A message is composed by its content, time of creation and source and target identification numbers (table \ref{table:message}).
-The message's target can reference any of object but our application is only covering messages to groups.
+A message is composed by its content, time of creation and source and target identification numbers (Table \ref{table:message}).
+The message's target could reference any object, but our application is only handling messages to groups.
 
@@
 In order to improve performance of queries over \emph{Relation} collection we have created indexes on the \emph{target} attribute and also on the pair composed by \emph{target} and \emph{\_id} attributes for finding and sorting the messages received by an entity more efficiently.
+%RP pela justificação não consigo perceber para que serve a chave no par target e _id
 
@@
 
-During a group conversation it is possible to create time annotations for making it easy to access that time either by searching or sharing with other users.
-
+During a group conversation, it is possible to create time annotations for making it easy to access that time either for searching or sharing with other users.
 A time annotation (table \ref{table:hyper}) contains a title, the correspondent group identification number and the time itself.
@@
 The hyper content is used to synchronize content among users during a conversation.
-Table \ref{table:hyper} shows that every hyper content must have a start and ending time, the correspondent group identification number and the content itself in the form of text.
-Beside those attributes, in order to perform queries over the \ac{HTML} contents with more precision, we have added an additional \emph{searchableContent} attribute that contains just the searchable content extracted from the content excluding the \ac{HTML} tags and parameters using \emph{Jsoup}\footnote{\url{http://jsoup.org/} (Accessed April 11, 2016)}.
+Table \ref{table:hyper} shows that every hyper content must have a start and ending time, the correspondent group identification number and the content itself, in the form of text.
+Beside those attributes, in order to perform queries over the \ac{HTML} contents with more precision, we have added an additional \emph{searchableContent} attribute.
+It contains just the searchable content extracted from the content, by excluding the \ac{HTML} tags and parameters using \emph{Jsoup}\footnote{\url{http://jsoup.org/} (Accessed April 11, 2016)}.
+
+%RP explica melhor a diferença entre HyperContent e Time Annotation do ponto de vista do utilizador / funcionalidade proporcionada.
 
@@
 \begin{itemize}
-\item{\emph{\{groupId,start,end\}} $\rightarrow$ finding visible contents for a given time  within a group (intersections).}
-\item{\emph{\{groupId,start\}} $\rightarrow$  finding contents that starts after a given time within a group (for pre-loading).}
-\item{\emph{\{groupId,searchableContent\}} $\rightarrow$ searching contents by keywords within a group.}  
+\item{\emph{\{groupId,start,end\}} $\rightarrow$ for finding visible contents for a given time  within a group (intersections).}
+\item{\emph{\{groupId,start\}} $\rightarrow$  for finding contents that starts after a given time within a group (for pre-loading).}
+\item{\emph{\{groupId,searchableContent\}} $\rightarrow$ for searching contents by keywords within a group.}  
 \end{itemize}
@@
 
-Within a conversation, users can write documents collaboratively. As we can see in table \ref{table:collaborative}, each document has a content and a reference to the correspondent group.
+Within a conversation, users can write documents collaboratively. As we can see in Table \ref{table:collaborative}, each document has a content and a reference to the correspondent group.
 
@@
 
-During a conversation, users may allow sharing their web cameras, by doing so their video is stored in recording chunks.
-Each chunk, as seen in table \ref{table:recording}, represents an interval of time  $T=\big[c^{start},c^{end}\big[$, it contains the media \ac{URL}, a reference to a group, an owner identification number and the correspondent \emph{WebSocket} session id.
+During a conversation, users may allow sharing their web cameras.
+By doing so, their video is stored in recording chunks.
+Each chunk, described in Table \ref{table:recording}, represents an interval of time $T=\big[c^{start},c^{end}\big[$ of the video stream.
+It also contains the media's \ac{URL}, a reference to a group, an owner identification number and the correspondent \emph{WebSocket} session id.
+    %RP ainda não tenho forma de perceber o que é o websocket session id. Nunca falaste de websockets antes, quanto mais disto! Nem do media URL, é gravado ou webrtc (dispositivo do utilizador)?
 
-In order to allow different devices within the same user, storing just the \ac{URL} with an associated user id is not enough as in this case it is needed one more parameter to differentiate the different \emph{webSocket} sessions opened by the same user.
+In order to allow the same user to have different devices, storing just the \ac{URL} with an associated user id is not enough.
+In this case, one more parameter is needed to differentiate the different \emph{webSocket} sessions opened by the same user.
 For this reason we had to associate a random session identification number to each \emph{webSocket}.
@@
 A set of chunks $S=[c_1,c_2,\ldots,c_n]$ is said continuous if $\forall c_i\in S, \exists c_j \in S$ where $j\neq i$ and $c_i^{start} = c_j^{end} \vee c_i^{end} = c_j^{start}$. 
-
 A recording interval represents a continuous set of recording chunks.
 
-In order to improve the performance of searching recording chunks and calculate which chunks are are going to be reproduced, we have created indexes over the following sequences of attributes: 
+In order to improve the performance of searching recording chunks and calculate which chunks are going to be reproduced, we have created indexes over the following sequences of attributes: 
 \begin{itemize}
-    \item{\emph{\{groupId,start,end\}} $\rightarrow$ finding all available chunks for a given time} 
-    \item{\emph{\{groupId,sessionId,end\}} $\rightarrow$ finding intersecting chunks with a given time within a session}
-    \item{\emph{\{groupId,owner,end\}} $\rightarrow$ finding intersecting chunks with a given time within a user (for a different session) or group}
-    \item{\emph{\{groupId,sessionId,\_id\}} $\rightarrow$ finding chunks that follow a given \emph{id} within a session}
-    \item{\emph{\{groupId,owner,\_id\}} $\rightarrow$ finding chunks that follow a given \emph{id} within a user or group}
+    \item{\emph{\{groupId,start,end\}} $\rightarrow$ for finding all available chunks for a given time} 
+    \item{\emph{\{groupId,sessionId,end\}} $\rightarrow$ for finding intersecting chunks with a given time within a session}
+    \item{\emph{\{groupId,owner,end\}} $\rightarrow$ for finding intersecting chunks with a given time within a user (for a different session) or group}
+    \item{\emph{\{groupId,sessionId,\_id\}} $\rightarrow$ for finding chunks that follow a given \emph{id} within a session}
+    \item{\emph{\{groupId,owner,\_id\}} $\rightarrow$ for finding chunks that follow a given \emph{id} within a user or group}
 \end{itemize}
+%RP os _id não são únicos? Para que servem os index que os incluem?
+%RP para a intercepção não precisaste de usar o start, apenas o end?
 
-Not less important we have also created an index over the \emph{RecordingInterval}'s \emph{groupId} attribute for listing all intervals withing a conference room.
+Not less important, we have also created an index over the \emph{RecordingInterval}'s \emph{groupId} attribute for listing all intervals within a conference room. 
+%RP olha outra coisa nova! Ainda não falaste de conference rooms!

### 2016-05-08 tese2/architecture.tex
[RFT] improving architecture

@@
 
+\section{Requirements}
+In a general way our system's goal is to provide a multi-party video and audio conference environment that supports chat, time manipulation, collaborative text edition and hyper-content creation.
+
 %RP isto pode ser muito expandido. Podes começar por descrever os goals (mais alto nível): audio e video conferência multi-party, possibilidade de rever o que foi dito antes, chat, etc etc.
@@
 
-The requirements for our web application are:
+For our system, the goals include:
 
 \begin{itemize}
- \item Text, Audio and Video communication using \ac{WebRTC}.
- \item Ability to record and playback interactive multimedia.
- \item Create annotations over video and voice content.
- \item Enrich communications by overlaying multiple media types.
- \item Search and navigate through annotations and contents.
- \item Ability to create and share time links among users.
- \item Mix multiple user streams into a single stream.
- \item Sound detection for showing the current speaker.
- \item Structure presentation of the content.
- \item Collaborative text edition.
- \item Easy interface for content creation and synchronization. 
+ \item \textbf{Instant Messaging} - Our application must provide a simple way to send instant text messages to the conference participants.
+ \item \textbf{Room management} - Conference rooms can be public or private.
+Public chat rooms are moderated by a group of clients which initially consists only of the room creator.
+This type of room has no access restrictions by default, but that can be changed by its moderators.
+Private chat rooms will only be visible to a defined list of clients or can be accessed by clients that have a link for that conference room.
+ \item \textbf{Stream Recording} - Ability to recording and playback recorded video including all the hyper-content displayed at that time.
+ \item \textbf{Stream Composition} - Our streaming server must support mixing multiple user streams into a single stream.
+The sound played on clients must be the composed sound of all participants but the video played can be switched to the individual or group view.
+ \item \textbf{Time annotations} - Our application must give the possibility to create annotations associated to a specified time and support viewing them on a time-line. 
+ \item \textbf{Content Overlay} - Our system must allow users to superimpose hyper-content to video given a range of time.
+This content can be continuous or discrete and interactive or non interactive.
+ \item \textbf{Search} - Unified search every objects related to a conference room such as hyper-content, time annotations and users.
+ \item \textbf{Content Sharing} - Our application must have the ability to share files and time links among users.
+ \item \textbf{Active talker detection} - The system must enable sound detection for showing the current speaker.
+ \item \textbf{Collaboration} - The web application must provide a collaborative text edition tool.
+ \item \textbf{QR Code detection} - The system must interpret \emph{QR codes} in order to ease content creation.
+ \item \textbf{NAT Traversal} - The system's services must be available to users that are inside a private network.
+ \item \textbf{Fault tolerance} - Our system's must use a database that supports data replication.
+ \item \textbf{Scalability} - Our system must be horizontally scalable, which means that load must be distributed across web, streaming and database servers.
 \end{itemize}
@@
 	Moreover we allow clients to discover chat rooms and other clients by navigating on the web pages provided by our web server.
-In addition users can create rooms for multi-party audio and video communication communication which is achieved by using \ac{WebRTC}'s \emph{MediaStream} and \emph{PeerConnection}.
+In addition users can create rooms for multi-party audio and video communication communication which is achieved by using \ac{WebRTC}'s \emph{PeerConnection}.
         %RP falas em peerconnection, mas na figura 3.1 apenas há um cliente, o que obviamente impede mostrar a comunicação entre peers.
@@
 A aí indicas o tipo de ligação entre cada par de componentes (HTTP, PeerConnectino, websocket, MediaStream ...)
-
-	Furthermore, chat rooms can be public or private.
-Public chat rooms are moderated by a group of clients which initially consistes only of the room creator.
-This type of room has no access restrictions by default, but that can be changed by its moderators.
-Private chat rooms will only be visible to a defined list of clients or can be accessed by clients that have a link for that conference room.
 
@@
 
-\item \underline{Application module} - responsible for providing information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client}.
+\item \textbf{Application module} - responsible for providing information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client} in the form of web pages in \ac{HTML} and \emph{JavaScript} libraries through \ac{HTTP}.
 
-\item \underline{Signaling module} - responsible for \emph{Client} and \emph{Stream} coordination.
+\item \textbf{Signaling module} - responsible for \emph{Client} and \emph{Stream} coordination which will be performed using \emph{WebSockets}.
 
- \item \underline{NAT Traversal module} - used by \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase which ends by establishing the connection between them.
+ \item \textbf{NAT Traversal module} - \ac{STUN} and \ac{TURN} techniques used by \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase which ends by establishing the connection between them.
 
- \item \underline{Stream module} - responsible to deliver and receive multimedia content from the \emph{Client}.
+ \item \textbf{Stream module} - responsible to deliver and receive multimedia content from the \emph{Client} using \ac{WebRTC}.
 
- \item \underline{Storage module} - provides two main functionalities: store the model information and media recorded.
+ \item \textbf{Storage module} - provides two main functionalities: store the model information and media recorded.
 This is the single module responsible for persistent storage.
@@
 
- \item \underline{Client module} - responsible for the interaction with the user.
+ \item \textbf{Client module} - responsible for the interaction with the user.
 
@@
 %RP também podes fazer um diagrama com vários utilizadores e mostrar quais os streams que vão de e para cada um (mostrar a mistura de som, video, etc).
-
-% Each room will be associated to multiple Media-Types, interactive and non-interactive, discrete and continuous.
-This media types can be structured with chains of events.
-Some components of the user interface may not be synchronized to all participants, for example an help or suggestion window.
 
@@
 To this set of modules, we often call \emph{backend}.
+
+	\subsection{Security and authorization}
 
@@
 
+\subsection{Client connections}
+	Although the delegation of processing work to clients can improve our system's scalability, we are concerned about using the least resources possible on the client side, as huge resource consumption may drain battery very fast or may even be impossible to run on mobile devices.
+We are aware that streaming video from clients is already a very intensive task which we cannot avoid but can improve by delegating the most intensive tasks to our servers.
+
+	\begin{figure}[!htb]
+	\centering
+	\includegraphics[width=0.4\textwidth]{figures/central.pdf}
+	\caption{Centralized Connections}
+	\label{fig:central}
+	\end{figure}
+
+	In this context, each client must only have one \emph{PeerConnection} to our streaming server and content shown to them is changed on demand either being it an individual or a composite view.
+This approach is centralized, as seen on figure \ref{fig:central}.
+
+	The same can be applied to instant message delivery, each client must have only one \emph{WebSocket} connection to the application server which consequently relays the messages to other users.
+
+	Relaying instant messages from clients through the our web servers is easier if all clients are connected to the same server because all messages can be directly delivered without sending messages across multiple web servers.
+
+	In the context of this thesis, we will not implement sending messages across web servers but, in order to allow our system to scale, we will consider that all conference participants are connected to the same server and our system is scaled by having conference rooms distributed across different servers.
+
+\subsection{Software choices}
+
 Furthermore, we have taken into account the compatibility between the streaming server, database and the operation transformation solutions, in order choose the appropriate framework to implement our web server.
@@
 One of our criteria for choosing the web framework is the ability to follow the \ac{MVC} paradigm which can help us to organize our code.
-Without any strong reason we have decided to implement our web server with the \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java}.
-%RP nem sequer por estares mais familiarizado com java ou a playframework? Ou porque querias aprender? Alguma razão há-de ter havido
+We have decided to implement our web server with the \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java} because of our previous experience with it.
 
@@
 \multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
-\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{~~~~~ot.js~~~~~} & \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} & \multicolumn{1}{c|}{adapter.js} &   \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} &\\[12pt]\hline
+\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{~~~~~ot.js~~~~~} & \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} & \multicolumn{1}{c|}{adapter.js} &   \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} &\\[12pt]\hline
 \multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
@@
 
-The synchronization between multimedia elements will be performed through chains of \emph{JavasSript} events or by specifying the interval of time which time content must be visible.
+With respect to displaying content, the synchronization between multimedia elements will be performed through chains of \emph{JavasSript} events or by specifying the interval of time which time content must be visible.
 Other animations can be implemented with \ac{SVG} embedded on \ac{HTML}.

### 2016-05-08 tese2/architecture.tex
[RFT] improving architecture

@@
 \label{chapter:architecture}
-%Esta secção deve englobar:
-
-%\begin{itemize}
-%    \item Uma análise de requisitos, onde devem “abrir” portas para as principais
-%        características da vossa arquitectura.
-%    \item Uma descrição de alto nível da arquitectura do vosso sistema. Esta descrição
-%        pode ser:
-%        \begin{itemize}
-%            \item A arquitectura de rede, explicando os principais componentes e as
-%                funções que executam.
-%            \item A arquitectura de SW da aplicação, explicando os principais
-%                componentes e as funções que executam.
-%        \end{itemize}
-%    \item Descrições detalhadas de todos os componentes da vossa arquitectura.
-%    \item As escolhas que efectuarem em termos de arquitectura global e de detalhe,
-%        devem ser justificadas, face aos requisitos que identificaram.
-%\end{itemize}
-
-%Sempre que possível, ilustrem a arquitectura com figuras, que descrevam
-%sucintamente o modelo como na Figure~\ref{fig:logoCNM}.
-
-%\begin{figure}[!htb]
-%  \centering
-%  \includegraphics[width=0.25\textwidth]{Figures/logoCNM.png}
-%  \caption[Caption for figure in TOC]{Caption for figure.}
-%  \label{fig:logoCNM}
-%\end{figure}
 
@@
 \centering
-	\includegraphics[width=0.4\textwidth]{figures/central.pdf}
-	\caption{Centralized Connections}
+\begin{minipage}[b]{0.45\linewidth}
+	\centering
+	\includegraphics[width=\textwidth]{figures/ccomposite.pdf}
+	\caption{Centralized connections}
 	\label{fig:central}
+\end{minipage}
+\quad
+\begin{minipage}[b]{0.45\linewidth}
+	\centering
+	\includegraphics[width=\textwidth]{figures/dcomposite.pdf}
+	\caption{Peer-to-Peer connections}
+	\label{fig:p2p}
+\end{minipage}
 \end{figure}
@@
 This approach is centralized, as seen on figure \ref{fig:central}.
+Otherwise clients could follow a peer-to-peer connection, as seen on figure \ref{fig:p2p}, which would result on maintaining more connections and performing the composition of videos on client side.
 
-	The same can be applied to instant message delivery, each client must have only one \emph{WebSocket} connection to the application server which consequently relays the messages to other users.
+	The composition of streams on client side is performed by receiving streams with the best quality possible but, due to undersizing the video of clients into a smaller region, this would result on wasting bandwidth on a quality that is not needed.
+
+	Although the peer-to-peer approach could be used on our system, we conclude that we need to record the video on our streaming server because web clients have a very limited storage and peer disconnections may result on recorded video loss.
+
+	The same can be concluded to instant message delivery, each client must have only one \emph{WebSocket} connection to the application server which consequently relays the messages to other users.
 

### 2016-05-08 tese2/evaluation.tex
[ADD] chapter summary
[RFT] scaling options, on tests

@@
 
-With this results we conclude that if we want to scale our solution's storage using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \emph{MongoDB} should be installed in the same machine because the loopback interface can handle bigger transfer rates than the remaining network interfaces.
-For the same reason another possible configuration for scaling the storage is to install both \emph{Kurento Repository} and \ac{KMS} on the same machine.
+With this results we conclude that if we want to scale our solution's storage using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \emph{MongoDB} should be installed in the same machine because the loopback interface can handle bigger transfer rates than the remaining network interfaces as seen on figure \ref{fig:scale1}.
+For the same reason another possible configuration for scaling the storage is to install both \emph{Kurento Repository} and \ac{KMS} on the same machine as seen on figure \ref{fig:scale2}.
+Installing the repository on the same machine as a database node does not ensure that recorded videos are stored in the same machine, for this reason we would prefer installing \ac{KMS} and \emph{Kurento Repository} on the same machine.
+
+\begin{figure}[!htb]
+\centering
+\begin{minipage}[b]{0.45\linewidth}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/scale1.pdf}
+  \caption{Repository and database in the same machine}
+  \label{fig:scale1}
+\end{minipage}
+\quad
+\begin{minipage}[b]{0.45\linewidth}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figures/scale2.pdf}
+  \caption{Repository and KMS in the same machine}
+  \label{fig:scale2}
+\end{minipage}
+\end{figure}
 
@@
 
-In this chapter we describe the test scenarios that we have created in order to evaluate our system.
+In this chapter we described the test scenarios that we have created in order to evaluate our system.
 

### 2016-05-09 tese2/10_related_work_hypermedia.tex
[ADD] template2
[RFT] rewriten model

--- 2016-03-12 tese1/10_related_work_hypermedia.tex
@@
-\subsection{Hypermedia: more than words, more than images}
+\section{Hypermedia: more than words, more than images}
 \label{hypermedia}
@@
 
-  \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Metainformation}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.
+  \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Meta Information}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.
 
@@
 
-  \item The \textbf{Metainformation} module contains elements and attributes that allows describing the \ac{SMIL} document.
+  \item The \textbf{Meta Information} module contains elements and attributes that allows describing the \ac{SMIL} document.
 For example, this module could be used to define movie details such as category, director, writers and cast.
@@
   \item The \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content.
-This module defines a \emph{head} element that contains non temporal behavior information defined by  \emph{Metainformation}, \emph{Layout} and \emph{Content Control} modules.
+This module defines a \emph{head} element that contains non temporal behavior information defined by  \emph{Meta Information}, \emph{Layout} and \emph{Content Control} modules.
 This module also defines the \emph{body} element, where all temporal related module information is contained.
@@
 
-%RP explicar melhor. Não vejo como low-latency pode ser mau
-  %HR troquei-me todo
-  With the emergence of \ac{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format.
+  With the emergence of \ac{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple \emph{codecs}, audio and subtitles in \ac{WebVTT} format.
 Another important tag is \emph{canvas} that allows drawing graphics with \emph{JavaScript} on a rectangle within a web page.
@@
 With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
- %XXX
-  %RP porque escolheste falar aogra de webgl se não falas antes?
-%HR já troquei, o WebGL usa o canvas do HTML5, talvez seja este o local mais adequado
 
@@
 Competitor's players, at that time were focused on playing video and audio, while \emph{flash} had vector graphics and focused on streaming \emph{on-demand} video across multiple platforms.
-\emph{VP6} was their choice of video codec, providing half the video size (in Byte) for the same quality and providing video quality adjusted to \emph{Internet} connection latency.
+\emph{VP6} was their choice of video \emph{codec}, providing half the video size (in Byte) for the same quality and providing video quality adjusted to \emph{Internet} connection latency.
   In 2010, \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
-  %RP ``was'' quando? Em 2010? Ontem?
-  %HR exacto.
 
@@
 Like \ac{WebRTC}, \emph{flash} players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
-  %RP as vezes fazes \emph{flash} e outras vezes não. Convém ser coerente.
 

### 2016-05-09 tese2/13_implementation_model.tex
[RFT] rewriten model

@@
 %RP antes desta secção fala um introdução ao capítulo a explicar o que vai ser apresentado.
+%HR done
 
@@
 During the course of this project, we had to redesign the model more than once in order to support new features.
-%RP mudei ``most important'' para critical
 
@@
 Que tens utilizadores, grupos, mensagens, chats, convites, etc...
+%HR não meti na arquitectura mas adicionei a linha seguinte
+
+In order to offer all the functionalities that we promise, some information about objects must be persistent such as users, groups,relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.
 
@@
 Needless to say that, by following this approach we don't need to create the collections with a well predefined schema.
-In fact we don't need to define it at all.
-Although, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure.
+In fact we don't need to define it at all but, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure as we will present in this section.
     %RP a última frase é confusa. O que queres dizer? Reescreve
+    %HR está melhor?
 
@@
 
-    In respect to attributes nullability, \emph{MongoDB} does not enforce a document's attribute to have a non nullable value, although, for sake of good functionality, we perform those constraints validation programmatically and as such we also represent them in our schema.
-    %RP nullable ou null?
+    In respect to attributes nullability, \emph{MongoDB} does not enforce a document's attribute to have a not null value, although, for sake of good functionality, we perform those constraints validation programmatically and as such we also represent them in our schema.
 
@@
 
-The identification number by itself is not sufficient to identify an object, as objects from different types can have the same identification number.
+The owner's identification number by itself is not sufficient to identify an object, as objects from different types can have the same identification number.
 In order to solve this problem, when an object is created, its correspondent data must contain the owner's object type.
 %RP mas o _id é apresentado com sendo uma chave primária na figura 4.2, logo única!
+%HR "The owner's identification number" corriji, não me referia ao "_id"
 
@@
 
+A conference room, which is the environment where users communicate among their selves, is represented persistently by a group of participating entities (in general users but our system allow other types of entities).
+
 A group is composed by an \emph{id}, \emph{inviteToken} and a \emph{visibility} as shown on table \ref{table:group}.
@@
 Entities that have a membership with a group can create more memberships by sharing an invite token or by specifying new group members.
+This invite token is used to create an invite \ac{URL} that if shared with other users allow them to join the group.
+Invite tokens can be deleted or regenerated with a different value.
 %RP isto quer dizer que quem tem o token se pode juntar ao grupo? Explicar. O que impede alguém de o partihar com outros?
+%HR done
 
@@
 
+In order to query for recent messages for a given target, we use the oldest message's \emph{\_id}, which is sequential, in order to find messages with a newer \emph{\_id}.
+
 In order to improve performance of queries over \emph{Relation} collection we have created indexes on the \emph{target} attribute and also on the pair composed by \emph{target} and \emph{\_id} attributes for finding and sorting the messages received by an entity more efficiently.
+
 %RP pela justificação não consigo perceber para que serve a chave no par target e _id
+%HR in order to find recent messages (não tinha explicado)
 
@@
 
+In the point o view of a user, a time annotation is just a simple way to associate a time stamp to a topic in order to make it easier to find.
+On the other hand, the hyper-content is more complete than time annotations, except they are not visible on the time-line.
+In addition to time annotations, hyper-contents have a duration and content that can be superimposed to video.
+
 %RP explica melhor a diferença entre HyperContent e Time Annotation do ponto de vista do utilizador / funcionalidade proporcionada.
+%HR done
 
@@
 It also contains the media's \ac{URL}, a reference to a group, an owner identification number and the correspondent \emph{WebSocket} session id.
+
+    The media's \ac{URL} is the location where \emph{Kurento Repository} stores one chunk of video (including audio) which is used to playback on users demand.
+
     %RP ainda não tenho forma de perceber o que é o websocket session id. Nunca falaste de websockets antes, quanto mais disto! Nem do media URL, é gravado ou webrtc (dispositivo do utilizador)?
@@
 
+In respect to find which chunks to play given a specified time (which is the current time in a user's perspective), we must select the chunk with an ending instant immediately after the current time, if the chunk's beginning is placed before the current time we have to play it instantaneously otherwise the chunk's beginning will occur after the current time and therefore we have to schedule the playback.
+
+On the other hand, we use the fact that \emph{\_id}'s are sequential in order to determine adjacent chunks, for example, if we need to find the next chunk to play, we have to sort all elements by \emph{\_id} (which on \emph{MlongoDB} is already sorted by default) and select a chunk with an \emph{\_id} immediately after the current one.
+
 In order to improve the performance of searching recording chunks and calculate which chunks are going to be reproduced, we have created indexes over the following sequences of attributes: 
@@
     \item{\emph{\{groupId,start,end\}} $\rightarrow$ for finding all available chunks for a given time} 
-    \item{\emph{\{groupId,sessionId,end\}} $\rightarrow$ for finding intersecting chunks with a given time within a session}
-    \item{\emph{\{groupId,owner,end\}} $\rightarrow$ for finding intersecting chunks with a given time within a user (for a different session) or group}
+    \item{\emph{\{groupId,sessionId,end\}} $\rightarrow$ for finding chunks that ends after a given time within a session}
+    \item{\emph{\{groupId,owner,end\}} $\rightarrow$ for finding chunks that ends after a given time within a user (for a different session) or group}
     \item{\emph{\{groupId,sessionId,\_id\}} $\rightarrow$ for finding chunks that follow a given \emph{id} within a session}
@@
 %RP os _id não são únicos? Para que servem os index que os incluem?
+%HR done, for finding adjacent chunks
+
 %RP para a intercepção não precisaste de usar o start, apenas o end?
+%HR exactly, if there is no intersection taking into account the "end" at least I can schedulle the next player (after the end)
 
@@
 %RP olha outra coisa nova! Ainda não falaste de conference rooms!
+%HR "conference room" dá-se no contexto de um "group", era para não estar sempre a repetir. Qualquer das formas adicionei esse pormenor na seccao do grupo.

### 2016-05-09 tese2/implementation.tex
[ADD] change architecture
[ADD] deployment and test description
[ADD] chapter summary
[RFT] rewriten model

@@
 
+In this chapter we are going to present our implementation choices, the difficulties that we have faced and the strategies available to overcome them.
+
+To this end, we are going to present and explain our database model, how we have improved it and how this model can help us reach our goals.
+
+Subsequently, we are going to describe our signaling protocol which will make it possible to use the \ac{WebRTC}'s functionalities in order to implement our system's streaming features.
+
+With the model and signaling protocol defined we are prepared to show the multiple approaches we have followed in order to implement stream recording, their drawbacks and our choices.
+
+Then are going to describe how we show hyper-content to users, namely the multiple ways to create hyper-content and the algorithm behind the user interface synchronization, the security flaws of our content displaying mechanism and a possible solution to overcome vulnerabilities.
+
+Not less important we also are going to describe how we have implemented our application time-line and the functionalities above it such as time manipulation and annotations management.
+Among other different functionalities we also are going to describe how we perform stream composition, synchronization of our collaborative text editor and implementation of our instant messenger.
+
+Lastly, we are going to describe the deployment of our solution in respect to the hardware that we have used and the software installed onto it.
+
 \input{13_implementation_model.tex}
@@
 \input{15_implementation_recording.tex}
+\input{16_implementation_content.tex}
+\input{17_implementation_deploy.tex}
+\input{18_implementation_summary.tex}

### 2016-05-10 tese2/14_implementation_signaling.tex
Reviewing implementation.

@@
 \section{Signaling Protocol}
+%RP devias começar com uma introdução a explicar que tiveste de desenvolver um protocolo para coordenar a interacção dos clientes com o servidor e entre clientes (?).
+%RP explicar quando é usado (começar chamada, terminar, juntar, chat(?), edição colaborativa (?), etc).
 
-The signaling protocol plays a fundamental role on our solution, after the web server validates the user access,
-it allows the users to directly connect to \ac{KMS} which is placed in a private network and let the server and users negotiate media types and encoding information to use during the conversation.
+The signaling protocol plays a fundamental role on our solution.
+After the web server validates the user access, it allows the users to directly connect to the \ac{KMS}, which is placed in a private network, and lets the server and users negotiate media types and encoding information to use during the conversation.
+%RP o it em ``it allows'' é quem? webserver ou signaling protocol?
 
-With respect to the signaling protocol, it could be implemented using \ac{HTTP} messages but they would transport extra information such as \ac{HTTP} headers and would follow a request-response signaling mechanism which would not be the best option, as multiple ice candidates can arrive at any time.
-Using a \ac{REST} \ac{API} over \ac{HTTP} for signaling and allowing other methods to be called at the same time would lead to open multiple \ac{TCP} connections at the same time.
+We considered implementing the signaling protocol using \ac{HTTP} messages.
+However, they would transport extra information such as \ac{HTTP} headers and would follow a request-response signaling mechanism which would not be the best option, as multiple ICE candidates can arrive at any time.
+Using a \ac{REST} \ac{API} over \ac{HTTP} for signaling and allowing other methods to be called at the same time would lead to opening multiple \ac{TCP} connections at the same time.
 
-Instead if we create a \emph{WebSocket} \ac{API} we only need one \ac{TCP} connection and at the same provide bi-directional communications without additional headers.
+Instead, if we create a \emph{WebSocket} \ac{API}, we only need one \ac{TCP} connection and, at the same, provide bi-directional communications without additional headers.
 Moreover, using \emph{WebSocket} allows the server to send messages to the client without the client having to request it, which can be slightly faster than using \ac{HTTP}.
+%RP em vez de slighhtly faster não queres antes enfatizar que pode ser antes o servidor a iniciar a comunicação, de forma assincrona?
 
-Our signaling protocol consists in send and receiving \ac{JSON} formated messages like listing \ref{lst:msgtype} over WebSockets by both the server and the client.
+Our signaling protocol consists of sending and receiving \ac{JSON} formated messages, e.g. listing \ref{lst:msgtype}, over WebSockets by both the server and the client.
 
@@
 
-When a user enters in a group conference, after the page is completely loaded it is created a WebSocket that maintains a connection with our web servers. 
-Before creating the web socket, we must identify the user and check if he has permissions to participate in the conference.
-The user identification is done by retrieving the session id from the cookie provided by the user through the \ac{HTTP} headers then we retrieve all the information need from database in order to check if the user has permissions to join the conference room.
-It's important to save the user identification before the socket is created because after the the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455} the \ac{HTTP} context is lost.
+When a user enters a group conference, after the page is completely loaded, a WebSocket is created to maintain a connection with our web servers. 
+But before creating the web socket, we must identify the user and check if he has permissions to participate in the conference.
+The user identification is done by retrieving the session id from the cookie provided by the user through the \ac{HTTP} headers.
+%RP não é bem o user que fornece o cookie. é o user-agent (browser). Convém ser preciso.
+The Web server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room.
+It is important to save the user identification before the socket is created because, after the handshake is performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \ac{HTTP} context is lost.
+%RP socket -> websocket?
 
-When the connection is established between the server and the client, a \emph{PeerConnection} is created on the client and immediately after an \ac{WebRTC} endpoint on the server specifying a possible set of \ac{ICE} servers to connect.
+When the connection is established between the server and the client, a \emph{PeerConnection} is created on the client and immediately after, an \ac{WebRTC} endpoint is created on the server, specifying a possible set of \ac{ICE} servers to connect.
+%RP Usa Application server e não apenas server. Há muitos servers.
 
-The user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server.
+At this stage, the Web Client's user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server.
 If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2015).} may ask to install a plug-in if the browser doesn't support screen sharing.
-
-If the user decides to share either from camera of screen, \emph{getUserMedia} is called with the correspondent constraints in order to obtain a local stream.
-Use the constraints presented on listing \ref{lst:constraints}.
+%RP web app user? OU apenas client user?
+%RP foot note com exemplos de browsers que aceitem screen sharing?
+If the user decides to share either from camera or screen, \emph{getUserMedia} is called with the correspondent constraints in order to obtain a local stream.
+We use the constraints presented on listing \ref{lst:constraints}.
 
@@
 
-A pop-up is raised in order to ask the user to give permission to share those resources.
-If the resources are shared successfully the user creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through WebSockets.
+If the user decides to share his camera or screen, a pop-up is raised in order to ask the user to give permission to share those resources.
+If the resources are shared successfully, the user creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the WebSocket to the Application Server.
+%RP usas ``user'' duas vezes, uma para te referires ao utilizador e outra ao browser (user agent).
 If the resources are not shared or the user specified to receive stream only, an offer is created specifying the constraints for receiving only video and audio.
+%RP offer -> SDP offer?
 
@@
 		"type":"offer",
-		"sdp":<sdp>	// this is huge
+		"sdp":<sdp>	// omitted for brevity
 	}
@@
 The \emph{local session description} contains the session identifier, codecs, containers, transport protocols and ports used per media type.
-The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} doesn't need to mix, record and analyze streams coming from the user \ac{WebRTC} endpoint.
+The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} doesn't need to mix, record nor analyze streams coming from the user's \ac{WebRTC} endpoint.
 
-The server receive and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint.
-Then a \emph{local session description} like listing \ref{lst:sig02} is created on the server and sent back to the client, after that the server tries to gather \ac{ICE} candidates.
+The server receives and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint.
+Then a \emph{local session description}, like the one presented in Listing \ref{lst:sig02}, is created on the server and sent back to the client, after that the server tries to gather \ac{ICE} candidates.
+%RP listing -> Listing
 
@@
 		"type":"answer",
-		"sdp":<sdp>	// this is huge
+		"sdp":<sdp>	// omitted for brevity
 	}
@@
 The client receives the server answer, sets the \emph{remote session description} and gets the ice candidates from the \ac{ICE} server.
+%RP tens ``ice'', ``\ac{ICE}''. Acho que também já vi ``ICE''. Convém ser sempre igual!
 
@@
 The same is done on the server which receives the \ac{ICE} candidates like on listing \ref{lst:sig03} from the client and propagates them to \ac{KMS}.
+%RP A figura 4.1 precisa de vir antes deste texto todo, para ser mais fácil de perceber.
+%RP também tens de explicar porque razão escolheste colocar o KMS numa rede privada. Se está numa rede privada, usas NAT ou um proxy (socket) para estabelecer a ligação?
 
@@
 An \ac{ICE} candidate contains an \ac{IP}, port, used transport protocol and an attribute named \emph{sdpMLineIndex} that is used for mapping to the \emph{remote session description} media type.
-
-In addition both intervenients test the connectivity of each \ac{ICE} candidate.
-When a connection is established the user and server start to interchange stream data but other \ac{ICE} candidates may arrive with better connections, when that happens the connection changes seamlessly.
+In addition, both intervenients test the connectivity of each \ac{ICE} candidate.
+When a connection is established, the user and server start to interchange stream data but other \ac{ICE} candidates may arrive with better connections.
+When that happens, the connection changes seamlessly.
 
@@
 \end{figure}
+
+%RP faz sentido haver dois ICE? Se queres colocar 2 dá-lhes nomes diferentes (nem que seja cliente e servidor ou 1 e 2).
+
+%RP não falas nada sobre chats, etc! Não há signalling para isso?

### 2016-05-10 tese2/15_implementation_recording.tex
Reviewing implementation.

@@
 In this section we describe our approaches for the stream recording implementation, namely recording on the client side and server side.
+%RP Innitially we experimented with local recording. Even thought it got it to work, it consumed to much bandwith. We then tried ....  In this section, we describe the several techniques evaluated.
 
 \subsection{Client side recording}
-	In order to record used shared streams, our first approach consisted on recording video and audio streams into blobs of limited duration.
+In order to record used shared streams, our first approach consisted in locally recording video and audio streams into blobs of limited duration.
 Because each user was recording directly from their shared local resources, we achieved the best stream quality.
+%RP used -> user?
 
-	After recording, each block was uploaded to our server through \ac{HTTP} and saved into the file system, the block meta-data was created and inserted on our database.
+After recording, each block was uploaded to our server through \ac{HTTP}, saved into the server's file system, the block meta-data was created and inserted on our database.
 Each block meta-data contained the file's location for the recorded block, the starting date, duration, user identification and group identification.
-When the file was completely saved the meta-data was advertised to the remaining users.
-This meta-data was simply used to refresh the user interface, the meta-data was completely discard after that because an huge amount of small blocks meta-data would use more and more memory as time went by.
+%RP referenciar o modelo na secção 3.xx?
+When the file was completely saved, the meta-data was advertised to the remaining users.
+This meta-data was simply used to refresh the user interface and was completely discarded after that, because an huge amount of small blocks' meta-data would use more and more memory as time went by.
 
-	The process to play a video was quite simple, the user specified which user and date was intended to play, the server calculated the intersection between the requested date and the block bounds and returned the file to the client.
+	The process to play a video was quite simple.
+The user specified which user and date he intended to play, the server calculated the intersection between the requested date and the block bounds and returned the file to the client.
+	Although this idea was fairly simple, we couldn't achieve seamless sequential block switching.
+Downloading the video file always took a noticeable amount of time.
+To solve this problem, while we were playing a block, the next one was downloading in parallel, so when the current block finished playing we would have an available block to play.
 
-	Although this idea was fairly simple, we couldn't achieve seamless sequential block switching.
-Downloading the video file always toke a noticeable amount of time.
-To solve this problem while we were playing a block, the next one was downloading in parallel so when the current block finished playing we would have an available block to play.
+        With this technique, block switching became more acceptable, but switching the \ac{URL} always produced a flash.
+We solved this problem by having two layers and changing the \ac{URL} on the back layer.
+When the front layer froze on the last frame, we changed the back layer to the front.
 
-	Block switching became more acceptable, but switching the \ac{URL} always produced a flash.
-We solved this problem by having two layers and set the \ac{URL} to the back layer, the front layer frozen with the last frame, then we changed the back layer to front.
-
-	After we implemented our recording solution using this approach, we tested locally and remotely.
-For remote connections we observed a fairly high bandwidth usage mainly because blocks were both sent and received at maximum quality.
+	After we implemented this recording solution using this approach, we tested locally (client and server on the same machine) and remotely.
+For remote connections we observed a fairly high bandwidth usage, mainly because blocks were both sent and received at maximum quality.
+        %RP We deemed this high bandwidth usage to be a significant bottleneck on our application and decided to explore other solutions.
 
 \subsection{Server side recording to file system}
+%RP começar por explicar razão para usar KMS: permite gravar e juntar várias streams numa só, o que torna a aplicação muito mais escalável.
 
-	Before recording any type of stream we had to analyze the user media offer in order to check if a video as really being received by \ac{KMS}, otherwise if we would not verify the user's offer, the recording video would be black.
+Before recording any type of stream, we had to analyze the user media offer in order to check if a video was really being received by \ac{KMS}, otherwise if we did not verify the user's offer, the recording video would be black.
+%RP Entras logo em grandes pormenores. Falta aqui uma introdução.
 
 The streaming content received on \ac{KMS} was already compressed due to \ac{WebRTC}'s exchanged quality of service metrics data.
-As a direct consequence our recorder solution used on most cases less disk space per block, but would never use more storage than client side recording.
+As a direct consequence, our recorder solution used on most cases less disk space per block, but would never use more storage than client side recording.
 \ac{KMS} allows recording files using \emph{webm} and \emph{mp4} containers.
+%RP escolheste qual?
 
 	With server side recording, the user would maintain always the same stream \ac{URL} even if it is playing real-time video or reproducing recorded video.
-When a user desires to play recorded video, a \emph{webSocket} message is sent specifying the time and the intended user id, the group identification is not sent because it is already associated to the \emph{webSocket}.
-The server performs the same calculations in order to find a block that intersects the requested time, plays it and when finished the next part is automatically played without the user intervention.
+It is \ac{KMS} that sends different content through that stream.
+When a user desires to play recorded video, a \emph{webSocket} message is sent specifying the time and the intended user id.
+The group identification is not sent because it is already associated to the \emph{webSocket}.
+The server performs the same calculations in order to find a block that intersects the requested time, plays it and when finished, the next part is automatically played without the user intervention.
 
-	We observed differences in image quality when switching parts, that was even noticeable if we set a short block duration.
-We also noticed a small gap on audio when switching blocks but it was acceptable and speech recognition was not very affected.
+	We still observed differences in image quality when switching parts.
+That was even noticeable if we set a short block duration.
+We also noticed a small gap on audio when switching blocks but it was acceptable and speech understanding was not very affected.
 
-	Back when we were implementing our solution, \ac{KMS} had not support for seeking videos, which meant that blocks would would always start playing from their beginning.
+	When we started implementing our solution, \ac{KMS} had no support for seeking videos, which meant that blocks would would always start playing from their beginning.
 This lead to a theoretic playing time error that can be at most half the duration of a block.
-In practice the maximum playing time error coincides with the block duration because we perform the intersection between the requested date and the block but we could decrease the error by half if we calculate the intersection with the requested time plus half the duration.
+In practice, the maximum playing time error coincides with the block duration because we perform the intersection between the requested date and the block.
+But we could decrease the error by half if we calculate the intersection with the requested time plus half the duration.
+        %RP já tinhamos falado sobre isto.
+O erro é no máximo o tamanho do bloco pois queres garantir que a pessoa ve sempre o que solicita.
+Assim começas sempre no bloco que fica antes e nunca no que fica depois.
+
+        %RP Figura tem de ser apresentada no texto e explicada. O que significam as letras?
 
@@
 
-	For playing video with an higher velocity we used \emph{ffmpeg}\footnote{\url{https://www.ffmpeg.org/}(accessed: 17 March 2016)} to convert the block into a new video with the desired velocity and seek time.
-Because the media duration is known, when the video started to being convert the headers located at the beginning of the file were already written and that made it possible to stream while converting.
-
+        \ac{KMS} does not support playing video at higher speeds, a useful feature that enables users to watch a video in less time than its duration.
+	For playing video with an higher velocity, we used \emph{ffmpeg}\footnote{\url{https://www.ffmpeg.org/}(accessed: 17 March 2016)} to convert the block into a new video with the desired velocity and seek time.
+Because the media duration is known, when the video started to being converted the headers located at the beginning of the file were already written and that made it possible to stream while converting.
 	%{\color{red} [TALK ABOUT COMPLEXITY OF DECODE VS DECODE+MANIPULATE+ENCODE]}
-
 	Although we implemented a solution that worked, we immediately noticed that \emph{ffmpeg} would take some time to initialize and that lead to pauses between switching parts.
@@
 	Later the \emph{Kurento} team released a version with support for seeking videos but we had to suspend the implementation of the fast forwarding feature as currently \ac{KMS} is not supporting that.
+        %RP fiquei sem perceber. O método com o ffmpeg deixou de funcionar? Ou desististe devido ao outros problemas (delay, qualidade, cpu)?
 
-	We could implement fast forwarding without real time conversion by creating multiple versions of the same video with different velocities after the recording of a block.
-When a user needed to play he would also need to specify one of the available velocities.
-We did not followed this approach has it would require a bigger disk space usage.
+	We could have implemented fast forwarding without real time conversion by creating multiple versions of the same video with different playback speeds after the recording of a block.
+When a user needed to play, he would also need to specify one of the available speed.
+We did not followed this approach has it would require larger disk space usage.
 
@@
 	One of our concerns during the development of our solution was the storage scalability.
-Saving files directly into file system would require an extra effort to distribute and replicate files among servers.
-For that purpose \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB}.
+Saving files directly into the file system would require an extra effort to distribute and replicate files among servers.
+For that purpose, the \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB}.
 
@@
 The same is true for recording, but because the file headers are in the beginning and the file is written until is stops, the headers don't contain the necessary information for seeking the file.
-
 	Although we gain with scalability with this approach, we lose access over the file for changing it to fit our needs, namely for using \emph{ffmpeg} or other video manipulation tool.
 
-	Although we did not implemented recorded file seeking, that could be achieved by waiting for full file recording and then proceed to database insertion with the correct headers.
-Another approach would be the specification of the file duration before recording so the correct file headers could be written a priori.
-Both approaches were not possible to implement using just the \emph{Kurento} clients, we would need change the source code of \emph{Kurento} in order to add those new features.
+	Although we did not implement recorded file seeking, that could be achieved by waiting for full file recording and then proceed to database re-insertion with the correct headers.
+Another approach would be the specification of the file duration before recording, so the correct file headers could be written a priori.
+Both approaches were not possible to implement using just the \emph{Kurento} clients.
+We would need change the source code of \emph{Kurento} in order to add those new features.
+        %RP explicar razões para não o fazer: perder upgrades, tornar instalação mais simples, maior complexidade manutenção app, etc.

### 2016-05-11 tese2/14_implementation_signaling.tex
[RFT] rewriten signalingimpl

@@
 \section{Signaling Protocol}
+
+As we have already mentioned on our related work, \ac{WebRTC} does not implement the signaling protocol, which is used to establish connections between peers.
+This connection may be direct using \ac{STUN} or relayed by a \ac{TURN} server in case of direct communications are not possible.
+
+The signaling protocol plays a fundamental role on our solution.
+Taking into account the choices we have based on our related work, we have decided to implement our own signaling protocol in order to use \ac{WebRTC} on our solution.
+
+Although we have mentioned that the signaling protocol is used to establish connections between peers, on our system our media server (\ac{KMS}) is a peer that receives video streams and sends to its connected clients.
+
+In a general view, the signaling protocol is used to share connection and media properties between two peers.
+In order to ease the understanding of our signaling protocol, we have created a sequence diagram that represented on figure \ref{fig:signaling2}\footnote{Although two \ac{ICE} servers are shown they are, in fact, the same.
+Showing just one \ac{ICE} server would be difficult to draw.}.
 %RP devias começar com uma introdução a explicar que tiveste de desenvolver um protocolo para coordenar a interacção dos clientes com o servidor e entre clientes (?).
 %RP explicar quando é usado (começar chamada, terminar, juntar, chat(?), edição colaborativa (?), etc).
+%HR done
 
-The signaling protocol plays a fundamental role on our solution.
-After the web server validates the user access, it allows the users to directly connect to the \ac{KMS}, which is placed in a private network, and lets the server and users negotiate media types and encoding information to use during the conversation.
+\begin{figure}[!htb]
+    \centering
+    \begin{subfigure}{}
+    	\includegraphics[width=0.9\textwidth]{figures/signaling}
+    \end{subfigure}
+    \caption{Signaling sequence diagram}
+    \label{fig:signaling2}
+\end{figure}
+
+%RP faz sentido haver dois ICE? Se queres colocar 2 dá-lhes nomes diferentes (nem que seja cliente e servidor ou 1 e 2).
+
+After the web application server validates the user access, the signaling protocol allows the users to directly connect to the \ac{KMS}, which is placed in a private network, and lets the application server and users negotiate media types and encoding information to use during the conversation.
 %RP o it em ``it allows'' é quem? webserver ou signaling protocol?
-
+%HR done
 We considered implementing the signaling protocol using \ac{HTTP} messages.
@@
 Instead, if we create a \emph{WebSocket} \ac{API}, we only need one \ac{TCP} connection and, at the same, provide bi-directional communications without additional headers.
-Moreover, using \emph{WebSocket} allows the server to send messages to the client without the client having to request it, which can be slightly faster than using \ac{HTTP}.
+Moreover, using \emph{WebSocket} allows the application server to send messages asynchronously to the client without the client having to request it, which can be slightly faster than using \ac{HTTP} which is based on a request followed by a response.
 %RP em vez de slighhtly faster não queres antes enfatizar que pode ser antes o servidor a iniciar a comunicação, de forma assincrona?
-
-Our signaling protocol consists of sending and receiving \ac{JSON} formated messages, e.g. listing \ref{lst:msgtype}, over WebSockets by both the server and the client.
+%HR done
+Our signaling protocol consists of sending and receiving \ac{JSON} formated messages, e.g. listing \ref{lst:msgtype}, over WebSockets by both the application server and the client.
 
@@
 But before creating the web socket, we must identify the user and check if he has permissions to participate in the conference.
-The user identification is done by retrieving the session id from the cookie provided by the user through the \ac{HTTP} headers.
+The user identification is done by retrieving the session id from the cookie provided by the user-agent (web browser) through the \ac{HTTP} headers.
 %RP não é bem o user que fornece o cookie. é o user-agent (browser). Convém ser preciso.
-The Web server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room.
-It is important to save the user identification before the socket is created because, after the handshake is performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \ac{HTTP} context is lost.
+%HR done
+The Web application server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room.
+It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake is performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \ac{HTTP} context is lost.
 %RP socket -> websocket?
 
-When the connection is established between the server and the client, a \emph{PeerConnection} is created on the client and immediately after, an \ac{WebRTC} endpoint is created on the server, specifying a possible set of \ac{ICE} servers to connect.
+When the connection is established between the application server and the client, a \emph{PeerConnection} is created on the client and immediately after, an \ac{WebRTC} endpoint is created on the server, specifying a possible set of \ac{ICE} servers to connect.
 %RP Usa Application server e não apenas server. Há muitos servers.
 
-At this stage, the Web Client's user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server.
-If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2015).} may ask to install a plug-in if the browser doesn't support screen sharing.
+At this stage, the web application's user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server.
+If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2016).} may ask to install a plug-in if the browser doesn't support screen sharing \footnote{\url{http://iswebrtcreadyyet.com/} (Accessed May 11, 2016)}.
 %RP web app user? OU apenas client user?
 %RP foot note com exemplos de browsers que aceitem screen sharing?
+%HR done
 If the user decides to share either from camera or screen, \emph{getUserMedia} is called with the correspondent constraints in order to obtain a local stream.
@@
 If the user decides to share his camera or screen, a pop-up is raised in order to ask the user to give permission to share those resources.
-If the resources are shared successfully, the user creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the WebSocket to the Application Server.
+If the resources are shared successfully, the user agent creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the WebSocket to the Application Server.
 %RP usas ``user'' duas vezes, uma para te referires ao utilizador e outra ao browser (user agent).
-If the resources are not shared or the user specified to receive stream only, an offer is created specifying the constraints for receiving only video and audio.
+If the resources are not shared or the user specified to receive stream only, an \ac{SDP} offer is created specifying the constraints for receiving only video and audio.
 %RP offer -> SDP offer?
@@
 %RP também tens de explicar porque razão escolheste colocar o KMS numa rede privada. Se está numa rede privada, usas NAT ou um proxy (socket) para estabelecer a ligação?
+%HR uso NAT, falei disso na arquitectura, serve para o cliente não aceder directamente ao servidor de streaming e realizar o seu proprio protocolo de sinalizacao (uso indevido por parte de atacantes)
 
@@
 
-\begin{figure}[!htb]
-    \centering
-    \begin{subfigure}{}
-    	\includegraphics[width=0.9\textwidth]{figures/signaling}
-    \end{subfigure}
-    \caption{Signaling sequence diagram}
-\end{figure}
-
-%RP faz sentido haver dois ICE? Se queres colocar 2 dá-lhes nomes diferentes (nem que seja cliente e servidor ou 1 e 2).
-
 %RP não falas nada sobre chats, etc! Não há signalling para isso?
+%HR não, só video e audio, chat é websockets

### 2016-05-11 tese2/architecture.tex
[RFT] rewriten signalingimpl

@@
 
+The placement of our streaming servers inside a \ac{NAT} also has an important role with respect to external misuse prevention.
+Otherwise, placing our streaming servers could allow external clients to perform their own signaling protocol and, as a consequence, use our infrastructure without our consent.
+
 \subsection{Client connections}

### 2016-05-11 tese2/15_implementation_recording.tex
[ADD] almost rewritten recording impl, gonna sleep Zzzz...

@@
 \section{Stream Recording}
-In this section we describe our approaches for the stream recording implementation, namely recording on the client side and server side.
+ Initially we experimented with local recording and synchronizing with our servers.
+Even thought it got it to work, it consumed to much bandwidth.
+We then tried recording on server side.
+In this section, we describe the several techniques evaluated.
+
 %RP Innitially we experimented with local recording. Even thought it got it to work, it consumed to much bandwith. We then tried ....  In this section, we describe the several techniques evaluated.
@@
 %RP referenciar o modelo na secção 3.xx? 
+%HR talvez não, não é o mesmo modelo, foi uma abordagem diferente
+
 When the file was completely saved, the meta-data was advertised to the remaining users.
@@
 For remote connections we observed a fairly high bandwidth usage, mainly because blocks were both sent and received at maximum quality.
+
+	We deemed this high bandwidth usage to be a significant bottleneck on our application and decided to explore other solutions.
+
         %RP We deemed this high bandwidth usage to be a significant bottleneck on our application and decided to explore other solutions.
@@
 \subsection{Server side recording to file system}
+
+In addition to perform as an \ac{MCU}, \ac{KMS} also allow recording individual incoming streams and the composition of streams which can be reproduced later using the existing \emph{Peer Connection} and as a consequence making our application more efficient and scalable.
+
+Without additional configurations, \ac{KMS} allows recording video directly to file system by specifying each file path.
+This approach makes the file management more complex as file paths are relative to one file system within a specific machine.
+If we use this approach with multiple instances of \ac{KMS} in order to scale our solution, we have to associate the machine location to each recording chunk and we also have to take into account replication so that we can provide fault tolerance.
+
 %RP começar por explicar razão para usar KMS: permite gravar e juntar várias streams numa só, o que torna a aplicação muito mais escalável.
+% HR better?
 
@@
 %RP Entras logo em grandes pormenores. Falta aqui uma introdução.
+% HR better?
 
@@
 As a direct consequence, our recorder solution used on most cases less disk space per block, but would never use more storage than client side recording.
-\ac{KMS} allows recording files using \emph{webm} and \emph{mp4} containers.
+\ac{KMS} allows recording files using the \emph{webm} container by default but \emph{mp4} is also available.
 %RP escolheste qual? 
+%HR webm, assim fica subentendido que deixei por default
 
@@
 	When we started implementing our solution, \ac{KMS} had no support for seeking videos, which meant that blocks would would always start playing from their beginning.
-This lead to a theoretic playing time error that can be at most half the duration of a block.
-In practice, the maximum playing time error coincides with the block duration because we perform the intersection between the requested date and the block.
-But we could decrease the error by half if we calculate the intersection with the requested time plus half the duration.
+
+	If we choose the closest chunk to play this lead to a playing time error that can be at most half the duration of a block, as it can be seen on figure \ref{fig:nearchunkerror} where the black dot represents the requested time to play.
+In this case, which is the worst, either $ block_{n}$ or $ block_{n+1}$ (with duration $d$) can be played from their beginning with a maximum error ($e$) correspondent to half their duration ($\frac{d}{2}$) .
+
+	In practice, the maximum playing time error coincides with the block duration because we must show always what the user requests, as it can be seen on figure \ref{fig:interchunkerror}.
+In this case, if a user wishes to play a time that is close to the end of $block_n$ the maximum playing error ($e$) is at most the chunk's duration ($d$) because we have to play $block_n$ from its beginning.
         %RP já tinhamos falado sobre isto.
@@
         %RP Figura tem de ser apresentada no texto e explicada. O que significam as letras?
-
+        %HR done
 	\begin{figure}
@@
 	\end{tikzpicture}
-		\caption{Theoretic maximum playing error $(e)$}
+		\caption{Maximum playing error $(e)$ if we play the closest chunk}
+		\label{fig:nearchunkerror}
 	\end{figure}
@@
 		\draw [brace_bottom] (1*1.5+0.03,-3.3) -- node [below, pos=0.5] {$e_{n}=d$} 	(3*1.5-0.03,-3.3);
+		\draw  (3*1.5+0.03,-3.3) -- node [below, pos=0.5] {$e_{n+1}=0$} 	(3*1.5-0.03,-3.3);
 		\draw [brace_bottom] (3*1.5+0.03,-3.3) -- node [below, pos=0.5] {$d$} 		(5*1.5-0.03,-3.3);
@@
 	\end{tikzpicture}
-		\caption{Practical maximum playing error $(e)$}
+		\caption{Maximum playing error $(e)$ if we play the chunk that intersect the user's time}
+		\label{fig:interchunkerror}
+
 	\end{figure}

### 2016-05-11 tese2/16_implementation_content.tex
[RFT] rewriten model
[ADD] composite figure
[FIX] pseudo code for QR code content creation and implementation, fix non return to real time

@@
 var ongoingCodesMap = {};
+var startedCodesMap = {}
+
 function onCodeFound(content) {
@@
 	var hash = md5(content);
-	if(ongoingCodesMap[hash] == null){
+
+	if(startedCodesMap[hash] == null) {
 		sendQRCodeToEditingUser(hash,content);
+		startedCodesMap[hash] = currentTime;
 	}
-	ongoingCodesMap[hash] = startingTime;
+
+	ongoingCodesMap[hash] = currentTime;
 	waitTwoSeconds(); // onCodeFound() may be called from other threads
 	var newestTime = ongoingCodesMap[hash]; 
-	if(newestTime == startingTime){
+	if(newestTime == currentTime){
+		var startingTime = startedCodesMap[hash];
 		var endingTime = getCurrentTime() - getUserOffset();
+
+		startedCodesMap[hash] = null; 
 		ongoingCodesMap[hash] = null;
+
 		sendQRCodeToEditingUser(hash,null); // remove from UI
@@
     while ( hasLocalEvents() ) {
-        if ( firstEventIsOlderThanNavigatedTime() ) {
+        if ( firstEventIsOlderThan(navigatedTime)) {
             if ( firstEventIsStart() ) {
@@
     }
-    if ( hasNoLocalEvents() ) {
+    if ( hasNoLocalEvents() && !serverHasMoreContent()) {
     	return;	// nothing to do
@@
 		\centering
-		\includegraphics[width=\textwidth]{figures/timeline.png}
+		\includegraphics[width=0.8\textwidth]{figures/timeline.png}
 		\caption{Interactive time-line}
@@
 		\centering
-		\includegraphics[width=0.7\textwidth]{figures/annotation.png}
+		\includegraphics[width=0.5\textwidth]{figures/annotation.png}
 		\caption{Creating a time annotation}
@@
 		\centering
-		\includegraphics[width=0.8\textwidth]{figures/timelink.png}
+		\includegraphics[width=0.6\textwidth]{figures/timelink.png}
 		\caption{Time link creation}
@@
 	\centering
-\begin{minipage}[b]{0.55\linewidth}
+	\begin{minipage}[b]{0.45\linewidth}
 		\centering
-
 		\includegraphics[width=\textwidth]{figures/search.png}
 		a) Result list
-\label{fig:minipage1}
 	\end{minipage}
 	\quad
-\begin{minipage}[b]{0.40\linewidth}
+	\begin{minipage}[b]{0.30\linewidth}
 		\centering
-
 		\includegraphics[width=\textwidth]{figures/search2.png}
 	    b) Selected result
-\label{fig:minipage2}
 	\end{minipage}
-
 	\caption{Example of search results}
@@
 
+		The composite operation, as shown on figure \ref{fig:wcomposite} works by decoding the incoming client streams, mixing audio and video and encoding again to send back to users.
+
+	\begin{figure}[!htb]
+		\centering
+		\includegraphics[width=0.6\textwidth]{figures/wcomposite.pdf}
+		\caption{Stream decoding, composition and encoding}
+		\label{fig:wcomposite}
+	\end{figure}
+
 		We provide a way for users to select the composite view of the conference room or the video shared by a particular user device as it can be seen on figure \ref{fig:devices}.
+
+	\begin{figure}[!htb]
+		\centering
+		\includegraphics[width=0.6\textwidth]{figures/devices.png}
+		\caption{Multiple devices per user}
+		\label{fig:devices}
+	\end{figure}
 
@@
 This list of devices is also sent to the users that are also in real time mode whenever a new user enters or leaves the conference room.
-
-	\begin{figure}[!htb]
-		\centering
-		\includegraphics[width=0.8\textwidth]{figures/devices.png}
-		\caption{Multiple devices per user}
-		\label{fig:devices}
-	\end{figure}
 

### 2016-05-11 project/httpserver/app/controllers/WSController.java

@@ -182,16 +182,6 @@ public class WSController extends Controller {
                                                 return;
                                             }
                                         break;
-                                        case "setRealTime":
-                                            if(args.has("uid")) {
-                                                String owner = args.getString("uid");
-                                                String sessionId = args.optString("sid", null);
-                                                userSession.setRealTime(owner, sessionId);
-                                                userSession.sendMessage(userSession.getContent());
-                                                return;
-                                            }
-
-                                        break;
                                         case "removeUser":
                                             if (args.has("uid")) {
                                                 String uid = args.getString("uid");
@@ -251,20 +241,18 @@ public class WSController extends Controller {
                                                 return;
                                             }
                                         break;
-                                        case "setHistoric":
+                                        case "setTime":
+                                            String owner = args.optString("uid",null);
+                                            String sessionId = args.optString("sid", null);
+
                                             if(args.has("offset")){
-                                                String owner = args.optString("uid",null);
-                                                String sessionId = args.optString("sid", null);
-                                                if(owner==null){
-                                                    owner = room.getGroupId();
-                                                    sessionId = "group";
-                                                }
                                                 userSession.setOffset(args.getLong("offset"));
                                                 userSession.setHistoric(owner,sessionId);
-                                                userSession.sendMessage(userSession.getContent());
-                                                return;
+                                            }else {
+                                                userSession.setRealTime(owner, sessionId);
                                             }
-                                        break;
+                                            userSession.sendMessage(userSession.getContent());
+                                            return;
                                         case "createTag":
                                             if(args.has("time") && args.has("title")){
                                                 try {

### 2016-05-11 project/httpserver/app/main/UserSession.java

@@ -51,7 +51,11 @@ public class UserSession implements Closeable, Comparable<UserSession> {
     private Timer scheduledPlayer;
     private Object scheduledPlayerLock = new Object();
 
-    private Map<String,Object> currentQRCodes = new TreeMap<String,Object>();
+    private Object QRLocker = new Object();
+    private Map<String,Date> startingQRCodes = new TreeMap<String,Date>();
+    private Map<String,Date> currentQRCodes = new TreeMap<String,Date>();
+
+
     private EventListener<CodeFoundEvent> codeListener;
 
     public UserSession(final String userId, final Room room, WebSocket.Out<String> out) throws ServiceException {
@@ -137,12 +141,13 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                 System.out.println(".");
                 final String content = event.getValue();
                 final String hash = Tools.md5(content);
-                final Date startTime = new Date(new Date().getTime() - timeOffset);
+                final Date currentTime = new Date(new Date().getTime() - timeOffset);
 
                 if (hash != null) {
                     boolean containsQR;
-                    synchronized (currentQRCodes) {
-                        containsQR = currentQRCodes.containsKey(hash);
+                    synchronized (QRLocker) {
+                        containsQR = startingQRCodes.containsKey(hash);
+
                     }
 
                     if (!containsQR) {
@@ -153,8 +158,11 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                         room.sendMessage(msg.toString());
                     }
 
-                    synchronized (currentQRCodes) {
-                        currentQRCodes.put(hash,startTime);
+                    synchronized (QRLocker) {
+                        currentQRCodes.put(hash,currentTime);
+                        if(!startingQRCodes.containsKey(hash)){
+                            startingQRCodes.put(hash,currentTime);
+                        }
                     }
 
                     new Thread(new Runnable() {
@@ -166,13 +174,15 @@ public class UserSession implements Closeable, Comparable<UserSession> {
                                 e.printStackTrace();
                             }
                             Object newestQRCode;
-                            synchronized (currentQRCodes) {
+                            synchronized (QRLocker) {
                                 newestQRCode = currentQRCodes.get(hash);
                             }
 
-                            if (startTime == newestQRCode) {
-                                synchronized (currentQRCodes) {
+                            if (currentTime == newestQRCode) {
+                                Date startingTime;
+                                synchronized (QRLocker) {
                                     currentQRCodes.remove(hash);
[…]

### 2016-05-11 project/httpserver/public/this/js/hyper.socket.js

@@ -198,7 +198,7 @@ var HyperWebSocket = new (function() {
 	this.receiveRealTime = function(userId,sessionId){
 		console.log("receiveRealTime",userId);
 		var msg = {
-            cmd : "setRealTime",
+            cmd : "setTime",
             uid:userId
         };
 		if(sessionId){
@@ -210,7 +210,7 @@ var HyperWebSocket = new (function() {
 	this.receiveHistoric = function(userId,offset,sessionId){
 		console.log("receiveHistoric",userId);
 		var msg = {
-            cmd : "setHistoric",
+            cmd : "setTime",
             uid:userId,
             offset:offset
         };

### 2016-05-13 tese2/15_implementation_recording.tex
until 1h test case

@@
 	%{\color{red} [TALK ABOUT COMPLEXITY OF DECODE VS DECODE+MANIPULATE+ENCODE]}
-	Although we implemented a solution that worked, we immediately noticed that \emph{ffmpeg} would take some time to initialize and that lead to pauses between switching parts.
+	Although we implemented a solution that worked, we immediately noticed that \emph{ffmpeg} would take some time to initialize and that lead to pauses between switching parts which represents an undesirable behavior.
 
@@
 
-	We could have implemented fast forwarding without real time conversion by creating multiple versions of the same video with different playback speeds after the recording of a block.
+	Although we have not implemented fast forwarding, we could implement it without real time conversion by creating multiple versions of the same video with different playback speeds after the recording of a block.
 When a user needed to play, he would also need to specify one of the available speed.

### 2016-05-13 tese2/16_implementation_content.tex
until 1h test case

--- 2016-04-26 tese2/16_implementation_content.tex
@@
 
-	In order to create content the user has the option to write simple movie captions without writing any code, otherwise, as mentioned before, it can write \ac{HTML}, \ac{CSS} and \emph{JavaScript}.
+	Our system supports creating superimposed content to video, which is achieved by creating \ac{HTML} tags on top of the video  with the same size.
+The decision of which content must be displayed to each user is performed by our content scheduler which uses the user's current time in order to synchronize which content is shown or removed from the user interface.
+
+	In order to create content, the user has the option to write simple movie captions without writing any code, otherwise, as mentioned before, it can write \ac{HTML}, \ac{CSS} and \emph{JavaScript}.
 The definition of the content's starting and ending time by the user it's not an easy task.
-Defining the content's time to appear in real-time would require a previous user plan.
-Otherwise the user could make a speech and add the content for later reproduction.
-
 	Figure \ref{fig:creation} shows the user interface for creating interactive hyper content manually.
@@
 		\centering
-		\includegraphics[width=0.6\textwidth]{figures/edition.png}
+		\includegraphics[width=0.5\textwidth]{figures/edition.png}
 		\caption{Manual content creation}
@@
 
-	Listing \ref{lst:createcontent} shows the structure of the content created by one user.
+	Listing \ref{lst:createcontent} shows the structure of the content created by one user%.
+and Listing \ref{lst:receivecontent} shows the format of messages that users receive in order to supply their content schedulers with fresh data.
 
 \begin{minipage}{\linewidth}
-\begin{lstlisting}[caption={Exampe of content created by one user},label={lst:createcontent},language=json]
+\begin{lstlisting}[caption={Example of content created by one user},label={lst:createcontent},language=json]
 {
@@
 \end{minipage}
-	In order to help the content creator to show and synchronize its content in real-time we allow the user to encode its content into \emph{QR codes} and show it to the camera in real time.
+
+	Defining the content's time to appear in real-time would require a previous user plan. Otherwise the user could make a speech and insert the content later.
+
+	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real-time, we allow the user to encode its content into \emph{QR codes} and show it to the camera in real time.
 
@@
 We had to implement our own mechanism for detecting those events.
-
 Each user session in the server maintains a map with the contents that are present on the screen.
 In order to apply our algorithm more efficiently we calculate the content's hash through the \emph{md5} method.
-
-	If the hash is not present in the map it means the \emph{QR code} was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real-time.
-By doing an analogy to the \emph{Nyquist} theorem we know that if the same \emph{QR code} is not detected after a time bigger then two periods we can conclude that the \emph{QR code} leaved the screen and we add the correspondent content into our database.
+If the hash is not present in the map, it means the \emph{QR code} was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real-time.
+If the same \emph{QR code} is not detected after a time bigger then two periods we can conclude that the \emph{QR code} left the screen and we add the correspondent content into our database.
 Listing \ref{lst:pseudo_qrcode} shows the pseudo code for \emph{QR code} leaving detection.
 
-\begin{minipage}[!htb]{\linewidth}
+\begin{minipage}{\linewidth}
 \begin{lstlisting}[caption={Pseudo code for QR code leaving detection},label={lst:pseudo_qrcode},language=JavaScript]
 var ongoingCodesMap = {};
+var startedCodesMap = {}
 function onCodeFound(content) {
@@
 	var hash = md5(content);
-	if(ongoingCodesMap[hash] == null){
-		sendQRCodeToEditingUser(hash,content);
+	if(startedCodesMap[hash] == null) {
+		sendQRCodeToEditingUser(hash,content); // insert UI
+		startedCodesMap[hash] = currentTime;
 	}
-	ongoingCodesMap[hash] = startingTime;
+	ongoingCodesMap[hash] = currentTime;
 	waitTwoSeconds(); // onCodeFound() may be called from other threads
 	var newestTime = ongoingCodesMap[hash]; 
-	if(newestTime == startingTime){
+	if(newestTime == currentTime){
+		var startingTime = startedCodesMap[hash];
 		var endingTime = getCurrentTime() - getUserOffset();
+		startedCodesMap[hash] = null; 
 		ongoingCodesMap[hash] = null; 
@@
 
-	Our main content synchronization mechanism is time based but with some programming knowledge it is possible to insert \emph{JavaScript} code that fires events on user interaction.
+	Our main content synchronization mechanism is time based but, with some programming knowledge, it is possible to insert \emph{JavaScript} code that fires events on user interaction.
 For example, after a teacher's lecture, it is possible to show a quiz to the users in order to understand what they learned and then submit the data to the server for further analysis.
@@
 
-	In our solution content is represented as simple text that can contain \ac{HTML}, \ac{CSS} or even \emph{JavaScript}, this content is displayed on defined intervals of time.
+	In order to allow users to visualize past and present communications, we express the user's position as the time offset between its navigated time and the current time.
 
-	Multiple contents can be displayed at the same time, in order to achieve that, we define layers above the video with the same size. Each layer is associated to the content.
-
-	We have taken into account that the amount of content tends to grow with time and the user should only have access to a subset of content instead all of them, which would be very inefficient.
-
-	The content to display for each user depends on the user's position on the time-line, this position is given by an offset between the current time and the navigated time.
-If the user is watching the content in real-time this offset is constant so there is no need to synchronize this value.
+	In our solution, content is represented as simple text that can contain \ac{HTML}, \ac{CSS} or even \emph{JavaScript}.
+This content is displayed on defined intervals of time.
+Multiple contents can be displayed at the same time.
+In order to achieve that, we define layers above the video with the same size.
+Each layer is associated to the content.	We have taken into account that the amount of content tends to grow with time and the user should only have access to a subset of the content instead of all of it, which would be very inefficient.
+The content to display for each user depends on the user's position on the time-line.
+This position is given by an offset between the current time and the navigated time.
+If the user is watching the content in real-time this offset is always zero.
+If the user is reproducing past communications in normal playback rate, this offset is constant and, as a consequence, there is need to re-synchronize this value.
+If, for instance, we had to implement communication reproduction with a faster or slower playback rate, we would need to take into account the fact of this time offset not being constant.
 
@@
 
-	The content to return is given by the union of two content subsets:
+	The users will receive their contents from the server on five different situations:
+
 	\begin{itemize}
-		\item The intersection of the content interval and user's time.
-		\item A subset of contents which starting time is immediately following the user's time.
+		\item Conference room entrance (advertised by server).
+		\item Set of events empty and server has more content (client requested).
+		\item Content is created (advertised by server).
+		\item Content is removed (advertised by server).
+		\item User navigates to different point in time (client requested).
+	\end{itemize}
+
+	The content to return is given by the union of the two following content subsets:
+	\begin{itemize}
+		\item Content that should be currently on display.
+		\item A subset of contents whose starting time is immediately following the user's time.
 	\end{itemize}
@@
 	"cmd":"content",
-	"data":[{
+	"data":[
+		{
 			"id":"56f9fd1aa986c615fab43d69",
@@
 			"content":"<h1>Hello</h1>"
-		},{
+		},
+		{
 			"id":"56f9fd1aa986c615fab43d69",
@@
 	When a user enters the conference room, immediately after the \emph{WebSocket} creation the server sends him the current content.
-
 The user receives the content, sorts all components by time and creates a set of events.
-All the events before the user's time are rendered and removed from the set of events.
+All the events before the user's time are displayed and removed from the set of events.
 A timer is scheduled for the first component on the event set and the process repeats while the set is not empty.
 
-	If the set is empty, there two options, if the server contained more content a new request for content is made and the process starts from the beginning, namely the server sends the correspondent content again.
-If the server has no more content the process is stopped until it sends more.
-
-	The users will receive their contents from the server on five different situations:
-
-	\begin{itemize}
-		\item Conference room entrance (advertised by server).
-		\item Set of events empty and server has more content (client requested).
-		\item Content is created (advertised by server).
-		\item Content is removed (advertised by server).
-		\item User navigates to different point in time (client requested).
-	\end{itemize}
+	If the event set is empty, there two options.
+If the server contained more content, a new request for content is made and the process starts from the beginning, namely the server sends the correspondent content again.
+If the server has no more content, the process is stopped until it sends more.
 
@@
 
-\begin{minipage}[!htb]{\linewidth}
+\begin{minipage}{\linewidth}
 \begin{lstlisting}[caption={Pseudo code for hyper content scheduler},label={lst:pseudo_render},language=JavaScript]
@@
     }
-    if ( hasNoLocalEvents() ) {
+    if ( hasNoLocalEvents() && serverHadNoMoreContent()) {
     	return;	// nothing to do
     } 
-    if ( hasNoStartEvents() && serverHasMoreContent() ) { 
+    if ( hasNoStartEvents() && serverHadMoreContent() ) { 
        	requestContentFromServer(navigatedTime);
@@
 
+	\subsection{Content Removal}
+		Each layer is identified by its content identification number which is used by our content scheduler in order to remove disappearing content.
+
+		If a user clicks over one layer's content the layer becomes selected (a frame is drawn around it) and its content identification number is added to a list containing the selected layers.
+When a user wishes to remove superimposed content, the selected layer ids are sent to the application server, with the same format as shown on Listing \ref{lst:deletecontent}, and subsequently all users will receive a message in order to update their user interface.
+
+\begin{minipage}{\linewidth}
+\begin{lstlisting}[caption={Example of removing content message sent by one user},label={lst:deletecontent},language=json]
+{
+	"cmd":"deleteContent",
+	"content":["56f9fd1aa986c615fab43d69","57339588ac687b423e5b3a1e"]
+}
+\end{lstlisting}
+\end{minipage}
+
 	\subsection{Security Concerns}
 
-	Our solution is flexible on what kind of interactions are possible to the users in real time but allowing users to write \emph{JavaScript} that is executed on the other users browser would attackers to misuse their resources and access to critical information.
-
+	Our solution is flexible on what kind of interactions are possible to the users in real time, but allowing users to write \emph{JavaScript} that is executed on the other users browser would attackers to misuse their resources and access critical information.
 We could solve this problem easily by escaping any \emph{script} tag present on the content, but we would sacrifice the kind of interactions that are possible.
-
-	By not allowing \emph{JavaScript} we would need to implement all actions a priori and fire them when a type of message is received.
-We decided to ignore the security vulnerabilities that are exposed by evaluating \emph{JavaScript} because in order to offer the same interactions we would have to do an exhaustive functional requirements gathering.
+By not allowing \emph{JavaScript} we would need to implement a subset of conceivable actions a priori and fire them when a type of message is received.
+We decided to ignore the security vulnerabilities that are exposed by evaluating \emph{JavaScript} because, in order to offer the same interactions, we would have to do an exhaustive functional requirements gathering and, in our prototype, our goal is to explore users behavior when interacting with our system.
 
 	Another way to solve this problem, which is not within the goals of this thesis, is to analyze the \emph{JavaScript} code and detect if it is malicious.
+Although we have not performed a detailed study on this field, we are aware of its complexity and the existence of solutions to deal with this problem, in particular \emph{EarlyBird}\cite{earlybird} came to our attention.
+
+\emph{EarlyBird} uses machine learning techniques in order to improve malicious \emph{JavaScript} code detection accuracy.
 
@@
 	We have used \emph{vis.js} to display out time-line.
-This library was created for content navigation through time, but it was not designed to be always moving automatically, we have created a background timer that performs the animation of moving the window of time bounds and user navigated time marker.
+This library was created for content navigation through time, but it was not designed to be always moving automatically.
+We have created a background timer that performs the animation of moving the window of time bounds and user navigated time marker.
 
-	Figure \ref{fig:timeline} shows the the graphic appearance of our interactive time-line, in order to navigate through time the user must drag and drop the time-line horizontally, when the user drops the time-line an event handler is called with the new user's time offset which will be used to send a message to the server in order to choose the correspondent content to display.
+	Figure \ref{fig:timeline} shows the graphical appearance of our interactive time-line.
+In order to navigate through time, the user must drag and drop the time-line horizontally.
+When the user drops the time-line an event handler is called with the new user's time offset, which will be used to send a message to the server in order to choose the correspondent content to display.
 
-	\begin{figure}[!htb]
+	\begin{figure}
 		\centering
-		\includegraphics[width=\textwidth]{figures/timeline.png}
+		\includegraphics[width=0.95\textwidth]{figures/timeline.png}
 		\caption{Interactive time-line}
@@
 
-	We have noticed the server time being different from client time and this created graphic inconsistencies such as showing existing recorded blocks of movie in the future.
-To solve this problem the server sends its time to client immediately after the \emph{webSocket} creation, we synchronize the time-line with the server and although it may exist a small error due to network transmission time the graphical error is negligible.
+	We have noticed the server time being different from client time and this created timing inconsistencies, such as showing existing recorded blocks of movie in the future.
+To solve this problem, the server sends its time to the client immediately after the \emph{webSocket} creation.
+We synchronize the time-line with the server and although it may exist a small error due to network transmission time, the graphical error is negligible.
 
@@
 
-	Time annotations are simpler way save points in time and share them with other users.
-When a user enters in a conference room the server sends all annotations so they can be displayed directly into the user time-line.
-Each time annotation is created all users are advertised so they can update their interfaces.
+	Time annotations are a simpler way to mark points in time and share them with other users.
+When a user enters a conference room, the server sends all annotations so they can be displayed directly into the user time-line.
+When each time annotation is created, all users are notified so they can update their interfaces.
 
-	Figure \ref{fig:annotation} shows the creation and placement of annotations on the time-line, to save the annotation the user must click on the floppy icon in order to save it and advertise the conference participants.
+	Figure \ref{fig:annotation} shows the creation and placement of annotations on the time-line.
+To save the annotation the user must click on the floppy icon in order to save it and notify the conference participants.
 
-	\begin{figure}[!htb]
+	\begin{figure}
 		\centering
@@
 
-	Besides the ability to create tags it is also possible to create time hyper-links, see figure \ref{fig:timelink}, that can be sent to other users externally so they can navigate directly to the content when entering in the conference room.
+	Besides the ability to create tags, it is also possible to create time hyper-links, see Figure \ref{fig:timelink}, that can be sent to other users externally so they can navigate directly to the content when entering the conference room.
 
-	\begin{figure}[!htb]
+	\begin{figure}
 		\centering
@@
 
-	Users can easily search for annotations and contents and travel to their correspondent times.
-In the case of hyper content after handling the result from database we ignore \ac{HTML} tags, extract the text with \emph{Jsoup}\footnote{\url{http://jsoup.org/} (Accessed 21 March 2016)} and apply the query again.
+	Users can search for annotations and contents and travel to their correspondent times.
+In the case of hyper content, after handling the result from database, we extract the text by discarding \ac{HTML} tags with \emph{Jsoup}\footnote{\url{http://jsoup.org/} (Accessed 21 March 2016)} and apply the query again.
 We extract the text from \ac{HTML} because accidentally searching for text contained in \ac{HTML} tags would lead to incorrect results.
@@
 
-\begin{figure}[!htb]
+\begin{figure}
 	\centering
@@
 		\centering
-
 		\includegraphics[width=\textwidth]{figures/search.png}
 		a) Result list
-\label{fig:minipage1}
 	\end{minipage}
@@
 		\centering
-
 		\includegraphics[width=\textwidth]{figures/search2.png}
 	    b) Selected result
-\label{fig:minipage2}
 	\end{minipage}
-
 	\caption{Example of search results}
@@
 
-	\subsection{Switching views}
+	\subsection{View composition and switching view}
+
+		In order to show all streams to all users, each conference participant could receive individual streams for each user that share video.
+In overall, if there were $n$ active participants in the conference room, the number of streams per user would be $n$ ($1$ for sending video and $n-1$ for receiving) and, consequently, the number of streams in that conference room, as observed in Figure \ref{fig:xcomposite}, would be $n^2$.
+
+\begin{figure}
+\centering
+\begin{minipage}[b]{0.45\linewidth}
+	\centering
+	\includegraphics[width=\textwidth]{figures/xcomposite.pdf}
+		\caption{Client-side stream composition}
+	\label{fig:xcomposite}
+\end{minipage}
+\quad
+\begin{minipage}[b]{0.45\linewidth}
+	\centering
+	\includegraphics[width=\textwidth]{figures/ecomposite.pdf}
+		\caption{Server-side stream composition}
+	\label{fig:ecomposite}
+\end{minipage}
+\end{figure}
+
+		By mixing all streams into a single one, users can receive a single stream with the representation of all active users.
+With this approach, the total number of streams within a conference room can be, at most, $2n$ (each user has $2$ streams, one to send and another to receive) as observed in Figure \ref{fig:ecomposite}.
+
+		The numbers we have presented above were calculated taking into account that all users would receive and send video at the same time, which is the worst case scenario, although users may choose just receiving streams.
+
+		When the amount of participants is just two, both solutions require the same amount of streams but as the number of participants increase, the composition operation tends to be a more efficient approach.
+
+		The composite operation, as shown on figure \ref{fig:wcomposite}, is performed at \ac{KMS} and works by decoding the incoming client streams, mixing audio and video and encoding again to send back to users.
+
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.7\textwidth]{figures/wcomposite.pdf}
+		\caption{Stream decoding, composition and encoding}
+		\label{fig:wcomposite}
+	\end{figure}
 
@@
 
-		In order to maintain a list of available user devices that can be played at the navigated time if a user plays or ends playing a block of recorded video it obtains at the same time a list of all devices available for each user at that time.
+	\begin{figure}
+		\centering
+		\includegraphics[width=0.7\textwidth]{figures/devices.png}
+		\caption{Multiple devices per user}
+		\label{fig:devices}
+	\end{figure}
+
+		When a user plays or ends playing a block of recorded video, it requests a list of available devices from each user that can be played at the navigated time by querying the database.
 
@@
 
-	\begin{figure}[!htb]
-		\centering
-		\includegraphics[width=0.8\textwidth]{figures/devices.png}
-		\caption{Multiple devices per user}
-		\label{fig:devices}
-	\end{figure}
-
-		We also provide an automatic mechanism for switching the view to the current speaker.
+		In addition to select a stream to view, which remains the same if not interrupted, we also provide an automatic mechanism for switching the view to the current speaker.
 With respect to sound analysis, the sound samples are analyzed on the client side through the web audio \ac{API}.
@@
 
+		If multiple participants are talking at the same time, the selected view will change indeterminately between them.
+Although such behavior is not the most desired, we could have done a dedicated composite view in order to show the talking participants.
+
 \subsection{Non interactive media synchronization}
-		By default each user receives the real time mixed video and audio of every users. As described on the previous section, users can switch the current view even for recorded media.
+		By default, each user receives the real time mixed video and audio of every users. As described on the previous section, users can switch the current view even for recorded media.
 
-		Whenever a user changes the current viewed time, our server calculates which chunks to reproduce.
+		Whenever a user changes the current viewed time, our server calculates which chunks to reproduce.In order to determine which chunks are going to be played we could find any chunk that matches the session id and intersects the given user time.
+Although this seems logical, we limit the size of the results to one and ignore the starting component of the recording chunk.
+As a consequence, we will obtain the recording chunk that ends right after the playing time.
 
-		In order to determine which chunks are going to be played we could find any chunk that matches the session id and intersects the given user time.
-Although this seems logical, we limit the size of the results to one and ignore the starting component of the recording chunk.
-As a consequence, we will obtain the recording chunk that ends after the playing time.
+		The obtained chunk may or not intersect the given time.
+If it intersects then we immediately start playing the chunk, otherwise we know that the chunks starting time is after the playing time and we schedule the player to reproduce after a time duration given by the difference of the playing time and chunk starting time.
 
-		The obtained chunk may or not intersect the given time, if it intersects than we immediately start playing the chunk, otherwise we know that the chunks starting time is after the playing time and we schedule the player to reproduce after a time duration given by the difference of the playing time and chunk starting time.
-
-		In order to reduce the delay between switching parts we perform a query on the database for finding the next chunk to play while the selected media is playing.
-In order to find the next chunk we use the chunk that has a bigger id immediately after the playing chunk.
+		In order to reduce the delay between switching parts we perform a query on the database for finding the next chunk to play while the selected media is still playing.
+In order to find the next chunk, we use the chunk that has a bigger id immediately after the playing chunk.
 
@@
 		If a user leaves a conference room, it is expected that the sequence of played blocks ends.
-When a block is not found for a given \emph{webSocket} session we try to find a block for the same user and if not possible we try to play a composite recorded block, lastly, if every attempt fails we wait for the next interval the contains media.
+When a block is not found for a given \emph{webSocket} session, we try to find a block for the same user and if not possible, we try to play a composite recorded block, lastly, if every attempt fails we wait for the next interval that contains media.
 
 \subsection{Chat}
-		Not less important the chat functionality was implemented using \emph{webSockets}, when a message is sent by a user, it is received by the server, which records it and sends to all the users.
+		Not less important the chat functionality was implemented using \emph{webSockets}. When a message is sent by a user, it is received by the server, which records it and sends to all the users.
 
 		In order to support receiving a huge amount of messages and not having a bottleneck when loading the conference room, we just load an amount of messages that fill the chat panel.
-When a user scrolls the chat to read old messages, we detect when the chat scroll bar reaches the limit and we request more older messages to the server automatically.
+When a user scrolls the chat to read old messages, we detect when the chat scroll bar reaches the limit and request older messages from the server automatically.
 
 		Moreover, when the clients receive messages, whether older or newer ones, their ids are used to place them in the correct position.
-By other words, older messages are placed in the top of the chat panel and newer ones in the bottom.
+In other words, older messages are placed in the top of the chat panel and newer ones in the bottom.
 
-		Besides that, we also support sending files to other users, which are stored on database and accessed through an \ac{URL}.
+		Besides that, we also support sending files to other users, which are stored on the database and accessed through an \ac{URL}.
 
@@
 
-The state of our collaborative editor is not saved on database every time it changes. Instead the users just synchronize the editor content among them and save it a priori.
+The state of our collaborative editor is not saved on the database every time it changes.
+Instead, the users just synchronize the editor content among themselves using the application server to relay editor changes and save on demand.
 
-If a user joins a conference room which editor was changed and the content differs from the version stored in the database, our server does not have the right content to deliver the user.
-As a consequence our server names a random participant to yield its collaborative content and delivers it to the joining user.
+When a user joins a conference room and the content differs from the version stored in the database, our server does not have the right content to deliver the user.
+As a consequence, our server names a random participant to yield its collaborative content and delivers it to the joining user.
 
@@
 
-Accordingly to \emph{ot.js} there are three types of operations:
+According to \emph{ot.js} there are three types of operations:
 \begin{itemize}
@@
 
-In summary, our server just ensures mutual exclusion and delivers the messages among users without even parsing the messages content.
+In summary, our server just ensures the ordering of operations and delivers the messages among users without even parsing the message's content.

### 2016-05-13 tese2/17_implementation_deploy.tex
until 1h test case

@@
 
-	The hardware was gently provided by INESC-ID\footnote{\url{http://www.inesc-id.pt/} (Accessed March 26, 2016)}, the specifications of the server that we used are specified on table \ref{table:hwspecs}.
+	The hardware was gently provided by INESC-ID\footnote{\url{http://www.inesc-id.pt/} (Accessed March 26, 2016)}.
+The specifications of the server that we used are described in Table \ref{table:hwspecs}.
 
-	\begin{table}[!htb]
+	\begin{table}
 \centering
@@
 
-	\begin{table}[!htb]
+	\begin{table}
 \centering
@@
 \hline
-Operating System & Linux version 3.16.0-4-amd64 \\ \hline
+Kernel & Linux version 3.16.0-4-amd64 \\ \hline
 Distribution & Debian GNU/Linux 8 (jessie)\footnotemark \\ \hline
@@
 \subsection{Software}
-	\emph{INESC-ID} provided us a restricted \emph{Linux} account without administration permissions, which we prevented us the installation of our solution directly on the machine because we would need administrative privileges to install all the software that our solution depends on.
-Although that limitation, they provided access to Docker\footnote{\url{https://www.docker.com/} (Accessed March 27, 2016)} which run on host operating system as an isolated process in user space.
+	\emph{INESC-ID} provided us a restricted \emph{Linux} account without administration permissions, which prevented us from installing our solution directly on the machine because we would need administrative privileges to install all the software that our solution depends on.
+Despite that limitation, they provided access to Docker\footnote{\url{https://www.docker.com/} (Accessed March 27, 2016)} which run on the host operating system as an isolated process in user space.
 
@@
 
-	We could use one docker image for each component but in order to reduce the network usage we preferred to install all the components within the same image.
+	We could use one docker image for each component but, in order to reduce the network usage, we preferred to install all the components within the same image.
 At the same we provide an all-in-one easy to deploy solution as all the \ac{IP}s are local.

### 2016-05-13 tese2/18_implementation_summary.tex
until 1h test case

@@
 
-Then, we have underlined the requirements of our signaling protocol and as consequence we have described, in detail, the protocol itself.
+Then, we underlined the requirements of our signaling protocol and, as a consequence, we have described, in detail, the protocol itself.
 
-With the signaling protocol implemented we observed the first outcomes of using \ac{WebRTC}, the basic functionality we have implemented was an echo of the video and audio sent by users.
+With the signaling protocol implemented, we observed the first outcomes of using \ac{WebRTC}, the basic functionality we have implemented was an echo of the video and audio sent by users.
 However we have implemented other features such as switching to other user streams, recording and mixing multiple streams into a single one.
@@
 Then, we have described how we have implemented our chat and collaborative editor.
-
-Lastly, although we have tested our solution with our own development machine during the implementation, we have described how we have deployed our system for allowing a public access and perform tests in a real case environment.

### 2016-05-13 tese2/abstract.tex
until 1h test case

@@
 
-The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to communication and collaboration services.
-This paradigm allows to synchronize, structure and navigate communication content, providing a way to integrate social media content and collaborative tools into voice and video calls.
-Voice and image together can express emotions and expose creativity like no other medium can. With hypermedia concepts, we can add more value to voice and video communications.
+The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content.
+This paradigm allows to synchronize, structure and navigate communication content integrated into voice and video calls.
 
-\emph{WebRTC} technology allows real-time communications between web browsers without the need to install and use additional applications or plug-ins.
+Voice and image together can express emotions like no other medium can. With hypermedia concepts, we can add more value to conference calls.
+
+\emph{WebRTC} technology allows real-time communications between web browsers without the need to install additional software.
 The nature of web browser applications already follows the hypermedia concept, which makes \emph{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
-{\color{blue}The web browser platform provides an abstraction layer that makes it possible to create applications that run independently from the operating system.}
+The web browser platform provides an abstraction layer that makes it possible to create applications that run independently from the operating system.
 The native support for \emph{WebRTC} in operating systems extends its usage to outside the web browser, allowing for the exploration of functionalities for which web browsers provide poor support, such as video recording and massive information storage.
@@
 
-{\color{red}Our goal in this project is to develop an application targeted to the web platform, resorting to \emph{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types.}
-
-{\color{red}In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor, time annotations, a chat that supports sending files to conference participants and mechanism to superimpose hyper-content to video.}
+Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging and a mechanism to superimpose hyper-content to video.
 
@@
 
-{\color{blue}
 In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose and implement an architecture for an hyper-linked communication application based on \emph{WebRTC}.
-This work was evaluated using a panel of 20 users, who reported that they liked to use it and thought it to be extremely innovative.
-}
+This work was evaluated by users, who reported that they liked to use it and thought it to be extremely innovative.
 

### 2016-05-13 tese2/evaluation.tex
[RFT] rewriten model
until 1h test case

@@
 \label{chapter:evaluation}
-  In the previous chapter we described how we implemented our solution namely the data model that we used, the description of the signaling protocol, how we performed stream recording, how we overlayed interactive content to video and lastly how we deployed our solution.
+  In the previous chapter we described how we implemented our solution namely the data model that we used, the description of the signaling protocol, how we performed stream recording, how we overlayed interactive content to video.
 
-  In this chapter we describe how we tested our solution, show and analyze the results in order to validate the contributions of this thesis.
+  In this chapter we describe how we deployed our solution, tested our prototype, show and analyze the results in order to validate the contributions of this thesis.
+
+\input{17_implementation_deploy.tex}
 
@@
   We have also tested the performance of our solution by measuring the used resources.
-Those performance tests are crucial to ensure that our solution is in fact stable and users can use it timelessly without decreasing the quality of their experience.
+Those performance tests are crucial to ensure that our solution is in fact stable and users can use it endlessly without decreasing the quality of their experience.
 
@@
 
-     In this section we describe performance test scenarios that we have applied and their respective results.
+     In this section we describe performance tests that we have applied and their respective results.
 
@@
 
-      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects with a periodicity of one second; CPU, physical memory information relative to each running process and network usage relative to each interface.
+      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects with a periodicity of one second: CPU, physical memory information relative to each running process and network usage relative to each interface.
 
       We would like to collect network information relative to each process, which \emph{nethogs}\footnote{\url{https://raboof.github.io/nethogs/} (Accessed March 27, 2016)} provide but it could not capture the network usage of some processes.
-Although this would be useful for a deep analysis, we know which network interfaces are used to establish connections between processes as we show on table \ref{table:interfacemap}.
+Although this would be useful for a deep analysis, we know which network interfaces are used to establish connections between processes as we show on Table \ref{table:interfacemap}.
 
-  \begin{table}[!htb]
+  \begin{table}
 \centering
@@
 
-      The performance test scenario that we have defined consists on two phases, the first phase consists only on having users, with similar computer and network specifications, entering sequentially on the conference room, the second phase consists on the users leaving the conference room.
+      The performance test scenario that we have defined consists on two phases, the first phase consists only on having users, with similar computer and network specifications, entering sequentially on the conference room.
+The second phase consists on the users leaving the conference room.
 Each event, joining and leaving, occurs with intervals of one minute in total of thirteen minutes (780 seconds).
@@
 
-      We are presenting in this section the results of our performance tests after implementing our solution.
+      In this section we discuss the results of our performance tests performed after implementing our solution.
 \subsubsection{Network usage}
 
-      Every time a user shares its camera or its screen in the context of a conference room, its offered video is sent to the server and independently from watching an individual stream or the mixed version the server just sends a unique stream back to the user.
+      Every time a user shares its camera or its screen in the context of a conference room, its offered video is sent to the server and, independently from watching an individual stream or the mixed version, the server just sends a unique stream back to the user.
 
-      From the media server prespective if there are $n$ clients connected each of them sending and receiving one stream, it is expected that server sends and receives also $n$ streams.
-By so we expect that the amount of network increases linearly as users join a conference room.
+      From the media server perspective if there are $n$ clients connected, each of them sending and receiving one stream, it is expected that server sends and receives also $n$ streams.
+As such, we expect that the amount of network traffic increases linearly as users join a conference room.
+Figure \ref{fig:test_full_features_net} confirms our expectations.
+Each vertical yellow line represents one event: the first seven events are users entering the conference room, the next ones represent users leaving the conversation.
 
-      Figure \ref{fig:test_full_features_net} confirms our expectations, each vertical yellow line represents each event, the first seven events are users entering the conference room the next ones represent users leaving the conversation.
-
-\begin{figure}[!htb]
+\begin{figure}
   \centering
@@
       The blue peaks are caused by the signaling phase and web page downloads, including resources such as images, stylesheets and javascript files. 
-      The green peaks are caused by that being transfered between \ac{KMS} and \emph{MongoDB} through \emph{Kurento Repository} each peak occurs each time a block of video is recorded which in this case is every ten seconds.
-      The recordings are synchronized so all user and mixed blocks starts and ends at the same time, that's why the amount of work done every ten seconds accumulates and because this is performed locally the maximum transfer rate is limited by the performance of the memory as buffers are written to buffers then to disks.
-      Client's sent data transfer rate has no significant peaks as signaling information contains few information.
+      The green peaks are caused by video and audio being transfered between \ac{KMS} and \emph{MongoDB} through \emph{Kurento Repository}.
+Each peak occurs every time a block of video is recorded, which in this case is every ten seconds.
+      The recordings are synchronized so all user and mixed blocks start and end at the same time.
+That is why the amount of work done every ten seconds accumulates, and because this is performed locally, the maximum transfer rate is limited by the performance of the memory as buffers are written to buffers then to disks.
+      Sent data transfer rate has no significant peaks as \ac{HTTP} requests and signaling information contains little information.
 
@@
 
-\begin{figure}[!htb]
+\begin{figure}
   \centering
@@
 
-With this results we conclude that if we want to scale our solution's storage using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \emph{MongoDB} should be installed in the same machine because the loopback interface can handle bigger transfer rates than the remaining network interfaces as seen on figure \ref{fig:scale1}.
-For the same reason another possible configuration for scaling the storage is to install both \emph{Kurento Repository} and \ac{KMS} on the same machine as seen on figure \ref{fig:scale2}.
+With this results we conclude that if we want to scale our storage solution using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \ac{KMS} should be installed on the same machine, as depicted on figure \ref{fig:scale2}, because the loopback interface can handle bigger transfer rates than the remaining network interfaces.
 Installing the repository on the same machine as a database node does not ensure that recorded videos are stored in the same machine, for this reason we would prefer installing \ac{KMS} and \emph{Kurento Repository} on the same machine.
 
-\begin{figure}[!htb]
+\begin{figure}
   \centering
-\begin{minipage}[b]{0.45\linewidth}
-  \centering
-  \includegraphics[width=0.8\textwidth]{figures/scale1.pdf}
-  \caption{Repository and database in the same machine}
-  \label{fig:scale1}
-\end{minipage}
-\quad
-\begin{minipage}[b]{0.45\linewidth}
-  \centering
-  \includegraphics[width=0.8\textwidth]{figures/scale2.pdf}
+  \includegraphics[width=0.5\textwidth]{figures/scale2.pdf}
   \caption{Repository and KMS in the same machine}
   \label{fig:scale2}
-\end{minipage}
 \end{figure}
 
-On the other hand figure \ref{fig:test_client_net} shows the respective network usage on the client side during our test case.
-We observe that int he first seconds the client adjusts the video quality that sends to \ac{KMS}, whenever a new client enters in the conference we observer that \ac{KMS} decreases the video quality in order to instantaneously integrate a new user into the conference room, after a while \ac{KMS} realizes that its network can handle the increase of clients and sends the video with a better quality to every participant.
+On the other hand, Figure \ref{fig:test_client_net} shows the respective network usage on the client side during our test case.
+We observe that in the first seconds the client adjusts the video quality it sends to \ac{KMS}.
+Whenever a new client enters the conference, we observe that \ac{KMS} decreases the video quality in order to instantaneously integrate a new user into the conference room.
+After a while, \ac{KMS} realizes that the network can handle the increase of clients and sends the video with a better quality to every participant.
 When a user leaves the conference room \ac{KMS} has no need to decrease the participant's video quality as less network bandwidth will be used.
 
-\begin{figure}[!htb]
+\begin{figure}
   \centering
@@
 
-Figure \ref{fig:test_full_features_mem} shows the memory usage during our performance test case.
+Figure \ref{fig:test_full_features_mem} shows the memory usage during our performance test.
 Both \ac{JVM}, \emph{MongoDB} and \ac{KMS} performs their own memory management by holding and recycling objects when needed.
-The expected and observed behavior of the memory usage is growth memory usage while the users are entering in the conference room and a memory usage stabilization afterwards.
+The expected and observed behavior of the memory usage is growth of memory usage while the users are entering the conference room and a memory usage stabilization afterwards.
 
-\begin{figure}[!htb]
+\begin{figure}
   \centering
@@
 \emph{MongoDB} checkpoints data to disk every 60 seconds or when journal data exceeds 2GB\footnote{\url{https://docs.mongodb.org/manual/faq/storage/}(Accessed March 28, 2016)}, that explains the small memory usage peaks during our test case.
-When the conference room is empty no there are no video recordings, which explains the memory stabilization at the end.
+When the conference room is empty there are no video recordings, which explains the memory stabilization at the end.
 
-\emph{KMS} memory usage was increasing and in fact we did not expected that, we have done more tests and we have observed that by disabling the recorder feature memory usage kept linearly related to the amount of clients present on the room.
-Therefore, we have verified our implementation and we have concluded that nothing was wrong with our code, we have confirmed that every allocated resource was indeed released after the end of recording.
+\emph{KMS} memory usage was increasing and in fact we did not expected that.
+We have done more tests and we have observed that by disabling the recorder feature, memory usage kept linearly related to the amount of clients present on the room.
+Therefore, we have verified our implementation and we have concluded that nothing was wrong with our code: we have confirmed that every allocated resource was indeed released after the end of recording.
 As a result of our intensive tests and consecutive failures we suspected that the problem was not ours.
@@
 Hence, in order to confirm our suspicions, we have decided to read the implementation of \ac{KMS} and we have found that \ac{KMS} was not releasing the memory if the recorder was stopped by us.
-Subsequently, we have changed the source code and submitted to \emph{Kurento} team through a \emph{GIT}'s \emph{"pull request"}.
+Subsequently, we have changed the source code and submitted to \emph{Kurento} team through a \emph{GitHub}'s \emph{"pull request"}, which was accepted.
 
-Figure \ref{fig:test_ram_fixed_mem} shows the memory usage during our performance test case taking into account the patch that we applied to \ac{KMS}.
+Figure \ref{fig:test_ram_fixed_mem} shows the memory usage during our performance test case using the patch that we applied to \ac{KMS}.
+As we can observe, there were \ac{KMS} and also \emph{Kurento Repository} memory usage improvements.
 
-\begin{figure}[!htb]
+\begin{figure}
   \centering
@@
 As we can see there are periodic processing peaks every ten seconds that are due to block recording.
+\ac{KMS} is responsible for most \ac{CPU} usage.
 
-\begin{figure}[!htb]
+\begin{figure}
   \centering
@@
 
-  Just for testing purposes we have done the same performance tests without detecting \emph{QR codes} in order to understand how \ac{CPU} intensive is this task.
+  Just for testing purposes, we performed the same performance tests disabling \emph{QR codes} detection in order to understand how \ac{CPU} intensive this task is.
 Figure \ref{fig:test_without_qrcode_cpu} shows the results for the same experience but with \emph{QR code} detection disabled.
 
-\begin{figure}[!htb]
+\begin{figure}
   \centering
@@
 
-We conclude that \emph{QR code} detection is a very intensive task, approximately doubling the amount of work performed by \ac{CPU}.
-Without this feature the network and memory usage had insignificant changes compared to \ac{CPU} usage.
+We conclude that \emph{QR code} detection is a very intensive task, approximately doubling the amount of work performed by the \ac{CPU}s.
+Without this feature, the network and memory usage had insignificant changes compared to \ac{CPU} usage.
+
+Even though, with this test results, we conclude that our solution's bottleneck is the \ac{CPU} usage at \ac{KMS}.
 
@@
 
-\begin{figure}[!htb]
+\begin{figure}
 \centering
@@
 
-In this chapter we described the test scenarios that we have created in order to evaluate our system.
+In this chapter we have described how we have deployed our system for allowing a public access and perform tests in a real case environment.
+
+Then, we described the test scenarios that we have created in order to evaluate our system.
 

### 2016-05-13 tese2/introduction.tex
until 1h test case

@@
 
-	{\color{red}Our goal in this project is to develop an application targeted to the web platform, resorting to \ac{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \emph{HTML5}, \emph{CSS3} and \emph{JavaScript} including continuous media types such as video, music and animations.}
+	{\color{red}Our goal in this project is to develop an application targeted to the web platform, resorting to \ac{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \ac{HTML}5, \ac{CSS}3 and \emph{JavaScript} including continuous media types such as video, music and animations.}
 

### 2016-05-13 tese2/xacronyms.tex
until 1h test case

\newacronym{AES}{AES}{Advanced Encryption Standard}
\newacronym{API}{API}{Application Programming Interface}
\newacronym{BOSH}{BOSH}{Bidirectional-streams Over Synchronous HTTP}
\newacronym{CBC}{CBC}{Cipher Block Chaining}
\newacronym{CPU}{CPU}{Central Processing Unit}
\newacronym{CRUD}{CRUD}{Create, Read, Update and Delete}
\newacronym{CSS}{CSS}{Cascading Style Sheets}
\newacronym{DOM}{DOM}{Document Object Model}
\newacronym{DTLS}{DTLS}{Datagram Transport Layer Security}
\newacronym{DVD}{DVD}{Digital Versatile Discs}
\newacronym{FMS}{FMS}{Flash Media Server}
\newacronym{HTML}{HTML}{HyperText Markup Language}
\newacronym{HTTP}{HTTP}{Hypertext Transfer Protocol}
\newacronym{ICE}{ICE}{Interactive Connectivity Establishment}
\newacronym{IM}{IM}{Instant Messenger}
\newacronym{IoT}{IoT}{Internet of Things}
\newacronym{IPv4}{IPv4}{Internet Protocol Version 4}
\newacronym{IPv6}{IPv6}{Internet Protocol Version 6}
\newacronym{IP}{IP}{Internet Protocol}
\newacronym{JSON}{JSON}{JavaScript Object Notation}
\newacronym{JVM}{JVM}{Java virtual machine}
\newacronym{KMS}{KMS}{Kurento Media Server}
\newacronym{MCU}{MCU}{Multipoint Control Unit}
\newacronym{MPEG}{MPEG}{Moving Picture Experts Group}
\newacronym{MVC}{MVC}{Model-View-Controller}
\newacronym{NAT}{NAT}{Network Address Translation}
\newacronym{OT}{OT}{Operational transformation}
\newacronym{P2P}{P2P}{Peer-to-peer}
\newacronym{PSTN}{PSTN}{Public Switched Telephone Network}
\newacronym{RAM}{RAM}{Random-access memory}
\newacronym{REST}{REST}{Representational State Transfer}
\newacronym{RFC}{RFC}{Request For Comments}
\newacronym{RTCP}{RTCP}{Real Time Control Protocol}
\newacronym{RTSP}{RTSP}{Real Time Streaming Protocol}
\newacronym{RTMFP}{RTMFP}{Real-Time Media Flow Protocol}
\newacronym{RTMP}{RTMP}{Real Time Messaging Protocol}
\newacronym{RTP}{RTP}{Real-time Transport Protocol}
\newacronym{SAMI}{SAMI}{Synchronized Accessible Media Interchange}
\newacronym{SCTP}{SCTP}{Stream Control Transmission Protocol}
\newacronym{SDP}{SDP}{Session Description Protocol}
\newacronym{SigOfly}{SigOfly}{Signaling-On-the-fly}
\newacronym{SIMPLE}{SIMPLE}{SIP for Instant Messaging and Presence Leveraging Extensions}
\newacronym{SIP}{SIP}{Session Initiation Protocol}
\newacronym{SMIL}{SMIL}{Synchronized Multimedia Integration Language}
\newacronym{SMS}{SMS}{Short Message Service}
\newacronym{SRTCP}{SRTCP}{Secure RTCP}
\newacronym{SRTP}{SRTP}{Secure Real-time Transport Protocol}
\newacronym{SRT}{SRT}{SubRip Text}
\newacronym{SSL}{SSL}{Secure Sockets Layer}
\newacronym{STUN}{STUN}{Session Traversal Utilities for NAT}
\newacronym{SVG}{SVG}{Scalable Vector Graphics}
\newacronym{TCP}{TCP}{Transmission Control Protocol}
\newacronym{TLS}{TLS}{Transport Layer Security}
\newacronym{TMN}{TMN}{This Means Nothing}
\newacronym{TURN}{TURN}{Traversal Using Relays around NAT}
\newacronym{UDP}{UDP}{User Datagram Protocol}
\newacronym{URL}{URL}{Uniform Resource Locator}
\newacronym{VoIP}{VoIP}{Voice Over IP}
\newacronym{WebRTC}{WebRTC}{Web Real-Time Communication}
\newacronym{WebVTT}{WebVTT}{Web Video Text Tracks}
\newacronym{XAML}{XAML}{eXtensible Application Markup Language}
\newacronym{XEP}{XEP}{XMPP Extensions}
\newacronym{XHTML}{XHTML}{eXtensible Hypertext Markup Language}
\newacronym{XML}{XML}{eXtensible Markup Language}
\newacronym{XMPP}{XMPP}{Extensible Messaging and Presence Protocol}

### 2016-05-13 tese2/xthesis.tex
until 1h test case

\documentclass[conference,compsoc,a4paper]{IEEEtran}
\ifCLASSOPTIONcompsoc
  \usepackage[nocompress]{cite}
\else
  \usepackage{cite}
\fi

\usepackage{array}
\usepackage[utf8]{inputenc}   % <<<<< Linux
\usepackage[nomain,acronym,xindy,toc]{glossaries} 
\usepackage{dirtytalk}
\usepackage{graphicx}
\usepackage{tabularx} 
\usepackage{hyperref}
%Includes "References" in the table of contents
%\usepackage[nottoc]{tocbibind}

\hyphenation{op-tical net-works semi-conduc-tor}
\loadglsentries[main]{xacronyms}
\begin{document}

\newcolumntype{Y}{>{\centering\arraybackslash}X}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{\linewidth/#1}}

\title{Hyper-linked Communications: WebRTC enabled asynchronous collaboration}

\author{\IEEEauthorblockN{Michael Shell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Homer Simpson}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678-2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

\maketitle

\begin{abstract}

The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content. This paradigm allows to synchronize, structure and navigate communication content integrated into voice and video calls.

Voice and image together can express emotions like no other medium can. With hypermedia concepts, we can add more value to conference calls.

\emph{WebRTC} technology allows real-time communications between web browsers without the need to install additional software. The nature of web browser applications already follows the hypermedia concept, which makes \emph{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
The web browser platform provides an abstraction layer that makes it possible to create applications that run independently from the operating system.
The native support for \emph{WebRTC} in operating systems extends its usage to outside the web browser, allowing for the exploration of functionalities for which web browsers provide poor support, such as video recording and massive information storage.

Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging and a mechanism to superimpose hyper-content to video.

In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose and implement an architecture for an hyper-linked communication application based on \emph{WebRTC}. This work was evaluated by users, who reported that they liked to use it and thought it to be extremely innovative.

\end{abstract}

\IEEEpeerreviewmaketitle

\section{Introduction}
\label{chapter:introduction}

\subsection{Background}
\label{section:background}

	As communications technologies appeared, we adapted the way we communicate. The purpose of this project is not the replacement of the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.

	With the advent of WebRTC and its successive integration with web browsers, it became possible to develop video conference web applications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.

    Furthermore, real-time communication applications can make a significant difference on business, education and health sectors by providing tools for developing teaching and learning online, teamworking and socializing web applications.

\subsection{Proposed Solution}
\label{section:proposed}

	Our goal in this project is to develop an application targeted to the web platform, resorting to \gls{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \gls{HTML}5, \gls{CSS}3 and \emph{JavaScript} including continuous media types such as video, music and animations.

	One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again or introduce hyper-content to it such as time annotations, interactive lists of topics and subtitles. In this context we also provide a simpler method for creating and synchronizing hyper-content using \emph{QR codes}.

	In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links and files to conference participants.

	Furthermore, another relevant feature is the possibility to compose multiple video streams into a single one, which enables adding more users to conference rooms without impacting on clients performance. Users can change to individual streams on demand or automatically to the talking users.

\subsection{Thesis Contribution}
\label{section:contribution}

Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g. creating time annotations for improving content search) in order to increase its value. It is also important to highlight another goal of this project which is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.

	We have presented an architecture that can meet our goals, implemented the respective prototype and tested it with real users and performance benchmarks.

	According to Martin Geddes, the quality of the interaction worsens as the number of users increase\cite{geddes}. In our testing phases we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.

	All the problems faced during the development and limitations were reported on the thesis so that a future project better then ours can be easily and better developed.

\subsection{Outline}

This rest of this document is structured as follows:

\begin{itemize}

\item \textbf{Chapter \ref{chapter:relatedwork}} describes the previous work in the field.
\item \textbf{Chapter \ref{chapter:architecture}} describes the system requirements and the architecture for an Web Application that fulfills the goals of this thesis.
\item \textbf{Chapter \ref{chapter:implementation}} describes the implementation of our Web Application and the technologies chosen.
\item \textbf{Chapter \ref{chapter:evaluation}} presents the evaluation tests performed and the corresponding results.
\item \textbf{Chapter \ref{chapter:conclusion}} summarizes the work developed and proposes future work.
\end{itemize}

\section{Related Work}
 This section is structured as follows.
 Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \gls{IPv4} address exhaustion and the client server model constraints. 
 Section \ref{rtc} describes the \gls{WebRTC} technology and the protocols needed to implement our project. 
 Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \gls{WebRTC} specifications. 
 Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
 Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.

\subsection{Early days of the Internet and its remaining flaws}\label{early}

The need to build a global communications network in an age when almost nobody had access to that technology and the unpredictability of the number of future users, lead to some protocols not being suitable for the explosive growth and proliferation of users that followed. \gls{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}.

To this end, one way to overcome the \gls{IPv4} address scarcity problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \gls{IP} address with a different port.

This technique is known as \gls{NAT} (figure \ref{fig:nat}).

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{figures/nat.pdf}
	\caption{Network Address Translation}
	\label{fig:nat}
\end{figure}

Initially \gls{NAT} offered an alternative to address exhaustion and a minimal sensation of security.
However, due to their current wide usage, \gls{NAT}s weaknesses are being exposing at the application layer, namely impacting applications that require direct communications between two private networks.

There are four types of \gls{NAT} implementations\cite{rfc3489}: \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.

\emph{Full Cone} \gls{NAT} maps each public \gls{IP} address and port to a private \gls{IP} address and port.
Any external host can communicate with private hosts through their mapped public address and port. This represents the least restrictive type of \gls{NAT} and, as we will see later, the unique type of \gls{NAT} that enables real time communications from point to point.

\emph{Restricted Cone} \gls{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host. With this type of \gls{NAT}, the private client can be contacted from any port of the same external host.

\emph{Port Restricted Cone} \gls{NAT} works in the same way as Restricted Cone \gls{NAT}, but it only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.

\emph{Symmetric} NAT maps different ports for each connection. As we will see later, this type of \gls{NAT} represents a problem on real time communications.

\emph{Non-Symmetric} \gls{NAT}s became the common configurations on the Internet. As a direct result, problems started to appear: the amount of ports that \gls{IP} makes available is also small compared to our current needs; worse than that, \gls{NAT} also difficults end-to-end communication, forcing most applications that follow this model to be implemented ineffectively.

Unless the router that performs as \gls{NAT} has forwarding rules to every desired ports of each user, applications behind a \gls{NAT} are prevented from receiving incoming connections from the public network, which forces them to behave as a client of a client-server model.

\gls{STUN} and \gls{TURN} \cite{natvoip} servers are a possible solution to overcome \gls{NAT}.

\gls{STUN} servers are quite simple. They receive requests from \gls{NAT}ed clients, with the source address of a request being the public address that \gls{NAT} mapped to the client. \gls{STUN} servers will then reply to the client, providing the mapped public address, so it knows its associated public \gls{IP} address and port. Symmetric \gls{NAT} changes \gls{IP} port for each different connection, for that reason, when the \gls{STUN} servers reply with the \gls{IP} address and port of their connection, it will be useless for clients to use them on other application connections. That is why Symmetric \gls{NAT} represents a problem for peer-to-peer communications.

On the other hand, \gls{TURN} uses public servers to relay traffic between private endpoints.
It may use a \gls{P2P} network relay to find the best peer, but after that, the behavior is much like client-server. Direct communication is only achieved by \gls{STUN} when \gls{NAT} is a type \emph{full cone}. \gls{ICE} is a technique that uses \gls{STUN} when direct communications are possible and \gls{TURN} while a direct communication isn't possible.

When connection is established, either in a direct or indirect way (via \gls{TURN} servers), \gls{WebRTC} came to simplify how audio and video are transmitted through web browsers.

\subsection{Real time communications}\label{rtc}

\gls{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \gls{API}s for web browser based real time communications without installing any additional application or plug-in. Table \ref{table:webrtcstack} shows the protocols that \gls{WebRTC} rely on.

Some operating systems such as \emph{Android}, \emph{iOS}, \emph{Linux}, \emph{OSX} and \emph{Windows} implement native \gls{WebRTC} libraries, extending the usage of \gls{WebRTC} to applications outside the web browser. This native support can help to implement applications that record video and audio streams for further playback.

\begin{table}
\centering
\caption{WebRTC protocol Stack}
\label{table:webrtcstack}
    \begin{tabularx}{\linewidth}{*8{Y}@{}m{0pt}@{}} 
    	\cline{5-8}
		&&&\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{MediaStream} & \multicolumn{2}{c|}{DataChannel} 								&\\[12pt]
		\hline 
		\multicolumn{1}{|c|}{XHR} & \multicolumn{1}{c|}{SSE} & \multicolumn{2}{c|}{WebSocket} & \multicolumn{2}{c|}{SRTP} & \multicolumn{2}{c|}{SCTP} 				&\\[12pt]
		\cline{1-3}\cline{6-8} 
		\multicolumn{3}{|c|}{HTTP 1.X/2.0} & \multicolumn{1}{c|}{} & \multicolumn{1}{c}{} & \multicolumn{3}{|c|}{Session (DTLS)} 	&\\[12pt]
		\hline
		\multicolumn{4}{|c|}{Session (TLS) - optional} & \multicolumn{4}{c|}{ICE, STUN, TURN} 																		&\\[12pt]
		\hline
		\multicolumn{4}{|c|}{Transport (TCP)} & \multicolumn{4}{c|}{Transport (UDP)} 																				&\\[12pt]
		\hline
		\multicolumn{8}{|c|}{Network (IP)}  																														&\\[12pt]
		\hline
	\end{tabularx}

\end{table}

\gls{WebRTC} defines three main \gls{API}s: MediaStream, PeerConnection and DataChannel.

\begin{itemize}
  \item \textbf{MediaStream} allows the browser to access the camera, microphone and the device's screen.

  \item \textbf{PeerConnection} acquires connection data and negotiates with peers.

  \item \textbf{DataChannel} provides a channel for exchanging arbitrary data with other peers.
\end{itemize}

\gls{WebRTC} uses \gls{UDP} for transporting data, which provides lower latencies than \gls{TCP}, but is not reliable and does not assure packet order and integrity. \gls{SCTP} and \gls{SRTP} are used for streaming data, providing a mechanism for congestion control and partial reliable delivery over \gls{UDP}. All transferred audio, data and video must be encrypted with \gls{DTLS} symmetric keys. \gls{DTLS} provides the same security guarantees as \gls{TLS}.

\gls{TLS} doesn't support independent packet decryption\cite{rfc6347}, for that it requires a reliable transport channel, typically \gls{TCP}. The decryption of a packet depends on the previous packet, which for unreliable transport protocols like \gls{UDP} may represent a problem, either due to packet loss or different reception order.

\gls{DTLS} is similar to \gls{TLS}, but is used on top of \gls{UDP}.

\gls{WebRTC}'s \emph{MediaStream} is built on top of \gls{SRTP}, which requires an external mechanism for key exchange. \gls{DTLS} keys are negotiated on handshake in order to achieve a secure connection. The new keys derived from \gls{DTLS} handshake are seized for \gls{SRTP} encryption, the remaining \gls{SRTP} communications are done through \gls{UDP} without using \gls{DTLS}.

\gls{WebRTC} aims to provide a standard platform for real-time audio and video on the Web. It arrives at a time when several proprietary products are well established.

\subsection{Signaling: meet and get to know}
\label{signaling}

  Signaling is the process by which applications exchange connection information about peers and servers, their capabilities and meta-data.
  In particular, \gls{WebRTC} does not implement signaling, as different applications may require different protocols and there is no single answer that fits all problems.
  As a consequence, multiple options are available for filling the missing \gls{WebRTC}'s signaling component, which can be performed using \gls{SIP}, \gls{XMPP}, \emph{WebSockets}, \emph{Socket.io}\footnote{\url{http://socket.io/}(accessed June 1, 2015).} or by implementing a custom protocol.

  \subsubsection{WebRTC}

  \gls{WebRTC} uses \gls{SDP} \cite{rfc4566} to define peer connection properties such as types of supported media, codecs, protocols used and network information. An \gls{SDP} offer describes to other peers the expected type of communication and its details, such as used transport protocols, codecs, security and other.

  One of \gls{WebRTC} signaling's requisites is bi-directional communication.

  \gls{HTTP} uses a request-response paradigm, where a request is sent by the client, followed by a server response.

  Sometimes it is required that some information be obtained in real time, but we saw, some \gls{NAT}'s do not support callbacks from servers, preventing them from notifying clients as soon as an event occures.
One technique to overcome this problem is long polling which consists on making the server hold the request until there is fresh information or expiring it after some time.
As soon as it receives the reply, the client makes another request. Long polling technique results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \gls{HTTP} headers, which add data overhead and can be noticed especially for short sized messages.

The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel \cite{rfc6455}, by other words it supports sending and receiving data simultaneously.

  \subsubsection{Session Initiation Protocol}

  \gls{SIP} \cite{rfc3261} is a protocol used for negotiation, creation, modification and finalization of communication sessions between users. \gls{SIP} follows a client/server architecture with \gls{HTTP} like messages and it can be used as a signaling protocol. The advantage of \gls{SIP} is the ability to make video and voice calls between the telephone network and applications over \gls{IP} networks.

  \gls{SIP} is used in \gls{VoIP} applications due to its compatibility with the \gls{PSTN}.

  Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}(accessed June 1, 2015).}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}(accessed June 1, 2015).} and \emph{sipML5}\footnote{\url{http://sipml5.org/}(accessed June 1, 2015).} are used on the client side to parse and encode \gls{SIP} messages, making \gls{SIP} accessible to web based applications.

  \gls{SIP} with \emph{WebSockets} can be used as a signaling method for \gls{WebRTC} applications, it allows web browsers to have audio, video and \gls{SMS} capabilities like mobile phones. For instance, it's possible to inter-operate web communications with \gls{SIP} networks, mobile and fixed phones.

\subsubsection{Extensible Messaging and Presence Protocol}

  \gls{XMPP} was initially developed for instant messaging and presence (Jabber\footnote{\url{http://jabber.org/}(accessed June 1, 2015).}). It is nowadays an open technology for standardized, decentralized, secure and extensible real-time communications.

  Today, multiple \gls{XMPP} server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}(accessed June 1, 2015).}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}(accessed June 1, 2015).}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}(accessed June 1, 2015).} and \emph{Prosody}\footnote{\url{http://prosody.im/}(accessed June 1, 2015).}. \emph{Ejabberd} is the server that implements more \gls{RFC} specifications and \gls{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}(accessed June 1, 2015).}.

    The connection to the \gls{XMPP} server can be done through the same web server as the user is connected, but this web server would handle a lot of connections and that would have a great impact on the overall system performance. On the other hand, users can directly connect directly to \gls{XMPP} servers from web applications using the \emph{JavaScript} library \emph{strophe.js} \footnote{\url{http://strophe.im/strophejs/} (accessed May 6, 2016).}

    Typically a web application consists in multiple web pages which are navigated by users. Each time a web page is accessed, either from a new context or a transition from a previous page, its context is cleared except for local storage and cookies. The \emph{JavaScript} context is cleared, including the \gls{XMPP} connections performed by \emph{strophe.js}. As such, an automatic mechanism is required for avoiding the implicit user reauthentication.
    One solution for reauthentication can be achieved by storing the user's JabberID and password on local storage and every time a page is accessed the authentication is performed without the users knowledge. Clearly this solution represents security flaws as the local storage can be easily accessed locally or by performing a cross site scripting attack to reveal all the local storage. The same problem arises if the credentials are stored on cookies.

    An alternative solution is known as Session Attachment \footnote{\url{https://metajack.im/2008/10/03/getting-attached-to-strophe/} (accessed May 6, 2016).}, which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} be passed to \emph{strophe.js} in order to re-connect to the same stream on \gls{XMPP} server. Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request making it worthless if a user maintains more than one tab opened, for example for multiple conversations at the same time.

\subsection{Hypermedia: more than words, more than images}
\label{hypermedia}

Since the early days of video technology, one of the problems raised consisted on how to add more information onto video without generating multiple versions. This section examines technologies that allows different ways to present multimedia content in such a way that it change based on synchronization amongst other multimedia elements or user interaction.

\emph{Hypermedia} concept brings the possibility to organize and overlay multimedia elements into a nonlinear linear structure holding the promise of future technology and features.

  Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it. An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time in a video or audio track. \emph{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe.

  In order to navigate through a dynamic video, we must be aware of time synchronization and the multiple time flows, it ils important that all time, causality and behavior rules are well defined.

  \emph{HyVAL}\cite{hyval} is an \gls{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia. HyVAL defines defines video structure, internal video and external media objects.

  \gls{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films. With \gls{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.

The \gls{DOM} is a standard \gls{API} that allows easy management of documents that are organized in a tree structure, by providing \gls{CRUD} operations over its elements and their attributes. \gls{DOM} makes it easy to inter-operate between imperative and declarative programming languages\cite{dom}.

Like \gls{DOM}, \gls{SMIL} \gls{DOM} is an \gls{API} for \gls{SMIL} documents. Allowing \gls{CRUD} operations over \gls{SMIL} documents is an important feature for extending \gls{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \emph{JavaScript} functions.

  \gls{SMIL} fits our goals for creating a multimedia rich hyper-call, but it lacks on browser compatibility. Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \gls{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plug-in.

  \emph{JavaScript} is an imperative object-oriented language based on \emph{ECMAScript}. It is used mainly on client-side and executed by a web browser. \emph{JavaScript} has its own implementation of \gls{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plug-ins.

  \emph{JavaScript} has compatibility issues among the different web browsers, leading to different behaviors. To solve that problem, there are libraries written in \emph{JavasSript}, namely \emph{jQuery}, that implements the same functionality for multiple browsers, masking most of the incompatibility issues.

  \gls{SVG} is an \gls{XML} based format that incorporates the animation module of \gls{SMIL}. Currently, \gls{SVG} allows adding movement and animating attributes of elements. When embedded on \gls{HTML}5, it allows dynamic changes to inner content in real-time through the \gls{DOM} \gls{API}. Besides that, it also allows calling \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.

  Using technologies that relies only on web standards, like \gls{CSS}, \gls{HTML}5, \emph{JavaScript} and \gls{SVG}, will make possible to develop an application that solves our problem with the advantage of being compatible with a greater amount of web browsers.

\subsection{Extending collaboration tools with time manipulation}
\label{collab}

Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity. Technologies like these are appearing along these days, but they were not be possible a few years ago because technology was limited or unavailable. Although today's technology is still limited on some aspects, progress is being done in order to improve the web ecosystem, by creating standards and migrating to newer technologies.

Section \ref{recstream} presents the RTP protocol how it can be used to record media streams. Section \ref{mediatype} describes the media types and what types of media should be streamed. Section \ref{intrecord} addresses how an interactive media can be recorded. Section \ref{collabenv} presents the collaborative environment and libraries to synchronize distributed object among users.

\subsubsection{Streaming and Recording}\label{recstream}~\\

 Our first concern on real time collaboration applications, besides the communication itself, is the data storage and representation. Storing multimedia content is not a viable solution because most browsers recommend limiting local storage to at most five megabytes per origin.

 	In order to provide a way to record and playback streams, additional servers will be required to process and record the large amount of data generated by audio and video streaming.

 \gls{RTP}\cite{rfc3550} is used for streaming audio and video over \gls{IP}.
 Multimedia content is transported on the payload of \gls{RTP} messages, that contain headers for payload identification. \gls{RTP} is independent from its payload type, allowing it to transport any kind of encoded multimedia. A sequence number is used for sorting received packets.

  \gls{RTP} allows to change its requirements and add extensions to it with profiles. One of the most used ones is the \gls{RTP} profile for audio and video \cite{rfc3551}, which lists the payload encoding and compression algorithms. 
  Another profile for \gls{RTP} is defined by \gls{SRTP}, which provides encryption, authentication and replay protection for \gls {RTP} traffic. The analogous secure protocol for \gls{RTCP} is \gls{SRTCP}.

  \gls{RTP} recorders are independent of payload encoding, they don't decode \gls{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.

	\gls{KMS} supports streaming over \emph{WebRTC}, \gls{HTTP} and \gls{RTP} endpoints. Another important component of \gls{KMS} is \emph{Kurento Repository}, which supports recording and playing directly from \emph{MongoDB}. That is important for providing a scalable media storage. Unlike \emph{Jitsi Video Bridge}\footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}(accessed June 2, 2015).}, \gls{KMS} does not enforce a specific signaling protocol.

\subsubsection{Media Types}\label{mediatype}~\\
	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuous, the second one describes it as interactive or non-interactive. A discrete media is characterized by not depending from time, and continuous media as depending on it. Interactive media is characterized by its state being changed by external events such as user interactions.

\begin{figure}
	\centering
	\includegraphics[width=0.3\textwidth]{figures/media_types.png}
	\caption{Media Types}
\end{figure}
	For example, an image is non-interactive and discrete, while a video is continuous and non-interactive. A simple collaborative editor with just text is interactive and discrete. An animation that changes in function of user behavior is interactive and continuous.

  \subsubsection{Collaborative Environment}\label{collabenv}~\\

        \gls{OT} technology was originally developed for consistency maintenance and concurrency control over distributed objects.
        \gls{OT} algorithms are mainly used in collaborative applications such as distributed document edition.

	Among mutiple \gls{OT} platforms and libraries we present \emph{ShareJS}\footnote{\url{http://sharejs.org/}(accessed June 2, 2015).}, \emph{TogetherJS}\footnote{\url{http://togetherjs.com/}(accessed June 2, 2015).}, \emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}(accessed June 2, 2015).}, \emph{Etherpad Lite}\footnote{\url{https://github.com/ether/etherpad-lite}(Accessed 20 March 2016)} and \emph{otJS}\footnote{\url{http://operational-transformation.github.io}(accessed March 10, 2016)} which we describe on Table \ref{table:otcomparision}.

\begin{table}
\centering
\caption{Comparision between Operational Transformation libraries}
\label{table:otcomparision}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Library} & \textbf{Own Server} & \textbf{Own Storage} & \textbf{Operations} \\ \hline
ShareJS          & yes                 & yes                  & text+objects        \\ \hline
TogetherJS       & yes                 & no                   & text+objects        \\ \hline
Goodow           & yes                 & yes                  & text+objects        \\ \hline
Etherpad Lite    & yes                 & yes                  & extendable 			    \\ \hline
OT.js            & no                  & no                   & text                \\ \hline
\end{tabular}
\end{table}

	\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations over plain text on the client side. An implication of implementing just the client side is the extra effort that is necessary to implement content's persistent storage. Besides this drawback, this library is very flexible because it's not tied to a specific database or server side technology.

\subsection{Chapter Summary}
\label{related:summary}

In order to implement an hyper-linked communication solution, several design decisions had to be made. The limitations imposed by the \emph{Internet's} structure, its protocols and a browser's capabilities are key factors to consider when implementing a solution that allows bi-directional communications, interactive media and collaboration environment.

Due to the use of \gls{NAT}, bi-directional communications between clients have different needs from request-response based communications between clients and servers.
This lead to the appearance of mechanisms such as \gls{STUN}, \gls{TURN} and \gls{ICE}, in order to bypass the limits imposed by \gls{NAT}.

\gls{WebRTC} introduced an \gls{API} for video, audio and data communications through web browsers without using plug-ins.
However, \gls{WebRTC} by itself does not define how users get to know each other nor how information flows between users. For this reason, we have studied the multiple ways we could implement this \emph{get-to-know} mechanism which is known as signaling protocol.

With the communications establishment issue solved, we had to discuss the different types of media and what can be done with each kind in order to increase the value of communications among users. In this context, we have studied solutions and libraries that allow us to implement our prototype with time manipulation features, collaborative text edition, record and playback interactive video.

\section{Conclusion}
The conclusion goes here.

\ifCLASSOPTIONcompsoc

  \section*{Acknowledgments}
\else

  \section*{Acknowledgment}
\fi

The authors would like to thank...

%Sets the bibliography style to UNSRT and imports the 
%bibliography file "samples.bib".
%\bibliographystyle{unsrt}
%\bibliography{references}

\input{thesis.bbl}
\end{document}

### 2016-05-13 tese2/conclusions.tex
[ADD] final version of statistics

@@
 \label{section:summary}
-In this thesis, our goal was the development of a web application using \emph{WebRTC} that complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time.
+In this thesis, our goal was the development of a web application using \emph{WebRTC} that complements current audio, text and video communications, in order to create rich and collaborative interfaces with the ability to add more content on a future time.
 Another important goal of this project is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.
 
-Having this in mind, we started to analyze the technologies necessary to implement our solution.
-In a first glance, we have analyzed the problems that real time communication applications are facing when using the current Internet infrastructure.
-Next we have we have analyzed the main components of the \emph{WebRTC} stack.
-Furthermore, we have analyzed how we could implement the signalization component that \emph{WebRTC} does not not define.
-Hence, with the transport component being analyzed, we have studied how we could expose content in a synchronized and interactive way.
-Moreover, we have performed an analysis on the kinds of media types the particularities of each one when manipulating time.
-Lastly, on the state of the art context, we have analyzed existing frameworks and libraries that could help us implementing a collaborative environment.
+Having this in mind, we started by analyzing the technologies necessary to implement our solution.
+In a first glance, we have analyzed the problems that real time communication applications face when using the current Internet infrastructure.
+Next, we analyzed the main components of the \emph{WebRTC} stack.
+Furthermore, we analyzed how we could implement the signaling component that \emph{WebRTC} does not not define.
+Hence, with the transport component analyzed, we have studied how we could expose content in a synchronized and interactive way.
+Moreover, we have performed an analysis on the kinds of media types ant the particularities of each one when manipulating time.
+Lastly, on the state of the art context, we analyzed existing frameworks and libraries that could help us implementing a collaborative environment.
 
@@
 
-Although we have used our own signaling protocol on our implementation, the first steps we have made during the development of our prototype were using \ac{XMPP}, we have discovered the drawbacks of this approach in practice and we moved our learnings to the state of the art.
-Besides starting our prototype using this approach, we have used \emph{strophe.js} for performing communications with \ac{XMPP} servers.
-We have found and fixed a bug present on \emph{strophe.js} library that silenced an error raised by duplicated user registrations.
-
-As a consequence of not using \ac{XMPP} for the signaling protocol, we have defined our own protocol based on \emph{WebSockets} for performing the stream between clients and \ac{KMS}.
-From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \emph{QR code} detection for creating hyper content.
+Although we have used our own signaling protocol we have made our first steps using \ac{XMPP} and we have discovered the drawbacks of this approach.
+Dor performing communications with \ac{XMPP} servers, we used \emph{strophe.js} which we found to have a bug that we have fixed.
 
@@
 \label{section:future}
-	Playing back video with a faster rate is not possible using the current version of \ac{KMS} which we expect the availability of that feature in a near future but we have proposed an alternative way to implement faster playback by using \emph{ffmpeg} to convert the video before playing it.
+	Playing back video with a faster rate is not possible using the current version of \ac{KMS}.
+Even though we expect the availability of that feature in a near future but we have proposed an alternative way to implement faster playback by using \emph{ffmpeg} to convert the video before playing it.
 
 	Although we have tested our solution in a powerful machine for the current time, our performance tests revealed that the streaming component uses a lot of resources.
-We left for a future work a deep analysis on the scalability of our system which we have proposed different approaches but we have not tested them.
+We left for a future work a deep analysis on the scalability of our system for which we have proposed different approaches but we have not tested them.
 

### 2016-05-13 tese2/evaluation.tex
[ADD] final version of statistics

@@
    Figure \ref{fig:test_hour_mem} shows the memory used by our solution with seven clients simultaneously on the same conference room during one hour.
-As we can observe \ac{KMS} memory usage stabilizes.
+As we can observe, \ac{KMS} memory usage stabilizes.
 
@@
 
-      In order to the usability of our solution, we haver performed usability tests with the help of real users with different backgrounds and ages.
+      In order to evaluate the usability of our solution, we have performed usability tests with the help of real users with different backgrounds and ages.
 
-      We handled a guide to the users with five tasks to perform, the metrics we used for each task were: number of clicks, number of errors including a description and time spent.
+      We handed a guide to the users with five tasks to perform. The metrics we used for each task were: number of clicks, number of errors (including a description) and time spent.
 
-      The tasks we asked to the user are:
+      The tasks we asked the user to perform were:
 
       \begin{enumerate}
-      \item Login into the system with the provided credentials and accept the received friendship request. Find the coordinator's brother and add him as a friend.
+      \item Login into the system with the provided credentials and accept the received friendship request. Find a given user by its name and add him as a friend.
 
@@
 
-      \item Enter in the conference room correspondent to the third task as an observer and navigate to the specified annotation.
+      \item Enter a conference room correspondent as an observer and navigate to the specified annotation.
 Watch the video until a list of topics is overlayed in the video and choose one of them and leave the conference room.
 
-      \item Enter in the conference room correspondent to the fourth task by sharing your camera and create a time annotation in the instant of time when you entered, navigate to the current time, search for the annotation and leave the conference room.
+      \item Enter a conference room by sharing their camera and create a time annotation in the instant of time when they entered, navigate to the current time, search for the annotation and leave the conference room.
 
-      \item Enter in the conference room correspondent to the fourth task by sharing your camera and add a subtitle in the video, preview it, specify an interval of time and save it.
+      \item Enter a conference room by sharing their camera and add a subtitle in the video, preview it, specify an interval of time and save it.
 Then show the provided \emph{QR code} to the camera and leave the leave the conference room.
@@
 
-      The goal of the first task consists on evaluating the interface for authentication and friendship management.
+      The goal of the first task is to evaluating the interface for authentication and friendship management.
 
-      The second task is used to make the users familiar with our tools for collaborative content edition.
+      The second task is used to evaluate our tools for collaborative content edition.
 
@@
 
-      The goal of the fifth task is to create synchronized overlayed content, demonstrate the difficulties of such task and present an easier way to synchronize content.
+      The goal of the fifth task is to create synchronized overlayed content and present an easier way to synchronize content.
 
@@
 The first time we tested our solution by providing tasks to users, we have observed that our solution was not perfect.
-We have faced usability problems during our tests, he had to improve our solution's usability and start the tests again.
+Having faced usability problems during our tests, he had to improve our solution's usability and start the tests again.
 
@@
 
-In the first testing phase we have performed just three tests with users and ceased for improvements.
+In the first testing phase we have performed tests with just three users and ceased for improvements.
 
-Both users that tested our solution had difficulties on the first task when finding new friendship requests and adding new friends, as friendship request were possible to access only from the top bar and new user profiles were only accessible through the search bar.
+The initial users that tested our solution had difficulties in the first task when finding new friendship requests and adding new friends, as friendship request were only possible to access from the top bar and new user profiles were only accessible through the search bar.
 In order to solve this problem we have added a menu for adding new users in the friends list and another near the friends list as it can be seen on figure \ref{fig:test_ui_01_02_03}.
@@
 
-On the second task one user clicked on the friends list for adding a friend to room and by consequence made him leave the conference room to the user profile.
+On the second task one user clicked on the friends list for adding a friend to a room and, as a consequence, left the conference room to the user profile.
 We solved this problem by showing a pop-up with the user profile and an additional button to add the user to the conference room as it can be seen on figure \ref{fig:test_ui_04}.
@@
 
-On the third task we noticed that both users were not expecting to search the content directly from the search bar, instead they used the time-line.
+On the third task we noticed that users were not expecting to search the content directly from the search bar, instead they used the time-line.
 
-On the fourth task both users didn't placed their annotations on right place, we only provide one way to place annotations, which is by dragging them on the time-line. 
-After creating an annotation users also didn't understand that they had to save their changes on the time-line, we solved this problem by saving automatically each time the user changed any annotation, the old user interface can be seen on figure \ref{fig:test_ui_05}.
+On the fourth task the users didn't placed their annotations on right place. We only provide one way to place annotations, which is by dragging them on the time-line. 
+After creating an annotation, users also didn't understand that they had to save their changes on the time-line.
+We solved this problem by saving automatically each time the user changed any annotation.
+The old user interface can be seen on figure \ref{fig:test_ui_05}.
 
@@
 
-Both users had success with fifth task but they revealed difficulties choosing the time interval for the hyper content, we are aware that this is a difficult task and that's why we have introduced the \emph{QR code} content creation mechanism which was very easy to learn.
+The three users had success with the fifth task, but they revealed difficulties choosing the time interval for the hyper content.
+We are aware that this is a difficult task and that's why we have introduced the \emph{QR code} content creation mechanism which was very easy to learn (with a \emph{QR code} already printed into paper).
+
+Although content creation through \emph{QR codes} revealed much easier to perform, this approach implies showing and, as a consequence, recording video with the \emph{QR codes}, making them visible to other users.
+In order to solve this problem, we could add a new video sharing mode that allow users to send video to \ac{KMS} without composing, recording and showing to other users.
 
@@
 \begin{itemize}
-\item Giving feedback to user when saving content and collaborative editors.
-\item Support file upload on our chat.
-\item Starting with our time-line more zoomed in order to give more sensation of time.
-\item Create content with starting time and duration instead ending time.
+\item Giving feedback to user when performing operations such as create superimposed content and save the collaborative editor content.
+\item Starting with our time-line more zoomed in order to give more sensation of time passing.
+\item Create content with starting time and duration, instead of ending time.
 \end{itemize}
@@
 
-On the second phase of our user interface tests, in a general view, we have noticed great improvements on the learning time.
+On the second phase of our user interface tests, in a general way, we have noticed great improvements on the learning time.
 
@@
 
-From the data that we have collected, namely the task duration (figure \ref{fig:user_times}), number of clicks (figure \ref{fig:user_clicks}), number of errors (figure \ref{fig:user_errors}), task difficulty (figure \ref{fig:user_diffs}) and qualitative evaluation (figure \ref{fig:user_evals}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.
+From the data that we have collected, namely the task duration (Figure \ref{fig:user_times}), number of clicks (Figure \ref{fig:user_clicks}), number of errors (Figure \ref{fig:user_errors}), task difficulty (Figure \ref{fig:user_diffs}) and qualitative evaluation (Figure \ref{fig:user_evals}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.
 
@@
 
-Where $\alpha=0.05$, $n$ is the number of samples (which may not coincide with the number of tests due to failed tasks), $\bar{x}$ is the sample mean, $x_i$ is the sample value and $s$ is the sample standard deviation that is given by:
+Where $\alpha=0.05$ \footnote{95\% ($\alpha=0.05=1-0.95$) is a common value, used by researchers, to calculate confidence intervals}, $n$ is the number of samples (which may not coincide with the number of tests due to failed tasks), $\bar{x}$ is the sample mean, $x_i$ is the sample value and $s$ is the sample standard deviation that is given by:
 \begin{equation}
@@
 
-Accordingly to figure \ref{fig:user_evals}, we can observe that most users had less difficulties with the first two tasks, which represents types of tasks that most users are familiar with.
-As soon as users had to navigate in time, manipulate annotations and create content (respectively \emph{task3}, \emph{task4} and \emph{task5}) we have observed that they revealed more difficulties.
+According to Figure \ref{fig:user_diffs}, we can observe that most users had less difficulties with the first two tasks, which represents types of tasks that most users are familiar with.
+As soon as users had to navigate in time, manipulate annotations and create content (respectively \emph{task3}, \emph{task4} and \emph{task5}) we observed that they revealed more difficulties.
 Most of those difficulties, based on the users feedback, were mainly due to those concepts not being familiar to them.
@@
 
-Although we had relatively bad results with some users, we have explained to the users that could not conclude the tasks or performed incorrectly the most efficient way to perform the requested tasks.
-Some users suggested to display more hints in order to achieve a faster learning but after all they were impressed and gave us a better evaluation (as seen on figure \ref{fig:user_evals}).
+As we had relatively bad results with some users, we explained to those users that could not conclude the tasks or performed them incorrectly, the most efficient way to perform the requested tasks.
+Some users suggested to display more hints in order to achieve a faster learning, but afterwards all were impressed and gave us a better evaluation (as seen on Figure \ref{fig:user_evals}).
 
@@
 
-Most users gave us worse evaluations on our user interface layout and content editor, which was due to having a lot of tools present in the same web page and some of them were hidden due the screen size.
+Most users gave us worse evaluations on our user interface layout and content editor, which was due to having a lot of tools present in the same web page and some of them being hidden due the screen size.
 In some cases users had to scroll down in order to find the tools they were looking for.
 
-Another weak aspect was our content editor, which in fact we recognize the difficulty to handle with, most due to the amount of information that is necessary to create a synchronized content (starting time, duration and content itself).
+Another weak aspect was our content editor, which, in fact, we recognize is difficult to work with, mostly due to the amount of information that is necessary to create a synchronized content (starting time, duration and content itself).
 Some users have suggested that the content should also be present on the time-line so they could be easily dragged and resized (on time).
 
-We are aware that placing content on the time-line will reduce our solutions performance, especially when there is a relatively big amount of content due to the content that is present on the time-line being loaded all in once.
-Although, we recognize that for some cases (relatively low amount of content), displaying the content on the time-line could not have a great impact on our solutions performance.
+We are aware that placing content on the time-line will reduce our solutions performance, especially when there is a relatively large amount of content, due to the content that is present on the time-line being loaded all at once.
+Although, we recognize that for some cases (relatively low amount of content), displaying the content on the time-line could not have a great impact on our solutions performance, we choose not to implement this.
 
-In conclusion, 100\% of our testers though that our solution was an innovation and 92.31\% recommended using our solution.
+In conclusion, 100\% of our testers though that our solution was an innovation and 95\% recommended using our solution.
 
@@
 The performance tests revealed unexpected results which were crucial to find and solve performance issues.
-Another important aspect of our performance test results, were the understanding of how data flows between our system modules.
+Another important aspect of our performance test results was the understanding of how data flows between our system modules.
 
@@
 In order to understand how good the results were for each performed task, we have defined optimal values for the different metrics.
-In general the results reveal a significant difference between observed values and optimal values which can be explained by the users lack of experience with systems such as we have created.
+In general the results reveal a significant difference between observed values and optimal values, which can be explained by the user's lack of experience with systems such as the one we have created.

### 2016-05-13 tese2/implementation.tex
until 1h test case
[ADD] final version of statistics

@@
 \input{13_implementation_model.tex}
+
+\section{XMPP experiments}
+Although we have used our own signaling protocol on our implementation, the first steps we made during the development of our prototype were using \ac{XMPP}.
+We discovered the drawbacks of this approach in practice and described our learnings in the state of the art.
+We have found and fixed a bug present on \emph{strophe.js} library that silenced an error raised by duplicated user registrations.
+
+As a consequence of not using \ac{XMPP} for the signaling protocol, we have defined our own protocol based on \emph{WebSockets} for performing the stream between clients and \ac{KMS}.
+From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \emph{QR code} detection for creating hyper content.
+
 \input{14_implementation_signaling.tex}
@@
 \input{16_implementation_content.tex}
-\input{17_implementation_deploy.tex}
 \input{18_implementation_summary.tex}

### 2016-05-13 tese2/evaluation.tex
gonna sleep

@@
         a) Friend list path
-\label{fig:minipage1}
 \end{minipage}
@@
          b) Friend requests path
-\label{fig:minipage2}
 \end{minipage}
@@
          c) Search bar path
-\label{fig:minipage2}
 \end{minipage}
@@
 
-From the data that we have collected, namely the task duration (Figure \ref{fig:user_times}), number of clicks (Figure \ref{fig:user_clicks}), number of errors (Figure \ref{fig:user_errors}), task difficulty (Figure \ref{fig:user_diffs}) and qualitative evaluation (Figure \ref{fig:user_evals}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.
+From the data collected with twenty tests with users, namely the task duration (Figure \ref{fig:user_times}), number of clicks (Figure \ref{fig:user_clicks}), number of errors (Figure \ref{fig:user_errors}), task difficulty (Figure \ref{fig:user_diffs}) and qualitative evaluation (Figure \ref{fig:user_evals}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.
 
@@
 
-Where $\alpha=0.05$ \footnote{95\% ($\alpha=0.05=1-0.95$) is a common value, used by researchers, to calculate confidence intervals}, $n$ is the number of samples (which may not coincide with the number of tests due to failed tasks), $\bar{x}$ is the sample mean, $x_i$ is the sample value and $s$ is the sample standard deviation that is given by:
+The youngest and oldest testers were, receptively, twenty two and thirty eight years old. Figure \ref{fig:user_ages} the ages of the users that tested our system.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{stats/user_ages.eps}
+  \caption{Ages of the users that tested our system}
+  \label{fig:user_ages}
+\end{figure}
+
+Where $\alpha=0.05$ \footnote{95\% ($\alpha=0.05=1-0.95$) is a common confidence level value used, by researchers, to calculate confidence intervals}, $n$ is the number of samples (which may not coincide with the number of tests due to failed tasks), $\bar{x}$ is the sample mean, $x_i$ is the sample value and $s$ is the sample standard deviation that is given by:
 \begin{equation}

### 2016-05-13 tese2/conclusions.tex
Reviewed conclusions

@@
 Next, we analyzed the main components of the \emph{WebRTC} stack.
-Furthermore, we analyzed how we could implement the signaling component that \emph{WebRTC} does not not define.
-Hence, with the transport component analyzed, we have studied how we could expose content in a synchronized and interactive way.
-Moreover, we have performed an analysis on the kinds of media types ant the particularities of each one when manipulating time.
+Furthermore, we analyzed how we could implement the signaling component that \emph{WebRTC} does not define.
+With the transport component analyzed, we then studied how we could expose content in a synchronized and interactive way.
+We also performed an analysis on the kinds of media types and the particularities of each one when manipulating time.
 Lastly, on the state of the art context, we analyzed existing frameworks and libraries that could help us implementing a collaborative environment.
@@
 Based on our evaluation of the state of the art, we have defined the architecture of our proposed solution.
+%RP podes falar um pouco da solução: aplicação web, servidor web para sinalização, KMS para transcoding e KS e MongoDB para guardar.
+App possibilita chat, multi-party, partilha de ficheiro, edição colaborativa.
+Funcionalidades inovadoras: tags, legendas, timeline com navegação, QR codes.
+Mistura de som, mistura de video.
+Trafego no client independente do numero de participantes.
 
-Although we have used our own signaling protocol we have made our first steps using \ac{XMPP} and we have discovered the drawbacks of this approach.
-Dor performing communications with \ac{XMPP} servers, we used \emph{strophe.js} which we found to have a bug that we have fixed.
+%RP avaliação, como nota positiva dos utilizadores.
+
+Although we ended up using our own signaling protocol, we made our first steps using \ac{XMPP} and we  discovered the drawbacks of this approach.
+For performing communications with \ac{XMPP} servers, we used \emph{strophe.js} which we found to have a bug that we have fixed.
 
@@
 
-\section{Achievements}
-\label{section:achievements}
+%\section{Achievements}
+%\label{section:achievements}
 	We have successfully implemented the basic functionalities of our prototype and spare some time for adding more valuable features such as the ability to create content by exposing \emph{QR codes} to the camera and lastly perform changes to the user interface in order to improve the quality of user's experience.
+%RP acho que estás a ser muito modesto. Tens de dizer que cumpriste todos (ou quase) os objectivos a que te tinhas proposto e ainda acrescentaste mais funcionalidades inovadoras
 
@@
 
-	Our usability tests show results that are considerably worse than the established optimal values due to our solution propose a different way to communicate that most people are not used to.
+	Our usability tests show results that are considerably worse than the established optimal values due to our solution proposing a different way to communicate that most people are not used to.
 Although we have obtained those results, in general our users gave us positive feedback and valuable advices which we have used to improve our system.
@@
 	Playing back video with a faster rate is not possible using the current version of \ac{KMS}.
-Even though we expect the availability of that feature in a near future but we have proposed an alternative way to implement faster playback by using \emph{ffmpeg} to convert the video before playing it.
+This is useful feature that enables users to go through large amounts of video in less time.
+Even though we expect the availability of that feature in a near future, we have proposed an alternative way to implement faster playback by using \emph{ffmpeg} to convert the video before playing it.
 
@@
 
-	Lastly, we have chosen functionality over security in respect to displaying content to users which lead to security flaws on our solution.
-Although we have not solved the security problems, we have proposed a solution which at the same time limits the flexibility of adding new functionalities to our prototype.
+	Lastly, we have chosen functionality over security in respect to displaying content to users, which lead to security flaws on our solution.
+Although we have not solved the security problems, we have proposed a solution that at the same time limits the flexibility of adding new functionalities to our prototype.
 \cleardoublepage

### 2016-05-13 tese2/conclusions.tex
[]

--- 2016-05-13 tese2/conclusions.tex
@@
 Although we have used our own signaling protocol we have made our first steps using \ac{XMPP} and we have discovered the drawbacks of this approach.
-Dor performing communications with \ac{XMPP} servers, we used \emph{strophe.js} which we found to have a bug that we have fixed.
+For performing communications with \ac{XMPP} servers, we used \emph{strophe.js} which we found to have a bug that we fixed.
 
@@
 \label{section:achievements}
-	We have successfully implemented the basic functionalities of our prototype and spare some time for adding more valuable features such as the ability to create content by exposing \emph{QR codes} to the camera and lastly perform changes to the user interface in order to improve the quality of user's experience.
 
-	The performance tests that we have executed showed that our system is stable and more importantly that our web server is lightweight and most of processing power is dedicated to the streaming server.
+	We have successfully implemented the architecture that we proposed, which made possible to create and use an innovative prototype for virtual meeting and e-learning.
+
+	Our prototype includes the basic functionalities that we planned to have.
+We dedicated additional effort to incorporate more valuable features such as the ability to create content by exposing \ac{QR} codes to the camera and lastly perform changes to the user interface in order to improve the quality of user's experience.
+
+	The performance tests let other researchers, in this area, to understand the expected behavior of each module present on our solution and the faced difficulties when implementing this kind of solutions.
+We have shown that our system is stable and, more importantly, that our web server is lightweight in comparison with streaming, which uses most of processing power.
 

### 2016-05-13 tese2/07_related_work_nat.tex
until 1h test case

@@
 Today we live in the \ac{IoT} era.
-  Users and companies may have multiple devices that require Internet connectivity, such as personal computers, smart phones, tablets, traffic sensors, public surveilance systems, health monitoring tools and so on.
+  Users and companies may have multiple devices that require Internet connectivity, such as personal computers, smartphones, tablets, traffic sensors, public surveilance systems, health monitoring tools and so on.
 All those devices combined cannot be publicly addressed by \ac{IPv4}.
@@
 Direct communication is only achieved by \ac{STUN} when \ac{NAT} is a type \emph{full cone}.
-\ac{ICE} is a technique that uses \ac{STUN} when direct communications are possible and \ac{TURN} while a direct communication isn't possible.
+\ac{ICE} is a technique that uses \ac{STUN} when direct communications are possible and \ac{TURN} while a direct communication is not possible.
 %RP explicar que para usar TURN as aplicações têm de estar cientes disso.
 
-Most of client-server applications aren't affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints.
+Most of client-server applications are not affected by \ac{NAT} when the servers are public, but they're inadequate for real time communication between two private endpoints.
 Clearly \ac{TURN} requires a more expensive relaying infrastructure and, in most cases, more network usage, leading to a worse quality of service.

### 2016-05-13 tese2/08_related_work_rtc.tex
[ADD] template2
[ADD] amazing explanation of network test results
[ADD] references to tables
[FIX] reference to stroph.js
[RFT] removed some blue text and some old RP comments
until 1h test case

--- 2016-03-18 tese1/08_related_work_rtc.tex
@@
-\subsection{Real time communications}\label{rtc}
+\section{Real time communications}\label{rtc}
 
-\ac{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \ac{API}s for web browser based real time communications without installing any additional application or plug-in.
+\ac{WebRTC} is an open source technology that defines a collection of standard protocols and \emph{JavaScript} \ac{API}s for web browser based real time communications without installing any additional application or plug-in.
+Table \ref{table:webrtcstack} shows the protocols that \ac{WebRTC} rely on.
 
@@
 
-%RP indicar que agora também já temos webrtc fora do browser devido ao seu sucesso e ao facto de reunir tecnologia state of the art?
-
-\newcolumntype{Y}{>{\centering\arraybackslash}X}
 %\renewcommand{\arraystretch}{2}
 
-\begin{table}[]
+\begin{table}
 \centering
-\label{my-label}
+\caption{WebRTC protocol Stack}
+\label{table:webrtcstack}
     \begin{tabularx}{\textwidth}{*8{Y}@{}m{0pt}@{}} 
     	\cline{5-8}
-		&&&\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{~~~~~~MediaStream~~~~~~} & \multicolumn{2}{c|}{~~~~~~DataChannel~~~~~~} 										&\\[12pt]
+		&&&\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{~~~~~~~~~~MediaStream~~~~~~~~~~} & \multicolumn{2}{c|}{~~~~~~~~~~DataChannel~~~~~~~~~~} 								&\\[12pt]
 		\hline 
@@
 	\end{tabularx}
-	\caption{WebRTC protocol Stack}
 
@@
   \item \textbf{PeerConnection} acquires connection data and negotiates with peers.
-%RP establishes and negociates a connection with a peer for transmitting real-time video or audio?
+
   \item \textbf{DataChannel} provides a channel for exchanging arbitrary data with other peers.
@@
 
-\ac{TLS} doesn't support independent packet decryption\cite{rfc6347}, for that it requires a reliable transport channel, typically \ac{TCP}.
+\ac{TLS} does not support independent packet decryption\cite{rfc6347}, for that it requires a reliable transport channel, typically \ac{TCP}.
 The decryption of a packet depends on the previous packet, which for unreliable transport protocols like \ac{UDP} may represent a problem, either due to packet loss or different reception order.
@@
 By knowing the sequence of messages that are sent and received in \ac{DTLS}, timers are used for packet retransmission avoiding acknowledgment messages.
+
+Table \ref{table:transport_protocols} shows an overview of the transport protocols.
+\ac{WebRTC}'s \emph{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}.
+\ac{DTLS} encapsulation provides confidentiality, authentication and integrity to the transfered data.
+A \emph{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication.
+Each data channel direction can be configured for reliable or unreliable transmission, the same can be done for order delivery and priority.
+which can also be defined for improving the quality of service of a particular stream over the others.
+
 %RP TLS?
 %\def\tabularxcolumn#1{m{#1}}
-\begin{table}[]
+\begin{table}
 \centering
+	\caption{Overview of transport protocols}
+	\label{table:transport_protocols}
     \begin{tabularx}{\textwidth}{X|Y|Y|Y|@{}m{0pt}@{}}
@@
 \end{tabularx}
-	\caption{Overview of transport protocols}
 \end{table}
-
-\ac{WebRTC}'s \emph{DataChannel} is built on top of \ac{SCTP}, which is encapsulated by \ac{DTLS}.
-\ac{DTLS} encapsulation provides confidentiality, authentication and integrity to the transfered data.
-A \emph{Data Channel} has one incoming stream and one outgoing stream, providing bidirectional communication.
-Each data channel direction can be configured for reliable or unreliable transmission, the same can be done for order delivery and priority.
-which can also be defined for improving the quality of service of a particular stream over the others.
 
@@
 The new keys derived from \ac{DTLS} handshake are seized for \ac{SRTP} encryption, the remaining \ac{SRTP} communications are done through \ac{UDP} without using \ac{DTLS}.
-%RP quais remaining?  A última frase é confusa
 
@@
 \emph{Skype}\footnote{\url{http://www.skype.com/}(accessed June 1, 2015).} is an application that allows video, voice, instant messaging and multi-party communication over proprietary protocols, its main strength are the amount of users that are using it nowadays and the ability to perform voice calls to the \ac{PSTN}.
-But compared to \emph{Skype}, \ac{WebRTC} applications don't need to be pre-installed.
+But compared to \emph{Skype}, \ac{WebRTC} applications do not need to be pre-installed.
 
 \emph{Google Hangouts}\footnote{\url{http://plus.google.com/hangouts}(accessed June 1, 2015).} is another popular video multi-party conference web application. 
-In the past, in order to use \emph{hangouts} on a web browser a plug-in had to be installed, nowadays hangouts is using \ac{WebRTC}.
+In the past, in order to use \emph{hangouts} on a web browser a plug-in had to be installed.
+Nowadays hangouts is using \ac{WebRTC}.
 \emph{Google Hangouts} supports viewing videos on \emph{youtube} synchronously, drawing collaboratively, creating music and playing multi-player games.

### 2016-05-13 tese2/09_related_work_signaling.tex
[RFT] removed some blue text and some old RP comments

@@
 
-  \ac{WebRTC} uses \ac{SDP} \cite{rfc4566} to define peer connection properties such as types of supported media, codecs, protocols used and network information.
+  \ac{WebRTC} uses \ac{SDP}~\cite{rfc4566} to define peer connection properties such as types of supported media, \emph{codecs}, protocols used and network information.
 An \ac{SDP} offer describes to other peers the expected type of communication and its details, such as used transport protocols, codecs, security and other.
@@
 One technique to overcome this problem is polling.
-  %RP num documento formal não se usam contrações como it's.
-  %RP o teu ingles é parco em artigos. Faltam muitos the, of, it, etc.
-
 Polling consists on sending periodic messages to which the server responds immediately with empty content or fresh information.
@@
 
-The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel \cite{rfc6455}, by other words it supports sending and receiving data simultaneously.
+The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel~\cite{rfc6455}, by other words it supports sending and receiving data simultaneously.
 %RP coloca as citações no final da frase. Pode haver excepções onde a citação apenas cobre parte da frase ou precisas de colocar várias a dizer coisas diferentes.
@@
 
-  \ac{BOSH}\cite{xep0124} is a technique based on long polling that uses two socket connections and allows sending client messages to the server while a previous request is held.
+  \ac{BOSH}~\cite{xep0124} is a technique based on long polling that uses two socket connections and allows sending client messages to the server while a previous request is held.
   The \ac{BOSH} specification assumes that a connection manager is implemented to handle \ac{HTTP} connections.
@@
 
-  \ac{SIP} \cite{rfc3261} is a protocol used for negotiation, creation, modification and finalization of communication sessions between users.
+  \ac{SIP}~\cite{rfc3261} is a protocol used for negotiation, creation, modification and finalization of communication sessions between users.
 \ac{SIP} follows a client/server architecture with \ac{HTTP} like messages and it can be used as a signaling protocol.
@@
 
-  The working group \ac{SIMPLE}\footnote{\url{https://datatracker.ietf.org/wg/simple/documents/}(accessed June 1, 2015).} proposed the creation of \ac{SIP} extensions, namely presence information \cite{rfc5263} and instant messaging \cite{rfc3428}.
+  The working group \ac{SIMPLE}\footnote{\url{https://datatracker.ietf.org/wg/simple/documents/}(accessed June 1, 2015).} proposed the creation of \ac{SIP} extensions, namely presence information~\cite{rfc5263} and instant messaging~\cite{rfc3428}.
 
@@
   \ac{XMPP} messages are \ac{XML} based, which is attractive for applications that need structured messages and rich hypermedia.
-Another advantage of \ac{XMPP} is the addition of extensions, for example \emph{XEP-0096}\cite{xep0096}, which adds file transfer capabilities between two entities, and \emph{XEP-0045}\cite{xep0045}, which enables multi-user chat.
-  \ac{XMPP}'s bi-directional communication over \ac{HTTP} is achieved through \ac{BOSH} \cite{xep0206}.
-  This kind of communication is also possible through WebSockets \cite{rfc7395}.
+Another advantage of \ac{XMPP} is the addition of extensions, for example \emph{XEP-0096}~\cite{xep0096}, which adds file transfer capabilities between two entities, and \emph{XEP-0045}~\cite{xep0045}, which enables multi-user chat.
+  \ac{XMPP}'s bi-directional communication over \ac{HTTP} is achieved through \ac{BOSH}~\cite{xep0206}.
+  This kind of communication is also possible through WebSockets~\cite{rfc7395}.
   Today, multiple \ac{XMPP} server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}(accessed June 1, 2015).}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}(accessed June 1, 2015).}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}(accessed June 1, 2015).} and \emph{Prosody}\footnote{\url{http://prosody.im/}(accessed June 1, 2015).}.
@@
 
-{\color{blue}
-
     One of the advantages that \ac{XMPP} based applications offer is the possibility to send messages across different servers.
@@
 
-    The user profile is represented in terms of Data Forms\cite{xep0004} protocol and managed (registration and update) through \textit{Info/Query} (IQ) \cite{rfc3920} requests to the \ac{XMPP} server, which responds with an IQ response containing information to retrieve the registration fields needed to fill its profile (eg username, password, telephone).
+    The user profile is represented in terms of Data Forms~\cite{xep0004} protocol and managed (registration and update) through \textit{Info/Query} (IQ)~\cite{rfc3920} requests to the \ac{XMPP} server, which responds with an IQ response containing information to retrieve the registration fields needed to fill its profile (eg username, password, telephone).
 %RP frase demasiado longa. Torna-se confusa.
@@
 
-    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that \ac{XMPP} already implement, but  we don't need all of them.
+    We know that an \ac{XMPP} based application could simplify our work just by using functionalities that \ac{XMPP} already implement, but  we do not need all of them.
 In fact, we just need a subset of \ac{XMPP} features, namely a simple way to register users, access and edit user's information, create chat rooms, send messages and access to presence information.
-}
 

### 2016-05-13 tese2/10_related_work_hypermedia.tex

@@
 An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time in a video or audio track.
-\emph{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe.
+\emph{HyperCafe}~\cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe.
 
   Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment.
-\emph{Hyper-Hitchcock}\cite{hitchcock} is an editor and player of detail-on-demand video.
+\emph{Hyper-Hitchcock}~\cite{hitchcock} is an editor and player of detail-on-demand video.
 
@@
 
-  \emph{HyVAL}\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia.
+  \emph{HyVAL}~\cite{hyval} is an \ac{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia.
 HyVAL defines defines video structure, internal video and external media objects.
@@
   \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically. 
-  This approach is quite restrictive if we want to apply hyper-video concepts to videos that don't follow this structure.
+  This approach is quite restrictive if we want to apply hyper-video concepts to videos that do not follow this structure.
 External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
@@
 
-  \ac{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films.
+  \ac{SMIL}~\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films.
 With \ac{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
@@
 The \ac{DOM} is a standard \ac{API} that allows easy management of documents that are organized in a tree structure, by providing \ac{CRUD} operations over its elements and their attributes.
-\ac{DOM} makes it easy to inter-operate between imperative and declarative programming languages\cite{dom}.
+\ac{DOM} makes it easy to inter-operate between imperative and declarative programming languages~\cite{dom}.
 %RP não arranjas uma referência para o dom?
@@
 
-  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, doesn't require a plug-in to be installed and shouldn't have incompatibility issues.
+  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \ac{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, does not require a plug-in to be installed and should not have incompatibility issues.
   SmillingWeb already takes advantage of \ac{HTML}5 and \ac{CSS}3.
-  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}(accessed June 2, 2015).}, a simple \emph{JavaScript} library that may require plug-ins if new features aren't supported.
+  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}(accessed June 2, 2015).}, a simple \emph{JavaScript} library that may require plug-ins if new features are not supported.
   But SmillingWeb just implements a subset of \ac{SMIL} 3.0 and their scheduler engine loads the \ac{SMIL} file only once, which could raise problems when dealing with \ac{SMIL} changes due to real time communications.
@@
 It is used mainly on client-side and executed by a web browser.
-\emph{JavaScript} has its own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plug-ins.
+\emph{JavaScript} has its own implementation of \ac{DOM} and one of its advantages is the ability to download and execute code on-the-fly without the need of pre-installed plug-ins.
 
@@
 
-  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} implement some of \emph{flash}'s features, it doesn't mean that \emph{flash} will be replaced soon.
+  Although \ac{HTML}5, \emph{JavaScript}, \ac{CSS} and \ac{WebRTC} implement some of \emph{flash}'s features, it does not mean that \emph{flash} will be replaced soon.
   Instead, both technologies can be used to develop rich \emph{Internet} applications.

### 2016-05-13 tese2/11_related_work_collaboration.tex
until 1h test case

@@
 
-  \ac{RTP} recorders are independent of payload encoding, they don't decode \ac{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.
+  \ac{RTP} recorders are independent of payload encoding, they do not decode \ac{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.
 
@@
   %RP e usa subsecções...
-\begin{figure}[!htb]
+\begin{figure}
 	\centering
@@
 	Streaming protocols like \ac {RTP} and \ac{RTSP} were designed for continuous and non-interactive media types, such as audio and video.
-Discrete and non-interactive media don't need to be streamed through \ac{RTP} because they don't change with time.
+Discrete and non-interactive media do not need to be streamed through \ac{RTP} because they do not change with time.
 For example, if an image appears in a specific time interval, just the \ac{HTML} or \emph{JavaScript} that will reference the image must be streamed, the image itself is then transferred through \ac{HTTP}.
@@
 
-\begin{table}[!htb]
+\begin{table}
 \centering

### 2016-05-13 tese2/13_implementation_model.tex
until 1h test case

@@
     Each \emph{MongoDB} document is represented by a \emph{JSON} object and, as a result, each document may have different attributes within a collection.
-Needless to say that, by following this approach we don't need to create the collections with a well predefined schema.
-In fact we don't need to define it at all but, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure as we will present in this section.
+Needless to say that, by following this approach we do not need to create the collections with a well predefined schema.
+In fact we do not need to define it at all but, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure as we will present in this section.
     %RP a última frase é confusa. O que queres dizer? Reescreve
@@
 This specification lets us create abstract authentication services.
-For example a user can login into our system by providing any attribute that identifies himself, e.g.
-the e-mail but others are possible like the user name or cellphone number.
+For example a user can login into our system by providing any attribute that identifies himself, \emph{e.g.} the e-mail but others are possible like the user name or cellphone number.
 

### 2016-05-13 tese2/14_implementation_signaling.tex
until 1h test case

@@
 
-\begin{figure}[!htb]
+\begin{figure}
     \centering
@@
 %HR done
-Our signaling protocol consists of sending and receiving \ac{JSON} formated messages, e.g. listing \ref{lst:msgtype}, over WebSockets by both the application server and the client.
+Our signaling protocol consists of sending and receiving \ac{JSON} formated messages, \emph{e.g.} listing \ref{lst:msgtype}, over WebSockets by both the application server and the client.
 
@@
 At this stage, the web application's user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server.
-If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2016).} may ask to install a plug-in if the browser doesn't support screen sharing \footnote{\url{http://iswebrtcreadyyet.com/} (Accessed May 11, 2016)}.
+If the user decides to perform a screen share, \emph{adapter.js}\footnote{\url{https://github.com/Temasys/AdapterJS}(accessed March 15, 2016).} may ask to install a plug-in if the browser does not support screen sharing \footnote{\url{http://iswebrtcreadyyet.com/} (Accessed May 11, 2016)}.
 %RP web app user? OU apenas client user?
@@
 The \emph{local session description} contains the session identifier, codecs, containers, transport protocols and ports used per media type.
-The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} doesn't need to mix, record nor analyze streams coming from the user's \ac{WebRTC} endpoint.
+The \emph{local session description} is useful to conclude if the client is receiving only, which means that \ac{KMS} does not need to mix, record nor analyze streams coming from the user's \ac{WebRTC} endpoint.
 

### 2016-05-13 tese2/15_implementation_recording.tex

@@
 The user specified which user and date he intended to play, the server calculated the intersection between the requested date and the block bounds and returned the file to the client.
-	Although this idea was fairly simple, we couldn't achieve seamless sequential block switching.
+	Although this idea was fairly simple, we could not achieve seamless sequential block switching.
 Downloading the video file always took a noticeable amount of time.
@@
 	One of the features that \emph{Kurento Repository} provides is the ability to play directly from the database without having to download the entire file to \ac{KMS}.
-The same is true for recording, but because the file headers are in the beginning and the file is written until is stops, the headers don't contain the necessary information for seeking the file.
+The same is true for recording, but because the file headers are in the beginning and the file is written until is stops, the headers do not contain the necessary information for seeking the file.
 	Although we gain with scalability with this approach, we lose access over the file for changing it to fit our needs, namely for using \emph{ffmpeg} or other video manipulation tool.

### 2016-05-13 tese2/16_implementation_content.tex

@@
 
-	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real-time, we allow the user to encode its content into \emph{QR codes} and show it to the camera in real time.
+	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real-time, we allow the user to encode its content into \ac{QR} codes and show it to the camera in real time.
 
-	\ac{KMS} lets registering event handlers for \emph{QR code} detection.
-The component that detects \emph{QR codes} on \ac{KMS} is called periodically and fires the handlers with the decoded content.
-This mechanism does not detect if the \emph{QR code} enters or leaves the screen.
+	\ac{KMS} lets registering event handlers for \ac{QR} code detection.
+The component that detects \ac{QR} codes on \ac{KMS} is called periodically and fires the handlers with the decoded content.
+This mechanism does not detect if the \ac{QR} code enters or leaves the screen.
 We had to implement our own mechanism for detecting those events.
@@
 In order to apply our algorithm more efficiently we calculate the content's hash through the \emph{md5} method.
-If the hash is not present in the map, it means the \emph{QR code} was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real-time.
-If the same \emph{QR code} is not detected after a time bigger then two periods we can conclude that the \emph{QR code} left the screen and we add the correspondent content into our database.
-Listing \ref{lst:pseudo_qrcode} shows the pseudo code for \emph{QR code} leaving detection.
+If the hash is not present in the map, it means the \ac{QR} code was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real-time.
+If the same \ac{QR} code is not detected after a time bigger then two periods we can conclude that the \ac{QR} code left the screen and we add the correspondent content into our database.
+Listing \ref{lst:pseudo_qrcode} shows the pseudo code for \ac{QR} code leaving detection.
 

### 2016-05-13 tese2/acknowledgements.tex

@@
 
-I would like to thank my dissertation supervisor Ricardo Pereira and co-supervisor Paulo Chainho for their active support and ideas that were crucial for the development of this thesis.
+I would like to thank my thesis supervisor Ricardo Pereira and co-supervisor Paulo Chainho for their active support and ideas that were crucial for the development of this thesis.
 
@@
 
-Not less important I should thank all IST employees that maintained the space I used during this journey.
+Not less important, I should thank all IST employees that maintained the space I used during this journey.
 

### 2016-05-13 tese2/evaluation.tex

@@
 
-  Just for testing purposes, we performed the same performance tests disabling \emph{QR codes} detection in order to understand how \ac{CPU} intensive this task is.
-Figure \ref{fig:test_without_qrcode_cpu} shows the results for the same experience but with \emph{QR code} detection disabled.
+  Just for testing purposes, we performed the same performance tests disabling \ac{QR} codes detection in order to understand how \ac{CPU} intensive this task is.
+Figure \ref{fig:test_without_qrcode_cpu} shows the results for the same experience but with \ac{QR} code detection disabled.
 
@@
 
-We conclude that \emph{QR code} detection is a very intensive task, approximately doubling the amount of work performed by the \ac{CPU}s.
+We conclude that \ac{QR} code detection is a very intensive task, approximately doubling the amount of work performed by the \ac{CPU}s.
 Without this feature, the network and memory usage had insignificant changes compared to \ac{CPU} usage.
@@
       \item Enter a conference room by sharing their camera and add a subtitle in the video, preview it, specify an interval of time and save it.
-Then show the provided \emph{QR code} to the camera and leave the leave the conference room.
+Then show the provided \ac{QR} code to the camera and leave the leave the conference room.
       \end{enumerate}
@@
 
-On the fourth task the users didn't placed their annotations on right place. We only provide one way to place annotations, which is by dragging them on the time-line. 
-After creating an annotation, users also didn't understand that they had to save their changes on the time-line.
+On the fourth task the users did not placed their annotations on right place. We only provide one way to place annotations, which is by dragging them on the time-line. 
+After creating an annotation, users also did not understand that they had to save their changes on the time-line.
 We solved this problem by saving automatically each time the user changed any annotation.
@@
 The three users had success with the fifth task, but they revealed difficulties choosing the time interval for the hyper content.
-We are aware that this is a difficult task and that's why we have introduced the \emph{QR code} content creation mechanism which was very easy to learn (with a \emph{QR code} already printed into paper).
+We are aware that this is a difficult task and that's why we have introduced the \ac{QR} code content creation mechanism which was very easy to learn (with a \ac{QR} code already printed into paper).
 
-Although content creation through \emph{QR codes} revealed much easier to perform, this approach implies showing and, as a consequence, recording video with the \emph{QR codes}, making them visible to other users.
+Although content creation through \ac{QR} codes revealed much easier to perform, this approach implies showing and, as a consequence, recording video with the \ac{QR} codes, making them visible to other users.
 In order to solve this problem, we could add a new video sharing mode that allow users to send video to \ac{KMS} without composing, recording and showing to other users.

### 2016-05-13 tese2/frontCover.tex

@@
 \hspace{-0.5cm}\includegraphics[height=50mm]{Figures/logo.png}
-% TODO - If you don't use an image uncomment the next line
+% TODO - If you do not use an image uncomment the next line
 % \vspace{5.0cm}
@@
 \vspace{1.9cm}
-{\FontLn Disserta\c{c}\~{a}o para a obten\c{c}\~{a}o de Grau de Mestre em} \\
+{\FontLn Dissertation submitted to obtain the Master Degree in} \\
 \vspace{0.3cm}
-{\FontLb Engenharia de Telecomunicações e Informática} \\
+{\FontLb Communications Networks Engineering} \\
 \vspace{1.9cm}
@@
 {\FontSn %
-Presidente:        Nome do Presidente \\
-Orientador:        Nome do Orientador \\
-Co-orientador:     Nome do Co-orientador \\
-Vogais:            Nome do Vogal 1 \\
-                   Nome do Vogal 2 \\
+Chairperson: 	   Prof. Paulo Jorge Pires Ferreira
+Supervisor:        Prof. Ricardo Jorge Feliciano Lopes Pereira \\
+Members of the Committee: Prof. João Pedro Barreto
 }
@@
 
-{\FontMb M\^{e}s e Ano} \\
+{\FontMb May 2016} \\
 %

### 2016-05-13 tese2/implementation.tex

@@
 As a consequence of not using \ac{XMPP} for the signaling protocol, we have defined our own protocol based on \emph{WebSockets} for performing the stream between clients and \ac{KMS}.
-From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \emph{QR code} detection for creating hyper content.
+From the \ac{KMS} functionalities, our solution uses the recording feature, composite endpoints that mix multiple audio and video streams into a single one and \ac{QR} code detection for creating hyper content.
 

### 2016-05-13 tese2/introduction.tex

--- 2016-05-05 tese2/introduction.tex
@@
 
+	As communications technologies appeared, we adapted the way we communicate.
+The purpose of this project is not the replacement of the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
+
+	With the advent of WebRTC and its successive integration with web browsers, it became possible to develop video conference web applications without plugins.
+This presents a range of possibilities on what can be implemented using already existing web technologies.
+
+    Furthermore, real-time communication applications can make a significant difference on business, education and health sectors by providing tools for developing teaching and learning online, teamworking and socializing web applications.
+
+        %RP Parece-me que até aqui ainda é background ou motivação. Não me parece que estejas já a descrever a solução proposta.
+
 \section{Proposed Solution}
@@
 
-	As communications technologies appeared, we adapted the way we communicate.
-{\color{blue} The purpose of this project is not the replacement of} the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
-
-	{\color{blue}With the advent of WebRTC and its successive integration with web browsers, it became possible to develop video conference web applications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.}
-
-    Furthermore, real-time communication applications can make a significant difference on business, education and health sectors by providing tools for developing teaching and learning online, teamworking and socializing web applications.
-
-	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they ear or see.
-        %RP Parece-me que até aqui ainda é background ou motivação. Não me parece que estejas já a descrever a solução proposta.
+	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they hear or see.
 
@@
 
-	{\color{blue}Another point to consider is making available collaborative tools for text edition, instant messaging and video superimposed content creation and visualization.}
+	Our goal in this project is to develop an application targeted to the web platform, resorting to \ac{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \ac{HTML}5, \ac{CSS}3 and \emph{JavaScript} including continuous media types such as video, music and animations.
+
+	One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again or introduce hyper-content to it such as time annotations, interactive lists of topics and subtitles.
+In this context we also provide a simpler method for creating and synchronizing hyper-content using \ac{QR} codes.
+
+	In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links and files to conference participants.
+
+	Furthermore, another relevant feature is the possibility to compose multiple video streams into a single one, which enables adding more users to conference rooms without impacting on clients performance.
+Users can change to individual streams on demand or automatically to the talking users.
 
@@
 
-Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (e.g.
-creating time annotations for improving content search) in order to increase its value.
+Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (\emph{e.g.} creating time annotations for improving content search) in order to increase its value.
 It is also important to highlight another goal of this project which is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.
@@
 
-A web application with an easy to learn user interface will be developed to accomplish solving our problem.
+A web application with an easy to learn user interface was developed to accomplish solving our problem.
 Our application, unnamed yet, is targeted at web browsers that are compatible with only standard technologies like JavaScript, \ac{WebRTC}, \ac{HTML}5 and \ac{CSS}3.
-Any additional plug-in is avoidable, \emph{JavaScript} libraries will be preferred as they can be downloaded on the fly.
+Any additional plug-in was avoided, \emph{JavaScript} libraries were preferred as they can be downloaded on-the-fly.
 

### 2016-05-13 tese2/resumo.tex

@@
 
-O conceito de comunicações hiper-ligadas aplica muitos dos conceitos de hiper-media, bastante utilizados no conteúdo \emph{Web}, para serviços se comunicação e colaboração.
-Este paradigma permite sincronizar, estruturar e navegar sobre conteúdo proveniente de communicações, colocando ao dispor formas de integrar conteúdo de mídia social e ferramentas colaborativas em chamadas de voz e vídeo.
-A voz e a imagem justas exprimem emoções e expoem a creatividade como nenhum outro tipo de media permite.
-Através dos conceitos de hiper-media, podemos adicionar mais valor às comunicações por voz e vídeo.
+O conceito de comunicações hiper-ligadas aplica muitos dos conceitos de hiper-media, bastante utilizados no conteúdo \emph{Web}.
+Este paradigma permite sincronizar, estruturar e navegar sobre conteúdo proveniente de communicações integrado em chamadas de voz e vídeo.
 
-A tecnologia \ac{WebRTC} permite realizar comunicações em tempo real entre navegadores web sem a necessidade de instalar aplicações adicionais ou \emph{plug-ins}.
+A voz e a imagem justas exprimem emoções como nenhum outro tipo de media permite. Através dos conceitos de hiper-media, podemos adicionar mais valor às comunicações por video-conferência.
+
+A tecnologia \ac{WebRTC} permite realizar comunicações em tempo real entre navegadores web sem a necessidade de instalar \emph{sofware} adicional.
 A natureza das aplicações \emph{web} já usufruem dos conceitos de hiper-media, o que faz do \ac{WebRTC} a tecnologia ideal para aplicar o conceito de comunicações hiper-ligadas.
@@
 
-O nosso objectivo neste projecto é o desenvolvimento de uma aplicação que utiliza o conceito de comunicações hiper-ligadas de forma a enriquecer ambientes colaborativos e comunicações audio e vídeo com vários tipos de media, que podem ser acedidos em tempo real ou mais tarde.
-Esta aplicação será desenvolvida sobre a plataforma \emph{web}, recorrendo ao \ac{WebRTC}.
+O nosso objectivo neste projecto é desenvolver de uma aplicação, recorrendo ao \ac{WebRTC}, que utiliza o conceito de comunicações hiper-ligadas de forma a enriquecer as comunicações em conferência com vários tipos de media, editores de texto colaborativos, anotações temporais, envio de mensagens e possilidade de sobrepor contéudo ao vídeo.
 
-Neste documento, apresentamos o actual Estado de Arte nas comunicações hiper-ligadas e as tecnologias relacionadas, propomos e implementamos uma arquitectura para uma aplicação com comunicações hiper-ligadas baseada no \emph{WebRTC} e mostramos os resultados da implementação e avaliação.
+Neste documento, apresentamos o actual Estado de Arte nas comunicações hiper-ligadas e as tecnologias relacionadas, propomos e implementamos uma arquitectura para uma aplicação com comunicações hiper-ligadas baseada no \emph{WebRTC}.
+O nosso trabalho foi avaliado por utilizadores que, para além de terem gostado da experiência com a nossa aplicação, acharam o nosso protótipo bastante inovador.
 

### 2016-05-13 tese2/xthesis.tex

@@
 \documentclass[conference,compsoc,a4paper]{IEEEtran}
-\ifCLASSOPTIONcompsoc
-  \usepackage[nocompress]{cite}
-\else
-  \usepackage{cite}
-\fi
+\usepackage{geometry}
+ \geometry{
+ a4paper,
+ total={170mm,257mm},
+ left=20mm,
+ top=20mm,
+ }
 
+\usepackage[backend=bibtex]{biblatex}
 \usepackage{array}
@@
 \usepackage[nomain,acronym,xindy,toc]{glossaries} 
-\usepackage{dirtytalk}
 \usepackage{graphicx}
 \usepackage{tabularx} 
-\usepackage{hyperref}
+\usepackage{float}
+
+\usepackage{color, colortbl}
+\definecolor{Gray}{gray}{0.9}
 %Includes "References" in the table of contents
@@
 
+\bibliography{references} % or
+
 \hyphenation{op-tical net-works semi-conduc-tor}
 \loadglsentries[main]{xacronyms}
-\begin{document}
 
@@
 
+\begin{document}
+
 \title{Hyper-linked Communications: WebRTC enabled asynchronous collaboration}
 
-\author{\IEEEauthorblockN{Michael Shell}
-\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
-Georgia Institute of Technology\\
-Atlanta, Georgia 30332--0250\\
-Email: http://www.michaelshell.org/contact.html}
+\author{\IEEEauthorblockN{Henrique Rocha}
+\IEEEauthorblockA{Instituto Superior T\'{e}cnico\\}
+\IEEEauthorblockN{Av. Prof. Dr. Anibal Cavaco Silva\\
+2744-016 Porto Salvo, Portugal\\
+henrique.rocha@tecnico.ulisboa.pt}
 \and
-\IEEEauthorblockN{Homer Simpson}
-\IEEEauthorblockA{Twentieth Century Fox\\
-Springfield, USA\\
-Email: homer@thesimpsons.com}
-\and
-\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
-\IEEEauthorblockA{Starfleet Academy\\
-San Francisco, California 96678-2391\\
-Telephone: (800) 555--1212\\
-Fax: (888) 555--1212}}
+\IEEEauthorblockN{Ricardo Pereira}
+\IEEEauthorblockA{Instituto Superior T\'{e}cnico\\}
+\IEEEauthorblockN{Av. Prof. Dr. Anibal Cavaco Silva\\
+2744-016 Porto Salvo, Portugal\\
+ricardo.pereira@inesc-id.pt}}
 
@@
 \begin{abstract}
-
 The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content.
@@
 Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging and a mechanism to superimpose hyper-content to video.
-
-In this document, we present the current State Of The Art in hyper-linked communications and related technologies, propose and implement an architecture for an hyper-linked communication application based on \emph{WebRTC}.
-This work was evaluated by users, who reported that they liked to use it and thought it to be extremely innovative.
-
 \end{abstract}
@@
 	One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again or introduce hyper-content to it such as time annotations, interactive lists of topics and subtitles.
-In this context we also provide a simpler method for creating and synchronizing hyper-content using \emph{QR codes}.
+In this context we also provide a simpler method for creating and synchronizing hyper-content using \gls{QR} codes.
 
@@
 \section{Related Work}
- This section is structured as follows.
- Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \gls{IPv4} address exhaustion and the client server model constraints. 
- Section \ref{rtc} describes the \gls{WebRTC} technology and the protocols needed to implement our project. 
- Section \ref{signaling} addresses the signaling component of chat applications, which is not defined on \gls{WebRTC} specifications. 
- Section \ref{hypermedia} presents the evolution of multimedia content until the hypermedia, its capabilities, synchronization mechanisms and interactivity. 
- Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
+\label{chapter:relatedwork}
 
-\subsection{Early days of the Internet and its remaining flaws}\label{early}
+One way to overcome the \gls{IPv4} address scarcity problem \cite{ipv4} was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \gls{IP} address with a different port.
+This technique is known as \gls{NAT}\cite{rfc3489}.
 
-The need to build a global communications network in an age when almost nobody had access to that technology and the unpredictability of the number of future users, lead to some protocols not being suitable for the explosive growth and proliferation of users that followed.
-\gls{IPv4} limits the number of public addresses in such a way that today they are scarce \cite{ipv4}.
-
-To this end, one way to overcome the \gls{IPv4} address scarcity problem was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \gls{IP} address with a different port.
-
-This technique is known as \gls{NAT} (figure \ref{fig:nat}).
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\linewidth]{figures/nat.pdf}
-	\caption{Network Address Translation}
-	\label{fig:nat}
-\end{figure}
-
-Initially \gls{NAT} offered an alternative to address exhaustion and a minimal sensation of security.
-However, due to their current wide usage, \gls{NAT}s weaknesses are being exposing at the application layer, namely impacting applications that require direct communications between two private networks.
-
-There are four types of \gls{NAT} implementations\cite{rfc3489}: \emph{Full Cone NAT}, \emph{Restricted Cone NAT}, \emph{Port Restricted Cone NAT}, \emph{Symmetric NAT}.
-
-\emph{Full Cone} \gls{NAT} maps each public \gls{IP} address and port to a private \gls{IP} address and port.
-Any external host can communicate with private hosts through their mapped public address and port.
-This represents the least restrictive type of \gls{NAT} and, as we will see later, the unique type of \gls{NAT} that enables real time communications from point to point.
-
-\emph{Restricted Cone} \gls{NAT} requires that a private client must first send a message to an external host before it can receive messages from the same host.
-With this type of \gls{NAT}, the private client can be contacted from any port of the same external host.
-
-\emph{Port Restricted Cone} \gls{NAT} works in the same way as Restricted Cone \gls{NAT}, but it only allows communications from the same external host's IP address and port, ignoring all messages from other applications within the same external host.
-
-\emph{Symmetric} NAT maps different ports for each connection. As we will see later, this type of \gls{NAT} represents a problem on real time communications.
-
-\emph{Non-Symmetric} \gls{NAT}s became the common configurations on the Internet.
-As a direct result, problems started to appear: the amount of ports that \gls{IP} makes available is also small compared to our current needs; worse than that, \gls{NAT} also difficults end-to-end communication, forcing most applications that follow this model to be implemented ineffectively.
-
-Unless the router that performs as \gls{NAT} has forwarding rules to every desired ports of each user, applications behind a \gls{NAT} are prevented from receiving incoming connections from the public network, which forces them to behave as a client of a client-server model.
-
-\gls{STUN} and \gls{TURN} \cite{natvoip} servers are a possible solution to overcome \gls{NAT}.
-
-\gls{STUN} servers are quite simple.
-They receive requests from \gls{NAT}ed clients, with the source address of a request being the public address that \gls{NAT} mapped to the client.
-\gls{STUN} servers will then reply to the client, providing the mapped public address, so it knows its associated public \gls{IP} address and port.
-Symmetric \gls{NAT} changes \gls{IP} port for each different connection, for that reason, when the \gls{STUN} servers reply with the \gls{IP} address and port of their connection, it will be useless for clients to use them on other application connections.
-That is why Symmetric \gls{NAT} represents a problem for peer-to-peer communications.
-
-On the other hand, \gls{TURN} uses public servers to relay traffic between private endpoints.
-It may use a \gls{P2P} network relay to find the best peer, but after that, the behavior is much like client-server.
-Direct communication is only achieved by \gls{STUN} when \gls{NAT} is a type \emph{full cone}.
-\gls{ICE} is a technique that uses \gls{STUN} when direct communications are possible and \gls{TURN} while a direct communication isn't possible.
-
-When connection is established, either in a direct or indirect way (via \gls{TURN} servers), \gls{WebRTC} came to simplify how audio and video are transmitted through web browsers.
-
-\subsection{Real time communications}\label{rtc}
-
-\gls{WebRTC} is an open source technology that defines a collection of standard protocols and JavaScript \gls{API}s for web browser based real time communications without installing any additional application or plug-in.
-Table \ref{table:webrtcstack} shows the protocols that \gls{WebRTC} rely on.
-
-Some operating systems such as \emph{Android}, \emph{iOS}, \emph{Linux}, \emph{OSX} and \emph{Windows} implement native \gls{WebRTC} libraries, extending the usage of \gls{WebRTC} to applications outside the web browser.
-This native support can help to implement applications that record video and audio streams for further playback.
-
-\begin{table}
-\centering
-\caption{WebRTC protocol Stack}
-\label{table:webrtcstack}
-    \begin{tabularx}{\linewidth}{*8{Y}@{}m{0pt}@{}} 
-    	\cline{5-8}
-		&&&\multicolumn{1}{c|}{} & \multicolumn{2}{c|}{MediaStream} & \multicolumn{2}{c|}{DataChannel} 								&\\[12pt]
-		\hline 
-		\multicolumn{1}{|c|}{XHR} & \multicolumn{1}{c|}{SSE} & \multicolumn{2}{c|}{WebSocket} & \multicolumn{2}{c|}{SRTP} & \multicolumn{2}{c|}{SCTP} 				&\\[12pt]
-		\cline{1-3}\cline{6-8} 
-		\multicolumn{3}{|c|}{HTTP 1.X/2.0} & \multicolumn{1}{c|}{} & \multicolumn{1}{c}{} & \multicolumn{3}{|c|}{Session (DTLS)} 	&\\[12pt]
-		\hline
-		\multicolumn{4}{|c|}{Session (TLS) - optional} & \multicolumn{4}{c|}{ICE, STUN, TURN} 																		&\\[12pt]
-		\hline
-		\multicolumn{4}{|c|}{Transport (TCP)} & \multicolumn{4}{c|}{Transport (UDP)} 																				&\\[12pt]
-		\hline
-		\multicolumn{8}{|c|}{Network (IP)}  																														&\\[12pt]
-		\hline
-	\end{tabularx}
-
-\end{table}
-
-\gls{WebRTC} defines three main \gls{API}s: MediaStream, PeerConnection and DataChannel.
-
-\begin{itemize}
-  \item \textbf{MediaStream} allows the browser to access the camera, microphone and the device's screen.
-
-  \item \textbf{PeerConnection} acquires connection data and negotiates with peers.
-
-  \item \textbf{DataChannel} provides a channel for exchanging arbitrary data with other peers.
-\end{itemize}
-
-\gls{WebRTC} uses \gls{UDP} for transporting data, which provides lower latencies than \gls{TCP}, but is not reliable and does not assure packet order and integrity.
-\gls{SCTP} and \gls{SRTP} are used for streaming data, providing a mechanism for congestion control and partial reliable delivery over \gls{UDP}.
-All transferred audio, data and video must be encrypted with \gls{DTLS} symmetric keys.
-\gls{DTLS} provides the same security guarantees as \gls{TLS}.
-
-\gls{TLS} doesn't support independent packet decryption\cite{rfc6347}, for that it requires a reliable transport channel, typically \gls{TCP}.
-The decryption of a packet depends on the previous packet, which for unreliable transport protocols like \gls{UDP} may represent a problem, either due to packet loss or different reception order.
-
-\gls{DTLS} is similar to \gls{TLS}, but is used on top of \gls{UDP}.
-
-\gls{WebRTC}'s \emph{MediaStream} is built on top of \gls{SRTP}, which requires an external mechanism for key exchange.
-\gls{DTLS} keys are negotiated on handshake in order to achieve a secure connection.
-The new keys derived from \gls{DTLS} handshake are seized for \gls{SRTP} encryption, the remaining \gls{SRTP} communications are done through \gls{UDP} without using \gls{DTLS}.
-
-\gls{WebRTC} aims to provide a standard platform for real-time audio and video on the Web. It arrives at a time when several proprietary products are well established.
-
-\subsection{Signaling: meet and get to know}
-\label{signaling}
-
-  Signaling is the process by which applications exchange connection information about peers and servers, their capabilities and meta-data.
-  In particular, \gls{WebRTC} does not implement signaling, as different applications may require different protocols and there is no single answer that fits all problems.
-  As a consequence, multiple options are available for filling the missing \gls{WebRTC}'s signaling component, which can be performed using \gls{SIP}, \gls{XMPP}, \emph{WebSockets}, \emph{Socket.io}\footnote{\url{http://socket.io/}(accessed June 1, 2015).} or by implementing a custom protocol.
-
-  \subsubsection{WebRTC}
-
-  \gls{WebRTC} uses \gls{SDP} \cite{rfc4566} to define peer connection properties such as types of supported media, codecs, protocols used and network information.
-An \gls{SDP} offer describes to other peers the expected type of communication and its details, such as used transport protocols, codecs, security and other.
-
-  One of \gls{WebRTC} signaling's requisites is bi-directional communication.
-
-  \gls{HTTP} uses a request-response paradigm, where a request is sent by the client, followed by a server response.
-
-  Sometimes it is required that some information be obtained in real time, but we saw, some \gls{NAT}'s do not support callbacks from servers, preventing them from notifying clients as soon as an event occures.
-One technique to overcome this problem is long polling which consists on making the server hold the request until there is fresh information or expiring it after some time.
-As soon as it receives the reply, the client makes another request.
-Long polling technique results on a better network usage and a faster server response, but both simple polling and long polling requests are sent with \gls{HTTP} headers, which add data overhead and can be noticed especially for short sized messages.
-
-The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel \cite{rfc6455}, by other words it supports sending and receiving data simultaneously.
-
-  \subsubsection{Session Initiation Protocol}
-
-  \gls{SIP} \cite{rfc3261} is a protocol used for negotiation, creation, modification and finalization of communication sessions between users.
-\gls{SIP} follows a client/server architecture with \gls{HTTP} like messages and it can be used as a signaling protocol.
-The advantage of \gls{SIP} is the ability to make video and voice calls between the telephone network and applications over \gls{IP} networks.
-
-  \gls{SIP} is used in \gls{VoIP} applications due to its compatibility with the \gls{PSTN}.
-
-  Frameworks like \emph{jsSIP}\footnote{\url{http://jssip.net/}(accessed June 1, 2015).}, \emph{QoffeeSIP}\footnote{\url{http://qoffeesip.quobis.com/}(accessed June 1, 2015).} and \emph{sipML5}\footnote{\url{http://sipml5.org/}(accessed June 1, 2015).} are used on the client side to parse and encode \gls{SIP} messages, making \gls{SIP} accessible to web based applications.
-
-  \gls{SIP} with \emph{WebSockets} can be used as a signaling method for \gls{WebRTC} applications, it allows web browsers to have audio, video and \gls{SMS} capabilities like mobile phones.
-For instance, it's possible to inter-operate web communications with \gls{SIP} networks, mobile and fixed phones.
-
-\subsubsection{Extensible Messaging and Presence Protocol}
-
-  \gls{XMPP} was initially developed for instant messaging and presence (Jabber\footnote{\url{http://jabber.org/}(accessed June 1, 2015).}).
-It is nowadays an open technology for standardized, decentralized, secure and extensible real-time communications.
-
-  Today, multiple \gls{XMPP} server implementations exists, such as: \emph{ejabberd}\footnote{\url{http://jabberd.im/}(accessed June 1, 2015).}, \emph{Metronome}\footnote{\url{http://lightwitch.org/metronome}(accessed June 1, 2015).}, \emph{Openfire}\footnote{\url{http://igniterealtime.org/projects/openfire/}(accessed June 1, 2015).} and \emph{Prosody}\footnote{\url{http://prosody.im/}(accessed June 1, 2015).}.
-\emph{Ejabberd} is the server that implements more \gls{RFC} specifications and \gls{XEP}s\footnote{\url{http://en.wikipedia.org/wiki/Comparison_of_XMPP_server_software}(accessed June 1, 2015).}.
-
-    The connection to the \gls{XMPP} server can be done through the same web server as the user is connected, but this web server would handle a lot of connections and that would have a great impact on the overall system performance.
-On the other hand, users can directly connect directly to \gls{XMPP} servers from web applications using the \emph{JavaScript} library \emph{strophe.js} \footnote{\url{http://strophe.im/strophejs/} (accessed May 6, 2016).}
-
-    Typically a web application consists in multiple web pages which are navigated by users.
-Each time a web page is accessed, either from a new context or a transition from a previous page, its context is cleared except for local storage and cookies.
-The \emph{JavaScript} context is cleared, including the \gls{XMPP} connections performed by \emph{strophe.js}.
-As such, an automatic mechanism is required for avoiding the implicit user reauthentication.
-    One solution for reauthentication can be achieved by storing the user's JabberID and password on local storage and every time a page is accessed the authentication is performed without the users knowledge.
-Clearly this solution represents security flaws as the local storage can be easily accessed locally or by performing a cross site scripting attack to reveal all the local storage.
-The same problem arises if the credentials are stored on cookies.
-
-    An alternative solution is known as Session Attachment \footnote{\url{https://metajack.im/2008/10/03/getting-attached-to-strophe/} (accessed May 6, 2016).}, which requires that a \textit{session identifier}(SID) together with a \textit{initial request identifier (RID)} be passed to \emph{strophe.js} in order to re-connect to the same stream on \gls{XMPP} server.
-Either SID and RID are unpredictable and, particularly, \textit{RID} changes on every request making it worthless if a user maintains more than one tab opened, for example for multiple conversations at the same time.
-
-\subsection{Hypermedia: more than words, more than images}
-\label{hypermedia}
-
-Since the early days of video technology, one of the problems raised consisted on how to add more information onto video without generating multiple versions.
-This section examines technologies that allows different ways to present multimedia content in such a way that it change based on synchronization amongst other multimedia elements or user interaction.
-
-\emph{Hypermedia} concept brings the possibility to organize and overlay multimedia elements into a nonlinear linear structure holding the promise of future technology and features.
-
-  Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
-An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time in a video or audio track.
-\emph{HyperCafe} \cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe.
-
-  In order to navigate through a dynamic video, we must be aware of time synchronization and the multiple time flows, it ils important that all time, causality and behavior rules are well defined.
-
-  \emph{HyVAL}\cite{hyval} is an \gls{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia.
-HyVAL defines defines video structure, internal video and external media objects.
-
-  \gls{SMIL}\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films.
-With \gls{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
-
-The \gls{DOM} is a standard \gls{API} that allows easy management of documents that are organized in a tree structure, by providing \gls{CRUD} operations over its elements and their attributes.
-\gls{DOM} makes it easy to inter-operate between imperative and declarative programming languages\cite{dom}.
-
-Like \gls{DOM}, \gls{SMIL} \gls{DOM} is an \gls{API} for \gls{SMIL} documents.
-Allowing \gls{CRUD} operations over \gls{SMIL} documents is an important feature for extending \gls{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \emph{JavaScript} functions.
-
-  \gls{SMIL} fits our goals for creating a multimedia rich hyper-call, but it lacks on browser compatibility.
-Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \gls{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plug-in.
-
-  \emph{JavaScript} is an imperative object-oriented language based on \emph{ECMAScript}.
-It is used mainly on client-side and executed by a web browser.
-\emph{JavaScript} has its own implementation of \gls{DOM} and one of its advantages is the ability to download and execute code on the fly without the need of pre-installed plug-ins.
-
-  \emph{JavaScript} has compatibility issues among the different web browsers, leading to different behaviors.
-To solve that problem, there are libraries written in \emph{JavasSript}, namely \emph{jQuery}, that implements the same functionality for multiple browsers, masking most of the incompatibility issues.
-
-  \gls{SVG} is an \gls{XML} based format that incorporates the animation module of \gls{SMIL}.
-Currently, \gls{SVG} allows adding movement and animating attributes of elements.
-When embedded on \gls{HTML}5, it allows dynamic changes to inner content in real-time through the \gls{DOM} \gls{API}.
-Besides that, it also allows calling \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.
-
-  Using technologies that relies only on web standards, like \gls{CSS}, \gls{HTML}5, \emph{JavaScript} and \gls{SVG}, will make possible to develop an application that solves our problem with the advantage of being compatible with a greater amount of web browsers.
-
-\subsection{Extending collaboration tools with time manipulation}
-\label{collab}
-
-Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity.
-Technologies like these are appearing along these days, but they were not be possible a few years ago because technology was limited or unavailable.
-Although today's technology is still limited on some aspects, progress is being done in order to improve the web ecosystem, by creating standards and migrating to newer technologies.
-
-Section \ref{recstream} presents the RTP protocol how it can be used to record media streams.
-Section \ref{mediatype} describes the media types and what types of media should be streamed.
-Section \ref{intrecord} addresses how an interactive media can be recorded.
-Section \ref{collabenv} presents the collaborative environment and libraries to synchronize distributed object among users.
-
-\subsubsection{Streaming and Recording}\label{recstream}~\\
-
- Our first concern on real time collaboration applications, besides the communication itself, is the data storage and representation.
-Storing multimedia content is not a viable solution because most browsers recommend limiting local storage to at most five megabytes per origin.
-
- 	In order to provide a way to record and playback streams, additional servers will be required to process and record the large amount of data generated by audio and video streaming.
-
- \gls{RTP}\cite{rfc3550} is used for streaming audio and video over \gls{IP}.
- Multimedia content is transported on the payload of \gls{RTP} messages, that contain headers for payload identification.
-\gls{RTP} is independent from its payload type, allowing it to transport any kind of encoded multimedia.
-A sequence number is used for sorting received packets.
-
-  \gls{RTP} allows to change its requirements and add extensions to it with profiles.
-One of the most used ones is the \gls{RTP} profile for audio and video \cite{rfc3551}, which lists the payload encoding and compression algorithms.
-  Another profile for \gls{RTP} is defined by \gls{SRTP}, which provides encryption, authentication and replay protection for \gls {RTP} traffic.
-The analogous secure protocol for \gls{RTCP} is \gls{SRTCP}.
-
-  \gls{RTP} recorders are independent of payload encoding, they don't decode \gls{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.
-
-	\gls{KMS} supports streaming over \emph{WebRTC}, \gls{HTTP} and \gls{RTP} endpoints.
-Another important component of \gls{KMS} is \emph{Kurento Repository}, which supports recording and playing directly from \emph{MongoDB}.
-That is important for providing a scalable media storage.
-Unlike \emph{Jitsi Video Bridge}\footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}(accessed June 2, 2015).}, \gls{KMS} does not enforce a specific signaling protocol.
-
-\subsubsection{Media Types}\label{mediatype}~\\
-	Media Types can be distinguished by two criteria, the first one describes a media as discrete or continuous, the second one describes it as interactive or non-interactive.
-A discrete media is characterized by not depending from time, and continuous media as depending on it.
-Interactive media is characterized by its state being changed by external events such as user interactions.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.3\textwidth]{figures/media_types.png}
-	\caption{Media Types}
-\end{figure}
-	For example, an image is non-interactive and discrete, while a video is continuous and non-interactive.
-A simple collaborative editor with just text is interactive and discrete.
-An animation that changes in function of user behavior is interactive and continuous.
-
-  \subsubsection{Collaborative Environment}\label{collabenv}~\\
-
-        \gls{OT} technology was originally developed for consistency maintenance and concurrency control over distributed objects.
-        \gls{OT} algorithms are mainly used in collaborative applications such as distributed document edition.
-
-	Among mutiple \gls{OT} platforms and libraries we present \emph{ShareJS}\footnote{\url{http://sharejs.org/}(accessed June 2, 2015).}, \emph{TogetherJS}\footnote{\url{http://togetherjs.com/}(accessed June 2, 2015).}, \emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}(accessed June 2, 2015).}, \emph{Etherpad Lite}\footnote{\url{https://github.com/ether/etherpad-lite}(Accessed 20 March 2016)} and \emph{otJS}\footnote{\url{http://operational-transformation.github.io}(accessed March 10, 2016)} which we describe on Table \ref{table:otcomparision}.
-
-\begin{table}
-\centering
-\caption{Comparision between Operational Transformation libraries}
-\label{table:otcomparision}
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Library} & \textbf{Own Server} & \textbf{Own Storage} & \textbf{Operations} \\ \hline
-ShareJS          & yes                 & yes                  & text+objects        \\ \hline
-TogetherJS       & yes                 & no                   & text+objects        \\ \hline
-Goodow           & yes                 & yes                  & text+objects        \\ \hline
-Etherpad Lite    & yes                 & yes                  & extendable 			    \\ \hline
-OT.js            & no                  & no                   & text                \\ \hline
-\end{tabular}
-\end{table}
-
-	\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations over plain text on the client side.
-An implication of implementing just the client side is the extra effort that is necessary to implement content's persistent storage.
-Besides this drawback, this library is very flexible because it's not tied to a specific database or server side technology.
-
-\subsection{Chapter Summary}
-\label{related:summary}
+\gls{NAT}s weaknesses are being exposing at the application layer, namely impacting applications that require direct communications between two private networks.
 
@@
 
-\gls{WebRTC} introduced an \gls{API} for video, audio and data communications through web browsers without using plug-ins.
+\gls{STUN}, \gls{TURN} and \gls{ICE} \cite{natvoip} servers are a possible solution to overcome \gls{NAT}'s exposed weaknesses to applications that require communications between two private networks.
+
+When connection is established, \gls{WebRTC} came to simplify how audio and video are transmitted through web browsers.
+\gls{WebRTC} is an open source technology that defines a collection of standard protocols and \emph{JavaScript} \gls{API}s for web browser based real time communications without installing any additional application or plug-in.
+
+\gls{WebRTC} uses \gls{SDP} \cite{rfc4566} to define peer connection properties such as types of supported media, \emph{codecs}, protocols used and network information.
+An \gls{SDP} offer describes to other peers the expected type of communication and its details, such as used transport protocols, codecs, security and other.
+
+Some operating systems such as \emph{Android}, \emph{iOS}, \emph{Linux}, \emph{OSX} and \emph{Windows} implement native \gls{WebRTC} libraries, extending the usage of \gls{WebRTC} to applications outside the web browser.
+This native support can help to implement applications that record video and audio streams for further playback.
+
 However, \gls{WebRTC} by itself does not define how users get to know each other nor how information flows between users.
 For this reason, we have studied the multiple ways we could implement this \emph{get-to-know} mechanism which is known as signaling protocol.
+
+  Signaling is the process by which applications exchange connection information about peers and servers, their capabilities and meta-data.
+  In particular, \gls{WebRTC} does not implement signaling, as different applications may require different protocols and there is no single answer that fits all problems.
+  As a consequence, multiple options are available for filling the missing \gls{WebRTC}'s signaling component, which can be performed using \gls{SIP}\cite{rfc3261}, \gls{XMPP}, \emph{WebSockets}, \emph{Socket.io}\footnote{\url{http://socket.io/}(accessed June 1, 2015).}, \gls{SigOfly}\cite{sigofly} or by implementing a custom protocol.
+
+  One of \gls{WebRTC} signaling's requisites is bi-directional communication.
+The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel \cite{rfc6455}, by other words it supports sending and receiving data simultaneously.
 
@@
 
-\section{Conclusion}
-The conclusion goes here.
+	\emph{Hypermedia} concept brings the possibility to organize and overlay multimedia elements into a nonlinear linear structure holding the promise of future technology and features.
+Languages such as \gls{SMIL}~\cite{hyval}, \emph{HyVAL}~\cite{hyval} and \gls{HTML} can be used to implement the \emph{Hypermedia} concept.
+Two examples of applications that captured our attention were \emph{HyperCafe}~\cite{hypercafe} and \emph{Hyper-Hitchcock}~\cite{hitchcock} which explored interactive video features.
 
-\ifCLASSOPTIONcompsoc
+  Using technologies that relies only on web standards, like \gls{CSS}, \gls{HTML}5, \emph{JavaScript} and \gls{SVG}, will make possible to develop an application that applies the hypermedia concept with the advantage of being compatible with a greater amount of web browsers.
 
-  \section*{Acknowledgments}
-\else
+\subsection{Extending collaboration tools with time manipulation}
+\label{collab}
 
-  \section*{Acknowledgment}
-\fi
+Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity.
 
-The authors would like to thank...
+ Our first concern on real time collaboration applications is the data storage and representation.
+Storing multimedia content on a web client is not a viable solution because the local storage is limited to at most five megabytes per origin.
+Additional servers will be required to process and record the large amount of data generated by audio and video streaming.
+
+\gls{KMS} supports streaming over \emph{WebRTC}.
+Another important component of \gls{KMS} is \emph{Kurento Repository}, which supports recording and playing directly from \emph{MongoDB}.
+That is important for providing a scalable media storage.
+
+     \gls{OT} technology was originally developed for consistency maintenance and concurrency control over distributed objects.
+     \gls{OT} algorithms are mainly used in collaborative applications such as distributed document edition.
+
+	Among mutiple \gls{OT} platforms and libraries we present \emph{ShareJS}\footnote{\url{http://sharejs.org/}(accessed June 2, 2015).}, \emph{TogetherJS}\footnote{\url{http://togetherjs.com/}(accessed June 2, 2015).}, \emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}(accessed June 2, 2015).}, \emph{Etherpad Lite}\footnote{\url{https://github.com/ether/etherpad-lite}(Accessed 20 March 2016)} and \emph{otJS}\footnote{\url{http://operational-transformation.github.io}(accessed March 10, 2016)}.
+
+	\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations over plain text on the client side and requires implementing the content's persistent storage.
+Besides this drawback, this library is very flexible because it's not tied to a specific database or server side technology.
+
+\section{Architecture}
+\label{chapter:architecture}
+
+Taking into account the goals of this project and all the technology presented so far.
+Our proposal is the development of a web application that provides communication and collaboration features in real-time.
+
+\subsection{Requirements}
+In a general way our system's goal is to provide a multi-party video and audio conference environment that supports chat, time manipulation, collaborative text edition and hyper-content creation.
+
+%RP isto pode ser muito expandido. Podes começar por descrever os goals (mais alto nível): audio e video conferência multi-party, possibilidade de rever o que foi dito antes, chat, etc etc.
+% Só depois traduzes isso em requisitos, tanto funcionais como não funcionais
+% Depois apresentas a arquitectura, fazendo um paralelo entre os módulo e escolhas realizadas e quais os requisitos que cada um deles satisfaz.
+
+For our system, our application must provide: a simple way to send instant text messages to the conference participants, ability to recording and playback recorded video including all the hyper-content displayed at that time, mixing multiple user streams into a single stream, create annotations associated to a specified time, allow users to superimpose hyper-content to video given a range of time, search every objects related to a conference room such as hyper-content, time annotations and users, share files and time links among users, sound detection for showing the current speaker, provide a collaborative text edition tool, interpret \gls{QR} codes in order to ease content creation and support database replication.
+%RP capacidade de lidar com NAT
+
+	Moreover we allow clients to discover chat rooms and other clients by navigating on the web pages provided by our web server.
+In addition users can create rooms for multi-party audio and video communication communication which is achieved by using \gls{WebRTC}'s \emph{PeerConnection}.
+        %RP falas em peerconnection, mas na figura 3.1 apenas há um cliente, o que obviamente impede mostrar a comunicação entre peers.
+        %RP acho que podes fazer outro diagrama com a comunicação entre os vários componentes, onde apareceriam vários clientes.
+A aí indicas o tipo de ligação entre cada par de componentes (HTTP, PeerConnectino, websocket, MediaStream ...)
+
+\subsection{Modules}
+In this section we present the several modules that were designed in order com fulfill the set requirements.
+	Figure~\ref{fig:modules} presents the structure of our system which was divided into six modules:
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\linewidth]{figures/modules.pdf}
+	\caption{System Modules}
+    \label{fig:modules}
+\end{figure}
+
+\begin{itemize}
+
+\item \textbf{Application module} - responsible for providing information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client} in the form of web pages in \gls{HTML} and \emph{JavaScript} libraries through \gls{HTTP}.
+
+\item \textbf{Signaling module} - responsible for \emph{Client} and \emph{Stream} coordination which will be performed using \emph{WebSockets}.
+
+ \item \textbf{NAT Traversal module} - \gls{STUN} and \gls{TURN} techniques used by \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase which ends by establishing the connection between them.
+
+ \item \textbf{Stream module} - responsible to deliver and receive multimedia content from the \emph{Client} using \gls{WebRTC}.
+
+ \item \textbf{Storage module} - provides two main functionalities: store the model information and media recorded.
+This is the single module responsible for persistent storage.
+It stores user and communication data as well as all the data required among user sessions.
+It is also use to store all the communication streams, so that they can be viewed later.
+
+ \item \textbf{Client module} - responsible for the interaction with the user.
+
+\end{itemize}
+
+%RP Acho que podes escrever muito mais sobre cada componente. Isto está muito curto. Fala de algumas das funções que proporcionam. Vê o que fiz com o Storage Module
+
+%RP também podes fazer um diagrama com vários utilizadores e mostrar quais os streams que vão de e para cada um (mostrar a mistura de som, video, etc).
+
+\subsection{Implementation Proposal}
+The infrastructure is composed by: web server, stream server, signaling server, database and video repository.
+%RP tens da fazer o paralelo entre isto e os componentes apresentados na secção anterior.
+O mesmo com a imagem, que parece não estar relacionada.
+Podes colocar os elementos na mesma posição, ou dentro das caixas da imagem 3.1
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=\linewidth]{figures/infrastructure.pdf}
+	\caption{System Infrastructure}
+\end{figure}
+
+In order to simplify our solution, we propose that the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules be implemented within the same server application, so it could be easy to deploy as a single image.
+To this set of modules, we often call \emph{backend}.
+
+	\subsubsection{Security and authorization}
+
+Having established that the \emph{backend} modules are placed in the same machine, that helps controlling which resources the client has permission to access as those modules are seen as a private network.
+%RP na mesma ``server application'' ou ``machine''?
+São coisa diferentes!
+Apresentas vantages de instalar tudo na mesma máquina mas não falas das desvantagens.
+Menor escalabilidade parece-me ser logo uma delas.
+Na figura 3.2 aparecem várias máquinas!
+
+To qualify the above we provide public access to \gls{HTTP} server ports, maintaining the access to other components restricted through firewall rules.
+%RP qualify?
+%RP Imagem com firewall e estes portos? Mas isto já me parece muito detalhe de implementação.
+In relation to the database there is no directly access from the outside. All the database information is accessed via our application server which validates the permissions of users on our system.
+
+On the other hand, the access to our streaming servers is also restricted, but clients can connect to them after concluding the signaling phase.
+This signaling phase may or not proceed in function of the client access permissions.
+For example, if a user is trying to access a private conference room that he is not a member of nor has an invitation link for, the signaling server refuses to start the signaling phase and the user cannot access the streaming server.
+
+The placement of our streaming servers inside a \gls{NAT} also has an important role with respect to external misuse prevention.
+Otherwise, placing our streaming servers could allow external clients to perform their own signaling protocol and, as a consequence, use our infrastructure without our consent.
+
+\subsubsection{Client connections}
+	Although the delegation of processing work to clients can improve our system's scalability, we are concerned about using the least resources possible on the client side, as huge resource consumption may drain battery very fast or may even be impossible to run on mobile devices.
+We are aware that streaming video from clients is already a very intensive task which we cannot avoid but can improve by delegating the most intensive tasks to our servers.
+
+	In this context, with a centralized approach, each client must only have one \emph{PeerConnection} to our streaming server and content shown to them is changed on demand either being it an individual or a composite view.
+Otherwise clients could follow a peer-to-peer connection which would result on maintaining more connections and performing the composition of videos on client side.
+
+	The composition of streams on client side is performed by receiving streams with the best quality possible but, due to undersizing the video of clients into a smaller region, this would result on wasting bandwidth on a quality that is not needed.
+
+	Although the peer-to-peer approach could be used on our system, we conclude that we need to record the video on our streaming server because web clients have a very limited storage and peer disconnections may result on recorded video loss.
+
+	The same can be concluded to instant message delivery, each client must have only one \emph{WebSocket} connection to the application server which consequently relays the messages to other users.
+
+	Relaying instant messages from clients through the our web servers is easier if all clients are connected to the same server because all messages can be directly delivered without sending messages across multiple web servers.
+
+	In the context of this thesis, we will not implement sending messages across web servers but, in order to allow our system to scale, we will consider that all conference participants are connected to the same server and our system is scaled by having conference rooms distributed across different servers.
+
+\subsubsection{Software choices}
+
+Furthermore, we have taken into account the compatibility between the streaming server, database and the operation transformation solutions, in order choose the appropriate framework to implement our web server.
+%RP esta secção devia começar aqui, a explicar a escolha das tecnologias/produtos a usar em cada componente.
+O que tens antes são pormenores que só faz sentido discutir no final.
+Tens de ir sempre do mais geral para o mais detalhado.
+
+We have decided that our solution must use \gls{KMS}.
+Our web server could be implemented easily with \emph{NodeJS} or \emph{Java} due to the fact \gls{KMS} provides clients for both technologies.
+But others could also be used as \gls{KMS} also exposes their \gls{API} via \emph{WebSockets}.
+
+Due to the fact we are going to use \gls{KMS} as streaming server solution, we could choose \emph{NodeJS} or any \emph{Java} based web framework for implementing our web application server.
+We have decided to implement our web server with the \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java} because of our previous experience with it.
+
+By default, \emph{Kurento Repository} is implemented over \emph{MongoDB}, for convenience our storage model will also use the same database.
+
+Importantly, for the collaborative text editor, we have chosen \emph{OT.js} due to its server and storage implementation choice independence.
+
+For the \emph{NAT Traversal} module a public \gls{STUN} server can be used for testing our solution.
+Nevertheless, we recognize that for a production environment we would need to maintain our own \gls{TURN} servers in order to ensure connectivity to all clients.
+
+Not less important, on the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} could be installed as web browsers.
+As such, both should be supported.
+Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Adapter.js}, \emph{OT.js} can be downloaded from the web server and executed on the client side using any of these two browsers.
+%RP relembrar que o IE e o Safari não suportam webrtc?
+
+\begin{table}[H]
+\centering
+	\caption{Application Architecture}
+	\label{table:apparch}
+
+\resizebox{\columnwidth}{!}{%
+    \begin{tabular}{cccccccc@{}m{0pt}@{}}
+	\hline 
+	\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
+	\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[12pt]\hline
+	\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
+	\end{tabular}
+}
+\end{table}
+
+%RP acho interessante o que queres fazer aqui mas acho que está mal conseguido. A linha de baixo, devia ter coisas ao mesmo nível, o que não aconteçe.
+% Qual a razão para HTTP estar por baixo do jquery e user interface estar ao lado?
+% Sugiro colocar a applicação no meio e as várias interfaces (webrtc, websocket e http) uma de cada lado (esquerda, baixo e direita), indicando o que fica do outro lado (outro cliente, servidor web, servidor stream, etc)
+
+Table~\ref{table:apparch} presents the application architecture and the underlying technologies seen from the user's perspective.
+\emph{Adapter.js} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.
+\emph{Bootstrap} will be used to make the user interface more appellative and responsive.
+With \emph{Bootstrap} it is quite easy to develop applications that adapt to mobile devices with different screen sizes.
+%RP bootstrap não aparece na figura!
+
+With respect to displaying content, the synchronization between multimedia elements will be performed through chains of \emph{JavasSript} events or by specifying the interval of time which time content must be visible.
+Other animations can be implemented with \gls{SVG} embedded on \gls{HTML}.
+%RP Caiu do céu! Tens de explicar e ligar melhor a sequencia de texto
+
+\section{Implementation}
+
+\subsection{Data Model}
+
+The data model is a critical component of our solution, as a badly designed model can imply serious difficulties when implementing new features that are not part of the plans.
+During the course of this project, we had to redesign the model more than once in order to support new features.
+
+In order to offer all the functionalities that we promise, some information about objects must be persistent such as users, groups,relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.
+For designing our model we have taken into account generic programming techniques.
+We observed that operations like searching for an object were quite repeated across different types of objects.
+
+\subsection{Signaling Protocol}
+
+Although we have mentioned that the signaling protocol is used to establish connections between peers, on our system our media server (\gls{KMS}) is a peer that receives video streams and sends to its connected clients.
+
+After the web application server validates the user access, the signaling protocol allows the users to directly connect to the \gls{KMS}, which is placed in a private network, and lets the application server and users negotiate media types and encoding information to use during the conversation.
+
+Our signaling protocol consists of sending and receiving \gls{JSON} formated messages over \emph{WebSockets} by both the application server and the client.
+
+When a user enters a group conference, after the page is completely loaded, a WebSocket is created to maintain a connection with our web servers. 
+But before creating the web socket, we must identify the user and check if he has permissions to participate in the conference.
+The user identification is done by retrieving the session id from the cookie provided by the user-agent (web browser) through the \gls{HTTP} headers.
+%RP não é bem o user que fornece o cookie. é o user-agent (browser). Convém ser preciso.
+%HR done
+The Web application server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room.
+It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake is performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \gls{HTTP} context is lost.
+%RP socket -> websocket?
+
+At this stage, the web application's user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server.
+
+If the user decides to share his camera or screen the user agent creates an offer, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the WebSocket to the Application Server.
+
+The server receives and processes the offer and sets the \emph{remote session description} to its client associated \gls{WebRTC} endpoint.
+Then a \emph{local session description} is created on the server and sent back to the client, after that the server tries to gather \gls{ICE} candidates.
+
+The client receives the server answer, sets the \emph{remote session description} and gets the ice candidates from the \gls{ICE} server.
+%RP tens ``ice'', ``\gls{ICE}''. Acho que também já vi ``ICE''. Convém ser sempre igual!
+
+Subsequently, after a while both the server and client receive the \gls{ICE} candidates that allow the client to connect directly to \gls{KMS} and vice-versa.
+The candidates are received at the client which sets them to its \emph{PeerConnection}.
+The same is done on the server which receives the \gls{ICE} candidates from the client and propagates them to \gls{KMS}.
+
+An \gls{ICE} candidate contains an \gls{IP}, port, used transport protocol and an attribute named \emph{sdpMLineIndex} that is used for mapping to the \emph{remote session description} media type.
+When a connection is established, the user and server start to interchange stream data but other \gls{ICE} candidates may arrive with better connections.
+
+Having the media session established, the server starts to record any received stream and the client creates an \gls{URL} correspondent to the stream location.
+
+\subsection{Stream Recording}
+ Initially we experimented with local recording and synchronizing with our servers. Even thought it got it to work, it consumed to much bandwidth. We then tried recording on server side.
+
+	One of our concerns during the development of our solution was the storage scalability.
+Saving files directly into the file system would require an extra effort to distribute and replicate files among servers.
+For that purpose, the \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB}.
+
+	With server side recording, the user would maintain always the same stream \gls{URL} even if it is playing real-time video or reproducing recorded video.
+It is \gls{KMS} that sends different content through that stream.
+When a user desires to play recorded video, a \emph{webSocket} message is sent specifying the time and the intended user id.
+The server performs the calculations in order to find a block that intersects the requested time, plays it and when finished, the next part is automatically played without the user intervention.
+
+%RP não falas nada sobre chats, etc! Não há signalling para isso?
+%HR não, só video e audio, chat é websockets
+
+\subsection{Hyper-Content}
+
+	Our system supports creating superimposed content to video, which is achieved by creating \gls{HTML} tags on top of the video  with the same size.
+The decision of which content must be displayed to each user is performed by our content scheduler which uses the user's current time in order to synchronize which content is shown or removed from the user interface.
+
+	In order to create content, the user has the option to write simple movie captions without writing any code, otherwise, as mentioned before, it can write \gls{HTML}, \gls{CSS} and \emph{JavaScript}.
+
+	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real-time, we allow the user to encode its content into \gls{QR} codes and show it to the camera in real time.
+
+\subsection{Collaborative editor}
+
+Our collaborative editor is a simple text editor that is synchronized with all participants within a conference room implemented using \emph{ot.js}.
+
+The state of our collaborative editor is not saved on the database every time it changes.
+Instead, the users just synchronize the editor content among themselves using the application server to relay editor changes and save on demand.
+
+\section{Evaluation}
+
+\subsection{Tests Objectives}
+
+  We have tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.
+
+  We have also tested the performance of our solution by measuring the used resources.
+Those performance tests are crucial to ensure that our solution is in fact stable and users can use it endlessly without decreasing the quality of their experience.
+
+  \subsection {Performance Tests}
+
+      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects with a periodicity of one second: CPU, physical memory information relative to each running process and network usage relative to each interface.
+
+      The performance test scenario that we have defined consists on two phases, the first phase consists only on having users, with similar computer and network specifications, entering sequentially on the conference room.
+The second phase consists on the users leaving the conference room.
+Each event, joining and leaving, occurs with intervals of one minute in total of thirteen minutes (780 seconds).
+
+      From the media server perspective if there are $n$ clients connected, each of them sending and receiving one stream, it is expected that server sends and receives also $n$ streams.
+As such, we expect that the amount of network traffic increases linearly as users join a conference room.
+Figure \ref{fig:test_full_features_net} confirms our expectations.
+Each vertical yellow line represents one event: the first seven events are users entering the conference room, the next ones represent users leaving the conversation.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{stats/test_full_features_net.eps}
+  \caption{Network usage after implementing all features}
+  \label{fig:test_full_features_net}
+\end{figure}
+
+      The blue peaks are caused by the signaling phase and web page downloads, including resources such as images, stylesheets and javascript files. 
+      The green peaks are caused by video and audio being transfered between \gls{KMS} and \emph{MongoDB} through \emph{Kurento Repository}.
+Each peak occurs every time a block of video is recorded, which in this case is every ten seconds.
+      The recordings are synchronized so all user and mixed blocks start and end at the same time.
+That is why the amount of work done every ten seconds accumulates, and because this is performed locally, the maximum transfer rate is limited by the performance of the memory as buffers are written to buffers then to disks.
+      Sent data transfer rate has no significant peaks as \gls{HTTP} requests and signaling information contains little information.
+
+	With this results we conclude that if we want to scale our storage solution using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \gls{KMS} should be installed on the same machine because the loopback interface can handle bigger transfer rates than the remaining network interfaces.
+Installing the repository on the same machine as a database node does not ensure that recorded videos are stored in the same machine, for this reason we would prefer installing \gls{KMS} and \emph{Kurento Repository} on the same machine.
+
+On the other hand, Figure \ref{fig:test_client_net} shows the respective network usage on the client side during our test case.
+We observe that in the first seconds the client adjusts the video quality it sends to \gls{KMS}.
+Whenever a new client enters the conference, we observe that \gls{KMS} decreases the video quality in order to instantaneously integrate a new user into the conference room.
+After a while, \gls{KMS} realizes that the network can handle the increase of clients and sends the video with a better quality to every participant.
+When a user leaves the conference room \gls{KMS} has no need to decrease the participant's video quality as less network bandwidth will be used.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{stats/test_client_net.eps}
+  \caption{Client Network usage during our test case}
+  \label{fig:test_client_net}
+\end{figure}
+
+\textbf{Memory usage}
+
+Figure \ref{fig:test_ram_fixed_mem} shows the memory usage during our performance test.
+Both \gls{JVM}, \emph{MongoDB} and \gls{KMS} performs their own memory management by holding and recycling objects when needed.
+The expected and observed behavior of the memory usage is growth of memory usage while the users are entering the conference room and a memory usage stabilization afterwards.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{stats/test_ram_fixed_mem.eps}
+  \caption{Memory usage after fixing recorder memory leak}
+  \label{fig:test_ram_fixed_mem}
+\end{figure}
+
+\emph{MongoDB} memory usage keeps increasing because it tries to fit part of the database on \gls{RAM} for fast read access.
+\emph{MongoDB} checkpoints data to disk every 60 seconds or when journal data exceeds 2GB\footnote{\url{https://docs.mongodb.org/manual/faq/storage/}(Accessed March 28, 2016)}, that explains the small memory usage peaks during our test case.
+When the conference room is empty there are no video recordings, which explains the memory stabilization at the end.
+
+\emph{KMS} memory management released resources as soon as users left the room.
+
+\textbf{CPU usage}
+
+Figure \ref{fig:test_full_features_cpu} shows the percentage of \gls{CPU} usage during our performance test case.
+Each 100\% represents one \gls{CPU} core, although that does not mean one \gls{CPU} is fully used, for example two cores at 60\% represent 120\% \gls{CPU} usage.
+As we can see, the percentage of \gls{CPU} used increases and decreases linearly in function of the amount of conference participants.
+\gls{KMS} is responsible for most \gls{CPU} usage.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{stats/test_full_features_cpu.eps}
+  \caption{Percentage of CPU used during the performance tests}
+  \label{fig:test_full_features_cpu}
+\end{figure}
+
+  Just for testing purposes, we performed the same performance tests disabling \gls{QR} codes detection.
+We concluded that \gls{QR} code detection is a very intensive task, approximately doubling the amount of work performed by the \gls{CPU}s.
+
+	Even though, with this test results, we conclude that our solution's bottleneck is the \gls{CPU} usage at \gls{KMS}.
+
+\subsection {Usability Tests}
+     In this section we describe usability test scenarios that we have applied and their respective results.
+
+    \subsubsection{Tests Scenarios}
+
+      In order to evaluate the usability of our solution, we have performed usability tests with the help of real users with different backgrounds and ages.
+
+      We handed a guide to the users with five tasks to perform. The metrics we used for each task were: number of clicks, number of errors (including a description) and time spent.
+
+  \subsubsection{Test Results}
+
+In this section we present the results of our usability tests.
+The first time we tested our solution by providing tasks to users, we have observed that our solution was not perfect.
+Having faced usability problems during our tests, he had to improve our solution's usability and start the tests again.
+
+In the first testing phase we have performed tests with just three users and ceased for improvements. We gathered comments and suggestions after letting the users explore our system.
+
+On the second phase of our user interface tests, in a general way, we have noticed great improvements on the learning time.
+
+In order to measure the users learning speed, we have performed tests with experienced users in order to retrieve the optimal task duration and minimal task clicks.
+
+To this end, with regard to optimal task duration and minimal clicks, we obtained the values shown in table \ref{table:optimal}.
+
+\begin{table}[H]
+\centering
+\caption{Metrics for an experienced user}
+\label{table:optimal}
+\begin{tabular}{|c|c|c|c|c|c|}
+\hline
+\textbf{Task} & 1 & 2 & 3 & 4 & 5 \\ \hline
+\textbf{Duration (seconds)} & 20 & 35 & 30 & 25 & 35 \\ \hline
+\end{tabular}
+\end{table}
+
+From the data collected with twenty tests with users, namely the task duration (Figure \ref{fig:user_times}) and task difficulty (Figure \ref{fig:user_diffs}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.
+
+The youngest and oldest testers were, receptively, twenty two and thirty eight years old. Figure \ref{fig:user_ages} the ages of the users that tested our system.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\linewidth]{stats/user_ages.eps}
+  \caption{Ages of the users that tested our system}
+  \label{fig:user_ages}
+\end{figure}
+
+As a result of both true average and variance being unknown and the usage of a relatively small amount of samples, we had to use \emph{t-distribution} to estimate our metrics confidence intervals.
+We have used a 95\% confidence level.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=0.8\linewidth]{stats/user_times.eps}
+    \caption{Time spent per task}
+    \label{fig:user_times}
+\end{figure}
+
+According to Figure \ref{fig:user_diffs}, we can observe that most users had less difficulties with the first two tasks, which represents types of tasks that most users are familiar with.
+As soon as users had to navigate in time, manipulate annotations and create content (respectively \emph{task3}, \emph{task4} and \emph{task5}) we observed that they revealed more difficulties.
+Most of those difficulties, based on the users feedback, were mainly due to those concepts not being familiar to them.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=0.8\linewidth]{stats/user_diffs.eps}
+  \caption{Difficulty per task}
+  \label{fig:user_diffs}
+\end{figure}
+
+As we had relatively bad results with some users, we explained to those users that could not conclude the tasks or performed them incorrectly, the most efficient way to perform the requested tasks.
+Some users suggested to display more hints in order to achieve a faster learning, but afterwards all were impressed and gave us a better evaluation.
+
+Most users gave us worse evaluations on our user interface layout and content editor, which was due to having a lot of tools present in the same web page and some of them being hidden due the screen size.
+In some cases users had to scroll down in order to find the tools they were looking for.
+
+Another weak aspect was our content editor, which, in fact, we recognize is difficult to work with, mostly due to the amount of information that is necessary to create a synchronized content (starting time, duration and content itself).
+Some users have suggested that the content should also be present on the time-line so they could be easily dragged and resized (on time).
+
+We are aware that placing content on the time-line will reduce our solutions performance, especially when there is a relatively large amount of content, due to the content that is present on the time-line being loaded all at once.
+Although, we recognize that for some cases (relatively low amount of content), displaying the content on the time-line could not have a great impact on our solutions performance, we choose not to implement this.
+
+In conclusion, 100\% of our testers though that our solution was an innovation and 95\% recommended using our solution.
+
+\section{Conclusions}
+\subsection{Achievements}
+\label{section:achievements}
+
+	We have successfully implemented the basic functionalities of our prototype and spare some time for adding more valuable features such as the ability to create content by exposing \gls{QR} codes to the camera and lastly perform changes to the user interface in order to improve the quality of user's experience.
+
+	The performance tests that we have executed showed that our system is stable and more importantly that our web server is lightweight and most of processing power is dedicated to the streaming server.
+
+	Our usability tests show results that are considerably worse than the established optimal values due to our solution propose a different way to communicate that most people are not used to.
+Although we have obtained those results, in general our users gave us positive feedback and valuable advices which we have used to improve our system.
+
+\subsection{Future Work}
+\label{section:future}
+	Playing back video with a faster rate is not possible using the current version of \gls{KMS}.
+Even though we expect the availability of that feature in a near future but we have proposed an alternative way to implement faster playback by using \emph{ffmpeg} to convert the video before playing it.
+
+	Although we have tested our solution in a powerful machine for the current time, our performance tests revealed that the streaming component uses a lot of resources.
+We left for a future work a deep analysis on the scalability of our system for which we have proposed different approaches but we have not tested them.
+
+	Another aspect we could have tested was the performance of our solution when using \gls{TURN} servers for relaying the traffic that fails using \gls{STUN}.
+
+	Lastly, we have chosen functionality over security in respect to displaying content to users which lead to security flaws on our solution.
+Although we have not solved the security problems, we have proposed a solution which at the same time limits the flexibility of adding new functionalities to our prototype.
 
@@
 
-\input{thesis.bbl}
+%\input{thesis.bbl}
+\printbibliography
 \end{document}

### 2016-05-13 tese2/abstract.tex
caneco

@@
 
-Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging and a mechanism to superimpose hyper-content to video.
+Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging, ability to superimpose hyper-content to video and the possibility to playback communications.
 
@@
 % TODO 4 to 6 keywords;
-\textbf{\Large Keywords:} WebRTC, asynchronous, communications, collaboration
+\textbf{\Large Keywords:} WebRTC, asynchronous communications, collaboration tools, timeline navigation, non-linear video-conferencing, video recording
 

### 2016-05-13 tese2/resumo.tex
caneco

@@
 
-A voz e a imagem justas exprimem emoções como nenhum outro tipo de media permite. Através dos conceitos de hiper-media, podemos adicionar mais valor às comunicações por video-conferência.
+A voz e a imagem justas exprimem emoções como nenhum outro tipo de media permite. Através dos conceitos de hiper-media, podemos adicionar mais valor às comunicações por vídeo-conferência.
 
@@
 
-O nosso objectivo neste projecto é desenvolver de uma aplicação, recorrendo ao \ac{WebRTC}, que utiliza o conceito de comunicações hiper-ligadas de forma a enriquecer as comunicações em conferência com vários tipos de media, editores de texto colaborativos, anotações temporais, envio de mensagens e possilidade de sobrepor contéudo ao vídeo.
+O nosso objectivo neste projecto é desenvolver de uma aplicação, recorrendo ao \ac{WebRTC}, que utiliza o conceito de comunicações hiper-ligadas de forma a enriquecer as comunicações em conferência com vários tipos de media, editores de texto colaborativos, anotações temporais, envio de mensagens, possilidade de sobrepor contéudo ao vídeo e voltar atrás no tempo.
 
@@
 % TODO 4 to 6 keywords;
-\textbf{\Large Palavras-chave:} WebRTC, comunicações, asíncronas, colaboração
+\textbf{\Large Palavras-chave:} WebRTC, comunicações asíncronas, ferramentas colaborativas, navegação temporal, vídeo-conferência não linear, gravação vídeo
 

### 2016-05-13 tese2/introduction.tex

@@
 
-	As Martin Geddes states\cite{geddes}, \say{No computer in our lifetimes will ever rival a human voice's capacity to conveying rich and complex social and emotional meaning}.
+	As Martin Geddes states, \say{No computer in our lifetimes will ever rival a human voice's capacity to conveying rich and complex social and emotional meaning}~\cite{geddes}.
 Although nothing replaces the physical contact with a person while we communicate, we are at a time when we can do more than just visual and verbal communication.
 Hypermedia can be added to video and voice in order to extend its value.
-The concept of structured voice and video synchronized with hypermedia is called hypervoice\cite{geddes}.
+The concept of structured voice and video synchronized with hypermedia is called hypervoice~\cite{geddes}.
 
@@
 	One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again or introduce hyper-content to it such as time annotations, interactive lists of topics and subtitles.
-In this context we also provide a simpler method for creating and synchronizing hyper-content using \ac{QR} codes.
+We call this non-linear video-conferencing.
+In this context, we also provide a simpler method for creating and synchronizing hyper-content using \ac{QR} codes.
 
-	In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links and files to conference participants.
+	In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links (to navigate to a location in time) and files to conference participants.
 
-	Furthermore, another relevant feature is the possibility to compose multiple video streams into a single one, which enables adding more users to conference rooms without impacting on clients performance.
+	Furthermore, another relevant feature is the possibility to compose multiple video streams into a single one, which enables adding more users to conference rooms without impacting the clients' performance.
 Users can change to individual streams on demand or automatically to the talking users.
@@
 	According to Martin Geddes, the quality of the interaction worsens as the number of users increase\cite{geddes}.
-In our testing phases we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.
+In our testing phases we quantifyed and qualifyed the impact of increasing users on the interface and performance of our prototype.
 

### 2016-05-22 tese2/06_related_work.tex
[ADD] template2
[FIX] front cover typo

--- 2016-03-12 tese1/06_related_work.tex
@@
-\section{Related Work}\label{related}
  This section is structured as follows.
- Section \ref{early} describes the problems that real-time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
+ Section \ref{early} describes the problems that real time communications face on nowadays internet, namely the \ac{IPv4} address exhaustion and the client server model constraints. 
  Section \ref{rtc} describes the \ac{WebRTC} technology and the protocols needed to implement our project. 
@@
  Section \ref{collab} explores streaming protocols for non-interactive multimedia and how to introduce the interactive component, another important aspects are the ability to control the time flux of a stream and collaborative application development.
-%RP este detalhe sobre a estrutura da sec. do estado da arte vai para o início dessa secção.

### 2016-05-22 tese2/07_related_work_nat.tex
[FIX] front cover typo

@@
 %RP o porto não identifica uma máquina (membro da rede privada) mas uma ligação.
-This technique is known as \ac{NAT} (figure \ref{fig:nat}).
+This technique is known as \ac{NAT} (Figure \ref{fig:nat}).
 
@@
 
-\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}. None of those can establish direct connections on multiple level \ac{NAT}s (figure \ref{fig:multinat}).
+\ac{STUN} and \ac{TURN} \cite{natvoip} servers are a possible solution to overcome \ac{NAT}. None of those can establish direct connections on multiple level \ac{NAT}s (Figure \ref{fig:multinat}).
 %RP explicar o que são multiple level pois não falaste disto antes? São ``nested''?
@@
 Clearly \ac{TURN} requires a more expensive relaying infrastructure and, in most cases, more network usage, leading to a worse quality of service.
-The requirements of real-time video communication makes this kind of model unsuitable.
+The requirements of real time video communication makes this kind of model unsuitable.
 

### 2016-05-22 tese2/09_related_work_signaling.tex
caneco
[FIX] front cover typo

@@
   \ac{SIP} with \emph{WebSockets} can be used as a signaling method for \ac{WebRTC} applications, it allows web browsers to have audio, video and \ac{SMS} capabilities like mobile phones.
-For instance, it's possible to inter-operate web communications with \ac{SIP} networks, mobile and fixed phones.
+For instance, it is possible to inter-operate web communications with \ac{SIP} networks, mobile and fixed phones.
 
@@
   \ac{XMPP} was initially developed for instant messaging and presence (Jabber\footnote{\url{http://jabber.org/}(accessed June 1, 2015).}).
-It is nowadays an open technology for standardized, decentralized, secure and extensible real-time communications.
+It is nowadays an open technology for standardized, decentralized, secure and extensible real time communications.
   \ac{XMPP} messages are \ac{XML} based, which is attractive for applications that need structured messages and rich hypermedia.
@@
     The connection to the \ac{XMPP} server can be done through the same web server as the user is connected, but this web server would handle a lot of connections and that would have a great impact on the overall system performance.
-On the other hand, users can directly connect directly to \ac{XMPP} servers from web applications using the \emph{JavaScript} library \emph{strophe.js} \footnote{\url{http://strophe.im/strophejs/} (accessed May 6, 2016).}
+On the other hand, users can directly connect directly to \ac{XMPP} servers from web applications using the \emph{JavaScript} library \emph{strophe.js}.
+\footnote{\url{http://strophe.im/strophejs/} (accessed May 6, 2016).}
 
@@
 
-  Another interesting approach for signaling would be \ac{SigOfly} which allows inter-domain real-time communications, while abstracting the protocol used~\cite{sigofly}.
+  Another interesting approach for signaling would be \ac{SigOfly} which allows inter-domain real time communications, while abstracting the protocol used~\cite{sigofly}.
 %RP o ~ antes do cite faz com que não haja um espaço mas não haja uma mudança de linha entre a última palavra e a citação

### 2016-05-22 tese2/10_related_work_hypermedia.tex
[FIX] front cover typo

@@
 
-  \item The\textbf{Linking} module contains elements and attributes for navigational \emph{hyperlinking}. Navigation can be triggered by events or user interaction.
+  \item The \textbf{Linking} module contains elements and attributes for navigational \emph{hyperlinking}. Navigation can be triggered by events or user interaction.
 
@@
 Currently, \ac{SVG} allows adding movement and animating attributes of elements.
-When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real-time through the \ac{DOM} \ac{API}.
+When embedded on \ac{HTML}5, it allows dynamic changes to inner content in real time through the \ac{DOM} \ac{API}.
 Besides that, it also allows calling \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.
@@
 Compared to \emph{flash}, which uses \emph{ActionScript}, \emph{Silverlight} applications can use languages like C\#, \emph{VisualBasic} and \ac{XAML}.
-\emph{Silverlight} uses a technique called \emph{Smooth Streaming} from \emph{IIS Media Service} that consists on delivering video in real-time with adjusted quality in function of bandwidth variations and \ac{CPU} usage.
+\emph{Silverlight} uses a technique called \emph{Smooth Streaming} from \emph{IIS Media Service} that consists on delivering video in real time with adjusted quality in function of bandwidth variations and \ac{CPU} usage.
 

### 2016-05-22 tese2/11_related_work_collaboration.tex
caneco
[FIX] front cover typo

@@
 \subsection{Streaming and Recording}\label{recstream}~\\
-	If, for instance, one wants to rewind a real-time video, recordings will be needed from whom is streaming the video. 
+	If, for instance, one wants to rewind a real time video, recordings will be needed from whom is streaming the video. 
  Our first concern on real time collaboration applications, besides the communication itself, is the data storage and representation.
-Storing multimedia content is not a viable solution because most browsers recommend limiting local storage to at most five megabytes per origin.
+Storing multimedia content is not a viable solution because most browsers recommend limiting local storage to at most five megabytes per origin\footnote{\url{http://www.html5rocks.com/en/tutorials/offline/quota-research/} (accessed May 13, 2016)} .
 
@@
 
-\ac{RTCP} is used for controlling \ac{RTP} multimedia streams, it provides bandwidth statistics and control information that can be used for changing the quality of the stream in real-time.
+\ac{RTCP} is used for controlling \ac{RTP} multimedia streams, it provides bandwidth statistics and control information that can be used for changing the quality of the stream in real time.
 
@@
 	Multi-party video calls can be achieved on \ac{WebRTC} by streaming video from each participant to all the other participants.
-Although this works, the bigger a conference room is, the bigger is the bandwidth used to stream video to all participants within the conference room.
+Although this works, the bigger a conference room is, the bigger is the used bandwidth to stream video to all participants within the conference room.
         In this scenario, a more efficient alternative to peer-to-peer is the use of a \ac{MCU}.
@@
  \subsection{Recording and Streaming Interactive Media}\label{intrecord}~\\
-    Mauve et al.
-proposed an \ac{RTP} profile for real-time transmission of interactive media\cite{interactive_stream}.
+    \emph{Mauve et al.} proposed an \ac{RTP} profile for real time transmission of interactive media\cite{interactive_stream}.
 This new profile reuses much of video and audio profile implementation, integrating the interactive component.
-Hilt et al.
-explained how to record interactive video with this new profile \cite{interactive_record}.
+\emph{Hilt et al.} explained how to record interactive video with this new profile \cite{interactive_record}.
         %RP não uses a citação como sujetio das frases. O trabalho não tem um nome?
@@
 	An \ac{RTP} recorder can have two operation modes, recording or playback.
-Traditional \ac{RTP} players can do random access, in contrast, interactive \ac{RTP} players must restore the environment and context at a given time.
+Traditional \ac{RTP} players can do random access.
+In contrast, interactive \ac{RTP} players must restore the environment and context at a given time.
 The environment is the initial state, so we can call it a non-interactive discrete media and handle it over \ac{HTTP}.
@@
 
-	Mauve et al.
-concluded that the ability to extract the objects state in order to synchronize them and the ability to intercept events in order to control remote objects can be realized using the \ac{MVC} concept\cite{interactive_stream}.
+	\emph{Mauve et al.} concluded that the ability to extract the objects state in order to synchronize them and the ability to intercept events in order to control remote objects can be realized using the \ac{MVC} concept\cite{interactive_stream}.
 %RP are realized ou ``can be realized''?
@@
 
-	\emph{ShareJS} is an \ac{OT} \emph{JavaScript} library, developed by the ex \emph{Google Wave} engineer Joseph Gentle, for collaborative text and \ac{JSON} documents edition in real-time.
+	\emph{ShareJS} is an \ac{OT} \emph{JavaScript} library, developed by the ex \emph{Google Wave} engineer Joseph Gentle, for collaborative text and \ac{JSON} documents edition in real time.
 It uses \emph{ShareDB} \footnote{\url{https://github.com/share/sharedb}(accessed March 10, 2016)} for its backend and data model, which supports simple integration with any database.
@@
 An implication of implementing just the client side is the extra effort that is necessary to implement content's persistent storage.
-Besides this drawback, this library is very flexible because it's not tied to a specific database or server side technology.
+Besides this drawback, this library is very flexible because it is not tied to a specific database or server side technology.
 
-	In summary, table \ref{table:otcomparision} shows a comparision of the \ac{OT} libraries presented so far.
+	In summary, Table \ref{table:otcomparision} shows a comparision of the \ac{OT} libraries presented so far.
 

### 2016-05-22 tese2/12_related_work_summary.tex
[RFT] state of art nearly closed
[FIX] front cover typo

@@
 In order to implement an hyper-linked communication solution, several design decisions had to be made.
-The limitations imposed by the \emph{Internet's} structure, its protocols and a browser's capabilities are key factors to consider when implementing a solution that allows bi-directional communications, interactive media and collaboration environment.
+The limitations imposed by the \emph{Internet's} structure, its protocols and browser's capabilities are key factors to consider when implementing a solution that allows bi-directional communications, interactive media and collaboration environment.
 
@@
 With the communications establishment issue solved, we had to discuss the different types of media and what can be done with each kind in order to increase the value of communications among users.
+In this context, we have studied solutions and libraries that allow us to implement our prototype with time manipulation features, collaborative text edition, record and playback interactive video.
 %RP falas de mais coisas: manipulação de tempo, gravação, edição colaborativa

### 2016-05-22 tese2/13_implementation_model.tex
caneco
[FIX] front cover typo

@@
 Needless to say that, by following this approach we do not need to create the collections with a well predefined schema.
-In fact we do not need to define it at all but, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure as we will present in this section.
+In fact, we do not need to define it at all but, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure as we will present in this section.
     %RP a última frase é confusa. O que queres dizer? Reescreve
@@
 
-    An example of schema representation can be seen on table \ref{table:schema}.
+    An example of schema representation can be seen on Table \ref{table:schema}.
 
@@
 This specification lets us create abstract authentication services.
-For example a user can login into our system by providing any attribute that identifies himself, \emph{e.g.} the e-mail but others are possible like the user name or cellphone number.
+For example a user can login into our system by providing any attribute that identifies himself, \emph{e.g.} the e-mail, but others are possible like the user name or cellphone number.
 
@@
 For example, the user role is an aggregator property that within users it allows the identifaction of each one is administrator.
-This aggregation specification is independent from the object type, so it's possible to search for the administrative role and return users and groups of users that contains that property.
+This aggregation specification is independent from the object type, so it is possible to search for the administrative role and return users and groups of users that contains that property.
 <= [currently this operation is possible but it is not specified]
@@
 
-A conference room, which is the environment where users communicate among their selves, is represented persistently by a group of participating entities (in general users but our system allow other types of entities).
+A conference room, which is the environment where users communicate among themselves, is represented persistently by a group of participating entities (in general users but our system allow other types of entities).
 
-A group is composed by an \emph{id}, \emph{inviteToken} and a \emph{visibility} as shown on table \ref{table:group}.
+A group is composed by an \emph{id}, \emph{inviteToken} and a \emph{visibility} as shown on Table \ref{table:group}.
 Moreover, a group can be public or private.
@@
 During a group conversation, it is possible to create time annotations for making it easy to access that time either for searching or sharing with other users.
-A time annotation (table \ref{table:hyper}) contains a title, the correspondent group identification number and the time itself.
+A time annotation (Table \ref{table:hyper}) contains a title, the correspondent group identification number and the time itself.
 
@@
 In the point o view of a user, a time annotation is just a simple way to associate a time stamp to a topic in order to make it easier to find.
-On the other hand, the hyper-content is more complete than time annotations, except they are not visible on the time-line.
+On the other hand, the hyper-content is more complete than time annotations, except they are not visible on the timeline.
 In addition to time annotations, hyper-contents have a duration and content that can be superimposed to video.
@@
 
-For example the content showed in listing \ref{lst:jsoup} would produce the searchable content "Click here!".
+For example the content showed in Listing \ref{lst:jsoup} would produce the searchable content "Click here!".
 

### 2016-05-22 tese2/14_implementation_signaling.tex
[FIX] front cover typo

@@
 In a general view, the signaling protocol is used to share connection and media properties between two peers.
-In order to ease the understanding of our signaling protocol, we have created a sequence diagram that represented on figure \ref{fig:signaling2}\footnote{Although two \ac{ICE} servers are shown they are, in fact, the same.
+In order to ease the understanding of our signaling protocol, we have created a sequence diagram that is represented on Figure \ref{fig:signaling2}\footnote{Although two \ac{ICE} servers are shown they are, in fact, the same.
 Showing just one \ac{ICE} server would be difficult to draw.}.
@@
 %HR done
-Our signaling protocol consists of sending and receiving \ac{JSON} formated messages, \emph{e.g.} listing \ref{lst:msgtype}, over WebSockets by both the application server and the client.
+Our signaling protocol consists of sending and receiving \ac{JSON} formated messages, \emph{e.g.} Listing \ref{lst:msgtype}, over WebSockets by both the application server and the client.
 
@@
 If the user decides to share either from camera or screen, \emph{getUserMedia} is called with the correspondent constraints in order to obtain a local stream.
-We use the constraints presented on listing \ref{lst:constraints}.
+We use the constraints presented on Listing \ref{lst:constraints}.
 
@@
 If the user decides to share his camera or screen, a pop-up is raised in order to ask the user to give permission to share those resources.
-If the resources are shared successfully, the user agent creates an offer like listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the WebSocket to the Application Server.
+If the resources are shared successfully, the user agent creates an offer like Listing \ref{lst:sig01}, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the WebSocket to the Application Server.
 %RP usas ``user'' duas vezes, uma para te referires ao utilizador e outra ao browser (user agent).
@@
 The server receives and processes the offer and sets the \emph{remote session description} to its client associated \ac{WebRTC} endpoint.
-Then a \emph{local session description}, like the one presented in Listing \ref{lst:sig02}, is created on the server and sent back to the client, after that the server tries to gather \ac{ICE} candidates.
+Then a \emph{local session description}, like the one presented in Listing \ref{lst:sig02}, is created on the server and sent back to the client.
+After that, the server tries to gather \ac{ICE} candidates.
 %RP listing -> Listing
@@
 The candidates are received at the client which sets them to its \emph{PeerConnection}.
-The same is done on the server which receives the \ac{ICE} candidates like on listing \ref{lst:sig03} from the client and propagates them to \ac{KMS}.
+The same is done on the server which receives the \ac{ICE} candidates like on Listing \ref{lst:sig03} from the client and propagates them to \ac{KMS}.
 %RP A figura 4.1 precisa de vir antes deste texto todo, para ser mais fácil de perceber.

### 2016-05-22 tese2/15_implementation_recording.tex
[FIX] front cover typo

@@
 
-	With server side recording, the user would maintain always the same stream \ac{URL} even if it is playing real-time video or reproducing recorded video.
+	With server side recording, the user would maintain always the same stream \ac{URL} even if it is playing real time video or reproducing recorded video.
 It is \ac{KMS} that sends different content through that stream.
@@
 
-	If we choose the closest chunk to play this lead to a playing time error that can be at most half the duration of a block, as it can be seen on figure \ref{fig:nearchunkerror} where the black dot represents the requested time to play.
+	If we choose the closest chunk to play this lead to a playing time error that can be at most half the duration of a block, as it can be seen on Figure \ref{fig:nearchunkerror} where the black dot represents the requested time to play.
 In this case, which is the worst, either $ block_{n}$ or $ block_{n+1}$ (with duration $d$) can be played from their beginning with a maximum error ($e$) correspondent to half their duration ($\frac{d}{2}$) .
 
-	In practice, the maximum playing time error coincides with the block duration because we must show always what the user requests, as it can be seen on figure \ref{fig:interchunkerror}.
+	In practice, the maximum playing time error coincides with the block duration because we must show always what the user requests, as it can be seen on Figure \ref{fig:interchunkerror}.
 In this case, if a user wishes to play a time that is close to the end of $block_n$ the maximum playing error ($e$) is at most the chunk's duration ($d$) because we have to play $block_n$ from its beginning.

### 2016-05-22 tese2/16_implementation_content.tex
caneco
[FIX] front cover typo

@@
 	In order to create content, the user has the option to write simple movie captions without writing any code, otherwise, as mentioned before, it can write \ac{HTML}, \ac{CSS} and \emph{JavaScript}.
-The definition of the content's starting and ending time by the user it's not an easy task.
+The definition of the content's starting and ending time by the user it is not an easy task.
 	Figure \ref{fig:creation} shows the user interface for creating interactive hyper content manually.
@@
 
-	Defining the content's time to appear in real-time would require a previous user plan. Otherwise the user could make a speech and insert the content later.
+	Defining the content's time to appear in real time would require a previous user plan. Otherwise the user could make a speech and insert the content later.
 
-	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real-time, we allow the user to encode its content into \ac{QR} codes and show it to the camera in real time.
+	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real time, we allow the user to encode its content into \ac{QR} codes and show it to the camera in real time.
 
@@
 In order to apply our algorithm more efficiently we calculate the content's hash through the \emph{md5} method.
-If the hash is not present in the map, it means the \ac{QR} code was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real-time.
+If the hash is not present in the map, it means the \ac{QR} code was entering the screen, we add that hash to the map and associate the current time to it, all the users are notified to watch that content in real time.
 If the same \ac{QR} code is not detected after a time bigger then two periods we can conclude that the \ac{QR} code left the screen and we add the correspondent content into our database.
@@
 Each layer is associated to the content.	We have taken into account that the amount of content tends to grow with time and the user should only have access to a subset of the content instead of all of it, which would be very inefficient.
-The content to display for each user depends on the user's position on the time-line.
+The content to display for each user depends on the user's position on the timeline.
 This position is given by an offset between the current time and the navigated time.
-If the user is watching the content in real-time this offset is always zero.
+If the user is watching the content in real time this offset is always zero.
 If the user is reproducing past communications in normal playback rate, this offset is constant and, as a consequence, there is need to re-synchronize this value.
@@
 
-	We have used \emph{vis.js} to display out time-line.
+	We have used \emph{vis.js} to display out timeline.
 This library was created for content navigation through time, but it was not designed to be always moving automatically.
@@
 
-	Figure \ref{fig:timeline} shows the graphical appearance of our interactive time-line.
-In order to navigate through time, the user must drag and drop the time-line horizontally.
-When the user drops the time-line an event handler is called with the new user's time offset, which will be used to send a message to the server in order to choose the correspondent content to display.
+	Figure \ref{fig:timeline} shows the graphical appearance of our interactive timeline.
+In order to navigate through time, the user must drag and drop the timeline horizontally.
+When the user drops the timeline an event handler is called with the new user's time offset, which will be used to send a message to the server in order to choose the correspondent content to display.
 
@@
 		\includegraphics[width=0.95\textwidth]{figures/timeline.png}
-		\caption{Interactive time-line}
+		\caption{Interactive timeline}
 		\label{fig:timeline}
@@
 To solve this problem, the server sends its time to the client immediately after the \emph{webSocket} creation.
-We synchronize the time-line with the server and although it may exist a small error due to network transmission time, the graphical error is negligible.
+We synchronize the timeline with the server and although it may exist a small error due to network transmission time, the graphical error is negligible.
 
@@
 	Time annotations are a simpler way to mark points in time and share them with other users.
-When a user enters a conference room, the server sends all annotations so they can be displayed directly into the user time-line.
+When a user enters a conference room, the server sends all annotations so they can be displayed directly into the user timeline.
 When each time annotation is created, all users are notified so they can update their interfaces.
 
-	Figure \ref{fig:annotation} shows the creation and placement of annotations on the time-line.
+	Figure \ref{fig:annotation} shows the creation and placement of annotations on the timeline.
 To save the annotation the user must click on the floppy icon in order to save it and notify the conference participants.
@@
 
-		The composite operation, as shown on figure \ref{fig:wcomposite}, is performed at \ac{KMS} and works by decoding the incoming client streams, mixing audio and video and encoding again to send back to users.
+		The composite operation, as shown on Figure \ref{fig:wcomposite}, is performed at \ac{KMS} and works by decoding the incoming client streams, mixing audio and video and encoding again to send back to users.
 
@@
 
-		We provide a way for users to select the composite view of the conference room or the video shared by a particular user device as it can be seen on figure \ref{fig:devices}.
+		We provide a way for users to select the composite view of the conference room or the video shared by a particular user device as it can be seen on Figure \ref{fig:devices}.
 

### 2016-05-22 tese2/17_implementation_deploy.tex
[FIX] front cover typo

@@
 
-On table \ref{table:osspecs} we present an overview of the operating system configurations of the machine we used for deployment.
+On Table \ref{table:osspecs} we present an overview of the operating system configurations of the machine we used for deployment.
 
@@
 
-	We present, on table \ref{table:softspecs}, the software we have installed inside our docker container including the versions that are in use.
+	We present, on Table \ref{table:softspecs}, the software we have installed inside our docker container including the versions that are in use.
 

### 2016-05-22 tese2/architecture.tex
until 1h test case
[FIX] front cover typo

@@
 
-Taking into account the goals of this project and all the technology presented so far.
-Our proposal is the development of a web application that provides communication and collaboration features in real-time.
+Taking into account the goals of this project and all the technology presented so far, our proposal is the development of a web application that provides communication and collaboration features in real time.
 
@@
 Private chat rooms will only be visible to a defined list of clients or can be accessed by clients that have a link for that conference room.
- \item \textbf{Stream Recording} - Ability to recording and playback recorded video including all the hyper-content displayed at that time.
+ \item \textbf{Stream Recording} - Ability to record and playback recorded video including all the hyper-content displayed at that time.
  \item \textbf{Stream Composition} - Our streaming server must support mixing multiple user streams into a single stream.
 The sound played on clients must be the composed sound of all participants but the video played can be switched to the individual or group view.
- \item \textbf{Time annotations} - Our application must give the possibility to create annotations associated to a specified time and support viewing them on a time-line. 
+ \item \textbf{Time annotations} - Our application must give the possibility to create annotations associated to a specified time and support viewing them on a timeline. 
  \item \textbf{Content Overlay} - Our system must allow users to superimpose hyper-content to video given a range of time.
@@
  \item \textbf{Collaboration} - The web application must provide a collaborative text edition tool.
- \item \textbf{QR Code detection} - The system must interpret \emph{QR codes} in order to ease content creation.
+ \item \textbf{QR Code detection} - The system must interpret \ac{QR} codes in order to ease content creation.
  \item \textbf{NAT Traversal} - The system's services must be available to users that are inside a private network.
@@
 
- \item \textbf{NAT Traversal module} - \ac{STUN} and \ac{TURN} techniques used by \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase which ends by establishing the connection between them.
+ \item \textbf{NAT Traversal module} - \ac{STUN} and \ac{TURN} are techniques used by \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase which ends by establishing the connection between them.
 
@@
 
-In order to simplify our solution, we propose that the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules be implemented within the same server application, so it could be easy to deploy as a single image.
+In order to simplify our solution, we propose that the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules be implemented within the same server application, so it could be easy to deploy as a single virtual machine image.
 To this set of modules, we often call \emph{backend}.
@@
 	In this context, each client must only have one \emph{PeerConnection} to our streaming server and content shown to them is changed on demand either being it an individual or a composite view.
-This approach is centralized, as seen on figure \ref{fig:central}.
-Otherwise clients could follow a peer-to-peer connection, as seen on figure \ref{fig:p2p}, which would result on maintaining more connections and performing the composition of videos on client side.
+This approach is centralized, as seen on Figure \ref{fig:central}.
+Otherwise clients could follow a peer-to-peer connection, as seen on Figure \ref{fig:p2p}, which would result on maintaining more connections and performing the composition of videos on client side.
 
@@
 
-In this chapter we have shown the modules needed in order to implement our solution.
+In this chapter we have described the modules needed in order to implement our solution.
 
-Several technologies were taken into account for implementing each module, we have studied the pros and cons of each technology and decided the architecture for our solution.
+Several technologies were taken into account for implementing each module. We have studied the pros and cons of each technology and decided the architecture for our solution.

### 2016-05-22 tese2/evaluation.tex
[FIX] front cover typo

@@
 
-With this results we conclude that if we want to scale our storage solution using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \ac{KMS} should be installed on the same machine, as depicted on figure \ref{fig:scale2}, because the loopback interface can handle bigger transfer rates than the remaining network interfaces.
+With this results we conclude that if we want to scale our storage solution using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \ac{KMS} should be installed on the same machine, as depicted on Figure \ref{fig:scale2}, because the loopback interface can handle bigger transfer rates than the remaining network interfaces.
 Installing the repository on the same machine as a database node does not ensure that recorded videos are stored in the same machine, for this reason we would prefer installing \ac{KMS} and \emph{Kurento Repository} on the same machine.
@@
 
-Figure \ref{fig:test_full_features_cpu_zoom} is the same as figure \ref{fig:test_full_features_cpu} but zoomed over \emph{MongoDB}, \emph{Java} and \emph{Kurento Repository}.
+Figure \ref{fig:test_full_features_cpu_zoom} is the same as Figure \ref{fig:test_full_features_cpu} but zoomed over \emph{MongoDB}, \emph{Java} and \emph{Kurento Repository}.
 As we can see there are periodic processing peaks every ten seconds that are due to block recording.
@@
 
-We have also performed two consecutive test cases in order to understand if we the influence of the first test case over the second, as observed in figure \ref{fig:test_two_times_cpu}, there is no \ac{CPU} influence between two test cases.
+We have also performed two consecutive test cases in order to understand if we the influence of the first test case over the second, as observed in Figure \ref{fig:test_two_times_cpu}, there is no \ac{CPU} influence between two test cases.
 
@@
 
-On the other hand due to memory recycling techniques we can observe in figure \ref{fig:test_two_times_mem} that some of the memory that was allocated in the first test was reused in the second case.
+On the other hand due to memory recycling techniques we can observe in Figure \ref{fig:test_two_times_mem} that some of the memory that was allocated in the first test was reused in the second case.
 
@@
 The initial users that tested our solution had difficulties in the first task when finding new friendship requests and adding new friends, as friendship request were only possible to access from the top bar and new user profiles were only accessible through the search bar.
-In order to solve this problem we have added a menu for adding new users in the friends list and another near the friends list as it can be seen on figure \ref{fig:test_ui_01_02_03}.
+In order to solve this problem we have added a menu for adding new users in the friends list and another near the friends list as it can be seen on Figure \ref{fig:test_ui_01_02_03}.
 
@@
 On the second task one user clicked on the friends list for adding a friend to a room and, as a consequence, left the conference room to the user profile.
-We solved this problem by showing a pop-up with the user profile and an additional button to add the user to the conference room as it can be seen on figure \ref{fig:test_ui_04}.
+We solved this problem by showing a pop-up with the user profile and an additional button to add the user to the conference room as it can be seen on Figure \ref{fig:test_ui_04}.
 
@@
 
-On the third task we noticed that users were not expecting to search the content directly from the search bar, instead they used the time-line.
+On the third task we noticed that users were not expecting to search the content directly from the search bar, instead they used the timeline.
 
-On the fourth task the users did not placed their annotations on right place. We only provide one way to place annotations, which is by dragging them on the time-line. 
-After creating an annotation, users also did not understand that they had to save their changes on the time-line.
+On the fourth task the users did not placed their annotations on right place. We only provide one way to place annotations, which is by dragging them on the timeline. 
+After creating an annotation, users also did not understand that they had to save their changes on the timeline.
 We solved this problem by saving automatically each time the user changed any annotation.
-The old user interface can be seen on figure \ref{fig:test_ui_05}.
+The old user interface can be seen on Figure \ref{fig:test_ui_05}.
 
@@
 \item Giving feedback to user when performing operations such as create superimposed content and save the collaborative editor content.
-\item Starting with our time-line more zoomed in order to give more sensation of time passing.
+\item Starting with our timeline more zoomed in order to give more sensation of time passing.
 \item Create content with starting time and duration, instead of ending time.
@@
 
-To this end, with regard to optimal task duration and minimal clicks, we obtained the values shown in table \ref{table:optimal}.
+To this end, with regard to optimal task duration and minimal clicks, we obtained the values shown in Table \ref{table:optimal}.
 
@@
 Another weak aspect was our content editor, which, in fact, we recognize is difficult to work with, mostly due to the amount of information that is necessary to create a synchronized content (starting time, duration and content itself).
-Some users have suggested that the content should also be present on the time-line so they could be easily dragged and resized (on time).
+Some users have suggested that the content should also be present on the timeline so they could be easily dragged and resized (on time).
 
-We are aware that placing content on the time-line will reduce our solutions performance, especially when there is a relatively large amount of content, due to the content that is present on the time-line being loaded all at once.
-Although, we recognize that for some cases (relatively low amount of content), displaying the content on the time-line could not have a great impact on our solutions performance, we choose not to implement this.
+We are aware that placing content on the timeline will reduce our solutions performance, especially when there is a relatively large amount of content, due to the content that is present on the timeline being loaded all at once.
+Although, we recognize that for some cases (relatively low amount of content), displaying the content on the timeline could not have a great impact on our solutions performance, we choose not to implement this.
 

### 2016-05-22 tese2/implementation.tex
[FIX] front cover typo

@@
 
-Then are going to describe how we show hyper-content to users, namely the multiple ways to create hyper-content and the algorithm behind the user interface synchronization, the security flaws of our content displaying mechanism and a possible solution to overcome vulnerabilities.
+Afterwards we are going to describe how we show hyper-content to users, namely the multiple ways to create hyper-content and the algorithm behind the user interface synchronization, the security flaws of our content displaying mechanism and a possible solution to overcome vulnerabilities.
 
-Not less important we also are going to describe how we have implemented our application time-line and the functionalities above it such as time manipulation and annotations management.
+Not less important we also are going to describe how we have implemented our application timeline and the functionalities above it such as time manipulation and annotations management.
 Among other different functionalities we also are going to describe how we perform stream composition, synchronization of our collaborative text editor and implementation of our instant messenger.

### 2016-05-22 tese2/introduction.tex
[FIX] front cover typo

@@
 	With the advent of the telephone network, communications have taken a very important step for us to feel more connected with whom we communicate.
-Still, only the human voice was not enough, and the invention of cameras and consequent video digitalization were a huge step for real-time communications.
+Still, only the human voice was not enough, and the invention of cameras and consequent video digitalization were a huge step for real time communications.
 
@@
 
-    Furthermore, real-time communication applications can make a significant difference on business, education and health sectors by providing tools for developing teaching and learning online, teamworking and socializing web applications.
+    Furthermore, real time communication applications can make a significant difference on business, education and health sectors by providing tools for developing teaching and learning online, teamworking and socializing web applications.
 
@@
 
-	Therefore, a real-time system is a huge source of information that requires much attention from its users.
+	Therefore, a real time system is a huge source of information that requires much attention from its users.
 In this context, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.

### 2016-05-22 tese2/xthesis.tex
caneco
[FIX] front cover typo

@@
 
-\emph{WebRTC} technology allows real-time communications between web browsers without the need to install additional software.
+\emph{WebRTC} technology allows real time communications between web browsers without the need to install additional software.
 The nature of web browser applications already follows the hypermedia concept, which makes \emph{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
@@
 	\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations over plain text on the client side and requires implementing the content's persistent storage.
-Besides this drawback, this library is very flexible because it's not tied to a specific database or server side technology.
+Besides this drawback, this library is very flexible because it is not tied to a specific database or server side technology.
 
@@
 Taking into account the goals of this project and all the technology presented so far.
-Our proposal is the development of a web application that provides communication and collaboration features in real-time.
+Our proposal is the development of a web application that provides communication and collaboration features in real time.
 
@@
 The server receives and processes the offer and sets the \emph{remote session description} to its client associated \gls{WebRTC} endpoint.
-Then a \emph{local session description} is created on the server and sent back to the client, after that the server tries to gather \gls{ICE} candidates.
+Then a \emph{local session description} is created on the server and sent back to the client.
+After that, the server tries to gather \gls{ICE} candidates.
 
@@
 
-	With server side recording, the user would maintain always the same stream \gls{URL} even if it is playing real-time video or reproducing recorded video.
+	With server side recording, the user would maintain always the same stream \gls{URL} even if it is playing real time video or reproducing recorded video.
 It is \gls{KMS} that sends different content through that stream.
@@
 
-	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real-time, we allow the user to encode its content into \gls{QR} codes and show it to the camera in real time.
+	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real time, we allow the user to encode its content into \gls{QR} codes and show it to the camera in real time.
 

### 2016-05-22 presentation2/video.sh
[ADD] starting final presentation

@@ -1,0 +1,2 @@
+#!/usr/bash
+mplayer -fs video.mp4

### 2016-05-24 project/httpserver/app/main/Room.java
[ADD] presentation implementation, minor bug fix

@@ -39,7 +39,6 @@ public class Room implements Closeable {
     
     /** The composite. */
     private Hub composite = null;
-    private long sequence = 0;
     /** The hub port. */
     private HubPort compositePort;
     private Recorder recorder;
@@ -139,7 +138,6 @@ public class Room implements Closeable {
                         Date end = new Date();
 
                         RecordingChunk rec = new RecordingChunk(gid,gid,start,end,"group",filepath);
-
                         rec.save();
 
                         interval.setEnd(end);
@@ -154,7 +152,7 @@ public class Room implements Closeable {
                         msg.put(interval.getId().toString(), array);
                         sendMessage(msg.toString());
 
-                        System.out.println("GRP REC: " + filepath);
+                        System.out.println("REC: group|"+filepath);
                         record(duration,true);
                     }
                 });

### 2016-05-25 tese2/architecture.tex
[RFT] removed extra slides

@@
 	\label{table:apparch}
-    \begin{tabular}{cccccccc@{}m{0pt}@{}}
+    \begin{tabular}{ccccccccc@{}m{0pt}@{}}
 	\hline 
-\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
-\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{~~~~~ot.js~~~~~} & \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} & \multicolumn{1}{c|}{adapter.js} &   \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} &\\[12pt]\hline
-\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
+\multicolumn{9}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-6}\cline{8-8}
+\multicolumn{2}{|c|}{~~~~~jQuery~~~~~} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{~~~~~ot.js~~~~~} & \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} & \multicolumn{1}{c|}{adapter.js} &   \multicolumn{1}{c|}{\cellcolor{Gray}~~~~~~~~~~~~~~~} &\\[12pt]\hline
+\multicolumn{1}{|c|}{HTTP} & \multicolumn{3}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
 \end{tabular}

### 2016-05-31 presentation2/present.sh
BU

@@ -1,0 +1,2 @@
+pdflatex project.tex
+pdfpc  project.pdf -n right

### 2016-05-31 project/httpserver/app/controllers/Application.java

@@ -49,7 +49,7 @@ public class Application extends Controller {
         System.out.println("RESET!");
         Service.reset();
 
-        User user1, user2, user3, user4 , user5,td1,td2;
+        User user1, user2, user3, user4 , user5,user6,user7;
         {
             List<AttributeDto> attributes = new ArrayList<AttributeDto>();
             attributes.add(new AttributeDto("email", "test", true, true, false,null,null));
@@ -90,26 +90,28 @@ public class Application extends Controller {
             attributes.add(new AttributeDto("photo", "/assets/images/me.png", false, false, false,null,null));
             user4 = new RegisterUserService("qazokm", attributes).execute();
         }
-/*
+
         {
             List<AttributeDto> attributes = new ArrayList<AttributeDto>();
-            attributes.add(new AttributeDto("email", "talkdesk1", true, true, false));
-            attributes.add(new AttributeDto("name", "Talkdesk 1", false, true, false));
-            attributes.add(new AttributeDto("photo", "/assets/images/td1.png", false, false, false));
-            td1 = new RegisterUserService("talkdesk",new ArrayList<PermissionDto>() ,attributes).execute();
+            attributes.add(new AttributeDto("email", "monster", true, true, false,null,null));
+            attributes.add(new AttributeDto("name", "Cookie Monster", false, true, false,null,null));
+            attributes.add(new AttributeDto("photo", "/assets/images/monster.jpeg", false, false, false,null,null));
+            user6 = new RegisterUserService("qazokm",attributes).execute();
         }
 
         {
             List<AttributeDto> attributes = new ArrayList<AttributeDto>();
-            attributes.add(new AttributeDto("email", "talkdesk2", true, true, false));
-            attributes.add(new AttributeDto("name", "Talkdesk 2", false, true, false));
-            attributes.add(new AttributeDto("photo", "/assets/images/td2.png", false, false, false));
-            td2 = new RegisterUserService("talkdesk",new ArrayList<PermissionDto>() ,attributes).execute();
+            attributes.add(new AttributeDto("email", "kermit", true, true, false,null,null));
+            attributes.add(new AttributeDto("name", "Kermit the Frog", false, true, false,null,null));
+            attributes.add(new AttributeDto("photo", "/assets/images/kermit.jpg", false, false, false,null,null));
+            user7 = new RegisterUserService("qazokm",attributes).execute();
         }
 
[…]

### 2016-06-07 curriculo/cv.tex
[CLS] removed unused files

@@
    \entry
-    {2013–now}
+    {2013–2016}
     {Master of Science}
     {Instituto Superior Técnico - Taguspark}
-    {Telecommunications and Informatics Engineering (Expected Grade: 16 in 20)}
+    {Telecommunications and Informatics Engineering (Final Grade: 16 in 20)}
    \entry

### 2016-06-07 implementation_research/Ejabberd Resource Constraint: Users are not allowed to register accounts so fast | Sid___files/a.js

@@ -1,1 +1,0 @@
-/**/skimlinksApplyHandlers({"country":"PT","sessid":"bd10f2249861afee148632498892277f","guid":"492594c6b114f0cdf04c0b628625ca94","merchant_domains":["vimeo.com"]});

### 2016-06-07 implementation_research/Ejabberd Resource Constraint: Users are not allowed to register accounts so fast | Sid___files/a_002.js

@@ -1,1 +1,0 @@
-/**/skimlinksApplyHandlers({"country":"PT","sessid":"bd10f2249861afee148632498892277f","guid":"492594c6b114f0cdf04c0b628625ca94","merchant_domains":[]});

### 2016-06-07 implementation_research/xmpp - Ejabberd and in-band registration - Stack Overflow_files/ados.js

@@ -1,34 +1,0 @@
-var html__72501 = '\x3ca\x20href\x3d\x22http\x3a\x2f\x2fengine\x2eadzerk\x2enet\x2fr\x3fe\x3deyJhdiI6NDE0LCJhdCI6NCwiYnQiOjAsImNtIjo5MTgyOCwiY2giOjExNzgsImNyIjo3MzgwMjksImRpIjoiODkwYTgzZDMzOWNhNGQ5ODkxMTBlOWIyYWY3MWRkZmMiLCJkbSI6MSwiZmMiOjc2OTA3NywiZmwiOjE2NTI5MywiaXAiOiI4NS4xMzkuNS4yNDAiLCJrdyI6InhtcHAsZWphYmJlcmQiLCJudyI6MjIsInBjIjowLCJwciI6NzY1LCJydCI6MSwicmYiOiJodHRwOi8vd3d3Lmdvb2dsZS5wdC91cmw_c2E9dCZyY3Q9aiZxPSZlc3JjPXMmc291cmNlPXdlYiZjZD0xJnZlZD0wQ0NBUUZqQUEmdXJsPWh0dHA6Ly9zdGFja292ZXJmbG93LmNvbS9xdWVzdGlvbnMvNTczMjE3My9lamFiYmVyZC1hbmQtaW4tYmFuZC1yZWdpc3RyYXRpb24mZWk9RTdxT1ZjaXZOb2E3VWJlZXI5QUcmdXNnPUFGUWpDTkdyekROYWJhMV81ZnNPdzRXR2M1OWdyV0FNdlEmc2lnMj1IaVF2aVhBNkExTW1ZbW1ZZFp3TnRBJmJ2bT1idi45Njc4MzQwNSxkLmQyNCIsInN0Ijo4Mjc3LCJ1ayI6InVlMS1jM2Y1ZmIxMzY0ZTE0MzhkYWM5OTk4ZWQxYzJiMGU3MSIsInpuIjo0MywidHMiOjE0MzU0MTcxMTgwMDcsImJmIjp0cnVlLCJwbiI6ImFkemVyazIxMjY4Mjk1MjkiLCJ1ciI6Imh0dHA6Ly9wdC5zdGFja292ZXJmbG93LmNvbS8_dXRtX3NvdXJjZT1zdGFja292ZXJmbG93LmNvbSZ1dG1fbWVkaXVtPWFkJnV0bV9jYW1wYWlnbj1wdH
[…]

### 2016-06-07 implementation_research/xmpp - Ejabberd and in-band registration - Stack Overflow_files/post-validation.js

@@ -1,1 +1,0 @@
-StackExchange.postValidation=function(){function e(e,t,n,i){var r=e.find('input[type="submit"]:visible'),a=r.length&&r.is(":enabled");a&&r.attr("disabled",!0),o(e,i),s(e,t,n,i),c(e),u(e),d(e);var f=function(){1!=t||e.find(C).length?(l(e,i),a&&r.attr("disabled",!1)):setTimeout(f,250)};f()}function t(t,i,o,s,l){e(t,i,s,o);var c,u=function(e){if(e.success)if(l)l(e);else{var n=window.location.href.split("#")[0],r=e.redirectTo.split("#")[0];0==r.indexOf("/")&&(r=window.location.protocol+"//"+window.location.hostname+r),c=!0,window.location=e.redirectTo,n.toLowerCase()==r.toLowerCase()&&window.location.reload(!0)}else e.captchaHtml?e.nocaptcha?StackExchange.nocaptcha.init(e.captchaHtml,u):StackExchange.captcha.init(e.captchaHtml,u):e.errors?(t.find("input[name=priorAttemptCount]").val(function(e,t){return(+t+1||0).toString()}),p(e.errors,t,i,o,e.warnings)):t.find('input[type="submit"]:visible').parent().showErrorMessage(e.message)};t.submit(function(){if(t.find("#answer-from-ask").is(":checked"))return!0;var e=t.find(E);if("[Edit removed during grace period]"==$.trim(e.val()))return g(e,["Comment reserved for system use. Please use an appropriate comment."],f()),!1;a(),StackExchange.navPrevention&&StackExchange.navPrevention.stop();var i=t.find('input[type="submit"]:visible');if(i.parent().addSpinner(),StackExchange.helpers.disableSubmitButton(t),StackExchange.options.site.enableNewTagCreationWarning){var s=t.find(C).parent().find("input#tagnames"),l=s.prop("defaultValue");if(s.val()!==l)return $.ajax({"type":"GET","url":"/posts/new-tags-warning","dataType":"json","data":{"tags":s.val()},"success":function(e){e.showWarning?i.loadPopup({"html":e.html,"dontShow":!0,"prepend":!0,"loaded":function(e){n(e,t,c,o,u)}}):r(t,o,c,u)}}),!1}return setTimeout(function(){r(t,o,c,u)},0),!1})}function n(e,t,n,a,o){e.bind("popupClose",function(){i(t,n)}),e.find(".submit-post").
[…]

### 2016-06-09 curriculo/cv.tex
[RW] until qr code detection (reversed)

@@
 \entry
-    {08 2015\\(standby)}
+    {08 2015\\(2 months)}
     {Systems Development}
     {Bullray-CIT}
-    {\emph{Development and configuration of a streaming server that supports receiving and recording video from authenticated users through \underline{ffmpeg}.
-I tried \underline{FFServer}, \underline{nginx with rtmp module} and \underline{Wowza Streaming Server}, the final decision was \underline{Kurento Media Server} with an \underline{RTP} endpoint for receiving video from network cameras and a \underline{WebRTC} endpoint for delivering the video to web browsers.
+    {\emph{Development and configuration of a streaming server that supports receiving and recording video from authenticated users.
+I tried \underline{FFServer}, \underline{nginx with rtmp module} and \underline{Wowza Streaming Server}, the final decision was \underline{Kurento Media Server} with \underline{WebRTC}.
 }}
@@
 \entry
-    {03 2015\\(ongoing)}
-    {Master Thesis}
+    {03 2015\\(1 year)}
+    {Master Thesis \\\href{https://citysdk.tagus.ist.utl.pt:9443/}{try it}}
     {IST}
@@
 \section{interests}
-    3D Studio Max, Photoshop, Game Development, Web Design, Network Security, Jogging, Karate and Yoga.
+    3D Studio Max, Photoshop, Game Development, Web Design, Network Security, Jogging, Kayaking, Karate and Yoga.
 

### 2016-06-09 tese2/16_implementation_content.tex
[RW] until qr code detection (reversed)

@@
 
-	\ac{KMS} lets registering event handlers for \ac{QR} code detection.
+	\ac{KMS} allows registering event handlers for \ac{QR} code detection.
 The component that detects \ac{QR} codes on \ac{KMS} is called periodically and fires the handlers with the decoded content.
+Although the \ac{QR} code detection period is defined by \ac{KMS} we observed to be one second.
 This mechanism does not detect if the \ac{QR} code enters or leaves the screen.
 We had to implement our own mechanism for detecting those events.
-Each user session in the server maintains a map with the contents that are present on the screen.
+Each user session in the server maintains a map with the contents that are present on the screen, which can be represented by multiple \ac{QR} codes simultaneously.
 In order to apply our algorithm more efficiently we calculate the content's hash through the \emph{md5} method.
@@
 
-	In order to allow users to visualize past and present communications, we express the user's position as the time offset between its navigated time and the current time.
+	In order to allow users to visualize past and present communications, we express the user's time position as the time offset between its reproducing time and the current time.
 
@@
 The content to display for each user depends on the user's position on the timeline.
-This position is given by an offset between the current time and the navigated time.
+This position is given by an offset between the current time and the reproducing time.
 If the user is watching the content in real time this offset is always zero.
@@
 
-	Each content is divided into two components, the \emph{start} and the \emph{end} which are sent to users.
+	Each content, that users receive from our application server, is divided into two components, the \emph{start} and the \emph{end}.
 The \emph{start} component contains a time stamp, the content identification number and the content itself in form of text.
@@
 		\item Content that should be currently on display.
-		\item A subset of contents whose starting time is immediately following the user's time.
+		\item A subset of contents whose starting time is immediately after the user's time.
 	\end{itemize}
@@
 
-	Our solution is flexible on what kind of interactions are possible to the users in real time, but allowing users to write \emph{JavaScript} that is executed on the other users browser would attackers to misuse their resources and access critical information.
+	Our solution is flexible on what kind of interactions are possible to the users in real time, but allowing users to write \emph{JavaScript} that is executed on the other users' browser would allow attackers to misuse their resources and access critical information.
 We could solve this problem easily by escaping any \emph{script} tag present on the content, but we would sacrifice the kind of interactions that are possible.
 By not allowing \emph{JavaScript} we would need to implement a subset of conceivable actions a priori and fire them when a type of message is received.
-We decided to ignore the security vulnerabilities that are exposed by evaluating \emph{JavaScript} because, in order to offer the same interactions, we would have to do an exhaustive functional requirements gathering and, in our prototype, our goal is to explore users behavior when interacting with our system.
+We decided to postpone, to future work, the correction of the security vulnerabilities that are exposed by evaluating \emph{JavaScript} because, in order to offer the same interactions, we would have to do an exhaustive functional requirements gathering and, in our prototype, our goal is to explore users behavior when interacting with our system.
 
@@
 		In order to show all streams to all users, each conference participant could receive individual streams for each user that share video.
-In overall, if there were $n$ active participants in the conference room, the number of streams per user would be $n$ ($1$ for sending video and $n-1$ for receiving) and, consequently, the number of streams in that conference room, as observed in Figure \ref{fig:xcomposite}, would be $n^2$.
+If there are $n$ active participants in the conference room, the number of streams per user is $n$ ($1$ for sending video and $n-1$ for receiving) and, consequently, the number of streams in that conference room, as observed in Figure \ref{fig:xcomposite}, is $n^2$.
 
@@
 
-		By mixing all streams into a single one, users can receive a single stream with the representation of all active users.
+		By mixing all streams into a single one, users can receive a single stream with the image (in a grid) and audio of all active users.
 With this approach, the total number of streams within a conference room can be, at most, $2n$ (each user has $2$ streams, one to send and another to receive) as observed in Figure \ref{fig:ecomposite}.

### 2016-06-09 tese2/evaluation.tex
[RW] until qr code detection (reversed)

@@
 
-  We have tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.
+  We have tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's user interface usability.
 
@@
 
-      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects with a periodicity of one second: CPU, physical memory information relative to each running process and network usage relative to each interface.
+      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects to a text file with a periodicity of one second: CPU, RAM usage relative to each running process and network usage relative to each interface.
 
@@
 
-      The performance test scenario that we have defined consists on two phases, the first phase consists only on having users, with similar computer and network specifications, entering sequentially on the conference room.
+      The performance test scenario that we have defined consists of two phases, the first phase consists only on having users, with similar computer and network specifications, entering sequentially on the conference room.
 The second phase consists on the users leaving the conference room.
-Each event, joining and leaving, occurs with intervals of one minute in total of thirteen minutes (780 seconds).
+Each event, joining and leaving, occurs with intervals of one minute in total of 13 minutes (780 seconds).
 
@@
 Due to the data being transfered also from \emph{Kurento Repo} to \emph{MongoDB} via localhost, the expected total rate of data transfered through localhost is $2((n+1)x) = (2n+2)x$, which explains the reason for the average amount of data transfered through localhost being more than twice the average amount of data received from users.
-
 With this results we conclude that if we want to scale our storage solution using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \ac{KMS} should be installed on the same machine, as depicted on Figure \ref{fig:scale2}, because the loopback interface can handle bigger transfer rates than the remaining network interfaces.
@@
 
-In conclusion, 100\% of our testers though that our solution was an innovation and 95\% recommended using our solution.
+%In conclusion, 100\% of our testers though that our solution was an innovation and 95\% recommended using our solution.
 
@@
 In order to understand how good the results were for each performed task, we have defined optimal values for the different metrics.
-In general the results reveal a significant difference between observed values and optimal values, which can be explained by the user's lack of experience with systems such as the one we have created.
+In general the results reveal a significant difference between observed values and optimal values, which can be explained by the user's lack of experience with with the tools that we provided.

### 2016-06-09 tese2/xthesis.sh

@@ -1,3 +1,4 @@
 pdflatex xthesis.tex 
 bibtex8 --wolfgang xthesis
 pdflatex xthesis.tex 
+pdflatex xthesis.tex 

### 2016-06-17 tese2/14_implementation_signaling.tex
[RFT] until page 26 (reversed)

@@
 \end{figure}
+
+Before the signaling protocol starts, the user must be authenticated.
+After the user provide the correct credentials, it receives an \ac{HTTP} \emph{cookie} from our server in order to be identified in the following \ac{HTTP} requests.
 
@@
 %HR done
-The Web application server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room.
+The web application server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room.
 It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake is performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \ac{HTTP} context is lost.

### 2016-06-17 tese2/15_implementation_recording.tex
[RFT] until page 26 (reversed)

@@
  Initially we experimented with local recording and synchronizing with our servers.
-Even thought it got it to work, it consumed to much bandwidth.
+Even thought it worked, it consumed too much bandwidth.
 We then tried recording on server side.
@@
 \subsection{Client side recording}
-In order to record used shared streams, our first approach consisted in locally recording video and audio streams into blobs of limited duration.
+In order to record user shared streams, our first approach consisted in locally recording video and audio streams into blobs of limited duration.
 Because each user was recording directly from their shared local resources, we achieved the best stream quality.
@@
 
-After recording, each block was uploaded to our server through \ac{HTTP}, saved into the server's file system, the block meta-data was created and inserted on our database.
+After recording, each block was uploaded to our server through \ac{HTTP}, saved into the server's file system, the block meta-data was created and inserted in our database.
 Each block meta-data contained the file's location for the recorded block, the starting date, duration, user identification and group identification.
@@
 When the file was completely saved, the meta-data was advertised to the remaining users.
-This meta-data was simply used to refresh the user interface and was completely discarded after that, because an huge amount of small blocks' meta-data would use more and more memory as time went by.
+This meta-data was simply used to refresh the user interface and was completely discarded after that, because a huge amount of small blocks' meta-data would use more and more memory as time went by.
+
+% HR-FINAL - streaming? gravação de x em x tempo, streaming em tempo real já foi
 
@@
 	For playing video with an higher velocity, we used \emph{ffmpeg}\footnote{\url{https://www.ffmpeg.org/}(accessed: 17 March 2016)} to convert the block into a new video with the desired velocity and seek time.
-Because the media duration is known, when the video started to being converted the headers located at the beginning of the file were already written and that made it possible to stream while converting.
+Because the media duration is known, when the video started being converted the headers located at the beginning of the file were already written and that made it possible to stream while converting.
 	%{\color{red} [TALK ABOUT COMPLEXITY OF DECODE VS DECODE+MANIPULATE+ENCODE]}
@@
 
-	Later the \emph{Kurento} team released a version with support for seeking videos but we had to suspend the implementation of the fast forwarding feature as currently \ac{KMS} is not supporting that.
+	Later the \emph{Kurento} team released a version with support for seeking videos.% but we had to suspend the implementation of the fast forwarding feature as currently \ac{KMS} is not supporting that.
         %RP fiquei sem perceber. O método com o ffmpeg deixou de funcionar? Ou desististe devido ao outros problemas (delay, qualidade, cpu)?
 
-	Although we have not implemented fast forwarding, we could implement it without real time conversion by creating multiple versions of the same video with different playback speeds after the recording of a block.
+	We have not implemented fast forwarding, as currently \ac{KMS} is not supporting that.
+We could implement fast forwarding without real time conversion by creating multiple versions of the same video with different playback speeds after the recording of a block.
 When a user needed to play, he would also need to specify one of the available speed.
@@
 Saving files directly into the file system would require an extra effort to distribute and replicate files among servers.
-For that purpose, the \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB}.
+For that purpose, the \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB} and, as a consequence, improve our solution's scalability by performing replication.
 
 	One of the features that \emph{Kurento Repository} provides is the ability to play directly from the database without having to download the entire file to \ac{KMS}.
-The same is true for recording, but because the file headers are in the beginning and the file is written until is stops, the headers do not contain the necessary information for seeking the file.
+The same is true for recording, but because the file headers are in the beginning and the file is written until it stops, the headers do not contain the necessary information for seeking the file.
 	Although we gain with scalability with this approach, we lose access over the file for changing it to fit our needs, namely for using \emph{ffmpeg} or other video manipulation tool.

### 2016-06-18 tese2/12_related_work_summary.tex
[RFT] last thesis commit?

@@
 In order to implement an hyper-linked communication solution, several design decisions had to be made.
-The limitations imposed by the \emph{Internet's} structure, its protocols and browser's capabilities are key factors to consider when implementing a solution that allows bi-directional communications, interactive media and collaboration environment.
+The limitations imposed by the \emph{Internet's} structure, its protocols and browser's capabilities are key factors to consider when implementing a solution that allows bi-directional communications, interactive media and a collaboration environment.
 

### 2016-06-18 tese2/abstract.tex
[FIX] front cover typo
[RFT] last thesis commit?

@@
 
-Voice and image together can express emotions like no other medium can. With hypermedia concepts, we can add more value to conference calls.
+%Voice and image together can express emotions like no other medium can. With hypermedia concepts, we can add more value to conference calls.
 
-\emph{WebRTC} technology allows real-time communications between web browsers without the need to install additional software.
+\emph{WebRTC} technology allows real time communications between web browsers without the need to install additional software.
 The nature of web browser applications already follows the hypermedia concept, which makes \emph{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
@@
 
-%\sout{Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real-time audio and video communication, enriched with other media types, that can be accessed live or accessed later.}
+%\sout{Our goal in this project is to develop an application that leverages the Hyper-linked communications concept to enable collaborative real time audio and video communication, enriched with other media types, that can be accessed live or accessed later.}
 %This application will target the web browser platform, resorting to \ac{WebRTC}.
 
-Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging, ability to superimpose hyper-content to video and the possibility to playback communications.
+Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a multi-party conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging, ability to superimpose hyper-content to video and the possibility to playback communications.
 

### 2016-06-18 tese2/architecture.tex
[RFT] last thesis commit?

@@
 This content can be continuous or discrete and interactive or non interactive.
- \item \textbf{Search} - Unified search every objects related to a conference room such as hyper-content, time annotations and users.
+ \item \textbf{Search} - Search every objects related to a conference room such as hyper-content, time annotations and users.
  \item \textbf{Content Sharing} - Our application must have the ability to share files and time links among users.
@@
 %RP capacidade de lidar com NAT
+% HR-FINAL - Collaboration VS QRCode, completamente diferentes
+% HR-FINAL - annotation vs content, specified time VS range of time, timeline VS superimposed video content (what else?)
 
@@
 
- \item \textbf{Stream module} - responsible to deliver and receive multimedia content from the \emph{Client} using \ac{WebRTC}.
+ \item \textbf{Stream module} - responsible to receive and deliver multimedia content to the \emph{Client} using \ac{WebRTC}.
 
@@
 \section{Implementation Proposal}
-The infrastructure is composed by: web server, stream server, signaling server, database and video repository.
+The infrastructure is composed by: web server, stream server, database and video repository.
 %RP tens da fazer o paralelo entre isto e os componentes apresentados na secção anterior.
@@
 Podes colocar os elementos na mesma posição, ou dentro das caixas da imagem 3.1
+
+% HR-FINAL - Corrigido
 
@@
 %RP Imagem com firewall e estes portos? Mas isto já me parece muito detalhe de implementação.
-In relation to the database there is no directly access from the outside. All the database information is accessed via our application server which validates the permissions of users on our system.
+In relation to the database there is no direct access from the outside. All the database information is accessed via our application server which validates the permissions of users on our system.
 
 On the other hand, the access to our streaming servers is also restricted, but clients can connect to them after concluding the signaling phase.
-This signaling phase may or not proceed in function of the client access permissions.
+This signaling phase may or not proceed in function of the client's access permissions.
 For example, if a user is trying to access a private conference room that he is not a member of nor has an invitation link for, the signaling server refuses to start the signaling phase and the user cannot access the streaming server.
+
+% HR-FINAL - Bastava o cliente saber o a API do KMS, ligar-se a ele e controla-lo como se fosse administrador (o servidor web faz isso).
+Uma forma de ter o servidor publico e ano ser vulneravel a ameaças passaria por configurar a firewall de forma a bloquear o acesso indevido.
 
@@
 
+% HR-FINAL - KMS? Não só, também as ligações por websocket são centralizadas, por isso é que não especifiquei na figura se era KMS ou WebServer.
+
 	In this context, each client must only have one \emph{PeerConnection} to our streaming server and content shown to them is changed on demand either being it an individual or a composite view.
@@
 
-	The composition of streams on client side is performed by receiving streams with the best quality possible but, due to undersizing the video of clients into a smaller region, this would result on wasting bandwidth on a quality that is not needed.
+% HR-FINAL - Já foi definido. Sim é o que está escrito.
+
+	The composition of streams on client side could be performed by receiving streams with the best quality possible from other users but, due to resizing the video of clients into a smaller region, this would result on wasting bandwidth on a quality that is not needed.
+
+% HR-FINAL - is VS could = could wins. Receive video with the best quality from other users... resizing
 
@@
 	In the context of this thesis, we will not implement sending messages across web servers but, in order to allow our system to scale, we will consider that all conference participants are connected to the same server and our system is scaled by having conference rooms distributed across different servers.
+
+% HR-FINAL - "all conference participants are connected to the same server" sim é mesmo isso, se não fosse teria que enviar as mensagens entre varios servidores (é o que está escrito)
 

### 2016-06-18 tese2/introduction.tex
[RFT] last thesis commit?

@@
 	As communications technologies appeared, we adapted the way we communicate.
-The purpose of this project is not the replacement of the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
+The purpose of this project is not the replacement of the current video and audio communications, but to enrich them with hyper-media content.
+%HR-FINAL - corrigido
 
@@
 
-	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they hear or see.
+%	For multiple reasons, we often need to repeat or postpone some of our tasks, some people tend to forget what they hear or see.
 
-	Therefore, a real time system is a huge source of information that requires much attention from its users.
+	A real time system is a huge source of information that requires much attention from its users.
 In this context, an application that provides a way to remember our past communications would be a strong tool for not only to catch what we lost but also to enhance our knowledge.
@@
 
-	In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links (to navigate to a location in time) and files to conference participants.
+% HR-FINAL - Não mencionei??? ("possibility to playback communications" VS "ability to navigate in time in order to reproduce the conversation ")
+
+	In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that support sending time hyper-links (to navigate to a location in time) and files to conference participants.
 
@@
 Users can change to individual streams on demand or automatically to the talking users.
+
+% HR-FINAL - Coloquei multi-party no abstract, não tenho muita margem para ecrever mais
 
@@
 Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (\emph{e.g.} creating time annotations for improving content search) in order to increase its value.
-It is also important to highlight another goal of this project which is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.
+It is also important to highlight another goal of this project, which is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.
 

### 2016-06-18 tese2/resumo.tex
[RFT] last thesis commit?

@@
 
-A voz e a imagem justas exprimem emoções como nenhum outro tipo de media permite. Através dos conceitos de hiper-media, podemos adicionar mais valor às comunicações por vídeo-conferência.
+%A voz e a imagem justas exprimem emoções como nenhum outro tipo de media permite. Através dos conceitos de hiper-media, podemos adicionar mais valor às comunicações por vídeo-conferência.
 
@@
 
-O nosso objectivo neste projecto é desenvolver de uma aplicação, recorrendo ao \ac{WebRTC}, que utiliza o conceito de comunicações hiper-ligadas de forma a enriquecer as comunicações em conferência com vários tipos de media, editores de texto colaborativos, anotações temporais, envio de mensagens, possilidade de sobrepor contéudo ao vídeo e voltar atrás no tempo.
+O nosso objectivo neste projecto é desenvolver uma aplicação, recorrendo ao \ac{WebRTC}, que utiliza o conceito de comunicações hiper-ligadas de forma a enriquecer as comunicações em conferência com vários tipos de media, editores colaborativos, anotações temporais, envio de mensagens, possilidade de sobrepor contéudo ao vídeo e voltar atrás no tempo.
 

### 2016-10-27 artigo/paper.tex
Initial version of paper to submitt to ICC 2017
Still needs to be shortened to 6 pages.
Working on the introduction.

--- 2016-05-22 tese2/xthesis.tex
@@
-\documentclass[conference,compsoc,a4paper]{IEEEtran}
-\usepackage{geometry}
- \geometry{
- a4paper,
- total={170mm,257mm},
- left=20mm,
- top=20mm,
- }
+%\documentclass[conference,compsoc,a4paper]{IEEEtran}
+\documentclass[10pt,conference]{IEEEtran}
 
-\usepackage[backend=bibtex]{biblatex}
 \usepackage{array}
@@
 \usepackage[nomain,acronym,xindy,toc]{glossaries} 
-\usepackage{graphicx}
 \usepackage{tabularx} 
@@
 
+\usepackage{cite}    
+\usepackage{graphicx,flafter}
+\graphicspath{{../tese2/}}
+
 \usepackage{color, colortbl}
 \definecolor{Gray}{gray}{0.9}
-%Includes "References" in the table of contents
-%\usepackage[nottoc]{tocbibind}
 
-\bibliography{references} % or
+\usepackage{url}
 
@@
 \author{\IEEEauthorblockN{Henrique Rocha}
-\IEEEauthorblockA{Instituto Superior T\'{e}cnico\\}
-\IEEEauthorblockN{Av. Prof. Dr. Anibal Cavaco Silva\\
+\IEEEauthorblockA{INESC-ID / Instituto Superior T\'{e}cnico\\
+Av. Prof. Dr. Anibal Cavaco Silva\\
 2744-016 Porto Salvo, Portugal\\
-henrique.rocha@tecnico.ulisboa.pt}
+Email: henrique.rocha@tecnico.ulisboa.pt}
 \and
-\IEEEauthorblockN{Ricardo Pereira}
-\IEEEauthorblockA{Instituto Superior T\'{e}cnico\\}
-\IEEEauthorblockN{Av. Prof. Dr. Anibal Cavaco Silva\\
+\IEEEauthorblockN{Ricardo Lopes Pereira}
+\IEEEauthorblockA{INESC-ID / Instituto Superior T\'ecnico \\
+Av. Prof. Dr. Cavaco Silva\\
 2744-016 Porto Salvo, Portugal\\
-ricardo.pereira@inesc-id.pt}}
+Email: ricardo.pereira@inesc-id.pt}
+}
 
@@
 \begin{abstract}
-The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content.
+The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to the communication realm.
 This paradigm allows to synchronize, structure and navigate communication content integrated into voice and video calls.
-
 Voice and image together can express emotions like no other medium can. With hypermedia concepts, we can add more value to conference calls.
 
-\emph{WebRTC} technology allows real time communications between web browsers without the need to install additional software.
-The nature of web browser applications already follows the hypermedia concept, which makes \emph{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
-The web browser platform provides an abstraction layer that makes it possible to create applications that run independently from the operating system.
-The native support for \emph{WebRTC} in operating systems extends its usage to outside the web browser, allowing for the exploration of functionalities for which web browsers provide poor support, such as video recording and massive information storage.
+This paper describes the design, implementation and evaluation of a hyper-linked communication platform targeting the web platform and making use of \emph{WebRTC} technology.
+Our prototype provides a hyper-linked, non-linear, multi-party video conferencing platform with support for multiple media types, collaborative text editor, time annotations, instant messaging and a mechanism to superimpose hyper-content to video.
 
-Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging and a mechanism to superimpose hyper-content to video.
+Our prototype was built using a hybrid peer-to-peer and client-server architecture, resorting to open source software.
+Usability tests using 20 users evaluated our prototype positively: 100\% considered it innovative and 95\% recommended its use. 
+From a deployment point of view, our performance tests showed that our solution was expensive to scale up, even though the cost could be significantly lower by disabling some of the functionality.
+
 \end{abstract}
@@
 
-\subsection{Background}
-\label{section:background}
+As communications technologies appeared, we adapted the way we communicate. 
+Multi-party video conference is today a commonly used tool in the enterprise, academia and personal live.
+The \gls{WWW} has made us familiar with the concept of text hyper-media, in the form of hyper-links.
+\gls{IPTV} and video hosting platforms, such as Youtube, made non-linear video watching a common practice.
+The later also use the concept of hyper-links superimposed over videos.
+However, these three concepts, video-conferencing (which is real-time), non-linear video watching and hyper-media, have largely remained independent concepts.
 
-	As communications technologies appeared, we adapted the way we communicate.
 The purpose of this project is not the replacement of the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
 
-	With the advent of WebRTC and its successive integration with web browsers, it became possible to develop video conference web applications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.
+The advent of \gls{WebRTC}, currently available on the major web browsers, made it possible to develop video conference web applications without plugins, greatly increasing the range of what can be implemented using web technologies.
 
@@
 
+\emph{WebRTC} technology allows real time communications between web browsers without the need to install additional software.
+The nature of web browser applications already follows the hypermedia concept, which makes \emph{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
+The web browser platform provides an abstraction layer that makes it possible to create applications that run independently from the operating system.
+The native support for \emph{WebRTC} in operating systems extends its usage to outside the web browser, allowing for the exploration of functionalities for which web browsers provide poor support, such as video recording and massive information storage.
+
 \subsection{Thesis Contribution}
@@
 
-\subsection{Outline}
-
-This rest of this document is structured as follows:
-
-\begin{itemize}
-
-\item \textbf{Chapter \ref{chapter:relatedwork}} describes the previous work in the field.
-\item \textbf{Chapter \ref{chapter:architecture}} describes the system requirements and the architecture for an Web Application that fulfills the goals of this thesis.
-\item \textbf{Chapter \ref{chapter:implementation}} describes the implementation of our Web Application and the technologies chosen.
-\item \textbf{Chapter \ref{chapter:evaluation}} presents the evaluation tests performed and the corresponding results.
-\item \textbf{Chapter \ref{chapter:conclusion}} summarizes the work developed and proposes future work.
-\end{itemize}
+The remainder of this paper is organized as follows.
+Section \ref{chapter:relatedwork} provides an overview of previous work in the field.
+Section \ref{chapter:architecture} describes the requirements we defined for our system and architecture for fulfilling them.
+The choices performed for the implementation of our prototype are presented in Section \ref{chapter:implementation}.
+In Section \ref{chapter:evaluation} we present the evaluation tests performed and their results.
+Section \ref{chapter:conclusion} summarizes the work developed and proposes future work.
 
@@
 
-One way to overcome the \gls{IPv4} address scarcity problem \cite{ipv4} was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \gls{IP} address with a different port.
-This technique is known as \gls{NAT}\cite{rfc3489}.
+Since the early days of video technology, one of the problems raised consisted on how to add more information onto video without generating multiple versions.
+This section examines technologies that allows different ways to present multimedia content in such a way that it change based on synchronization amongst other multimedia elements or user interaction.
 
-\gls{NAT}s weaknesses are being exposing at the application layer, namely impacting applications that require direct communications between two private networks.
+\emph{Hypertext} is a type of text that provides links to texts or other types of content, these links are known by \emph{hyperlinks}.
+\emph{Hypermedia} is an evolution of \emph{hypertext}, it includes audio, images, text and video.
 
-In order to implement an hyper-linked communication solution, several design decisions had to be made.
-The limitations imposed by the \emph{Internet's} structure, its protocols and a browser's capabilities are key factors to consider when implementing a solution that allows bi-directional communications, interactive media and collaboration environment.
+\emph{Hypermedia} concept brings the possibility to organize and overlay multimedia elements into a nonlinear linear structure.
 
-Due to the use of \gls{NAT}, bi-directional communications between clients have different needs from request-response based communications between clients and servers.
-This lead to the appearance of mechanisms such as \gls{STUN}, \gls{TURN} and \gls{ICE}, in order to bypass the limits imposed by \gls{NAT}.
+In the beginning of analog video technology, the navigation over it was quite limited to simple operations such as play, stop, rewind and fast-forward.
+As video started to be digitalized, new operations over video emerged, such as random jumps and chapter navigation through interactive menus present on \gls{DVD}.
 
-\gls{STUN}, \gls{TURN} and \gls{ICE} \cite{natvoip} servers are a possible solution to overcome \gls{NAT}'s exposed weaknesses to applications that require communications between two private networks.
+%RP depois disto falta uma introdução a explicar que precisas de controlar a apresentação e execução de diversos conteúdos e que para isso vais precisar de uma tecnologia que permite sincronizar e controlar.
+Esta secção examina isso.
 
-When connection is established, \gls{WebRTC} came to simplify how audio and video are transmitted through web browsers.
+Some \gls{MPEG} implementations, like \cite{embedded}, added hypermedia information to empty space present on \gls{MPEG} frames in order to provide interactive television, modifying the \gls{MPEG} encoder and decoder in order to handle hypermedia content.
+%RP esta última frase, com detalhes de implementação fica algo deslocada antes de apresentares os exemplos de hypermedia. Talvez deslocar para mais tarde.
+Hypermedia is a concept that holds the promise of future technology and features but it is also already present in our daily lives.
+%RP por exemplo na publicidade no youtube
+
+  Subtitles are an example of information that might be required.
+  The need to translate movies, raised the problem whether it is appropriate to change the original video or audio.
+For example, subtitles should be an entity independent from the video, in order to be personalized or replaced easily.
+
+  \gls{SAMI}, and \gls{SRT} are two of the multiple formats for subtitles commonly supported by video players. Although those formats have styling available, they are quite limited to text.
+
+  Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
+An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time in a video or audio track.
+\emph{HyperCafe}~\cite{hypercafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe.
+
+  Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment.
+\emph{Hyper-Hitchcock}~\cite{hitchcock} is an editor and player of detail-on-demand video.
+
+  In order to navigate through a dynamic video, we must be aware of time synchronization and the multiple time flows, it ils important that all time, causality and behavior rules are well defined.
+  %RP one must be aware? O utilizador? Ou queres dizer que é preciso definir as relações? O que falas a seguir é sobre linguagens para isso.
+
+  \emph{HyVAL}~\cite{hyval} is an \gls{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia.
+HyVAL defines defines video structure, internal video and external media objects.
+  \emph{HyVAL} uses a primary video stream, around which all other elements are organized and synchronized.
+  \emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically. 
+  This approach is quite restrictive if we want to apply hyper-video concepts to videos that do not follow this structure.
+External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
+  %RP não percebo bem a última frase. O que é primary video?
+
+  \gls{SMIL}~\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films.
+With \gls{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
+  %RP synchronize multiple sections of video? vários videos apresentados em simultâneo?
+
+  \gls{SMIL} is an \gls{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Meta Information}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.
+
+\begin{itemize}
+
+  \item The \textbf{Animation} module contains elements and attributes that define a time based mechanism for composing the effects of animations.
+For example, this module can perform changes on \gls{XML} or \gls{CSS} attributes like color and dimensions.
+
+  \item The \textbf{Content Control} module contains elements and attributes that provide optimized alternatives for content delivery.
+For example, it could be used to change audio language in function of user's nationality, for videos with multiple audio channels.
+
+  \item The \textbf{Layout} module contains elements and attributes for coloring and positioning media content. Other layout mechanisms are also possible, such as \gls{CSS}.
+
+  \item The \textbf{Linking} module contains elements and attributes for navigational \emph{hyperlinking}. Navigation can be triggered by events or user interaction.
+
+  \item The \textbf{Media Object} module contains elements and attributes for referencing rendering behavior of external multimedia or control objects.
+
+  \item The \textbf{SmilText} module contains elements and attributes that define and control timed text. For example, this module could be used to create labels and captions.
+
+  \item The \textbf{Meta Information} module contains elements and attributes that allows describing the \gls{SMIL} document.
+For example, this module could be used to define movie details such as category, director, writers and cast.
+
+  \item The \textbf{Structure} module defines the basic elements and attributes for structuring \gls{SMIL} content.
+This module defines a \emph{head} element that contains non temporal behavior information defined by  \emph{Meta Information}, \emph{Layout} and \emph{Content Control} modules.
+This module also defines the \emph{body} element, where all temporal related module information is contained.
+
+  \item The \textbf{Timing} module is the most important module on \gls{SMIL} specification.
+Due to its complexity, it is divided into seventeen sub-modules for coordination and synchronization of media over time.
+The three main elements are \emph{seq}, \emph{excl} and \emph{par}, that, respectively, play child elements in sequence, one at a time and all at the same time.
+
+  \item The \textbf{Time Manipulations} module adds time behavior attributes to \gls{SMIL} elements, such as speed, rate or time.
+
+  \item The \textbf{State} module defines attributes that define the state of \gls{SMIL} elements, such as element visibility, current element time, amount of repeated loops, playing state and many others.
+
+  \item The \textbf{Transitions} module defines attributes and elements for transitions across multiple \gls{SMIL} elements according to the \emph{Timing} module.
+
+\end{itemize}
+
+The \gls{DOM} is a standard \gls{API} that allows easy management of documents that are organized in a tree structure, by providing \gls{CRUD} operations over its elements and their attributes.
+\gls{DOM} makes it easy to inter-operate between imperative and declarative programming languages~\cite{dom}.
+%RP não arranjas uma referência para o dom?
+
+Like \gls{DOM}, \gls{SMIL} \gls{DOM} is an \gls{API} for \gls{SMIL} documents.
+Allowing \gls{CRUD} operations over \gls{SMIL} documents is an important feature for extending \gls{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \emph{JavaScript} functions.
+
+  \gls{SMIL}'s modules can be used to synchronize and animate \gls{XHTML} and \gls{SVG} elements.
+  %RP ``are used'' ou ``can be used''?
+
+  \gls{SMIL} fits our goals for creating a multimedia rich hyper-call, but it lacks on browser compatibility.
+Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \gls{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plug-in.
+
+  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \gls{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, does not require a plug-in to be installed and should not have incompatibility issues.
+  SmillingWeb already takes advantage of \gls{HTML}5 and \gls{CSS}3.
+  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}(accessed June 2, 2015).}, a simple \emph{JavaScript} library that may require plug-ins if new features are not supported.
+  But SmillingWeb just implements a subset of \gls{SMIL} 3.0 and their scheduler engine loads the \gls{SMIL} file only once, which could raise problems when dealing with \gls{SMIL} changes due to real time communications.
+  Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to high latency networks leading to  pauses during playback.
+
+% XXX
+  %RP fazes uma mudanças de tema muito bruscas. Podes usar \subsubsection ou \paragraph para dividir isto melhor. E uma frase entre temas a servir de ``cola'' ou ``separação'' também não fica mal
+
+  An alternative to \gls{SMIL}'s \emph{Layout} module is to use \gls{HTML} which is a markup language based on \gls{XML} that is used for creating web pages.
+\gls{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages.
+Languages like \gls{CSS} and \emph{JavaScript} are typically used along with \gls{HTML} for improving the interaction and appearance of a web page.
+
+  %RP esta parte está um pouco confusa.
+Primeiro falas de smil, depois de html, depois de smil novamente.
+Devias agrupar tudo.
+Devias agrupar as soluçoes por tecnologia.
+Se preciso, primeiro fala de html, css, etc, e depois de smil e das implementação de smil (incluindo as que são implementadas em html e js).
+
+  \gls{CSS}'s goal is to separate the structure of an \gls{XML} document from its appearance. \gls{CSS} defines styles for \gls{XML} tags based on their name, class, identifier or position.
+  Besides static styling \gls{CSS} also supports animation and transitions leading to more dynamic content.
+  %RP Tens muitas alturas em que não começas frases novas. Usas virgulas para separar conteúdo que não faz sentido ficar na mesma frase pois começa com um novo sujeito seguido de verbo.
+
+  \emph{JavaScript} is an imperative object-oriented language based on \emph{ECMAScript}.
+It is used mainly on client-side and executed by a web browser.
+\emph{JavaScript} has its own implementation of \gls{DOM} and one of its advantages is the ability to download and execute code on-the-fly without the need of pre-installed plug-ins.
+
+  \emph{JavaScript} has compatibility issues among the different web browsers, leading to different behaviors.
+To solve that problem, there are libraries written in \emph{JavasSript}, namely \emph{jQuery}, that implements the same functionality for multiple browsers, masking most of the incompatibility issues.
+
+  With the emergence of \gls{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple \emph{codecs}, audio and subtitles in \gls{WebVTT} format.
+Another important tag is \emph{canvas} that allows drawing graphics with \emph{JavaScript} on a rectangle within a web page.
+
+  For example, with \gls{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}(accessed June 2, 2015).}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.
+Another example would be a collaborative spreadsheet using \gls{WebRTC}.
+With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
+
+  \gls{SVG} is an \gls{XML} based format that incorporates the animation module of \gls{SMIL}.
+Currently, \gls{SVG} allows adding movement and animating attributes of elements.
+When embedded on \gls{HTML}5, it allows dynamic changes to inner content in real time through the \gls{DOM} \gls{API}.
+Besides that, it also allows calling \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.
+
+  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}(accessed June 2, 2015).} was developed for web-based animations.
+Introducing video support in 2002, \emph{flash} started to grow after that.
+Competitor's players, at that time were focused on playing video and audio, while \emph{flash} had vector graphics and focused on streaming \emph{on-demand} video across multiple platforms.
+\emph{VP6} was their choice of video \emph{codec}, providing half the video size (in Byte) for the same quality and providing video quality adjusted to \emph{Internet} connection latency.
+  In 2010, \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \gls{HTTP} and streaming using \gls{RTMP}.
+
+  \gls{RTMP} is a \gls{TCP} based protocol used for streaming audio, video and data between a \gls{FMS} and \emph{flash} players.
+A bidirectional connection is established between the two in order to allow real time communications.
+A \emph{flash} player can stream a webcam video to a \gls{FMS} using \gls{RTMP} or it can request a video stream to \gls{FMS} that can either be a pre-recorded stream, live stream or data.
+Multiple \gls{FMS} servers can be used in parallel to increase capacity and handle more streams simultaneously.
+  %RP chained ou usados em paralelo?
+
+  \gls{FMS} can stream video and audio to one or more subscribers by sending a separate copy for each subscriber.
+With \gls{RTMFP} it is possible to stream video directly between \emph{flash} players, allowing a publisher to break up a stream into pieces that can be cooperatively distributed in a P2P mesh.
+\gls{RTMFP} uses \gls{UDP} to speed packet delivery, which although it is not reliable, is well suited for video streaming.
+Like \gls{WebRTC}, \emph{flash} players also need to apply techniques like \gls{STUN} and \gls{TURN} for \gls{NAT} traversal.
+
+  Although \gls{HTML}5, \emph{JavaScript}, \gls{CSS} and \gls{WebRTC} implement some of \emph{flash}'s features, it does not mean that \emph{flash} will be replaced soon.
+  Instead, both technologies can be used to develop rich \emph{Internet} applications.
+  It is also important to note that \gls{HTML}5 is better supported in mobile devices than \emph{Adobe Flash}.
+
+  Like \emph{Flash}, \emph{Microsoft Silverlight}\footnote{\url{http://www.microsoft.com/silverlight/}(accessed June 2, 2015).} is a cross browser plug-in and platform that is used to develop rich \emph{Internet} applications.
+It supports vector graphics, animation and video.
+Compared to \emph{flash}, which uses \emph{ActionScript}, \emph{Silverlight} applications can use languages like C\#, \emph{VisualBasic} and \gls{XAML}.
+\emph{Silverlight} uses a technique called \emph{Smooth Streaming} from \emph{IIS Media Service} that consists on delivering video in real time with adjusted quality in function of bandwidth variations and \gls{CPU} usage.
+
+  Using technologies that relies only on web standards, like \gls{CSS}, \gls{HTML}5, \emph{JavaScript} and \gls{SVG}, will make possible to develop an application that solves our problem with the advantage of being compatible with a greater amount of web browsers.
+
 \gls{WebRTC} is an open source technology that defines a collection of standard protocols and \emph{JavaScript} \gls{API}s for web browser based real time communications without installing any additional application or plug-in.
@@
 
-  One of \gls{WebRTC} signaling's requisites is bi-directional communication.
-The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel \cite{rfc6455}, by other words it supports sending and receiving data simultaneously.
-
-With the communications establishment issue solved, we had to discuss the different types of media and what can be done with each kind in order to increase the value of communications among users.
-In this context, we have studied solutions and libraries that allow us to implement our prototype with time manipulation features, collaborative text edition, record and playback interactive video.
-
 	\emph{Hypermedia} concept brings the possibility to organize and overlay multimedia elements into a nonlinear linear structure holding the promise of future technology and features.
@@
 Two examples of applications that captured our attention were \emph{HyperCafe}~\cite{hypercafe} and \emph{Hyper-Hitchcock}~\cite{hitchcock} which explored interactive video features.
-
-  Using technologies that relies only on web standards, like \gls{CSS}, \gls{HTML}5, \emph{JavaScript} and \gls{SVG}, will make possible to develop an application that applies the hypermedia concept with the advantage of being compatible with a greater amount of web browsers.
 
@@
 
+\subsection{Streaming and Recording}\label{recstream}~\\
+	If, for instance, one wants to rewind a real time video, recordings will be needed from whom is streaming the video. 
+ Our first concern on real time collaboration applications, besides the communication itself, is the data storage and representation.
+Storing multimedia content is not a viable solution because most browsers recommend limiting local storage to at most five megabytes per origin\footnote{\url{http://www.html5rocks.com/en/tutorials/offline/quota-research/} (accessed May 13, 2016)} .
+
+ 	In order to provide a way to record and playback streams, additional servers will be required to process and record the large amount of data generated by audio and video streaming.
+
+ \gls{RTP}\cite{rfc3550} is used for streaming audio and video over \gls{IP}.
+ Multimedia content is transported on the payload of \gls{RTP} messages, that contain headers for payload identification.
+\gls{RTP} is independent from its payload type, allowing it to transport any kind of encoded multimedia.
+A sequence number is used for sorting received packets.
+
+\gls{RTCP} is used for controlling \gls{RTP} multimedia streams, it provides bandwidth statistics and control information that can be used for changing the quality of the stream in real time.
+
+  \gls{RTP} allows to change its requirements and add extensions to it with profiles.
+One of the most used ones is the \gls{RTP} profile for audio and video \cite{rfc3551}, which lists the payload encoding and compression algorithms.
+This profile also assigns a name to each encoding which may be used with other protocols like \gls{SDP}.
+  Another profile for \gls{RTP} is defined by \gls{SRTP}, which provides encryption, authentication and replay protection for \gls {RTP} traffic.
+The analogous secure protocol for \gls{RTCP} is \gls{SRTCP}.
+
+  Both \gls{SRTP} and \gls{SRTCP} use \gls{AES} by default, which is a symmetric-key algorithm for data encryption.
+Each packet is encrypted using a distinct key-stream, as otherwise, using a single key-stream with \gls{AES} on \gls{CBC} would make it impossible to recover from packet loss.
+
+  Two key-stream generators for \gls{AES} were defined: \emph{Segmented Integer Counter Mode} and \emph{f8-mode}.
+If a packet is lost, there is no impact on other packets, as the initialization vector is obtained through those key-stream generators and it is fixed for each packet.
+  %RP isto está tudo no RFC3550? Se não, faltam muitas citações.
+  %HR esse RFC é grande
+
+  \gls{RTP} recorders are independent of payload encoding, they do not decode \gls{RTP} packets, they record packets instead, allowing to record all video and audio formats even if they're encrypted.
+
+        %RP lá vais tu mudar de tema da conversa sem explicar porquê....
+        Even though one on one calls are common, there are occasions when several people take part of the same video call.
+	Multi-party video calls can be achieved on \gls{WebRTC} by streaming video from each participant to all the other participants.
+Although this works, the bigger a conference room is, the bigger is the used bandwidth to stream video to all participants within the conference room.
+        In this scenario, a more efficient alternative to peer-to-peer is the use of a \gls{MCU}.
+
+	\emph{Jitsi Video Bridge} \footnote{\url{http://jitsi.org/Projects/JitsiVideobridge}(accessed June 2, 2015).} receives one stream from every participant on a conference, either from a \emph{jitsi} client or a \emph{WebRTC} application, and redirects it to all the other conference participants, reducing the amount of data that each peer sends.
+Although all the participants still need to download all the streams from the \emph{Jitsi Video Bridge} server, typically download rates are much bigger than upload rates, making this solution more feasible.
+	\emph{Jitsi Video Bridge} uses \gls{XMPP} as a signaling protocol and its \emph{colibri} extension \cite{xep0340} to reserve channels for video transmission.
+Despite this choice for signaling protocol, \emph{Jitsi Video Bridge} also supports \gls{SIP}.
+
+	  \gls{KMS} supports trans-coding, group communications, recording, mixing, broadcasting, applying filters and performing image and sound analysis (for example it allows face and QRcode detection).
+\gls{KMS} functionalities are exposed through \emph{JSON-RPC} over \emph{WebSocket}.
+          There are three clients available: \emph{JavaScript} client for web browsers, \emph{Java} client for \emph{Java EE} servers and \emph{JavaScript} client for \emph{Node.js} servers. 
+	\gls{KMS} supports streaming over \emph{WebRTC}, \gls{HTTP} and \gls{RTP} endpoints.
+Another important component of \gls{KMS} is \emph{Kurento Repository}, which supports recording and playing directly from \emph{MongoDB}.
+That is important for providing a scalable media storage.
+Unlike \emph{Jitsi Video Bridge}, \gls{KMS} does not enforce a specific signaling protocol.
+
 \section{Architecture}
@@
 \subsection{Requirements}
+\label{chapter:requirements}
 In a general way our system's goal is to provide a multi-party video and audio conference environment that supports chat, time manipulation, collaborative text edition and hyper-content creation.
@@
 \section{Implementation}
-
+\label{chapter:implementation}
 \subsection{Data Model}
@@
 \section{Evaluation}
+\label{chapter:evaluation}
 
@@
 \section{Conclusions}
+\label{chapter:conclusion}
 \subsection{Achievements}
@@
 
-%Sets the bibliography style to UNSRT and imports the 
-%bibliography file "samples.bib".
-%\bibliographystyle{unsrt}
-%\bibliography{references}
+% trigger a \newpage just before the given reference
+% number - used to balance the columns on the last page
+% adjust value as needed - may need to be readjusted if
+% the document is modified later
+%\IEEEtriggeratref{8}
+% The "triggered" command can be changed if desired:
+%\IEEEtriggercmd{\enlargethispage{-5in}}
 
-%\input{thesis.bbl}
-\printbibliography
+\bibliographystyle{IEEEtran}
+\bibliography{../tese2/references}
+
 \end{document}

### 2016-10-29 artigo/paper.tex
Submission to ICC 2017

--- 2016-05-22 tese2/xthesis.tex
@@
 
+\usepackage{cite}    
+\usepackage{graphicx,flafter}
+\graphicspath{{../tese2/} {.}}
+
 \usepackage{color, colortbl}
 \definecolor{Gray}{gray}{0.9}
-%Includes "References" in the table of contents
-%\usepackage[nottoc]{tocbibind}
 
-\bibliography{references} % or
+\usepackage{url}
 
@@
 \begin{abstract}
-The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content.
+The Hyper-linked communications concept applies much of the hypermedia concepts, widely used on Web content, to the communications realm.
 This paradigm allows to synchronize, structure and navigate communication content integrated into voice and video calls.
+Voice and image together can express emotions like no other medium can. 
+With hypermedia concepts, we can add even more value to conference calls.
 
-Voice and image together can express emotions like no other medium can. With hypermedia concepts, we can add more value to conference calls.
+This paper describes the design, implementation and evaluation of a hyper-media communication platform targeting the web platform and making use of \emph{WebRTC} technology.
+Our prototype provides a hyper-linked, non-linear, multi-party video conferencing platform with support for multiple media types, collaborative text editor, time annotations, instant messaging and a mechanism to superimpose hyper-media to video.
 
-\emph{WebRTC} technology allows real time communications between web browsers without the need to install additional software.
-The nature of web browser applications already follows the hypermedia concept, which makes \emph{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
-The web browser platform provides an abstraction layer that makes it possible to create applications that run independently from the operating system.
-The native support for \emph{WebRTC} in operating systems extends its usage to outside the web browser, allowing for the exploration of functionalities for which web browsers provide poor support, such as video recording and massive information storage.
+%Our prototype was built using a hybrid peer-to-peer and client-server architecture, resorting to open source software.
+Usability tests using 20 users evaluated our prototype positively: 100\% considered it innovative and 95\% recommended its use. 
+From a deployment point of view, our performance tests showed that our solution was expensive to scale up, even though the cost could be significantly lowered by disabling some of the advanced functionality.
 
-Our goal was the development of an application targeted to the web platform, resorting to \emph{WebRTC}, that leveraged the hyper-linked communications by providing a conference environment enriched with multiple media types, collaborative text editors, time annotations, instant messaging and a mechanism to superimpose hyper-content to video.
 \end{abstract}
@@
 
-\subsection{Background}
-\label{section:background}
+As communications technologies evolved, we adapted the way we communicate. 
+Multi-party video conference is today a commonly used tool in the enterprise, academia and personal live.
+The \gls{WWW} has made us familiar with the concept of text hyper-media, in the form of hyper-links.
+\gls{IPTV} and video hosting platforms, such as Youtube, made non-linear video watching a common practice.
+The latter also use the concept of hyper-links superimposed over videos.
+However, these three concepts, video-conferencing (which is real-time), non-linear video watching and hyper-media, have largely remained independent concepts.
 
-	As communications technologies appeared, we adapted the way we communicate.
-The purpose of this project is not the replacement of the current video and audio communications, but to enrich them with hyper-media content and make them a more natural and easy to learn process.
+This paper describes our efforts to develop a prototype that combines these three concepts, providing a video-conferencing web application that allows users to navigate back and forth in time to catch up on what was said, while the video-conferencing is going on.
+We developed a collaboration application, targeted at the web platform resorting to \gls{WebRTC}, that leverages the hyper-linked communications concept by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \gls{HTML}5, \gls{CSS}3 and \emph{JavaScript}.
+Our collaboration application also includes a collaborative text editor and a chat that supports sending time hyper-links and files to conference participants.
 
-	With the advent of WebRTC and its successive integration with web browsers, it became possible to develop video conference web applications without plugins, this presents a range of possibilities on what can be implemented using already existing web technologies.
+One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again.
+Users can use a time bar to navigate in time: rewinding communications, fast-forward or jump to a certain point in time.
+Time hyper-links also enable accessing the collaboration (video, chat, text editor, etc) at a given instance in time.
+Users are capable of enriching the video content, in real-time or at a latter time, by adding content to it, such as timed annotations, interactive lists of topics and subtitles, thus also improving content search abilities.
+In this context, we also provide a method for real-time creation and synchronizing of hyper-linked content using \gls{QR} codes.
 
-    Furthermore, real time communication applications can make a significant difference on business, education and health sectors by providing tools for developing teaching and learning online, teamworking and socializing web applications.
+We envision the use of our collaboration platform in many situations, such as: virtual classroom, where students can interact with the professor during class and later review the subjects, while the latter can enrich the recording with extra materials and annotation after the lecture; recording meetings, where the video stream can be annotated and bookmarked with the agenda points, and support documents added.
 
-\subsection{Proposed Solution}
-\label{section:proposed}
+The remainder of this paper is organized as follows.
+Section \ref{chapter:relatedwork} provides a background on the field.
+Section \ref{chapter:architecture} describes the requirements we defined for our system and architecture for fulfilling them.
+The choices performed for the implementation of our prototype are presented in Section \ref{chapter:implementation}.
+In Section \ref{chapter:evaluation} we present the evaluation tests performed and their results.
+Section \ref{chapter:conclusion} summarizes the work developed and proposes future work.
 
-	Our goal in this project is to develop an application targeted to the web platform, resorting to \gls{WebRTC}, that leverages the hyper-linked communications by providing a video conference environment enriched with interactive and non-interactive discrete media types such as images, subtitles, forms and all types of content that can be added using \gls{HTML}5, \gls{CSS}3 and \emph{JavaScript} including continuous media types such as video, music and animations.
-
-	One of the key features of this project is the ability to navigate in time in order to reproduce the conversation again or introduce hyper-content to it such as time annotations, interactive lists of topics and subtitles.
-In this context we also provide a simpler method for creating and synchronizing hyper-content using \gls{QR} codes.
-
-	In addition to this conference environment, which provides different functionalities than traditional conference environments such as \emph{Skype} and \emph{Google Hangouts}, we also enable a collaborative text editor and a chat that supports sending time hyper-links and files to conference participants.
-
-	Furthermore, another relevant feature is the possibility to compose multiple video streams into a single one, which enables adding more users to conference rooms without impacting on clients performance.
-Users can change to individual streams on demand or automatically to the talking users.
-
-\subsection{Thesis Contribution}
-\label{section:contribution}
-
-Making it clear, this project aims to complement current audio, text and video communications in order to create rich and collaborative interfaces with the ability to add more content on a future time (\emph{e.g.} creating time annotations for improving content search) in order to increase its value.
-It is also important to highlight another goal of this project which is the ability to navigate in time by rewinding communications, fast-forward and jump to certain points.
-
-	We have presented an architecture that can meet our goals, implemented the respective prototype and tested it with real users and performance benchmarks.
-
-	According to Martin Geddes, the quality of the interaction worsens as the number of users increase\cite{geddes}.
-In our testing phases we will quantify and qualify the impact of increasing users on the interface and performance of our prototype.
-
-	All the problems faced during the development and limitations were reported on the thesis so that a future project better then ours can be easily and better developed.
-
-\subsection{Outline}
-
-This rest of this document is structured as follows:
-
-\begin{itemize}
-
-\item \textbf{Chapter \ref{chapter:relatedwork}} describes the previous work in the field.
-\item \textbf{Chapter \ref{chapter:architecture}} describes the system requirements and the architecture for an Web Application that fulfills the goals of this thesis.
-\item \textbf{Chapter \ref{chapter:implementation}} describes the implementation of our Web Application and the technologies chosen.
-\item \textbf{Chapter \ref{chapter:evaluation}} presents the evaluation tests performed and the corresponding results.
-\item \textbf{Chapter \ref{chapter:conclusion}} summarizes the work developed and proposes future work.
-\end{itemize}
-
-\section{Related Work}
+\section{Background}
 \label{chapter:relatedwork}
 
-One way to overcome the \gls{IPv4} address scarcity problem \cite{ipv4} was the development of a mechanism that groups multiple address into a single one, the machine that is assigned that address is then responsible for redirecting messages to members of its group using their private addresses, each connection in the private network is identified publicly by the same \gls{IP} address with a different port.
-This technique is known as \gls{NAT}\cite{rfc3489}.
+\emph{Hypertext} is a type of text that provides links to texts or other types of content, these links are known by \emph{hyperlinks}. 
+\emph{Hypermedia} is an evolution of \emph{hypertext}, it includes audio, images, text and video. 
+\emph{Hypermedia} concept brings the possibility to organize and overlay multimedia elements into a nonlinear structure~\cite{geddes}.
 
-\gls{NAT}s weaknesses are being exposing at the application layer, namely impacting applications that require direct communications between two private networks.
+Some \gls{MPEG} implementations added hypermedia information to empty space present on \gls{MPEG} frames in order to provide interactive television, modifying the \gls{MPEG} encoder and decoder in order to handle hypermedia content~\cite{embedded}.
 
-In order to implement an hyper-linked communication solution, several design decisions had to be made.
-The limitations imposed by the \emph{Internet's} structure, its protocols and a browser's capabilities are key factors to consider when implementing a solution that allows bi-directional communications, interactive media and collaboration environment.
+Hyper-video is a kind of video that contains links to any kind of hypermedia, including links to skip part of it.
+An example of hypermedia application could be a search engine over hypermedia content, like subtitles, in order to jump to a specific time in a video or audio track.
+\emph{HyperCafe} was an experimental project to expose hyper-video concepts that consisted of an interactive film that enabled switching between different conversations taking place inside a cafe~\cite{hypercafe}.
 
-Due to the use of \gls{NAT}, bi-directional communications between clients have different needs from request-response based communications between clients and servers.
-This lead to the appearance of mechanisms such as \gls{STUN}, \gls{TURN} and \gls{ICE}, in order to bypass the limits imposed by \gls{NAT}.
+Detail-on-demand is a subset of hyper-video that allow us to obtain additional information about something that appears along the video, like obtaining information about a painting that appears in a particular segment.
+\emph{Hyper-Hitchcock} is an editor and player of detail-on-demand video~\cite{hitchcock}.
 
-\gls{STUN}, \gls{TURN} and \gls{ICE} \cite{natvoip} servers are a possible solution to overcome \gls{NAT}'s exposed weaknesses to applications that require communications between two private networks.
+\emph{HyVAL}~\cite{hyval} is an \gls{XML} based language that was proposed for modeling composition, synchronization and interaction of hypermedia. 
+HyVAL defines defines video structure, internal video and external media objects. 
+\emph{HyVAL} uses a primary video stream, around which all other elements are organized and synchronized.
+\emph{HyVAL}'s video structure object defines a structure derived from traditional video, which divides video into segments, scenes, shots and frames hierarchically. 
+  This approach is quite restrictive if we want to apply hyper-video concepts to videos that do not follow this structure. 
+%External media objects are linked by primary video, those objects can represent other videos, images, text, animation and sound.
+  %RP não percebo bem a última frase. O que é primary video?
 
-When connection is established, \gls{WebRTC} came to simplify how audio and video are transmitted through web browsers.
+  \gls{SMIL}~\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films. 
+With \gls{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
+  %RP synchronize multiple sections of video? vários videos apresentados em simultâneo?
+
+ % \gls{SMIL} is an \gls{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Meta Information}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.
+
+%The \gls{DOM} is a standard \gls{API} that allows easy management of documents that are organized in a tree structure, by providing \gls{CRUD} operations over its elements and their attributes.
+\gls{DOM} makes it easy to inter-operate between imperative and declarative programming languages~\cite{dom}.
+%RP não arranjas uma referência para o dom?
+
+%Like \gls{DOM}, \gls{SMIL} \gls{DOM} is an \gls{API} for \gls{SMIL} documents.
+Allowing \gls{CRUD} operations over \gls{SMIL} documents is an important feature for extending \gls{SMIL} capabilities, for example for creating non-linear animations and triggering external events like \emph{JavaScript} functions.
+
+%  \gls{SMIL} fits our goals for creating a multimedia rich hyper-call, but it lacks on browser compatibility.
+Ambulant \cite{ambulant} was one of the SMIL players that were developed for browsers, although this player implements most of \gls{SMIL} 3.0 \cite{smil3} specifications, it needs to be installed on browsers as a plug-in.
+
+%  SmillingWeb \cite{smillingweb} attempts to implement a cross platform multimedia player designed for \gls{SMIL} 3.0 presentations with \emph{JavaScript} and \emph{jQuery} which, unlike \cite{ambulant}, does not require a plug-in to be installed and should not have incompatibility issues.
+%  SmillingWeb already takes advantage of \gls{HTML}5 and \gls{CSS}3.
+%  It takes into account unsupported web browsers through the use of \emph{Modernizr}\footnote{\url{http://modernizr.com/}(accessed June 2, 2015).}, a simple \emph{JavaScript} library that may require plug-ins if new features are not supported.
+%  But SmillingWeb just implements a subset of \gls{SMIL} 3.0 and their scheduler engine loads the \gls{SMIL} file only once, which could raise problems when dealing with \gls{SMIL} changes due to real time communications.
+ % Another problem with SmillingWeb is pre-loading and playing elements at the correct interval of time, which is not always possible due to high latency networks leading to  pauses during playback.
+
+ % An alternative to \gls{SMIL}'s \emph{Layout} module is to use \gls{HTML} which is a markup language based on \gls{XML} that is used for creating web pages.
+\gls{HTML} alone is a very poor language when we are focused on visual appealing and interactive web pages.
+Languages like \gls{CSS} and \emph{JavaScript} are typically used along with \gls{HTML} for improving the interaction and appearance of a web page.
+
+  With the emergence of \gls{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple \emph{codecs}, audio and subtitles in \gls{WebVTT} format.
+Another important tag is \emph{canvas} that allows drawing graphics with \emph{JavaScript} on a rectangle within a web page.
+
+%  For example, with \gls{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}(accessed June 2, 2015).}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call.
+Another example would be a collaborative spreadsheet using \gls{WebRTC}.
+With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
+
+  \gls{SVG} is an \gls{XML} based format that incorporates the animation module of \gls{SMIL}.
+Currently, \gls{SVG} allows adding movement and animating attributes of elements.
+When embedded on \gls{HTML}5, it allows dynamic changes to inner content in real time through the \gls{DOM} \gls{API}.
+Besides that, it also allows calling \emph{JavaScript} functions on events such as animation end, mouse over and mouse click.
+
 \gls{WebRTC} is an open source technology that defines a collection of standard protocols and \emph{JavaScript} \gls{API}s for web browser based real time communications without installing any additional application or plug-in.
-
 \gls{WebRTC} uses \gls{SDP} \cite{rfc4566} to define peer connection properties such as types of supported media, \emph{codecs}, protocols used and network information.
@@
 
-Some operating systems such as \emph{Android}, \emph{iOS}, \emph{Linux}, \emph{OSX} and \emph{Windows} implement native \gls{WebRTC} libraries, extending the usage of \gls{WebRTC} to applications outside the web browser.
+%Some operating systems such as \emph{Android}, \emph{iOS}, \emph{Linux}, \emph{OSX} and \emph{Windows} implement native \gls{WebRTC} libraries, extending the usage of \gls{WebRTC} to applications outside the web browser.
 This native support can help to implement applications that record video and audio streams for further playback.
@@
 However, \gls{WebRTC} by itself does not define how users get to know each other nor how information flows between users. 
-For this reason, we have studied the multiple ways we could implement this \emph{get-to-know} mechanism which is known as signaling protocol.
-
   Signaling is the process by which applications exchange connection information about peers and servers, their capabilities and meta-data.
@@
 
-  One of \gls{WebRTC} signaling's requisites is bi-directional communication.
-The \emph{WebSocket} protocol allows bi-directional communications over a full-duplex socket channel \cite{rfc6455}, by other words it supports sending and receiving data simultaneously.
-
-With the communications establishment issue solved, we had to discuss the different types of media and what can be done with each kind in order to increase the value of communications among users.
-In this context, we have studied solutions and libraries that allow us to implement our prototype with time manipulation features, collaborative text edition, record and playback interactive video.
-
-	\emph{Hypermedia} concept brings the possibility to organize and overlay multimedia elements into a nonlinear linear structure holding the promise of future technology and features.
-Languages such as \gls{SMIL}~\cite{hyval}, \emph{HyVAL}~\cite{hyval} and \gls{HTML} can be used to implement the \emph{Hypermedia} concept.
-Two examples of applications that captured our attention were \emph{HyperCafe}~\cite{hypercafe} and \emph{Hyper-Hitchcock}~\cite{hitchcock} which explored interactive video features.
-
-  Using technologies that relies only on web standards, like \gls{CSS}, \gls{HTML}5, \emph{JavaScript} and \gls{SVG}, will make possible to develop an application that applies the hypermedia concept with the advantage of being compatible with a greater amount of web browsers.
-
-\subsection{Extending collaboration tools with time manipulation}
-\label{collab}
-
-Real time collaboration applications have become a huge help on team tasks, providing a great boost on business, research and investigation velocity.
-
- Our first concern on real time collaboration applications is the data storage and representation.
-Storing multimedia content on a web client is not a viable solution because the local storage is limited to at most five megabytes per origin.
-Additional servers will be required to process and record the large amount of data generated by audio and video streaming.
-
-\gls{KMS} supports streaming over \emph{WebRTC}.
-Another important component of \gls{KMS} is \emph{Kurento Repository}, which supports recording and playing directly from \emph{MongoDB}.
-That is important for providing a scalable media storage.
-
-     \gls{OT} technology was originally developed for consistency maintenance and concurrency control over distributed objects.
-     \gls{OT} algorithms are mainly used in collaborative applications such as distributed document edition.
-
-	Among mutiple \gls{OT} platforms and libraries we present \emph{ShareJS}\footnote{\url{http://sharejs.org/}(accessed June 2, 2015).}, \emph{TogetherJS}\footnote{\url{http://togetherjs.com/}(accessed June 2, 2015).}, \emph{Goodow}\footnote{\url{http://realtimeplayground.goodow.com/}(accessed June 2, 2015).}, \emph{Etherpad Lite}\footnote{\url{https://github.com/ether/etherpad-lite}(Accessed 20 March 2016)} and \emph{otJS}\footnote{\url{http://operational-transformation.github.io}(accessed March 10, 2016)}.
-
-	\emph{otJS} is a \emph{JavaScript} library that only implements operation transformations over plain text on the client side and requires implementing the content's persistent storage.
-Besides this drawback, this library is very flexible because it is not tied to a specific database or server side technology.
-
 \section{Architecture}
@@
 
-Taking into account the goals of this project and all the technology presented so far.
-Our proposal is the development of a web application that provides communication and collaboration features in real time.
+From an user experience point of view, we designed our collaboration application around the concept of group conferences (or conference rooms).
+Users can create group conferences and invite other users to participate in these, either by adding the other user or by sharing a \gls{URL} (invitation link).
+In order to meet user expectation, a group conference provides functionality similar to those found in established applications such as \emph{Skype} or \emph{Google Hangouts}.
+Among other common features, one can view a small video of each participant or a large video of a selected participant.
+Each participant may use more than one camera, in order to show different angles or objects, e.g. a white board and himself.
+The source of the large video can be selected manually or automatically assigned to the participant talking.
+Participating users may choose to view only, share their microphone, camera or screen.
 
-\subsection{Requirements}
-In a general way our system's goal is to provide a multi-party video and audio conference environment that supports chat, time manipulation, collaborative text edition and hyper-content creation.
+Besides the videos, a group conference includes a chat, collaborative text editor, list of users and document sharing.
+The major differentiating feature is a time-line bar, for navigating in time.
+While the group conference is taking place, this time bar is continuously progressing as time goes by.
+The user is free to slide over to any instant in the past or back to real-time.
+Changing the instant of time updates the state of not only the video but of other collaboration elements.
 
-%RP isto pode ser muito expandido. Podes começar por descrever os goals (mais alto nível): audio e video conferência multi-party, possibilidade de rever o que foi dito antes, chat, etc etc.
-% Só depois traduzes isso em requisitos, tanto funcionais como não funcionais
-% Depois apresentas a arquitectura, fazendo um paralelo entre os módulo e escolhas realizadas e quais os requisitos que cada um deles satisfaz.
+Either in real-time or later, users can create bookmarks, displayed in the time-line, for easier navigation in time.
+Users can also add content to superimpose on the video.
+This content can be as simple as subtitles or as complex as a \gls{HTML} document or JavaScript code to perform an animation.
+All the content is searchable, allowing for time navigation by content.
 
-For our system, our application must provide: a simple way to send instant text messages to the conference participants, ability to recording and playback recorded video including all the hyper-content displayed at that time, mixing multiple user streams into a single stream, create annotations associated to a specified time, allow users to superimpose hyper-content to video given a range of time, search every objects related to a conference room such as hyper-content, time annotations and users, share files and time links among users, sound detection for showing the current speaker, provide a collaborative text edition tool, interpret \gls{QR} codes in order to ease content creation and support database replication.
-%RP capacidade de lidar com NAT
+From an early stage, we decided to target the browser platform, due to its ubiquity and breadth of devices, from desktops, to tablets and smartphones.
+The nature of web browser applications already follows the hypermedia concept, which makes \emph{WebRTC} the ideal technology to apply the hyper-linked communications concepts.
+The advent of \gls{WebRTC}, currently available on the major web browsers, makes it possible to develop video conference web applications without plugins.
+However, mobile devices have limited resources for handling a large number of simultaneous video stream.
 
-	Moreover we allow clients to discover chat rooms and other clients by navigating on the web pages provided by our web server.
-In addition users can create rooms for multi-party audio and video communication communication which is achieved by using \gls{WebRTC}'s \emph{PeerConnection}.
-        %RP falas em peerconnection, mas na figura 3.1 apenas há um cliente, o que obviamente impede mostrar a comunicação entre peers.
-        %RP acho que podes fazer outro diagrama com a comunicação entre os vários componentes, onde apareceriam vários clientes.
-A aí indicas o tipo de ligação entre cada par de componentes (HTTP, PeerConnectino, websocket, MediaStream ...)
+In order to limit the processing requirement in the client application (browser), we decided to use a server to combine the audio and video streams from all the participants into a single video and a single audio track.
+This also helps conserve battery on mobile devices.
+Each user listens to the combined audio track and may choose to see a single participant or all of them, in this case using the combined video track.
 
-\subsection{Modules}
-In this section we present the several modules that were designed in order com fulfill the set requirements.
-	Figure~\ref{fig:modules} presents the structure of our system which was divided into six modules:
+We realized that it would be useful to add some content, such as bookmarks, an agenda or links to referenced documents, while participating in a conference.
+We also realized it would be difficult for the same person to do it while performing a presentation.
+As such we devised a method, where the user displays \gls{QR} codes in order to ease content creation.
+
+Figure~\ref{fig:modules} presents the structure of our system, which was divided into six modules:
 
@@
 
-\begin{itemize}
+\paragraph{Application module} responsible for providing information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client} in the form of web pages in \gls{HTML} and \emph{JavaScript} libraries through \gls{HTTP}.
 
-\item \textbf{Application module} - responsible for providing information about the relevant modules (\emph{NAT Traversal} and \emph{Signaling}) and user interface to the \emph{Client} in the form of web pages in \gls{HTML} and \emph{JavaScript} libraries through \gls{HTTP}.
+\paragraph{Signaling module} responsible for \emph{Client} and \emph{Stream} coordination which will be performed using \emph{WebSockets}.
 
-\item \textbf{Signaling module} - responsible for \emph{Client} and \emph{Stream} coordination which will be performed using \emph{WebSockets}.
+\paragraph{NAT Traversal module} \gls{STUN} and \gls{TURN} techniques (part of \gls{ICE}) are used by the \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase that ends with the establishment of the connection between them.
 
- \item \textbf{NAT Traversal module} - \gls{STUN} and \gls{TURN} techniques used by \emph{Client} and \emph{Stream} modules during the \emph{Signaling} phase which ends by establishing the connection between them.
+\paragraph{Stream module} responsible to deliver and receive multimedia content from the \emph{Client} using \gls{WebRTC}.
+This is where audio and video are combined into single streams.
+It is also here that \gls{QR} codes present in the client streams are detected.
 
- \item \textbf{Stream module} - responsible to deliver and receive multimedia content from the \emph{Client} using \gls{WebRTC}.
-
- \item \textbf{Storage module} - provides two main functionalities: store the model information and media recorded.
+\paragraph{Storage module} provides two main functionalities: store the model information and media recorded.
 This is the single module responsible for persistent storage.
@@
 
- \item \textbf{Client module} - responsible for the interaction with the user.
+\paragraph{Client module} responsible for the interaction with the user. Runs on the browser on the user device.
 
-\end{itemize}
+Figure \ref{fig:infrastructure} depicts the architecture of our system, being composed by: web server, streaming server (\gls{KMS}), signaling server (\gls{ICE} server), database (MongoDB) and video repository (Kurento Repository).
 
-%RP Acho que podes escrever muito mais sobre cada componente. Isto está muito curto. Fala de algumas das funções que proporcionam. Vê o que fiz com o Storage Module
-
-%RP também podes fazer um diagrama com vários utilizadores e mostrar quais os streams que vão de e para cada um (mostrar a mistura de som, video, etc).
-
-\subsection{Implementation Proposal}
-The infrastructure is composed by: web server, stream server, signaling server, database and video repository.
-%RP tens da fazer o paralelo entre isto e os componentes apresentados na secção anterior.
-O mesmo com a imagem, que parece não estar relacionada.
-Podes colocar os elementos na mesma posição, ou dentro das caixas da imagem 3.1
-
-\begin{figure}[H]
+\begin{figure}
 	\centering
@@
 	\caption{System Infrastructure}
+        \label{fig:infrastructure}
 \end{figure}
 
-In order to simplify our solution, we propose that the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules be implemented within the same server application, so it could be easy to deploy as a single image.
-To this set of modules, we often call \emph{backend}.
+In order to simplify our solution, the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules are deployed within the same  server, so to facilitate deployment.
+To this set of modules, we call \emph{backend}.
+The \emph{Application} module is split between client and server part.
+The \emph{Signaling} module is implemented together with the application server.
 
-	\subsubsection{Security and authorization}
+(Application) clients can only access the web server, all other servers are in a firewalled private network.
+Access to our streaming servers is given to clients after concluding the signaling phase.
+This signaling phase may or not proceed in function of the client access permissions to the particular group conference.
 
-Having established that the \emph{backend} modules are placed in the same machine, that helps controlling which resources the client has permission to access as those modules are seen as a private network.
-%RP na mesma ``server application'' ou ``machine''?
-São coisa diferentes!
-Apresentas vantages de instalar tudo na mesma máquina mas não falas das desvantagens.
-Menor escalabilidade parece-me ser logo uma delas.
-Na figura 3.2 aparecem várias máquinas!
+With the use of a centralized architecture, each client only has one \emph{PeerConnection} to our streaming server and content shown to them is changed on demand, either being it an individual or a composite view.
+A peer-to-peer approach would have required each peer to maintain a connection to each of the other participants an one to the streaming server, for recording.
+This would seriously limit the size of each group conference by placing high bandwidth and processing requirements on the clients. 
+The same can be concluded about instant message delivery.
+Each client has only one \emph{WebSocket} connection to the application server which consequently relays the messages to other users.
 
-To qualify the above we provide public access to \gls{HTTP} server ports, maintaining the access to other components restricted through firewall rules.
-%RP qualify?
-%RP Imagem com firewall e estes portos? Mas isto já me parece muito detalhe de implementação.
-In relation to the database there is no directly access from the outside. All the database information is accessed via our application server which validates the permissions of users on our system.
+Our architecture can be scaled by having conference rooms distributed across different servers.
+However, all of a group conference participants must be connected to the same server.
 
-On the other hand, the access to our streaming servers is also restricted, but clients can connect to them after concluding the signaling phase.
-This signaling phase may or not proceed in function of the client access permissions.
-For example, if a user is trying to access a private conference room that he is not a member of nor has an invitation link for, the signaling server refuses to start the signaling phase and the user cannot access the streaming server.
+\section{Implementation}
+\label{chapter:implementation}
 
-The placement of our streaming servers inside a \gls{NAT} also has an important role with respect to external misuse prevention.
-Otherwise, placing our streaming servers could allow external clients to perform their own signaling protocol and, as a consequence, use our infrastructure without our consent.
+Our prototype was built using \gls{KMS} as our media streamer, as \gls{KMS} supports \gls{WebRTC}.
+The server side of the \emph{Application} and \emph{Signaling} modules are implemented as a server side web application. 
+Our server could be implemented with JavaScript (e.g. with \emph{NodeJS}) or \emph{Java} due to the fact \gls{KMS} provides clients for both technologies. 
+We decided to implement our web server using the \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java}.
+For the collaborative text editor, we resorted to \emph{OT.js}\footnote{\url{http://operational-transformation.github.io}(accessed March 10, 2016)} due to its server and storage implementation choice independence.
 
-\subsubsection{Client connections}
-	Although the delegation of processing work to clients can improve our system's scalability, we are concerned about using the least resources possible on the client side, as huge resource consumption may drain battery very fast or may even be impossible to run on mobile devices.
-We are aware that streaming video from clients is already a very intensive task which we cannot avoid but can improve by delegating the most intensive tasks to our servers.
+By default, \emph{Kurento Repository}, used for storing the recorded video and audio streams, is implemented over \emph{MongoDB}.
+For convenience, our storage model also uses the same database.
+In order to offer all the functionality, some information about objects must be persistent such as users, groups, relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.
 
-	In this context, with a centralized approach, each client must only have one \emph{PeerConnection} to our streaming server and content shown to them is changed on demand either being it an individual or a composite view.
-Otherwise clients could follow a peer-to-peer connection which would result on maintaining more connections and performing the composition of videos on client side.
-
-	The composition of streams on client side is performed by receiving streams with the best quality possible but, due to undersizing the video of clients into a smaller region, this would result on wasting bandwidth on a quality that is not needed.
-
-	Although the peer-to-peer approach could be used on our system, we conclude that we need to record the video on our streaming server because web clients have a very limited storage and peer disconnections may result on recorded video loss.
-
-	The same can be concluded to instant message delivery, each client must have only one \emph{WebSocket} connection to the application server which consequently relays the messages to other users.
-
-	Relaying instant messages from clients through the our web servers is easier if all clients are connected to the same server because all messages can be directly delivered without sending messages across multiple web servers.
-
-	In the context of this thesis, we will not implement sending messages across web servers but, in order to allow our system to scale, we will consider that all conference participants are connected to the same server and our system is scaled by having conference rooms distributed across different servers.
-
-\subsubsection{Software choices}
-
-Furthermore, we have taken into account the compatibility between the streaming server, database and the operation transformation solutions, in order choose the appropriate framework to implement our web server.
-%RP esta secção devia começar aqui, a explicar a escolha das tecnologias/produtos a usar em cada componente.
-O que tens antes são pormenores que só faz sentido discutir no final.
-Tens de ir sempre do mais geral para o mais detalhado.
-
-We have decided that our solution must use \gls{KMS}.
-Our web server could be implemented easily with \emph{NodeJS} or \emph{Java} due to the fact \gls{KMS} provides clients for both technologies.
-But others could also be used as \gls{KMS} also exposes their \gls{API} via \emph{WebSockets}.
-
-Due to the fact we are going to use \gls{KMS} as streaming server solution, we could choose \emph{NodeJS} or any \emph{Java} based web framework for implementing our web application server.
-We have decided to implement our web server with the \emph{PlayFramework}\footnote{\url{https://www.playframework.com/}(acessed March 25, 2016)} using \emph{Java} because of our previous experience with it.
-
-By default, \emph{Kurento Repository} is implemented over \emph{MongoDB}, for convenience our storage model will also use the same database.
-
-Importantly, for the collaborative text editor, we have chosen \emph{OT.js} due to its server and storage implementation choice independence.
-
-For the \emph{NAT Traversal} module a public \gls{STUN} server can be used for testing our solution.
+For the \emph{NAT Traversal} module a public \gls{STUN} server was used for testing our solution.
 Nevertheless, we recognize that for a production environment we would need to maintain our own \gls{TURN} servers in order to ensure connectivity to all clients.
 
-Not less important, on the client computers, both \emph{Mozilla Firefox} and \emph{Google Chrome} could be installed as web browsers.
-As such, both should be supported.
-Libraries such as \emph{jQuery}, \emph{Bootstrap}, \emph{Adapter.js}, \emph{OT.js} can be downloaded from the web server and executed on the client side using any of these two browsers.
-%RP relembrar que o IE e o Safari não suportam webrtc?
+Table~\ref{table:apparch} presents the client application architecture and the underlying technologies used.
+Both \emph{Mozilla Firefox} and \emph{Google Chrome} browsers are supported.
+\emph{Adapter.js} and \emph{jQuery} ensure that our application is compatible with the most popular web browsers.
+\emph{Bootstrap} is used to make the user interface more appellative and responsive, adapting to mobile devices with different screen sizes.
+\emph{OT.js} is used to build the collaborative text editor.
 
-\begin{table}[H]
+\begin{table}
 \centering
@@
 
-Table~\ref{table:apparch} presents the application architecture and the underlying technologies seen from the user's perspective.
-\emph{Adapter.js} and \emph{jQuery} will ensure that our application is compatible with the most popular web browsers.
-\emph{Bootstrap} will be used to make the user interface more appellative and responsive.
-With \emph{Bootstrap} it is quite easy to develop applications that adapt to mobile devices with different screen sizes.
-%RP bootstrap não aparece na figura!
+Interaction and synchronization between the streaming server, the application server and the clients is achieved using a signaling protocol.
+After the web application server validates the user access, the client creates a WebSocket to exchange signaling information with that server. 
+Our signaling protocol consists of sending and receiving \gls{JSON} formated messages over that \emph{WebSockets} by both the application server and the client.
 
-With respect to displaying content, the synchronization between multimedia elements will be performed through chains of \emph{JavasSript} events or by specifying the interval of time which time content must be visible.
-Other animations can be implemented with \gls{SVG} embedded on \gls{HTML}.
-%RP Caiu do céu! Tens de explicar e ligar melhor a sequencia de texto
-
-\section{Implementation}
-
-\subsection{Data Model}
-
-The data model is a critical component of our solution, as a badly designed model can imply serious difficulties when implementing new features that are not part of the plans.
-During the course of this project, we had to redesign the model more than once in order to support new features.
-
-In order to offer all the functionalities that we promise, some information about objects must be persistent such as users, groups,relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.
-For designing our model we have taken into account generic programming techniques.
-We observed that operations like searching for an object were quite repeated across different types of objects.
-
-\subsection{Signaling Protocol}
-
-Although we have mentioned that the signaling protocol is used to establish connections between peers, on our system our media server (\gls{KMS}) is a peer that receives video streams and sends to its connected clients.
-
-After the web application server validates the user access, the signaling protocol allows the users to directly connect to the \gls{KMS}, which is placed in a private network, and lets the application server and users negotiate media types and encoding information to use during the conversation.
-
-Our signaling protocol consists of sending and receiving \gls{JSON} formated messages over \emph{WebSockets} by both the application server and the client.
-
-When a user enters a group conference, after the page is completely loaded, a WebSocket is created to maintain a connection with our web servers. 
-But before creating the web socket, we must identify the user and check if he has permissions to participate in the conference.
-The user identification is done by retrieving the session id from the cookie provided by the user-agent (web browser) through the \gls{HTTP} headers.
-%RP não é bem o user que fornece o cookie. é o user-agent (browser). Convém ser preciso.
-%HR done
 The Web application server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room. 
 It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake is performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \gls{HTTP} context is lost.
-%RP socket -> websocket?
 
 At this stage, the web application's user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server. 
+If the user decides to share his camera or screen the user agent creates an offer, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the WebSocket to the Application Server.
+From a \gls{WebRTC} point of view, our media server (\gls{KMS}) is a peer that receives video streams and sends it to its connected peers (the clients).
 
-If the user decides to share his camera or screen the user agent creates an offer, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the WebSocket to the Application Server.
+Figure \ref{fig:signaling2} shows the sequence of messages, intermediated by the server, by which both client and \gls{KMS} determine and exchange \gls{ICE} candidates for establishing a direct connection despite \gls{NAT} firewalls.
 
-The server receives and processes the offer and sets the \emph{remote session description} to its client associated \gls{WebRTC} endpoint.
-Then a \emph{local session description} is created on the server and sent back to the client.
-After that, the server tries to gather \gls{ICE} candidates.
+%The server receives and processes the offer and sets the \emph{remote session description} to its client associated \gls{WebRTC} endpoint.
+%Then a \emph{local session description} is created on the server and sent back to the client.
+%After that, the server tries to gather \gls{ICE} candidates.
 
-The client receives the server answer, sets the \emph{remote session description} and gets the ice candidates from the \gls{ICE} server.
-%RP tens ``ice'', ``\gls{ICE}''. Acho que também já vi ``ICE''. Convém ser sempre igual!
+%The client receives the server answer, sets the \emph{remote session description} and gets the \gls{ICE} candidates from the \gls{ICE} server.
+%Subsequently, both the server and client receive the \gls{ICE} candidates that allow the client to connect directly to \gls{KMS} and vice-versa. 
+%The candidates are received at the client which sets them to its \emph{PeerConnection}. 
+%The same is done on the server which receives the \gls{ICE} candidates from the client and propagates them to \gls{KMS}.
 
-Subsequently, after a while both the server and client receive the \gls{ICE} candidates that allow the client to connect directly to \gls{KMS} and vice-versa.
-The candidates are received at the client which sets them to its \emph{PeerConnection}.
-The same is done on the server which receives the \gls{ICE} candidates from the client and propagates them to \gls{KMS}.
+%An \gls{ICE} candidate contains an \gls{IP}, port, used transport protocol and an attribute named \emph{sdpMLineIndex} that is used for mapping to the \emph{remote session description} media type.
+%When a connection is established, the user and server start to interchange stream data but other \gls{ICE} candidates may arrive with better connections.
 
-An \gls{ICE} candidate contains an \gls{IP}, port, used transport protocol and an attribute named \emph{sdpMLineIndex} that is used for mapping to the \emph{remote session description} media type.
-When a connection is established, the user and server start to interchange stream data but other \gls{ICE} candidates may arrive with better connections.
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/signaling}
+    \caption{Signaling sequence diagram}
+    \label{fig:signaling2}
+\end{figure}
 
-Having the media session established, the server starts to record any received stream and the client creates an \gls{URL} correspondent to the stream location.
-
-\subsection{Stream Recording}
- Initially we experimented with local recording and synchronizing with our servers. Even thought it got it to work, it consumed to much bandwidth. We then tried recording on server side.
-
-	One of our concerns during the development of our solution was the storage scalability.
-Saving files directly into the file system would require an extra effort to distribute and replicate files among servers.
-For that purpose, the \emph{Kurento} team developed \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} which is based on \emph{MongoDB}.
+Having the media session established, \gls{KMS} starts to record any received stream and the client creates an \gls{URL} correspondent to the stream location.
+It also combines all the streams into a single one, which is also recorded.
+Figure \ref{fig:devices} shows a combined stream with two different views of the same user with two cameras.
+%One of our concerns was the storage scalability. 
+%Saving files directly into the file system would require an extra effort to distribute and replicate files among servers.
+\emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} provides a scalable, replicated storage solution for streams, based on \emph{MongoDB}.
+Streams are saved in 10s blocks.
 
@@
 It is \gls{KMS} that sends different content through that stream.
-When a user desires to play recorded video, a \emph{webSocket} message is sent specifying the time and the intended user id.
+When a user desires to play recorded video, he uses the time bar, shown in Figure \ref{fig:timeline}, to select a bookmark or simply drags it backwards.
+In response, the client uses the \emph{webSocket} to send a message specifying the time and the intended user id.
 The server performs the calculations in order to find a block that intersects the requested time, plays it and when finished, the next part is automatically played without the user intervention.
 
-%RP não falas nada sobre chats, etc! Não há signalling para isso?
-%HR não, só video e audio, chat é websockets
+	\begin{figure}
+		\centering
+		\includegraphics[width=\linewidth]{figures/devices.png}
+		\caption{Multiple devices per user}
+		\label{fig:devices}
+	\end{figure}
 
-\subsection{Hyper-Content}
+	\begin{figure}
+		\centering
+		\includegraphics[width=\linewidth]{figures/timeline.png}
+		\caption{Interactive timeline}
+		\label{fig:timeline}
+	\end{figure}
 
-	Our system supports creating superimposed content to video, which is achieved by creating \gls{HTML} tags on top of the video  with the same size.
-The decision of which content must be displayed to each user is performed by our content scheduler which uses the user's current time in order to synchronize which content is shown or removed from the user interface.
+Our system supports creating content superimposed on the video.
+This is achieved by creating \gls{HTML} tags on top of the video, with the same size.
+The decision of which content to be displayed to each user is performed by our content scheduler, that uses the user's current time in order to synchronize which content is shown or removed from the user interface.
 
@@
 
-	As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.	In order to help the content creator to introduce and synchronize its content in real time, we allow the user to encode its content into \gls{QR} codes and show it to the camera in real time.
+As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.
+In order to help the content creator to introduce and synchronize its content in real time, we allow the user to encode its content into \gls{QR} codes and show it to the camera in real time.
+The \gls{QR} code processing is performed at \gls{KMS}.
 
-\subsection{Collaborative editor}
-
-Our collaborative editor is a simple text editor that is synchronized with all participants within a conference room implemented using \emph{ot.js}.
-
+Our collaborative editor is a simple text editor, that is synchronized with all participants within a conference room, implemented using \emph{ot.js}.
 The state of our collaborative editor is not saved on the database every time it changes.
@@
 \section{Evaluation}
+\label{chapter:evaluation}
 
-\subsection{Tests Objectives}
+We tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.
+We have also tested the performance of our solution by measuring the used resources.
+Those performance tests are crucial to ensure that our solution is in fact stable and scalable.
 
-  We have tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.
+We evaluated our system by running all the backend components as Docker containers on a Debian 8 machine with 128GB of RAM, 2 Intel Xeon E5-2640V2 CPUs, hybrid SSD, HDD storage and 1 Gbps Ethernet. 
+We implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects, every second, CPU and physical memory information relative to each running process and network usage relative to each interface.
 
-  We have also tested the performance of our solution by measuring the used resources.
-Those performance tests are crucial to ensure that our solution is in fact stable and users can use it endlessly without decreasing the quality of their experience.
+The performance test scenario consists of having 7 users join the conference room sequentially and then leaving it, one at a time.
+Each event, occurs with intervals of one minute, for a total of thirteen minutes (780 seconds).
 
-  \subsection {Performance Tests}
-
-      In order to benchmark our system, we have implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects with a periodicity of one second: CPU, physical memory information relative to each running process and network usage relative to each interface.
-
-      The performance test scenario that we have defined consists on two phases, the first phase consists only on having users, with similar computer and network specifications, entering sequentially on the conference room.
-The second phase consists on the users leaving the conference room.
-Each event, joining and leaving, occurs with intervals of one minute in total of thirteen minutes (780 seconds).
-
-      From the media server perspective if there are $n$ clients connected, each of them sending and receiving one stream, it is expected that server sends and receives also $n$ streams.
+From the media server perspective, if there are $n$ clients connected, each of them sending and receiving one stream, it is expected that server sends and receives also $n$ streams.
 As such, we expect that the amount of network traffic increases linearly as users join a conference room.
@@
   \centering
-  \includegraphics[width=\linewidth]{stats/test_full_features_net.eps}
-  \caption{Network usage after implementing all features}
+  %\includegraphics[width=\linewidth]{stats/test_full_features_net.eps}
+\includegraphics[width=\linewidth]{net_usage.png}
+  \caption{Server network usage}
   \label{fig:test_full_features_net}
@@
 
-      The blue peaks are caused by the signaling phase and web page downloads, including resources such as images, stylesheets and javascript files. 
-      The green peaks are caused by video and audio being transfered between \gls{KMS} and \emph{MongoDB} through \emph{Kurento Repository}.
-Each peak occurs every time a block of video is recorded, which in this case is every ten seconds.
-      The recordings are synchronized so all user and mixed blocks start and end at the same time.
-That is why the amount of work done every ten seconds accumulates, and because this is performed locally, the maximum transfer rate is limited by the performance of the memory as buffers are written to buffers then to disks.
-      Sent data transfer rate has no significant peaks as \gls{HTTP} requests and signaling information contains little information.
+The blue peaks are caused by the signaling phase and web page download when a new client joins.
+The green peaks inter-container traffic, caused by video and audio being transfered between \gls{KMS} and \emph{MongoDB} through \emph{Kurento Repository}.
+Peaks occurs every 10s, when a block of video is recorded.
+The recordings are synchronized, so all user and mixed blocks start and end at the same time.
+%That is why the amount of work done every ten seconds accumulates, and because this is performed locally, the maximum transfer rate is limited by the performance of the memory as buffers are written to buffers then to disks.
+Sent data transfer rate has no significant peaks as \gls{HTTP} requests and signaling information contains little information, being mostly made up of video streams sent to the clients.
+As expected, send and received traffic grows linearly with the number of users.
 
-	With this results we conclude that if we want to scale our storage solution using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \gls{KMS} should be installed on the same machine because the loopback interface can handle bigger transfer rates than the remaining network interfaces.
+%candidato a apagar
+%With this results we conclude that if we want to scale our storage solution using the \emph{MongoDB}'s cluster configuration, both \emph{Kurento Repository} and \gls{KMS} should be installed on the same machine because the loopback interface can handle bigger transfer rates than the remaining network interfaces.
 Installing the repository on the same machine as a database node does not ensure that recorded videos are stored in the same machine, for this reason we would prefer installing \gls{KMS} and \emph{Kurento Repository} on the same machine.
 
-On the other hand, Figure \ref{fig:test_client_net} shows the respective network usage on the client side during our test case.
-We observe that in the first seconds the client adjusts the video quality it sends to \gls{KMS}.
+Figure \ref{fig:test_client_net} shows the same events from the respective of network usage on the client side.
+We observe that in the first seconds the client adjusts the video quality it sends to \gls{KMS} and maintains that upload rate throughout.
 Whenever a new client enters the conference, we observe that \gls{KMS} decreases the video quality in order to instantaneously integrate a new user into the conference room.
@@
 When a user leaves the conference room \gls{KMS} has no need to decrease the participant's video quality as less network bandwidth will be used.
+As expected, the maximum bandwidth used is constant and independent of the number of participants.
 
@@
   \includegraphics[width=\linewidth]{stats/test_client_net.eps}
-  \caption{Client Network usage during our test case}
+  \caption{Client network usage}
   \label{fig:test_client_net}
@@
 
-\textbf{Memory usage}
-
 Figure \ref{fig:test_ram_fixed_mem} shows the memory usage during our performance test.
-Both \gls{JVM}, \emph{MongoDB} and \gls{KMS} performs their own memory management by holding and recycling objects when needed.
+\gls{JVM}, \emph{MongoDB} and \gls{KMS} each perform their own memory management by holding and recycling objects when needed.
 The expected and observed behavior of the memory usage is growth of memory usage while the users are entering the conference room and a memory usage stabilization afterwards.
@@
   \includegraphics[width=\linewidth]{stats/test_ram_fixed_mem.eps}
-  \caption{Memory usage after fixing recorder memory leak}
+  \caption{Server memory usage}
   \label{fig:test_ram_fixed_mem}
@@
 \emph{MongoDB} checkpoints data to disk every 60 seconds or when journal data exceeds 2GB\footnote{\url{https://docs.mongodb.org/manual/faq/storage/}(Accessed March 28, 2016)}, that explains the small memory usage peaks during our test case.
-When the conference room is empty there are no video recordings, which explains the memory stabilization at the end.
-
+When the conference room is empty, there are no video recordings, which explains the memory stabilization at the end.
 \emph{KMS} memory management released resources as soon as users left the room.
 
-\textbf{CPU usage}
-
 Figure \ref{fig:test_full_features_cpu} shows the percentage of \gls{CPU} usage during our performance test case. 
-Each 100\% represents one \gls{CPU} core, although that does not mean one \gls{CPU} is fully used, for example two cores at 60\% represent 120\% \gls{CPU} usage.
-As we can see, the percentage of \gls{CPU} used increases and decreases linearly in function of the amount of conference participants.
-\gls{KMS} is responsible for most \gls{CPU} usage.
+Each 100\% represents one \gls{CPU} core, although that does not mean one \gls{CPU} is fully used, as e.g. two cores at 60\% represent 120\% \gls{CPU} usage.
+As we can see, the percentage of \gls{CPU} used increases and decreases linearly in function of the number of conference participants.
+As expected due to the video decoding and enconding and \gls{QR} code detection, \gls{KMS} is responsible for most of the \gls{CPU} usage.
 
@@
 
-  Just for testing purposes, we performed the same performance tests disabling \gls{QR} codes detection.
-We concluded that \gls{QR} code detection is a very intensive task, approximately doubling the amount of work performed by the \gls{CPU}s.
+We conclude that our solution's bottleneck is the \gls{CPU} usage by \gls{KMS}.
+We repeated the same test with \gls{QR} code detection disabled and verified that \gls{CPU} usage dropped to half.
 
-	Even though, with this test results, we conclude that our solution's bottleneck is the \gls{CPU} usage at \gls{KMS}.
+In order to evaluate the usability of our solution, we performed usability tests with the help of 20 users with different backgrounds and ages between 22 and 38 years. 
+Users were given a guide with five tasks to perform.
+%The metrics we used were: number of clicks, number of errors and time spent. 
+As a baseline, we used experienced users in order to retrieve the optimal task duration.
 
-\subsection {Usability Tests}
-     In this section we describe usability test scenarios that we have applied and their respective results.
+From the data collected with twenty tests with users, we have calculated the confidence intervals in order to understand the most plausible values for each metric.
+As a result of both true average and variance being unknown and the usage of a relatively small amount of samples, we had to use the \emph{t-distribution} to estimate our metrics confidence intervals.
+We have used a 95\% confidence level.
 
-    \subsubsection{Tests Scenarios}
-
-      In order to evaluate the usability of our solution, we have performed usability tests with the help of real users with different backgrounds and ages.
-
-      We handed a guide to the users with five tasks to perform. The metrics we used for each task were: number of clicks, number of errors (including a description) and time spent.
-
-  \subsubsection{Test Results}
-
-In this section we present the results of our usability tests.
-The first time we tested our solution by providing tasks to users, we have observed that our solution was not perfect.
-Having faced usability problems during our tests, he had to improve our solution's usability and start the tests again.
-
-In the first testing phase we have performed tests with just three users and ceased for improvements. We gathered comments and suggestions after letting the users explore our system.
-
-On the second phase of our user interface tests, in a general way, we have noticed great improvements on the learning time.
-
-In order to measure the users learning speed, we have performed tests with experienced users in order to retrieve the optimal task duration and minimal task clicks.
-
-To this end, with regard to optimal task duration and minimal clicks, we obtained the values shown in Table \ref{table:optimal}.
-
-\begin{table}[H]
-\centering
-\caption{Metrics for an experienced user}
-\label{table:optimal}
-\begin{tabular}{|c|c|c|c|c|c|}
-\hline
-\textbf{Task} & 1 & 2 & 3 & 4 & 5 \\ \hline
-\textbf{Duration (seconds)} & 20 & 35 & 30 & 25 & 35 \\ \hline
-\end{tabular}
-\end{table}
-
-From the data collected with twenty tests with users, namely the task duration (Figure \ref{fig:user_times}) and task difficulty (Figure \ref{fig:user_diffs}), we have calculated the confidence intervals in order to understand the most plausible values for each metric.
-
-The youngest and oldest testers were, receptively, twenty two and thirty eight years old. Figure \ref{fig:user_ages} the ages of the users that tested our system.
+Task duration, shown in Figure \ref{fig:user_times}, as expected, was larger than that of an experienced user.
+Curiously, there is no correlation between time taken by experienced and novice users.
 
@@
   \centering
-  \includegraphics[width=0.8\linewidth]{stats/user_ages.eps}
-  \caption{Ages of the users that tested our system}
-  \label{fig:user_ages}
-\end{figure}
-
-As a result of both true average and variance being unknown and the usage of a relatively small amount of samples, we had to use \emph{t-distribution} to estimate our metrics confidence intervals.
-We have used a 95\% confidence level.
-
-\begin{figure}
-  \centering
-    \includegraphics[width=0.8\linewidth]{stats/user_times.eps}
+    \includegraphics[width=0.75\linewidth]{stats/user_times.eps}
     \caption{Time spent per task}
@@
 
-According to Figure \ref{fig:user_diffs}, we can observe that most users had less difficulties with the first two tasks, which represents types of tasks that most users are familiar with.
-As soon as users had to navigate in time, manipulate annotations and create content (respectively \emph{task3}, \emph{task4} and \emph{task5}) we observed that they revealed more difficulties.
+In Figure \ref{fig:user_diffs}, we can observe that most users had less difficulties with the task one (login in, accept and create friends) and two (create conference, share screen and use text editor), which represents types of tasks that most users are familiar with.
+New concepts, such as navigate in time, manipulate annotations and create content (respectively \emph{task3}, \emph{task4} and \emph{task5}) turned out to be more difficult.
  Most of those difficulties, based on the users feedback, were mainly due to those concepts not being familiar to them.
@@
   \centering
-    \includegraphics[width=0.8\linewidth]{stats/user_diffs.eps}
+    \includegraphics[width=0.75\linewidth]{stats/user_diffs.eps}
   \caption{Difficulty per task}
@@
 As we had relatively bad results with some users, we explained to those users that could not conclude the tasks or performed them incorrectly, the most efficient way to perform the requested tasks.
-Some users suggested to display more hints in order to achieve a faster learning, but afterwards all were impressed and gave us a better evaluation.
+Some users suggested to display more hints in order to achieve a faster learning.
+Afterwards, all said they were impressed with the new functionality.
 
@@
 
-Another weak aspect was our content editor, which, in fact, we recognize is difficult to work with, mostly due to the amount of information that is necessary to create a synchronized content (starting time, duration and content itself).
+Another weak aspect was our content editor, which, in fact, we recognize is difficult to work with, mostly due to the amount of information that is necessary to create a synchronized content (starting time, duration and the content itself).
 Some users have suggested that the content should also be present on the timeline so they could be easily dragged and resized (on time).
-
-We are aware that placing content on the timeline will reduce our solutions performance, especially when there is a relatively large amount of content, due to the content that is present on the timeline being loaded all at once.
-Although, we recognize that for some cases (relatively low amount of content), displaying the content on the timeline could not have a great impact on our solutions performance, we choose not to implement this.
 
@@
 \section{Conclusions}
-\subsection{Achievements}
-\label{section:achievements}
+\label{chapter:conclusion}
 
-	We have successfully implemented the basic functionalities of our prototype and spare some time for adding more valuable features such as the ability to create content by exposing \gls{QR} codes to the camera and lastly perform changes to the user interface in order to improve the quality of user's experience.
+We have successfully implemented a collaboration application that provides video-conferencing, chat, document editing and file sharing, with a novel concept: the ability to navigate back and forth in time, enabling non-linear visualization of video-conferencing.
+This also makes it possible for users to enrich the video-content with hyper-media after it is created.
 
-	The performance tests that we have executed showed that our system is stable and more importantly that our web server is lightweight and most of processing power is dedicated to the streaming server.
+The performance tests that we have executed showed that our system is stable and more importantly, that our application server is lightweight and most of the processing power is dedicated to the streaming server.
 
-	Our usability tests show results that are considerably worse than the established optimal values due to our solution propose a different way to communicate that most people are not used to.
-Although we have obtained those results, in general our users gave us positive feedback and valuable advices which we have used to improve our system.
+Our usability tests showed that users had difficulty using this new concept, even though they found it useful and were willing to use it in the future.
+This suggests that the user interface has to be improved, possibly by reducing some of the functionality in order to simplify it.
 
-\subsection{Future Work}
-\label{section:future}
-	Playing back video with a faster rate is not possible using the current version of \gls{KMS}.
-Even though we expect the availability of that feature in a near future but we have proposed an alternative way to implement faster playback by using \emph{ffmpeg} to convert the video before playing it.
+We are also aware of other limitation that would prevent our prototype from being deployed on a wide scale.
+We have purposely chosen functionality over security when allowing users to provide code that is executed by other users.
+A balance can be stroke by liming the flexibility afforded to users.
 
-	Although we have tested our solution in a powerful machine for the current time, our performance tests revealed that the streaming component uses a lot of resources.
-We left for a future work a deep analysis on the scalability of our system for which we have proposed different approaches but we have not tested them.
+Our performance tests revealed that the streaming component uses a lot of resources, increasing the costs of a large deployment.
+We left for a future work a deep analysis of the scalability of our system.
 
-	Another aspect we could have tested was the performance of our solution when using \gls{TURN} servers for relaying the traffic that fails using \gls{STUN}.
+A useful feature when reviewing past video, is playback at a faster rate.
+This is not possible using the current version of \gls{KMS}.
+Even though we expect the availability of that feature in a near future, an alternative way to implement faster playback would be to use an external application, such as \emph{ffmpeg} to convert the video before playing it.
 
-	Lastly, we have chosen functionality over security in respect to displaying content to users which lead to security flaws on our solution.
-Although we have not solved the security problems, we have proposed a solution which at the same time limits the flexibility of adding new functionalities to our prototype.
+% trigger a \newpage just before the given reference
+% number - used to balance the columns on the last page
+% adjust value as needed - may need to be readjusted if
+% the document is modified later
+%\IEEEtriggeratref{8}
+% The "triggered" command can be changed if desired:
+%\IEEEtriggercmd{\enlargethispage{-5in}}
 
-%Sets the bibliography style to UNSRT and imports the 
-%bibliography file "samples.bib".
-%\bibliographystyle{unsrt}
-%\bibliography{references}
+\bibliographystyle{IEEEtran}
+\bibliography{../tese2/references}
 
-%\input{thesis.bbl}
-\printbibliography
 \end{document}

### 2017-02-20 artigo/paper.tex
Added finantial acknowledgments
Shrinked some images and tables

@@
 	\hline 
-	\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
-	\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[12pt]\hline
-	\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
+	\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[2pt]\cline{1-5}\cline{7-7}
+	\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[2pt]\hline
+	\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[2pt]\hline
 	\end{tabular}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/signaling}
+    \includegraphics[width=\linewidth]{figures/signaling2}
     \caption{Signaling sequence diagram}
@@
 		\centering
-		\includegraphics[width=\linewidth]{figures/devices.png}
+		\includegraphics[width=\linewidth]{figures/devices2.png}
 		\caption{Multiple devices per user}
@@
 
+\section*{Acknowledgments}
+
+This work was partially supported by national funds through Fundação para a Ciência e a Tecnologia (FCT) with reference UID/CEC/50021/2013.
+This work has received funding from the European Union’s Horizon 2020 research and innovation program under grant agreement No 645342; project reTHINK.
+
 % trigger a \newpage just before the given reference
@@
 \end{document}
+
+\grid

### 2017-02-21 artigo/paper.tex
Shrinked some images and tables
reviewed paper

@@
 As communications technologies evolved, we adapted the way we communicate. 
-Multi-party video conference is today a commonly used tool in the enterprise, academia and personal live.
+Multi-party video conference is today a commonly used tool in the enterprise, academia and individually.
 The \gls{WWW} has made us familiar with the concept of text hyper-media, in the form of hyper-links.
@@
 
-  \gls{SMIL}~\cite{smil} was introduced to describe temporal behavior of multimedia content, in particular, it could be used to overlay subtitles on films. 
-With \gls{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements with hyper-links, among multiple other features.
+  \gls{SMIL}~\cite{smil} was introduced to describe temporal behavior of multimedia content%, in particular, it could be used to overlay subtitles on films%
+  . 
+With \gls{SMIL} it is possible to synchronize multiple videos, either in parallel or in sequence, reproduce a different audio track, overlay user interface elements like subtitles or hyper-links, among multiple other features.
   %RP synchronize multiple sections of video? vários videos apresentados em simultâneo?
@@
 In order to meet user expectation, a group conference provides functionality similar to those found in established applications such as \emph{Skype} or \emph{Google Hangouts}.
-Among other common features, one can view a small video of each participant or a large video of a selected participant.
+Among other common features, one can view a video composed by all participants or a video of a selected participant.
 Each participant may use more than one camera, in order to show different angles or objects, e.g. a white board and himself.
-The source of the large video can be selected manually or automatically assigned to the participant talking.
+The source of the video can be selected manually or automatically by the talking participant.
 Participating users may choose to view only, share their microphone, camera or screen.
@@
 The user is free to slide over to any instant in the past or back to real-time.
-Changing the instant of time updates the state of not only the video but of other collaboration elements.
+Changing the instant of time updates the state of not only the video but of other collaboration elements. % ("other collaboration elements", nao foi implementado, apenas para conteudo sobreposto...) %
 
 Either in real-time or later, users can create bookmarks, displayed in the time-line, for easier navigation in time.
-Users can also add content to superimpose on the video.
-This content can be as simple as subtitles or as complex as a \gls{HTML} document or JavaScript code to perform an animation.
+Users can also superimpose content to video.
+This content can be as simple as subtitles or as complex as a \gls{HTML} document or JavaScript code to perform an animation or interaction.
 All the content is searchable, allowing for time navigation by content.
@@
 
+\gls{SFU} is a technique that consists on having a server relaying each stream, as it is, to other users. Leaving the stream selection and composition operations to the user.
+
 In order to limit the processing requirement in the client application (browser), we decided to use a server to combine the audio and video streams from all the participants into a single video and a single audio track.
+This techinique is called \gls{MCU}.
 This also helps conserve battery on mobile devices.
@@
 	\centering
-	\includegraphics[width=\linewidth]{figures/modules.pdf}
+	\includegraphics[width=\linewidth]{figures/modules2.png}
 	\caption{System Modules}
@@
 It stores user and communication data as well as all the data required among user sessions.
-It is also use to store all the communication streams, so that they can be viewed later.
+It is used as well to store all the communication streams, so that they can be viewed later.
 
-\paragraph{Client module} responsible for the interaction with the user. Runs on the browser on the user device.
+\paragraph{Client module} responsible for the interaction with the user. It runs on the user device's browser.
 
@@
 With the use of a centralized architecture, each client only has one \emph{PeerConnection} to our streaming server and content shown to them is changed on demand, either being it an individual or a composite view.
-A peer-to-peer approach would have required each peer to maintain a connection to each of the other participants an one to the streaming server, for recording.
+A peer-to-peer approach would have required each peer to maintain a connection to each of the other participants and one to the streaming server, for recording.
 This would seriously limit the size of each group conference by placing high bandwidth and processing requirements on the clients. 
@@
 For convenience, our storage model also uses the same database.
-In order to offer all the functionality, some information about objects must be persistent such as users, groups, relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.
+In order to offer all the functionality, some information about objects must be stored on database, such as users, groups, relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.
 
-For the \emph{NAT Traversal} module a public \gls{STUN} server was used for testing our solution.
+% persistent -> stored on database, reviewer2 confundiu persistent com fixed
+
+For the \emph{NAT Traversal} module, a public \gls{STUN} server was used for testing our solution.
 Nevertheless, we recognize that for a production environment we would need to maintain our own \gls{TURN} servers in order to ensure connectivity to all clients.
@@
 \emph{Bootstrap} is used to make the user interface more appellative and responsive, adapting to mobile devices with different screen sizes.
-\emph{OT.js} is used to build the collaborative text editor.
+\emph{OT.js} is used to synchronize the collaborative text editor.
 
@@
 The Web application server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room. 
-It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake is performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \gls{HTTP} context is lost.
+It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \gls{HTTP} context is lost.
 
 At this stage, the web application's user is asked if he wants to share its camera and microphone, share screen or just receive streams from the server. 
-If the user decides to share his camera or screen the user agent creates an offer, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the WebSocket to the Application Server.
+If the user decides to share his camera or screen the user agent creates an offer, sets a \emph{local session description} to its \emph{PeerConnection} and sends it through the \emph{WebSocket} to the Application Server.
 From a \gls{WebRTC} point of view, our media server (\gls{KMS}) is a peer that receives video streams and sends it to its connected peers (the clients).
@@
 
-Having the media session established, \gls{KMS} starts to record any received stream and the client creates an \gls{URL} correspondent to the stream location.
-It also combines all the streams into a single one, which is also recorded.
+Having the media session established, \gls{KMS} starts to record any received stream, combines all the streams into a single one (which is also recorded) and the client creates an \gls{URL} correspondent to the stream location.
 Figure \ref{fig:devices} shows a combined stream with two different views of the same user with two cameras.
@@
 \emph{Kurento Repository}\footnote{\url{http://doc-kurento-repository.readthedocs.org} (accessed on 17 March 2016)} provides a scalable, replicated storage solution for streams, based on \emph{MongoDB}.
-Streams are saved in 10s blocks.
+Streams are saved into blocks of 10 seconds.
 
-With server side recording, the user would maintain always the same stream \gls{URL} even if it is playing real time video or reproducing recorded video. 
+With server side recording, the user keeps the same stream \gls{URL} even if it is playing real time video or reproducing recorded video. 
 It is \gls{KMS} that sends different content through that stream.
 When a user desires to play recorded video, he uses the time bar, shown in Figure \ref{fig:timeline}, to select a bookmark or simply drags it backwards.
-In response, the client uses the \emph{webSocket} to send a message specifying the time and the intended user id.
+In response, the client uses the \emph{WebSocket} to send a message specifying the time and the intended user id.
 The server performs the calculations in order to find a block that intersects the requested time, plays it and when finished, the next part is automatically played without the user intervention.
@@
 
-As manually content insertion is a laborious task that can be realized after the video is recorder, we provide an alternative mechanism for real time introduction of superimposed content.
+As manually content insertion is a laborious task that can be realized after the video is recorded, we provide an alternative mechanism for real time introduction of superimposed content.
 In order to help the content creator to introduce and synchronize its content in real time, we allow the user to encode its content into \gls{QR} codes and show it to the camera in real time.
-The \gls{QR} code processing is performed at \gls{KMS}.
+The \gls{QR} code processing is performed at \gls{KMS} and its content is returned by our content scheduler.
 
@@
 We tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.
-We have also tested the performance of our solution by measuring the used resources.
+We have tested as well the performance of our solution by measuring the used resources.
 Those performance tests are crucial to ensure that our solution is in fact stable and scalable.
 
-We evaluated our system by running all the backend components as Docker containers on a Debian 8 machine with 128GB of RAM, 2 Intel Xeon E5-2640V2 CPUs, hybrid SSD, HDD storage and 1 Gbps Ethernet. 
+We evaluated our system by running all the backend components as Docker containers on a Debian 8 machine with 128GB of \emph{RAM}, 2 Intel Xeon E5-2640V2 CPUs, hybrid \emph{SSD}, \emph{HDD} storage and 1 Gbps Ethernet.
 We implemented a small \emph{Python} script using \emph{psutil}\footnote{\url{https://github.com/giampaolo/psutil} (Accessed March 27, 2016)} that collects, every second, CPU and physical memory information relative to each running process and network usage relative to each interface.
@@
 The blue peaks are caused by the signaling phase and web page download when a new client joins.
-The green peaks inter-container traffic, caused by video and audio being transfered between \gls{KMS} and \emph{MongoDB} through \emph{Kurento Repository}.
-Peaks occurs every 10s, when a block of video is recorded.
+The green peaks are inter-container traffic, caused by video and audio being transfered between \gls{KMS} and \emph{MongoDB} through \emph{Kurento Repository}.
+Peaks occurs every 10 seconds, when a block of video is recorded.
 The recordings are synchronized, so all user and mixed blocks start and end at the same time.
 %That is why the amount of work done every ten seconds accumulates, and because this is performed locally, the maximum transfer rate is limited by the performance of the memory as buffers are written to buffers then to disks.
-Sent data transfer rate has no significant peaks as \gls{HTTP} requests and signaling information contains little information, being mostly made up of video streams sent to the clients.
-As expected, send and received traffic grows linearly with the number of users.
+Received data transfer rate has no significant peaks as \gls{HTTP} requests and signaling information contains little information, being mostly made up of video streams sent by the clients.
+% Esta frase estava um bocado sem sentido (olhando para a figura, no grafico mostra o ponto de vista do servidor, a frase estava no ponto de vista do utilizador)
+As expected, sent and received traffic grows linearly with the number of users.
 
@@
 
-Figure \ref{fig:test_client_net} shows the same events from the respective of network usage on the client side.
+Figure \ref{fig:test_client_net} shows the same events from the respective network usage on the client side.
 We observe that in the first seconds the client adjusts the video quality it sends to \gls{KMS} and maintains that upload rate throughout.
@@
 When a user leaves the conference room \gls{KMS} has no need to decrease the participant's video quality as less network bandwidth will be used.
-As expected, the maximum bandwidth used is constant and independent of the number of participants.
+As expected, the maximum bandwidth used is constant and independent from the amount of participants.
 

### 2017-02-22 artigo/paper.tex
A few english revisions.

@@
 Each participant may use more than one camera, in order to show different angles or objects, e.g. a white board and himself.
-The source of the video can be selected manually or automatically by the talking participant.
+The source of the video can be selected manually or be automatically assigned to the talking participant.
 Participating users may choose to view only, share their microphone, camera or screen.
@@
 The user is free to slide over to any instant in the past or back to real-time.
-Changing the instant of time updates the state of not only the video but of other collaboration elements. % ("other collaboration elements", nao foi implementado, apenas para conteudo sobreposto...) %
+Changing the instant of time updates the state of not only the video but also of other collaboration elements.
+% ("other collaboration elements", nao foi implementado, apenas para conteudo sobreposto...) %
 
@@
 In order to limit the processing requirement in the client application (browser), we decided to use a server to combine the audio and video streams from all the participants into a single video and a single audio track.
-This techinique is called \gls{MCU}.
+This technique is called \gls{MCU}.
 This also helps conserve battery on mobile devices.
@@
 	\centering
-	\includegraphics[width=\linewidth]{figures/modules2.png}
+	\includegraphics[width=\linewidth]{figures/modules.pdf}
 	\caption{System Modules}
@@
 
-In order to simplify our solution, the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules are deployed within the same server, so to facilitate deployment.
+In order to simplify our solution, the \emph{Application}, \emph{Signaling}, \emph{Stream} and \emph{Storage} modules are deployed within the same server, as to facilitate deployment.
 To this set of modules, we call \emph{backend}.
@@
 For convenience, our storage model also uses the same database.
-In order to offer all the functionality, some information about objects must be stored on database, such as users, groups, relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.
+In order to offer all the functionality, some information about objects must be stored on the database, such as users, groups, relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.
 
@@
 	\hline 
-	\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[2pt]\cline{1-5}\cline{7-7}
-	\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[2pt]\hline
-	\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[2pt]\hline
+	\multicolumn{8}{|c|}{\cellcolor{Gray}Application}  &\\[12pt]\cline{1-5}\cline{7-7}
+	\multicolumn{1}{|c|}{jQuery} & \multicolumn{1}{c|}{HTML5} & \multicolumn{1}{c|}{CSS3 (Bootstrap)} & \multicolumn{1}{c|}{Signaling} & \multicolumn{1}{c|}{ot.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} & \multicolumn{1}{c|}{adapter.js} & \multicolumn{1}{c|}{\cellcolor{Gray}} &\\[12pt]\hline
+	\multicolumn{1}{|c|}{HTTP} & \multicolumn{2}{c|}{User Interface}  & \multicolumn{3}{c|}{WebSocket}    & \multicolumn{2}{c|}{WebRTC}      &\\[12pt]\hline
 	\end{tabular}
@@
 The Web application server retrieves all the information needed from the database in order to check if the user has permissions to join that conference room. 
-It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \gls{HTTP} context is lost.
+It is important to save the user identification before the \emph{WebSocket} connection is created because, after the handshake is performed by the \emph{WebSocket} protocol\cite{rfc6455}, the \gls{HTTP} context is lost.
 
@@
 		\centering
-		\includegraphics[width=\linewidth]{figures/devices2.png}
+		\includegraphics[width=\linewidth]{figures/devices.png}
 		\caption{Multiple devices per user}
@@
 We tested our solution with real users for a better understanding of their difficulties and what can be done in order to improve our solution's usability.
-We have tested as well the performance of our solution by measuring the used resources.
+We have also tested the performance of our solution by measuring the used resources.
 Those performance tests are crucial to ensure that our solution is in fact stable and scalable.
@@
 
-Figure \ref{fig:test_client_net} shows the same events from the respective network usage on the client side.
+Figure \ref{fig:test_client_net} shows the same events regarding network usage from the perspective of the client.
 We observe that in the first seconds the client adjusts the video quality it sends to \gls{KMS} and maintains that upload rate throughout.
